### 2015-07-17 crystal-gel/crystal_gel2.tex
Initial commit from John

%\documentclass[twocolumn,showpacs,amsmath,amssymb,pre,superscriptaddress]{revtex4}
%\documentclass[aps,prl,twocolumn,showpacs,amsmath,amssymb,superscriptaddress]{revtex4}
%\documentclass[a4paper,showpacs,aps,twocolumn,superscriptaddress]{revtex4}
\documentclass[a4paper,preprint,superscriptaddress]{revtex4}
\usepackage{graphicx}% Include figure files
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color}
\usepackage{comment}
%\usepackage{dcolumn}% Align table columns on decimal point
%\usepackage{bm}% bold math

%\graphicspath{{./FIGS/}}

\begin{document}

\title{Mechanism of crystal-gel formation}
\author{Hideyo Tsurosawa}
\affiliation{Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}
\author{John Russo}
%\email{russoj@iis.u-tokyo.ac.jp}
\affiliation{Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}
\author{Hajime Tanaka}
%\email{tanaka@iis.u-tokyo.ac.jp}
\affiliation{Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}

\date{\today}

\begin{abstract}
{\bf 
Colloidal gels are usually formed by viscoelastic phase separation and its dynamical arrest by glass transition \cite{tanaka1999colloid,tanaka2000viscoelastic,foffi2002,lu2008gelation,zaccarelli2008gelation,testard2011}.
% Thus, the dynamical arrests of colloidal gels and glasses have been believed to originate from a dynamical transition to the same non-ergodic disordered glass state, i.e., vitrification. The only difference comes from the fact that the former is formed by densification induced by phase separation and thus is spatially heterogeneous whereas the latter is formed by homogeneous densification and thus spatially homogeneous.
With confocal measurements of a PMMA colloid and polymer mixture, we experimentally access the formation of \emph{crystal-gels}, a novel arrested phase in which the dense branches of a gel are replaced by crystal aggregates~\cite{fortini2008crystallization,perez2011pathways,sabin2012}. 
For the first time we are able to observe the formation mechanism of crystal-gels at a single-particle level, giving experimental access to the elementary process of this novel route to gelation, in which phase separation is not arrested by slow dynamics but by crystallization.
The formation of the crystal gel  involves a new kinetic pathway, in which crystal seeds form inside the liquid branches of the spinodal aggregate, but then grow by direct condensation from the gas phase.
We argue that crystal gels can form as the result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems, whose dynamics is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
% , which is necessary to form a network structure of a minority liquid phase in a majority gas phase, and at the same time fast enough to prevent immediate vitrification.
For example, our experiment provides a colloidal model of the Bergeron process (or ``cold-rain process''), which regulates ice crystal growth that occurs in mixed phase clouds containing a mixture of supercooled water and ice~\cite{glickman2000glossary}.
}
\end{abstract}

\maketitle
Gels and glasses are two important non-ergodic disordered states, which are very important 
in our life and nature~\cite{anderson2002insights,lekkerkerker2011colloids}.
The dynamical arrest of colloidal gels and glasses is believed to originate from a dynamical transition to the same non-ergodic disordered glass state, i.e., vitrification. The only difference comes from the fact that the former is formed by densification induced by phase separation and thus is spatially heterogeneous, whereas the latter is formed by uniform densification and is thus spatially homogeneous.

Gels are distinct from glasses because of their heterogeneous network structure, which gives them properties intermediate between
those of a solid and a fluid phase, and allows the coexistence of two properties usually incompatible with each other, 
elasticity and fluid transport. 
The elasticity of gels originates from the percolated network structure and the resulting space-spanning connectivity. 
At the same time a fluid component can easily flow through the elastic network.
There are two types of gels with such characteristics: one is a network structure stabilized by stable crosslinks, and the other is a phase-separated structure frozen by dynamical arrest due to vitrification.
Many chemical and physical gels belong to the former category, in which the size of network pores is rather small and 
the fluid transport is not so efficient. On the other hand, gels made of spherical colloids and proteins 
with attractive interactions belong to the latter category. 
This type of gels are formed by viscoelastic phase separation \cite{tanaka1999colloid,tanaka2000viscoelastic} into a dense liquid phase with slow dynamics and a dilute gas phase 
with fast dynamics. This difference in the viscoelastic properties between the two phases allows the system to form a space-spanning 
network structure of the liquid phase, even if it is the minority phase. During phase separation, the density of the liquid phase 
increases towards the glass-transition density, leading to slow glassy dynamics.
At the glass transition point, the percolated network structure is dynamically stabilised by vitrification 
of the dense liquid phase~\cite{pusey1993dynamics,ilett1995phase,verhaegh1997transient,tanaka1999colloid,foffi2002,zaccarelli2007,lu2008gelation,zaccarelli2008gelation,testard2011}.  
More precisely, the dynamical stabilization is due to percolation of locally favoured structures, which are locally stable non-crystalline structures \cite{royall2008g}.
% Since ``glass'' is the term used for bulk, this may be a more appropriate way to express the mechanism of dynamical arrest. 
% This type of gels are made by phase separation. Thus, the structure is intrinsically heterogeneous, and the network pore size is usually 
% much larger than gels made of stable crosslinks, resulting in a high fluid transport ability.

Unlike the above standard scenario of colloidal gelation, it is known that phase separation can also be arrested by crystallization, if the process takes place below the melting point of a component of a mixture~\cite{tanaka1985new}. 
Indeed the possibility of a different class of gels, which are stabilised by crystallization, 
was suggested by numerical simulations~\cite{fortini2008crystallization,perez2011pathways} and observed by space experiments \cite{sabin2012}.
For colloidal systems, we can in principle access structural evolution in real time at a single-particle level; however, there has so far been no confocal microscopy  
study on the process of crystal-gel formation. The main difficulty consists in the preparation protocol, where a colloidal-polymer mixture undergoes
shaking or shear before observation of its dynamics. This introduces turbolent flows at the beginning of the process, and also does not allow the
observation of the first stages of crystallization.
In our experiments we instead succeed in building a novel setup that allows the formation mechanism of crystal-gels to be observed directly
under the confocal microscope from the very early stages and without introducing fluid flows.

In our experimental setup, the sample cell is put in contact with a reservoir of a salt solution through a semi-permeable membrane.
At $t=0$, the increase in the concentration of salt ions progressively screens the Coulombic repulsion between colloidal particles,
which are then subject to attractive depletion forces, making the system thermodynamically unstable and leading to phase separation (see Methods).
In the following we discuss all the stages of the formation of the crystal gels: phase separation, stress-driven agein and crystallization.
The experimental data is taken at two different volume fractions ($\phi=0.10$ and $\phi=0.25$) and for different values
of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ for $\phi=0.25$ and $c_p=0.82,1.36$ for $\phi=0.10$.
For each state points, scans at early times (every $10s$) and at the late stage (every $30s$) of the gelation process.

\subsection*{Phase separation}

In Fig.~\ref{fig:gel}a we show the cluster size distribution for all state points at their gelation point. The decay of the distribution
follows different power laws $n(s)\sim s^\tau$, where $\tau$ is the Fisher exponent \cite{family2012kinetics}.
For state points with $\phi=0.25$ the decay is well described by the three dimensional (3D) random percolation exponent $\tau=2.18$ \cite{family2012kinetics}.
For state points with $\phi=0.10$, on the other hand, the decay appears to be slower, close to the diffusion limited cluster aggregation (DLCA) 
exponent $\tau=1.8$ \cite{family2012kinetics}. While it is plausible that gelation at lower volume fraction occurs through cluster aggregation, especially given 
the high polymer concentration of these state points, we have to warn that the cluster size distribution is very sensitive to
experimental conditions and statistical noise (which is high in our case since the distributions are extracted from a single frame).
Other measures of fractal dimensions (see below) will in fact suggest that also the $\phi=0.10$ states belong to the random gel universality class.

Next we show in Fig.~\ref{fig:gel}b the radius of gyration $R_{\rm g}$ for the clusters of size $s$ at the gel point. All clusters, irrespective
of the state point, follow the power-law: $R_{\rm g}\sim s^{2/D}$, with $D=2.53$ the fractal dimension in the random gelation
universality class \cite{family2012kinetics}. So far, we have thus proven that, in the initial stages of phase separation, the liquid branch is
formed by colloidal particles that aggregate to form a random gel.

Even after a gel is formed, the system keeps evolving.
However, since a gel has an elasticity due to its space-spanning connectivity, the process should be markedly different from 
ordinary fluid phase separation. Now we consider this unique feature of the coarsening of a gel network.

\subsection*{Stress-driven ageing}

Previous simulation results~\cite{tanaka2000,tanaka2007spontaneous,furukawa2010key} have suggested that hydrodynamic interactions play
a key role in the process of colloidal gelation. Without hydrodynamic interactions, particles have the tendency to
aggregate in compact structures and subsequently form thick network structures. This also limits the possibility to
form arrested gel states only at relatively high volume fractions. With hydrodynamic interactions, particles first form
a transient gel even at very low volume fractions, and only later they increase the number of
nearest neighbours to minimize the energy of the structure. 
Thus, hydrodynamic interactions lead to the formation of gels that are very far from equilibrium and under a strong thermodynamic driving force 
towards more stable compact structures. 
The resulting transition from open to more compact networks occurs only through
the breaking of the bonds that have accumulated more stress. The process was called stress-driven ageing~\cite{tanaka2007spontaneous}, 
which is characteristic of viscoelastic phase separation \cite{tanaka2000viscoelastic}. 
Below we provide the first experimental proof of this process.

In Fig.~\ref{fig:stress}a, we plot the fraction of liquid and gas particles as a function of time during phase separation
at the state points of $\phi=0.10$. We observe that at high polymer concentration ($c_p=1.36$, dashed lines) both the
liquid and gas branch change monotonically with time as the phase separation proceeds. At low polymer concentration
($c_p=0.82$, continuous lines), on the other hand, we find a non-monotonic evolution of both the liquid and gas branches, which is signalling
a structural reorganization of the network. To prove that this is indeed a stress-driven ageing process, in Fig.~\ref{fig:stress}b 
we plot the average number of bond-breaking events for the low-polymer concentration state point. The figure clearly shows that
the network restructuring follows a peak in the number of bond-breaking events. We have thus found direct evidence of the
stress-driven ageing process in colloidal gels that is a direct consequence of the formation of a thin network with the help of hydrodynamic interactions in the first stages
of the aggregation process.

In Fig.~\ref{fig:stress}c we plot the time evolution of the structure factor, $S(q)$, for the state point
undergoing network reorganization $\phi=0.10$ and $c_p=0.82$. The time of network reorganization as found in Fig.~\ref{fig:stress}a and b
corresponds to a steep increase of the low wavenumber scattering from the network, that signals the coarsening of the network structure.

The same conclusions also hold for the state point at $\phi=0.25$. We observe network restructuring only for the state point
at lower polymer concentration, $c_p=0.38$. Interestingly, this is also the only state point that undergoes crystallization,
meaning that the coarsening of the network of the dense liquid phase and the resulting formation of a wider liquid branch is a necessary condition for crystallization.

\subsection*{Crystallization and arrest}

After the creation of the spinodal liquid branch, crystallization events start occurring for the state point $\phi=0.25$ and $c_p=0.38$, which has the lowest
polymer concentration of all state points. This is shown in Fig.~\ref{fig:crystals}a where the black line indicates the average size of the crystallites,
while the red line the average number of crystallites as a function of time. The number of crystallites first rapidly increases as nucleation events
start occurring inside the liquid branches, but eventually starts decreasing as the different crystallites grow and merge with each other.
%The average size of the crystals shows an interesting linear growth regime for $100<t<600$. This linear growth regime goes past the time
%of coalescence of nuclei, so a first speculation is that it could be due to the compensation of the following two effects:
%\begin{itemize}
% \item diffusive growth of the nuclei, which scales like $t^{1/2}$
% \item coalescence growth, which should scale faster than $t$, maybe $t^2?$
%\end{itemize}
All other state points show also signs of crystallization, but in negligible amounts. In particular, for state points with $\phi=0.25$, increasing
polymer concentration drastically reduces the amount of crystals. This is in agreement with the
idea of enhanced crystallization rates near metastable critical points~\cite{ten1997enhancement,olmsted1998spinodal} and the
results of Ref.~\cite{fortini2008crystallization,perez2011pathways},
which speculated two different arrest mechanism: crystallization at low polymer concentration, and dynamic arrest at high polymer concentrations.

To show the presence of two different arrest mechanisms, we show in Fig.~\ref{fig:crystals}b the average local volume fraction, computed
through Voronoi diagrams, as a function of the number of neighbours. For all state points, we can see that the average volume fraction of a colloidal
particle increases with its number of neighbours. Approaching close packing (12 neighbours), we clearly see two family of curves. The state point
at $\phi=0.25$ and $c_p=0.38$ reaches close packing at a volume fraction of around $\sim 60\%$, which is in reasonable agreement with the
volume fraction of the stable crystal phase. All other state point, instead, reach close packing at volume fraction $\sim 55\%$, which is
close to the volume fraction of the glass state. 
%(\textbf{the absolute values of these volume fraction are not totally convincing, probably because of a not optimal estimate of the size of the colloids. We should run the code of Mathieau with size detection}). 
The figure then clearly shows that the state point at $\phi=0.25$ and $c_p=0.38$ follows a different arrest mechanism, in which phase separation
is arrested by crystallization and not glassiness. Although our estimation of the absolute volume fraction may involves errors of a few \%, the trend should be robust.

To confirm the different arrest mechanisms, crystallization and vitrification, we compute the structure factor for the final stages of the gelation process for all state points,
as shown in Fig.~\ref{fig:crystals}c. Calculation of the structure factor is done with the Hanning window function, to minimize boundary effects.
The figure shows that at low wavenumber $q$, the structure factor displays fractal scaling compatible with Guinier law, $S(q)\sim q^{-D}$. But a difference
in the fractal dimension $D$ between the $\phi=0.25$ and $c_p=0.38$ state point and other state points is visible. Non-crystallizing state
points have $D=2.53$, of the random gel universality class, while the crystallizing state point shows a scaling compatible with $D=3$, which
signals compact crystalline nuclei. At high $q$, the first signals from diffraction planes are also evident in the $\phi=0.25$ and $c_p=0.38$ state
point. This suggests that the space filling network is composed of crystalline aggregates, differently from the fractal branches which characterize the
spinodal liquid. Also this suggests that the growth mechanism of the crystal is not simply due to filling of the spinodal liquid network, as was
instead observed in Ref.~\cite{fortini2008crystallization,perez2011pathways} for brownian dynamics simulations.

To investigate this process in more detail, we plot in Fig.~\ref{fig:crystals}c the gyration radius of individual crystalline nuclei for the state point
$\phi=0.25$ and $c_p=0.38$. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with
the dimension of random percolation $D=2.53$, while at large sizes it scales as $D=3$ of compact crystal growth. The figure thus shows that
small crystalline nuclei are nucleated inside the phase-separated liquid branches; but once they reach the transverse size of the liquid branch, the growth is continued
by direct addition of colloidal particles from the gas.

Now we show in Fig.~\ref{fig:transitions}a slabs of 10 particle diameter thickness for configurations at $\phi=0.25$ and $c_p=0.38$ at different times.
Particles are coloured according to their phase. Green: HCP; Red: FCC; blue: BCC; white: liquid or gas. The figure shows the first nucleation
events inside the liquid network (left panel); then these crystalline nuclei grow directly from the gas phase (middle panel), and finally
the different nuclei coalesce (right panel).

Next we show in Fig.~\ref{fig:transitions}b the fraction of particles in each different phase for the crystallization trajectory at
$\phi=0.25$ and $c_p=0.38$, after the liquid-gas phase separation has occurred. The process of crystallization is characterized
by a steep decrease in liquid particles, as they transform in small crystalline nuclei inside the liquid domains. This decrease
is accompanied by an increase in the fraction of gas particles: as the first crystals start to appear, liquid particles evaporate
to the gas phase due to the highest vapour pressure of the liquid phase compared to the crystalline phase. After the onset
of a steady-state gas population, the crystallization proceeds by adding surface particles to the small crystallites. To confirm this process we
examine in Fig.~\ref{fig:transitions}c the transfer matrix probabilities $P_{\alpha\beta}$, where $\alpha$ and $\beta$ are each of the
phases gas, liquid and crystal (which includes bcc, hcp, fcc and surface particles). Liquid particles undergo two processes, on one
hand they transform to crystallites growing inside the liquid branches of the network, and also increasingly transform to gas particles.
Gas particles first undergo a transition to liquid particles as the final stage of the liquid-gas phase separation, but once
crystalline nuclei outgrow the liquid pockets in which they are formed, the probability for the transition from gas to liquid
abruptly decreases. Instead, gas particles start depositing on the crystal, and at later stages the pace of the gas-to-crystal transformation
is larger than the one of the liquid-to-crystal one.

We can thus conclude that the crystallization process proceeds via two distinct mechanisms:
(i) freezing of the fluid; (ii) evaporation of the fluid and condensation of the vapor on the crystals.
The first process dominates the initial stages of crystallization, while the second one becomes effective once
crystalline nuclei have outgrown the liquid regions in which they form. 
% 
% Here we mention the simulations closer to our experimental results~\cite{fortini2008crystallization}.
% It was shown there that close to the metastable liquid-gas critical point, the gelation process is arrested by
% crystallization, while at higher polymer concentration by slow dynamics. This is roughly what we observe in
% our experiments, but there is a crucial difference. In the simulation results~\cite{fortini2008crystallization} the kinetic
% pathway at volume fraction $\phi=0.25$ consists of the following three steps: 
% spinodal decomposition; nucleation of crystalline nuclei within the fluid branches;
% growth of crystalline clusters until the whole spinodal structure is crystalline.
% As shown above, the last step appears to be markedly different in our experiments.
% More precisely, at $\phi=0.25$ and polymer concentration $c_p=0.38$ we also observe the first two steps, 
% but the growth process is entirely different. Instead of growing inside
% the spinodal aggregate, the small nuclei grow by adding colloids directly from the gas phase.

\subsection*{Discussion and conclusions}
The first stages of gelation always involve spinodal decomposition with the formation
of a fractal network corresponding to the liquid phase.
Due to hydrodynamic interaction, the network is locally stretched, and the accumulation of mechanical stress on some bonds causes some
of them to break and the network to reorganize in a more compact and larger structure. This occurs only at low polymer concentration,
while at high polymer concentrations the bonds are too strong to break. This will eventually dictate which arrest mechanism
is available to the system.
Depending on the polymer concentration, there are two possible arrest mechanisms.
(i) Crystallization, observed at $\phi=0.25$ and $c_p=0.38$, small crystalline nuclei appear inside the liquid, reach the surface
 of the liquid branches, and then grow by addition of particles from the gas phase. The final structure is a network of crystal aggregates,
 as confirmed by the fractal dimension of the branches, the volume fraction of the particles within the branches, and bond orientational analysis.
(ii) Dynamic arrest, particle arrest when the dynamics inside the liquid branch becomes slow, which should happen at the intersection of the
 glass line with the liquid side of the coexistence curve.

The process of crystal-gel formation is akin to the Bergeron process which is the primary mechanism for the formation of raindrops in clouds~\cite{glickman2000glossary}.
In clouds there is a mixture of ice crystals and supercooled water. The vapor phase is in coexistence with the liquid phase, but is supersaturated
with respect to the ice crystals. This causes the water droplets to evaporate and sublimate directly on the ice crystals. Our system can then
be regarded as a colloidal analogue for this important process, which can now be studied at the single-particle level.
The basic mechanism creating ``crystal gels'' may be generic to many other
systems. The requirements are (i) the presence of gas-liquid phase separation below the melting point of a crystal, (ii) 
weak or little frustration against crystallization (in our case, the use of monodisperse colloids), 
(ii) dynamical slowing down in a supercooled liquid state, which is necessary to induce viscoelastic phase separation leading to the formation 
of a network structure of the minority phase, and (iii) the degree of supercooling is low enough to avoid a high rate of densification of the liquid phase 
leading to vitrification.   
Many monoatomic and single-component molecular systems can satisfy all these conditions in a certain range of the temperature and pressure. 
This can be seen, for example, by looking at the phase diagram of a Lennard-Jones liquid, which represents many molecular systems 
without specific directional interactions. So we believe that crystal gels are an important class of heterogeneous non-ergodic states in nature 
and industrial applications, although they have not attracted much attention so far.

\vspace{2cm}
\noindent
{\bf \large Methods}

\noindent
{\bf Experimental.}

\noindent
{\bf State points studied. }
The experimental data is taken at two different volume fractions ($\phi=0.10$ and $\phi=0.25$) and for different values
of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ for $\phi=0.25$ and $c_p=0.82,1.36$ for $\phi=0.10$.
For each state points, scans at early times (every $10s$) and at the late stage (every $30s$) of the gelation process.

We will consider two particles bonded if their center-to-center distance is within $12$ pixels. The radius of a colloidal
particle is $5.365$ pixels.

\noindent
{\bf Structural analysis}
To detect percolation, two different methods have been employed. In the first one we detect percolation
by looking at which frame the cluster size distribution has a more extended power-law decay. In the second technique,
we simply measure the spatial extent of the largest cluster and consider it percolating when it is comparable to the
size of the field of view of the microscope. Both methods lead to essentially identical percolation time for each state point.

\bibliographystyle{naturemag4}
\bibliography{biblio}

\clearpage

\begin{figure}[!t]
 \centering
 \includegraphics[width=8cm,clip]{fig1a}
  \includegraphics[width=8cm]{fig1b}
 % csd_beforegrl.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
 \caption{{\bf Characterisation of gels formed by phase separation.} {\bf a,} 
 Cluster size distribution for all state points at percolation, where $s$ is the size of the cluster. Different power-laws are represented with dashed lines. 
{\bf b,} Radius of gyration ($R_g$) for clusters at the gel point.}
 \label{fig:gel}
\end{figure}

\clearpage

\begin{figure}[!t]
 \centering
 \includegraphics[width=8cm]{fig2a}
  \includegraphics[width=8cm]{fig2b}
  \includegraphics[width=8cm]{fig2c}
 % figure_restructuring.eps: 990x765 pixel, 300dpi, 8.38x6.48 cm, bb=0 -1 1453 480
 \caption{{\bf Stress-induced ageing.} {\bf a,} Fraction of liquid and gas particles as a function of time during phase separation for state points
 $\phi=0.10$, $c_p=0.82$ (continuous lines) and $c_p=1.36$ (dashed lines). The two vertical (green) lines are located at time $t=600\,s$ and $t=700\,s$
 respectively. {\bf b,} Average number of bond breaking events
 per particle for the state point $\phi=0.10$, $c_p=1.36$. The fraction of liquid particles for the same state point
 is also depicted for clarity. The two vertical (green) lines are located at time $t=600\,s$ and $t=700\,s$
 respectively.
{\bf c,} Time evolution of the structure factor, $S(q)$, for the state point $\phi=0.10$ and $c_p=0.82$.}
 \label{fig:stress}
\end{figure}

\clearpage

\begin{figure}[!t]
 \centering
 \includegraphics[width=8cm]{fig3a}
 \includegraphics[width=8cm]{fig3b}
 \includegraphics[width=8cm]{fig3c}
 \includegraphics[width=8cm]{fig3d}
 % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
 \caption{{\bf Two types of arrest mechanism leading to gel formation.} {\bf a,} Time evolution of the average size of the crystals (black line) and the number of crystals (red line) for the state point $\phi=0.25$ and $c_p=0.38$. {\bf b,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle.
 Different lines represent the different state points. {\bf c,} Structure factors for the late stages of the gelation process. Calculation of the structure factor is done with the Hanning window function,  to minimize boundary effects. {\bf d,} Radius of gyration of crystalline nuclei at the late stages of gelation for the state point $\phi=0.25$ and $c_p=0.38$.} 
 \label{fig:crystals}
\end{figure}

\begin{figure}[!t]
 \centering
 \includegraphics[width=16cm]{fig4a}
 \includegraphics[width=8cm]{fig4b}
 \includegraphics[width=8cm]{fig4c}
 % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735
 \caption{
{\bf Crystal gel formation.} 
{\bf a,} Snapshots of configurations at $\phi=0.25$ and $c_p=0.38$, for t=4856 (left), 12956 (center), 27956 (right). Particles are colored according to their phase. Green: HCP; Red: FCC; blue: BCC; white: liquid or gas. 
 {\bf b,} Fraction of particles for each phase (between bcc, hcp, fcc, surface, liquid or gas) at $\phi=0.25$ and $c_p=0.38$.
 Crystalline phases are detected through bond orientational analysis~\cite{russo2012microscopic}; surface particles are particles
 with at least two crystalline neighbours; liquid particles have at least 4 neighbours; gas particles are particles with less than 4
 nearest neighbours.
{\bf c,} Transition probabilities for the same process as in {\bf b}. Crystal includes all crystalline phases (bcc, hcp, fcc) and surface particles.}
 \label{fig:transitions}
\end{figure}

\end{document}

\section{Future things to do}
Is the absence of network restructuring that is preventing crystallization?

### 2015-07-17 crystal-gel/crystal_gel_letter.tex
Letter layout Reorganise the figures according to what was decided with John.

--- 2015-07-17 crystal-gel/crystal_gel2.tex
@@
-%\documentclass[twocolumn,showpacs,amsmath,amssymb,pre,superscriptaddress]{revtex4}
-%\documentclass[aps,prl,twocolumn,showpacs,amsmath,amssymb,superscriptaddress]{revtex4}
-%\documentclass[a4paper,showpacs,aps,twocolumn,superscriptaddress]{revtex4}
-\documentclass[a4paper,preprint,superscriptaddress]{revtex4}
+\documentclass[twocolumn,superscriptaddress,showpacs,preprintnumbers,
+amsmath,amssymb,prl]{revtex4-1}
+\usepackage[utf8]{inputenc}
 \usepackage{graphicx}% Include figure files
@@
 \title{Mechanism of crystal-gel formation}
-\author{Hideyo Tsurosawa}
+\author{Hideyo Tsurusawa}
 \affiliation{Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}
@@
 \affiliation{Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}
+\author{Mathieu Leocmach}
+\affiliation{Institut Lumière Matière, CNRS UMR 5306, Université Claude Bernard Lyon 1, Université de Lyon, Lyon, 69622 Villeurbanne Cedex, France}
 \author{Hajime Tanaka}
@@
 \begin{abstract}
-{\bf 
+{
 Colloidal gels are usually formed by viscoelastic phase separation and its dynamical arrest by glass transition \cite{tanaka1999colloid,tanaka2000viscoelastic,foffi2002,lu2008gelation,zaccarelli2008gelation,testard2011}.
@@
 
+\begin{figure}
+ \centering
+ \framebox{Phase diagram\hfill}
+ \includegraphics[width=8cm]{oldfig/fig3b}
+ \includegraphics[width=8cm]{oldfig/fig3c}
+ % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
+ \caption{{\bf Two types of arrest mechanism leading to gel formation.} \textbf{a,} Phase diagram \textbf{b,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle.
+ Different lines represent the different state points. {\bf c,} Structure factors for the late stages of the gelation process.} 
+ \label{fig:statepoints}
+\end{figure}
+
+\begin{figure}[!t]
+ \centering
+ \includegraphics[width=8cm]{oldfig/fig3a}
+ \includegraphics[width=8cm]{oldfig/fig3d}
+ % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
+ \caption{{\bf Characterisation of the crystal-gel.} {\bf a,} Time evolution of the average size of the crystals (black line) and the number of crystals (red line) for the state point $\phi=0.25$ and $c_p=0.38$.
+{\bf d,} Radius of gyration of crystalline nuclei at the late stages of gelation for the same state point.}
+ \label{fig:crystals}
+\end{figure}
+
 After the creation of the spinodal liquid branch, crystallization events start occurring for the state point $\phi=0.25$ and $c_p=0.38$, which has the lowest
@@
 by direct addition of colloidal particles from the gas.
+
+\begin{figure*}
+ \centering
+ \includegraphics[width=16cm]{oldfig/fig4a}
+ \includegraphics[width=8cm]{oldfig/fig4b}
+ \includegraphics[width=8cm]{oldfig/fig4c}
+ % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735
+ \caption{
+{\bf Crystal gel formation.} 
+\textbf{a,} Snapshots of configurations at $\phi=0.25$ and $c_p=0.38$, for t=4856 (left), 12956 (center), 27956 (right).
+Particles are colored according to their phase.
+Green: HCP; Red: FCC; blue: BCC; white: liquid or gas.
+ \textbf{b,} Fraction of particles for each phase (between bcc, hcp, fcc, surface, liquid or gas) at $\phi=0.25$ and $c_p=0.38$.
+ Crystalline phases are detected through bond orientational analysis~\cite{russo2012microscopic}; surface particles are particles
+ with at least two crystalline neighbours; liquid particles have at least 4 neighbours; gas particles are particles with less than 4
+ nearest neighbours.
+\textbf{c,} Transition probabilities for the same process as in \textbf{b}. Crystal includes all crystalline phases (bcc, hcp, fcc) and surface particles.}
+ \label{fig:transitions}
+\end{figure*}
 
@@
 
-\begin{figure}[!t]
- \centering
- \includegraphics[width=8cm,clip]{fig1a}
-  \includegraphics[width=8cm]{fig1b}
- % csd_beforegrl.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
- \caption{{\bf Characterisation of gels formed by phase separation.} {\bf a,} 
- Cluster size distribution for all state points at percolation, where $s$ is the size of the cluster. Different power-laws are represented with dashed lines. 
-{\bf b,} Radius of gyration ($R_g$) for clusters at the gel point.}
- \label{fig:gel}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[!t]
- \centering
- \includegraphics[width=8cm]{fig2a}
-  \includegraphics[width=8cm]{fig2b}
-  \includegraphics[width=8cm]{fig2c}
- % figure_restructuring.eps: 990x765 pixel, 300dpi, 8.38x6.48 cm, bb=0 -1 1453 480
- \caption{{\bf Stress-induced ageing.} {\bf a,} Fraction of liquid and gas particles as a function of time during phase separation for state points
- $\phi=0.10$, $c_p=0.82$ (continuous lines) and $c_p=1.36$ (dashed lines). The two vertical (green) lines are located at time $t=600\,s$ and $t=700\,s$
- respectively. {\bf b,} Average number of bond breaking events
- per particle for the state point $\phi=0.10$, $c_p=1.36$. The fraction of liquid particles for the same state point
- is also depicted for clarity. The two vertical (green) lines are located at time $t=600\,s$ and $t=700\,s$
- respectively.
-{\bf c,} Time evolution of the structure factor, $S(q)$, for the state point $\phi=0.10$ and $c_p=0.82$.}
- \label{fig:stress}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[!t]
- \centering
- \includegraphics[width=8cm]{fig3a}
- \includegraphics[width=8cm]{fig3b}
- \includegraphics[width=8cm]{fig3c}
- \includegraphics[width=8cm]{fig3d}
- % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
- \caption{{\bf Two types of arrest mechanism leading to gel formation.} {\bf a,} Time evolution of the average size of the crystals (black line) and the number of crystals (red line) for the state point $\phi=0.25$ and $c_p=0.38$.
-{\bf b,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle.
- Different lines represent the different state points.
-{\bf c,} Structure factors for the late stages of the gelation process.
-Calculation of the structure factor is done with the Hanning window function,  to minimize boundary effects.
-{\bf d,} Radius of gyration of crystalline nuclei at the late stages of gelation for the state point $\phi=0.25$ and $c_p=0.38$.}
- \label{fig:crystals}
-\end{figure}
-
-\begin{figure}[!t]
- \centering
- \includegraphics[width=16cm]{fig4a}
- \includegraphics[width=8cm]{fig4b}
- \includegraphics[width=8cm]{fig4c}
- % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735
- \caption{
-{\bf Crystal gel formation.} 
-{\bf a,} Snapshots of configurations at $\phi=0.25$ and $c_p=0.38$, for t=4856 (left), 12956 (center), 27956 (right).
-Particles are colored according to their phase.
-Green: HCP; Red: FCC; blue: BCC; white: liquid or gas.
- {\bf b,} Fraction of particles for each phase (between bcc, hcp, fcc, surface, liquid or gas) at $\phi=0.25$ and $c_p=0.38$.
- Crystalline phases are detected through bond orientational analysis~\cite{russo2012microscopic}; surface particles are particles
- with at least two crystalline neighbours; liquid particles have at least 4 neighbours; gas particles are particles with less than 4
- nearest neighbours.
-{\bf c,} Transition probabilities for the same process as in {\bf b}. Crystal includes all crystalline phases (bcc, hcp, fcc) and surface particles.}
- \label{fig:transitions}
-\end{figure}
-
 \end{document}

### 2015-07-23 crystal_gel_letter.tex
This is trial version with title, abstract and parts of the introduction modified to include the idea of "colloidal rain". If the idea is dismissed we should revert to the previous version.

--- 2015-07-17 crystal-gel/crystal_gel_letter.tex
@@
-\documentclass[twocolumn,superscriptaddress,showpacs,preprintnumbers,
-amsmath,amssymb,prl]{revtex4-1}
+%\documentclass[twocolumn,superscriptaddress,showpacs,preprintnumbers,amsmath,amssymb,prl]{revtex4-1}
+\documentclass[twocolumn,showpacs,amsmath,amssymb,pre,superscriptaddress]{revtex4}
 \usepackage[utf8]{inputenc}
@@
 
-\title{Mechanism of crystal-gel formation}
+%\title{``Colloidal rain'', an alternative route to gelation during viscoelastic phase separation}
+%\title{Colloidal rain: a novel crystallization pathway following viscoelastic phase separation}
+\title{Novel crystallization pathway\\and formation of ``colloidal rain'' during viscoelastic phase separation}
 \author{Hideyo Tsurusawa}
@@
 {
-Colloidal gels are usually formed by viscoelastic phase separation and its dynamical arrest by glass transition \cite{tanaka1999colloid,tanaka2000viscoelastic,foffi2002,lu2008gelation,zaccarelli2008gelation,testard2011}.
-% Thus, the dynamical arrests of colloidal gels and glasses have been believed to originate from a dynamical transition to the same non-ergodic disordered glass state, i.e., vitrification.
-The only difference comes from the fact that the former is formed by densification induced by phase separation and thus is spatially heterogeneous whereas the latter is formed by homogeneous densification and thus spatially homogeneous.
-With confocal measurements of a PMMA colloid and polymer mixture, we experimentally access the formation of \emph{crystal-gels}, a novel arrested phase in which the dense branches of a gel are replaced by crystal aggregates~\cite{fortini2008crystallization,perez2011pathways,sabin2012}.
-For the first time we are able to observe the formation mechanism of crystal-gels at a single-particle level, giving experimental access to the elementary process of this novel route to gelation, in which phase separation is not arrested by slow dynamics but by crystallization.
-The formation of the crystal gel  involves a new kinetic pathway, in which crystal seeds form inside the liquid branches of the spinodal aggregate, but then grow by direct condensation from the gas phase.
-We argue that crystal gels can form as the result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems, whose dynamics is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
-% , which is necessary to form a network structure of a minority liquid phase in a majority gas phase, and at the same time fast enough to prevent immediate vitrification.
-For example, our experiment provides a colloidal model of the Bergeron process (or ``cold-rain process''), which regulates ice crystal growth that occurs in mixed phase clouds containing a mixture of supercooled water and ice~\cite{glickman2000glossary}.
+With confocal measurements of a PMMA colloid and polymer mixture, we experimentally access the formation of a novel arrested state,
+in which the dense branches of a gel are replaced by crystal aggregates~\cite{fortini2008crystallization,perez2011pathways,sabin2012}, which then grow into droplets by direct condensation of the gas phase on the surface of the crystals.
+We call this new state \emph{colloidal rain}, as it represents the colloidal analogue of rain formation from ice crystals that occurs in mixed phase clouds (the Bergeron process)~\cite{glickman2000glossary}.
+For the first time we are able to observe the dynamics of the system at single-particle level, giving full experimental access to the entire kinetic
+pathway of droplet formation. The formation of colloidal rain occurs at low polymer concentrations, close to the metastable critical point~\cite{ten1997enhancement}, where
+the strength of the interparticle attractions is low enough to allow bond reorganization and nucleation. For high polymer concentrations the system
+forms colloidal gels~\cite{lu2008gelation}, 
+where viscoelastic phase separation is interrupted by the glass transition.
+We thus argue that similar states can form as the result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems, whose dynamics is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
 }
@@
 observation of the first stages of crystallization.
-In our experiments we instead succeed in building a novel setup that allows the formation mechanism of crystal-gels to be observed directly
+In our experiments we instead succeed in building a novel setup that allows the time evolution of the system to be observed directly
 under the confocal microscope from the very early stages and without introducing fluid flows.
+
+Our results show that, for low polymer concentration, the system can undergo a novel crystallization pathway that leads to the formation
+of crystalline droplets.
+These droplets originate from crystal seeds that form inside the liquid branches of the spinodal aggregate, but then grow by direct condensation from the gas phase.
+The process is analogous to the Bergeron process of ice formation in mixed phase clouds, where ice droplets grow at the
+expense of the supercooled liquid droplets due to their lower saturated vapour pressure. In our colloidal system, we then call this process ``colloidal rain''
+formation.
+
+\subsection{Experiments}
 
@@
 For each state points, scans at early times (every $10s$) and at the late stage (every $30s$) of the gelation process.
+{\color{blue}We need more details of the experiments here}.
 
-\subsection*{Phase separation}
-
-In Fig.~\ref{fig:gel}a we show the cluster size distribution for all state points at their gelation point. The decay of the distribution
-follows different power laws $n(s)\sim s^\tau$, where $\tau$ is the Fisher exponent \cite{family2012kinetics}.
-For state points with $\phi=0.25$ the decay is well described by the three dimensional (3D) random percolation exponent $\tau=2.18$ \cite{family2012kinetics}.
-For state points with $\phi=0.10$, on the other hand, the decay appears to be slower, close to the diffusion limited cluster aggregation (DLCA) 
-exponent $\tau=1.8$ \cite{family2012kinetics}. While it is plausible that gelation at lower volume fraction occurs through cluster aggregation, especially given 
-the high polymer concentration of these state points, we have to warn that the cluster size distribution is very sensitive to
-experimental conditions and statistical noise (which is high in our case since the distributions are extracted from a single frame).
-Other measures of fractal dimensions (see below) will in fact suggest that also the $\phi=0.10$ states belong to the random gel universality class.
-
-Next we show in Fig.~\ref{fig:gel}b the radius of gyration $R_{\rm g}$ for the clusters of size $s$ at the gel point. All clusters, irrespective
-of the state point, follow the power-law: $R_{\rm g}\sim s^{2/D}$, with $D=2.53$ the fractal dimension in the random gelation
-universality class \cite{family2012kinetics}. So far, we have thus proven that, in the initial stages of phase separation, the liquid branch is
-formed by colloidal particles that aggregate to form a random gel.
-
-Even after a gel is formed, the system keeps evolving.
-However, since a gel has an elasticity due to its space-spanning connectivity, the process should be markedly different from 
-ordinary fluid phase separation. Now we consider this unique feature of the coarsening of a gel network.
-
-\subsection*{Stress-driven ageing}
-
-Previous simulation results~\cite{tanaka2000,tanaka2007spontaneous,furukawa2010key} have suggested that hydrodynamic interactions play
-a key role in the process of colloidal gelation. Without hydrodynamic interactions, particles have the tendency to
-aggregate in compact structures and subsequently form thick network structures. This also limits the possibility to
-form arrested gel states only at relatively high volume fractions. With hydrodynamic interactions, particles first form
-a transient gel even at very low volume fractions, and only later they increase the number of
-nearest neighbours to minimize the energy of the structure. 
-Thus, hydrodynamic interactions lead to the formation of gels that are very far from equilibrium and under a strong thermodynamic driving force 
-towards more stable compact structures. 
-The resulting transition from open to more compact networks occurs only through
-the breaking of the bonds that have accumulated more stress. The process was called stress-driven ageing~\cite{tanaka2007spontaneous}, 
-which is characteristic of viscoelastic phase separation \cite{tanaka2000viscoelastic}. 
-Below we provide the first experimental proof of this process.
-
-In Fig.~\ref{fig:stress}a, we plot the fraction of liquid and gas particles as a function of time during phase separation
-at the state points of $\phi=0.10$. We observe that at high polymer concentration ($c_p=1.36$, dashed lines) both the
-liquid and gas branch change monotonically with time as the phase separation proceeds. At low polymer concentration
-($c_p=0.82$, continuous lines), on the other hand, we find a non-monotonic evolution of both the liquid and gas branches, which is signalling
-a structural reorganization of the network. To prove that this is indeed a stress-driven ageing process, in Fig.~\ref{fig:stress}b 
-we plot the average number of bond-breaking events for the low-polymer concentration state point. The figure clearly shows that
-the network restructuring follows a peak in the number of bond-breaking events. We have thus found direct evidence of the
-stress-driven ageing process in colloidal gels that is a direct consequence of the formation of a thin network with the help of hydrodynamic interactions in the first stages
-of the aggregation process.
-
-In Fig.~\ref{fig:stress}c we plot the time evolution of the structure factor, $S(q)$, for the state point
-undergoing network reorganization $\phi=0.10$ and $c_p=0.82$. The time of network reorganization as found in Fig.~\ref{fig:stress}a and b
-corresponds to a steep increase of the low wavenumber scattering from the network, that signals the coarsening of the network structure.
-
-The same conclusions also hold for the state point at $\phi=0.25$. We observe network restructuring only for the state point
-at lower polymer concentration, $c_p=0.38$. Interestingly, this is also the only state point that undergoes crystallization,
-meaning that the coarsening of the network of the dense liquid phase and the resulting formation of a wider liquid branch is a necessary condition for crystallization.
+%
+%\subsection*{Phase separation}
+%
+%In Fig.~\ref{fig:gel}a we show the cluster size distribution for all state points at their gelation point. The decay of the distribution
+%follows different power laws $n(s)\sim s^\tau$, where $\tau$ is the Fisher exponent \cite{family2012kinetics}.
+%For state points with $\phi=0.25$ the decay is well described by the three dimensional (3D) random percolation exponent $\tau=2.18$ \cite{family2012kinetics}.
+%For state points with $\phi=0.10$, on the other hand, the decay appears to be slower, close to the diffusion limited cluster aggregation (DLCA) 
+%exponent $\tau=1.8$ \cite{family2012kinetics}. While it is plausible that gelation at lower volume fraction occurs through cluster aggregation, especially given 
+%the high polymer concentration of these state points, we have to warn that the cluster size distribution is very sensitive to
+%experimental conditions and statistical noise (which is high in our case since the distributions are extracted from a single frame).
+%Other measures of fractal dimensions (see below) will in fact suggest that also the $\phi=0.10$ states belong to the random gel universality class.
+%
+%Next we show in Fig.~\ref{fig:gel}b the radius of gyration $R_{\rm g}$ for the clusters of size $s$ at the gel point. All clusters, irrespective
+% of the state point, follow the power-law: $R_{\rm g}\sim s^{2/D}$, with $D=2.53$ the fractal dimension in the random gelation
+% universality class \cite{family2012kinetics}. So far, we have thus proven that, in the initial stages of phase separation, the liquid branch is
+% formed by colloidal particles that aggregate to form a random gel.
+% 
+% Even after a gel is formed, the system keeps evolving.
+% However, since a gel has an elasticity due to its space-spanning connectivity, the process should be markedly different from 
+% ordinary fluid phase separation. Now we consider this unique feature of the coarsening of a gel network.  
+% 
+% \subsection*{Stress-driven ageing}
+% 
+% Previous simulation results~\cite{tanaka2000,tanaka2007spontaneous,furukawa2010key} have suggested that hydrodynamic interactions play
+% a key role in the process of colloidal gelation. Without hydrodynamic interactions, particles have the tendency to
+% aggregate in compact structures and subsequently form thick network structures. This also limits the possibility to
+% form arrested gel states only at relatively high volume fractions. With hydrodynamic interactions, particles first form
+% a transient gel even at very low volume fractions, and only later they increase the number of
+% nearest neighbours to minimize the energy of the structure. 
+% Thus, hydrodynamic interactions lead to the formation of gels that are very far from equilibrium and under a strong thermodynamic driving force 
+% towards more stable compact structures. 
+% The resulting transition from open to more compact networks occurs only through
+% the breaking of the bonds that have accumulated more stress. The process was called stress-driven ageing~\cite{tanaka2007spontaneous}, 
+% which is characteristic of viscoelastic phase separation \cite{tanaka2000viscoelastic}. 
+% Below we provide the first experimental proof of this process.
+% 
+% In Fig.~\ref{fig:stress}a, we plot the fraction of liquid and gas particles as a function of time during phase separation
+% at the state points of $\phi=0.10$. We observe that at high polymer concentration ($c_p=1.36$, dashed lines) both the
+% liquid and gas branch change monotonically with time as the phase separation proceeds. At low polymer concentration
+% ($c_p=0.82$, continuous lines), on the other hand, we find a non-monotonic evolution of both the liquid and gas branches, which is signalling
+% a structural reorganization of the network. To prove that this is indeed a stress-driven ageing process, in Fig.~\ref{fig:stress}b 
+% we plot the average number of bond-breaking events for the low-polymer concentration state point. The figure clearly shows that
+% the network restructuring follows a peak in the number of bond-breaking events. We have thus found direct evidence of the
+% stress-driven ageing process in colloidal gels that is a direct consequence of the formation of a thin network with the help of hydrodynamic interactions in the first stages
+% of the aggregation process.
+% 
+% In Fig.~\ref{fig:stress}c we plot the time evolution of the structure factor, $S(q)$, for the state point
+% undergoing network reorganization $\phi=0.10$ and $c_p=0.82$. The time of network reorganization as found in Fig.~\ref{fig:stress}a and b
+% corresponds to a steep increase of the low wavenumber scattering from the network, that signals the coarsening of the network structure. 
+% 
+% The same conclusions also hold for the state point at $\phi=0.25$. We observe network restructuring only for the state point
+% at lower polymer concentration, $c_p=0.38$. Interestingly, this is also the only state point that undergoes crystallization,
+% meaning that the coarsening of the network of the dense liquid phase and the resulting formation of a wider liquid branch is a necessary condition for crystallization.
 
@@
 After the creation of the spinodal liquid branch, crystallization events start occurring for the state point $\phi=0.25$ and $c_p=0.38$, which has the lowest
-polymer concentration of all state points. This is shown in Fig.~\ref{fig:crystals}a where the black line indicates the average size of the crystallites,
+polymer concentration of all state points. This is shown in Fig.~\ref{fig:statepoints}a where the black line indicates the average size of the crystallites,
 while the red line the average number of crystallites as a function of time. The number of crystallites first rapidly increases as nucleation events

### 2015-07-24 crystal-gel/crystal_gel_letter.tex
New version, with new figure layout, changed text.

@@
 {
-Colloidal gels are usually formed by viscoelastic phase separation and its dynamical arrest by glass transition \cite{tanaka1999colloid,tanaka2000viscoelastic,foffi2002,lu2008gelation,zaccarelli2008gelation,testard2011}.
-% Thus, the dynamical arrests of colloidal gels and glasses have been believed to originate from a dynamical transition to the same non-ergodic disordered glass state, i.e., vitrification.
-The only difference comes from the fact that the former is formed by densification induced by phase separation and thus is spatially heterogeneous whereas the latter is formed by homogeneous densification and thus spatially homogeneous.
-With confocal measurements of a PMMA colloid and polymer mixture, we experimentally access the formation of \emph{crystal-gels}, a novel arrested phase in which the dense branches of a gel are replaced by crystal aggregates~\cite{fortini2008crystallization,perez2011pathways,sabin2012}.
-For the first time we are able to observe the formation mechanism of crystal-gels at a single-particle level, giving experimental access to the elementary process of this novel route to gelation, in which phase separation is not arrested by slow dynamics but by crystallization.
-The formation of the crystal gel  involves a new kinetic pathway, in which crystal seeds form inside the liquid branches of the spinodal aggregate, but then grow by direct condensation from the gas phase.
-We argue that crystal gels can form as the result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems, whose dynamics is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
-% , which is necessary to form a network structure of a minority liquid phase in a majority gas phase, and at the same time fast enough to prevent immediate vitrification.
-For example, our experiment provides a colloidal model of the Bergeron process (or ``cold-rain process''), which regulates ice crystal growth that occurs in mixed phase clouds containing a mixture of supercooled water and ice~\cite{glickman2000glossary}.
+With confocal measurements of a PMMA colloid and polymer mixture, we experimentally access the formation of a novel arrested state,
+formed by crystal droplets nucleating inside the dense branches of a gel, and growing by direct condensation of the gas phase on the surface of the crystals.
+% in which the dense branches of a gel are replaced by crystal aggregates~\cite{fortini2008crystallization,perez2011pathways,sabin2012}, which then grow into droplets by direct condensation of the gas phase on the surface of the crystals.
+We call this new state \emph{colloidal rain}, as it represents the colloidal analogue of rain formation from ice crystals that occurs in mixed phase clouds (the Bergeron process)~\cite{glickman2000glossary}.
+We observe the dynamics of the system at single-particle level, giving for the first time full experimental access to the kinetic
+pathway of droplet formation. The formation of colloidal rain occurs at low polymer concentrations, close to the metastable critical point~\cite{ten1997enhancement}, where
+the strength of the interparticle attractions is low enough to allow bond reorganization and nucleation. For high polymer concentrations, instead, the system
+forms colloidal gels~\cite{lu2008gelation}, 
+where viscoelastic phase separation is interrupted by the glass transition.
+We thus argue that states similar to colloidal rain can form as the result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems, whose dynamics is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
 }
@@
 The only difference comes from the fact that the former is formed by densification induced by phase separation and thus is spatially heterogeneous, whereas the latter is formed by uniform densification and is thus spatially homogeneous.
-
 Gels are distinct from glasses because of their heterogeneous network structure, which gives them properties intermediate between
@@
 observation of the first stages of crystallization.
-In our experiments we instead succeed in building a novel setup that allows the formation mechanism of crystal-gels to be observed directly
+In our experiments we instead succeed in building a novel setup that allows the time evolution of the system to be observed directly
 under the confocal microscope from the very early stages and without introducing fluid flows.
+
+Our results show that, for low polymer concentration, the system can undergo a novel crystallization pathway that leads to the formation
+of crystalline droplets.
+These droplets originate from crystal seeds that form inside the liquid branches of the spinodal aggregate, but then grow by direct condensation from the gas phase.
+The process is analogous to the Bergeron process of ice formation in mixed phase clouds, where ice droplets grow at the
+expense of the supercooled liquid droplets due to their lower saturated vapour pressure. In our colloidal system, we then call this process ``colloidal rain''
+formation. Due to the difference in the nucleation and growth processes, the droplets display different scaling laws depending on their
+size. For small nuclei, the droplets have the same fractal dimension of the embedding gel network. But during the growth process, droplets
+grow with an integer fractal dimension due to direct condensation of gas particles.
+This is in contrast to the standard crystal-gel scenario, where crystals nucleate and grow inside the dense branches of the gel, becoming dynamically arrested.
+We also observe that increasing the strength of the interparticle attraction
+(by increasing the polymer concentration), instead of forming colloidal rain, the system undergoes the classical gelation process, in which
+phase separation is interrupted by vitrification~\cite{lu2008gelation}.
+
+\subsection*{Experiments}
 
@@
 which are then subject to attractive depletion forces, making the system thermodynamically unstable and leading to phase separation (see Methods).
-In the following we discuss all the stages of the formation of the crystal gels: phase separation, stress-driven agein and crystallization.
+%In the following we discuss all the stages of the formation of the crystal gels: phase separation, stress-driven agein and crystallization.
 The experimental data is taken at two different volume fractions ($\phi=0.10$ and $\phi=0.25$) and for different values
 of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ for $\phi=0.25$ and $c_p=0.82,1.36$ for $\phi=0.10$.
+In Fig.XXX we superimpose the state points with the phase diagram calculated from the YYY equation of state.
 For each state points, scans at early times (every $10s$) and at the late stage (every $30s$) of the gelation process.
+{\color{blue}We need more details of the experiments here}.
 
-\subsection*{Phase separation}
+%
+%\subsection*{Phase separation}
+%
+%In Fig.~\ref{fig:gel}a we show the cluster size distribution for all state points at their gelation point. The decay of the distribution
+%follows different power laws $n(s)\sim s^\tau$, where $\tau$ is the Fisher exponent \cite{family2012kinetics}.
+%For state points with $\phi=0.25$ the decay is well described by the three dimensional (3D) random percolation exponent $\tau=2.18$ \cite{family2012kinetics}.
+%For state points with $\phi=0.10$, on the other hand, the decay appears to be slower, close to the diffusion limited cluster aggregation (DLCA) 
+%exponent $\tau=1.8$ \cite{family2012kinetics}. While it is plausible that gelation at lower volume fraction occurs through cluster aggregation, especially given 
+%the high polymer concentration of these state points, we have to warn that the cluster size distribution is very sensitive to
+%experimental conditions and statistical noise (which is high in our case since the distributions are extracted from a single frame).
+%Other measures of fractal dimensions (see below) will in fact suggest that also the $\phi=0.10$ states belong to the random gel universality class.
+%
+%Next we show in Fig.~\ref{fig:gel}b the radius of gyration $R_{\rm g}$ for the clusters of size $s$ at the gel point. All clusters, irrespective
+% of the state point, follow the power-law: $R_{\rm g}\sim s^{2/D}$, with $D=2.53$ the fractal dimension in the random gelation
+% universality class \cite{family2012kinetics}. So far, we have thus proven that, in the initial stages of phase separation, the liquid branch is
+% formed by colloidal particles that aggregate to form a random gel.
+% 
+% Even after a gel is formed, the system keeps evolving.
+% However, since a gel has an elasticity due to its space-spanning connectivity, the process should be markedly different from 
+% ordinary fluid phase separation. Now we consider this unique feature of the coarsening of a gel network.  
+% 
+% \subsection*{Stress-driven ageing}
+% 
+% Previous simulation results~\cite{tanaka2000,tanaka2007spontaneous,furukawa2010key} have suggested that hydrodynamic interactions play
+% a key role in the process of colloidal gelation. Without hydrodynamic interactions, particles have the tendency to
+% aggregate in compact structures and subsequently form thick network structures. This also limits the possibility to
+% form arrested gel states only at relatively high volume fractions. With hydrodynamic interactions, particles first form
+% a transient gel even at very low volume fractions, and only later they increase the number of
+% nearest neighbours to minimize the energy of the structure. 
+% Thus, hydrodynamic interactions lead to the formation of gels that are very far from equilibrium and under a strong thermodynamic driving force 
+% towards more stable compact structures. 
+% The resulting transition from open to more compact networks occurs only through
+% the breaking of the bonds that have accumulated more stress. The process was called stress-driven ageing~\cite{tanaka2007spontaneous}, 
+% which is characteristic of viscoelastic phase separation \cite{tanaka2000viscoelastic}. 
+% Below we provide the first experimental proof of this process.
+% 
+% In Fig.~\ref{fig:stress}a, we plot the fraction of liquid and gas particles as a function of time during phase separation
+% at the state points of $\phi=0.10$. We observe that at high polymer concentration ($c_p=1.36$, dashed lines) both the
+% liquid and gas branch change monotonically with time as the phase separation proceeds. At low polymer concentration
+% ($c_p=0.82$, continuous lines), on the other hand, we find a non-monotonic evolution of both the liquid and gas branches, which is signalling
+% a structural reorganization of the network. To prove that this is indeed a stress-driven ageing process, in Fig.~\ref{fig:stress}b 
+% we plot the average number of bond-breaking events for the low-polymer concentration state point. The figure clearly shows that
+% the network restructuring follows a peak in the number of bond-breaking events. We have thus found direct evidence of the
+% stress-driven ageing process in colloidal gels that is a direct consequence of the formation of a thin network with the help of hydrodynamic interactions in the first stages
+% of the aggregation process.
+% 
+% In Fig.~\ref{fig:stress}c we plot the time evolution of the structure factor, $S(q)$, for the state point
+% undergoing network reorganization $\phi=0.10$ and $c_p=0.82$. The time of network reorganization as found in Fig.~\ref{fig:stress}a and b
+% corresponds to a steep increase of the low wavenumber scattering from the network, that signals the coarsening of the network structure. 
+% 
+% The same conclusions also hold for the state point at $\phi=0.25$. We observe network restructuring only for the state point
+% at lower polymer concentration, $c_p=0.38$. Interestingly, this is also the only state point that undergoes crystallization,
+% meaning that the coarsening of the network of the dense liquid phase and the resulting formation of a wider liquid branch is a necessary condition for crystallization.
 
-In Fig.~\ref{fig:gel}a we show the cluster size distribution for all state points at their gelation point. The decay of the distribution
-follows different power laws $n(s)\sim s^\tau$, where $\tau$ is the Fisher exponent \cite{family2012kinetics}.
-For state points with $\phi=0.25$ the decay is well described by the three dimensional (3D) random percolation exponent $\tau=2.18$ \cite{family2012kinetics}.
-For state points with $\phi=0.10$, on the other hand, the decay appears to be slower, close to the diffusion limited cluster aggregation (DLCA) 
-exponent $\tau=1.8$ \cite{family2012kinetics}. While it is plausible that gelation at lower volume fraction occurs through cluster aggregation, especially given 
-the high polymer concentration of these state points, we have to warn that the cluster size distribution is very sensitive to
-experimental conditions and statistical noise (which is high in our case since the distributions are extracted from a single frame).
-Other measures of fractal dimensions (see below) will in fact suggest that also the $\phi=0.10$ states belong to the random gel universality class.
+\subsection*{Results}
 
-Next we show in Fig.~\ref{fig:gel}b the radius of gyration $R_{\rm g}$ for the clusters of size $s$ at the gel point. All clusters, irrespective
-of the state point, follow the power-law: $R_{\rm g}\sim s^{2/D}$, with $D=2.53$ the fractal dimension in the random gelation
-universality class \cite{family2012kinetics}. So far, we have thus proven that, in the initial stages of phase separation, the liquid branch is
-formed by colloidal particles that aggregate to form a random gel.
+\begin{figure*}[!t]
+ \centering
+ \includegraphics[width=8cm]{./oldfig/fig1b.pdf}
+ \includegraphics[width=8cm]{./oldfig/fig1b.pdf}
+ \includegraphics[width=8cm]{./oldfig/fig3c}
+ \includegraphics[width=8cm]{./oldfig/fig3b}
+ % fig1b.pdf: 687x479 pixel, 72dpi, 24.24x16.90 cm, bb=0 0 687 479
+ \caption{{\bf Different routes of viscoelastic phase separation.}
+ \textbf{a,} {\color{blue}Phase diagram.}
+ \textbf{b,} Radius of gyration ($R_g$) as a function of cluster size ($s$) for all state points at percolation. The
+ dashed line represents the exponent of the random gelation universality class.
+ \textbf{c,} Structure factors for the late stages of the gelation process. Calculation of the structure factor is done with the Hanning window function,  to minimize boundary effects.
+ \textbf{d,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle.
+ Different lines represent the different state points.
+ }
+ \label{fig:phase_separation}
+\end{figure*}
 
-Even after a gel is formed, the system keeps evolving.
-However, since a gel has an elasticity due to its space-spanning connectivity, the process should be markedly different from 
-ordinary fluid phase separation. Now we consider this unique feature of the coarsening of a gel network.
+All samples share the same early time stages of spinodal decomposition. First the colloidal particles start aggregating,
+eventually forming a percolating network. Then liquid-gas phase separation takes place, in which a dense network (liquid)
+coexists with freely diffusing monomers (gas). The topology of the network is similar to all our samples at the initial stages.
+This is shown for example in Fig.~\ref{fig:phase_separation}b, where we plot the radius of gyration of colloidal clusters ($R_g$)
+as a function of the cluster size ($s$) at the percolation point. Two particles belong to the same cluster if their distance is
+within the first peak of the radial distribution function. All samples show the same scaling law, which is compatible with
+the random gelation universality class exponent ($D=2.53$) as shown by the dashed line in Fig.~\ref{fig:phase_separation}b.
 
-\subsection*{Stress-driven ageing}
+While sharing similar structural properties at early times, the samples differ during the later stages of the phase separation process.
+To examin this in more detail, we plot in Fig.~\ref{fig:phase_separation}c the structure factor for the final stages of the gelation process for all state points.
+Calculation of the structure factor is done with the Hanning window function, to minimize boundary effects.
+The figure shows that at low wavenumber $q$, the structure factor displays fractal scaling compatible with Guinier law, $S(q)\sim q^{-D}$.
+But a difference
+in the fractal dimension $D$ between the $\phi=0.25$ and $c_p=0.38$ state point and other state points starts to become visible. In fact, while states
+with high polymer concentration retain the exponent $D=2.53$,
+which is the random gel universality class exponent we also observed in the early stages of the gelation process (see Fig.~\ref{fig:phase_separation}b),
+the state point with low polymer concentration ($\phi=0.25$ and $c_p=0.38$) displays the biggest deviation from that exponent. In the figure we also
+plot the exponent $D=3$, which corresponds to volume growth, which we'll show in the following analysis to be the correct exponent for this state.
+The difference is due to the onset of crystallization in the low-polymer concentration sample. This is already evident in the high $q$ behaviour of
+the structure factors (Fig.~\ref{fig:phase_separation}c), where the diffraction planes appear as sharper peaks for $\phi=0.25$ and $c_p=0.38$.
 
-Previous simulation results~\cite{tanaka2000,tanaka2007spontaneous,furukawa2010key} have suggested that hydrodynamic interactions play
-a key role in the process of colloidal gelation. Without hydrodynamic interactions, particles have the tendency to
-aggregate in compact structures and subsequently form thick network structures. This also limits the possibility to
-form arrested gel states only at relatively high volume fractions. With hydrodynamic interactions, particles first form
-a transient gel even at very low volume fractions, and only later they increase the number of
-nearest neighbours to minimize the energy of the structure. 
-Thus, hydrodynamic interactions lead to the formation of gels that are very far from equilibrium and under a strong thermodynamic driving force 
-towards more stable compact structures. 
-The resulting transition from open to more compact networks occurs only through
-the breaking of the bonds that have accumulated more stress. The process was called stress-driven ageing~\cite{tanaka2007spontaneous}, 
-which is characteristic of viscoelastic phase separation \cite{tanaka2000viscoelastic}. 
-Below we provide the first experimental proof of this process.
+To gain additional insight, in Fig.~\ref{fig:phase_separation}d we plot the average local volume fraction, computed
+through Voronoi diagrams, as a function of the number of neighbours. For all state points, we can see that the average volume fraction of a colloidal
+particle increases with its number of neighbours. Approaching close packing (12 neighbours), we clearly see two family of curves. The state point
+at $\phi=0.25$ and $c_p=0.38$ reaches close packing at a volume fraction of around $\sim 60\%$, which is in reasonable agreement with the
+volume fraction of the stable crystal phase. All other state point, instead, reach close packing at volume fraction $\sim 55\%$, which is
+close to the volume fraction of the glass state. 
+%(\textbf{the absolute values of these volume fraction are not totally convincing, probably because of a not optimal estimate of the size of the colloids.
+We should run the code of Mathieau with size detection}).
+The figure then clearly shows that the state point at $\phi=0.25$ and $c_p=0.38$ follows a different arrest mechanism, in which phase separation
+is arrested by crystallization and not glassiness. Although our estimation of the absolute volume fraction may involves errors of a few \%, the trend should be robust.
 
-In Fig.~\ref{fig:stress}a, we plot the fraction of liquid and gas particles as a function of time during phase separation
-at the state points of $\phi=0.10$. We observe that at high polymer concentration ($c_p=1.36$, dashed lines) both the
-liquid and gas branch change monotonically with time as the phase separation proceeds. At low polymer concentration
-($c_p=0.82$, continuous lines), on the other hand, we find a non-monotonic evolution of both the liquid and gas branches, which is signalling
-a structural reorganization of the network. To prove that this is indeed a stress-driven ageing process, in Fig.~\ref{fig:stress}b 
-we plot the average number of bond-breaking events for the low-polymer concentration state point. The figure clearly shows that
-the network restructuring follows a peak in the number of bond-breaking events. We have thus found direct evidence of the
-stress-driven ageing process in colloidal gels that is a direct consequence of the formation of a thin network with the help of hydrodynamic interactions in the first stages
-of the aggregation process.
+% 
+%  REINTREGRARE
+% while the crystallizing state point shows a scaling compatible with $D=3$, which
+% signals compact crystalline nuclei. At high $q$, the first signals from diffraction planes are also evident in the $\phi=0.25$ and $c_p=0.38$ state
+% point. This suggests that the space filling network is composed of crystalline aggregates, differently from the fractal branches which characterize the
+% spinodal liquid. Also this suggests that the growth mechanism of the crystal is not simply due to filling of the spinodal liquid network, as was
+% instead observed in Ref.~\cite{fortini2008crystallization,perez2011pathways} for brownian dynamics simulations.
 
-In Fig.~\ref{fig:stress}c we plot the time evolution of the structure factor, $S(q)$, for the state point
-undergoing network reorganization $\phi=0.10$ and $c_p=0.82$. The time of network reorganization as found in Fig.~\ref{fig:stress}a and b
-corresponds to a steep increase of the low wavenumber scattering from the network, that signals the coarsening of the network structure.
-
-The same conclusions also hold for the state point at $\phi=0.25$. We observe network restructuring only for the state point
-at lower polymer concentration, $c_p=0.38$. Interestingly, this is also the only state point that undergoes crystallization,
-meaning that the coarsening of the network of the dense liquid phase and the resulting formation of a wider liquid branch is a necessary condition for crystallization.
-
-\subsection*{Crystallization and arrest}
-
-\begin{figure}
- \centering
- \framebox{Phase diagram\hfill}
- \includegraphics[width=8cm]{oldfig/fig3b}
- \includegraphics[width=8cm]{oldfig/fig3c}
- % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
- \caption{{\bf Two types of arrest mechanism leading to gel formation.} \textbf{a,} Phase diagram \textbf{b,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle.
- Different lines represent the different state points. {\bf c,} Structure factors for the late stages of the gelation process.} 
- \label{fig:statepoints}
-\end{figure}
+% 
+% 
+% \begin{figure}
+%  \centering
+%  \framebox{Phase diagram\hfill}
+%  \includegraphics[width=8cm]{oldfig/fig3b}
+%  \includegraphics[width=8cm]{oldfig/fig3c}
+%  % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
+%  \caption{{\bf Two types of arrest mechanism leading to gel formation.} \textbf{a,} Phase diagram \textbf{b,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle.
+%  Different lines represent the different state points. {\bf c,} Structure factors for the late stages of the gelation process.} 
+%  \label{fig:statepoints}
+% \end{figure}
 
@@
 
-After the creation of the spinodal liquid branch, crystallization events start occurring for the state point $\phi=0.25$ and $c_p=0.38$, which has the lowest
+To confirm the formation of crystalline seeds we adopt bond orientational analysis to detect which particles are in crystalline environments~\cite{russo2012microscopic}.
+With this analysis we confirm that crystallization events start occurring for the state point $\phi=0.25$ and $c_p=0.38$, which has the lowest
 polymer concentration of all state points. This is shown in Fig.~\ref{fig:crystals}a where the black line indicates the average size of the crystallites,
@@
 %\end{itemize}
-All other state points show also signs of crystallization, but in negligible amounts. In particular, for state points with $\phi=0.25$, increasing
+All other state points show only negligible signs of crystallization. In particular, for state points with $\phi=0.25$, increasing
 polymer concentration drastically reduces the amount of crystals. This is in agreement with the
@@
 
-To show the presence of two different arrest mechanisms, we show in Fig.~\ref{fig:crystals}b the average local volume fraction, computed
-through Voronoi diagrams, as a function of the number of neighbours. For all state points, we can see that the average volume fraction of a colloidal
-particle increases with its number of neighbours. Approaching close packing (12 neighbours), we clearly see two family of curves. The state point
-at $\phi=0.25$ and $c_p=0.38$ reaches close packing at a volume fraction of around $\sim 60\%$, which is in reasonable agreement with the
-volume fraction of the stable crystal phase. All other state point, instead, reach close packing at volume fraction $\sim 55\%$, which is
-close to the volume fraction of the glass state. 
-%(\textbf{the absolute values of these volume fraction are not totally convincing, probably because of a not optimal estimate of the size of the colloids.
-We should run the code of Mathieau with size detection}).
-The figure then clearly shows that the state point at $\phi=0.25$ and $c_p=0.38$ follows a different arrest mechanism, in which phase separation
-is arrested by crystallization and not glassiness. Although our estimation of the absolute volume fraction may involves errors of a few \%, the trend should be robust.
-
-To confirm the different arrest mechanisms, crystallization and vitrification, we compute the structure factor for the final stages of the gelation process for all state points,
-as shown in Fig.~\ref{fig:crystals}c. Calculation of the structure factor is done with the Hanning window function, to minimize boundary effects.
-The figure shows that at low wavenumber $q$, the structure factor displays fractal scaling compatible with Guinier law, $S(q)\sim q^{-D}$. But a difference
-in the fractal dimension $D$ between the $\phi=0.25$ and $c_p=0.38$ state point and other state points is visible. Non-crystallizing state
-points have $D=2.53$, of the random gel universality class, while the crystallizing state point shows a scaling compatible with $D=3$, which
-signals compact crystalline nuclei. At high $q$, the first signals from diffraction planes are also evident in the $\phi=0.25$ and $c_p=0.38$ state
-point. This suggests that the space filling network is composed of crystalline aggregates, differently from the fractal branches which characterize the
-spinodal liquid. Also this suggests that the growth mechanism of the crystal is not simply due to filling of the spinodal liquid network, as was
-instead observed in Ref.~\cite{fortini2008crystallization,perez2011pathways} for brownian dynamics simulations.
-
-To investigate this process in more detail, we plot in Fig.~\ref{fig:crystals}c the gyration radius of individual crystalline nuclei for the state point
+While nucleation always occurs inside the liquid branches of the phase separating network, we suggest here the presence of a new growth channel
+which was not observed before.
+In the following we will show that the growth mechanism of the crystal is not only due to filling of the spinodal liquid network, as was observed in Ref.~\cite{fortini2008crystallization,perez2011pathways} for brownian dynamics simulations, but involves direct condensation of the gas phase
+on the crystalline seeds.
+To investigate this process, we plot in Fig.~\ref{fig:crystals}c the gyration radius of individual crystalline nuclei for the state point
 $\phi=0.25$ and $c_p=0.38$. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with
-the dimension of random percolation $D=2.53$, while at large sizes it scales as $D=3$ of compact crystal growth. The figure thus shows that
-small crystalline nuclei are nucleated inside the phase-separated liquid branches; but once they reach the transverse size of the liquid branch, the growth is continued
-by direct addition of colloidal particles from the gas.
+the dimension of random percolation $D=2.53$, while at large sizes it scales as $D=3$ of compact crystal growth. This demonstrates that, while
+small crystalline nuclei are nucleated inside the phase-separated liquid branches, once they reach the transverse size of the liquid branch, the growth
+follows a volume expansion with the formation of isotropic droplets (colloidal rain). The same scaling law was suggested in the analysis of the
+structure factors, Fig.~\ref{fig:phase_separation}c, but it is shown directly by the analys of the gyration radius.
 
@@
 
-Now we show in Fig.~\ref{fig:transitions}a slabs of 10 particle diameter thickness for configurations at $\phi=0.25$ and $c_p=0.38$ at different times.
+In the next section we demonstrate that the volume growth is due to direct condensation of the gas phase on the crystalline seeds.
+First we show in Fig.~\ref{fig:transitions}a slabs of 10 particle diameter thickness for configurations at $\phi=0.25$ and $c_p=0.38$ at different times.
 Particles are coloured according to their phase. Green: HCP; Red: FCC; blue: BCC; white: liquid or gas. The figure shows the first nucleation
@@
 The first stages of gelation always involve spinodal decomposition with the formation
-of a fractal network corresponding to the liquid phase.
-Due to hydrodynamic interaction, the network is locally stretched, and the accumulation of mechanical stress on some bonds causes some
-of them to break and the network to reorganize in a more compact and larger structure. This occurs only at low polymer concentration,
-while at high polymer concentrations the bonds are too strong to break. This will eventually dictate which arrest mechanism
-is available to the system.
+of a fractal network and gas-liquid phase separation.
+% Due to hydrodynamic interaction, the network is locally stretched, and the accumulation of mechanical stress on some bonds causes some
+% of them to break and the network to reorganize in a more compact and larger structure. This occurs only at low polymer concentration,
+% while at high polymer concentrations the bonds are too strong to break. This will eventually dictate which arrest mechanism
+% is available to the system.
 Depending on the polymer concentration, there are two possible arrest mechanisms.
 (i) Crystallization, observed at $\phi=0.25$ and $c_p=0.38$, small crystalline nuclei appear inside the liquid, reach the surface
- of the liquid branches, and then grow by addition of particles from the gas phase. The final structure is a network of crystal aggregates,
+ of the liquid branches, and then grow by addition of particles from the gas phase. The final structure is a network of crystal droplets,
  as confirmed by the fractal dimension of the branches, the volume fraction of the particles within the branches, and bond orientational analysis.
@@
 
-The process of crystal-gel formation is akin to the Bergeron process which is the primary mechanism for the formation of raindrops in clouds~\cite{glickman2000glossary}.
+The process we observed is akin to the Bergeron process which is the primary mechanism for the formation of raindrops in clouds~\cite{glickman2000glossary}.
 In clouds there is a mixture of ice crystals and supercooled water. The vapor phase is in coexistence with the liquid phase, but is supersaturated

### 2015-08-07 crystal_gel_letter.tex
Abstract centred on colloidal rain space -> microgravity -process proceeds

@@
 {
-With confocal measurements of a PMMA colloid and polymer mixture, we experimentally access the formation of a novel arrested state,
-in which the dense branches of a gel are replaced by crystal aggregates~\cite{fortini2008crystallization,perez2011pathways,sabin2012}, which then grow into droplets by direct condensation of the gas phase on the surface of the crystals.
-We call this new state \emph{colloidal rain}, as it represents the colloidal analogue of rain formation from ice crystals that occurs in mixed phase clouds (the Bergeron process)~\cite{glickman2000glossary}.
-For the first time we are able to observe the dynamics of the system at single-particle level, giving full experimental access to the entire kinetic
-pathway of droplet formation. The formation of colloidal rain occurs at low polymer concentrations, close to the metastable critical point~\cite{ten1997enhancement}, where
-the strength of the interparticle attractions is low enough to allow bond reorganization and nucleation. For high polymer concentrations the system
-forms colloidal gels~\cite{lu2008gelation}, 
-where viscoelastic phase separation is interrupted by the glass transition.
-We thus argue that similar states can form as the result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems, whose dynamics is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
+Colloidal gels form when spinodal phase separation is interrupted by the glass transition~\cite{lu2008gelation}.
+It has been shown that crystallization can equivalently interrupts the phase separation by freezing the existing network.
+However the particle-level details of this process have not yet been confirmed experimentally.
+Here we show by confocal measurements that crystals nucleate inside the dense branches of a gel, and grow by evaporation of the liquid network and direct condensation of the gas phase on the surface of the crystals.
+This process represents the colloidal analogue of the Bergeron process~\cite{glickman2000glossary}, i.e.
+ice crystals formation that occurs in mixed phase clouds and is at the origin of rain.
+Our \emph{colloidal rain} gives us full experimental access to the kinetic pathway of this phenomenon important for climate science.
+We argue that states similar to colloidal rain can form as the result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems, whose dynamics is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
 }
@@
 Indeed the possibility of a different class of gels, which are stabilised by crystallization, 
-was suggested by numerical simulations~\cite{fortini2008crystallization,perez2011pathways} and observed by space experiments \cite{sabin2012}.
+was suggested by numerical simulations~\cite{fortini2008crystallization,perez2011pathways} and observed in microgravity experiments \cite{sabin2012}.
 For colloidal systems, we can in principle access structural evolution in real time at a single-particle level; however, there has so far been no confocal microscopy  
@@
 expense of the supercooled liquid droplets due to their lower saturated vapour pressure. In our colloidal system, we then call this process ``colloidal rain''
-formation.
+formation. Due to the difference in the nucleation and growth processes, the droplets display different scaling laws depending on their
+size. For small nuclei, the droplets have the same fractal dimension of the embedding gel network. But during the growth process, droplets
+grow with an integer fractal dimension due to direct condensation of gas particles.
+This is in contrast to the standard crystal-gel scenario, where crystals nucleate and grow inside the dense branches of the gel, becoming dynamically arrested.
+We also observe that increasing the strength of the interparticle attraction
+(by increasing the polymer concentration), instead of forming colloidal rain, the system undergoes the classical gelation process, in which
+phase separation is interrupted by vitrification~\cite{lu2008gelation}.
 
-\subsection{Experiments}
+\subsection*{Experiments}
 
@@
 which are then subject to attractive depletion forces, making the system thermodynamically unstable and leading to phase separation (see Methods).
-In the following we discuss all the stages of the formation of the crystal gels: phase separation, stress-driven agein and crystallization.
+%In the following we discuss all the stages of the formation of the crystal gels: phase separation, stress-driven agein and crystallization.
 The experimental data is taken at two different volume fractions ($\phi=0.10$ and $\phi=0.25$) and for different values
 of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ for $\phi=0.25$ and $c_p=0.82,1.36$ for $\phi=0.10$.
+In Fig.XXX we superimpose the state points with the phase diagram calculated from the YYY equation of state.
 For each state points, scans at early times (every $10s$) and at the late stage (every $30s$) of the gelation process.
@@
 
-\subsection*{Crystallization and arrest}
+\subsection*{Results}
 
-\begin{figure}
+\begin{figure*}[!t]
  \centering
- \framebox{Phase diagram\hfill}
- \includegraphics[width=8cm]{oldfig/fig3b}
- \includegraphics[width=8cm]{oldfig/fig3c}
- % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
- \caption{{\bf Two types of arrest mechanism leading to gel formation.} \textbf{a,} Phase diagram \textbf{b,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle.
- Different lines represent the different state points. {\bf c,} Structure factors for the late stages of the gelation process.} 
- \label{fig:statepoints}
-\end{figure}
+ \includegraphics[width=8cm]{./oldfig/fig1b.pdf}
+ \includegraphics[width=8cm]{./oldfig/fig1b.pdf}
+ \includegraphics[width=8cm]{./oldfig/fig3c}
+ \includegraphics[width=8cm]{./oldfig/fig3b}
+ % fig1b.pdf: 687x479 pixel, 72dpi, 24.24x16.90 cm, bb=0 0 687 479
+ \caption{{\bf Different routes of viscoelastic phase separation.}
+ \textbf{a,} {\color{blue}Phase diagram.}
+ \textbf{b,} Radius of gyration ($R_g$) as a function of cluster size ($s$) for all state points at percolation. The
+ dashed line represents the exponent of the random gelation universality class.
+ \textbf{c,} Structure factors for the late stages of the gelation process. Calculation of the structure factor is done with the Hanning window function,  to minimize boundary effects.
+ \textbf{d,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle.
+ Different lines represent the different state points.
+ }
+ \label{fig:phase_separation}
+\end{figure*}
+
+All samples share the same early time stages of spinodal decomposition. First the colloidal particles start aggregating,
+eventually forming a percolating network. Then liquid-gas phase separation takes place, in which a dense network (liquid)
+coexists with freely diffusing monomers (gas). The topology of the network is similar to all our samples at the initial stages.
+This is shown for example in Fig.~\ref{fig:phase_separation}b, where we plot the radius of gyration of colloidal clusters ($R_g$)
+as a function of the cluster size ($s$) at the percolation point. Two particles belong to the same cluster if their distance is
+within the first peak of the radial distribution function. All samples show the same scaling law, which is compatible with
+the random gelation universality class exponent ($D=2.53$) as shown by the dashed line in Fig.~\ref{fig:phase_separation}b.
+
+While sharing similar structural properties at early times, the samples differ during the later stages of the phase separation process.
+To examin this in more detail, we plot in Fig.~\ref{fig:phase_separation}c the structure factor for the final stages of the gelation process for all state points.
+Calculation of the structure factor is done with the Hanning window function, to minimize boundary effects.
+The figure shows that at low wavenumber $q$, the structure factor displays fractal scaling compatible with Guinier law, $S(q)\sim q^{-D}$.
+But a difference
+in the fractal dimension $D$ between the $\phi=0.25$ and $c_p=0.38$ state point and other state points starts to become visible. In fact, while states
+with high polymer concentration retain the exponent $D=2.53$,
+which is the random gel universality class exponent we also observed in the early stages of the gelation process (see Fig.~\ref{fig:phase_separation}b),
+the state point with low polymer concentration ($\phi=0.25$ and $c_p=0.38$) displays the biggest deviation from that exponent. In the figure we also
+plot the exponent $D=3$, which corresponds to volume growth, which we'll show in the following analysis to be the correct exponent for this state.
+The difference is due to the onset of crystallization in the low-polymer concentration sample. This is already evident in the high $q$ behaviour of
+the structure factors (Fig.~\ref{fig:phase_separation}c), where the diffraction planes appear as sharper peaks for $\phi=0.25$ and $c_p=0.38$.
+
+To gain additional insight, in Fig.~\ref{fig:phase_separation}d we plot the average local volume fraction, computed
+through Voronoi diagrams, as a function of the number of neighbours. For all state points, we can see that the average volume fraction of a colloidal
+particle increases with its number of neighbours. Approaching close packing (12 neighbours), we clearly see two family of curves. The state point
+at $\phi=0.25$ and $c_p=0.38$ reaches close packing at a volume fraction of around $\sim 60\%$, which is in reasonable agreement with the
+volume fraction of the stable crystal phase. All other state point, instead, reach close packing at volume fraction $\sim 55\%$, which is
+close to the volume fraction of the glass state. 
+%(\textbf{the absolute values of these volume fraction are not totally convincing, probably because of a not optimal estimate of the size of the colloids.
+We should run the code of Mathieau with size detection}).
+The figure then clearly shows that the state point at $\phi=0.25$ and $c_p=0.38$ follows a different arrest mechanism, in which phase separation
+is arrested by crystallization and not glassiness. Although our estimation of the absolute volume fraction may involves errors of a few \%, the trend should be robust.
+
+% 
+%  REINTREGRARE
+% while the crystallizing state point shows a scaling compatible with $D=3$, which
+% signals compact crystalline nuclei. At high $q$, the first signals from diffraction planes are also evident in the $\phi=0.25$ and $c_p=0.38$ state
+% point. This suggests that the space filling network is composed of crystalline aggregates, differently from the fractal branches which characterize the
+% spinodal liquid. Also this suggests that the growth mechanism of the crystal is not simply due to filling of the spinodal liquid network, as was
+% instead observed in Ref.~\cite{fortini2008crystallization,perez2011pathways} for brownian dynamics simulations.
+
+% 
+% 
+% \begin{figure}
+%  \centering
+%  \framebox{Phase diagram\hfill}
+%  \includegraphics[width=8cm]{oldfig/fig3b}
+%  \includegraphics[width=8cm]{oldfig/fig3c}
+%  % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
+%  \caption{{\bf Two types of arrest mechanism leading to gel formation.} \textbf{a,} Phase diagram \textbf{b,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle.
+%  Different lines represent the different state points. {\bf c,} Structure factors for the late stages of the gelation process.} 
+%  \label{fig:statepoints}
+% \end{figure}
 
@@
 
-After the creation of the spinodal liquid branch, crystallization events start occurring for the state point $\phi=0.25$ and $c_p=0.38$, which has the lowest
-polymer concentration of all state points. This is shown in Fig.~\ref{fig:statepoints}a where the black line indicates the average size of the crystallites,
+To confirm the formation of crystalline seeds we adopt bond orientational analysis to detect which particles are in crystalline environments~\cite{russo2012microscopic}.
+With this analysis we confirm that crystallization events start occurring for the state point $\phi=0.25$ and $c_p=0.38$, which has the lowest
+polymer concentration of all state points. This is shown in Fig.~\ref{fig:crystals}a where the black line indicates the average size of the crystallites,
 while the red line the average number of crystallites as a function of time. The number of crystallites first rapidly increases as nucleation events
@@
 
-We can thus conclude that the crystallization process proceeds via two distinct mechanisms:
+We can thus conclude that the crystallization proceeds via two distinct mechanisms:
 (i) freezing of the fluid; (ii) evaporation of the fluid and condensation of the vapor on the crystals.

### 2015-08-07 crystal_gel_letter.tex
network structure different from the original phase separation pattern

@@
 Colloidal gels form when spinodal phase separation is interrupted by the glass transition~\cite{lu2008gelation}.
-It has been shown that crystallization can equivalently interrupts the phase separation by freezing the existing network.
+It has been shown that crystallization may equivalently interrupt the phase separation by freezing the dense phase.
 However the particle-level details of this process have not yet been confirmed experimentally.
-Here we show by confocal measurements that crystals nucleate inside the dense branches of a gel, and grow by evaporation of the liquid network and direct condensation of the gas phase on the surface of the crystals.
+Here we show by confocal microscopy measurements that crystallisation leads to a network structure different from the original phase separation pattern.
+Crystals nucleate inside the liquid network, and grow past it by direct condensation of the gas phase on their surface, driving evaporation of nearby liquid.
 This process represents the colloidal analogue of the Bergeron process~\cite{glickman2000glossary}, i.e.

### 2015-08-07 crystal-gel/crystal_gel_letter.tex
network-forming spinodal

@@
 {
-With confocal measurements of a PMMA colloid and polymer mixture, we experimentally access the formation of a novel arrested state,
-formed by crystal droplets nucleating inside the dense branches of a gel, and growing by direct condensation of the gas phase on the surface of the crystals.
-% in which the dense branches of a gel are replaced by crystal aggregates~\cite{fortini2008crystallization,perez2011pathways,sabin2012}, which then grow into droplets by direct condensation of the gas phase on the surface of the crystals.
-We call this new state \emph{colloidal rain}, as it represents the colloidal analogue of rain formation from ice crystals that occurs in mixed phase clouds (the Bergeron process)~\cite{glickman2000glossary}.
-We observe the dynamics of the system at single-particle level, giving for the first time full experimental access to the kinetic
-pathway of droplet formation. The formation of colloidal rain occurs at low polymer concentrations, close to the metastable critical point~\cite{ten1997enhancement}, where
-the strength of the interparticle attractions is low enough to allow bond reorganization and nucleation. For high polymer concentrations, instead, the system
-forms colloidal gels~\cite{lu2008gelation}, 
-where viscoelastic phase separation is interrupted by the glass transition.
-We thus argue that states similar to colloidal rain can form as the result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems, whose dynamics is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
+Colloidal gels form when network-forming spinodal phase separation is interrupted by the glass transition~\cite{lu2008gelation}.
+It has been shown that crystallization may equivalently interrupt the phase separation by freezing the dense phase.
+However the particle-level details of this process have not yet been confirmed experimentally.
+Here we show by confocal microscopy measurements that crystallisation leads to a network structure different from the original phase separation pattern.
+Crystals nucleate inside the liquid network, and grow past it by direct condensation of the gas phase on their surface, driving evaporation of nearby liquid.
+This process represents the colloidal analogue of the Bergeron process~\cite{glickman2000glossary}, i.e.
+ice crystals formation that occurs in mixed phase clouds and is at the origin of rain.
+Our \emph{colloidal rain} gives us full experimental access to the kinetic pathway of this phenomenon important for climate science.
+We argue that states similar to colloidal rain can form as the result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems, whose dynamics is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
 }

### 2015-08-10 crystal_gel_letter.tex
Methods section with all necessary experimental details

@@
 \usepackage{comment}
+\usepackage{siunitx}
 %\usepackage{dcolumn}% Align table columns on decimal point
@@
 {
-Colloidal gels form when spinodal phase separation is interrupted by the glass transition~\cite{lu2008gelation}.
+Colloidal gels form when network-forming spinodal phase separation is interrupted by the glass transition~\cite{lu2008gelation}.
 It has been shown that crystallization may equivalently interrupt the phase separation by freezing the dense phase.
@@
 
-\vspace{2cm}
-\noindent
-{\bf \large Methods}
+\section*{Methods}
 
-\noindent
-{\bf Experimental.}
+\paragraph*{Experimental}
+We use \textsc{pmma} (poly(methyl methacrylate)) colloids sterically stabilized with methacryloxypropyl terminated \textsc{pdms}(poly(dimethyl siloxane)) and fluorescently labelled with rhodamine isothiocyanate chemically bonded to the \textsc{pmma}.
+We disperse the particles in refractive index and density matching mixture of cis-decalin (Tokyo Kasei) and bromocyclohexane (Sigma-Aldrich).
 
-\noindent
-{\bf State points studied. }
-The experimental data is taken at two different volume fractions ($\phi=0.10$ and $\phi=0.25$) and for different values
-of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ for $\phi=0.25$ and $c_p=0.82,1.36$ for $\phi=0.10$.
-For each state points, scans at early times (every $10s$) and at the late stage (every $30s$) of the gelation process.
+To induce short-ranged depletion attraction, we use polystyrene (TOSOH) of molecular weight \SI{3.8}{\mega\dalton} as non-adsorbing polymer.
+Experiments are conducted at \SI{27}{\celsius}, some \SI{80}{\celsius} above the theta temperature in this solvents mixture~\cite{Royall2007}.
+A Flory scaling of the measurements of~\cite{lu2008gelation} yields a radius of gyration $R_g=\SI{76}{\nano\metre}$.
 
-We will consider two particles bonded if their center-to-center distance is within $12$ pixels. The radius of a colloidal
-particle is $5.365$ pixels.
+In the absence of salt, the Debye length is expected to reach several \si{\micro\metre} and the (weakly) charged colloids experience a long range electrostatic repulsion.
+We confirm that colloids never come close enough to feel the short-ranged attraction.
+Screening by tetrabutylammonium bromide (Fluka) at saturated concentration brings down the Debye length to about \SI{100}{\nano\metre}, practically discarding the repulsion.
 
-\noindent
-{\bf Structural analysis}
+The data are collected on a Leica SP5 confocal microscope, using \SI{532}{\nano\metre} laser excitation.
+We assess that the size distribution of our particle is Gaussian with a polydispersity below 5\% via direct confocal measurements~\cite{Leocmach2013}.
+The mean diameter is $\sigma=\SI{2.1}{\micro\metre}$ XXX to be checked XXX yielding a size ratio $q_R=2R_g/\sigma=0.07$.
+The Brownian time is $\tau_B \approx \SI{4.5}{\second}$.
+To be able to follow individual trajectories, we perform a 3D scan every \SI{10}{\second} at early time and every \SI{30}{\second} later.
+
+%\paragraph*{State points studied}
+%The experimental data is taken at two different volume fractions ($\phi=0.10$ and $\phi=0.25$) and for different values
+%of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ for $\phi=0.25$ and $c_p=0.82,1.36$ for $\phi=0.10$.
+%For each state points, scans at early times (every $10s$) and at the late stage (every $30s$) of the gelation process.
+%
+%We will consider two particles bonded if their center-to-center distance is within $12$ pixels. The radius of a colloidal
+%particle is $5.365$ pixels.
+
+\paragraph*{Structural analysis}
+To account for imprecision in particle localisation close to contact, we consider two particles bonded if their center-to-center distance is within $\SI{2.3}{\micro\metre}>\sigma+2R_g$.
+
 To detect percolation, two different methods have been employed. In the first one we detect percolation

### 2015-09-14 phasediag.py
phase diagram

@@ -1,0 +1,95 @@
+from matplotlib.pyplot import *
+from colloids.phase import *
+from os.path import isfile
+
+qR = 0.071
+cpov = 1
+q = qR2q(qR)
+
+#experimental points
+phixp = np.array([0.25]*4+[0.1]*2)
+cpxp = np.array([0.38, 0.48, 0.57, 1.07, 0.82, 1.36])
+
+#load or compute theoretical phase diagram
+fc, pivc = Liu().critical_point(q)
+if isfile('phasediag.txt'):
+    GL = np.loadtxt('phasediag.txt')
+else:
+    GL = Liu().all_GL(q, maxpiv=3500)
+    np.savetxt('phasediag.txt', GL)
+
+
+
+fig = figure('experimental phase diagram')
+clf()
+
+#draw experimental points
+scatter(phixp, cpxp, marker='s')
+
+#draw critical point
+scatter(f2vf(fc), piv2y(pivc, q) * alpha(fc, qR) * cpov, c='r', marker='o')
+
+#draw binodal and spinodals
+for f,l in zip(GL.T[1:], ['--']*2+['-']*2):
+    cp = piv2y(GL[:,0], q) * alpha(f, qR)
+    plot(f2vf(f), cp * cpov, 'k'+l)
+#draw tie lines
+for piv, fG, fL in GL[[50,60,70,80,85,88]+range(90,98),:3]:
+    cpG = piv2y(piv, q) * alpha(fG, qR) * cpov
+    cpL = piv2y(piv, q) * alpha(fL, qR) * cpov
+    plot(f2vf(np.array([fG,fL])), [cpG, cpL], '-', color=(0.5,0.5,0.5))
+
+#draw equicomposition lines
+for x in np.arange(1,10)/10.:
+    phiG, phiL = f2vf(GL[:,1:3]).T
+    phi = x*phiG+(1-x)*phiL
+    cp = piv2y(GL[:,0], q) * alpha(vf2f(phi), qR)
+    plot(phi, cp * cpov, '-', color=(0.5,0.5,0.5))
+
+ylabel(r'$c_p$ (g/L)')
+xlabel(r'$\phi$ (%)')
+ylim(0,2)
+xlim(0,0.65)
+draw()
+savefig('phasediag.pdf')
+
+#save binodal and spinodal
+for f,l in zip(GL.T[1:], ['bg', 'bl', 'sg', 'sl']):
[…]

### 2015-09-15 phasediag.py
Add fluid-crystal and subfigure labels

@@ -17,7 +17,11 @@ if isfile('phasediag.txt'):
 else:
     GL = Liu().all_GL(q, maxpiv=3500)
     np.savetxt('phasediag.txt', GL)
-
+if isfile('phasediagFS.txt'):
+    FS = np.loadtxt('phasediagFS.txt')
+else:
+    FS = all_coexistence(q, Liu(), Hall(), maxpiv=3500)
+    np.savetxt('phasediagFS.txt', FS)
 
 
 fig = figure('experimental phase diagram')
@@ -45,11 +49,16 @@ for x in np.arange(1,10)/10.:
     phi = x*phiG+(1-x)*phiL
     cp = piv2y(GL[:,0], q) * alpha(vf2f(phi), qR)
     plot(phi, cp * cpov, '-', color=(0.5,0.5,0.5))
+    
+#draw fluid-crystal
+for f in FS.T[1:]:
+    cp = piv2y(FS[:,0], q) * alpha(f, qR)
+    plot(f2vf(f), cp * cpov, 'k.')
 
 ylabel(r'$c_p$ (g/L)')
 xlabel(r'$\phi$ (%)')
 ylim(0,2)
-xlim(0,0.65)
+xlim(0,0.75)
 draw()
 savefig('phasediag.pdf')
 
@@ -61,7 +70,14 @@ for f,l in zip(GL.T[1:], ['bg', 'bl', 'sg', 'sl']):
         np.column_stack((f2vf(f), cp * cpov)),
         header='phi\tcp',
         )
-
+#save fluid-crystal
+for f,l in zip(FS.T[1:], ['f', 'x']):
+    cp = piv2y(FS[:,0], q) * alpha(f, qR)
+    np.savetxt(
+        'fluidcrystal_%s.phd'%l,
+        np.column_stack((f2vf(f), cp * cpov)),
+        header='phi\tcp',
+        )
 
 fig = figure('theoretical phase diagram')
 clf()

### 2015-09-15 crystal-gel/crystal_gel_letter.tex
Include fig:phase_separation

@@
  \centering
- \includegraphics[width=8cm]{./oldfig/fig1b.pdf}
- \includegraphics[width=8cm]{./oldfig/fig1b.pdf}
- \includegraphics[width=8cm]{./oldfig/fig3c}
- \includegraphics[width=8cm]{./oldfig/fig3b}
- % fig1b.pdf: 687x479 pixel, 72dpi, 24.24x16.90 cm, bb=0 0 687 479
+ \includegraphics{./fig/phase_separation.pdf}
  \caption{{\bf Different routes of viscoelastic phase separation.}
- \textbf{a,} {\color{blue}Phase diagram.}
+ \textbf{a,} Phase diagram.
+Dashed lines are fluid-solid equilibrium coexistence boundaries.
+The continuous line is the gas side of the metastable gas-liquid spinodal region with the black diamond marking the critical point.
+The other symbols are the experimental points, consistently used in all panels.
  \textbf{b,} Radius of gyration ($R_g$) as a function of cluster size ($s$) for all state points at percolation. The
- dashed line represents the exponent of the random gelation universality class.
- \textbf{c,} Structure factors for the late stages of the gelation process. Calculation of the structure factor is done with the Hanning window function,  to minimize boundary effects.
- \textbf{d,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle.
- Different lines represent the different state points.
+ gray line represents the fractal dimension of the random gelation universality class.
+ \textbf{c,} Structure factors for the late stages of the gelation process. The gray and black lines represents respectively random gelation and compact fractal dimensions.
+ \textbf{d,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle \textcolor{blue}{at late stage?}.
+Horizontal lines stress the value at 12 neighbours.
  }

### 2015-09-23 crystal_gel_letter.tex
insert fig:characterisation

@@
  \centering
- \includegraphics[width=8cm]{oldfig/fig3a}
- \includegraphics[width=8cm]{oldfig/fig3d}
- % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
- \caption{{\bf Characterisation of the crystal-gel.} {\bf a,} Time evolution of the average size of the crystals (black line) and the number of crystals (red line) for the state point $\phi=0.25$ and $c_p=0.38$.
+ \includegraphics{fig/characterisation.pdf}
+ \caption{{\bf Characterisation of the crystal-gel.} {\bf a,} Time evolution of the average size of the crystals (blue line) and the number of crystals (red line) for the state point $\phi=0.25$ and $c_p=0.38$.
 {\bf d,} Radius of gyration of crystalline nuclei at the late stages of gelation for the same state point.}

### 2015-09-23 data/slices.py
Generate 3D views (slices) for povray

@@ -1,0 +1,23 @@
+from colloids.povray import *
+import subprocess
+
+for t in [120, 390, 890]:
+    pos = np.loadtxt('pos_%03d'%t, skiprows=1)
+    sl = (pos[:,-1]>100) & (pos[:,-1]<150)
+    pos = pos[sl]
+    phase = np.loadtxt('studyphase_%03d'%t, dtype=int, usecols=[1])[sl]
+    
+    f = File("slice_t%03d.pov"%t,"colors.inc", "header.inc")
+    crystals = Union(*[
+        Sphere( (x,y,z), 5)
+        for x,y,z in pos[phase<3]
+        ]+[Texture(Pigment(color=(0.5,0,0)))])
+    f.write(crystals)
+    for i, c in zip([3,4,5], [(1,0.33,0), (1,0.66,0), (0.5,0,1)]):
+        f.write(Union(*[
+            Sphere( (x,y,z), 5)
+            for x,y,z in pos[phase==i]
+            ]+[Texture(Pigment(color=c))
+            ]))
+    f.file.flush()
+    subprocess.call(['povray', '-Oslice_t%03d.png'%t, '+W512', '+H512' , 'slice_t%03d.pov'%t])

### 2015-09-23 crystal_gel_letter.tex
fig:crystal

@@
  \centering
- \includegraphics[width=16cm]{oldfig/fig4a}
- \includegraphics[width=8cm]{oldfig/fig4b}
- \includegraphics[width=8cm]{oldfig/fig4c}
- % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735
+ \includegraphics{fig/crystal.pdf}
  \caption{
 {\bf Crystal gel formation.} 
-\textbf{a,} Snapshots of configurations at $\phi=0.25$ and $c_p=0.38$, for t=4856 (left), 12956 (center), 27956 (right).
-Particles are colored according to their phase.
-Green: HCP; Red: FCC; blue: BCC; white: liquid or gas.
- \textbf{b,} Fraction of particles for each phase (between bcc, hcp, fcc, surface, liquid or gas) at $\phi=0.25$ and $c_p=0.38$.
- Crystalline phases are detected through bond orientational analysis~\cite{russo2012microscopic}; surface particles are particles
- with at least two crystalline neighbours; liquid particles have at least 4 neighbours; gas particles are particles with less than 4
- nearest neighbours.
-\textbf{c,} Transition probabilities for the same process as in \textbf{b}. Crystal includes all crystalline phases (bcc, hcp, fcc) and surface particles.}
+\textbf{a,} Snapshots of configurations at $\phi=0.25$ and $c_p=0.38$. Particles are coloured according to their phase. Dark red: crystal; orange: surface of crystal; yellow: liquid; blue: gas. 
+ \textbf{b,} Fraction of particles for each phase.
+\textbf{c,} Transition probabilities for the same process as in \textbf{b}. Here crystal includes surface particles.}
  \label{fig:transitions}
@@
 First we show in Fig.~\ref{fig:transitions}a slabs of 10 particle diameter thickness for configurations at $\phi=0.25$ and $c_p=0.38$ at different times.
-Particles are coloured according to their phase. Green: HCP; Red: FCC; blue: BCC; white: liquid or gas. The figure shows the first nucleation
+Particles are coloured according to their phase. The figure shows the first nucleation
 events inside the liquid network (left panel); then these crystalline nuclei grow directly from the gas phase (middle panel), and finally
@@
 
+Crystalline phases are detected through bond orientational analysis~\cite{russo2012microscopic}, they are mainly FCC and HCP in comparable proportions.
+Surface particles are particles with at least two crystalline neighbours; liquid particles have at least 4 neighbours; gas particles are particles with less than 4 nearest neighbours.
+
 \bibliographystyle{naturemag4}

### 2015-09-23 crystal-gel/crystal_gel_letter.tex
fig:crystal Reconstruction from confocal coordinates. Gray vertical lines

@@
  \centering
- \includegraphics[width=16cm]{oldfig/fig4a}
- \includegraphics[width=8cm]{oldfig/fig4b}
- \includegraphics[width=8cm]{oldfig/fig4c}
- % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735
+ \includegraphics{fig/crystal.pdf}
  \caption{
 {\bf Crystal gel formation.} 
-\textbf{a,} Snapshots of configurations at $\phi=0.25$ and $c_p=0.38$, for t=4856 (left), 12956 (center), 27956 (right).
-Particles are colored according to their phase.
-Green: HCP; Red: FCC; blue: BCC; white: liquid or gas.
- \textbf{b,} Fraction of particles for each phase (between bcc, hcp, fcc, surface, liquid or gas) at $\phi=0.25$ and $c_p=0.38$.
- Crystalline phases are detected through bond orientational analysis~\cite{russo2012microscopic}; surface particles are particles
- with at least two crystalline neighbours; liquid particles have at least 4 neighbours; gas particles are particles with less than 4
- nearest neighbours.
-\textbf{c,} Transition probabilities for the same process as in \textbf{b}. Crystal includes all crystalline phases (bcc, hcp, fcc) and surface particles.}
+\textbf{a,} Reconstructions from confocal coordinates at $\phi=0.25$ and $c_p=0.38$.
+Depth of view is $5\sigma$.
+Particles are drawn to scale and coloured according to their phase.
+Dark red: crystal; orange: surface of crystal; yellow: liquid; blue: gas.
+ \textbf{b,} Fraction of particles for each phase. Gray vertical lines indicate the times shown in \textbf{a.}
+\textbf{c,} Transition probabilities for the same process as in \textbf{b}. Here crystal includes surface particles.}
  \label{fig:transitions}

### 2015-09-25 crystal-gel/science/SOM2.tex

%\documentclass[12pt]{article}
\documentclass[12pt,onecolumn]{revtex4}
%\documentstyle[aps,multicol,eincludegraphics]{revtex}
%\documentclass[twocolumn]{revtex4}
\usepackage{graphicx}
%\usepackage{scicite}
% \usepackage{epstopdf}
\usepackage{amsmath}
%\usepackage{times}

\topmargin 0.0cm
\oddsidemargin 0.2cm
\textwidth 16cm 
\textheight 21cm
\footskip 1.0cm

\begin{document}
\title{``Colloidal cold rain'' as a novel route to crystallization via viscoelastic phase separation}

\author{Hideyo Tsurusawa$1$, John Russo$^1$, Mathieu Leocmach$^{2}$ and Hajime Tanaka$^{1,\ast}$ 
\\
\normalsize{$^1$Institute of Industrial Science, University of Tokyo,}\\
\normalsize{4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}\\
\normalsize{$^2$Institut Lumière Matière, CNRS UMR 5306, Université Claude Bernard Lyon 1, }\\
\normalsize{Université de Lyon, Lyon, 69622 Villeurbanne Cedex, France}\\
\normalsize{$^\ast$Corresponding author. E-mail: tanaka@iis.u-tokyo.ac.jp}}

%\date{Received August 11, 2011}
\maketitle

%\centerline{\bf \Large Supporting Online Material}

\noindent
{\bf METHODS}

\noindent
{\bf Experimental.}
We use \textsc{pmma} (poly(methyl methacrylate)) colloids sterically stabilized with methacryloxypropyl terminated \textsc{pdms}(poly(dimethyl siloxane)) and fluorescently labelled with rhodamine isothiocyanate chemically bonded to the \textsc{pmma}. 
The average colloid diameter is 2.1 $\mu$m. 
We assess that the size distribution of our particle is Gaussian with a polydispersity below 5\% via direct confocal measurements~\cite{Leocmach2013}.
This small polydispersity allows crystallization.
We disperse the particles in refractive index and density matching mixture of cis-decalin (Tokyo Kasei) and bromocyclohexane (Sigma-Aldrich).

To induce short-ranged depletion attraction, we use polystyrene (TOSOH) of molecular weight 3.8 Mdalton as non-adsorbing polymer. 
Experiments are conducted at 27 $^\circ$C, some 80 $^\circ$C above the theta temperature in this solvents mixture~\cite{Royall2007}. A Flory scaling of the measurements of~\cite{lu2008gelation} yields a radius of gyration $R_g=76$ nm, and thus the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.07$.

In the absence of salt, the Debye length is expected to reach several $\mu$m and the (weakly) charged colloids experience a long range electrostatic repulsion. We confirm that colloids never come close enough to feel the short-ranged attraction. Screening by tetrabutylammonium bromide (Fluka) at saturated concentration brings down the Debye length to about 100 nm, practically discarding the repulsion. 
Thus, salt injection can screen the Coulomb repulsion and make the polymer-induced depletion attraction effective, initiating phase separation.

Our sample cell has two layer compartments separated by membrane filters. The first layer is a sample container with thickness of 200 $\mu$m. A mixture of colloids, polymer, and solvent without salt is  set in the first layer for microscopy observation. The second layer is a salt-reservoir with a half-opened structure, which allos us to exchange or insert a reservoir solution. The volume ratio of the first and second compartments is approximately 1:100. These two compartments are separated by a membrane filter with its pore size of 0.1 $\mu$m, which passes only salt ions.

In our experiments, the reservoir solution was initially a polymer solution without salt and electrostatic repulsion by the unscreened surface charges inhibited colloidal aggregation. Under microscopy observation, we quickly exchanged the reservoir solution into a polymer solution with salt at saturated concentration (4 mM) and sealed the half-opened reservoir with cover glass to avoid evaporation of solvent. Salt diffused into the first layer within a few minutes, typically 2 minutes, screened the surface charges, and initiated colloidal aggregation.

The data are collected on a Leica SP5 confocal microscope, using 532 nm laser excitation. The scanning volume is 98 $\times$ 98 $\times$ 53 $\mu$m$^3$, which contains $\sim 10^4$ colloid particles.  The Brownian time is $\tau_B \approx 4.5$ s. To be able to follow individual trajectories, we perform a 3D scan every 10 s at early time and every 30 s later.

%\paragraph*{State points studied}
%The experimental data is taken at two different volume fractions ($\phi=0.10$ and $\phi=0.25$) and for different values
%of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ for $\phi=0.25$ and $c_p=0.82,1.36$ for $\phi=0.10$.
%For each state points, scans at early times (every $10s$) and at the late stage (every $30s$) of the gelation process.
%
%We will consider two particles bonded if their center-to-center distance is within $12$ pixels. The radius of a colloidal
%particle is $5.365$ pixels.

\noindent
{\bf Structural analysis.}
To account for imprecision in particle localisation close to contact, we consider two particles bonded if their center-to-center distance is within 2.3 $\mu$m $>\sigma+2R_g$.

To detect percolation, two different methods have been employed. In the first one we detect percolation
by looking at which frame the cluster size distribution has a more extended power-law decay. In the second technique,
we simply measure the spatial extent of the largest cluster and consider it percolating when it is comparable to the
size of the field of view of the microscope. Both methods lead to essentially identical percolation time for each state point.

%\bibliographystyle{Science}
\bibliographystyle{apsrev}
\bibliography{biblio}

\end{document}

### 2015-09-25 crystal-gel/science/crystal_gel_letter2.tex

--- 2015-09-23 crystal_gel_letter.tex
@@
-%\documentclass[twocolumn,superscriptaddress,showpacs,preprintnumbers,amsmath,amssymb,prl]{revtex4-1}
-\documentclass[twocolumn,showpacs,amsmath,amssymb,pre,superscriptaddress]{revtex4}
-\usepackage[utf8]{inputenc}
-\usepackage{graphicx}% Include figure files
+\documentclass[12pt]{article}
+\usepackage{graphicx}
+\usepackage{epstopdf}
 \usepackage{amsmath}
 \usepackage{amssymb}
+\usepackage{scicite}
+
+% \usepackage{hyperref}
+\usepackage{booktabs}
 \usepackage{color}
-\usepackage{comment}
-\usepackage{siunitx}
-%\usepackage{dcolumn}% Align table columns on decimal point
-%\usepackage{bm}% bold math
 
-%\graphicspath{{./FIGS/}}
+\topmargin 0.0cm
+\oddsidemargin 0.2cm
+\textwidth 16cm 
+\textheight 21cm
+\footskip 1.0cm
+
+\newenvironment{sciabstract}{%
+\begin{quote} \bf}
+{\end{quote}}
+
+\newcounter{lastnote}
+\newenvironment{scilastnote}{%
+\setcounter{lastnote}{\value{enumiv}}%
+\addtocounter{lastnote}{+1}%
+\begin{list}%
+{\arabic{lastnote}.}
+{\setlength{\leftmargin}{.22in}}
+{\setlength{\labelsep}{.5em}}}
+{\end{list}}
+
+\title{``Colloidal rain'' as a novel route to crystallization via viscoelastic phase separation}
+
+\author{Hideyo Tsurusawa$1$, John Russo$^1$, Mathieu Leocmach$^{2}$ and Hajime Tanaka$^{1,\ast}$ 
+\\
+\normalsize{$^1$Institute of Industrial Science, University of Tokyo,}\\
+\normalsize{4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}\\
+\normalsize{$^2$Institut Lumière Matière, CNRS UMR 5306, Université Claude Bernard Lyon 1, }\\
+\normalsize{Université de Lyon, Lyon, 69622 Villeurbanne Cedex, France}\\
+\normalsize{$^\ast$Corresponding author. E-mail: tanaka@iis.u-tokyo.ac.jp}}
+
+\date{Received}
 
@@
 
-%\title{``Colloidal rain'', an alternative route to gelation during viscoelastic phase separation}
-%\title{Colloidal rain: a novel crystallization pathway following viscoelastic phase separation}
-\title{Novel crystallization pathway\\and formation of ``colloidal rain'' during viscoelastic phase separation}
-\author{Hideyo Tsurusawa}
-\affiliation{Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}
-\author{John Russo}
-%\email{russoj@iis.u-tokyo.ac.jp}
-\affiliation{Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}
-\author{Mathieu Leocmach}
-\affiliation{Institut Lumière Matière, CNRS UMR 5306, Université Claude Bernard Lyon 1, Université de Lyon, Lyon, 69622 Villeurbanne Cedex, France}
-\author{Hajime Tanaka}
-%\email{tanaka@iis.u-tokyo.ac.jp}
-\affiliation{Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}
+\baselineskip24pt
 
-\date{\today}
+% Make the title.
 
-\begin{abstract}
-{
-Colloidal gels form when network-forming spinodal phase separation is interrupted by the glass transition~\cite{lu2008gelation}.
+\maketitle
+\vspace{-1cm}
+\begin{sciabstract}
+%Colloidal gels form when network-forming spinodal phase separation is interrupted by the glass transition~\cite{lu2008gelation}.
 It has been shown that crystallization may equivalently interrupt the phase separation by freezing the dense phase.
 However the particle-level details of this process have not yet been confirmed experimentally.
-Here we show by confocal microscopy measurements that crystallisation leads to a network structure different from the original phase separation pattern.
+Here we show by confocal microscopy measurements that crystallization leads to a network structure different from the original phase separation pattern.
 Crystals nucleate inside the liquid network, and grow past it by direct condensation of the gas phase on their surface, driving evaporation of nearby liquid.
-This process represents the colloidal analogue of the Bergeron process~\cite{glickman2000glossary}, i.e.
-ice crystals formation that occurs in mixed phase clouds and is at the origin of rain.
-Our \emph{colloidal rain} gives us full experimental access to the kinetic pathway of this phenomenon important for climate science.
-We argue that states similar to colloidal rain can form as the result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems, whose dynamics is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
-}
-\end{abstract}
+This process represents the colloidal analogue of the Bergeron process, i.e., ice crystals formation that occurs in mixed phase clouds and is at the origin of rain.
+Our \emph{colloidal cold rain} gives us full experimental access to the kinetic pathway of this phenomenon important for climate science.
+We argue that states similar to colloidal cold rain can form as the result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems, whose dynamics is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
+\end{sciabstract}
 
 \maketitle
+
+\clearpage
+
 Gels and glasses are two important non-ergodic disordered states, which are very important 
@@
 those of a solid and a fluid phase, and allows the coexistence of two properties usually incompatible with each other, 
-elasticity and fluid transport. 
+elasticity and fluidity. 
 The elasticity of gels originates from the percolated network structure and the resulting space-spanning connectivity. 
@@
 of the dense liquid phase~\cite{pusey1993dynamics,ilett1995phase,verhaegh1997transient,tanaka1999colloid,foffi2002,zaccarelli2007,lu2008gelation,zaccarelli2008gelation,testard2011}.  
-More precisely, the dynamical stabilization is due to percolation of locally favoured structures, which are locally stable non-crystalline structures \cite{royall2008g}.
+More precisely, the dynamical stabilization is due to percolation of locally favored structures, which are locally stable non-crystalline structures \cite{royall2008g}.
 % Since ``glass'' is the term used for bulk, this may be a more appropriate way to express the mechanism of dynamical arrest. 
@@
 Unlike the above standard scenario of colloidal gelation, it is known that phase separation can also be arrested by crystallization, if the process takes place below the melting point of a component of a mixture~\cite{tanaka1985new}.
-Indeed the possibility of a different class of gels, which are stabilised by crystallization, 
+Indeed the possibility of a different class of gels, which are stabilized by crystallization, 
 was suggested by numerical simulations~\cite{fortini2008crystallization,perez2011pathways} and observed in microgravity experiments \cite{sabin2012}.
 For colloidal systems, we can in principle access structural evolution in real time at a single-particle level; however, there has so far been no confocal microscopy  
-study on the process of crystal-gel formation. The main difficulty consists in the preparation protocol, where a colloidal-polymer mixture undergoes
-shaking or shear before observation of its dynamics. This introduces turbolent flows at the beginning of the process, and also does not allow the
-observation of the first stages of crystallization.
+study on the dynamical process of crystal-gel formation. 
+For confocal microscopy experiments on colloid gelation, colloid-polymer mixtures have so far been used as a model system. 
+The main difficulty consists in the preparation protocol of such a colloid-polymer mixture, where before observation of its dynamics shaking or applying shear is necessary to prepare an initial homogeneous state.
+This introduces turbulent flows at the beginning of the process, and also does not allow the
+observation of the initial stages of crystallization. This initial perturbation may even alter the selection of the final non-equilibrium arrested state.  
 In our experiments we instead succeed in building a novel setup that allows the time evolution of the system to be observed directly
-under the confocal microscope from the very early stages and without introducing fluid flows.
+under the confocal microscope from the very early stages and in a quiescent situation without introducing fluid flows.
 
-Our results show that, for low polymer concentration, the system can undergo a novel crystallization pathway that leads to the formation
+In this Letter, we show that, for low polymer concentration, the system can undergo a novel crystallization pathway that leads to the formation
 of crystalline droplets.
@@
 The process is analogous to the Bergeron process of ice formation in mixed phase clouds, where ice droplets grow at the
-expense of the supercooled liquid droplets due to their lower saturated vapour pressure. In our colloidal system, we then call this process ``colloidal rain''
+expense of the supercooled liquid droplets due to their lower saturated vapor pressure. 
+This is speculated to be the primary mechanism for the formation of raindrops. 
+In our colloidal system, we then call this process ``colloidal cold rain''
 formation. Due to the difference in the nucleation and growth processes, the droplets display different scaling laws depending on their
@@
 We also observe that increasing the strength of the interparticle attraction
-(by increasing the polymer concentration), instead of forming colloidal rain, the system undergoes the classical gelation process, in which
-phase separation is interrupted by vitrification~\cite{lu2008gelation}.
+(by increasing the polymer concentration), instead of forming colloidal cold rain, the system undergoes the classical gelation process, in which
+phase separation is interrupted by vitrification~\cite{verhaegh1997transient,tanaka1999colloid,Royall2007,lu2008gelation}.
 
-\subsection*{Experiments}
-
-In our experimental setup, the sample cell is put in contact with a reservoir of a salt solution through a semi-permeable membrane.
+In our experimental setup, the sample cell is put in contact with a reservoir of a salt solution through a semi-permeable membrane (see Methods 
+on the details of our experiments).
 At $t=0$, the increase in the concentration of salt ions progressively screens the Coulombic repulsion between colloidal particles,
-which are then subject to attractive depletion forces, making the system thermodynamically unstable and leading to phase separation (see Methods).
+which are then subject to attractive depletion forces, making the system thermodynamically unstable and leading to phase separation. 
 %In the following we discuss all the stages of the formation of the crystal gels: phase separation, stress-driven agein and crystallization.
@@
 of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ for $\phi=0.25$ and $c_p=0.82,1.36$ for $\phi=0.10$.
-In Fig.XXX we superimpose the state points with the phase diagram calculated from the YYY equation of state.
+In Fig. 1A, we superimpose the state points with the phase diagram calculated from the YYY equation of state.
 For each state points, scans at early times (every $10s$) and at the late stage (every $30s$) of the gelation process.
-{\color{blue}We need more details of the experiments here}.
 
@@
 
-\subsection*{Results}
-
-\begin{figure*}[!t]
- \centering
- \includegraphics{./fig/phase_separation.pdf}
- \caption{{\bf Different routes of viscoelastic phase separation.}
- \textbf{a,} Phase diagram.
-Dashed lines are fluid-solid equilibrium coexistence boundaries.
-The continuous line is the gas side of the metastable gas-liquid spinodal region with the black diamond marking the critical point.
-The other symbols are the experimental points, consistently used in all panels.
- \textbf{b,} Radius of gyration ($R_g$) as a function of cluster size ($s$) for all state points at percolation. The
- gray line represents the fractal dimension of the random gelation universality class.
- \textbf{c,} Structure factors for the late stages of the gelation process. The gray and black lines represents respectively random gelation and compact fractal dimensions.
- \textbf{d,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle \textcolor{blue}{at late stage?}.
-Horizontal lines stress the value at 12 neighbours.
- }
- \label{fig:phase_separation}
-\end{figure*}
-
-All samples share the same early time stages of spinodal decomposition. First the colloidal particles start aggregating,
-eventually forming a percolating network. Then liquid-gas phase separation takes place, in which a dense network (liquid)
-coexists with freely diffusing monomers (gas). The topology of the network is similar to all our samples at the initial stages.
-This is shown for example in Fig.~\ref{fig:phase_separation}b, where we plot the radius of gyration of colloidal clusters ($R_g$)
+As described above and also in Methods, salt injection initiates liquid-gas phase separation of the colloidal suspension. 
+All samples share the same early stages of spinodal decomposition. 
+Due to strong dynamical asymmetry between colloids and the solvent \cite{tanaka1999colloid}, the colloidal particles start aggregating, 
+eventually forming a percolating network. Thus, a dense network (liquid) coexists with freely diffusing monomers (gas). 
+The topology of the network is similar to all our samples at the initial stages.
+This is shown for example in Fig.~\ref{fig:phase_separation}B, where we plot the radius of gyration of colloidal clusters ($R_g$)
 as a function of the cluster size ($s$) at the percolation point. Two particles belong to the same cluster if their distance is
 within the first peak of the radial distribution function. All samples show the same scaling law, which is compatible with
-the random gelation universality class exponent ($D=2.53$) as shown by the dashed line in Fig.~\ref{fig:phase_separation}b.
+the random gelation universality class exponent ($D=2.53$) as shown by the dashed line in Fig.~\ref{fig:phase_separation}B.
 
 While sharing similar structural properties at early times, the samples differ during the later stages of the phase separation process.
-To examin this in more detail, we plot in Fig.~\ref{fig:phase_separation}c the structure factor for the final stages of the gelation process for all state points.
+To examine this in more detail, we plot in Fig.~\ref{fig:phase_separation}C the structure factor for the final stages of the gelation process for all state points.
 Calculation of the structure factor is done with the Hanning window function, to minimize boundary effects.
@@
 with high polymer concentration retain the exponent $D=2.53$,
-which is the random gel universality class exponent we also observed in the early stages of the gelation process (see Fig.~\ref{fig:phase_separation}b),
+which is the random gel universality class exponent, as we also observed in the early stages of the gelation process (see Fig.~\ref{fig:phase_separation}B),
 the state point with low polymer concentration ($\phi=0.25$ and $c_p=0.38$) displays the biggest deviation from that exponent. In the figure we also
-plot the exponent $D=3$, which corresponds to volume growth, which we'll show in the following analysis to be the correct exponent for this state.
-The difference is due to the onset of crystallization in the low-polymer concentration sample. This is already evident in the high $q$ behaviour of
-the structure factors (Fig.~\ref{fig:phase_separation}c), where the diffraction planes appear as sharper peaks for $\phi=0.25$ and $c_p=0.38$.
+plot the exponent $D=3$, which corresponds to volume growth, which we will show in the following analysis to be the correct exponent for this state.
+The difference is due to the onset of crystallization in the low-polymer concentration sample. This is already evident in the high $q$ behavior of
+the structure factors (Fig.~\ref{fig:phase_separation}C), where the diffraction planes appear as sharper peaks for $\phi=0.25$ and $c_p=0.38$.
 
-To gain additional insight, in Fig.~\ref{fig:phase_separation}d we plot the average local volume fraction, computed
-through Voronoi diagrams, as a function of the number of neighbours. For all state points, we can see that the average volume fraction of a colloidal
-particle increases with its number of neighbours. Approaching close packing (12 neighbours), we clearly see two family of curves. The state point
+To gain additional insight, in Fig.~\ref{fig:phase_separation}D we plot the average local volume fraction, computed
+through Voronoi diagrams, as a function of the number of neighbors. For all state points, we can see that the average volume fraction of a colloidal
+particle increases with its number of neighbors. Approaching close packing (12 neighbors), we clearly see two family of curves. The state point
 at $\phi=0.25$ and $c_p=0.38$ reaches close packing at a volume fraction of around $\sim 60\%$, which is in reasonable agreement with the
-volume fraction of the stable crystal phase. All other state point, instead, reach close packing at volume fraction $\sim 55\%$, which is
+volume fraction of the stable crystal phase. All other state points, instead, reach close packing at volume fraction $\sim 55\%$, which is
 close to the volume fraction of the glass state. 
@@
 We should run the code of Mathieau with size detection}).
-The figure then clearly shows that the state point at $\phi=0.25$ and $c_p=0.38$ follows a different arrest mechanism, in which phase separation
+This result clearly indicates that the state point at $\phi=0.25$ and $c_p=0.38$ follows a different arrest mechanism, in which phase separation
 is arrested by crystallization and not glassiness. Although our estimation of the absolute volume fraction may involves errors of a few \%, the trend should be robust.
@@
 
-\begin{figure}[!t]
- \centering
- \includegraphics{fig/characterisation.pdf}
- \caption{{\bf Characterisation of the crystal-gel.} {\bf a,} Time evolution of the average size of the crystals (blue line) and the number of crystals (red line) for the state point $\phi=0.25$ and $c_p=0.38$.
-{\bf d,} Radius of gyration of crystalline nuclei at the late stages of gelation for the same state point.}
- \label{fig:crystals}
-\end{figure}
-
 To confirm the formation of crystalline seeds we adopt bond orientational analysis to detect which particles are in crystalline environments~\cite{russo2012microscopic}.
 With this analysis we confirm that crystallization events start occurring for the state point $\phi=0.25$ and $c_p=0.38$, which has the lowest
-polymer concentration of all state points. This is shown in Fig.~\ref{fig:crystals}a where the black line indicates the average size of the crystallites,
+polymer concentration of all state points. This is shown in Fig.~\ref{fig:crystals}A, where the black line indicates the average size of the crystallites,
 while the red line the average number of crystallites as a function of time. The number of crystallites first rapidly increases as nucleation events
@@
 idea of enhanced crystallization rates near metastable critical points~\cite{ten1997enhancement,olmsted1998spinodal} and the
-results of Ref.~\cite{fortini2008crystallization,perez2011pathways},
-which speculated two different arrest mechanism: crystallization at low polymer concentration, and dynamic arrest at high polymer concentrations.
+results of Refs.~\cite{fortini2008crystallization,perez2011pathways},
+which speculated two different arrest mechanisms: crystallization at low polymer concentration, and dynamic arrest at high polymer concentrations.
 
-While nucleation always occurs inside the liquid branches of the phase separating network, we suggest here the presence of a new growth channel
-which was not observed before.
-In the following we will show that the growth mechanism of the crystal is not only due to filling of the spinodal liquid network, as was observed in Ref.~\cite{fortini2008crystallization,perez2011pathways} for brownian dynamics simulations, but involves direct condensation of the gas phase
+While nucleation always occurs inside the liquid branches of the phase-separating network, we suggest here the presence of a new growth channel
+which was neither considered nor observed before.
+In the following we will show that the growth mechanism of the crystal is not only due to filling of the spinodal liquid network, as was observed in Refs.~\cite{fortini2008crystallization,perez2011pathways} for Brownian Dynamics simulations, but also involves direct condensation of the gas phase
 on the crystalline seeds.
-To investigate this process, we plot in Fig.~\ref{fig:crystals}c the gyration radius of individual crystalline nuclei for the state point
+To investigate this process, we plot in Fig.~\ref{fig:crystals}C the gyration radius of individual crystalline nuclei for the state point
 $\phi=0.25$ and $c_p=0.38$. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with
@@
 small crystalline nuclei are nucleated inside the phase-separated liquid branches, once they reach the transverse size of the liquid branch, the growth
-follows a volume expansion with the formation of isotropic droplets (colloidal rain). The same scaling law was suggested in the analysis of the
-structure factors, Fig.~\ref{fig:phase_separation}c, but it is shown directly by the analys of the gyration radius.
-
-\begin{figure*}
- \centering
- \includegraphics{fig/crystal.pdf}
- \caption{
-{\bf Crystal gel formation.} 
-\textbf{a,} Snapshots of configurations at $\phi=0.25$ and $c_p=0.38$. Particles are coloured according to their phase. Dark red: crystal; orange: surface of crystal; yellow: liquid; blue: gas. 
- \textbf{b,} Fraction of particles for each phase.
-\textbf{c,} Transition probabilities for the same process as in \textbf{b}. Here crystal includes surface particles.}
- \label{fig:transitions}
-\end{figure*}
+follows a volume expansion with the formation of isotropic droplets (colloidal cold rain). The same scaling law was suggested in the analysis of the
+structure factors, Fig.~\ref{fig:phase_separation}C, but it is shown directly by the analysis of the gyration radius.
 
 In the next section we demonstrate that the volume growth is due to direct condensation of the gas phase on the crystalline seeds.
-First we show in Fig.~\ref{fig:transitions}a slabs of 10 particle diameter thickness for configurations at $\phi=0.25$ and $c_p=0.38$ at different times.
-Particles are coloured according to their phase. The figure shows the first nucleation
+First we show in Fig.~\ref{fig:transitions}A slabs of 10 particle diameter thickness for configurations at $\phi=0.25$ and $c_p=0.38$ at different times.
+Particles are colored according to their phase. The figure shows the first nucleation
 events inside the liquid network (left panel); then these crystalline nuclei grow directly from the gas phase (middle panel), and finally
@@
 
-Next we show in Fig.~\ref{fig:transitions}b the fraction of particles in each different phase for the crystallization trajectory at
+%In the next section we show evidence for the fact that the volume growth is due to direct condensation of the gas phase on the crystalline seeds.
+%First we show in Fig.~\ref{fig:transitions}a slabs of 10 particle diameter thickness for configurations at $\phi=0.25$ and $c_p=0.38$ at different times.
+%Particles are coloured according to their phase as follows. Green: hcp; red: fcc; blue: bcc; white: liquid or gas. The figure shows the first nucleation
+%events inside the liquid network (left panel); then these crystalline nuclei grow directly from the gas phase (middle panel), and finally
+%the different nuclei coalesce (right panel).
+
+Next we show in Fig.~\ref{fig:transitions}B the fraction of particles in each different phase for the crystallization trajectory at
 $\phi=0.25$ and $c_p=0.38$, after the liquid-gas phase separation has occurred. The process of crystallization is characterized
@@
 is accompanied by an increase in the fraction of gas particles: as the first crystals start to appear, liquid particles evaporate
-to the gas phase due to the highest vapour pressure of the liquid phase compared to the crystalline phase. After the onset
-of a steady-state gas population, the crystallization proceeds by adding surface particles to the small crystallites. To confirm this process we
-examine in Fig.~\ref{fig:transitions}c the transfer matrix probabilities $P_{\alpha\beta}$, where $\alpha$ and $\beta$ are each of the
+to the gas phase due to the higher vapor pressure of the liquid phase compared to the crystalline phase. After the onset
+of a steady-state gas population, the crystallization proceeds by adding surface particles from the gas phase to the small crystallites. To confirm this process we
+examine in Fig.~\ref{fig:transitions}C the transfer matrix probabilities $P_{\alpha\beta}$, where $\alpha$ and $\beta$ are each of the
 phases gas, liquid and crystal (which includes bcc, hcp, fcc and surface particles). Liquid particles undergo two processes, on one
@@
 
-\subsection*{Discussion and conclusions}
+%\subsection*{Discussion and conclusions}
 The first stages of gelation always involve spinodal decomposition with the formation
@@
 
-The process we observed is akin to the Bergeron process which is the primary mechanism for the formation of raindrops in clouds~\cite{glickman2000glossary}.
+The process we observed is akin to the Bergeron process which is the primary mechanism for the formation of cold rain drops in clouds~\cite{glickman2000glossary}.
 In clouds there is a mixture of ice crystals and supercooled water. The vapor phase is in coexistence with the liquid phase, but is supersaturated
@@
 be regarded as a colloidal analogue for this important process, which can now be studied at the single-particle level.
+
 The basic mechanism creating ``crystal gels'' may be generic to many other
@@
 This can be seen, for example, by looking at the phase diagram of a Lennard-Jones liquid, which represents many molecular systems 
-without specific directional interactions. So we believe that crystal gels are an important class of heterogeneous non-ergodic states in nature 
+without specific directional interactions. 
+Usually, monoatomic systems are very poor glass-formers and thus have not been expected to form gels. 
+However, our mechanism provides a novel kinetic pathway to form network structures made of crystals.  
+So we believe that crystal gels are an important class of heterogeneous non-ergodic states in nature 
 and industrial applications, although they have not attracted much attention so far.
 
-\section*{Methods}
+\bibliographystyle{Science}
+\bibliography{biblio}
 
-\paragraph*{Experimental}
-We use \textsc{pmma} (poly(methyl methacrylate)) colloids sterically stabilized with methacryloxypropyl terminated \textsc{pdms}(poly(dimethyl siloxane)) and fluorescently labelled with rhodamine isothiocyanate chemically bonded to the \textsc{pmma}.
-We disperse the particles in refractive index and density matching mixture of cis-decalin (Tokyo Kasei) and bromocyclohexane (Sigma-Aldrich).
-
-To induce short-ranged depletion attraction, we use polystyrene (TOSOH) of molecular weight \SI{3.8}{\mega\dalton} as non-adsorbing polymer.
-Experiments are conducted at \SI{27}{\celsius}, some \SI{80}{\celsius} above the theta temperature in this solvents mixture~\cite{Royall2007}.
-A Flory scaling of the measurements of~\cite{lu2008gelation} yields a radius of gyration $R_g=\SI{76}{\nano\metre}$.
-
-In the absence of salt, the Debye length is expected to reach several \si{\micro\metre} and the (weakly) charged colloids experience a long range electrostatic repulsion.
-We confirm that colloids never come close enough to feel the short-ranged attraction.
-Screening by tetrabutylammonium bromide (Fluka) at saturated concentration brings down the Debye length to about \SI{100}{\nano\metre}, practically discarding the repulsion.
-
-The data are collected on a Leica SP5 confocal microscope, using \SI{532}{\nano\metre} laser excitation.
-We assess that the size distribution of our particle is Gaussian with a polydispersity below 5\% via direct confocal measurements~\cite{Leocmach2013}.
-The mean diameter is $\sigma=\SI{2.1}{\micro\metre}$ XXX to be checked XXX yielding a size ratio $q_R=2R_g/\sigma=0.07$.
-The Brownian time is $\tau_B \approx \SI{4.5}{\second}$.
-To be able to follow individual trajectories, we perform a 3D scan every \SI{10}{\second} at early time and every \SI{30}{\second} later.
-
-%\paragraph*{State points studied}
-%The experimental data is taken at two different volume fractions ($\phi=0.10$ and $\phi=0.25$) and for different values
-%of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ for $\phi=0.25$ and $c_p=0.82,1.36$ for $\phi=0.10$.
-%For each state points, scans at early times (every $10s$) and at the late stage (every $30s$) of the gelation process.
-%
-%We will consider two particles bonded if their center-to-center distance is within $12$ pixels. The radius of a colloidal
-%particle is $5.365$ pixels.
-
-\paragraph*{Structural analysis}
-To account for imprecision in particle localisation close to contact, we consider two particles bonded if their center-to-center distance is within $\SI{2.3}{\micro\metre}>\sigma+2R_g$.
-
-To detect percolation, two different methods have been employed. In the first one we detect percolation
-by looking at which frame the cluster size distribution has a more extended power-law decay. In the second technique,
-we simply measure the spatial extent of the largest cluster and consider it percolating when it is comparable to the
-size of the field of view of the microscope. Both methods lead to essentially identical percolation time for each state point.
-
-Crystalline phases are detected through bond orientational analysis~\cite{russo2012microscopic}, they are mainly FCC and HCP in comparable proportions.
-Surface particles are particles with at least two crystalline neighbours; liquid particles have at least 4 neighbours; gas particles are particles with less than 4 nearest neighbours.
-
-\bibliographystyle{naturemag4}
-\bibliography{biblio}
+\noindent
+{\bf Acknowledgments}
+\noindent
+H.Ts. and J.R. contributed equally to this work. 
+H.Ta. conceived and supervised the project, H.Ts. performed experiments, J.R. analyzed the data, M.L. calculated the phase diagram, and all the authors discussed and wrote the manuscript. 
+This study was partly supported by Grants-in-Aid for Scientific Research (S) and Specially Promoted Research from JSPS.
 
@@
 
+\begin{figure*}[!t]
+ \centering
+ \includegraphics{phase_separation.pdf}
+ \caption{Different routes of viscoelastic phase separation. 
+(A) Phase diagram.
+Dashed lines are fluid-solid equilibrium coexistence boundaries.
+The continuous line is the gas side of the metastable gas-liquid spinodal region with the black diamond marking the critical point.
+The other symbols are the experimental points, consistently used in all panels.
+(B) Radius of gyration ($R_g$) as a function of cluster size ($s$) for all state points at percolation. The
+ gray line represents the fractal dimension of the random gelation universality class.
+(C) Structure factors for the late stages of the gelation process. The gray and black lines represents respectively random gelation and compact fractal dimensions.
+(D) Local volume fraction around a colloidal particle, as a function of the number of nearest neighbors of the particle for the state points in the very late stage.
+Horizontal lines stress the value at 12 neighbors.
+ }
+ \label{fig:phase_separation}
+\end{figure*}
+
+\clearpage
+
+\begin{figure}[!t]
+ \centering
+ \includegraphics[width=12cm]{characterisation}
+ % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
+\caption{Characterization of the crystal-gel.
+(A) Time evolution of the average size of the crystals (blue line) and the number of crystals (red line) for the state point $\phi=0.25$ and $c_p=0.38$.
+(B) Radius of gyration of crystalline nuclei at the late stages of gelation for the same state point.}
+ \label{fig:crystals}
+\end{figure}
+
+\clearpage
+
+\begin{figure*}
+ \centering
+ \includegraphics[width=16cm]{crystal}
+ % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735
+ \caption{
+Crystal gel formation.  
+(A) Reconstructions from confocal coordinates at $\phi=0.25$ and $c_p=0.38$.
+Depth of view is $5\sigma$.
+Particles are drawn to scale and colored according to their phase.
+Dark red: crystal; orange: surface of crystal; yellow: liquid; blue: gas.
+(B) Fraction of particles for each phase. Gray vertical lines indicate the times shown in (A). 
+(C) Transition probabilities for the same process as in (B). Here crystal includes surface particles.}
+\label{fig:transitions}
+\end{figure*}
+
 \end{document}
-
-\section{Future things to do}
-Is the absence of network restructuring that is preventing crystallization?

### 2015-09-25 data/slices.py
more saturation gap between the 4 colors to be print ready

@@ -11,9 +11,9 @@ for t in [120, 390, 890]:
     crystals = Union(*[
         Sphere( (x,y,z), 5)
         for x,y,z in pos[phase<3]
-        ]+[Texture(Pigment(color=(0.5,0,0)))])
+        ]+[Texture(Pigment(color=(0.33,0,0)))])
     f.write(crystals)
-    for i, c in zip([3,4,5], [(1,0.33,0), (1,0.66,0), (0.5,0,1)]):
+    for i, c in zip([3,4,5], [(0.66,0.33,0), (1,0.66,0.1), (0.75,0.75,1)]):
         f.write(Union(*[
             Sphere( (x,y,z), 5)
             for x,y,z in pos[phase==i]

### 2015-09-29 science/crystal_gel_letter2.tex
Updated some references. Minor changes to the text. Written first draft of the cover letter.

--- 2015-09-25 crystal-gel/science/crystal_gel_letter2.tex
@@
 
-\author{Hideyo Tsurusawa$1$, John Russo$^1$, Mathieu Leocmach$^{2}$ and Hajime Tanaka$^{1,\ast}$ 
+\author{Hideyo Tsurusawa$^1$, John Russo$^1$, Mathieu Leocmach$^{2}$ and Hajime Tanaka$^{1,\ast}$ 
 \\
@@
 This process represents the colloidal analogue of the Bergeron process, i.e., ice crystals formation that occurs in mixed phase clouds and is at the origin of rain.
-Our \emph{colloidal cold rain} gives us full experimental access to the kinetic pathway of this phenomenon important for climate science.
-We argue that states similar to colloidal cold rain can form as the result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems, whose dynamics is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
+Our \emph{colloidal rain} gives us full experimental access to the kinetic pathway of this phenomenon important for climate science.
+We argue that states similar to colloidal rain can form as the result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems, whose dynamics is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
 \end{sciabstract}
@@
 with fast dynamics. This difference in the viscoelastic properties between the two phases allows the system to form a space-spanning 
-network structure of the liquid phase, even if it is the minority phase. During phase separation, the density of the liquid phase 
+network structure of the liquid phase, even if it is the minority phase. The network is fractal and its dynamic structure factor shows a stretched-exponential decay to a finite
+plateau~\cite{krall1998internal,solomon2001dynamic,romer2000sol}. During phase separation, the density of the liquid phase 
 increases towards the glass-transition density, leading to slow glassy dynamics.
 At the glass transition point, the percolated network structure is dynamically stabilised by vitrification 
-of the dense liquid phase~\cite{pusey1993dynamics,ilett1995phase,verhaegh1997transient,tanaka1999colloid,foffi2002,zaccarelli2007,lu2008gelation,zaccarelli2008gelation,testard2011}.  
+of the dense liquid phase~\cite{pusey1993dynamics,ilett1995phase,verhaegh1997transient,tanaka1999colloid,foffi2002,buzzaccaro2007sticky,zaccarelli2007,lu2008gelation,zaccarelli2008gelation,testard2011}.
 More precisely, the dynamical stabilization is due to percolation of locally favored structures, which are locally stable non-crystalline structures \cite{royall2008g}.
@@
 This is speculated to be the primary mechanism for the formation of raindrops. 
-In our colloidal system, we then call this process ``colloidal cold rain''
+In our colloidal system, we then call this process ``colloidal rain''
 formation. Due to the difference in the nucleation and growth processes, the droplets display different scaling laws depending on their
@@
 We also observe that increasing the strength of the interparticle attraction
-(by increasing the polymer concentration), instead of forming colloidal cold rain, the system undergoes the classical gelation process, in which
+(by increasing the polymer concentration), instead of forming colloidal rain, the system undergoes the classical gelation process, in which
 phase separation is interrupted by vitrification~\cite{verhaegh1997transient,tanaka1999colloid,Royall2007,lu2008gelation}.
@@
 
-To confirm the formation of crystalline seeds we adopt bond orientational analysis to detect which particles are in crystalline environments~\cite{russo2012microscopic}.
+To confirm the formation of crystalline seeds we adopt bond orientational analysis to detect which particles are in crystalline environments~\cite{russo2013interplay}.
 With this analysis we confirm that crystallization events start occurring for the state point $\phi=0.25$ and $c_p=0.38$, which has the lowest
@@
 small crystalline nuclei are nucleated inside the phase-separated liquid branches, once they reach the transverse size of the liquid branch, the growth
-follows a volume expansion with the formation of isotropic droplets (colloidal cold rain). The same scaling law was suggested in the analysis of the
+follows a volume expansion with the formation of isotropic droplets (colloidal rain). The same scaling law was suggested in the analysis of the
 structure factors, Fig.~\ref{fig:phase_separation}C, but it is shown directly by the analysis of the gyration radius.
@@
 
-The process we observed is akin to the Bergeron process which is the primary mechanism for the formation of cold rain drops in clouds~\cite{glickman2000glossary}.
+The process we observed is akin to the Bergeron process which is the primary mechanism for the formation of rain drops in clouds~\cite{glickman2000glossary,morrison2012resilience,}.
 In clouds there is a mixture of ice crystals and supercooled water. The vapor phase is in coexistence with the liquid phase, but is supersaturated
@@
  \centering
- \includegraphics[width=12cm]{characterisation}
+ \includegraphics{characterisation}
  % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
@@
  \centering
- \includegraphics[width=16cm]{crystal}
+ \includegraphics{crystal}
  % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735

### 2015-09-29 crystal-gel/science/coverl_letter.tex
Updated some references. Minor changes to the text. Written first draft of the cover letter.

\documentclass[11pt]{article}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage{color}
\usepackage{setspace}

\begin{document}

\doublespacing

\noindent
Dear Editor,
\vskip 0.3cm
please find enclosed in the present submission our manuscript entitled
\emph{``Colloidal rain'' as a novel route to crystallization via viscoelastic phase separation},
that we would like you to consider for publication in \emph{Science}.

The Letter represents the first particle-level study of crystal growth via the Wegener-Bergeron-Findeisen mechanism.
The process occurs naturally in mixed phase clouds, between 0$^\circ$ C and -38$^\circ$ C, where supercooled water and ice crystals coexist,
and is responsible for the growth of ice crystals in clouds and rain formation. But we argue that this
mechanism of crystal growth is much more common than previously thought,
and show that it can be accessed directly in a Soft Matter model system.
We study a mixture of poly(methyl methacrylate) colloids with non-adorbing polymer (polystyrene)
with confocal microscopy experiments, that give us access to the full kinetic process with
particle resolution. We have built a novel experimental setup that allows the evolution of
the system to be observed directly from the very early stages and in absence of spurious fluid flows.
This allows us to directly study the dynamics of phase separation and of crystal growth at single-particle resolutions,
and for the first time characterize the Bergeron process at a microscopic level.

Understanding the process of crystal growth in mixed-phase systems is of fundamental importance,
and has several technological applications. As an example, understanding
water crystallization in clouds is recognized as one of the main challenges in the
modeling of Earth's radiation budget and climate. The difficulties arise both from our poor understanding of small-scale microphysical effects,
and to the difficulty of experimental investigations. In our colloidal system, both
difficulties are overcome due to the colloid's accessible length and time scales.

Our manuscript thus represents a fundamental investigation on one of the most important
crystal-growth mechanisms, that will resonate with the broad Science community.
Moreover, it highlights a very deep connection between the disciplines of Soft Matter and Climate Modeling,
and we thus believe that will have broad appeal to the readership of Science.

We hope that you will consider it for publication.

\vskip 0.3cm
\noindent
For your convenience, we provide here a list of possible Referees who are widely considered to be experts in the field.

\begin{itemize}
 \item Prof. Francesco Sciortino, University of Rome ``La Sapienza'':
 francesco.sciortino@phys.uniroma1.it
\end{itemize}

\vskip 0.3cm
\noindent
Warm thanks to you for your kind attention.

\vskip 0.8cm

\noindent
Best regards,
\vskip 0.3cm
\indent Hideyo Tsurusawa\\
\indent John Russo\\
\indent Mathieu Leocmach\\
\indent Hajime Tanaka

\end{document}

### 2015-10-05 crystal-gel/crystal_gel_letter.tex
Fixed the equation for the radius of gyration.

@@
 \usepackage{comment}
-\usepackage{siunitx}
+%\usepackage{siunitx}
 %\usepackage{dcolumn}% Align table columns on decimal point
@@
  \textbf{c,} Structure factors for the late stages of the gelation process. The gray and black lines represents respectively random gelation and compact fractal dimensions.
+ Inset: same data in linear scales, shifted vertically for clarity.
  \textbf{d,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle \textcolor{blue}{at late stage?}.
@@
 
-\paragraph*{Experimental}
-We use \textsc{pmma} (poly(methyl methacrylate)) colloids sterically stabilized with methacryloxypropyl terminated \textsc{pdms}(poly(dimethyl siloxane)) and fluorescently labelled with rhodamine isothiocyanate chemically bonded to the \textsc{pmma}.
+% \paragraph*{Experimental}
+% We use \textsc{pmma} (poly(methyl methacrylate)) colloids sterically stabilized with methacryloxypropyl terminated \textsc{pdms}(poly(dimethyl siloxane)) and fluorescently labelled with rhodamine isothiocyanate chemically bonded to the \textsc{pmma}.
 We disperse the particles in refractive index and density matching mixture of cis-decalin (Tokyo Kasei) and bromocyclohexane (Sigma-Aldrich).
-
-To induce short-ranged depletion attraction, we use polystyrene (TOSOH) of molecular weight \SI{3.8}{\mega\dalton} as non-adsorbing polymer.
+% 
+% To induce short-ranged depletion attraction, we use polystyrene (TOSOH) of molecular weight \SI{3.8}{\mega\dalton} as non-adsorbing polymer.
 Experiments are conducted at \SI{27}{\celsius}, some \SI{80}{\celsius} above the theta temperature in this solvents mixture~\cite{Royall2007}.
 A Flory scaling of the measurements of~\cite{lu2008gelation} yields a radius of gyration $R_g=\SI{76}{\nano\metre}$.
-
-In the absence of salt, the Debye length is expected to reach several \si{\micro\metre} and the (weakly) charged colloids experience a long range electrostatic repulsion.
+% 
+% In the absence of salt, the Debye length is expected to reach several \si{\micro\metre} and the (weakly) charged colloids experience a long range electrostatic repulsion.
 We confirm that colloids never come close enough to feel the short-ranged attraction.
 Screening by tetrabutylammonium bromide (Fluka) at saturated concentration brings down the Debye length to about \SI{100}{\nano\metre}, practically discarding the repulsion.
-
-The data are collected on a Leica SP5 confocal microscope, using \SI{532}{\nano\metre} laser excitation.
+% 
+% The data are collected on a Leica SP5 confocal microscope, using \SI{532}{\nano\metre} laser excitation.
 We assess that the size distribution of our particle is Gaussian with a polydispersity below 5\% via direct confocal measurements~\cite{Leocmach2013}.
@@
 
-\paragraph*{Structural analysis}
-To account for imprecision in particle localisation close to contact, we consider two particles bonded if their center-to-center distance is within $\SI{2.3}{\micro\metre}>\sigma+2R_g$.
-
-To detect percolation, two different methods have been employed. In the first one we detect percolation
-by looking at which frame the cluster size distribution has a more extended power-law decay. In the second technique,
-we simply measure the spatial extent of the largest cluster and consider it percolating when it is comparable to the
-size of the field of view of the microscope. Both methods lead to essentially identical percolation time for each state point.
+% \paragraph*{Structural analysis}
+% To account for imprecision in particle localisation close to contact, we consider two particles bonded if their center-to-center distance is within $\SI{2.3}{\micro\metre}>\sigma+2R_g$.
+% 
+% To detect percolation, two different methods have been employed. In the first one we detect percolation
+% by looking at which frame the cluster size distribution has a more extended power-law decay. In the second technique,
+% we simply measure the spatial extent of the largest cluster and consider it percolating when it is comparable to the
+% size of the field of view of the microscope. Both methods lead to essentially identical percolation time for each state point.
 
@@
 \end{document}
-
-\section{Future things to do}
-Is the absence of network restructuring that is preventing crystallization?

### 2015-10-05 science/crystal_gel_letter2.tex
Fixed the equation for the radius of gyration.

@@
 as a function of the cluster size ($s$) at the percolation point. Two particles belong to the same cluster if their distance is
-within the first peak of the radial distribution function. All samples show the same scaling law, which is compatible with
+within the first peak of the radial distribution function. All samples show the same scaling law, $R^2_{\rm g}\sim s^{2/D}$, which is compatible with
 the random gelation universality class exponent ($D=2.53$) as shown by the dashed line in Fig.~\ref{fig:phase_separation}B.

### 2015-10-05 science/coverl_letter.tex

--- 2015-09-29 crystal-gel/science/coverl_letter.tex
@@
 \noindent
-Dear Editor,
+{\bf Dear Editor,}
 \vskip 0.3cm
-please find enclosed in the present submission our manuscript entitled
+Please find enclosed in the present submission our manuscript entitled
 \emph{``Colloidal rain'' as a novel route to crystallization via viscoelastic phase separation},
@@
 
-The Letter represents the first particle-level study of crystal growth via the Wegener-Bergeron-Findeisen mechanism.
+Our Report represents the first particle-level study of crystal growth via the Wegener-Bergeron-Findeisen mechanism.
 The process occurs naturally in mixed phase clouds, between 0$^\circ$ C and -38$^\circ$ C, where supercooled water and ice crystals coexist,
@@
 \vskip 0.3cm
-\noindent
 For your convenience, we provide here a list of possible Referees who are widely considered to be experts in the field.
-
 \begin{itemize}
- \item Prof. Francesco Sciortino, University of Rome ``La Sapienza'':
- francesco.sciortino@phys.uniroma1.it
+\item Daniel Bonn, University of Amsterdam, D.Bonn@uva.nl
+\item Prof. Henk Lekkerkerker, University of Utrecht, H.N.W.Lekkerkerker@uu.nl
+\item Prof. Francesco Sciortino, University of Rome ``La Sapienza'', \\ francesco.sciortino@phys.uniroma1.it
+\item Prof. Michael Solomon, University of Michigan, mjsolo@umich.edu
+\item Prof. David Weitz, Harvard University, weitz­@seas.harvard.edu
 \end{itemize}
@@
 
+{\bf 
 \noindent
-Best regards,
+Sincerely yours,
 \vskip 0.3cm
-\indent Hideyo Tsurusawa\\
-\indent John Russo\\
-\indent Mathieu Leocmach\\
-\indent Hajime Tanaka
+\noindent
+Hideyo Tsurusawa, John Russo, Mathieu Leocmach, \\ 
+and Hajime Tanaka
+}
 

### 2015-10-05 science/crystal_gel_letter2.tex

@@
 of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ for $\phi=0.25$ and $c_p=0.82,1.36$ for $\phi=0.10$.
-In Fig. 1A, we superimpose the state points with the phase diagram calculated from the YYY equation of state.
+In Fig. 1A, we superimpose the state points with the phase diagram calculated from the equation of state.
 For each state points, scans at early times (every $10s$) and at the late stage (every $30s$) of the gelation process.
@@
 on the crystalline seeds.
-To investigate this process, we plot in Fig.~\ref{fig:crystals}C the gyration radius of individual crystalline nuclei for the state point
+To investigate this process, we plot in Fig.~\ref{fig:crystals}B the gyration radius of individual crystalline nuclei for the state point
 $\phi=0.25$ and $c_p=0.38$. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with
@@
 
-The process we observed is akin to the Bergeron process which is the primary mechanism for the formation of rain drops in clouds~\cite{glickman2000glossary,morrison2012resilience,}.
+The process we observed is akin to the Bergeron process which is the primary mechanism for the formation of rain drops in clouds~\cite{glickman2000glossary,morrison2012resilience}.
 In clouds there is a mixture of ice crystals and supercooled water. The vapor phase is in coexistence with the liquid phase, but is supersaturated
@@
 weak or little frustration against crystallization (in our case, the use of monodisperse colloids), 
-(ii) dynamical slowing down in a supercooled liquid state, which is necessary to induce viscoelastic phase separation leading to the formation 
-of a network structure of the minority phase, and (iii) the degree of supercooling is low enough to avoid a high rate of densification of the liquid phase 
+(iii) dynamical slowing down in a supercooled liquid state, which is necessary to induce viscoelastic phase separation leading to the formation 
+of a network structure of the minority phase, and (iv) the degree of supercooling is low enough to avoid a high rate of densification of the liquid phase 
 leading to vitrification.   

### 2015-10-05 science/SOM2.tex

--- 2015-09-25 crystal-gel/science/SOM2.tex
@@
 We disperse the particles in refractive index and density matching mixture of cis-decalin (Tokyo Kasei) and bromocyclohexane (Sigma-Aldrich). 
+The precision of density matching is $\sim 10^{-4}$ mg/ml, for which the gravity effect on our gelation processes is negligible up to 12 hours.
 
@@
 
-In the absence of salt, the Debye length is expected to reach several $\mu$m and the (weakly) charged colloids experience a long range electrostatic repulsion.
+In the absence of salt, the Debye length is expected to reach a few $\mu$m and the (weakly) charged colloids experience a long range electrostatic repulsion.
 We confirm that colloids never come close enough to feel the short-ranged attraction.
-Screening by tetrabutylammonium bromide (Fluka) at saturated concentration brings down the Debye length to about 100 nm, practically discarding the repulsion.
+Screening by tetrabutylammonium bromide (Fluka) at saturated concentration brings down the Debye length to about several nanometers practically discarding the repulsion.
 Thus, salt injection can screen the Coulomb repulsion and make the polymer-induced depletion attraction effective, initiating phase separation.
@@
 The scanning volume is 98 $\times$ 98 $\times$ 53 $\mu$m$^3$, which contains $\sim 10^4$ colloid particles.
+Our spatial resolution is 192 nm / pixel.
 The Brownian time is $\tau_B \approx 4.5$ s.

### 2015-10-05 crystal-gel/science/crystal_gel_letter2.tex
Adapt figures to science layout: sans serif and labels in capital fig:characterisation horizontal

@@
 The experimental data is taken at two different volume fractions ($\phi=0.10$ and $\phi=0.25$) and for different values
-of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ for $\phi=0.25$ and $c_p=0.82,1.36$ for $\phi=0.10$.
-In Fig. 1A, we superimpose the state points with the phase diagram calculated from the YYY equation of state.
+of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ mg/g for $\phi=0.25$ and $c_p=0.82,1.36$ mg/g for $\phi=0.10$.
+In Fig. 1A, we superimpose the state points with the phase diagram calculated from the equation of state.
 For each state points, scans at early times (every $10s$) and at the late stage (every $30s$) of the gelation process.
@@
 But a difference
-in the fractal dimension $D$ between the $\phi=0.25$ and $c_p=0.38$ state point and other state points starts to become visible. In fact, while states
+in the fractal dimension $D$ between the $\phi=0.25$ and $c_p=0.38$ mg/g state point and other state points starts to become visible. In fact, while states
 with high polymer concentration retain the exponent $D=2.53$,
 which is the random gel universality class exponent, as we also observed in the early stages of the gelation process (see Fig.~\ref{fig:phase_separation}B),
-the state point with low polymer concentration ($\phi=0.25$ and $c_p=0.38$) displays the biggest deviation from that exponent. In the figure we also
+the state point with low polymer concentration ($\phi=0.25$ and $c_p=0.38$ mg/g) displays the biggest deviation from that exponent. In the figure we also
 plot the exponent $D=3$, which corresponds to volume growth, which we will show in the following analysis to be the correct exponent for this state.
 The difference is due to the onset of crystallization in the low-polymer concentration sample. This is already evident in the high $q$ behavior of
-the structure factors (Fig.~\ref{fig:phase_separation}C), where the diffraction planes appear as sharper peaks for $\phi=0.25$ and $c_p=0.38$.
+the structure factors (Fig.~\ref{fig:phase_separation}C), where the diffraction planes appear as sharper peaks for $\phi=0.25$ and $c_p=0.38$ mg/g.
 
@@
 particle increases with its number of neighbors. Approaching close packing (12 neighbors), we clearly see two family of curves. The state point
-at $\phi=0.25$ and $c_p=0.38$ reaches close packing at a volume fraction of around $\sim 60\%$, which is in reasonable agreement with the
+at $\phi=0.25$ and $c_p=0.38$ mg/g reaches close packing at a volume fraction of around $\sim 60\%$, which is in reasonable agreement with the
 volume fraction of the stable crystal phase. All other state points, instead, reach close packing at volume fraction $\sim 55\%$, which is
@@
 We should run the code of Mathieau with size detection}).
-This result clearly indicates that the state point at $\phi=0.25$ and $c_p=0.38$ follows a different arrest mechanism, in which phase separation
+This result clearly indicates that the state point at $\phi=0.25$ and $c_p=0.38$ mg/g follows a different arrest mechanism, in which phase separation
 is arrested by crystallization and not glassiness. Although our estimation of the absolute volume fraction may involves errors of a few \%, the trend should be robust.
@@
 
-To confirm the formation of crystalline seeds we adopt bond orientational analysis to detect which particles are in crystalline environments~\cite{russo2012microscopic}.
-With this analysis we confirm that crystallization events start occurring for the state point $\phi=0.25$ and $c_p=0.38$, which has the lowest
+To confirm the formation of crystalline seeds we adopt bond orientational analysis to detect which particles are in crystalline environments~\cite{russo2013interplay}.
+With this analysis we confirm that crystallization events start occurring for the state point $\phi=0.25$ and $c_p=0.38$ mg/g, which has the lowest
 polymer concentration of all state points. This is shown in Fig.~\ref{fig:crystals}A, where the black line indicates the average size of the crystallites,
@@
 on the crystalline seeds.
-To investigate this process, we plot in Fig.~\ref{fig:crystals}C the gyration radius of individual crystalline nuclei for the state point
-$\phi=0.25$ and $c_p=0.38$. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with
+To investigate this process, we plot in Fig.~\ref{fig:crystals}B the gyration radius of individual crystalline nuclei for the state point
+$\phi=0.25$ and $c_p=0.38$ mg/g. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with
 the dimension of random percolation $D=2.53$, while at large sizes it scales as $D=3$ of compact crystal growth. This demonstrates that, while
 small crystalline nuclei are nucleated inside the phase-separated liquid branches, once they reach the transverse size of the liquid branch, the growth
-follows a volume expansion with the formation of isotropic droplets (colloidal cold rain). The same scaling law was suggested in the analysis of the
+follows a volume expansion with the formation of isotropic droplets (colloidal rain). The same scaling law was suggested in the analysis of the
 structure factors, Fig.~\ref{fig:phase_separation}C, but it is shown directly by the analysis of the gyration radius.
@@
 In the next section we demonstrate that the volume growth is due to direct condensation of the gas phase on the crystalline seeds.
-First we show in Fig.~\ref{fig:transitions}A slabs of 10 particle diameter thickness for configurations at $\phi=0.25$ and $c_p=0.38$ at different times.
+First we show in Fig.~\ref{fig:transitions}A slabs of 10 particle diameter thickness for configurations at $\phi=0.25$ and $c_p=0.38$ mg/g at different times.
 Particles are colored according to their phase. The figure shows the first nucleation
@@
 Next we show in Fig.~\ref{fig:transitions}B the fraction of particles in each different phase for the crystallization trajectory at
-$\phi=0.25$ and $c_p=0.38$, after the liquid-gas phase separation has occurred. The process of crystallization is characterized
+$\phi=0.25$ and $c_p=0.38$ mg/g, after the liquid-gas phase separation has occurred. The process of crystallization is characterized
 by a steep decrease in liquid particles, as they transform in small crystalline nuclei inside the liquid domains. This decrease
@@
 Depending on the polymer concentration, there are two possible arrest mechanisms.
-(i) Crystallization, observed at $\phi=0.25$ and $c_p=0.38$, small crystalline nuclei appear inside the liquid, reach the surface
+(i) Crystallization, observed at $\phi=0.25$ and $c_p=0.38$ mg/g, small crystalline nuclei appear inside the liquid, reach the surface
  of the liquid branches, and then grow by addition of particles from the gas phase. The final structure is a network of crystal droplets,
@@
 
-The process we observed is akin to the Bergeron process which is the primary mechanism for the formation of cold rain drops in clouds~\cite{glickman2000glossary}.
+The process we observed is akin to the Bergeron process which is the primary mechanism for the formation of rain drops in clouds~\cite{glickman2000glossary,morrison2012resilience}.
 In clouds there is a mixture of ice crystals and supercooled water. The vapor phase is in coexistence with the liquid phase, but is supersaturated
@@
  \centering
- \includegraphics[width=12cm]{characterisation}
+ \includegraphics{characterisation}
  % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
 \caption{Characterization of the crystal-gel.
-(A) Time evolution of the average size of the crystals (blue line) and the number of crystals (red line) for the state point $\phi=0.25$ and $c_p=0.38$.
+(A) Time evolution of the average size of the crystals (blue line) and the number of crystals (red line) for the state point $\phi=0.25$ and $c_p=0.38$ mg/g.
 (B) Radius of gyration of crystalline nuclei at the late stages of gelation for the same state point.}
@@
 Crystal gel formation.  
-(A) Reconstructions from confocal coordinates at $\phi=0.25$ and $c_p=0.38$.
+(A) Reconstructions from confocal coordinates at $\phi=0.25$ and $c_p=0.38$ mg/g.
 Depth of view is $5\sigma$.

### 2015-10-05 crystal-gel/phasediag.py
Cp*=2.87 mg/g

@@ -3,7 +3,7 @@ from colloids.phase import *
 from os.path import isfile
 
 qR = 0.071
-cpov = 1
+cpov = 2.8
 q = qR2q(qR)
 
 #experimental points
@@ -35,8 +35,8 @@ scatter(f2vf(fc), piv2y(pivc, q) * alpha(fc, qR) * cpov, c='r', marker='o')
 
 #draw binodal and spinodals
 for f,l in zip(GL.T[1:], ['--']*2+['-']*2):
-    cp = piv2y(GL[:,0], q) * alpha(f, qR)
-    plot(f2vf(f), cp * cpov, 'k'+l)
+    y = piv2y(GL[:,0], q) * alpha(f, qR)
+    plot(f2vf(f), y * cpov, 'k'+l)
 #draw tie lines
 for piv, fG, fL in GL[[50,60,70,80,85,88]+range(90,98),:3]:
     cpG = piv2y(piv, q) * alpha(fG, qR) * cpov
@@ -55,7 +55,7 @@ for f in FS.T[1:]:
     cp = piv2y(FS[:,0], q) * alpha(f, qR)
     plot(f2vf(f), cp * cpov, 'k.')
 
-ylabel(r'$c_p$ (g/L)')
+ylabel(r'$c_p$ (mg/g)')
 xlabel(r'$\phi$ (%)')
 ylim(0,2)
 xlim(0,0.75)
@@ -64,18 +64,18 @@ savefig('phasediag.pdf')
 
 #save binodal and spinodal
 for f,l in zip(GL.T[1:], ['bg', 'bl', 'sg', 'sl']):
-    cp = piv2y(GL[:,0], q) * alpha(f, qR)
+    y = piv2y(GL[:,0], q) * alpha(f, qR)
     np.savetxt(
         'gasliquid_%s.phd'%l,
-        np.column_stack((f2vf(f), cp * cpov)),
+        np.column_stack((f2vf(f), y * cpov)),
         header='phi\tcp',
         )
 #save fluid-crystal
 for f,l in zip(FS.T[1:], ['f', 'x']):
-    cp = piv2y(FS[:,0], q) * alpha(f, qR)
+    y = piv2y(FS[:,0], q) * alpha(f, qR)
     np.savetxt(
         'fluidcrystal_%s.phd'%l,
-        np.column_stack((f2vf(f), cp * cpov)),
+        np.column_stack((f2vf(f), y * cpov)),
         header='phi\tcp',
         )
 

### 2015-10-12 science/SOM2.tex
Update phase diagram with colloid diameter=2.2um and polymer radius 82nm. Update volume fractions accordingly. Cite Fleer2008

@@
 We use \textsc{pmma} (poly(methyl methacrylate)) colloids sterically stabilized with methacryloxypropyl terminated \textsc{pdms}(poly(dimethyl siloxane)) and fluorescently labelled with rhodamine isothiocyanate chemically bonded to the \textsc{pmma}.
-The average colloid diameter is 2.1 $\mu$m. 
+The average colloid diameter is 2.2 $\mu$m, measured via the diffusion constant in dilute conditions without polymer. The Brownian time is $\tau_B \approx 2.3$ s. 
 We assess that the size distribution of our particle is Gaussian with a polydispersity below 5\% via direct confocal measurements~\cite{Leocmach2013}.
@@
 Experiments are conducted at 27 $^\circ$C, some 80 $^\circ$C above the theta temperature in this solvents mixture~\cite{Royall2007}.
-A Flory scaling of the measurements of~\cite{lu2008gelation} yields a radius of gyration $R_g=76$ nm, and thus the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.07$.
+A Flory scaling of the measurements of~\cite{lu2008gelation} yields a radius of gyration $R_g=82$ nm, and thus the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.075$ and the overlap mass fraction of polymer $2.1$~mg/g.
 
@@
 Our spatial resolution is 192 nm / pixel.
-The Brownian time is $\tau_B \approx 4.5$ s.
-To be able to follow individual trajectories, we perform a 3D scan every 10 s at early time and every 30 s later.
+To be able to follow individual trajectories, we perform a 3D scan every 10~s ($\approx 4\tau_B$) at early time and every 30 s later.
 

### 2015-10-12 phasediag.py

@@ -1,17 +1,31 @@
 from matplotlib.pyplot import *
 from colloids.phase import *
 from os.path import isfile
+from scipy.constants import N_A
 
-qR = 0.071
-cpov = 2.8
+#radius of gyration in nm
+R = 82.
+#Molecular weight (Dalton or g/mol)
+Mw = 3.6e6
+#colloid diameter in nm
+sigma = 2200.
+#density of the solvent in g/mL
+d = 1.24
+
+qR = 2*R/sigma
+cpov = 1000 * (3 * Mw)/(4 * np.pi * (R*1e-9)**3 * N_A * d*1e6)
 q = qR2q(qR)
+print 'qR = %0.3f, q = %0.3f, Cov = %0.3f mg/g'%(qR, q, cpov)
 
 #experimental points
-phixp = np.array([0.25]*4+[0.1]*2)
-cpxp = np.array([0.38, 0.48, 0.57, 1.07, 0.82, 1.36])
+phixpG = np.array([0.069, 0.128, 0.135, 0.126, 0.333, 0.313, 0.294, 0.305])
+cpxpG = np.array([0.9, 0.786, 0.82, 1.36, 1.07, 0.38, 0.57, 0.48])
+phixpF = np.array([0.141, 0.141, 0.139, 0.084, 0.307, 0.309])
+cpxpF = np.array([0.25,0.48,0.57, 0.76, 0.28, 0.34])
 
 #load or compute theoretical phase diagram
 fc, pivc = Liu().critical_point(q)
+print 'At critical point colloid volume fraction is %0.3f and osmotic insertion work is %0.3f kT'%(f2vf(fc), pivc)
 if isfile('phasediag.txt'):
     GL = np.loadtxt('phasediag.txt')
 else:
@@ -28,32 +42,43 @@ fig = figure('experimental phase diagram')
 clf()
 
 #draw experimental points
-scatter(phixp, cpxp, marker='s')
+scatter(phixpG, cpxpG, marker='s')
+scatter(phixpF, cpxpF, color='k')
 
 #draw critical point
-scatter(f2vf(fc), piv2y(pivc, q) * alpha(fc, qR) * cpov, c='r', marker='o')
+scatter(f2vf(fc), piv2y(pivc, qR) * alpha(fc, q) * cpov, c='r', marker='o')
 
-#draw binodal and spinodals
[…]

### 2015-10-12 crystal-gel/science/crystal_gel_letter2.tex
Update phase diagram with colloid diameter=2.2um and polymer radius 82nm. Update volume fractions accordingly. Cite Fleer2008

@@
 %In the following we discuss all the stages of the formation of the crystal gels: phase separation, stress-driven agein and crystallization.
-The experimental data is taken at two different volume fractions ($\phi=0.10$ and $\phi=0.25$) and for different values
-of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ mg/g for $\phi=0.25$ and $c_p=0.82,1.36$ mg/g for $\phi=0.10$.
-In Fig. 1A, we superimpose the state points with the phase diagram calculated from the equation of state.
-For each state points, scans at early times (every $10s$) and at the late stage (every $30s$) of the gelation process.
+The experimental data is taken at two different volume fractions ($\phi\approx 0.13$ and $\phi\approx 0.30$) and for different values
+of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ mg/g for $\phi\approx 0.30$ and $c_p=0.82,1.36$ mg/g for $\phi\approx 0.13$.
+In Fig.~\ref{fig:phase_separation}A, we superimpose the state points with the phase diagram calculated from generalised free volume theory~\cite{Fleer2008}.
+For each state points, scans at early times (every $10$~s, i.e. $4\tau_B$) and at the late stage (every $30$~s) of the gelation process.
 
@@
 %follows different power laws $n(s)\sim s^\tau$, where $\tau$ is the Fisher exponent \cite{family2012kinetics}.
-%For state points with $\phi=0.25$ the decay is well described by the three dimensional (3D) random percolation exponent $\tau=2.18$ \cite{family2012kinetics}.
-%For state points with $\phi=0.10$, on the other hand, the decay appears to be slower, close to the diffusion limited cluster aggregation (DLCA) 
+%For state points with $\phi\approx 0.30$ the decay is well described by the three dimensional (3D) random percolation exponent $\tau=2.18$ \cite{family2012kinetics}.
+%For state points with $\phi\approx 0.13$, on the other hand, the decay appears to be slower, close to the diffusion limited cluster aggregation (DLCA) 
 %exponent $\tau=1.8$ \cite{family2012kinetics}. While it is plausible that gelation at lower volume fraction occurs through cluster aggregation, especially given 
@@
 %experimental conditions and statistical noise (which is high in our case since the distributions are extracted from a single frame).
-%Other measures of fractal dimensions (see below) will in fact suggest that also the $\phi=0.10$ states belong to the random gel universality class.
+%Other measures of fractal dimensions (see below) will in fact suggest that also the $\phi\approx 0.13$ states belong to the random gel universality class.
 %
@@
 % In Fig.~\ref{fig:stress}a, we plot the fraction of liquid and gas particles as a function of time during phase separation
-% at the state points of $\phi=0.10$. We observe that at high polymer concentration ($c_p=1.36$, dashed lines) both the
+% at the state points of $\phi\approx 0.13$. We observe that at high polymer concentration ($c_p=1.36$, dashed lines) both the
 % liquid and gas branch change monotonically with time as the phase separation proceeds. At low polymer concentration
@@
 % In Fig.~\ref{fig:stress}c we plot the time evolution of the structure factor, $S(q)$, for the state point
-% undergoing network reorganization $\phi=0.10$ and $c_p=0.82$. The time of network reorganization as found in Fig.~\ref{fig:stress}a and b
+% undergoing network reorganization $\phi\approx 0.13$ and $c_p=0.82$. The time of network reorganization as found in Fig.~\ref{fig:stress}a and b
 % corresponds to a steep increase of the low wavenumber scattering from the network, that signals the coarsening of the network structure. 
 % 
-% The same conclusions also hold for the state point at $\phi=0.25$. We observe network restructuring only for the state point
+% The same conclusions also hold for the state point at $\phi\approx 0.30$. We observe network restructuring only for the state point
 % at lower polymer concentration, $c_p=0.38$. Interestingly, this is also the only state point that undergoes crystallization,
@@
 But a difference
-in the fractal dimension $D$ between the $\phi=0.25$ and $c_p=0.38$ mg/g state point and other state points starts to become visible. In fact, while states
+in the fractal dimension $D$ between the $\phi\approx 0.30$ and $c_p=0.38$ mg/g state point and other state points starts to become visible. In fact, while states
 with high polymer concentration retain the exponent $D=2.53$,
 which is the random gel universality class exponent, as we also observed in the early stages of the gelation process (see Fig.~\ref{fig:phase_separation}B),
-the state point with low polymer concentration ($\phi=0.25$ and $c_p=0.38$ mg/g) displays the biggest deviation from that exponent. In the figure we also
+the state point with low polymer concentration ($\phi\approx 0.30$ and $c_p=0.38$ mg/g) displays the biggest deviation from that exponent. In the figure we also
 plot the exponent $D=3$, which corresponds to volume growth, which we will show in the following analysis to be the correct exponent for this state.
 The difference is due to the onset of crystallization in the low-polymer concentration sample. This is already evident in the high $q$ behavior of
-the structure factors (Fig.~\ref{fig:phase_separation}C), where the diffraction planes appear as sharper peaks for $\phi=0.25$ and $c_p=0.38$ mg/g.
+the structure factors (Fig.~\ref{fig:phase_separation}C), where the diffraction planes appear as sharper peaks for $\phi\approx 0.30$ and $c_p=0.38$ mg/g.
 
@@
 particle increases with its number of neighbors. Approaching close packing (12 neighbors), we clearly see two family of curves. The state point
-at $\phi=0.25$ and $c_p=0.38$ mg/g reaches close packing at a volume fraction of around $\sim 60\%$, which is in reasonable agreement with the
+at $\phi\approx 0.30$ and $c_p=0.38$ mg/g reaches close packing at a volume fraction of around $\sim 60\%$, which is in reasonable agreement with the
 volume fraction of the stable crystal phase. All other state points, instead, reach close packing at volume fraction $\sim 55\%$, which is
@@
 We should run the code of Mathieau with size detection}).
-This result clearly indicates that the state point at $\phi=0.25$ and $c_p=0.38$ mg/g follows a different arrest mechanism, in which phase separation
+This result clearly indicates that the state point at $\phi\approx 0.30$ and $c_p=0.38$ mg/g follows a different arrest mechanism, in which phase separation
 is arrested by crystallization and not glassiness. Although our estimation of the absolute volume fraction may involves errors of a few \%, the trend should be robust.
@@
 % while the crystallizing state point shows a scaling compatible with $D=3$, which
-% signals compact crystalline nuclei. At high $q$, the first signals from diffraction planes are also evident in the $\phi=0.25$ and $c_p=0.38$ state
+% signals compact crystalline nuclei. At high $q$, the first signals from diffraction planes are also evident in the $\phi\approx 0.30$ and $c_p=0.38$ state
 % point. This suggests that the space filling network is composed of crystalline aggregates, differently from the fractal branches which characterize the
@@
 To confirm the formation of crystalline seeds we adopt bond orientational analysis to detect which particles are in crystalline environments~\cite{russo2013interplay}.
-With this analysis we confirm that crystallization events start occurring for the state point $\phi=0.25$ and $c_p=0.38$ mg/g, which has the lowest
+With this analysis we confirm that crystallization events start occurring for the state point $\phi\approx 0.30$ and $c_p=0.38$ mg/g, which has the lowest
 polymer concentration of all state points. This is shown in Fig.~\ref{fig:crystals}A, where the black line indicates the average size of the crystallites,
@@
 %\end{itemize}
-All other state points show only negligible signs of crystallization. In particular, for state points with $\phi=0.25$, increasing
+All other state points show only negligible signs of crystallization. In particular, for state points with $\phi\approx 0.30$, increasing
 polymer concentration drastically reduces the amount of crystals. This is in agreement with the
@@
 To investigate this process, we plot in Fig.~\ref{fig:crystals}B the gyration radius of individual crystalline nuclei for the state point
-$\phi=0.25$ and $c_p=0.38$ mg/g. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with
+$\phi\approx 0.30$ and $c_p=0.38$ mg/g. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with
 the dimension of random percolation $D=2.53$, while at large sizes it scales as $D=3$ of compact crystal growth. This demonstrates that, while
@@
 In the next section we demonstrate that the volume growth is due to direct condensation of the gas phase on the crystalline seeds.
-First we show in Fig.~\ref{fig:transitions}A slabs of 10 particle diameter thickness for configurations at $\phi=0.25$ and $c_p=0.38$ mg/g at different times.
+First we show in Fig.~\ref{fig:transitions}A slabs of 10 particle diameter thickness for configurations at $\phi\approx 0.30$ and $c_p=0.38$ mg/g at different times.
 Particles are colored according to their phase. The figure shows the first nucleation
@@
 %In the next section we show evidence for the fact that the volume growth is due to direct condensation of the gas phase on the crystalline seeds.
-%First we show in Fig.~\ref{fig:transitions}a slabs of 10 particle diameter thickness for configurations at $\phi=0.25$ and $c_p=0.38$ at different times.
+%First we show in Fig.~\ref{fig:transitions}a slabs of 10 particle diameter thickness for configurations at $\phi\approx 0.30$ and $c_p=0.38$ at different times.
 %Particles are coloured according to their phase as follows. Green: hcp; red: fcc; blue: bcc; white: liquid or gas. The figure shows the first nucleation
@@
 Next we show in Fig.~\ref{fig:transitions}B the fraction of particles in each different phase for the crystallization trajectory at
-$\phi=0.25$ and $c_p=0.38$ mg/g, after the liquid-gas phase separation has occurred. The process of crystallization is characterized
+$\phi\approx 0.30$ and $c_p=0.38$ mg/g, after the liquid-gas phase separation has occurred. The process of crystallization is characterized
 by a steep decrease in liquid particles, as they transform in small crystalline nuclei inside the liquid domains. This decrease
@@
 % our experiments, but there is a crucial difference. In the simulation results~\cite{fortini2008crystallization} the kinetic
-% pathway at volume fraction $\phi=0.25$ consists of the following three steps: 
+% pathway at volume fraction $\phi\approx 0.30$ consists of the following three steps: 
 % spinodal decomposition; nucleation of crystalline nuclei within the fluid branches;
@@
 % As shown above, the last step appears to be markedly different in our experiments.
-% More precisely, at $\phi=0.25$ and polymer concentration $c_p=0.38$ we also observe the first two steps, 
+% More precisely, at $\phi\approx 0.30$ and polymer concentration $c_p=0.38$ we also observe the first two steps, 
 % but the growth process is entirely different. Instead of growing inside
@@
 Depending on the polymer concentration, there are two possible arrest mechanisms.
-(i) Crystallization, observed at $\phi=0.25$ and $c_p=0.38$ mg/g, small crystalline nuclei appear inside the liquid, reach the surface
+(i) Crystallization, observed at $\phi\approx 0.30$ and $c_p=0.38$ mg/g, small crystalline nuclei appear inside the liquid, reach the surface
  of the liquid branches, and then grow by addition of particles from the gas phase. The final structure is a network of crystal droplets,
@@
 The continuous line is the gas side of the metastable gas-liquid spinodal region with the black diamond marking the critical point.
-The other symbols are the experimental points, consistently used in all panels.
+Empty black triangles are experimental point showing a fluid behaviour.
+The other symbols are the experimental points showing gel behaviour, consistently used in all panels.
 (B) Radius of gyration ($R_g$) as a function of cluster size ($s$) for all state points at percolation. The
@@
 \caption{Characterization of the crystal-gel.
-(A) Time evolution of the average size of the crystals (blue line) and the number of crystals (red line) for the state point $\phi=0.25$ and $c_p=0.38$ mg/g.
+(A) Time evolution of the average size of the crystals (blue line) and the number of crystals (red line) for the state point $\phi\approx 0.30$ and $c_p=0.38$ mg/g.
 (B) Radius of gyration of crystalline nuclei at the late stages of gelation for the same state point.}
@@
 Crystal gel formation.  
-(A) Reconstructions from confocal coordinates at $\phi=0.25$ and $c_p=0.38$ mg/g.
+(A) Reconstructions from confocal coordinates at $\phi\approx 0.30$ and $c_p=0.38$ mg/g.
 Depth of view is $5\sigma$.

### 2015-10-13 crystal-gel/science/crystal_gel_letter2.tex
Small change.

@@
 In the next section we demonstrate that the volume growth is due to direct condensation of the gas phase on the crystalline seeds.
-First we show in Fig.~\ref{fig:transitions}A slabs of 10 particle diameter thickness for configurations at $\phi\approx 0.30$ and $c_p=0.38$ mg/g at different times.
+Supplementary Movie 2 shows the confocal microscopy video for the crystallizing sample at $\phi\approx 0.30$ and $c_p=0.38$ mg/g,
+where the process of crystal growth by direct attachment of gas particles can be observed directly.
+From the trajectory, we extract slabs of 10 particle diameter thickness for configurations at $\phi\approx 0.30$ and $c_p=0.38$ mg/g at different times,
+shown in Fig.~\ref{fig:transitions}A slabs.
 Particles are colored according to their phase. The figure shows the first nucleation

### 2015-10-13 crystal-gel/science/SOM2.tex
Corrected the SOM2

@@
 \noindent
-{\bf Experimental.}
+{\bf Experiments}
+
 We use \textsc{pmma} (poly(methyl methacrylate)) colloids sterically stabilized with methacryloxypropyl terminated \textsc{pdms}(poly(dimethyl siloxane)) and fluorescently labelled with rhodamine isothiocyanate chemically bonded to the \textsc{pmma}.
-The average colloid diameter is 2.1 $\mu$m. 
+The average colloid diameter is 2.2 $\mu$m, measured via the diffusion constant in dilute conditions without polymer. The Brownian time is $\tau_B \approx 2.3$ s. 
 We assess that the size distribution of our particle is Gaussian with a polydispersity below 5\% via direct confocal measurements~\cite{Leocmach2013}.
@@
 Experiments are conducted at 27 $^\circ$C, some 80 $^\circ$C above the theta temperature in this solvents mixture~\cite{Royall2007}.
-A Flory scaling of the measurements of~\cite{lu2008gelation} yields a radius of gyration $R_g=76$ nm, and thus the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.07$.
+A Flory scaling of the measurements of~\cite{lu2008gelation} yields a radius of gyration $R_g=82$ nm, and thus the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.075$ and the overlap mass fraction of polymer $2.1$~mg/g.
 
-In the absence of salt, the Debye length is expected to reach several $\mu$m and the (weakly) charged colloids experience a long range electrostatic repulsion.
+In the absence of salt, the Debye length is expected to reach a few $\mu$m and the (weakly) charged colloids experience a long range electrostatic repulsion.
 We confirm that colloids never come close enough to feel the short-ranged attraction.
-Screening by tetrabutylammonium bromide (Fluka) at saturated concentration brings down the Debye length to about 100 nm, practically discarding the repulsion.
+Screening by tetrabutylammonium bromide (Fluka) at saturated concentration brings down the Debye length to about several nanometers practically discarding the repulsion.
 Thus, salt injection can screen the Coulomb repulsion and make the polymer-induced depletion attraction effective, initiating phase separation.
@@
 The scanning volume is 98 $\times$ 98 $\times$ 53 $\mu$m$^3$, which contains $\sim 10^4$ colloid particles.
-The Brownian time is $\tau_B \approx 4.5$ s.
-To be able to follow individual trajectories, we perform a 3D scan every 10 s at early time and every 30 s later.
+Our spatial resolution is 192 nm / pixel.
+To be able to follow individual trajectories, we perform a 3D scan every 10~s ($\approx 4\tau_B$) at early time and every 30 s later.
 
@@
 \noindent
-{\bf Structural analysis.}
+{\bf Structural analysis}
+
 To account for imprecision in particle localisation close to contact, we consider two particles bonded if their center-to-center distance is within 2.3 $\mu$m $>\sigma+2R_g$.
@@
 
+\noindent
+{\bf Supplementary Movies}
+
+\emph{Supplementary Movie 1}.
+This movie shows the early stage of gelation process of $\phi\approx 0.30$ and $c_p=0.38$ mg/g, cropped from the original confocal images at the same z-position as Fig.~3A with scanning duration of 1160 s (~580~$\tau$).
+We quickly started scanning the sample after exchanging salt-reservoir with the solvent including salt.
+At the beginning, salt had not reached the scanning volume and colloids were dispersed by electrostatic repulsion.
+At 0:02 in the movie, salt diffused into the scanning volume and screened the surface charges of colloids.
+Then, colloids immediately initiated phase separation by short-ranged attractions.
+
+\emph{Supplementary Movie 2}.
+This movie shows the late stage after Supplemental Movie 1 with scanning duration of 27000~s (~13500$\tau$).
+
 %\bibliographystyle{Science}

### 2015-10-15 science/crystal_gel_letter2.tex
New introduction. Additions to the References. Other small changes to the text.

@@
 \usepackage{color}
+
+\usepackage[normalem]{ulem}
 
@@
 However the particle-level details of this process have not yet been confirmed experimentally.
-Here we show by confocal microscopy measurements that crystallization leads to a network structure different from the original phase separation pattern.
+\textcolor{blue}{
+With confocal microscopy measurements we access the microscopic mechanism of crystal growth in a three phase system of a colloid-polymer mixture
+undergoing viscoelastic phase separation.} We show that crystallization leads to a network structure different from the original phase separation pattern.
 Crystals nucleate inside the liquid network, and grow past it by direct condensation of the gas phase on their surface, driving evaporation of nearby liquid.
@@
 
-Gels and glasses are two important non-ergodic disordered states, which are very important 
-in our life and nature~\cite{anderson2002insights,lekkerkerker2011colloids}.
-The dynamical arrest of colloidal gels and glasses is believed to originate from a dynamical transition to the same non-ergodic disordered glass state, i.e., vitrification.
-The only difference comes from the fact that the former is formed by densification induced by phase separation and thus is spatially heterogeneous, whereas the latter is formed by uniform densification and is thus spatially homogeneous.
-Gels are distinct from glasses because of their heterogeneous network structure, which gives them properties intermediate between
-those of a solid and a fluid phase, and allows the coexistence of two properties usually incompatible with each other, 
-elasticity and fluidity. 
-The elasticity of gels originates from the percolated network structure and the resulting space-spanning connectivity. 
-At the same time a fluid component can easily flow through the elastic network.
-There are two types of gels with such characteristics: one is a network structure stabilized by stable crosslinks, and the other is a phase-separated structure frozen by dynamical arrest due to vitrification.
-Many chemical and physical gels belong to the former category, in which the size of network pores is rather small and 
-the fluid transport is not so efficient. On the other hand, gels made of spherical colloids and proteins 
-with attractive interactions belong to the latter category. 
-This type of gels are formed by viscoelastic phase separation \cite{tanaka1999colloid,tanaka2000viscoelastic} into a dense liquid phase with slow dynamics and a dilute gas phase 
-with fast dynamics. This difference in the viscoelastic properties between the two phases allows the system to form a space-spanning 
+\textcolor{blue}{Crystallization plays a fundamental role in many processes occurring in nature, such as cloud composition, and in technologial applications
+that are at the core of the chemical, pharmaceutical and food industries.
+Many of the properties of crystals, like the polymorph type and size distribution of the crystallites, depend on the conditions at which
+the nucleation process took place. Controlling the early stages of crystallization is thus of fundamental importance in order to
+obtain in a reproducible manner crystals with the desired properties.}
+
+\textcolor{blue}{Under usual conditions, the nucleation stage involves a very small number of molecules, of the order of 10-1000 molecules.
+This has severily limited the possibility to observe the nucleation process directly, and has also posed limits on
+models based on macroscopic thermodynamic properties.
+The inability to directly observe the nucleation process is also a limit to our understanding of its kinetic pathway.
+Classical Nucleation Theory assumes the crystallization processes to occurr in one step,
+where the transition can be described by just one reaction coordinate. An example is the formation of an ordered crystalline nucleus
+directly from the supersatured solution. But the possibility of different pathways has recently come to prominence with
+the discovery of two-step nucleation pathways, where crystallization is preceeded by the formation of dense liquid droplets~\cite{ten1997enhancement,SearR,savage2009experimental,vekilov2010two,palberg2014crystallization}.
+Understanding the process of crystal formation in mixed-phase systems (composed of gas, liquid, and solid phases) is thus
+of great importance for a variety of systems, from protein solutions to clouds.}
+
+\textcolor{blue}{Colloidal solutions offer a system where the crystallization process in a mixed-phase environment can be observed with single particle
+resolution, and at the same time scales over which nucleation takes place. In Colloids with short-range attractions the liquid-gas transition
+becomes metastable with respect to crystallization~\cite{anderson2002insights,lekkerkerker2011colloids}, and can form gels
+by viscoelastic phase separation \cite{tanaka1999colloid,tanaka2000viscoelastic}, with the coexistence of a dense liquid phase with slow dynamics and a dilute gas phase with fast dynamics.}
+This difference in the viscoelastic properties between the two phases allows the system to form a space-spanning 
 network structure of the liquid phase, even if it is the minority phase. The network is fractal and its dynamic structure factor shows a stretched-exponential decay to a finite
-plateau~\cite{krall1998internal,solomon2001dynamic,romer2000sol}. During phase separation, the density of the liquid phase 
+plateau~\cite{krall1998internal,solomon2001dynamic,romer2000sol}.
+During phase separation, the density of the liquid phase 
 increases towards the glass-transition density, leading to slow glassy dynamics.
@@
 More precisely, the dynamical stabilization is due to percolation of locally favored structures, which are locally stable non-crystalline structures \cite{royall2008g}.
+
+% Gels and glasses are two important non-ergodic disordered states, which are very important 
+% in our life and nature~\cite{anderson2002insights,lekkerkerker2011colloids}.
+% The dynamical arrest of colloidal gels and glasses is believed to originate from a dynamical transition to the same non-ergodic disordered glass state, i.e., vitrification.
+The only difference comes from the fact that the former is formed by densification induced by phase separation and thus is spatially heterogeneous, whereas the latter is formed by uniform densification and is thus spatially homogeneous.
+% Gels are distinct from glasses because of their heterogeneous network structure, which gives them properties intermediate between
+% those of a solid and a fluid phase, and allows the coexistence of two properties usually incompatible with each other, 
+% elasticity and fluidity. 
+% The elasticity of gels originates from the percolated network structure and the resulting space-spanning connectivity. 
+% At the same time a fluid component can easily flow through the elastic network.
+% There are two types of gels with such characteristics: one is a network structure stabilized by stable crosslinks, and the other is a phase-separated structure frozen by dynamical arrest due to vitrification.
+% Many chemical and physical gels belong to the former category, in which the size of network pores is rather small and 
+% the fluid transport is not so efficient. On the other hand, gels made of spherical colloids and proteins 
+% with attractive interactions belong to the latter category. 
+% This type of gels are formed by viscoelastic phase separation \cite{tanaka1999colloid,tanaka2000viscoelastic} into a dense liquid phase with slow dynamics and a dilute gas phase 
+% with fast dynamics. This difference in the viscoelastic properties between the two phases allows the system to form a space-spanning 
+% network structure of the liquid phase, even if it is the minority phase. The network is fractal and its dynamic structure factor shows a stretched-exponential decay to a finite
+% plateau~\cite{krall1998internal,solomon2001dynamic,romer2000sol}. During phase separation, the density of the liquid phase 
+% increases towards the glass-transition density, leading to slow glassy dynamics.
+% At the glass transition point, the percolated network structure is dynamically stabilised by vitrification 
+% of the dense liquid phase~\cite{pusey1993dynamics,ilett1995phase,verhaegh1997transient,tanaka1999colloid,foffi2002,buzzaccaro2007sticky,zaccarelli2007,lu2008gelation,zaccarelli2008gelation,testard2011}.
+% More precisely, the dynamical stabilization is due to percolation of locally favored structures, which are locally stable non-crystalline structures \cite{royall2008g}.
 % Since ``glass'' is the term used for bulk, this may be a more appropriate way to express the mechanism of dynamical arrest. 
@@
 
-In this Letter, we show that, for low polymer concentration, the system can undergo a novel crystallization pathway that leads to the formation
+In this Report, we show that, for low polymer concentration, the system can undergo a novel crystallization pathway that leads to the formation
 of crystalline droplets.
@@
 In our colloidal system, we then call this process ``colloidal rain''
-formation. Due to the difference in the nucleation and growth processes, the droplets display different scaling laws depending on their
+formation.
+\textcolor{blue}{\sout{Due to the difference in the nucleation and growth processes, the droplets display different scaling laws depending on their
 size. For small nuclei, the droplets have the same fractal dimension of the embedding gel network. But during the growth process, droplets
-grow with an integer fractal dimension due to direct condensation of gas particles.
-This is in contrast to the standard crystal-gel scenario, where crystals nucleate and grow inside the dense branches of the gel, becoming dynamically arrested.
+grow with an integer fractal dimension due to direct condensation of gas particles.}} This is in contrast to the standard crystal-gel scenario, where crystals nucleate and grow inside the dense branches of the gel, becoming dynamically arrested.
 We also observe that increasing the strength of the interparticle attraction
@@
 %In the following we discuss all the stages of the formation of the crystal gels: phase separation, stress-driven agein and crystallization.
-The experimental data is taken at two different volume fractions ($\phi=0.10$ and $\phi=0.25$) and for different values
-of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ for $\phi=0.25$ and $c_p=0.82,1.36$ for $\phi=0.10$.
-In Fig. 1A, we superimpose the state points with the phase diagram calculated from the equation of state.
-For each state points, scans at early times (every $10s$) and at the late stage (every $30s$) of the gelation process.
+The experimental data is taken at two different volume fractions ($\phi\approx 0.13$ and $\phi\approx 0.30$) and for different values
+of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ mg/g for $\phi\approx 0.30$ and $c_p=0.82,1.36$ mg/g for $\phi\approx 0.13$.
+In Fig.~\ref{fig:phase_separation}A, we superimpose the state points with the phase diagram \textcolor{blue}{which we calculated from generalised free volume theory~\cite{Fleer2008} using the experimental values for the size of the colloids, the radius of gyration of the polymer and the polymer concentration (see Supplementary Information)}.
+For each state points, scans at early times (every $10$~s, i.e. $4\tau_B$) and at the late stage (every $30$~s) of the gelation process.
 
@@
 eventually forming a percolating network. Thus, a dense network (liquid) coexists with freely diffusing monomers (gas).
+\textcolor{blue}{In Fig.~\ref{fig:phase_separation}A, the state points which didn't phase separate are indicated with open triangles, while
+the other symbols indicate samples undergoing phase separation. The (black) continous line is the gas spinodal computed from free volume theory,
+and it is in good agreement with our experimental points (the small discrepancy at small colloidal volume fractions
+is common in free volume theory)~\cite{Royall2007,lu2008gelation}}.
 The topology of the network is similar to all our samples at the initial stages.
@@
 But a difference
-in the fractal dimension $D$ between the $\phi=0.25$ and $c_p=0.38$ state point and other state points starts to become visible. In fact, while states
+in the fractal dimension $D$ between the $\phi\approx 0.30$ and $c_p=0.38$ mg/g state point and other state points starts to become visible. In fact, while states
 with high polymer concentration retain the exponent $D=2.53$,
 which is the random gel universality class exponent, as we also observed in the early stages of the gelation process (see Fig.~\ref{fig:phase_separation}B),
-the state point with low polymer concentration ($\phi=0.25$ and $c_p=0.38$) displays the biggest deviation from that exponent. In the figure we also
+the state point with low polymer concentration ($\phi\approx 0.30$ and $c_p=0.38$ mg/g) displays the biggest deviation from that exponent. In the figure we also
 plot the exponent $D=3$, which corresponds to volume growth, which we will show in the following analysis to be the correct exponent for this state.
 The difference is due to the onset of crystallization in the low-polymer concentration sample. This is already evident in the high $q$ behavior of
-the structure factors (Fig.~\ref{fig:phase_separation}C), where the diffraction planes appear as sharper peaks for $\phi=0.25$ and $c_p=0.38$.
+the structure factors (Fig.~\ref{fig:phase_separation}C), where the diffraction planes appear as sharper peaks for $\phi\approx 0.30$ and $c_p=0.38$ mg/g.
 
@@
 particle increases with its number of neighbors. Approaching close packing (12 neighbors), we clearly see two family of curves. The state point
-at $\phi=0.25$ and $c_p=0.38$ reaches close packing at a volume fraction of around $\sim 60\%$, which is in reasonable agreement with the
-volume fraction of the stable crystal phase. All other state points, instead, reach close packing at volume fraction $\sim 55\%$, which is
-close to the volume fraction of the glass state. 
+at $\phi\approx 0.30$ and $c_p=0.38$ mg/g reaches \textcolor{blue}{close packing at a volume fraction of around $\sim 68\%$, which is in reasonable agreement with the
+volume fraction of the stable crystal phase~\cite{ilett1995phase}. All other state points, instead, reach close packing at volume fraction $\sim 64\%$, which is
+indeed close to the volume fraction of the attractive glass state~\cite{pham2002multiple}}. 
 %(\textbf{the absolute values of these volume fraction are not totally convincing, probably because of a not optimal estimate of the size of the colloids.
 We should run the code of Mathieau with size detection}).
-This result clearly indicates that the state point at $\phi=0.25$ and $c_p=0.38$ follows a different arrest mechanism, in which phase separation
+This result clearly indicates that the state point at $\phi\approx 0.30$ and $c_p=0.38$ mg/g follows a different arrest mechanism, in which phase separation
 is arrested by crystallization and not glassiness. Although our estimation of the absolute volume fraction may involves errors of a few \%, the trend should be robust.
@@
 To confirm the formation of crystalline seeds we adopt bond orientational analysis to detect which particles are in crystalline environments~\cite{russo2013interplay}.
-With this analysis we confirm that crystallization events start occurring for the state point $\phi=0.25$ and $c_p=0.38$, which has the lowest
+With this analysis we confirm that crystallization events start occurring for the state point $\phi\approx 0.30$ and $c_p=0.38$ mg/g, which has the lowest
 polymer concentration of all state points. This is shown in Fig.~\ref{fig:crystals}A, where the black line indicates the average size of the crystallites,
@@
 which was neither considered nor observed before.
-In the following we will show that the growth mechanism of the crystal is not only due to filling of the spinodal liquid network, as was observed in Refs.~\cite{fortini2008crystallization,perez2011pathways} for Brownian Dynamics simulations, but also involves direct condensation of the gas phase
-on the crystalline seeds.
+In the following we will show that the growth mechanism of the crystal is not only due to filling of the spinodal liquid network, as was observed in Refs.~\cite{fortini2008crystallization,perez2011pathways} for Brownian Dynamics simulations, but also involves
+\textcolor{blue}{evaporation of the liquid phase and subsequent direct condensation of the gas phase
+on the crystalline seeds.}
 To investigate this process, we plot in Fig.~\ref{fig:crystals}B the gyration radius of individual crystalline nuclei for the state point
-$\phi=0.25$ and $c_p=0.38$. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with
+$\phi\approx 0.30$ and $c_p=0.38$ mg/g. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with
 the dimension of random percolation $D=2.53$, while at large sizes it scales as $D=3$ of compact crystal growth. This demonstrates that, while
@@
 In the next section we demonstrate that the volume growth is due to direct condensation of the gas phase on the crystalline seeds.
-First we show in Fig.~\ref{fig:transitions}A slabs of 10 particle diameter thickness for configurations at $\phi=0.25$ and $c_p=0.38$ at different times.
+Supplementary Movie 2 shows the confocal microscopy video for the crystallizing sample at $\phi\approx 0.30$ and $c_p=0.38$ mg/g,
+where the process of crystal growth by direct attachment of gas particles can be observed directly.
+From the trajectory, we extract slabs of 10 particle diameter thickness for configurations at $\phi\approx 0.30$ and $c_p=0.38$ mg/g at different times,
+shown in Fig.~\ref{fig:transitions}A slabs.
 Particles are colored according to their phase. The figure shows the first nucleation
@@
 Next we show in Fig.~\ref{fig:transitions}B the fraction of particles in each different phase for the crystallization trajectory at
-$\phi=0.25$ and $c_p=0.38$, after the liquid-gas phase separation has occurred. The process of crystallization is characterized
+$\phi\approx 0.30$ and $c_p=0.38$ mg/g, after the liquid-gas phase separation has occurred. The process of crystallization is characterized
 by a steep decrease in liquid particles, as they transform in small crystalline nuclei inside the liquid domains. This decrease
@@
 examine in Fig.~\ref{fig:transitions}C the transfer matrix probabilities $P_{\alpha\beta}$, where $\alpha$ and $\beta$ are each of the
-phases gas, liquid and crystal (which includes bcc, hcp, fcc and surface particles). Liquid particles undergo two processes, on one
+phases gas, liquid and crystal (which includes hcp, fcc and surface particles). Liquid particles undergo two processes, on one
 hand they transform to crystallites growing inside the liquid branches of the network, and also increasingly transform to gas particles.
@@
 Depending on the polymer concentration, there are two possible arrest mechanisms.
-(i) Crystallization, observed at $\phi=0.25$ and $c_p=0.38$, small crystalline nuclei appear inside the liquid, reach the surface
+(i) Crystallization, observed at $\phi\approx 0.30$ and $c_p=0.38$ mg/g, small crystalline nuclei appear inside the liquid, reach the surface
  of the liquid branches, and then grow by addition of particles from the gas phase. The final structure is a network of crystal droplets,
@@
 with respect to the ice crystals. This causes the water droplets to evaporate and sublimate directly on the ice crystals. Our system can then
-be regarded as a colloidal analogue for this important process, which can now be studied at the single-particle level.
+be regarded as a colloidal analogue for this important process, \textcolor{blue}{which, for the first time, we can observe at the single-particle level.}
 
@@
 \caption{Characterization of the crystal-gel.
-(A) Time evolution of the average size of the crystals (blue line) and the number of crystals (red line) for the state point $\phi=0.25$ and $c_p=0.38$.
+(A) Time evolution of the average size of the crystals (blue line) and the number of crystals (red line) for the state point $\phi\approx 0.30$ and $c_p=0.38$ mg/g.
 (B) Radius of gyration of crystalline nuclei at the late stages of gelation for the same state point.}
@@
 Crystal gel formation.  
-(A) Reconstructions from confocal coordinates at $\phi=0.25$ and $c_p=0.38$.
+(A) Reconstructions from confocal coordinates at $\phi\approx 0.30$ and $c_p=0.38$ mg/g.
 Depth of view is $5\sigma$.

### 2015-10-16 crystal-gel/science/crystal_gel_letter2.tex

@@
 
-\title{``Colloidal rain'' as a novel route to crystallization via viscoelastic phase separation}
+\title{A novel route to crystallization via viscoelastic phase separation}
 
@@
 
-\date{Received}
+%\date{Received}
 
@@
 \maketitle
-\vspace{-1cm}
+\vspace{-1.5cm}
 \begin{sciabstract}
@@
 However the particle-level details of this process have not yet been confirmed experimentally.
-Here we show by confocal microscopy measurements that crystallization leads to a network structure different from the original phase separation pattern.
-Crystals nucleate inside the liquid network, and grow past it by direct condensation of the gas phase on their surface, driving evaporation of nearby liquid.
-This process represents the colloidal analogue of the Bergeron process, i.e., ice crystals formation that occurs in mixed phase clouds and is at the origin of rain.
-Our \emph{colloidal rain} gives us full experimental access to the kinetic pathway of this phenomenon important for climate science.
-We argue that states similar to colloidal rain can form as the result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems, whose dynamics is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
+\textcolor{blue}{
+With confocal microscopy measurements we find a novel crystallization pathway of a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation.} We show that crystallization leads to a network structure different from the original phase separation pattern.
+Crystals nucleate inside the liquid network, \textcolor{blue}{but} grow past it by direct condensation of the gas phase on their surface, driving evaporation of nearby liquid.
+This process represents the colloidal analogue of the Bergeron process, i.e., ice crystals formation in mixed phase clouds and is at the origin of rain.
+\textcolor{blue}{Our finding} gives us full experimental access to the kinetic pathway of this phenomenon important for climate science.
+We argue that similar states can form as the result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems, if dynamics of the liquid phase is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
 \end{sciabstract}
@@
 
-Gels and glasses are two important non-ergodic disordered states, which are very important 
-in our life and nature~\cite{anderson2002insights,lekkerkerker2011colloids}.
-The dynamical arrest of colloidal gels and glasses is believed to originate from a dynamical transition to the same non-ergodic disordered glass state, i.e., vitrification.
-The only difference comes from the fact that the former is formed by densification induced by phase separation and thus is spatially heterogeneous, whereas the latter is formed by uniform densification and is thus spatially homogeneous.
-Gels are distinct from glasses because of their heterogeneous network structure, which gives them properties intermediate between
-those of a solid and a fluid phase, and allows the coexistence of two properties usually incompatible with each other, 
-elasticity and fluidity. 
-The elasticity of gels originates from the percolated network structure and the resulting space-spanning connectivity. 
-At the same time a fluid component can easily flow through the elastic network.
-There are two types of gels with such characteristics: one is a network structure stabilized by stable crosslinks, and the other is a phase-separated structure frozen by dynamical arrest due to vitrification.
-Many chemical and physical gels belong to the former category, in which the size of network pores is rather small and 
-the fluid transport is not so efficient. On the other hand, gels made of spherical colloids and proteins 
-with attractive interactions belong to the latter category. 
-This type of gels are formed by viscoelastic phase separation \cite{tanaka1999colloid,tanaka2000viscoelastic} into a dense liquid phase with slow dynamics and a dilute gas phase 
-with fast dynamics. This difference in the viscoelastic properties between the two phases allows the system to form a space-spanning 
+Crystallization plays a fundamental role in many processes occurring in nature, such as cloud composition, and in technological applications
+that are at the core of the chemical, pharmaceutical, and food industries.
+Many of the properties of crystals, like the \textcolor{blue}{shape, spatial arrangement}, polymorph type, and size distribution of the crystallites, depend on the conditions at which
+the nucleation process took place. Controlling the early stages of crystallization is thus of fundamental importance in order to
+obtain in a reproducible manner crystals with the desired properties.
+
+Under usual conditions, the nucleation stage involves a very small number of molecules, of the order of 10-1000 molecules.
+This has severely limited the possibility to observe the nucleation process directly, and has also posed limits on
+models based on macroscopic thermodynamic properties. 
+The inability to directly observe the nucleation process is also a limit to our understanding of its kinetic pathway. 
+Classical Nucleation Theory assumes the crystallization processes to occur in one step,
+where the transition can be described by just one reaction coordinate. An example is the formation of an ordered crystalline nucleus
+directly from the supersaturated solution. But the possibility of different pathways has recently come to prominence with
+the discovery of two-step nucleation pathways, where crystallization is proceeded by the formation of dense liquid  droplets \textcolor{blue}{as an intermediate step}~\cite{ten1997enhancement,SearR,savage2009experimental,vekilov2010two,palberg2014crystallization}.
+Understanding the process of crystal formation in mixed-phase systems (composed of gas, liquid, and solid phases) is thus
+of great importance for a variety of systems, from protein solutions to clouds.
+
+Colloidal suspensions offer a system where the crystallization process in a mixed-phase environment can be observed with single-particle
+resolution, and at the same time scales over which nucleation takes place. In colloids with short-range attractions the liquid-gas transition 
+becomes metastable with respect to crystallization~\cite{anderson2002insights,lekkerkerker2011colloids}, and can form gels
+by viscoelastic phase separation \cite{tanaka1999colloid,tanaka2000viscoelastic}, with the coexistence of a dense liquid phase with slow dynamics and a dilute gas phase with fast dynamics. 
+This difference in the viscoelastic properties between the two phases allows the system to form a space-spanning 
 network structure of the liquid phase, even if it is the minority phase. The network is fractal and its dynamic structure factor shows a stretched-exponential decay to a finite
-plateau~\cite{krall1998internal,solomon2001dynamic,romer2000sol}. During phase separation, the density of the liquid phase 
-increases towards the glass-transition density, leading to slow glassy dynamics.
+plateau~\cite{krall1998internal,solomon2001dynamic,romer2000sol}. 
+During phase separation, the density of the liquid phase increases towards the glass-transition density, leading to slow glassy dynamics.
 At the glass transition point, the percolated network structure is dynamically stabilised by vitrification 
@@
 observation of the initial stages of crystallization. This initial perturbation may even alter the selection of the final non-equilibrium arrested state.  
-In our experiments we instead succeed in building a novel setup that allows the time evolution of the system to be observed directly
+In our experiments we instead succeed in building a novel \textcolor{blue}{protocol} that allows the time evolution of the system to be observed directly
 under the confocal microscope from the very early stages and in a quiescent situation without introducing fluid flows.
 
-In this Letter, we show that, for low polymer concentration, the system can undergo a novel crystallization pathway that leads to the formation
+In this Report, we show that, for low polymer concentration, the system can undergo a novel crystallization pathway that leads to the formation
 of crystalline droplets.
-These droplets originate from crystal seeds that form inside the liquid branches of the spinodal aggregate, but then grow by direct condensation from the gas phase.
+These droplets originate from crystal seeds that form inside the liquid branches of the spinodal aggregate \textcolor{blue}{of network morphology}, but then grow by direct condensation from the gas phase.
 The process is analogous to the Bergeron process of ice formation in mixed phase clouds, where ice droplets grow at the
@@
 This is speculated to be the primary mechanism for the formation of raindrops. 
-In our colloidal system, we then call this process ``colloidal rain''
-formation. Due to the difference in the nucleation and growth processes, the droplets display different scaling laws depending on their
-size. For small nuclei, the droplets have the same fractal dimension of the embedding gel network. But during the growth process, droplets
-grow with an integer fractal dimension due to direct condensation of gas particles.
-This is in contrast to the standard crystal-gel scenario, where crystals nucleate and grow inside the dense branches of the gel, becoming dynamically arrested.
+\textcolor{blue}{The mechanism found here} is in contrast to the standard crystal-gel scenario, where crystals nucleate and grow inside the dense branches of the gel, becoming dynamically arrested.
 We also observe that increasing the strength of the interparticle attraction
-(by increasing the polymer concentration), instead of forming colloidal rain, the system undergoes the classical gelation process, in which
+(by increasing the polymer concentration), instead of forming \textcolor{blue}{crystal gels}, the system undergoes the classical gelation process, in which
 phase separation is interrupted by vitrification~\cite{verhaegh1997transient,tanaka1999colloid,Royall2007,lu2008gelation}.
@@
 %In the following we discuss all the stages of the formation of the crystal gels: phase separation, stress-driven agein and crystallization.
-The experimental data is taken at two different volume fractions ($\phi\approx 0.13$ and $\phi\approx 0.30$) and for different values
+The experimental data are taken at two different volume fractions ($\phi\approx 0.13$ and $\phi\approx 0.30$) and for different values
 of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ mg/g for $\phi\approx 0.30$ and $c_p=0.82,1.36$ mg/g for $\phi\approx 0.13$.
-In Fig.~\ref{fig:phase_separation}A, we superimpose the state points with the phase diagram calculated from generalised free volume theory~\cite{Fleer2008}.
-For each state points, scans at early times (every $10$~s, i.e. $4\tau_B$) and at the late stage (every $30$~s) of the gelation process.
+In Fig.~\ref{fig:phase_separation}A, we superimpose the state points with the phase diagram, which we calculated from generalised free volume theory~\cite{Fleer2008} using the experimental values for the size of the colloids, the radius of gyration of the polymer, and the polymer concentration (see Supplementary Information).
+For each state point, scans \textcolor{blue}{were made} at early times (every $10$~s, i.e. $4\tau_B$) and at the late stage (every $30$~s) of the gelation process.
 
@@
 eventually forming a percolating network. Thus, a dense network (liquid) coexists with freely diffusing monomers (gas).
+In Fig.~\ref{fig:phase_separation}A, the state points which did not phase separate are indicated with open triangles, while
+the other symbols indicate samples undergoing phase separation. The (black) continuous line is the gas spinodal computed from free volume theory,
+and it is in good agreement with our experimental points (the small discrepancy at small colloidal volume fractions 
+is common in free volume theory)~\cite{Royall2007,lu2008gelation}. 
 The topology of the network is similar to all our samples at the initial stages. 
@@
 
-While sharing similar structural properties at early times, the samples differ during the later stages of the phase separation process.
+While sharing similar structural properties at early times, the samples \textcolor{blue}{show significantly different behaviors in} the later stages of the phase separation process.
 To examine this in more detail, we plot in Fig.~\ref{fig:phase_separation}C the structure factor for the final stages of the gelation process for all state points.
@@
 The figure shows that at low wavenumber $q$, the structure factor displays fractal scaling compatible with Guinier law, $S(q)\sim q^{-D}$.
-But a difference
-in the fractal dimension $D$ between the $\phi\approx 0.30$ and $c_p=0.38$ mg/g state point and other state points starts to become visible. In fact, while states
-with high polymer concentration retain the exponent $D=2.53$,
+But a difference in the fractal dimension $D$ between the $\phi\approx 0.30$ and $c_p=0.38$ mg/g state point and other state points starts to become visible. In fact, while states
+with high polymer concentration retain the exponent $D=2.5(3)$, 
 which is the random gel universality class exponent, as we also observed in the early stages of the gelation process (see Fig.~\ref{fig:phase_separation}B),
@@
 through Voronoi diagrams, as a function of the number of neighbors. For all state points, we can see that the average volume fraction of a colloidal
-particle increases with its number of neighbors. Approaching close packing (12 neighbors), we clearly see two family of curves. The state point
-at $\phi\approx 0.30$ and $c_p=0.38$ mg/g reaches close packing at a volume fraction of around $\sim 60\%$, which is in reasonable agreement with the
-volume fraction of the stable crystal phase. All other state points, instead, reach close packing at volume fraction $\sim 55\%$, which is
-close to the volume fraction of the glass state. 
+particle increases with its number of neighbors. Approaching close packing (12 neighbors), we clearly see two families of curves. The state point
+at $\phi\approx 0.30$ and $c_p=0.38$ mg/g reaches close packing at a volume fraction of around $\sim 68\%$, which is in reasonable agreement with the
+volume fraction of the stable crystal phase~\cite{ilett1995phase}. All other state points, instead, reach close packing at a volume fraction $\sim 64\%$, which is
+indeed close to the volume fraction of the attractive glass state~\cite{pham2002multiple}. 
 %(\textbf{the absolute values of these volume fraction are not totally convincing, probably because of a not optimal estimate of the size of the colloids.
@@
 This result clearly indicates that the state point at $\phi\approx 0.30$ and $c_p=0.38$ mg/g follows a different arrest mechanism, in which phase separation
-is arrested by crystallization and not glassiness. Although our estimation of the absolute volume fraction may involves errors of a few \%, the trend should be robust.
+is arrested by crystallization and not by glassiness. Although our estimation of the absolute volume fraction may involves errors of a few \%, the trend should be robust.
 
@@
 
-To confirm the formation of crystalline seeds we adopt bond orientational analysis to detect which particles are in crystalline environments~\cite{russo2013interplay}.
-With this analysis we confirm that crystallization events start occurring for the state point $\phi\approx 0.30$ and $c_p=0.38$ mg/g, which has the lowest
+To confirm the formation of crystalline seeds, we adopt bond orientational analysis to detect which particles are in crystalline environments~\cite{russo2013interplay}. 
+With this analysis we reveal that crystallization events indeed start occurring for the state point $\phi\approx 0.30$ and $c_p=0.38$ mg/g, which has the lowest
 polymer concentration of all state points. This is shown in Fig.~\ref{fig:crystals}A, where the black line indicates the average size of the crystallites, 
-while the red line the average number of crystallites as a function of time. The number of crystallites first rapidly increases as nucleation events
-start occurring inside the liquid branches, but eventually starts decreasing as the different crystallites grow and merge with each other.
+while the red line indicates the average number of crystallites as a function of time. The number of crystallites first rapidly increases as nucleation events
+start occurring inside the liquid network, but eventually starts decreasing as the different crystallites grow and merge with each other.
 %The average size of the crystals shows an interesting linear growth regime for $100<t<600$. This linear growth regime goes past the time
@@
 which was neither considered nor observed before.
-In the following we will show that the growth mechanism of the crystal is not only due to filling of the spinodal liquid network, as was observed in Refs.~\cite{fortini2008crystallization,perez2011pathways} for Brownian Dynamics simulations, but also involves direct condensation of the gas phase
-on the crystalline seeds.
-To investigate this process, we plot in Fig.~\ref{fig:crystals}B the gyration radius of individual crystalline nuclei for the state point
+In the following we will show that the growth mechanism of the crystal is not only due to filling of the spinodal liquid network, as was observed in Refs.~\cite{fortini2008crystallization,perez2011pathways} for Brownian Dynamics simulations, but also involves
+evaporation of the liquid phase and subsequent direct condensation of the gas phase on the crystalline seeds. 
+To \textcolor{blue}{confirm the presence of the two processes}, we plot in Fig.~\ref{fig:crystals}B the gyration radius of individual crystalline nuclei for the state point 
 $\phi\approx 0.30$ and $c_p=0.38$ mg/g. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with 
@@
 small crystalline nuclei are nucleated inside the phase-separated liquid branches, once they reach the transverse size of the liquid branch, the growth 
-follows a volume expansion with the formation of isotropic droplets (colloidal rain). The same scaling law was suggested in the analysis of the
-structure factors, Fig.~\ref{fig:phase_separation}C, but it is shown directly by the analysis of the gyration radius.
+follows a volume \textcolor{blue}{growth} with the formation of isotropic \textcolor{blue}{crystal} droplets. The same scaling law was suggested in the analysis of the
+structure factors, Fig.~\ref{fig:phase_separation}C, but \textcolor{blue}{here} it is shown directly by the analysis of the gyration radius.
 
-In the next section we demonstrate that the volume growth is due to direct condensation of the gas phase on the crystalline seeds.
+\textcolor{blue}{Now} we demonstrate that the volume growth is due to direct condensation of the gas phase on the crystalline seeds.
 Supplementary Movie 2 shows the confocal microscopy video for the crystallizing sample at $\phi\approx 0.30$ and $c_p=0.38$ mg/g,
 where the process of crystal growth by direct attachment of gas particles can be observed directly. 
-From the trajectory, we extract slabs of 10 particle diameter thickness for configurations at $\phi\approx 0.30$ and $c_p=0.38$ mg/g at different times,
-shown in Fig.~\ref{fig:transitions}A slabs.
-Particles are colored according to their phase. The figure shows the first nucleation
+From the trajectory, we extract slabs of 10 particle diameter thickness for configurations at different times, shown in Fig.~\ref{fig:transitions}A. 
+Particles are colored according to their phases (gas, liquid, crystal, and crystal surface). The figure shows the first nucleation 
 events inside the liquid network (left panel); then these crystalline nuclei grow directly from the gas phase (middle panel), and finally
@@
 by a steep decrease in liquid particles, as they transform in small crystalline nuclei inside the liquid domains. This decrease
-is accompanied by an increase in the fraction of gas particles: as the first crystals start to appear, liquid particles evaporate
+is then accompanied by an increase in the fraction of gas particles: as the first crystals start to appear \textcolor{blue}{in the gas phase}, liquid particles evaporate 
 to the gas phase due to the higher vapor pressure of the liquid phase compared to the crystalline phase. After the onset
@@
 examine in Fig.~\ref{fig:transitions}C the transfer matrix probabilities $P_{\alpha\beta}$, where $\alpha$ and $\beta$ are each of the 
-phases gas, liquid and crystal (which includes bcc, hcp, fcc and surface particles). Liquid particles undergo two processes, on one
+phases, gas, liquid and crystal (which includes hcp, fcc and surface particles). Liquid particles undergo two processes: on one
 hand they transform to crystallites growing inside the liquid branches of the network, and also increasingly transform to gas particles.
@@
 crystalline nuclei outgrow the liquid pockets in which they are formed, the probability for the transition from gas to liquid
-abruptly decreases. Instead, gas particles start depositing on the crystal, and at later stages the pace of the gas-to-crystal transformation
+abruptly decreases. Instead, gas particles start depositing on the crystal, and at later stages the pace of the gas-to-crystal transformation \textcolor{blue}{(sublimation)}  
 is larger than the one of the liquid-to-crystal one.
@@
 We can thus conclude that the crystallization proceeds via two distinct mechanisms:
-(i) freezing of the fluid; (ii) evaporation of the fluid and condensation of the vapor on the crystals.
+(i) freezing of the fluid; (ii) evaporation of the fluid and condensation of the vapor on the crystals, \textcolor{blue}{i.e., a coupled process of evaporation and sublimation.} 
 The first process dominates the initial stages of crystallization, while the second one becomes effective once
@@
 The first stages of gelation always involve spinodal decomposition with the formation
-of a fractal network and gas-liquid phase separation.
+of \textcolor{blue}{liquid network by viscoelastic gas-liquid phase separation.} 
 % Due to hydrodynamic interaction, the network is locally stretched, and the accumulation of mechanical stress on some bonds causes some
@@
 Depending on the polymer concentration, there are two possible arrest mechanisms. 
-(i) Crystallization, observed at $\phi\approx 0.30$ and $c_p=0.38$ mg/g, small crystalline nuclei appear inside the liquid, reach the surface
+(i) Crystallization, observed at $\phi\approx 0.30$ and $c_p=0.38$ mg/g, small crystalline nuclei appear inside the liquid network, reach the surface 
  of the liquid branches, and then grow by addition of particles from the gas phase. The final structure is a network of crystal droplets,
@@
 with respect to the ice crystals. This causes the water droplets to evaporate and sublimate directly on the ice crystals. Our system can then
-be regarded as a colloidal analogue for this important process, which can now be studied at the single-particle level.
+be regarded as a colloidal analogue for this important process, which, for the first time, we can observe at the single-particle level.
 
-The basic mechanism creating ``crystal gels'' may be generic to many other
+\textcolor{blue}{This novel} mechanism creating ``crystal gels'' may be generic to many other 
 systems. The requirements are (i) the presence of gas-liquid phase separation below the melting point of a crystal, (ii) 
@@
 (iii) dynamical slowing down in a supercooled liquid state, which is necessary to induce viscoelastic phase separation leading to the formation 
-of a network structure of the minority phase, and (iv) the degree of supercooling is low enough to avoid a high rate of densification of the liquid phase 
+of a network structure of the minority \textcolor{blue}{liquid} phase, and (iv) the degree of supercooling is low enough to avoid a high rate of densification of the liquid phase 
 leading to vitrification.   
@@
 and industrial applications, although they have not attracted much attention so far.  
+\textcolor{blue}{We speculate that our scenario is valid also for crystal network formation in dynamically asymmetric mixtures, 
+which includes magma \cite{philpotts1998role}, biominerals \cite{rousseau2005multiscale}, and foods \cite{deman1987fat}  
+({\bf Hajime: This need to be checked carefully)}. }
 

### 2015-10-17 crystal-gel/science/coverl_letter.tex
Small adjustements.

@@
 water crystallization in clouds is recognized as one of the main challenges in the
-modeling of Earth's radiation budget and climate. The difficulties arise both from our poor understanding of small-scale microphysical effects,
+modeling of Earth's radiation budget and climate. \textcolor{red}{Moreover, our study addresses the formation of crystal networks,
+which are observed in many systems, including magma, biominerals and foods, where the origin
+and formation of the networks still lacks fundamental understanding.}
+The difficulties arise both from our poor understanding of small-scale microphysical effects,
 and to the difficulty of experimental investigations. In our colloidal system, both

### 2015-10-17 crystal-gel/science/crystal_gel_letter2.tex
Small adjustements.

@@
 \begin{sciabstract}
-%Colloidal gels form when network-forming spinodal phase separation is interrupted by the glass transition~\cite{lu2008gelation}.
-It has been shown that crystallization may equivalently interrupt the phase separation by freezing the dense phase.
-However the particle-level details of this process have not yet been confirmed experimentally.
-\textcolor{blue}{
-With confocal microscopy measurements we find a novel crystallization pathway of a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation.} We show that crystallization leads to a network structure different from the original phase separation pattern.
-Crystals nucleate inside the liquid network, \textcolor{blue}{but} grow past it by direct condensation of the gas phase on their surface, driving evaporation of nearby liquid.
-This process represents the colloidal analogue of the Bergeron process, i.e., ice crystals formation in mixed phase clouds and is at the origin of rain.
-\textcolor{blue}{Our finding} gives us full experimental access to the kinetic pathway of this phenomenon important for climate science.
-We argue that similar states can form as the result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems, if dynamics of the liquid phase is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
+
+With confocal microscopy measurements we find a novel crystallization pathway of a colloid-polymer mixture undergoing viscoelastic
+gas-liquid phase separation. We show that crystallization leads to a network structure different from the original phase separation pattern.
+Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface,
+driving evaporation of nearby liquid. This process represents the colloidal analogue of the Bergeron process, i.e.,
+ice crystals formation in mixed phase clouds and is at the origin of rain. Our finding gives us full experimental
+access to the kinetic pathway of this phenomenon important for climate science. We argue that similar states can form as the
+result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems,
+if dynamics of the liquid phase is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate 
+vitrification.
 \end{sciabstract}
@@
 that are at the core of the chemical, pharmaceutical, and food industries.
-Many of the properties of crystals, like the \textcolor{blue}{shape, spatial arrangement}, polymorph type, and size distribution of the crystallites, depend on the conditions at which
+Many of the properties of crystals, like the shape, spatial arrangement, polymorph type, and size distribution of the crystallites, depend on the conditions at which
 the nucleation process took place. Controlling the early stages of crystallization is thus of fundamental importance in order to
@@
 directly from the supersaturated solution. But the possibility of different pathways has recently come to prominence with
-the discovery of two-step nucleation pathways, where crystallization is proceeded by the formation of dense liquid  droplets \textcolor{blue}{as an intermediate step}~\cite{ten1997enhancement,SearR,savage2009experimental,vekilov2010two,palberg2014crystallization}.
+the discovery of two-step nucleation pathways, where crystallization is proceeded by the formation of dense liquid  droplets as an intermediate step~\cite{ten1997enhancement,SearR,savage2009experimental,vekilov2010two,palberg2014crystallization}.
 Understanding the process of crystal formation in mixed-phase systems (composed of gas, liquid, and solid phases) is thus
@@
 observation of the initial stages of crystallization. This initial perturbation may even alter the selection of the final non-equilibrium arrested state.  
-In our experiments we instead succeed in building a novel \textcolor{blue}{protocol} that allows the time evolution of the system to be observed directly
+In our experiments we instead succeed in building a novel protocol that allows the time evolution of the system to be observed directly
 under the confocal microscope from the very early stages and in a quiescent situation without introducing fluid flows.
@@
 of crystalline droplets.
-These droplets originate from crystal seeds that form inside the liquid branches of the spinodal aggregate \textcolor{blue}{of network morphology}, but then grow by direct condensation from the gas phase.
+These droplets originate from crystal seeds that form inside the liquid branches of the spinodal aggregate of network morphology, but then grow by direct condensation from the gas phase.
 The process is analogous to the Bergeron process of ice formation in mixed phase clouds, where ice droplets grow at the
@@
 This is speculated to be the primary mechanism for the formation of raindrops. 
-\textcolor{blue}{The mechanism found here} is in contrast to the standard crystal-gel scenario, where crystals nucleate and grow inside the dense branches of the gel, becoming dynamically arrested.
+The mechanism found here is in contrast to the standard crystal-gel scenario, where crystals nucleate and grow inside the dense branches of the gel, becoming dynamically arrested.
 We also observe that increasing the strength of the interparticle attraction
-(by increasing the polymer concentration), instead of forming \textcolor{blue}{crystal gels}, the system undergoes the classical gelation process, in which
+(by increasing the polymer concentration), instead of forming crystal gels, the system undergoes the classical gelation process, in which
 phase separation is interrupted by vitrification~\cite{verhaegh1997transient,tanaka1999colloid,Royall2007,lu2008gelation}.
@@
 In Fig.~\ref{fig:phase_separation}A, we superimpose the state points with the phase diagram, which we calculated from generalised free volume theory~\cite{Fleer2008} using the experimental values for the size of the colloids, the radius of gyration of the polymer, and the polymer concentration (see Supplementary Information).
-For each state point, scans \textcolor{blue}{were made} at early times (every $10$~s, i.e. $4\tau_B$) and at the late stage (every $30$~s) of the gelation process.
+For each state point, scans were made at early times (every $10$~s, i.e. $4\tau_B$) and at the late stage (every $30$~s) of the gelation process.
 
@@
 
-While sharing similar structural properties at early times, the samples \textcolor{blue}{show significantly different behaviors in} the later stages of the phase separation process.
+While sharing similar structural properties at early times, the samples show significantly different behaviors in the later stages of the phase separation process.
 To examine this in more detail, we plot in Fig.~\ref{fig:phase_separation}C the structure factor for the final stages of the gelation process for all state points.
@@
 evaporation of the liquid phase and subsequent direct condensation of the gas phase on the crystalline seeds. 
-To \textcolor{blue}{confirm the presence of the two processes}, we plot in Fig.~\ref{fig:crystals}B the gyration radius of individual crystalline nuclei for the state point 
+To confirm the presence of the two processes, we plot in Fig.~\ref{fig:crystals}B the gyration radius of individual crystalline nuclei for the state point 
 $\phi\approx 0.30$ and $c_p=0.38$ mg/g. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with 
@@
 small crystalline nuclei are nucleated inside the phase-separated liquid branches, once they reach the transverse size of the liquid branch, the growth 
-follows a volume \textcolor{blue}{growth} with the formation of isotropic \textcolor{blue}{crystal} droplets. The same scaling law was suggested in the analysis of the
-structure factors, Fig.~\ref{fig:phase_separation}C, but \textcolor{blue}{here} it is shown directly by the analysis of the gyration radius.
+follows a volume growth with the formation of isotropic crystal droplets. The same scaling law was suggested in the analysis of the
+structure factors, Fig.~\ref{fig:phase_separation}C, but here it is shown directly by the analysis of the gyration radius.
 
-\textcolor{blue}{Now} we demonstrate that the volume growth is due to direct condensation of the gas phase on the crystalline seeds.
+Now we demonstrate that the volume growth is due to direct condensation of the gas phase on the crystalline seeds.
 Supplementary Movie 2 shows the confocal microscopy video for the crystallizing sample at $\phi\approx 0.30$ and $c_p=0.38$ mg/g,
@@
 by a steep decrease in liquid particles, as they transform in small crystalline nuclei inside the liquid domains. This decrease
-is then accompanied by an increase in the fraction of gas particles: as the first crystals start to appear \textcolor{blue}{in the gas phase}, liquid particles evaporate 
+is then accompanied by an increase in the fraction of gas particles: as the first crystals start to appear in the gas phase, liquid particles evaporate 
 to the gas phase due to the higher vapor pressure of the liquid phase compared to the crystalline phase. After the onset
@@
 crystalline nuclei outgrow the liquid pockets in which they are formed, the probability for the transition from gas to liquid
-abruptly decreases. Instead, gas particles start depositing on the crystal, and at later stages the pace of the gas-to-crystal transformation \textcolor{blue}{(sublimation)}  
+abruptly decreases. Instead, gas particles start depositing on the crystal, and at later stages the pace of the gas-to-crystal transformation (sublimation)
 is larger than the one of the liquid-to-crystal one.
@@
 We can thus conclude that the crystallization proceeds via two distinct mechanisms:
-(i) freezing of the fluid; (ii) evaporation of the fluid and condensation of the vapor on the crystals, \textcolor{blue}{i.e., a coupled process of evaporation and sublimation.} 
+(i) freezing of the fluid; (ii) evaporation of the fluid and condensation of the vapor on the crystals, i.e., a coupled process of evaporation and sublimation.
 The first process dominates the initial stages of crystallization, while the second one becomes effective once
@@
 The first stages of gelation always involve spinodal decomposition with the formation
-of \textcolor{blue}{liquid network by viscoelastic gas-liquid phase separation.} 
+of liquid network by viscoelastic gas-liquid phase separation. 
 % Due to hydrodynamic interaction, the network is locally stretched, and the accumulation of mechanical stress on some bonds causes some
@@
 
-\textcolor{blue}{This novel} mechanism creating ``crystal gels'' may be generic to many other 
+This novel mechanism creating ``crystal gels'' may be generic to many other 
 systems. The requirements are (i) the presence of gas-liquid phase separation below the melting point of a crystal, (ii) 
@@
 (iii) dynamical slowing down in a supercooled liquid state, which is necessary to induce viscoelastic phase separation leading to the formation 
-of a network structure of the minority \textcolor{blue}{liquid} phase, and (iv) the degree of supercooling is low enough to avoid a high rate of densification of the liquid phase 
+of a network structure of the minority liquid phase, and (iv) the degree of supercooling is low enough to avoid a high rate of densification of the liquid phase 
 leading to vitrification.   
@@
 and industrial applications, although they have not attracted much attention so far.  
-\textcolor{blue}{We speculate that our scenario is valid also for crystal network formation in dynamically asymmetric mixtures, 
-which includes magma \cite{philpotts1998role}, biominerals \cite{rousseau2005multiscale}, and foods \cite{deman1987fat}  
-({\bf Hajime: This need to be checked carefully)}. }
+\textcolor{red}{We speculate that our scenario could play a role in the formation of crystal networks observed in dynamically asymmetric mixtures, 
+which includes magma \cite{philpotts1998role}, biominerals \cite{rousseau2005multiscale}, and foods \cite{deman1987fat}.}
 

### 2015-10-17 science/crystal_gel_letter2.tex
Other small adjustements.

@@
 \maketitle
-\vspace{-1cm}
+\vspace{-1.5cm}
 \begin{sciabstract}
-%Colloidal gels form when network-forming spinodal phase separation is interrupted by the glass transition~\cite{lu2008gelation}.
-It has been shown that crystallization may equivalently interrupt the phase separation by freezing the dense phase.
-However the particle-level details of this process have not yet been confirmed experimentally.
-\textcolor{blue}{
-With confocal microscopy measurements we access the microscopic mechanism of crystal growth in a three phase system of a colloid-polymer mixture
-undergoing viscoelastic phase separation.} We show that crystallization leads to a network structure different from the original phase separation pattern.
-Crystals nucleate inside the liquid network, and grow past it by direct condensation of the gas phase on their surface, driving evaporation of nearby liquid.
-This process represents the colloidal analogue of the Bergeron process, i.e., ice crystals formation that occurs in mixed phase clouds and is at the origin of rain.
-Our \emph{colloidal rain} gives us full experimental access to the kinetic pathway of this phenomenon important for climate science.
-We argue that states similar to colloidal rain can form as the result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems, whose dynamics is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
+
+With confocal microscopy measurements we find a novel crystallization pathway of a colloid-polymer mixture undergoing viscoelastic
+gas-liquid phase separation. We show that crystallization leads to a network structure different from the original phase separation pattern.
+Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface,
+driving evaporation of nearby liquid. This process represents the colloidal analogue of the Bergeron process, i.e.,
+ice crystals formation in mixed phase clouds and is at the origin of rain. Our finding gives us full experimental
+access to the kinetic pathway of this phenomenon important for climate science. We argue that similar states can form as the
+result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems,
+if dynamics of the liquid phase is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate 
+vitrification.
 \end{sciabstract}
@@
 
-\textcolor{blue}{Crystallization plays a fundamental role in many processes occurring in nature, such as cloud composition, and in technologial applications
-that are at the core of the chemical, pharmaceutical and food industries.
-Many of the properties of crystals, like the polymorph type and size distribution of the crystallites, depend on the conditions at which
+Crystallization plays a fundamental role in many processes occurring in nature, such as cloud composition, and in technological applications
+that are at the core of the chemical, pharmaceutical, and food industries.
+Many of the properties of crystals, like the shape, spatial arrangement, polymorph type, and size distribution of the crystallites, depend on the conditions at which
 the nucleation process took place. Controlling the early stages of crystallization is thus of fundamental importance in order to
-obtain in a reproducible manner crystals with the desired properties.}
+obtain in a reproducible manner crystals with the desired properties.
 
-\textcolor{blue}{Under usual conditions, the nucleation stage involves a very small number of molecules, of the order of 10-1000 molecules.
-This has severily limited the possibility to observe the nucleation process directly, and has also posed limits on
+Under usual conditions, the nucleation stage involves a very small number of molecules, of the order of 10-1000 molecules.
+This has severely limited the possibility to observe the nucleation process directly, and has also posed limits on
 models based on macroscopic thermodynamic properties. 
 The inability to directly observe the nucleation process is also a limit to our understanding of its kinetic pathway. 
-Classical Nucleation Theory assumes the crystallization processes to occurr in one step,
+Classical Nucleation Theory assumes the crystallization processes to occur in one step,
 where the transition can be described by just one reaction coordinate. An example is the formation of an ordered crystalline nucleus
-directly from the supersatured solution. But the possibility of different pathways has recently come to prominence with
-the discovery of two-step nucleation pathways, where crystallization is preceeded by the formation of dense liquid droplets~\cite{ten1997enhancement,SearR,savage2009experimental,vekilov2010two,palberg2014crystallization}.
+directly from the supersaturated solution. But the possibility of different pathways has recently come to prominence with
+the discovery of two-step nucleation pathways, where crystallization is proceeded by the formation of dense liquid  droplets as an intermediate step~\cite{ten1997enhancement,SearR,savage2009experimental,vekilov2010two,palberg2014crystallization}.
 Understanding the process of crystal formation in mixed-phase systems (composed of gas, liquid, and solid phases) is thus
-of great importance for a variety of systems, from protein solutions to clouds.}
+of great importance for a variety of systems, from protein solutions to clouds.
 
-\textcolor{blue}{Colloidal solutions offer a system where the crystallization process in a mixed-phase environment can be observed with single particle
-resolution, and at the same time scales over which nucleation takes place. In Colloids with short-range attractions the liquid-gas transition
+Colloidal suspensions offer a system where the crystallization process in a mixed-phase environment can be observed with single-particle
+resolution, and at the same time scales over which nucleation takes place. In colloids with short-range attractions the liquid-gas transition 
 becomes metastable with respect to crystallization~\cite{anderson2002insights,lekkerkerker2011colloids}, and can form gels
-by viscoelastic phase separation \cite{tanaka1999colloid,tanaka2000viscoelastic}, with the coexistence of a dense liquid phase with slow dynamics and a dilute gas phase with fast dynamics.}
+by viscoelastic phase separation \cite{tanaka1999colloid,tanaka2000viscoelastic}, with the coexistence of a dense liquid phase with slow dynamics and a dilute gas phase with fast dynamics. 
 This difference in the viscoelastic properties between the two phases allows the system to form a space-spanning 
@@
 plateau~\cite{krall1998internal,solomon2001dynamic,romer2000sol}. 
-During phase separation, the density of the liquid phase 
-increases towards the glass-transition density, leading to slow glassy dynamics.
+During phase separation, the density of the liquid phase increases towards the glass-transition density, leading to slow glassy dynamics.
 At the glass transition point, the percolated network structure is dynamically stabilised by vitrification 
@@
 observation of the initial stages of crystallization. This initial perturbation may even alter the selection of the final non-equilibrium arrested state.  
-In our experiments we instead succeed in building a novel setup that allows the time evolution of the system to be observed directly
+In our experiments we instead succeed in building a novel protocol that allows the time evolution of the system to be observed directly
 under the confocal microscope from the very early stages and in a quiescent situation without introducing fluid flows.
@@
 of crystalline droplets.
-These droplets originate from crystal seeds that form inside the liquid branches of the spinodal aggregate, but then grow by direct condensation from the gas phase.
+These droplets originate from crystal seeds that form inside the liquid branches of the spinodal aggregate of network morphology, but then grow by direct condensation from the gas phase.
 The process is analogous to the Bergeron process of ice formation in mixed phase clouds, where ice droplets grow at the
@@
 This is speculated to be the primary mechanism for the formation of raindrops. 
-In our colloidal system, we then call this process ``colloidal rain''
-formation.
-\textcolor{blue}{\sout{Due to the difference in the nucleation and growth processes, the droplets display different scaling laws depending on their
-size. For small nuclei, the droplets have the same fractal dimension of the embedding gel network. But during the growth process, droplets
-grow with an integer fractal dimension due to direct condensation of gas particles.}} This is in contrast to the standard crystal-gel scenario, where crystals nucleate and grow inside the dense branches of the gel, becoming dynamically arrested.
+The mechanism found here is in contrast to the standard crystal-gel scenario, where crystals nucleate and grow inside the dense branches of the gel, becoming dynamically arrested.
 We also observe that increasing the strength of the interparticle attraction
-(by increasing the polymer concentration), instead of forming colloidal rain, the system undergoes the classical gelation process, in which
+(by increasing the polymer concentration), instead of forming crystal gels, the system undergoes the classical gelation process, in which
 phase separation is interrupted by vitrification~\cite{verhaegh1997transient,tanaka1999colloid,Royall2007,lu2008gelation}.
@@
 %In the following we discuss all the stages of the formation of the crystal gels: phase separation, stress-driven agein and crystallization.
-The experimental data is taken at two different volume fractions ($\phi\approx 0.13$ and $\phi\approx 0.30$) and for different values
+The experimental data are taken at two different volume fractions ($\phi\approx 0.13$ and $\phi\approx 0.30$) and for different values
 of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ mg/g for $\phi\approx 0.30$ and $c_p=0.82,1.36$ mg/g for $\phi\approx 0.13$.
-In Fig.~\ref{fig:phase_separation}A, we superimpose the state points with the phase diagram \textcolor{blue}{which we calculated from generalised free volume theory~\cite{Fleer2008} using the experimental values for the size of the colloids, the radius of gyration of the polymer and the polymer concentration (see Supplementary Information)}.
-For each state points, scans at early times (every $10$~s, i.e. $4\tau_B$) and at the late stage (every $30$~s) of the gelation process.
+In Fig.~\ref{fig:phase_separation}A, we superimpose the state points with the phase diagram, which we calculated from generalised free volume theory~\cite{Fleer2008} using the experimental values for the size of the colloids, the radius of gyration of the polymer, and the polymer concentration (see Supplementary Information).
+For each state point, scans were made at early times (every $10$~s, i.e. $4\tau_B$) and at the late stage (every $30$~s) of the gelation process.
 
@@
 eventually forming a percolating network. Thus, a dense network (liquid) coexists with freely diffusing monomers (gas).
-\textcolor{blue}{In Fig.~\ref{fig:phase_separation}A, the state points which didn't phase separate are indicated with open triangles, while
-the other symbols indicate samples undergoing phase separation. The (black) continous line is the gas spinodal computed from free volume theory,
+In Fig.~\ref{fig:phase_separation}A, the state points which did not phase separate are indicated with open triangles, while
+the other symbols indicate samples undergoing phase separation. The (black) continuous line is the gas spinodal computed from free volume theory,
 and it is in good agreement with our experimental points (the small discrepancy at small colloidal volume fractions 
-is common in free volume theory)~\cite{Royall2007,lu2008gelation}}.
+is common in free volume theory)~\cite{Royall2007,lu2008gelation}. 
 The topology of the network is similar to all our samples at the initial stages. 
@@
 within the first peak of the radial distribution function. All samples show the same scaling law, $R^2_{\rm g}\sim s^{2/D}$, which is compatible with
-the random gelation universality class exponent ($D=2.53$) as shown by the dashed line in Fig.~\ref{fig:phase_separation}B.
+the random gelation universality class exponent ($D=2.53$) as shown by the dashed line in Fig.~\ref{fig:phase_separation}B
+(\textcolor{red}{we don't imply that the formation of the network follows the random gelation universality class, just that our
+fractal dimension is compatible with it}).
 
-While sharing similar structural properties at early times, the samples differ during the later stages of the phase separation process.
+While sharing similar structural properties at early times, the samples show significantly different behaviors in the later stages of the phase separation process.
 To examine this in more detail, we plot in Fig.~\ref{fig:phase_separation}C the structure factor for the final stages of the gelation process for all state points.
@@
 through Voronoi diagrams, as a function of the number of neighbors. For all state points, we can see that the average volume fraction of a colloidal
-particle increases with its number of neighbors. Approaching close packing (12 neighbors), we clearly see two family of curves. The state point
-at $\phi\approx 0.30$ and $c_p=0.38$ mg/g reaches \textcolor{blue}{close packing at a volume fraction of around $\sim 68\%$, which is in reasonable agreement with the
-volume fraction of the stable crystal phase~\cite{ilett1995phase}. All other state points, instead, reach close packing at volume fraction $\sim 64\%$, which is
-indeed close to the volume fraction of the attractive glass state~\cite{pham2002multiple}}. 
+particle increases with its number of neighbors. Approaching close packing (12 neighbors), we clearly see two families of curves. The state point
+at $\phi\approx 0.30$ and $c_p=0.38$ mg/g reaches close packing at a volume fraction of around $\sim 68\%$, which is in reasonable agreement with the
+volume fraction of the stable crystal phase~\cite{ilett1995phase}. All other state points, instead, reach close packing at a volume fraction $\sim 64\%$, which is
+indeed close to the volume fraction of the attractive glass state~\cite{pham2002multiple}. 
 %(\textbf{the absolute values of these volume fraction are not totally convincing, probably because of a not optimal estimate of the size of the colloids.
@@
 In the following we will show that the growth mechanism of the crystal is not only due to filling of the spinodal liquid network, as was observed in Refs.~\cite{fortini2008crystallization,perez2011pathways} for Brownian Dynamics simulations, but also involves
-\textcolor{blue}{evaporation of the liquid phase and subsequent direct condensation of the gas phase
-on the crystalline seeds.}
-To investigate this process, we plot in Fig.~\ref{fig:crystals}B the gyration radius of individual crystalline nuclei for the state point
+evaporation of the liquid phase and subsequent direct condensation of the gas phase on the crystalline seeds. 
+To confirm the presence of the two processes, we plot in Fig.~\ref{fig:crystals}B the gyration radius of individual crystalline nuclei for the state point 
 $\phi\approx 0.30$ and $c_p=0.38$ mg/g. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with 
-the dimension of random percolation $D=2.53$, while at large sizes it scales as $D=3$ of compact crystal growth. This demonstrates that, while
+\textcolor{red}{a fractal dimension close to random percolation ($D=2.53$)}, while at large sizes it scales as $D=3$, \textcolor{red}{as in} compact crystal growth. This demonstrates that, while 
 small crystalline nuclei are nucleated inside the phase-separated liquid branches, once they reach the transverse size of the liquid branch, the growth 
-follows a volume expansion with the formation of isotropic droplets (colloidal rain). The same scaling law was suggested in the analysis of the
-structure factors, Fig.~\ref{fig:phase_separation}C, but it is shown directly by the analysis of the gyration radius.
+follows a volume growth with the formation of isotropic crystal droplets. The same scaling law was suggested in the analysis of the
+structure factors, Fig.~\ref{fig:phase_separation}C, but here it is shown directly by the analysis of the gyration radius.
 
-In the next section we demonstrate that the volume growth is due to direct condensation of the gas phase on the crystalline seeds.
+Now we demonstrate that the volume growth is due to direct condensation of the gas phase on the crystalline seeds.
 Supplementary Movie 2 shows the confocal microscopy video for the crystallizing sample at $\phi\approx 0.30$ and $c_p=0.38$ mg/g,
 where the process of crystal growth by direct attachment of gas particles can be observed directly. 
-From the trajectory, we extract slabs of 10 particle diameter thickness for configurations at $\phi\approx 0.30$ and $c_p=0.38$ mg/g at different times,
-shown in Fig.~\ref{fig:transitions}A slabs.
-Particles are colored according to their phase. The figure shows the first nucleation
+From the trajectory, we extract slabs of 10 particle diameter thickness for configurations at different times, shown in Fig.~\ref{fig:transitions}A. 
+Particles are colored according to their phases (gas, liquid, crystal, and crystal surface). The figure shows the first nucleation 
 events inside the liquid network (left panel); then these crystalline nuclei grow directly from the gas phase (middle panel), and finally
@@
 by a steep decrease in liquid particles, as they transform in small crystalline nuclei inside the liquid domains. This decrease
-is accompanied by an increase in the fraction of gas particles: as the first crystals start to appear, liquid particles evaporate
+is then accompanied by an increase in the fraction of gas particles: as the first crystals start to appear in the gas phase, liquid particles evaporate 
 to the gas phase due to the higher vapor pressure of the liquid phase compared to the crystalline phase. After the onset
@@
 examine in Fig.~\ref{fig:transitions}C the transfer matrix probabilities $P_{\alpha\beta}$, where $\alpha$ and $\beta$ are each of the 
-phases gas, liquid and crystal (which includes hcp, fcc and surface particles). Liquid particles undergo two processes, on one
+phases, gas, liquid and crystal (which includes hcp, fcc and surface particles). Liquid particles undergo two processes: on one
 hand they transform to crystallites growing inside the liquid branches of the network, and also increasingly transform to gas particles.
@@
 crystalline nuclei outgrow the liquid pockets in which they are formed, the probability for the transition from gas to liquid
-abruptly decreases. Instead, gas particles start depositing on the crystal, and at later stages the pace of the gas-to-crystal transformation
+abruptly decreases. Instead, gas particles start depositing on the crystal, and at later stages the pace of the gas-to-crystal transformation (sublimation)
 is larger than the one of the liquid-to-crystal one.
@@
 We can thus conclude that the crystallization proceeds via two distinct mechanisms:
-(i) freezing of the fluid; (ii) evaporation of the fluid and condensation of the vapor on the crystals.
+(i) freezing of the fluid; (ii) evaporation of the fluid and condensation of the vapor on the crystals, i.e., a coupled process of evaporation and sublimation.
 The first process dominates the initial stages of crystallization, while the second one becomes effective once
@@
 The first stages of gelation always involve spinodal decomposition with the formation
-of a fractal network and gas-liquid phase separation.
+of liquid network by viscoelastic gas-liquid phase separation. 
 % Due to hydrodynamic interaction, the network is locally stretched, and the accumulation of mechanical stress on some bonds causes some
@@
 with respect to the ice crystals. This causes the water droplets to evaporate and sublimate directly on the ice crystals. Our system can then
-be regarded as a colloidal analogue for this important process, \textcolor{blue}{which, for the first time, we can observe at the single-particle level.}
+be regarded as a colloidal analogue for this important process, which, for the first time, we can observe at the single-particle level.
 
-The basic mechanism creating ``crystal gels'' may be generic to many other
+This novel mechanism creating ``crystal gels'' may be generic to many other 
 systems. The requirements are (i) the presence of gas-liquid phase separation below the melting point of a crystal, (ii) 
@@
 (iii) dynamical slowing down in a supercooled liquid state, which is necessary to induce viscoelastic phase separation leading to the formation 
-of a network structure of the minority phase, and (iv) the degree of supercooling is low enough to avoid a high rate of densification of the liquid phase 
+of a network structure of the minority liquid phase, and (iv) the degree of supercooling is low enough to avoid a high rate of densification of the liquid phase 
 leading to vitrification.   
@@
 and industrial applications, although they have not attracted much attention so far.  
+\textcolor{red}{We speculate that our scenario could play a role in the formation of crystal networks observed in dynamically asymmetric mixtures, 
+which includes magma \cite{philpotts1998role}, biominerals \cite{rousseau2005multiscale}, and foods \cite{deman1987fat}.}
 

### 2015-10-18 crystal-gel/science/crystal_gel_letter2.tex
Poon2012 systematic errors define tau_B reach gas phase

@@
 on the details of our experiments).
-At $t=0$, the increase in the concentration of salt ions progressively screens the Coulombic repulsion between colloidal particles,
+At $t=0$, the increase in the concentration of salt ions \textcolor{blue}{within a few Brownian times ($\tau_B$)} screens the Coulombic repulsion between colloidal particles,
 which are then subject to attractive depletion forces, making the system thermodynamically unstable and leading to phase separation. 
@@
 within the first peak of the radial distribution function. All samples show the same scaling law, $R^2_{\rm g}\sim s^{2/D}$, which is compatible with
-the random gelation universality class exponent ($D=2.53$) as shown by the dashed line in Fig.~\ref{fig:phase_separation}B.
+the random gelation universality class exponent ($D=2.53$) as shown by the dashed line in Fig.~\ref{fig:phase_separation}B
+(\textcolor{red}{we don't imply that the formation of the network follows the random gelation universality class, just that our
+fractal dimension is compatible with it}).
 
@@
 This result clearly indicates that the state point at $\phi\approx 0.30$ and $c_p=0.38$ mg/g follows a different arrest mechanism, in which phase separation
-is arrested by crystallization and not by glassiness. Although our estimation of the absolute volume fraction may involves errors of a few \%, the trend should be robust.
+is arrested by crystallization and not by glassiness.
+Although our estimation of the absolute volume fraction may involves \textcolor{blue}{systematic errors of a few \%, the trend should be robust~\cite{Poon2012}}.
 
@@
 $\phi\approx 0.30$ and $c_p=0.38$ mg/g. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with 
-the dimension of random percolation $D=2.53$, while at large sizes it scales as $D=3$ of compact crystal growth. This demonstrates that, while 
+\textcolor{red}{a fractal dimension close to random percolation ($D=2.53$)}, while at large sizes it scales as $D=3$, \textcolor{red}{as in} compact crystal growth. This demonstrates that, while 
 small crystalline nuclei are nucleated inside the phase-separated liquid branches, once they reach the transverse size of the liquid branch, the growth 
@@
 by a steep decrease in liquid particles, as they transform in small crystalline nuclei inside the liquid domains. This decrease
-is then accompanied by an increase in the fraction of gas particles: as the first crystals start to appear in the gas phase, liquid particles evaporate 
+is then accompanied by an increase in the fraction of gas particles: as the first crystals start to \textcolor{blue}{reach} the gas phase, liquid particles evaporate 
 to the gas phase due to the higher vapor pressure of the liquid phase compared to the crystalline phase. After the onset

### 2015-10-18 science/SOM2.tex
correct tauB and formulation of supplementary movies description

@@
 \noindent
-{\bf Experimental.}
+{\bf Experiments}
+
 We use \textsc{pmma} (poly(methyl methacrylate)) colloids sterically stabilized with methacryloxypropyl terminated \textsc{pdms}(poly(dimethyl siloxane)) and fluorescently labelled with rhodamine isothiocyanate chemically bonded to the \textsc{pmma}.
@@
 
+\noindent
+{\bf Supplementary Movies}
+
+\emph{Supplementary Movie 1}.
+This movie shows the early stage of gelation process of $\phi\approx 0.30$ and $c_p=0.38$ mg/g, from the original confocal images at the same z-position as Fig.~3A.
+Total duration is 1160 s (~530~$\tau$).
+We quickly started scanning the sample after exchanging the salt-reservoir with the solvent including salt.
+At the beginning, salt had not reached the scanning volume and colloids were dispersed by electrostatic repulsion.
+At 0:02 in the movie, salt diffused into the scanning volume and screened the surface charges of colloids.
+Then, colloids immediately initiated phase separation by short-ranged attractions.
+
+\emph{Supplementary Movie 2}.
+This movie shows the late stage after Supplemental Movie 1 with scanning duration of 27000~s (~12300$\tau$).
+
 %\bibliographystyle{Science}

### 2015-10-19 science/crystal_gel_letter3.tex

--- 2015-10-17 science/crystal_gel_letter2.tex
@@
 the random gelation universality class exponent ($D=2.53$) as shown by the dashed line in Fig.~\ref{fig:phase_separation}B
-(\textcolor{red}{we don't imply that the formation of the network follows the random gelation universality class, just that our
-fractal dimension is compatible with it}).
+(we do not imply that the formation of the network follows the random gelation universality class, just that our
+fractal dimension is compatible with it).
 
@@
 $\phi\approx 0.30$ and $c_p=0.38$ mg/g. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with 
-\textcolor{red}{a fractal dimension close to random percolation ($D=2.53$)}, while at large sizes it scales as $D=3$, \textcolor{red}{as in} compact crystal growth. This demonstrates that, while 
+a fractal dimension close to random percolation ($D=2.53$), while at large sizes it scales as $D=3$, as in compact crystal growth. This demonstrates that, while 
 small crystalline nuclei are nucleated inside the phase-separated liquid branches, once they reach the transverse size of the liquid branch, the growth 
@@
 and industrial applications, although they have not attracted much attention so far.  
-\textcolor{red}{We speculate that our scenario could play a role in the formation of crystal networks observed in dynamically asymmetric mixtures, 
-which includes magma \cite{philpotts1998role}, biominerals \cite{rousseau2005multiscale}, and foods \cite{deman1987fat}.}
+We speculate that our scenario could play a role in the formation of crystal networks observed in dynamically asymmetric mixtures, 
+which includes magma \cite{philpotts1998role}, biominerals \cite{rousseau2005multiscale}, and foods \cite{deman1987fat}.
 
@@
 H.Ta. conceived and supervised the project, H.Ts. performed experiments, J.R. analyzed the data, M.L. calculated the phase diagram, and all the authors discussed and wrote the manuscript. 
-This study was partly supported by Grants-in-Aid for Scientific Research (S) and Specially Promoted Research from JSPS.
+This study was partly supported by Grants-in-Aid for Scientific Research (S) (21224011) and Specially Promoted Research (25000002) from the Japan Society of the Promotion of Science (JSPS).
 

### 2015-10-19 science/crystal_gel_letter3.tex
Merged Mathieu's and sensei versions.
The working copy is crystal_gel_letter3.tex

@@
 on the details of our experiments).
-At $t=0$, the increase in the concentration of salt ions progressively screens the Coulombic repulsion between colloidal particles,
+At $t=0$, the increase in the concentration of salt ions within a few Brownian times ($\tau_B$) screens the Coulombic repulsion between colloidal particles,
 which are then subject to attractive depletion forces, making the system thermodynamically unstable and leading to phase separation. 
@@
 This result clearly indicates that the state point at $\phi\approx 0.30$ and $c_p=0.38$ mg/g follows a different arrest mechanism, in which phase separation
-is arrested by crystallization and not by glassiness. Although our estimation of the absolute volume fraction may involves errors of a few \%, the trend should be robust.
+is arrested by crystallization and not by glassiness. Although our estimation of the absolute volume fraction may involves systematic errors of a few \%, the trend should be robust~\cite{Poon2012}.
 
@@
 by a steep decrease in liquid particles, as they transform in small crystalline nuclei inside the liquid domains. This decrease
-is then accompanied by an increase in the fraction of gas particles: as the first crystals start to appear in the gas phase, liquid particles evaporate 
+is then accompanied by an increase in the fraction of gas particles: as the first crystals start to reach the gas phase, liquid particles evaporate 
 to the gas phase due to the higher vapor pressure of the liquid phase compared to the crystalline phase. After the onset

### 2015-10-20 crystal-gel/science/coverl_letter.tex

@@
 water crystallization in clouds is recognized as one of the main challenges in the
-modeling of Earth's radiation budget and climate. \textcolor{red}{Moreover, our study addresses the formation of crystal networks,
+modeling of Earth's radiation budget and climate. Moreover, our study addresses the formation of crystal networks,
 which are observed in many systems, including magma, biominerals and foods, where the origin
-and formation of the networks still lacks fundamental understanding.}
+and formation of the networks still lacks fundamental understanding. 
 The difficulties arise both from our poor understanding of small-scale microphysical effects,

### 2015-10-20 crystal-gel/science/crystal_gel_letter3.tex

--- 2015-10-19 science/crystal_gel_letter3.tex
@@
 \begin{sciabstract}
-
-With confocal microscopy measurements we find a novel crystallization pathway of a colloid-polymer mixture undergoing viscoelastic
-gas-liquid phase separation. We show that crystallization leads to a network structure different from the original phase separation pattern.
-Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface,
-driving evaporation of nearby liquid. This process represents the colloidal analogue of the Bergeron process, i.e.,
-ice crystals formation in mixed phase clouds and is at the origin of rain. Our finding gives us full experimental
-access to the kinetic pathway of this phenomenon important for climate science. We argue that similar states can form as the
-result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems,
-if dynamics of the liquid phase is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate 
-vitrification.
+With confocal microscopy measurements we find a novel crystallization pathway of a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation.
+We show that crystallization leads to a network structure different from the original phase separation pattern.
+Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface, driving evaporation of nearby liquid.
+This process represents the colloidal analogue of the Bergeron process, i.e., ice crystals formation in mixed phase clouds and is at the origin of rain.
+Our finding gives us full experimental access to the kinetic pathway of this phenomenon important for climate science.
+We argue that similar states can form as the result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems, if dynamics of the liquid phase is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
 \end{sciabstract}

### 2015-11-19 crystal-gel/nature/Supplementary.tex

--- 2015-10-05 science/SOM2.tex
@@
 \begin{document}
-\title{``Colloidal cold rain'' as a novel route to crystallization via viscoelastic phase separation}
-
-\author{Hideyo Tsurusawa$1$, John Russo$^1$, Mathieu Leocmach$^{2}$ and Hajime Tanaka$^{1,\ast}$ 
-\\
-\normalsize{$^1$Institute of Industrial Science, University of Tokyo,}\\
-\normalsize{4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}\\
-\normalsize{$^2$Institut Lumière Matière, CNRS UMR 5306, Université Claude Bernard Lyon 1, }\\
-\normalsize{Université de Lyon, Lyon, 69622 Villeurbanne Cedex, France}\\
-\normalsize{$^\ast$Corresponding author. E-mail: tanaka@iis.u-tokyo.ac.jp}}
-
-%\date{Received August 11, 2011}
-\maketitle
-
-%\centerline{\bf \Large Supporting Online Material}
 
 \noindent
-{\bf METHODS}
+{\bf Supplementary Information}
 
+\vspace{5mm}
 \noindent
-{\bf Experimental.}
-We use \textsc{pmma} (poly(methyl methacrylate)) colloids sterically stabilized with methacryloxypropyl terminated \textsc{pdms}(poly(dimethyl siloxane)) and fluorescently labelled with rhodamine isothiocyanate chemically bonded to the \textsc{pmma}.
-The average colloid diameter is 2.1 $\mu$m. 
-We assess that the size distribution of our particle is Gaussian with a polydispersity below 5\% via direct confocal measurements~\cite{Leocmach2013}.
-This small polydispersity allows crystallization.
-We disperse the particles in refractive index and density matching mixture of cis-decalin (Tokyo Kasei) and bromocyclohexane (Sigma-Aldrich). 
-The precision of density matching is $\sim 10^{-4}$ mg/ml, for which the gravity effect on our gelation processes is negligible up to 12 hours.
+\emph{\bf Supplementary Movie 1}.
+This movie shows the early stage of gelation process of $\phi\approx 0.30$ and $c_p=0.38$ mg/g, from the original confocal images at the same z-position as Fig.~3a.
+Total duration is 1160 s (530~$\tau$).
+We quickly started scanning the sample after exchanging the salt-reservoir with the solvent including salt.
+At the beginning, salt had not reached the scanning volume and colloids were dispersed by electrostatic repulsion.
+At 0:02 in the movie, salt diffused into the scanning volume and screened the surface charges of colloids.
+Then, colloids immediately initiated phase separation by short-ranged attractions.
 
-To induce short-ranged depletion attraction, we use polystyrene (TOSOH) of molecular weight 3.8 Mdalton as non-adsorbing polymer. 
-Experiments are conducted at 27 $^\circ$C, some 80 $^\circ$C above the theta temperature in this solvents mixture~\cite{Royall2007}.
-A Flory scaling of the measurements of~\cite{lu2008gelation} yields a radius of gyration $R_g=76$ nm, and thus the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.07$.
-
-In the absence of salt, the Debye length is expected to reach a few $\mu$m and the (weakly) charged colloids experience a long range electrostatic repulsion.
-We confirm that colloids never come close enough to feel the short-ranged attraction.
-Screening by tetrabutylammonium bromide (Fluka) at saturated concentration brings down the Debye length to about several nanometers practically discarding the repulsion.
-Thus, salt injection can screen the Coulomb repulsion and make the polymer-induced depletion attraction effective, initiating phase separation.
-
-Our sample cell has two layer compartments separated by membrane filters.
-The first layer is a sample container with thickness of 200 $\mu$m.
-A mixture of colloids, polymer, and solvent without salt is  set in the first layer for microscopy observation.
-The second layer is a salt-reservoir with a half-opened structure, which allos us to exchange or insert a reservoir solution.
-The volume ratio of the first and second compartments is approximately 1:100.
-These two compartments are separated by a membrane filter with its pore size of 0.1 $\mu$m, which passes only salt ions.
-
-In our experiments, the reservoir solution was initially a polymer solution without salt and electrostatic repulsion by the unscreened surface charges inhibited colloidal aggregation.
-Under microscopy observation, we quickly exchanged the reservoir solution into a polymer solution with salt at saturated concentration (4 mM) and sealed the half-opened reservoir with cover glass to avoid evaporation of solvent.
-Salt diffused into the first layer within a few minutes, typically 2 minutes, screened the surface charges, and initiated colloidal aggregation.
-
-The data are collected on a Leica SP5 confocal microscope, using 532 nm laser excitation.
-The scanning volume is 98 $\times$ 98 $\times$ 53 $\mu$m$^3$, which contains $\sim 10^4$ colloid particles.
-Our spatial resolution is 192 nm / pixel.
-The Brownian time is $\tau_B \approx 4.5$ s.
-To be able to follow individual trajectories, we perform a 3D scan every 10 s at early time and every 30 s later.
-
-%\paragraph*{State points studied}
-%The experimental data is taken at two different volume fractions ($\phi=0.10$ and $\phi=0.25$) and for different values
-%of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ for $\phi=0.25$ and $c_p=0.82,1.36$ for $\phi=0.10$.
-%For each state points, scans at early times (every $10s$) and at the late stage (every $30s$) of the gelation process.
-%
-%We will consider two particles bonded if their center-to-center distance is within $12$ pixels. The radius of a colloidal
-%particle is $5.365$ pixels.
-
+\vspace{5mm}
 \noindent
-{\bf Structural analysis.}
-To account for imprecision in particle localisation close to contact, we consider two particles bonded if their center-to-center distance is within 2.3 $\mu$m $>\sigma+2R_g$.
-
-To detect percolation, two different methods have been employed. In the first one we detect percolation
-by looking at which frame the cluster size distribution has a more extended power-law decay. In the second technique,
-we simply measure the spatial extent of the largest cluster and consider it percolating when it is comparable to the
-size of the field of view of the microscope. Both methods lead to essentially identical percolation time for each state point.
-
-%\bibliographystyle{Science}
-\bibliographystyle{apsrev}
-\bibliography{biblio}
+\emph{\bf Supplementary Movie 2}.
+This movie shows the late stage after Supplemental Movie 1 with scanning duration of 27000~s (12300~$\tau$).
 

### 2015-11-19 nature/SOM2.tex

--- 2015-10-18 science/SOM2.tex
@@
 \begin{document}
-\title{``Colloidal cold rain'' as a novel route to crystallization via viscoelastic phase separation}
 
-\author{Hideyo Tsurusawa$1$, John Russo$^1$, Mathieu Leocmach$^{2}$ and Hajime Tanaka$^{1,\ast}$ 
-\\
-\normalsize{$^1$Institute of Industrial Science, University of Tokyo,}\\
-\normalsize{4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}\\
-\normalsize{$^2$Institut Lumière Matière, CNRS UMR 5306, Université Claude Bernard Lyon 1, }\\
-\normalsize{Université de Lyon, Lyon, 69622 Villeurbanne Cedex, France}\\
-\normalsize{$^\ast$Corresponding author. E-mail: tanaka@iis.u-tokyo.ac.jp}}
-
-%\date{Received August 11, 2011}
-\maketitle
-
-%\centerline{\bf \Large Supporting Online Material}
-
-\noindent
-{\bf METHODS}
-
-\noindent
-{\bf Experiments}
-
-We use \textsc{pmma} (poly(methyl methacrylate)) colloids sterically stabilized with methacryloxypropyl terminated \textsc{pdms}(poly(dimethyl siloxane)) and fluorescently labelled with rhodamine isothiocyanate chemically bonded to the \textsc{pmma}.
-The average colloid diameter is 2.2 $\mu$m, measured via the diffusion constant in dilute conditions without polymer. The Brownian time is $\tau_B \approx 2.3$ s. 
-We assess that the size distribution of our particle is Gaussian with a polydispersity below 5\% via direct confocal measurements~\cite{Leocmach2013}.
-This small polydispersity allows crystallization.
-We disperse the particles in refractive index and density matching mixture of cis-decalin (Tokyo Kasei) and bromocyclohexane (Sigma-Aldrich). 
-The precision of density matching is $\sim 10^{-4}$ mg/ml, for which the gravity effect on our gelation processes is negligible up to 12 hours.
-
-To induce short-ranged depletion attraction, we use polystyrene (TOSOH) of molecular weight 3.8 Mdalton as non-adsorbing polymer. 
-Experiments are conducted at 27 $^\circ$C, some 80 $^\circ$C above the theta temperature in this solvents mixture~\cite{Royall2007}.
-A Flory scaling of the measurements of~\cite{lu2008gelation} yields a radius of gyration $R_g=82$ nm, and thus the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.075$ and the overlap mass fraction of polymer $2.1$~mg/g.
-
-In the absence of salt, the Debye length is expected to reach a few $\mu$m and the (weakly) charged colloids experience a long range electrostatic repulsion.
-We confirm that colloids never come close enough to feel the short-ranged attraction.
-Screening by tetrabutylammonium bromide (Fluka) at saturated concentration brings down the Debye length to about several nanometers practically discarding the repulsion.
-Thus, salt injection can screen the Coulomb repulsion and make the polymer-induced depletion attraction effective, initiating phase separation.
-
-Our sample cell has two layer compartments separated by membrane filters.
-The first layer is a sample container with thickness of 200 $\mu$m.
-A mixture of colloids, polymer, and solvent without salt is  set in the first layer for microscopy observation.
-The second layer is a salt-reservoir with a half-opened structure, which allos us to exchange or insert a reservoir solution.
-The volume ratio of the first and second compartments is approximately 1:100.
-These two compartments are separated by a membrane filter with its pore size of 0.1 $\mu$m, which passes only salt ions.
-
-In our experiments, the reservoir solution was initially a polymer solution without salt and electrostatic repulsion by the unscreened surface charges inhibited colloidal aggregation.
-Under microscopy observation, we quickly exchanged the reservoir solution into a polymer solution with salt at saturated concentration (4 mM) and sealed the half-opened reservoir with cover glass to avoid evaporation of solvent.
-Salt diffused into the first layer within a few minutes, typically 2 minutes, screened the surface charges, and initiated colloidal aggregation.
-
-The data are collected on a Leica SP5 confocal microscope, using 532 nm laser excitation.
-The scanning volume is 98 $\times$ 98 $\times$ 53 $\mu$m$^3$, which contains $\sim 10^4$ colloid particles.
-Our spatial resolution is 192 nm / pixel.
-To be able to follow individual trajectories, we perform a 3D scan every 10~s ($\approx 4\tau_B$) at early time and every 30 s later.
-
-%\paragraph*{State points studied}
-%The experimental data is taken at two different volume fractions ($\phi=0.10$ and $\phi=0.25$) and for different values
-%of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ for $\phi=0.25$ and $c_p=0.82,1.36$ for $\phi=0.10$.
-%For each state points, scans at early times (every $10s$) and at the late stage (every $30s$) of the gelation process.
-%
-%We will consider two particles bonded if their center-to-center distance is within $12$ pixels. The radius of a colloidal
-%particle is $5.365$ pixels.
-
-\noindent
-{\bf Structural analysis}
-
-To account for imprecision in particle localisation close to contact, we consider two particles bonded if their center-to-center distance is within 2.3 $\mu$m $>\sigma+2R_g$.
-
-To detect percolation, two different methods have been employed. In the first one we detect percolation
-by looking at which frame the cluster size distribution has a more extended power-law decay. In the second technique,
-we simply measure the spatial extent of the largest cluster and consider it percolating when it is comparable to the
-size of the field of view of the microscope. Both methods lead to essentially identical percolation time for each state point.
+\centerline{\bf \Large Supplementary Information}
 
@@
 \emph{Supplementary Movie 1}.
-This movie shows the early stage of gelation process of $\phi\approx 0.30$ and $c_p=0.38$ mg/g, from the original confocal images at the same z-position as Fig.~3A.
+This movie shows the early stage of gelation process of $\phi\approx 0.30$ and $c_p=0.38$ mg/g, from the original confocal images at the same z-position as Fig.~3a.
 Total duration is 1160 s (~530~$\tau$).

### 2015-11-19 nature/crystal_gel2.tex

--- 2015-10-19 science/crystal_gel_letter3.tex
@@
-\documentclass[12pt]{article}
-\usepackage{graphicx}
-\usepackage{epstopdf}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{scicite}
+\documentclass[preprint,amsmath,amssymb,superscriptaddress]{revtex4-1}
+\usepackage{graphicx,amsmath}
+\usepackage{color}
+\begin{document}
+\draft
 
-% \usepackage{hyperref}
-\usepackage{booktabs}
-\usepackage{color}
+\title{A novel route to the spontaneous formation of porous crystals via viscoelastic phase separation}
 
-\usepackage[normalem]{ulem}
+\author{Hideyo Tsurusawa} 
+\affiliation{Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}
+\author{John Russo}
+\affiliation{Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}
+\author{Mathieu Leocmach}
+\affiliation{Institut Lumière Matière, CNRS UMR 5306, Université Claude Bernard Lyon 1, Université de Lyon, Lyon, 69622 Villeurbanne Cedex, France}
+\author{Hajime Tanaka}
+\affiliation{Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}
 
-\topmargin 0.0cm
-\oddsidemargin 0.2cm
-\textwidth 16cm 
-\textheight 21cm
-\footskip 1.0cm
+\date{Received \today}
 
-\newenvironment{sciabstract}{%
-\begin{quote} \bf}
-{\end{quote}}
-
-\newcounter{lastnote}
-\newenvironment{scilastnote}{%
-\setcounter{lastnote}{\value{enumiv}}%
-\addtocounter{lastnote}{+1}%
-\begin{list}%
-{\arabic{lastnote}.}
-{\setlength{\leftmargin}{.22in}}
-{\setlength{\labelsep}{.5em}}}
-{\end{list}}
-
-\title{A novel route to crystallization via viscoelastic phase separation}
-
-\author{Hideyo Tsurusawa$^1$, John Russo$^1$, Mathieu Leocmach$^{2}$ and Hajime Tanaka$^{1,\ast}$ 
-\\
-\normalsize{$^1$Institute of Industrial Science, University of Tokyo,}\\
-\normalsize{4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}\\
-\normalsize{$^2$Institut Lumière Matière, CNRS UMR 5306, Université Claude Bernard Lyon 1, }\\
-\normalsize{Université de Lyon, Lyon, 69622 Villeurbanne Cedex, France}\\
-\normalsize{$^\ast$Corresponding author. E-mail: tanaka@iis.u-tokyo.ac.jp}}
-
-%\date{Received}
-
-\begin{document}
-
-\baselineskip24pt
-
-% Make the title.
+\begin{abstract}
+{\bf
+Viscoelastic phase separation of colloidal suspensions can be interrupted either by the glass transition, with the formation of colloidal gels~\cite{poon2002,zaccarelli2007,piazza1994phase,verhaegh1997transient,tanaka1999colloid,lu2008gelation}, or by crystallization, which results in the formation of a crystalline network~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways,sabin2012}.
+But while the former process has been studied in detail, the latter has never been observed experimentally at a single-particle level and it is thus lacking a microscopic understanding.
+With a new confocal microscopy protocol, we unveil a novel crystallization pathway to sponge-like porous structures in a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation.
+We show that crystallization leads to a network structure different from the original phase separation pattern.
+Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface, driving evaporation of nearby liquid.
+This process represents the colloidal analogue of the Bergeron process~\cite{glickman2000glossary}, i.e., ice crystals formation in mixed phase clouds and is at the origin of rain.
+Our finding gives us full experimental access to the kinetic pathway of this poorly understood phenomenon important for climate science.
+\textcolor{blue}{The key is complex dynamic interplay involving the three phases upon simultaneous occurrence of two non-equilibrium phenomena, phase separation and crystallization.
+We argue that similar crystal-gel states can be formed as the result of the gas-liquid phase separation of any crystallizable components, such as monoatomic and molecular systems, if dynamics of the liquid phase is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
+This novel mechanism not only realizes single-step spontaneous formation of nano-porous crystals of metals and semiconductors without dealloying \cite{erlebacher2001evolution}, which may be important for catalytic, optical, sensing, and filtration applications \cite{ding2004nanoporous,ding2009nanoporous, wittstock2010nanoporous}, but also sheds new light on crystal network formation in more complex systems such as magma \cite{philpotts1998role}, biominerals \cite{rousseau2005multiscale}, and foods \cite{deman1987fat}.} }
+\end{abstract}
 
 \maketitle
-\vspace{-1.5cm}
-\begin{sciabstract}
-
-With confocal microscopy measurements we find a novel crystallization pathway of a colloid-polymer mixture undergoing viscoelastic
-gas-liquid phase separation. We show that crystallization leads to a network structure different from the original phase separation pattern.
-Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface,
-driving evaporation of nearby liquid. This process represents the colloidal analogue of the Bergeron process, i.e.,
-ice crystals formation in mixed phase clouds and is at the origin of rain. Our finding gives us full experimental
-access to the kinetic pathway of this phenomenon important for climate science. We argue that similar states can form as the
-result of the gas-liquid phase separation of crystallizable components, such as monoatomic and molecular systems,
-if dynamics of the liquid phase is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate 
-vitrification.
-\end{sciabstract}
-
-\maketitle
-
-\clearpage
 
@@
 At the glass transition point, the percolated network structure is dynamically stabilised by vitrification 
-of the dense liquid phase~\cite{pusey1993dynamics,ilett1995phase,verhaegh1997transient,tanaka1999colloid,foffi2002,buzzaccaro2007sticky,zaccarelli2007,lu2008gelation,zaccarelli2008gelation,testard2011}.
-More precisely, the dynamical stabilization is due to percolation of locally favored structures, which are locally stable non-crystalline structures \cite{royall2008g}.
+of the dense liquid phase~\cite{pusey1993dynamics,piazza1994phase,ilett1995phase,verhaegh1997transient,tanaka1999colloid,foffi2002,buzzaccaro2007sticky,zaccarelli2007,lu2008gelation,zaccarelli2008gelation,testard2011}.
+More precisely, the dynamical stabilization is due to percolation of locally favoured structures, which are locally stable non-crystalline structures \cite{royall2008g}.
 
@@
 Indeed the possibility of a different class of gels, which are stabilized by crystallization,
-was suggested by numerical simulations~\cite{fortini2008crystallization,perez2011pathways} and observed in microgravity experiments \cite{sabin2012}. 
+was suggested by numerical simulations~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways} and observed in microgravity experiments \cite{sabin2012}. 
 For colloidal systems, we can in principle access structural evolution in real time at a single-particle level; however, there has so far been no confocal microscopy  
@@
 
-In this Report, we show that, for low polymer concentration, the system can undergo a novel crystallization pathway that leads to the formation
+In this Letter, we show that, for low polymer concentration, the system can undergo a novel crystallization pathway that leads to the formation
 of crystalline droplets.
@@
 The process is analogous to the Bergeron process of ice formation in mixed phase clouds, where ice droplets grow at the
-expense of the supercooled liquid droplets due to their lower saturated vapor pressure. 
+expense of the supercooled liquid droplets due to their lower saturated vapour pressure. 
 This is speculated to be the primary mechanism for the formation of raindrops. 
@@
 of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ mg/g for $\phi\approx 0.30$ and $c_p=0.82,1.36$ mg/g for $\phi\approx 0.13$.
-In Fig.~\ref{fig:phase_separation}A, we superimpose the state points with the phase diagram, which we calculated from generalised free volume theory~\cite{Fleer2008} using the experimental values for the size of the colloids, the radius of gyration of the polymer, and the polymer concentration (see Supplementary Information).
+In Fig.~\ref{fig:phase_separation}a, we superimpose the state points with the phase diagram, which we calculated from generalised free volume theory~\cite{Fleer2008} using the experimental values for the size of the colloids, the radius of gyration of the polymer, and the polymer concentration (see Methods).
 For each state point, scans were made at early times (every $10$~s, i.e. $4\tau_B$) and at the late stage (every $30$~s) of the gelation process.
@@
 %
-% Previous simulation results~\cite{tanaka2000,tanaka2007spontaneous,furukawa2010key} have suggested that hydrodynamic interactions play
-% a key role in the process of colloidal gelation. Without hydrodynamic interactions, particles have the tendency to
-% aggregate in compact structures and subsequently form thick network structures. This also limits the possibility to
-% form arrested gel states only at relatively high volume fractions. With hydrodynamic interactions, particles first form
-% a transient gel even at very low volume fractions, and only later they increase the number of
-% nearest neighbours to minimize the energy of the structure. 
-% Thus, hydrodynamic interactions lead to the formation of gels that are very far from equilibrium and under a strong thermodynamic driving force 
-% towards more stable compact structures. 
-% The resulting transition from open to more compact networks occurs only through
-% the breaking of the bonds that have accumulated more stress. The process was called stress-driven ageing~\cite{tanaka2007spontaneous}, 
-% which is characteristic of viscoelastic phase separation \cite{tanaka2000viscoelastic}. 
-% Below we provide the first experimental proof of this process.
-% 
-% In Fig.~\ref{fig:stress}a, we plot the fraction of liquid and gas particles as a function of time during phase separation
-% at the state points of $\phi\approx 0.13$. We observe that at high polymer concentration ($c_p=1.36$, dashed lines) both the
-% liquid and gas branch change monotonically with time as the phase separation proceeds. At low polymer concentration
-% ($c_p=0.82$, continuous lines), on the other hand, we find a non-monotonic evolution of both the liquid and gas branches, which is signalling
-% a structural reorganization of the network. To prove that this is indeed a stress-driven ageing process, in Fig.~\ref{fig:stress}b 
-% we plot the average number of bond-breaking events for the low-polymer concentration state point. The figure clearly shows that
-% the network restructuring follows a peak in the number of bond-breaking events. We have thus found direct evidence of the
-% stress-driven ageing process in colloidal gels that is a direct consequence of the formation of a thin network with the help of hydrodynamic interactions in the first stages
-% of the aggregation process.
-% 
-% In Fig.~\ref{fig:stress}c we plot the time evolution of the structure factor, $S(q)$, for the state point
-% undergoing network reorganization $\phi\approx 0.13$ and $c_p=0.82$. The time of network reorganization as found in Fig.~\ref{fig:stress}a and b
-% corresponds to a steep increase of the low wavenumber scattering from the network, that signals the coarsening of the network structure. 
-% 
-% The same conclusions also hold for the state point at $\phi\approx 0.30$. We observe network restructuring only for the state point
-% at lower polymer concentration, $c_p=0.38$. Interestingly, this is also the only state point that undergoes crystallization,
-% meaning that the coarsening of the network of the dense liquid phase and the resulting formation of a wider liquid branch is a necessary condition for crystallization.
 
@@
 All samples share the same early stages of spinodal decomposition. 
+A typical phase ordering process observed at $\phi\approx 0.30$ and $c_p=0.38$ mg/g can be seen in Supplementary Movies 1 and 2. 
 Due to strong dynamical asymmetry between colloids and the solvent \cite{tanaka1999colloid}, the colloidal particles start aggregating, 
 eventually forming a percolating network. Thus, a dense network (liquid) coexists with freely diffusing monomers (gas).
-In Fig.~\ref{fig:phase_separation}A, the state points which did not phase separate are indicated with open triangles, while
+In Fig.~\ref{fig:phase_separation}a, the state points which did not phase separate are indicated with open triangles, while
 the other symbols indicate samples undergoing phase separation. The (black) continuous line is the gas spinodal computed from free volume theory,
@@
 The topology of the network is similar to all our samples at the initial stages. 
-This is shown for example in Fig.~\ref{fig:phase_separation}B, where we plot the radius of gyration of colloidal clusters ($R_g$)
+This is shown for example in Fig.~\ref{fig:phase_separation}b, where we plot the radius of gyration of colloidal clusters ($R_g$)
 as a function of the cluster size ($s$) at the percolation point. Two particles belong to the same cluster if their distance is
 within the first peak of the radial distribution function. All samples show the same scaling law, $R^2_{\rm g}\sim s^{2/D}$, which is compatible with
-the random gelation universality class exponent ($D=2.53$) as shown by the dashed line in Fig.~\ref{fig:phase_separation}B
+the random gelation universality class exponent ($D=2.53$) as shown by the dashed line in Fig.~\ref{fig:phase_separation}b
 (we do not imply that the formation of the network follows the random gelation universality class, just that our
@@
 
-While sharing similar structural properties at early times, the samples show significantly different behaviors in the later stages of the phase separation process.
-To examine this in more detail, we plot in Fig.~\ref{fig:phase_separation}C the structure factor for the final stages of the gelation process for all state points.
+%%%%
+%Previous simulation results~\cite{tanaka2000,tanaka2007spontaneous,furukawa2010key} have suggested that hydrodynamic interactions play a key role in the process of colloidal gelation.
+Without hydrodynamic interactions, particles have the tendency to aggregate in compact structures and subsequently form thick network structures.
+This also limits the possibility to form arrested gel states only at relatively high volume fractions.
+With hydrodynamic interactions, particles first form a transient gel even at very low volume fractions, and only later they increase the number of nearest neighbours to minimize the energy of the structure.
+Thus, hydrodynamic interactions lead to the formation of gels that are very far from equilibrium and under a strong thermodynamic driving force towards more stable compact structures.
+The resulting transition from open to more compact networks occurs only through the breaking of the bonds that have accumulated more stress.
+The process was called stress-driven ageing~\cite{tanaka2007spontaneous}, which is characteristic of viscoelastic phase separation \cite{tanaka2000viscoelastic}.
+Below we provide the first experimental proof of this process.
+
+%In Fig.~\ref{fig:stress}a, we plot the fraction of liquid and gas particles as a function of time during phase separation at the state points of $\phi\approx 0.13$.
+We observe that at high polymer concentration ($c_p=1.36$, dashed lines) both the liquid and gas branch change monotonically with time as the phase separation proceeds.
+At low polymer concentration ($c_p=0.82$, continuous lines), on the other hand, we find a non-monotonic evolution of both the liquid and gas branches, which is signalling a structural reorganization of the network.
+To prove that this is indeed a stress-driven ageing process, in Fig.~\ref{fig:stress}b we plot the average number of bond-breaking events for the low-polymer concentration state point.
+The figure clearly shows that the network restructuring follows a peak in the number of bond-breaking events.
+We have thus found direct evidence of the stress-driven ageing process in colloidal gels that is a direct consequence of the formation of a thin network with the help of hydrodynamic interactions in the first stages of the aggregation process.
+
+%In Fig.~\ref{fig:stress}c we plot the time evolution of the structure factor, $S(q)$, for the state point undergoing network reorganization $\phi\approx 0.13$ and $c_p=0.82$.
+The time of network reorganization as found in Fig.~\ref{fig:stress}a and b corresponds to a steep increase of the low wavenumber scattering from the network, that signals the coarsening of the network structure.
+
+%The same conclusions also hold for the state point at $\phi\approx 0.30$.
+We observe network restructuring only for the state point at lower polymer concentration, $c_p=0.38$.
+Interestingly, this is also the only state point that undergoes crystallization, meaning that the coarsening of the network of the dense liquid phase and the resulting formation of a wider liquid branch is a necessary condition for crystallization.
+
+%%%
+
+While sharing similar structural properties at early times, the samples show significantly different behaviours in the later stages of the phase separation process.
+To examine this in more detail, we plot in Fig.~\ref{fig:phase_separation}c the structure factor for the final stages of the gelation process for all state points.
 Calculation of the structure factor is done with the Hanning window function, to minimize boundary effects.
@@
 with high polymer concentration retain the exponent $D=2.5(3)$, 
-which is the random gel universality class exponent, as we also observed in the early stages of the gelation process (see Fig.~\ref{fig:phase_separation}B),
+which is the random gel universality class exponent, as we also observed in the early stages of the gelation process (see Fig.~\ref{fig:phase_separation}b),
 the state point with low polymer concentration ($\phi\approx 0.30$ and $c_p=0.38$ mg/g) displays the biggest deviation from that exponent. In the figure we also
 plot the exponent $D=3$, which corresponds to volume growth, which we will show in the following analysis to be the correct exponent for this state. 
-The difference is due to the onset of crystallization in the low-polymer concentration sample. This is already evident in the high $q$ behavior of
-the structure factors (Fig.~\ref{fig:phase_separation}C), where the diffraction planes appear as sharper peaks for $\phi\approx 0.30$ and $c_p=0.38$ mg/g.
+The difference is due to the onset of crystallization in the low-polymer concentration sample. This is already evident in the high $q$ behaviour of
+the structure factors (Fig.~\ref{fig:phase_separation}c), where the diffraction planes appear as sharper peaks for $\phi\approx 0.30$ and $c_p=0.38$ mg/g.
 
-To gain additional insight, in Fig.~\ref{fig:phase_separation}D we plot the average local volume fraction, computed
-through Voronoi diagrams, as a function of the number of neighbors. For all state points, we can see that the average volume fraction of a colloidal
-particle increases with its number of neighbors. Approaching close packing (12 neighbors), we clearly see two families of curves. The state point
+To gain additional insight, in Fig.~\ref{fig:phase_separation}d we plot the average local volume fraction, computed
+through Voronoi diagrams, as a function of the number of neighbours. For all state points, we can see that the average volume fraction of a colloidal
+particle increases with its number of neighbours. Approaching close packing (12 neighbours), we clearly see two families of curves. The state point
 at $\phi\approx 0.30$ and $c_p=0.38$ mg/g reaches close packing at a volume fraction of around $\sim 68\%$, which is in reasonable agreement with the
@@
 % spinodal liquid. Also this suggests that the growth mechanism of the crystal is not simply due to filling of the spinodal liquid network, as was
-% instead observed in Ref.~\cite{fortini2008crystallization,perez2011pathways} for brownian dynamics simulations.
+% instead observed in Ref.~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways} for brownian dynamics simulations.
 
@@
 With this analysis we reveal that crystallization events indeed start occurring for the state point $\phi\approx 0.30$ and $c_p=0.38$ mg/g, which has the lowest
-polymer concentration of all state points. This is shown in Fig.~\ref{fig:crystals}A, where the black line indicates the average size of the crystallites, 
+polymer concentration of all state points. This is shown in Fig.~\ref{fig:crystals}a, where the black line indicates the average size of the crystallites, 
 while the red line indicates the average number of crystallites as a function of time. The number of crystallites first rapidly increases as nucleation events
@@
 idea of enhanced crystallization rates near metastable critical points~\cite{ten1997enhancement,olmsted1998spinodal} and the
-results of Refs.~\cite{fortini2008crystallization,perez2011pathways},
+results of Refs.~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways},
 which speculated two different arrest mechanisms: crystallization at low polymer concentration, and dynamic arrest at high polymer concentrations.
@@
 which was neither considered nor observed before.
-In the following we will show that the growth mechanism of the crystal is not only due to filling of the spinodal liquid network, as was observed in Refs.~\cite{fortini2008crystallization,perez2011pathways} for Brownian Dynamics simulations, but also involves
+In the following we will show that the growth mechanism of the crystal is not only due to filling of the spinodal liquid network, as was observed in Refs.~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways} for Brownian Dynamics simulations, but also involves
 evaporation of the liquid phase and subsequent direct condensation of the gas phase on the crystalline seeds. 
-To confirm the presence of the two processes, we plot in Fig.~\ref{fig:crystals}B the gyration radius of individual crystalline nuclei for the state point 
+To confirm the presence of the two processes, we plot in Fig.~\ref{fig:crystals}b the gyration radius of individual crystalline nuclei for the state point 
 $\phi\approx 0.30$ and $c_p=0.38$ mg/g. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with 
@@
 follows a volume growth with the formation of isotropic crystal droplets. The same scaling law was suggested in the analysis of the
-structure factors, Fig.~\ref{fig:phase_separation}C, but here it is shown directly by the analysis of the gyration radius.
+structure factors, Fig.~\ref{fig:phase_separation}c, but here it is shown directly by the analysis of the gyration radius.
 
@@
 where the process of crystal growth by direct attachment of gas particles can be observed directly. 
-From the trajectory, we extract slabs of 10 particle diameter thickness for configurations at different times, shown in Fig.~\ref{fig:transitions}A. 
+From the trajectory, we extract slabs of 10 particle diameter thickness for configurations at different times, shown in Fig.~\ref{fig:transitions}a. 
 Particles are colored according to their phases (gas, liquid, crystal, and crystal surface). The figure shows the first nucleation 
@@
 
-Next we show in Fig.~\ref{fig:transitions}B the fraction of particles in each different phase for the crystallization trajectory at
+Next we show in Fig.~\ref{fig:transitions}b the fraction of particles in each different phase for the crystallization trajectory at
 $\phi\approx 0.30$ and $c_p=0.38$ mg/g, after the liquid-gas phase separation has occurred. The process of crystallization is characterized 
@@
 is then accompanied by an increase in the fraction of gas particles: as the first crystals start to reach the gas phase, liquid particles evaporate 
-to the gas phase due to the higher vapor pressure of the liquid phase compared to the crystalline phase. After the onset
+to the gas phase due to the higher vapour pressure of the liquid phase compared to the crystalline phase. After the onset
 of a steady-state gas population, the crystallization proceeds by adding surface particles from the gas phase to the small crystallites. To confirm this process we
-examine in Fig.~\ref{fig:transitions}C the transfer matrix probabilities $P_{\alpha\beta}$, where $\alpha$ and $\beta$ are each of the 
+examine in Fig.~\ref{fig:transitions}c the transfer matrix probabilities $P_{\alpha\beta}$, where $\alpha$ and $\beta$ are each of the 
 phases, gas, liquid and crystal (which includes hcp, fcc and surface particles). Liquid particles undergo two processes: on one
@@
 We can thus conclude that the crystallization proceeds via two distinct mechanisms:
-(i) freezing of the fluid; (ii) evaporation of the fluid and condensation of the vapor on the crystals, i.e., a coupled process of evaporation and sublimation.
+(1) freezing of the fluid; (2) evaporation of the fluid and condensation of the vapour on the crystals, i.e., a coupled process of evaporation and sublimation.
 The first process dominates the initial stages of crystallization, while the second one becomes effective once
-crystalline nuclei have outgrown the liquid regions in which they form. 
+crystalline nuclei have outgrown the liquid regions in which they form. \textcolor{blue}{Process (2) has been completely overlooked so far, although it plays a crucial role 
+in not only in the kinetic pathway but also the morphology, as we discussed above. We note that this is the crucial step  
+in the Bergeron process~\cite{glickman2000glossary,morrison2012resilience} (see below). }
 % 
@@
 Depending on the polymer concentration, there are two possible arrest mechanisms. 
-(i) Crystallization, observed at $\phi\approx 0.30$ and $c_p=0.38$ mg/g, small crystalline nuclei appear inside the liquid network, reach the surface 
+(a) Crystallization, observed at $\phi\approx 0.30$ and $c_p=0.38$ mg/g, small crystalline nuclei appear inside the liquid network, reach the surface 
  of the liquid branches, and then grow by addition of particles from the gas phase. The final structure is a network of crystal droplets,
  as confirmed by the fractal dimension of the branches, the volume fraction of the particles within the branches, and bond orientational analysis. 
-(ii) Dynamic arrest, particle arrest when the dynamics inside the liquid branch becomes slow, which should happen at the intersection of the
+(b) Dynamic arrest, particle arrest when the dynamics inside the liquid branch becomes slow, which should happen at the intersection of the
  glass line with the liquid side of the coexistence curve.
@@
 The process we observed is akin to the Bergeron process which is the primary mechanism for the formation of rain drops in clouds~\cite{glickman2000glossary,morrison2012resilience}.
-In clouds there is a mixture of ice crystals and supercooled water. The vapor phase is in coexistence with the liquid phase, but is supersaturated
+In clouds there is a mixture of ice crystals and supercooled water. The vapour phase is in coexistence with the liquid phase, but is supersaturated
 with respect to the ice crystals. This causes the water droplets to evaporate and sublimate directly on the ice crystals. Our system can then
@@
 leading to vitrification.   
-Many monoatomic and single-component molecular systems can satisfy all these conditions in a certain range of the temperature and pressure. 
-This can be seen, for example, by looking at the phase diagram of a Lennard-Jones liquid, which represents many molecular systems 
+\textcolor{blue}{Many monoatomic and single-component molecular systems can satisfy all these conditions in a certain range of the temperature and pressure. 
+This can be seen, for example, by looking at the phase diagram of a Lennard-Jones liquid \cite{lodge1997brownian}, which represents many molecular systems 
 without specific directional interactions. 
@@
 and industrial applications, although they have not attracted much attention so far.  
-We speculate that our scenario could play a role in the formation of crystal networks observed in dynamically asymmetric mixtures, 
+For example, nano-porous crystals of noble metals such as Au have special functions, which are relevant to catalytic, optical, sensing, and filtration applications 
+\cite{ding2004nanoporous, ding2009nanoporous, wittstock2010nanoporous}. Usually such nano-porous materials are formed via at least two steps: for example, phase separation and dealloying of a mixture
+\cite{erlebacher2001evolution}. Unlike such a method, our novel mechanism allows us to spontaneously form sponge-like nano-porous crystals \emph{in a single step}, 
+which may have an impact on many applications. 
+We note that laser ablation of metals is a promising method for this purpose \cite{povarnitsyn2013mechanisms}.}
+We also speculate that our scenario could play a role in the formation of crystal networks observed in dynamically asymmetric mixtures, 
 which includes magma \cite{philpotts1998role}, biominerals \cite{rousseau2005multiscale}, and foods \cite{deman1987fat}.
 
-\bibliographystyle{Science}
+\vspace{1cm}
+\noindent
+{\bf METHODS}
+
+\noindent
+{\bf Experiments}
+
+We use \textsc{pmma} (poly(methyl methacrylate)) colloids sterically stabilized with methacryloxypropyl terminated \textsc{pdms}(poly(dimethyl siloxane)) and fluorescently labelled with rhodamine isothiocyanate chemically bonded to the \textsc{pmma}.
+The average colloid diameter is 2.2 $\mu$m, measured via the diffusion constant in dilute conditions without polymer. The Brownian time is $\tau_B \approx 2.3$ s. 
+We assess that the size distribution of our particle is Gaussian with a polydispersity below 5\% via direct confocal measurements~\cite{Leocmach2013}.
+This small polydispersity allows crystallization.
+We disperse the particles in refractive index and density matching mixture of cis-decalin (Tokyo Kasei) and bromocyclohexane (Sigma-Aldrich). 
+The precision of density matching is $\sim 10^{-4}$ mg/ml, for which the gravity effect on our gelation processes is negligible up to 12 hours.
+
+To induce short-ranged depletion attraction, we use polystyrene (TOSOH) of molecular weight 3.8 Mdalton as non-adsorbing polymer. 
+Experiments are conducted at 27 $^\circ$C, some 80 $^\circ$C above the theta temperature in this solvents mixture~\cite{Royall2007}.
+A Flory scaling of the measurements of~\cite{lu2008gelation} yields a radius of gyration $R_g=82$ nm, and thus the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.075$ and the overlap mass fraction of polymer $2.1$~mg/g.
+
+In the absence of salt, the Debye length is expected to reach a few $\mu$m and the (weakly) charged colloids experience a long range electrostatic repulsion.
+We confirm that colloids never come close enough to feel the short-ranged attraction.
+Screening by tetrabutylammonium bromide (Fluka) at saturated concentration brings down the Debye length to about several nanometers practically discarding the repulsion.
+Thus, salt injection can screen the Coulomb repulsion and make the polymer-induced depletion attraction effective, initiating phase separation.
+
+Our sample cell has two layer compartments separated by membrane filters.
+The first layer is a sample container with thickness of 200 $\mu$m.
+A mixture of colloids, polymer, and solvent without salt is  set in the first layer for microscopy observation.
+The second layer is a salt-reservoir with a half-opened structure, which allos us to exchange or insert a reservoir solution.
+The volume ratio of the first and second compartments is approximately 1:100.
+These two compartments are separated by a membrane filter with its pore size of 0.1 $\mu$m, which passes only salt ions.
+
+In our experiments, the reservoir solution was initially a polymer solution without salt and electrostatic repulsion by the unscreened surface charges inhibited colloidal aggregation.
+Under microscopy observation, we quickly exchanged the reservoir solution into a polymer solution with salt at saturated concentration (4 mM) and sealed the half-opened reservoir with cover glass to avoid evaporation of solvent.
+Salt diffused into the first layer within a few minutes, typically 2 minutes, screened the surface charges, and initiated colloidal aggregation.
+
+The data are collected on a Leica SP5 confocal microscope, using 532 nm laser excitation.
+The scanning volume is 98 $\times$ 98 $\times$ 53 $\mu$m$^3$, which contains $\sim 10^4$ colloid particles.
+Our spatial resolution is 192 nm / pixel.
+To be able to follow individual trajectories, we perform a 3D scan every 10~s ($\approx 4\tau_B$) at early time and every 30 s later.
+
+%\paragraph*{State points studied}
+%The experimental data is taken at two different volume fractions ($\phi=0.10$ and $\phi=0.25$) and for different values
+%of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ for $\phi=0.25$ and $c_p=0.82,1.36$ for $\phi=0.10$.
+%For each state points, scans at early times (every $10s$) and at the late stage (every $30s$) of the gelation process.
+%
+%We will consider two particles bonded if their center-to-center distance is within $12$ pixels. The radius of a colloidal
+%particle is $5.365$ pixels.
+
+\noindent
+{\bf Structural analysis}
+
+To account for imprecision in particle localisation close to contact, we consider two particles bonded if their center-to-center distance is within 2.3 $\mu$m $>\sigma+2R_g$.
+
+To detect percolation, two different methods have been employed. In the first one we detect percolation
+by looking at which frame the cluster size distribution has a more extended power-law decay. In the second technique,
+we simply measure the spatial extent of the largest cluster and consider it percolating when it is comparable to the
+size of the field of view of the microscope. Both methods lead to essentially identical percolation time for each state point.
+
+\bibliographystyle{naturemag2}
 \bibliography{biblio}
 
+\vspace{1cm}
 \noindent
-{\bf Acknowledgments}
+{\bf Acknowledgements.} This study was partly supported by Grants-in-Aid for Scientific Research (S) (Grand No.
+21224011) and Specially Promoted Research (Grand No.
+25000002) from the Japan Society for the Promotion of Science (JSPS).
+
+%\vspace{0.3cm}
 \noindent
+{\bf Author Contributions.} 
 H.Ts. and J.R. contributed equally to this work. 
 H.Ta. conceived and supervised the project, H.Ts. performed experiments, J.R. analyzed the data, M.L. calculated the phase diagram, and all the authors discussed and wrote the manuscript.
-This study was partly supported by Grants-in-Aid for Scientific Research (S) (21224011) and Specially Promoted Research (25000002) from the Japan Society of the Promotion of Science (JSPS).
+
+%\vspace{0.3cm}
+\noindent
+{\bf Author Information.} 
+%Reprints and permissions information is available at www.nature.com/reprints.
+Correspondence and requests for materials should be addressed to H.T. ~(tanaka@iis.u-tokyo.ac.jp).
 
@@
  \includegraphics{phase_separation.pdf}
- \caption{Different routes of viscoelastic phase separation. 
-(A) Phase diagram.
+ \caption{{\bf Different routes of viscoelastic phase separation.} 
+{\bf a,} Phase diagram.
 Dashed lines are fluid-solid equilibrium coexistence boundaries.
@@
 The other symbols are the experimental points showing gel behaviour, consistently used in all panels.
-(B) Radius of gyration ($R_g$) as a function of cluster size ($s$) for all state points at percolation. The
+{\bf b,} Radius of gyration ($R_g$) as a function of cluster size ($s$) for all state points at percolation. The
  gray line represents the fractal dimension of the random gelation universality class.
-(C) Structure factors for the late stages of the gelation process. The gray and black lines represents respectively random gelation and compact fractal dimensions.
-(D) Local volume fraction around a colloidal particle, as a function of the number of nearest neighbors of the particle for the state points in the very late stage.
-Horizontal lines stress the value at 12 neighbors.
+{\bf c,} Structure factors for the late stages of the gelation process. The gray and black lines represents respectively random gelation and compact fractal dimensions.
+{\bf d,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle for the state points in the very late stage.
+Horizontal lines stress the value at 12 neighbours.
  }
@@
  % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
-\caption{Characterization of the crystal-gel.
-(A) Time evolution of the average size of the crystals (blue line) and the number of crystals (red line) for the state point $\phi\approx 0.30$ and $c_p=0.38$ mg/g.
-(B) Radius of gyration of crystalline nuclei at the late stages of gelation for the same state point.}
+\caption{{\bf Characterization of the crystal-gel.} {\bf a,} Time evolution of the average size of the crystals (blue line) and the number of crystals (red line) for the state point $\phi\approx 0.30$ and $c_p=0.38$ mg/g.
+{\bf b,} Radius of gyration of crystalline nuclei at the late stages of gelation for the same state point.}
  \label{fig:crystals}
@@
  \caption{
-Crystal gel formation.  
-(A) Reconstructions from confocal coordinates at $\phi\approx 0.30$ and $c_p=0.38$ mg/g.
+{\bf Crystal gel formation.}  
+{\bf a,} Reconstructions from confocal coordinates at $\phi\approx 0.30$ and $c_p=0.38$ mg/g.
 Depth of view is $5\sigma$.
@@
 Dark red: crystal; orange: surface of crystal; yellow: liquid; blue: gas.
-(B) Fraction of particles for each phase. Gray vertical lines indicate the times shown in (A). 
-(C) Transition probabilities for the same process as in (B). Here crystal includes surface particles.}
+{\bf b,} Fraction of particles for each phase. Gray vertical lines indicate the times shown in panel a. 
+{\bf c,} Transition probabilities for the same process as in panel b. Here crystal includes surface particles.}
 \label{fig:transitions}

### 2015-11-19 nature/crystal_gel2.tex
New figure.

@@
-\documentclass[preprint,amsmath,amssymb,superscriptaddress]{revtex4-1}
+\documentclass[preprint,amsmath,amssymb,superscriptaddress]{revtex4}
 \usepackage{graphicx,amsmath}
@@
 Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface, driving evaporation of nearby liquid.
-This process represents the colloidal analogue of the Bergeron process~\cite{glickman2000glossary}, i.e., ice crystals formation in mixed phase clouds and is at the origin of rain.
+This process represents the colloidal analogue of the Bergeron process~\cite{glickman2000glossary}, which explains the formation of ice crystals in mixed phase clouds and is at the origin of rain.
 Our finding gives us full experimental access to the kinetic pathway of this poorly understood phenomenon important for climate science.
-\textcolor{blue}{The key is complex dynamic interplay involving the three phases upon simultaneous occurrence of two non-equilibrium phenomena, phase separation and crystallization.
+The key is complex dynamic interplay involving the three phases upon simultaneous occurrence of two non-equilibrium phenomena, phase separation and crystallization.
 We argue that similar crystal-gel states can be formed as the result of the gas-liquid phase separation of any crystallizable components, such as monoatomic and molecular systems, if dynamics of the liquid phase is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
-This novel mechanism not only realizes single-step spontaneous formation of nano-porous crystals of metals and semiconductors without dealloying \cite{erlebacher2001evolution}, which may be important for catalytic, optical, sensing, and filtration applications \cite{ding2004nanoporous,ding2009nanoporous, wittstock2010nanoporous}, but also sheds new light on crystal network formation in more complex systems such as magma \cite{philpotts1998role}, biominerals \cite{rousseau2005multiscale}, and foods \cite{deman1987fat}.} }
+This novel mechanism not only realizes single-step spontaneous formation of nano-porous crystals of metals and semiconductors without dealloying \cite{erlebacher2001evolution}, which may be important for catalytic, optical, sensing, and filtration applications \cite{ding2004nanoporous,ding2009nanoporous, wittstock2010nanoporous}, but also sheds new light on crystal network formation in more complex systems such as magma \cite{philpotts1998role}, biominerals \cite{rousseau2005multiscale}, and foods \cite{deman1987fat}.}
 \end{abstract}
@@
 
+\textcolor{red}{
+A direct analysis of the colloidal positions at long times confirms that indeed the new structure obtained at
+$\phi\approx 0.30$ and $c_p=0.38$ mg/g has different morphological properties than usual colloidal gels.
+We directly compare the two structures in Fig.~\ref{fig:network}, where Gaussian filtering is used to depict the network structure as a continuous field.
+Panel a) presents the results for high
+polymer concentration gel, $c_p=1.07$ mg/g; panel b) presents the results for the low polymer concentration sample, $c_p=0.38$ mg/g.
+The comparison shows that the new arrest mechanism produces a network structure with bigger pores and thicker strands (panel b) compared
+to the colloidal gel sample (panel a). The bottom row shows the individual particle positions for slabs with a thickness of 5 particle diameters.
+From the figure it is immediately clear that the network strands at low polymer concentration are crystalline. The new arrest mechanism thus
+involve the formation of a crystal-gel network. In order to explain the morphological differences between colloidal gels and crystal-gels,
+we now consider the kinetics of formation in our samples.}
+
 To confirm the formation of crystalline seeds, we adopt bond orientational analysis to detect which particles are in crystalline environments~\cite{russo2013interplay}. 
@@
 The first process dominates the initial stages of crystallization, while the second one becomes effective once
-crystalline nuclei have outgrown the liquid regions in which they form. \textcolor{blue}{Process (2) has been completely overlooked so far, although it plays a crucial role 
-in not only in the kinetic pathway but also the morphology, as we discussed above. We note that this is the crucial step  
-in the Bergeron process~\cite{glickman2000glossary,morrison2012resilience} (see below). }
+crystalline nuclei have outgrown the liquid regions in which they form. Process (2) has been completely overlooked so far, although it plays a crucial role 
+not only in the kinetic pathway, but also in determining the morphology of the porous crystal, as we discussed above. We note that this is the crucial step  
+in the Bergeron process~\cite{glickman2000glossary,morrison2012resilience} (see below).
 % 
@@
 leading to vitrification.   
-\textcolor{blue}{Many monoatomic and single-component molecular systems can satisfy all these conditions in a certain range of the temperature and pressure. 
+Many monoatomic and single-component molecular systems can satisfy all these conditions in a certain range of the temperature and pressure. 
 This can be seen, for example, by looking at the phase diagram of a Lennard-Jones liquid \cite{lodge1997brownian}, which represents many molecular systems 
@@
 which may have an impact on many applications. 
-We note that laser ablation of metals is a promising method for this purpose \cite{povarnitsyn2013mechanisms}.}
+We note that laser ablation of metals is a promising method for this purpose \cite{povarnitsyn2013mechanisms}.
 We also speculate that our scenario could play a role in the formation of crystal networks observed in dynamically asymmetric mixtures, 
@@
 
+\begin{figure*}[!t]
+ \centering
+ \includegraphics[width=12cm]{new_fig2}
+ % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
+\caption{{\bf Percolated networks: a) gel; b) crystal-gel.} The gel state is obtained from the state
+point $\phi\approx 0.30$ and $c_p=1.07$ mg/g, while the crystal-gel is obtained at the same $\phi$ but
+at lower polymer concentration, $c_p=0.38$ mg/g. The top row shows depicts the networks as a continuous field that has
+been obtained by applying a gaussian filter to the position of the particles (of width equal to the inter-particle distance) and
+plotting the surface which bounds the highest 20\% values of the field. The lower row depicts slabs of 5 particle diameter thickness,
+with particles colored according to the following code: red (fcc), green (hcp), white (crystal surface particles), cyan (liquid particles), magenta (gas particles).} 
+ \label{fig:network}
+\end{figure*}
+
+\clearpage
+
 \begin{figure}[!t]

### 2015-11-20 crystal-gel/nature/crystal_gel2.tex

--- 2015-11-19 nature/crystal_gel2.tex
@@
 We show that crystallization leads to a network structure different from the original phase separation pattern.
-Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface, driving evaporation of nearby liquid.
-This process represents the colloidal analogue of the Bergeron process~\cite{glickman2000glossary}, which explains the formation of ice crystals in mixed phase clouds and is at the origin of rain.
+Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface, \textcolor{blue}{i.e.
+desublimation,} driving evaporation of nearby liquid.
+\textcolor{blue}{The latter processes via the gas phase have been overlooked so far.} This process represents the colloidal analogue of the Bergeron process~\cite{glickman2000glossary}, which explains the formation of ice crystals in mixed phase clouds and is at the origin of rain.
 Our finding gives us full experimental access to the kinetic pathway of this poorly understood phenomenon important for climate science.
@@
 
-\textcolor{red}{
 A direct analysis of the colloidal positions at long times confirms that indeed the new structure obtained at
@@
 We directly compare the two structures in Fig.~\ref{fig:network}, where Gaussian filtering is used to depict the network structure as a continuous field.
-Panel a) presents the results for high
-polymer concentration gel, $c_p=1.07$ mg/g; panel b) presents the results for the low polymer concentration sample, $c_p=0.38$ mg/g.
-The comparison shows that the new arrest mechanism produces a network structure with bigger pores and thicker strands (panel b) compared
+Panel a presents the results for high
+polymer concentration gel, $c_p=1.07$ mg/g; panel b presents the results for the low polymer concentration sample, $c_p=0.38$ mg/g.
+The comparison shows that the new arrest mechanism produces a network structure with bigger pores, \textcolor{blue}{smoother surfaces}, and thicker strands (panel b) compared
 to the colloidal gel sample (panel a). The bottom row shows the individual particle positions for slabs with a thickness of 5 particle diameters.
@@
 involve the formation of a crystal-gel network. In order to explain the morphological differences between colloidal gels and crystal-gels,
-we now consider the kinetics of formation in our samples.}
+we now consider the kinetics of formation in our samples.
 
@@
 The first process dominates the initial stages of crystallization, while the second one becomes effective once
-crystalline nuclei have outgrown the liquid regions in which they form. Process (2) has been completely overlooked so far, although it plays a crucial role 
+crystalline nuclei have outgrown the liquid regions in which they form.
+\textcolor{blue}{Process (2), or the kinetic path via the gas phase (liquid$\rightarrow$gas$\rightarrow$crystal),} has been completely overlooked so far, although it plays a crucial role
 not only in the kinetic pathway, but also in determining the morphology of the porous crystal, as we discussed above. We note that this is the crucial step  
@@
 without specific directional interactions.
+
 Usually, monoatomic systems are very poor glass-formers and thus have not been expected to form gels. 
-However, our mechanism provides a novel kinetic pathway to form network structures made of crystals.  
-So we believe that crystal gels are an important class of heterogeneous non-ergodic states in nature 
+However, our mechanism provides a novel kinetic pathway to spontaneously form network or porous structures made of crystals.
+\textcolor{blue}{The fact that crystals outgrow from the liquid network means that well-ordered crystal planes appear on the surface of the porous structure (see Fig.
+2b), which is crucial for catalytic and sensing applications.} So we believe that crystal gels are an important class of heterogeneous non-ergodic states in nature
 and industrial applications, although they have not attracted much attention so far.  
-For example, nano-porous crystals of noble metals such as Au have special functions, which are relevant to catalytic, optical, sensing, and filtration applications 
-\cite{ding2004nanoporous, ding2009nanoporous, wittstock2010nanoporous}. Usually such nano-porous materials are formed via at least two steps: for example, phase separation and dealloying of a mixture
+For example, nano-porous crystals of noble metals such as Au have special functions \textcolor{blue}{associated with ultra-high interfacial area and connectivity of pores}, which are relevant to catalytic, optical, sensing, super-capacitor, and filtration applications
+\cite{ding2004nanoporous, ding2009nanoporous, wittstock2010nanoporous,fujita2012atomic}.
+Usually such nano-porous materials are formed via at least two steps: for example, phase separation and dealloying of a mixture
 \cite{erlebacher2001evolution}. Unlike such a method, our novel mechanism allows us to spontaneously form sponge-like nano-porous crystals \emph{in a single step}, 
@@
  \centering
- \includegraphics[width=12cm]{new_fig2}
+ \includegraphics[width=14cm]{figure2_n}
  % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
-\caption{{\bf Percolated networks: a) gel; b) crystal-gel.} The gel state is obtained from the state
+\caption{{\bf Percolated networks: {\bf a:} gel; {\bf b:} crystal-gel.} The gel state is obtained from the state
 point $\phi\approx 0.30$ and $c_p=1.07$ mg/g, while the crystal-gel is obtained at the same $\phi$ but
@@
 been obtained by applying a gaussian filter to the position of the particles (of width equal to the inter-particle distance) and
-plotting the surface which bounds the highest 20\% values of the field. The lower row depicts slabs of 5 particle diameter thickness,
-with particles colored according to the following code: red (fcc), green (hcp), white (crystal surface particles), cyan (liquid particles), magenta (gas particles).} 
+plotting the surface which bounds the highest 20\% values of the field. \textcolor{blue}{We can clearly see not only the percolated nature of both network, 
+but also the difference in the smoothness of the network surface between amorphous and crystal gels.} 
+The lower row depicts slabs of 5 particle diameter thickness,
+with particles colored according to the following code: red (fcc), green (hcp), white (crystal surface particles), cyan (liquid particles), magenta (gas particles). 
+\textcolor{blue}{We note that there remain a very few liquid particles and the network is almost crystalline.}} 
 \label{fig:network}

### 2015-11-20 crystal-gel/nature/crystal_gel2.tex
Small adjustements.

@@
 We show that crystallization leads to a network structure different from the original phase separation pattern.
-Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface, \textcolor{blue}{i.e.
-desublimation,} driving evaporation of nearby liquid.
-\textcolor{blue}{The latter processes via the gas phase have been overlooked so far.} This process represents the colloidal analogue of the Bergeron process~\cite{glickman2000glossary}, which explains the formation of ice crystals in mixed phase clouds and is at the origin of rain.
+Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface, i.e.
+desublimation, driving evaporation of nearby liquid.
+The latter processes via the gas phase has been overlooked so far.
+This process represents the colloidal analogue of the Bergeron process~\cite{glickman2000glossary}, which explains the formation of ice crystals in mixed phase clouds and is at the origin of rain.
 Our finding gives us full experimental access to the kinetic pathway of this poorly understood phenomenon important for climate science.
@@
 polymer concentration gel, $c_p=1.07$ mg/g; panel b presents the results for the low polymer concentration sample, $c_p=0.38$ mg/g.
-The comparison shows that the new arrest mechanism produces a network structure with bigger pores, \textcolor{blue}{smoother surfaces}, and thicker strands (panel b) compared
+The comparison shows that the new arrest mechanism produces a network structure with bigger pores, smoother surfaces, and thicker strands (panel b) compared
 to the colloidal gel sample (panel a). The bottom row shows the individual particle positions for slabs with a thickness of 5 particle diameters.
@@
 crystalline nuclei have outgrown the liquid regions in which they form.
-\textcolor{blue}{Process (2), or the kinetic path via the gas phase (liquid$\rightarrow$gas$\rightarrow$crystal),} has been completely overlooked so far, although it plays a crucial role
+Process (2), or the kinetic path via the gas phase (liquid$\rightarrow$gas$\rightarrow$crystal), has been completely overlooked so far, although it plays a crucial role
 not only in the kinetic pathway, but also in determining the morphology of the porous crystal, as we discussed above. We note that this is the crucial step  
@@
 However, our mechanism provides a novel kinetic pathway to spontaneously form network or porous structures made of crystals.
-\textcolor{blue}{The fact that crystals outgrow from the liquid network means that well-ordered crystal planes appear on the surface of the porous structure (see Fig.
-2b), which is crucial for catalytic and sensing applications.} So we believe that crystal gels are an important class of heterogeneous non-ergodic states in nature
+The fact that crystals outgrow from the liquid network means that well-ordered crystal planes appear on the surface of the porous structure (see Fig.
+2b), which is crucial for catalytic and sensing applications.
+So we believe that crystal gels are an important class of heterogeneous non-ergodic states in nature
 and industrial applications, although they have not attracted much attention so far.  
-For example, nano-porous crystals of noble metals such as Au have special functions \textcolor{blue}{associated with ultra-high interfacial area and connectivity of pores}, which are relevant to catalytic, optical, sensing, super-capacitor, and filtration applications
+For example, nano-porous crystals of noble metals such as Au have special functions associated with ultra-high interfacial area and connectivity of pores, which are relevant to catalytic, optical, sensing, super-capacitor, and filtration applications
 \cite{ding2004nanoporous, ding2009nanoporous, wittstock2010nanoporous,fujita2012atomic}.

### 2015-11-20 crystal-gel/nature/crystal_gel2.tex

@@
 been obtained by applying a gaussian filter to the position of the particles (of width equal to the inter-particle distance) and
-plotting the surface which bounds the highest 20\% values of the field. \textcolor{blue}{We can clearly see not only the percolated nature of both network, 
-but also the difference in the smoothness of the network surface between amorphous and crystal gels.} 
+plotting the surface which bounds the highest 20\% values of the field. We can clearly see not only the percolated nature of both network, 
+but also the difference in the smoothness of the network surface between amorphous and crystal gels. 
 The lower row depicts slabs of 5 particle diameter thickness,
 with particles colored according to the following code: red (fcc), green (hcp), white (crystal surface particles), cyan (liquid particles), magenta (gas particles). 
-\textcolor{blue}{We note that there remain a very few liquid particles and the network is almost crystalline.}} 
+We note that very few liquid particles remain and the network is almost crystalline.} 
 \label{fig:network}

### 2016-02-05 phasediag.py
2to3 and tie lines for crystal gel

@@ -3,6 +3,10 @@ from colloids.phase import *
 from os.path import isfile
 from scipy.constants import N_A
 
+def xp2th(cp, phi, qR=0.1, cpov=1.):
+    """Converts experimental polymer concentration in osmotic pressure"""
+    return y2piv(cp/cpov/alpha(vf2f(phi), qR2q(qR)), qR)
+
 #radius of gyration in nm
 R = 82.
 #Molecular weight (Dalton or g/mol)
@@ -15,7 +19,7 @@ d = 1.24
 qR = 2*R/sigma
 cpov = 1000 * (3 * Mw)/(4 * np.pi * (R*1e-9)**3 * N_A * d*1e6)
 q = qR2q(qR)
-print 'qR = %0.3f, q = %0.3f, Cov = %0.3f mg/g'%(qR, q, cpov)
+print('qR = %0.3f, q = %0.3f, Cov = %0.3f mg/g'%(qR, q, cpov))
 
 #experimental points
 phixpG = np.array([0.069, 0.128, 0.135, 0.126, 0.333, 0.313, 0.294, 0.305])
@@ -25,7 +29,7 @@ cpxpF = np.array([0.25,0.48,0.57, 0.76, 0.28, 0.34])
 
 #load or compute theoretical phase diagram
 fc, pivc = Liu().critical_point(q)
-print 'At critical point colloid volume fraction is %0.3f and osmotic insertion work is %0.3f kT'%(f2vf(fc), pivc)
+print('At critical point colloid volume fraction is %0.3f and osmotic insertion work is %0.3f kT'%(f2vf(fc), pivc))
 if isfile('phasediag.txt'):
     GL = np.loadtxt('phasediag.txt')
 else:
@@ -58,10 +62,25 @@ for f,l,ph in zip(GL.T[1:], ['--']*2+['-']*2, ['bg', 'bl', 'sg', 'sl']):
         header='phi\tcp',
         )
 #draw tie lines
-for piv, fG, fL in GL[[50,60,70,80,85,88]+range(90,98),:3]:
+for piv, fG, fL in GL[[50,60,70,80,85,88]+list(range(90,98)),:3]:
     cpG = piv2y(piv, qR) * alpha(fG, q) * cpov
     cpL = piv2y(piv, qR) * alpha(fL, q) * cpov
     plot(f2vf(np.array([fG,fL])), [cpG, cpL], '-', color=(0.5,0.5,0.5))
[…]

### 2016-02-08 nature/crystal_gel2.tex

@@
 \author{Hajime Tanaka}
-\affiliation{Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}
+\affiliation{Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan \footnote{e-mail: tanaka@iis.u-tokyo.ac.jp}}
 
@@
 {\bf
-Viscoelastic phase separation of colloidal suspensions can be interrupted either by the glass transition, with the formation of colloidal gels~\cite{poon2002,zaccarelli2007,piazza1994phase,verhaegh1997transient,tanaka1999colloid,lu2008gelation}, or by crystallization, which results in the formation of a crystalline network~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways,sabin2012}.
-But while the former process has been studied in detail, the latter has never been observed experimentally at a single-particle level and it is thus lacking a microscopic understanding.
+Viscoelastic phase separation of colloidal suspensions can be interrupted to form gels either by 
+%the glass transition, with the formation of colloidal gels~\cite{poon2002,zaccarelli2007,piazza1994phase,verhaegh1997transient,tanaka1999colloid,lu2008gelation}, or by crystallization, which results in the formation of a crystalline network~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways,sabin2012}.
+glass transition~\cite{poon2002,zaccarelli2007,piazza1994phase,verhaegh1997transient,tanaka1999colloid,lu2008gelation} or by crystallization~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways,sabin2012}.
+%But while the former process has been studied in detail, the latter has never been observed experimentally at a single-particle level and it is thus lacking a microscopic understanding. 
 With a new confocal microscopy protocol, we unveil a novel crystallization pathway to sponge-like porous structures in a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation.
 We show that crystallization leads to a network structure different from the original phase separation pattern.
-Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface, driving evaporation of nearby liquid.
-This process represents the colloidal analogue of the Bergeron process~\cite{glickman2000glossary}, which explains the formation of ice crystals in mixed phase clouds and is at the origin of rain.
+Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface, i.e.
+desublimation, driving evaporation of nearby liquid.
+%The latter processes via the gas phase has been overlooked so far.
+%This process represents the colloidal analogue of the Bergeron process~\cite{glickman2000glossary}, which explains the formation of ice crystals in mixed phase clouds and is at the origin of rain.
 Our finding gives us full experimental access to the kinetic pathway of this poorly understood phenomenon important for climate science.
-The key is complex dynamic interplay involving the three phases upon simultaneous occurrence of two non-equilibrium phenomena, phase separation and crystallization.
+%The key is complex dynamic interplay involving the three phases upon simultaneous occurrence of two non-equilibrium phenomena, phase separation and crystallization. 
 We argue that similar crystal-gel states can be formed as the result of the gas-liquid phase separation of any crystallizable components, such as monoatomic and molecular systems, if dynamics of the liquid phase is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
-This novel mechanism not only realizes single-step spontaneous formation of nano-porous crystals of metals and semiconductors without dealloying \cite{erlebacher2001evolution}, which may be important for catalytic, optical, sensing, and filtration applications \cite{ding2004nanoporous,ding2009nanoporous, wittstock2010nanoporous}, but also sheds new light on crystal network formation in more complex systems such as magma \cite{philpotts1998role}, biominerals \cite{rousseau2005multiscale}, and foods \cite{deman1987fat}.}
+This mechanism will provide a novel single-step pathway to form nano-porous crystals of metals and semiconductors without dealloying \cite{erlebacher2001evolution}, which may be important for catalytic, optical, sensing, and filtration applications \cite{ding2004nanoporous,ding2009nanoporous, wittstock2010nanoporous}.
+%, but also sheds new light on crystal network formation in more complex systems such as magma \cite{philpotts1998role}, biominerals \cite{rousseau2005multiscale}, and foods \cite{deman1987fat}.}
 \end{abstract}
@@
 We directly compare the two structures in Fig.~\ref{fig:network}, where Gaussian filtering is used to depict the network structure as a continuous field.
-Panel a) presents the results for high
-polymer concentration gel, $c_p=1.07$ mg/g; panel b) presents the results for the low polymer concentration sample, $c_p=0.38$ mg/g.
-The comparison shows that the new arrest mechanism produces a network structure with bigger pores and thicker strands (panel b) compared
+Panel a presents the results for high
+polymer concentration gel, $c_p=1.07$ mg/g; panel b presents the results for the low polymer concentration sample, $c_p=0.38$ mg/g.
+The comparison shows that the new arrest mechanism produces a network structure with bigger pores, smoother surfaces, and thicker strands (panel b) compared
 to the colloidal gel sample (panel a). The bottom row shows the individual particle positions for slabs with a thickness of 5 particle diameters.
@@
 The first process dominates the initial stages of crystallization, while the second one becomes effective once
-crystalline nuclei have outgrown the liquid regions in which they form. Process (2) has been completely overlooked so far, although it plays a crucial role 
+crystalline nuclei have outgrown the liquid regions in which they form.
+Process (2), or the kinetic path via the gas phase (liquid$\rightarrow$gas$\rightarrow$crystal), has been completely overlooked so far, although it plays a crucial role
 not only in the kinetic pathway, but also in determining the morphology of the porous crystal, as we discussed above. We note that this is the crucial step  
@@
 without specific directional interactions.
+
 Usually, monoatomic systems are very poor glass-formers and thus have not been expected to form gels. 
-However, our mechanism provides a novel kinetic pathway to form network structures made of crystals.  
+However, our mechanism provides a novel kinetic pathway to spontaneously form network or porous structures made of crystals.
+The fact that crystals outgrow from the liquid network means that well-ordered crystal planes appear on the surface of the porous structure (see Fig.
+2b), which is crucial for catalytic and sensing applications.
 So we believe that crystal gels are an important class of heterogeneous non-ergodic states in nature
 and industrial applications, although they have not attracted much attention so far.  
-For example, nano-porous crystals of noble metals such as Au have special functions, which are relevant to catalytic, optical, sensing, and filtration applications 
-\cite{ding2004nanoporous, ding2009nanoporous, wittstock2010nanoporous}. Usually such nano-porous materials are formed via at least two steps: for example, phase separation and dealloying of a mixture
+For example, nano-porous crystals of noble metals such as Au have special functions associated with ultra-high interfacial area and connectivity of pores, which are relevant to catalytic, optical, sensing, super-capacitor, and filtration applications
+\cite{ding2004nanoporous, ding2009nanoporous, wittstock2010nanoporous,fujita2012atomic}.
+Usually such nano-porous materials are formed via at least two steps: for example, phase separation and dealloying of a mixture
 \cite{erlebacher2001evolution}. Unlike such a method, our novel mechanism allows us to spontaneously form sponge-like nano-porous crystals \emph{in a single step}, 
@@
 \noindent
-{\bf Acknowledgements.} This study was partly supported by Grants-in-Aid for Scientific Research (S) (Grand No.
+{\bf Acknowledgements} 
+This study was partly supported by Grants-in-Aid for Scientific Research (S) (Grand No.
 21224011) and Specially Promoted Research (Grand No.
@@
 
-%\vspace{0.3cm}
+\vspace{0.3cm}
 \noindent
-{\bf Author Contributions.} 
+{\bf Author Contributions} 
 H.Ts. and J.R. contributed equally to this work. 
@@
 
-%\vspace{0.3cm}
+\vspace{0.3cm}
 \noindent
-{\bf Author Information.} 
+{\bf Additional information} 
 %Reprints and permissions information is available at www.nature.com/reprints.
-Correspondence and requests for materials should be addressed to H.T. ~(tanaka@iis.u-tokyo.ac.jp).
+Correspondence and requests for materials should be addressed to H.T.
+
+\vspace{0.3cm}
+\noindent
+{\bf Competing financial interests}
+The authors declare no competing financial interests
 
@@
 been obtained by applying a gaussian filter to the position of the particles (of width equal to the inter-particle distance) and
-plotting the surface which bounds the highest 20\% values of the field. The lower row depicts slabs of 5 particle diameter thickness,
-with particles colored according to the following code: red (fcc), green (hcp), white (crystal surface particles), cyan (liquid particles), magenta (gas particles).} 
+plotting the surface which bounds the highest 20\% values of the field. We can clearly see not only the percolated nature of both network, 
+but also the difference in the smoothness of the network surface between amorphous and crystal gels. 
+The lower row depicts slabs of 5 particle diameter thickness,
+with particles colored according to the following code: red (fcc), green (hcp), white (crystal surface particles), cyan (liquid particles), magenta (gas particles). 
+We note that very few liquid particles remain and the network is almost crystalline.} 
 \label{fig:network}

### 2016-02-29 nature/crystal_gel2.tex
John. First major changes following the Referees comments. Figures 2 and 4 have been updated, and the text changed accordingly.

@@
-\documentclass[preprint,amsmath,amssymb,superscriptaddress]{revtex4}
+\documentclass[preprint,amsmath,amssymb,superscriptaddress]{revtex4-1}
 \usepackage{graphicx,amsmath}
@@
 %, but also sheds new light on crystal network formation in more complex systems such as magma \cite{philpotts1998role}, biominerals \cite{rousseau2005multiscale}, and foods \cite{deman1987fat}.}
+}
 \end{abstract}
@@
 of crystalline droplets.
-These droplets originate from crystal seeds that form inside the liquid branches of the spinodal aggregate of network morphology, but then grow by direct condensation from the gas phase.
+These droplets originate from crystal seeds that form inside the liquid branches of the spinodal aggregate of network morphology, and then grow by direct condensation from the gas phase.
 The process is analogous to the Bergeron process of ice formation in mixed phase clouds, where ice droplets grow at the
@@
 This is speculated to be the primary mechanism for the formation of raindrops. 
-The mechanism found here is in contrast to the standard crystal-gel scenario, where crystals nucleate and grow inside the dense branches of the gel, becoming dynamically arrested.
-We also observe that increasing the strength of the interparticle attraction
+Differently from the standard crystal-gel scenario, where crystals only nucleate and grow inside the dense branches of the gel, becoming dynamically arrested, we show that the kinetic
+pathways involving the gas phase (predominanlty the Bergeron process, but also Ostwald ripening) play a fundamental role. We also observe that increasing the strength of the interparticle attraction
 (by increasing the polymer concentration), instead of forming crystal gels, the system undergoes the classical gelation process, in which
@@
 
-Now we demonstrate that the volume growth is due to direct condensation of the gas phase on the crystalline seeds.
+\textcolor{red}{Next we analyze in detail the crystallization trajectory, distinguishing between the different process responsible for the
+crystallization of the phase separated network.
 Supplementary Movie 2 shows the confocal microscopy video for the crystallizing sample at $\phi\approx 0.30$ and $c_p=0.38$ mg/g,
@@
 From the trajectory, we extract slabs of 10 particle diameter thickness for configurations at different times, shown in Fig.~\ref{fig:transitions}a. 
-Particles are colored according to their phases (gas, liquid, crystal, and crystal surface). The figure shows the first nucleation 
-events inside the liquid network (left panel); then these crystalline nuclei grow directly from the gas phase (middle panel), and finally
-the different nuclei coalesce (right panel).
+Particles are colored according to their phases (gas, liquid, and crystal). For the classification criteria see the \emph{Methods} section. The figure shows the first nucleation 
+events inside the liquid network (left panel). At the same time liquid droples that have not crystallized start evaporating, and the gas phase contributes to the nuclei growth (middle panel). Finally
+the different nuclei coalesce (right panel). Note that the network pattern changes between $t=2055\,\tau_B$ (left panel) and $t=5577,\tau_B$ (middle panel): not all liquid
+branches of the network crystallize due to evaporation. The crystal-gel structure instead persists, as seen by comparing $t=5577,\tau_B$ (middle panel) and $t=12098,\tau_B$ (right panel).}
 
-%In the next section we show evidence for the fact that the volume growth is due to direct condensation of the gas phase on the crystalline seeds.
-%First we show in Fig.~\ref{fig:transitions}a slabs of 10 particle diameter thickness for configurations at $\phi\approx 0.30$ and $c_p=0.38$ at different times.
-%Particles are coloured according to their phase as follows. Green: hcp; red: fcc; blue: bcc; white: liquid or gas. The figure shows the first nucleation
-%events inside the liquid network (left panel); then these crystalline nuclei grow directly from the gas phase (middle panel), and finally
-%the different nuclei coalesce (right panel).
-
+\textcolor{red}{
 Next we show in Fig.~\ref{fig:transitions}b the fraction of particles in each different phase for the crystallization trajectory at
-$\phi\approx 0.30$ and $c_p=0.38$ mg/g, after the liquid-gas phase separation has occurred. The process of crystallization is characterized 
+$\phi\approx 0.30$ and $c_p=0.38$ mg/g, after the liquid-gas phase separation has occurred.
+For the classification criteria see \emph{Methods}.
+The process of crystallization is characterized 
 by a steep decrease in liquid particles, as they transform in small crystalline nuclei inside the liquid domains. This decrease
@@
 to the gas phase due to the higher vapour pressure of the liquid phase compared to the crystalline phase. After the onset
-of a steady-state gas population, the crystallization proceeds by adding surface particles from the gas phase to the small crystallites. To confirm this process we
-examine in Fig.~\ref{fig:transitions}c the transfer matrix probabilities $P_{\alpha\beta}$, where $\alpha$ and $\beta$ are each of the 
-phases, gas, liquid and crystal (which includes hcp, fcc and surface particles). Liquid particles undergo two processes: on one
-hand they transform to crystallites growing inside the liquid branches of the network, and also increasingly transform to gas particles.
-Gas particles first undergo a transition to liquid particles as the final stage of the liquid-gas phase separation, but once
-crystalline nuclei outgrow the liquid pockets in which they are formed, the probability for the transition from gas to liquid
-abruptly decreases. Instead, gas particles start depositing on the crystal, and at later stages the pace of the gas-to-crystal transformation (sublimation)
-is larger than the one of the liquid-to-crystal one.
+of a steady-state gas population, there are three growth channels for the crystal. Direct crystallization is the process by which
+crystals grow by incorporating nearby liquid particles. In the Bergeron process, liquid droplets first evaporate and the resulting gas phase
+contributes to the growth of crystalline regions. Ostwald ripening is instead the process by which small crystallites sublimate, and colloidal
+particles are transfered to larger nuclei.
+In order to assess the relative importance of these three growth channels, we analyze the trajectory, and, for every time frame, assign each particle a state between gas, liquid and crystal.
+In order to minimize short
+term fluctuations, the state of each colloidal particle is time averaged for $50\,\tau_B$. We then measure the fraction of trajectories with different transition histories.
+Direct crystallization accounts for 54\% of particles trajectories in which
+gas or liquid particles transition to the crystal state without liquid evaporation or crystal de-sublimation. The Bergeron process accounts for 35\% of particle trajectories in which liquid particles
+transition to the gas state before crystallizing. Ostwald ripening accounts for 11\% of particle trajectories in which crystal particle transition to the gas state before returning
+to the crystal state. The different crystallization channels are depicted in the diagram of Fig.~\ref{fig:transitions}c.}
 
-We can thus conclude that the crystallization proceeds via two distinct mechanisms:
-(1) freezing of the fluid; (2) evaporation of the fluid and condensation of the vapour on the crystals, i.e., a coupled process of evaporation and sublimation.
-The first process dominates the initial stages of crystallization, while the second one becomes effective once
-crystalline nuclei have outgrown the liquid regions in which they form.
-Process (2), or the kinetic path via the gas phase (liquid$\rightarrow$gas$\rightarrow$crystal), has been completely overlooked so far, although it plays a crucial role
-not only in the kinetic pathway, but also in determining the morphology of the porous crystal, as we discussed above. We note that this is the crucial step  
-in the Bergeron process~\cite{glickman2000glossary,morrison2012resilience} (see below).
+\textcolor{red}{
+While the direct freezing of the fluid represents the major contribution to the nuclei growth, the kinetic path via the gas phase (gas$\rightarrow$crystal), also plays a crucial role,
+especially in determining the morphology of the porous crystal, as we discussed above. In this context the Bergeron process~\cite{glickman2000glossary,morrison2012resilience}
+plays a considerable more important role than Ostwald ripening.}
 % 
@@
 
-The process we observed is akin to the Bergeron process which is the primary mechanism for the formation of rain drops in clouds~\cite{glickman2000glossary,morrison2012resilience}.
+The Bergeron process is also the primary mechanism for the formation of rain drops in clouds~\cite{glickman2000glossary,morrison2012resilience}.
 In clouds there is a mixture of ice crystals and supercooled water. The vapour phase is in coexistence with the liquid phase, but is supersaturated
@@
 
-This novel mechanism creating ``crystal gels'' may be generic to many other 
+The process of formation of ``crystal gels'' may be generic to many other 
 systems. The requirements are (i) the presence of gas-liquid phase separation below the melting point of a crystal, (ii) 
@@
  \centering
- \includegraphics[width=14cm]{figure2_n}
+ \includegraphics[width=12cm]{fig2}
  % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
-\caption{{\bf Percolated networks: {\bf a:} gel; {\bf b:} crystal-gel.} The gel state is obtained from the state
+\caption{\textcolor{red}{{\bf Percolated networks: {\bf a:} gel; {\bf b:} crystal-gel.} The gel state is obtained from the state
 point $\phi\approx 0.30$ and $c_p=1.07$ mg/g, while the crystal-gel is obtained at the same $\phi$ but
-at lower polymer concentration, $c_p=0.38$ mg/g. The top row shows depicts the networks as a continuous field that has
+at lower polymer concentration, $c_p=0.38$ mg/g. On the left, the network is represented as a continuous field that has
 been obtained by applying a gaussian filter to the position of the particles (of width equal to the inter-particle distance) and
-plotting the surface which bounds the highest 20\% values of the field. We can clearly see not only the percolated nature of both network, 
+plotting the surface which bounds the highest 20\% values of the field. We can clearly see not only the percolated nature of both networks, 
 but also the difference in the smoothness of the network surface between amorphous and crystal gels. 
-The lower row depicts slabs of 5 particle diameter thickness,
-with particles colored according to the following code: red (fcc), green (hcp), white (crystal surface particles), cyan (liquid particles), magenta (gas particles). 
-We note that very few liquid particles remain and the network is almost crystalline.} 
+On the right, slabs of 5 particle diameter thickness are plotted,
+with particles colored with the same color scheme as in Fig.~\ref{fig:transitions}: red (crystal), green (gas particles), white (liquid particles). 
+We note that very few liquid particles remain and the network is almost crystalline.}} 
 \label{fig:network}
@@
 
-\begin{figure*}
+\begin{figure*}[!t]
  \centering
- \includegraphics{crystal}
+ \includegraphics[width=12cm]{fig4}
  % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735
- \caption{
+ \caption{\textcolor{red}{
 {\bf Crystal gel formation.}  
@@
 Particles are drawn to scale and colored according to their phase.
-Dark red: crystal; orange: surface of crystal; yellow: liquid; blue: gas.
+Red: crystal; white: liquid; green: gas.
 {\bf b,} Fraction of particles for each phase. Gray vertical lines indicate the times shown in panel a. 
-{\bf c,} Transition probabilities for the same process as in panel b. Here crystal includes surface particles.}
+{\bf c,} Each particle during its trajectory changes between three states: gas, liquid and crystal. Direct crystallization accounts for 54\% of particles trajectories in which
+gas or liquid particles transition to the crystal state without liquid evaporation or crystal de-sublimation. The Bergeron process accounts for 35\% of particle trajectories in which liquid particles
+transition to the gas state before crystallizing. Ostwald ripening accounts for 11\% of particle trajectories in which crystal particle transition to the gas state before returning
+to the crystal state.}}
 \label{fig:transitions}

### 2016-03-01 nature/crystal_gel2.tex

@@
 
-\textcolor{red}{Next we analyze in detail the crystallization trajectory, distinguishing between the different process responsible for the
+\textcolor{red}{Next we analyse in detail the crystallization trajectory, distinguishing between the different process responsible for the
 crystallization of the phase separated network.
@@
 From the trajectory, we extract slabs of 10 particle diameter thickness for configurations at different times, shown in Fig.~\ref{fig:transitions}a. 
-Particles are colored according to their phases (gas, liquid, and crystal). For the classification criteria see the \emph{Methods} section. The figure shows the first nucleation 
-events inside the liquid network (left panel). At the same time liquid droples that have not crystallized start evaporating, and the gas phase contributes to the nuclei growth (middle panel). Finally
-the different nuclei coalesce (right panel). Note that the network pattern changes between $t=2055\,\tau_B$ (left panel) and $t=5577,\tau_B$ (middle panel): not all liquid
+Particles are coloured according to their phases (gas, liquid, and crystal). For the classification criteria see the \emph{Methods} section. The figure shows the first nucleation 
+events inside the liquid network (left panel).
+At the same time liquid regions that have not crystallized start evaporating, and the gas phase contributes to the growth of nuclei (middle panel).
+Finally the different nuclei coalesce (right panel).
+Note that the network pattern changes between $t=2055\,\tau_B$ (left panel) and $t=5577,\tau_B$ (middle panel): not all liquid
 branches of the network crystallize due to evaporation. The crystal-gel structure instead persists, as seen by comparing $t=5577,\tau_B$ (middle panel) and $t=12098,\tau_B$ (right panel).}
@@
 The process of crystallization is characterized 
-by a steep decrease in liquid particles, as they transform in small crystalline nuclei inside the liquid domains. This decrease
+by a steep decrease in liquid particles, as they transform into small crystalline nuclei inside the liquid domains. This decrease
 is then accompanied by an increase in the fraction of gas particles: as the first crystals start to reach the gas phase, liquid particles evaporate 
@@
 contributes to the growth of crystalline regions. Ostwald ripening is instead the process by which small crystallites sublimate, and colloidal
-particles are transfered to larger nuclei.
-In order to assess the relative importance of these three growth channels, we analyze the trajectory, and, for every time frame, assign each particle a state between gas, liquid and crystal.
-In order to minimize short
-term fluctuations, the state of each colloidal particle is time averaged for $50\,\tau_B$. We then measure the fraction of trajectories with different transition histories.
+particles are transferred to larger nuclei.
+In order to assess the relative importance of these three growth channels, we analyse the trajectory, and, for every time frame, assign each particle a state between gas, liquid and crystal.
+In order to minimize short-term fluctuations, the state of each colloidal particle is time averaged for $50\,\tau_B$.
+We then measure the fraction of trajectories with different transition histories.
 Direct crystallization accounts for 54\% of particles trajectories in which
 gas or liquid particles transition to the crystal state without liquid evaporation or crystal de-sublimation. The Bergeron process accounts for 35\% of particle trajectories in which liquid particles
-transition to the gas state before crystallizing. Ostwald ripening accounts for 11\% of particle trajectories in which crystal particle transition to the gas state before returning
+transition to the gas state before crystallizing. Ostwald ripening accounts for 11\% of particle trajectories in which crystal particles transition to the gas state before returning
 to the crystal state. The different crystallization channels are depicted in the diagram of Fig.~\ref{fig:transitions}c.}
@@
 especially in determining the morphology of the porous crystal, as we discussed above. In this context the Bergeron process~\cite{glickman2000glossary,morrison2012resilience}
-plays a considerable more important role than Ostwald ripening.}
+plays a considerably more important role than Ostwald ripening.}
 % 
@@
  % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
-\caption{\textcolor{red}{{\bf Percolated networks: {\bf a:} gel; {\bf b:} crystal-gel.} The gel state is obtained from the state
+\caption{\textcolor{red}{{\bf Percolated network structures.} {\bf a:} gel; {\bf b:} crystal-gel. The gel state is obtained from the state
 point $\phi\approx 0.30$ and $c_p=1.07$ mg/g, while the crystal-gel is obtained at the same $\phi$ but
@@
 been obtained by applying a gaussian filter to the position of the particles (of width equal to the inter-particle distance) and
-plotting the surface which bounds the highest 20\% values of the field. We can clearly see not only the percolated nature of both networks, 
+plotting the surface which bounds the highest 20 \% values of the field. We can clearly see not only the percolated nature of both networks, 
 but also the difference in the smoothness of the network surface between amorphous and crystal gels. 
 On the right, slabs of 5 particle diameter thickness are plotted,
-with particles colored with the same color scheme as in Fig.~\ref{fig:transitions}: red (crystal), green (gas particles), white (liquid particles). 
+with particles coloured with the same colour scheme as in Fig.~\ref{fig:transitions}: red (crystal), green (gas particles), white (liquid particles). 
 We note that very few liquid particles remain and the network is almost crystalline.}} 
@@
  \centering
- \includegraphics[width=12cm]{fig4}
+ \includegraphics[width=14cm]{fig4}
  % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735
- \caption{\textcolor{red}{
-{\bf Crystal gel formation.}  
+ \caption{\textcolor{red}{{\bf Crystal gel formation.}  
 {\bf a,} Reconstructions from confocal coordinates at $\phi\approx 0.30$ and $c_p=0.38$ mg/g.
 Depth of view is $5\sigma$.
-Particles are drawn to scale and colored according to their phase.
+Particles are drawn to scale and coloured according to their phase.
 Red: crystal; white: liquid; green: gas.

### 2016-03-05 nature/response.tex
Expanded supplementary and response letter.

\documentclass[11pt]{article}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage{color}
\usepackage{setspace}

\begin{document}

\doublespacing

\doublespacing

\noindent
Dear Editor,
\vskip 0.3cm
we would like to thank you for your comments to our manuscript ``A novel route to the spontaneous formation of porous crystals via viscoelastic phase separation'' (NM15113574),
and for providing us with three Referee reports.
We were happy to see that all Referees appreciated our work, and at least two of them suggested
resubmitting it after some technical clarifications. Referee 1 describes our results as ``very nice and beautiful'' but does not recommend our work
based on the claim that it is not novel. He refers in particular to a series of works by the Edinburgh group in which the kinetics of phase
transitions is classified according to the free energy landscape. These work are by themselves very important and relevant, but contrary
to the Referee claim, do not contain a description or results on the crystallization channel we describe in our work. A crucial difference is
the fact that those previous works
study the case in which gas-liquid phase separation is stable against crystellization, while our crystallization scenario depends crucially
on the metastability of liquid-gas phase separation against crystallization. Both Referees 2 and 3 have appreciated this fact.

To address the technical questions of Referee 2 and 3 we have conducted novel experiments, that clearly demonstrate the reproducibility
of our work. In particular, we believe that the concerns of Referee 3 are simply due to the fact that the two movies in Supplementary Informations
are shot with different framerates. This gives the impression of abundance of gas phase, that probably led the Referee into thinking that the Ostwald
ripening process is dominating. We have now made this issue more clear, and are confident that it will answer the Referee's concern fully.
We also broadened our analysis, to include a direct study of the trajectory history to determine the different crystallization
channels. As a result, new sections have been added to the Supplementary Informations, providing for a in-depth demonstration of the generality of our results.

We hope that you will find the revised version of the manuscript to be fully addressing the Referee questions, and
that you will find the new manuscript to be suitable for publication in \emph{Nature Materials}.

\vskip 0.8cm

\noindent
Yours Sincerely,
\vskip 0.3cm
\indent Hideyo Tsurusawa\\
\indent John Russo\\
\indent Mathieu Leocmach\\
\indent Hajime Tanaka

\clearpage

\noindent
\begin{Large}
{\bf Response to Referee \#1}
\end{Large}

\vspace{1em}

\singlespacing

{\bf
This manuscript describes experimental work on a colloid-polymer mixture that undergoes viscoelastic phase separation thereby forming a spinodal structure, which is interrupted by crystallization. The authors argue that this mechanism can be used to form nano-porous crystals, which has potential applications. However, this phenomenon has been shown in extensive previous studies, see Ref. [7-10], and is therefore not novel. Another key result of the paper is that the authors demonstrate that the crystalline branches of the spinodal network can also grow by the attachment of "gas"-like particles onto the surfaces. Even this message is not completely new as the Edinburgh group (Poon, Evans, Renth, Cates) has made extensive studies on the precise kinetic pathways of the phase transformations in colloid-polymer mixtures and explained this by means of free-energy landscapes. An extensive body of work that the authors apparently missed. Although the experimental study and analysis is very nice and beautiful, the referee hesitate to recommend the manuscript for publication in Nature Materials due to a lack of novelty (in terms of material and fundamental insight) of the present work.
}

\bigskip
\doublespacing

First We would like to thank the Referee for her/his careful reading of our manuscript, for her/his comments.

The main concern of the reviewer is the novelty of our work in the light of Ref. [7-10] and the previous pioneering works on the related issue by the Edinburgh group. We should have cited the works by Poon, Evans, and Renth besides the review by Poon, which were also kindly pointed out by Reviewer 2. We should have cited these pioneering works on the pathway whereby a system phase separate into macroscopically coexisting colloidal gas, liquid, and crystal regions and its link to the free energy landscape. We are afraid that this made the novelty of our work a bit obscure in the original version of our paper.

However, we have very carefully read these previous references and confirmed that our new mechanism was not mentioned before. This is supported by the kind comments on this point made by Reviewer 2. In the paper by Renth et al., the closed example to ours is their “Sample 5” belonging to the kinetic regime called “G”. In this case, the process is initiated by spinodal decomposition. They inferred that any crystallites nucleated at the early stage must have gas coatings, on the basis of the non-existence of a liquid-crystal common tangent. But this does not tell anything on the kinetic pathway on a microscopic level. We stress that in these previous studies there was no assess to particle-level information and the discussion is based on macroscopic experiments under an influence of gravity. The subsequent theoretical paper by Evans et al. also stated that spinodal decomposition takes place in the region G before crystallization takes place. Then they cited the simulation work by Soga et al. (Ref. [XX]) and stated that spinodal-like instability leads to a texture of high and low density amorphous regions and then crystals subsequently formed in one of these regions. This statement means that crystallites are formed in either the gas or the liquid phase. Thus, this is clearly different from what we found in our paper as a new pathway involving all the three phases in a non-trivial manner.

We hope that the above explanations would convince the reviewer that our paper reports a novel kinetic pathway to porous crystals, which has not been discussed before.

As explained in our replies to the other reviewers, we have revised our manuscript incorporating all the reviewers’ comments including the above concern on the novelty of our work. We hope that the Reviewer would find that our revised manuscript is now suitable for publication in Nature Materials.

\clearpage

\noindent
\begin{Large}
{\bf Response to Referee \#2}
\end{Large}

\vspace{1em}

\singlespacing

{\bf
The authors report on a very interesting and new result: systems with short-range attractions that undergo freezing from a gas phase can form a solid with short-rang crystalline order but long-range, gel-like amorphous (porous) structure. The experiments support the conclusions quite well and the results are explained in a convincing way using the idea of metastable states. While previous researchers have shown that a short-range attraction can lead to gels (which is well known) or an intermediate fluid state on the way to crystal formation, this is the first report of early-time crystal formation that crosses over to spinodal-like gel structures at later times. I note in passing that this is different from early-time crystallization followed by random aggregation of these crystallites, which would be expected to yield fractal aggregates of crystalline domains, and not the porous or bicontinuous structures reported here. Therefore, I think the results are quite new. I think the results are relevent to the materials community and I recommend publication after a few comments are addressed.
}

\bigskip
\doublespacing

First We would like to thank the Referee for her/his careful reading of our manuscript, for providing us with valuable comments. We are very glad to see such kind positive comments on our work.

\vspace{1em}

\singlespacing

{\bf
The authors propose that this method might work in a variety of other materials and be useful as a generalizable, rational method to make materials that are crystalline yet porous. I agree with them about the potential advantages of such materials. However, I am less optimistic than they are about the possibility of this mechanism in small-molecule systems. For example, the authors propose that this could work for metals - this is mentioned in the abstract and in the conclusion, on page 9. Success in this area requires meeting the 4 conditions listed on pages 8-9 but this may not be possible with nobel metals (especially with regard to iii). Choosing a different example might make this point more credible and the article more valuable. Alternatively, if the authors have good arguments for why this could work for noble metals, then these could be a good addition to the article.
}

\bigskip
\doublespacing

Below we will answer the comments one by one.

\vspace{1em}

\singlespacing

{\bf
Overall the methods are appropriate to the conclusions. The article is quite clear other than the points listed below.
}

\bigskip
\doublespacing

Below we will answer the comments one by one.

\vspace{1em}

\singlespacing

{\bf
A few technical points:

1) Concerning references: Page 8, line 7: "kinetic path via the gas phase (liquid -- > gas -- > crystal) has been completely overlooked so far" - does this statement refer to experiments? I believe that the literature has covered a wide range of theory on this topic, for example in articles by Evans et al or Renth et al with WCK Poon (2 articles in PRE, 2001). One of those articles also reports experiments, which might be worth citing here (though I do not think those experiments are the same as these findings).
}

\bigskip
\doublespacing
We should have cited these pioneering works on the pathway whereby a system phase separate into macroscopically coexisting colloidal gas, liquid, and crystal regions and its link to the free energy landscape. We have carefully read these papers. In the paper by Renth et al., the closed example to ours is their “Sample 5” belonging to the kinetic regime called “G”. In this case, the process is initiated by spinodal decomposition. They inferred that any crystallites nucleated at the early stage must have gas coatings, on the basis of the non-existence of a liquid-crystal common tangent. But this does not tell anything on the kinetic pathway on a microscopic level. We stress that in these previous studies there was no assess to particle-level information and the discussion is based on macroscopic experiments under an influence of gravity. The subsequent theoretical paper by Evans et al. also stated that spinodal decomposition takes place in the region G before crystallization takes place. Then they cited the simulation work by Soga et al. (Ref. [XX]) and stated that spinodal-like instability leads to a texture of high and low density amorphous regions and then crystals subsequently formed in one of these regions. This statement means that crystallites are formed in either the gas or the liquid phase. Thus, this is clearly different from what we found in our paper as a new pathway involving all the three phases in a non-trivial manner. 
Nonetheless, our statement "kinetic path via the gas phase (liquid -- > gas -- > crystal) has been completely overlooked so far" was obviously too strong. So We have replaced “completely” by “largely” and have cited these pioneering works, which considered the role of the complex free energy landscape for the first time.

\vspace{1em}

\singlespacing

{\bf
2) Page 5, second paragraph: the text states that the vol frac of 68\% corresponds to the vol frac of the stable crystal phase. This contradicts the phase diagram shown in Fig 1A, which has the coexisting crystal vol frac at about 0.72 or so - far over toward the right-hand axis. This should be quite distinct from 0.68. There may be good reasons for the difference between the measured 0.68 and the expected equilibrium value, but this should be discussed.
}

\bigskip
\doublespacing

We were not clear enough on this point. In the experimental representation of the phase diagram (vertical axis is $c_p$), the tie lines are not horizontal. We have now drawn the slanted gas-crystal tie line in the figure as a guide. We hope that this would make the situation much clearer. This is not enough according to Mathieu!!

\vspace{1em}

\singlespacing

{\bf
3) How were crystalline particles defined? Page 6 refers to measurements of crystallite size and number, and the definitions should be given. Also does "amount of crystals" (p6) refer to number or volume?

4) How is the local volume fraction defined? Is it the particle volume divided by the volume of the Voronoi cell?

5) On pages 4 and 5, the authors refer to "random gelation universality class" and the associated exponent D=2.53. To what do the authors refer? I would have thought that the appropriate comparison would be a particle-aggregation model such as DLCA or RLCA, which have D=1.75 and 2.0 or greater. Please explain and justify this comparison.
}

\bigskip
\doublespacing

It is known that three-dimensional random percolation clusters and diffusion-limited aggregation clusters have both about the same fractal dimension of about 2.5, for which we have cited the review article by Herrmann. We agree that DLCA and RLCA provide smaller fractal dimensions; however, DLCA is observed in a much lower volume fraction with stronger bonding energy (after bonding there is little diffusion, which leads to the formation of really fractal objects). Our case is very different from such a limit. We hope that this clarifies the situation.

\vspace{1em}

\singlespacing

{\bf
6) A small technical point: 'sublimation' is used to refer to the formation of a crystal from vapor at the bottom of page 7 and on page 8 (8 lines from the bottom). Sublimation is a transition from solid to vapor, not the other way.
}

\bigskip
\doublespacing

We thank the Reviewer for pointing out this mistake. We have corrected it as suggested.

We hope that the Reviewer would find that the revised manuscript is now suitable for publication in Nature Materials.

\clearpage

\noindent
\begin{Large}
{\bf Response to Referee \#3}
\end{Large}

We would like to thank the Referee for her/his careful reading of our
manuscript, and for her/his comments.

\vspace{1em}

\singlespacing

{\bf
Tsurusawa et al. report measurements of crystal gels. Crystal gels are described as crystalline networks that form by a two-step mechanism. First, a metastable spinodal crystallization boundary induces liquid-gas phase separation; crystals then nucleate within the liquid phase. Second, the crystals grow by a process in which liquid droplets evaporate into the gas phase, then desublime to the crystal. This process is viewed as akin to the Bergeron process of ice crystal growth in clouds. I have the following specific comments:
}

\bigskip
\doublespacing

First we would like to thank the Referee for her/his careful reading of our manuscript, for providing us with valuable comments. Below, we will answer the comments one by one.

\vspace{1em}

\singlespacing

{\bf
1) The evidence for this crystal gel mechanism appears to derive from measurements at a single state point. For this sample, what happens at the transition between the two SI movies, at about 500 Brownian times? At early times, I see very few free particles, and no crystallization. A few frames into the longer movie, I see many more free particles, with exchange between the dense and dilute regions, mostly from one crystallite to the other. What, within the scenario proposed, would lead to such an abrupt increase in the abundance of free particles at ~ 500 Brownian times? It is as if the system itself changed its state point and shifted over a boundary that then allowed the crystallization to accelerate. (The increase in free particles occurs more at ~ 500 Brownian times occurs more rapidly than I think would be expected based on the kinetics of spinodal decomposition.) I think that additional observation of the phenomena of interest at different state points and/or evidence for replication at the indicated state points are needed to rule out idiosyncratic (time-dependent) physical chemical explanations for the abrupt increase in gas phase colloids.
}

\bigskip
\doublespacing

We think that we are not clear enough about the explanations of our movies. Movie 1 and Movie 2 were made with different frame rates and there was an interval of 90 s between them. This is because the dynamic is rather fast in the initial stage driven by spinodal decomposition. But in the late stage, the dynamics slows down and thus we reduced the frame rate of 3D scan to cover the long-time period. We are afraid that the Reviewer’s impression on a sudden increase in free gas particles and a possible discontinuity between the two movies originate from this difference in the frame rate. 
Concerning the reproducibility of this kinetic pathway, we already made an independent experiment by using a sample with the similar composition which is inserted into an ordinary glass capillary tube and confirmed the reproducibility of this type of crystallization pathway. This we added in Supplementary Information.

\vspace{1em}

\singlespacing

{\bf
2) From the SI movie, I don't recognize the final reported structure to be a crystal gel. These final structures seem instead to be crystallites of finite size at equilibrium. The crystallites diffusive minimally because they are jammed against their neighbors. The results seem similar to Anderson and Lekkerkerker's (ref 20) explanation of Ostwald ripening. The changes in crystal size at late times, when the liquid to gas to crystal mechanism is proposed, appear that they could be equally well described as a process in which larger crystallites grow at the expense of smaller ones, through the gas phase.

}

\bigskip
\doublespacing

The impression of the reviewer that the final structures are made of crystallites of finite size may come from that the percolated network structure cannot be seen by 2D sectioning. To show the connectivity of the network clearly, we have also made Supplementary Movies of a thin slice with a finite thickness. We hope that this will clarify the situation and convince the reviewer of the fact that our sample indeed forms a porous network of crystals.

\vspace{1em}

\singlespacing

{\bf
3) The data quality of Figure 1c, Figure 1d appear insufficient to classify the cp = 0.38 mg/mg state point as different from the others. The power law scaling in Figure 1c is generated over at most half a decade and the difference between 68% and 64% in Figure 1d at 12 nearest neighbors seems marginally resolvable given what the errors might be. The authors should add replications, errors, and significance testing if they wish to distinguish the dimensionality (Fig 1c) and maximum volume fraction (Fig 1d) of the 0.38 mg/ml state point from the others.

4) In panel 4b, about ~ 25% of the particles are identified as in the surface state; however, transition probabilities are not recorded for this state in Figure 4c. From the SI movies, it seems that most of the kinetics are between surface and gas states at long times (as in Ostwald ripening). I did not find explanation in the paper about how these transition probabilities are defined and measured, and how the surface particles are excluded from the analysis. Also, do the kinetics measured satisfy detailed balance at long times? These data should include crystal-gas and crystal-liquid transition probabilities so that the author's mechanism of liquid to gas to crystal can be evaluated against the alternative that the sublimination/desublimination transitions are from crystal to gas to crystal.

5) Looking at Figure 4b and 4c together: although the gas crystal transition probability is large, the abundance of gas particles is never very significant. In the key window between 2 and 6 thousand Brownian times, when the crystal forms, I do not see how Figure 4c discriminates between a liquid to gas to crystal scenario and a direct liquid to gas scenario. (I agree that at long times there are abundant gas to crystal and crystal to gas transitions.) But, at four thousand Brownian times, the low gas concentration relative to liquid concentration and the equivalence of the transition probabilities suggests to me that the direct liquid to crystal route is what actually builds this crystal. The authors describe this as an initial nucleation process; however, from these figures it appears that ~ 80% of the crystallization actually occurs in this window of initial nucleation, and this process is the typical one (reviewed for example in ref 20).

The evidence for the formation of crystals in this paper is convincing (Figure 2d and Figure 4b). However, because of the weaknesses in the evidence presented in Figure 1c,d and Figure 4b,c, as well as from watching the supplemental movies, I do not think that the authors have adequately proven that they have observed a new crystallization pathway. First, at about ~ 4000 Brownian times, when the crystals form, it seems like a liquid to crystal transition pathway explains the data as well as the proposed liquid to gas to crystal pathway, and the bulk of the crystallization occurs at this point (rather than later through the proposed desublimation pathway). Second, at longer times, these solutions seem to be understandable as diffusing crystallites - the connection to gelation seems unnecessary to understand the physics. The late state dynamics instead seems like Ostwald ripening, in which smaller crystallites transition to larger one through sublimination; the liquid does not appear to play an intermediate role here and there is little evidence for the liquid to gas to crystal route that is the paper's primary message.
}

\bigskip
\doublespacing

We are afraid that the above interpretation basically originates from some misunderstandings, which is caused by our poor explanations on the details of Supplementary Movies and the lack of more 3D representations of our results. We hope that the above explanations and new Movies would convince the reviewer of the formation of a percolated network structure composed of crystals (not isolated crystals diffusing around) and the novel kinetic pathway leading to porous crystalline structures.

\vspace{1em}

\singlespacing

{\bf

Minor points:

A) I think the previous literature about Bernal spirals and other one-dimensional structures in gels (e.g. PRL 94 208301 2005) is relevant to the present study.
}

\bigskip
\doublespacing

We thank the reviewer for notifying this important reference. We have cited it in our revised manuscript.

\vspace{1em}

\singlespacing

{\bf
B) The SI movies would be more effective if they were merged and a time stamp added; that would allow the reader to better compare them to Figure 4.
}

\bigskip
\doublespacing
We have followed the kind suggestion of the reviewer.

\vspace{1em}

\singlespacing

{\bf
C) What time point or condition is Figure 3b at? Please add this information to distinguish these data from Figure 1b.
}

\bigskip
\doublespacing

We have followed the kind suggestion of the reviewer.

We hope that the Reviewer would find that the revised manuscript is now suitable for publication in Nature Materials.

\end{document}

### 2016-03-05 nature/Supplementary.tex
John. First major changes following the Referees comments. Figures 2 and 4 have been updated, and the text changed accordingly.
Expanded supplementary and response letter.

--- 2015-11-19 crystal-gel/nature/Supplementary.tex
@@
-%\documentclass[12pt]{article}
-\documentclass[12pt,onecolumn]{revtex4}
-%\documentstyle[aps,multicol,eincludegraphics]{revtex}
-%\documentclass[twocolumn]{revtex4}
+% %\documentclass[12pt]{article}
+% \documentclass[12pt,onecolumn]{revtex4-1}
+% %\documentstyle[aps,multicol,eincludegraphics]{revtex}
+% %\documentclass[twocolumn]{revtex4}
+% \usepackage{graphicx}
+% %\usepackage{scicite}
+% % \usepackage{epstopdf}
+% \usepackage{amsmath}
+% %\usepackage{times}
+% 
+% \topmargin 0.0cm
+% \oddsidemargin 0.2cm
+% \textwidth 16cm 
+% \textheight 21cm
+% \footskip 1.0cm
+% 
+% 
+% 
+% 
+% \begin{document}
+% 
+% 
+% \noindent
+% {\bf Supplementary Information}
+
+\documentclass[preprint,superscriptaddress]{revtex4-1}
 \usepackage{graphicx}
-%\usepackage{scicite}
-% \usepackage{epstopdf}
+\usepackage{epstopdf}
 \usepackage{amsmath}
-%\usepackage{times}
+% \usepackage{hyperref}
+\usepackage{booktabs}
+\usepackage{color}
 
-\topmargin 0.0cm
-\oddsidemargin 0.2cm
-\textwidth 16cm 
-\textheight 21cm
-\footskip 1.0cm
+\setlength{\tabcolsep}{10pt}
+
+\newenvironment{sistema}%
+  {\left\lbrace\begin{array}{@{}l@{}}}%
+  {\end{array}\right.}
+
+\renewcommand{\figurename}{Fig.} 
+\renewcommand{\thefigure}{S\arabic{figure}}
+
+% \setcounter {figure} {S}
+% \usecounter{figure}
 
@@
 
-\noindent
-{\bf Supplementary Information}
+\centerline{\bf \large Supplementary Information for}
+\centerline{\bf \large``A novel route to the spontaneous formation of porous crystals via}
+\centerline{\bf \large viscoelastic phase separation''}
+\vspace{0.5cm}
 
-\vspace{5mm}
-\noindent
-\emph{\bf Supplementary Movie 1}.
+\centerline{Hideyo Tsurusawa$^1$, John Russo$^{1,2}$, Mathieu Leocmach$^3$ and Hajime Tanaka$^1$} 
+\centerline{$^1${\it Institute of Industrial Science, University of Tokyo}}
+\centerline{ {\it 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan} }
+\centerline{$^2$ {\it School of Mathematics, University of Bristol, Bristol BS8 1TW, United Kingdom}}
+\centerline{$^3$ {\it Institut Lumière Matière, CNRS UMR 5306, Université Claude Bernard Lyon 1,}}
+\centerline{ {\it Université de Lyon, Lyon, 69622 Villeurbanne Cedex, France}}
+% \centerline{ {\it Oxford, OX1 3QZ, United Kingdom}}
+
+\section*{Supplementary Movie 1}
 This movie shows the early stage of gelation process of $\phi\approx 0.30$ and $c_p=0.38$ mg/g, from the original confocal images at the same z-position as Fig.~3a.
@@
 
-\vspace{5mm}
-\noindent
-\emph{\bf Supplementary Movie 2}.
+\section*{Supplementary Movie 2}
 This movie shows the late stage after Supplemental Movie 1 with scanning duration of 27000~s (12300~$\tau$). 
+Please note that there is a large difference in the frame rate between Movie 1 and 2.
+
+\section*{Definition of states}
+The configurations obtained by tracking the position of the colloids in the samples
+are analyzed in order to determine whether a colloidal particles is in the
+crystal, surface, liquid or gas state. In the following we describe the classification
+criteria
+
+\subsection*{Crystal}
+In order to detect crystalline particle we use bond orientational analysis,
+as described in Ref.~\cite{russo2013interplay}.
+A $(2l+1)$ dimensional complex vector ($\mathbf{q}_l$) is defined for each
+particle $i$ as $q_{lm}(i)=\frac{1}{N_b(i)}\sum_{j=1}^{N_b(i)}
+Y_{lm}(\mathbf{\hat{r}_{ij}})$, where we set $l=12$, and $m$ is an integer that
+runs from $m=-l$ to $m=l$. The functions $Y_{lm}$ are the spherical harmonics
+and $\mathbf{\hat{r}_{ij}}$ is the normalised vector from the oxygen of
+molecule $i$ to the one of molecule $j$.  The sum goes over the first
+$N_b(i)=12$ neighbours of molecule $i$. This choice accounts for
+the first coordination shell of close packed crystals (fcc or hcp).
+The scalar product between $q_{6,m}$ of two particles
+is defined as $\mathbf{q}_{6}(i)\cdot\mathbf{q}_{6}(j)=\sum_m q_{6,m}(i)q_{6,m}(j)$.
+For each pair $i$ and $j$ of neighbouring particles we define a \emph{connection}
+if the scalar product $(\mathbf{q}_{6}(i)/|\mathbf{q}_{6}(i)|)\cdot(\mathbf{q}_{6}(j)/|\mathbf{q}_{6}(j)|)>0.7$.
+A molecule is then identified as crystalline if it has at least $7$ connected neighbours.
+
+The criteria outlined above are commonly used to identify crystalline arrangements in hard sphere systems.
+
+\subsection*{Surface, gas and liquid particles}
+
+\begin{figure}[!t]
+ \centering
+ \includegraphics[width=10cm]{./fractions.pdf}
+ % fractions.pdf: 681x503 pixel, 72dpi, 24.02x17.74 cm, bb=0 0 681 503
+ \caption{{\bf Fraction of particles for each phase at $\phi\approx 0.30$ and $c_p=0.38$ mg/g}. The liquid
+ fraction is depicted as a dashed black line, the gas fraction as a dotted-dashed green line, the surface
+ fraction as a double-dotted-dashed blue line, while the crystal with a continuous red line.}
+ \label{fig:fractions}
+\end{figure}
+
+Surface, gas and liquid particles are identified from the subset of non-crystalline particles depending
+on their neighbouring particles. A particle with at least two neighbouring crystal particle is classified as
+\emph{surface}. A particle that is neither solid or surface, and has at least four neighbouring particles
+is classified as \emph{liquid}. A particle which is neither solid, surface or liquid is classified as \emph{gas}.
+Fig.~\ref{fig:fractions} plots the fraction of each state for the state point at $\phi\approx 0.30$ and $c_p=0.38$ mg/g
+
+The distinction between gas and liquid phases is consistent with the analysis of volume fraction and composition
+of the gas and liquid phases just after phase separation (and before crystallization). To analyse the fraction
+of liquid and gas after phase separation we convert the position of particles into a contiunous field smearing
+the position of each particle with a gaussian field of variance $5$ pixels, and then mapping the field
+to a cubic grid with $10$ pixels spacing. In this way liquid and gas regions can be identified by
+an appropriate threshold of the field.
+
+\begin{figure}[!t]
+ \centering
+ \includegraphics[width=10cm]{./field.pdf}
+ % field.pdf: 681x483 pixel, 72dpi, 24.02x17.04 cm, bb=0 0 681 483
+ \caption{{\bf Distribution function of the gaussian field at $t=490\,\tau_B$.} The gaussian field is applied to
+ a configuration just after phase separation and before crystallization. The vertical dashed line represents the
+ field threshold $f=0.06$ used to distinguish between gas and liquid particles.}
+ \label{fig:field}
+\end{figure}
+
+Fig.~\ref{fig:field} shows the probability distribution of the field, where two peaks clearly indicate the presence
+of gas/liquid coexistence. By choosing a threshold over the field of $f=0.06$ (dashed line) we compute the fraction
+of particles in the solid and gas phases at $t=490\,\tau_B$, just after phase separation. The results are consistent
+with Fig.~\ref{fig:fractions}, where the number fraction of particles in the liquid state accounts for more than
+$90\%$ of the particles after phase separation.
+
+\section*{Transition probabilities}
+\begin{figure}[!t]
+ \centering
+ \includegraphics[width=14cm]{./sfig1.pdf}
+ % sfig1.pdf: 1398x1026 pixel, 72dpi, 49.32x36.19 cm, bb=0 0 1398 1026
+ \caption{{\bf Transition probabilities during the formation of the crystal-gel.}
+ {\bf a,} Transition probabilities from the liquid state;
+ {\bf b,} Transition probabilities from the gas state;
+ {\bf c,} Transition probabilities from the surface state;
+ {\bf d,} Transition probabilities from the crystal state. In all panels the possible
+ transition states are surface (blue), liquid (black), gas (green) and crystal (red).
+ See the text for a definition of the states.
+ }
+ \label{fig:probabilities}
+\end{figure}
+
+In Fig.~\ref{fig:probabilities} we plot the transition probabilities between the gas, liquid, crystal
+and surface states, as defined in the previous Section. Each panel represents the transition probability
+of a different state: {\bf a} for the liquid state, {\bf b} for the gas state, {\bf c} for the surface state,
+and {\bf d} for the crystal state. The results indicate that the liquid phase preferentially transforms into
+the surface state (as a precursor to crystallization), which accounts for the \emph{direct crystallization channel}
+that was described in the main text. A high percentage of liquid transforms also in the gas phase. The gas phase itself
+transforms back into liquid or into surface particles, a process which
+is linked to the \emph{Bergeron process}. These results support the idea that both channels are active
+crystallization pathways. Direct transformation of either liquid or gas into crystalline states is almost absent,
+indicating that our surface state indeed captures the precursor particles that attach to nuclei and later crystallize.
+Surface particles in fact, first transform preferentially into the liquid state, but at later time instead transform
+more prominently into the crystal state. The transformation of crystals into surface state, represents the first
+step of the \emph{Ostwald process}, but this channel is limited by the small rate of conversion of surface particles
+into the gas state.
+
+\bibliographystyle{naturemag2}
+\bibliography{biblio}
 

### 2016-03-05 nature/crystal_gel2.tex
Expanded supplementary and response letter.

@@
 \affiliation{Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}
+\affiliation{ {School of Mathematics, University of Bristol, Bristol BS8 1TW, United Kingdom} }
 \author{Mathieu Leocmach}
 \affiliation{Institut Lumière Matière, CNRS UMR 5306, Université Claude Bernard Lyon 1, Université de Lyon, Lyon, 69622 Villeurbanne Cedex, France}
-\author{Hajime Tanaka}
-\affiliation{Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan \footnote{e-mail: tanaka@iis.u-tokyo.ac.jp}}
+\author{Hajime Tanaka  \footnote{e-mail: tanaka@iis.u-tokyo.ac.jp}}
+\affiliation{Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}
 
@@
 \textcolor{red}{
+In Supplementary Information we confirm these results with a detailed analysis of transition probabilities, including a state for surface particles that are in contact.
+to crystalline particles.
 While the direct freezing of the fluid represents the major contribution to the nuclei growth, the kinetic path via the gas phase (gas$\rightarrow$crystal), also plays a crucial role,
@@
 plays a considerably more important role than Ostwald ripening.}
+
 % 

### 2016-03-06 nature/response.tex
Updated version.

@@
 transitions is classified according to the free energy landscape. These work are by themselves very important and relevant, but contrary
-to the Referee claim, do not contain a description or results on the crystallization channel we describe in our work. A crucial difference is
-the fact that those previous works
-study the case in which gas-liquid phase separation is stable against crystellization, while our crystallization scenario depends crucially
-on the metastability of liquid-gas phase separation against crystallization. Both Referees 2 and 3 have appreciated this fact.
+to the Referee claim, do not contain a description or results on the crystallization channel we describe in our work.
+Both Referees 2 and 3 have appreciated this fact.
 
@@
 
-We hope that you will find the revised version of the manuscript to be fully addressing the Referee questions, and
+We hope that you will find that we have fully addressed the Referee questions, and
 that you will find the new manuscript to be suitable for publication in \emph{Nature Materials}.
@@
 An extensive body of work that the authors apparently missed.
-Although the experimental study and analysis is very nice and beautiful, the referee hesitate to recommend the manuscript for publication in Nature Materials due to a lack of novelty (in terms of material and fundamental insight) of the present work.
+Although the experimental study and analysis is very nice and beautiful, the
+referee hesitate to recommend the manuscript for publication in Nature Materials due to a lack of novelty (in terms of material and fundamental insight) of the present work.
 }
@@
 [7-10] and the previous pioneering works on the related issue by the Edinburgh group.
-We should have cited the works by Poon, Evans, and Renth besides the review by Poon, which were also kindly pointed out by Reviewer 2.
-We should have cited these pioneering works on the pathway whereby a system phase separate into macroscopically coexisting colloidal gas, liquid, and crystal regions and its link to the free energy landscape.
-We are afraid that this made the novelty of our work a bit obscure in the original version of our paper.
+In the original manuscript we only cited a review by Poon, but, as the Referee correctly points out, we should have
+also included papers by Poon, Evans, and Renth and the Edinburgh group on phase separation into coexisting colloidal gas, liquid, and crystal regions, and its link to the free energy landscape.
 
@@
 This is supported by the kind comments on this point made by Reviewer 2.
-In the paper by Renth et al., the closed example to ours is their “Sample 5” belonging to the kinetic regime called “G”.
+The references predict the kinetic pathway from the free energy landscape, and the discussion
+is based on macroscopic experiments under the influence of gravity. These experiments do not contain information on the kinetic pathway at a microscopic level.
+In the paper by Renth et al., the closest example to our crystalizzation mechanism is “Sample 5” belonging to the kinetic regime called “G”.
 In this case, the process is initiated by spinodal decomposition.
 They inferred that any crystallites nucleated at the early stage must have gas coatings, on the basis of the non-existence of a liquid-crystal common tangent.
-But this does not tell anything on the kinetic pathway on a microscopic level.
-We stress that in these previous studies there was no assess to particle-level information and the discussion is based on macroscopic experiments under an influence of gravity.
 The subsequent theoretical paper by Evans et al.
@@
 (Ref.
-[XX]) and stated that spinodal-like instability leads to a texture of high and low density amorphous regions and then crystals subsequently formed in one of these regions.
+7) and
+stated that spinodal-like instability leads to a texture of high and low density amorphous regions and then crystals subsequently formed in one of these regions.
 This statement means that crystallites are formed in either the gas or the liquid phase.
-Thus, this is clearly different from what we found in our paper as a new pathway involving all the three phases in a non-trivial manner.
+Overall, the process is different from what we found in our paper, which is a pathway involving all the three phases in a non-trivial manner.
+Unveiling the exact kinetic pathway of crystal-gel formation cannot
+be done exclusively from the free energy-landscape, and requires a particle-level study of the process, which is achieved for the first time by our experiments.
 
@@
 Therefore, I think the results are quite new.
-I think the results are relevent to the materials community and I recommend publication after a few comments are addressed.
+I think the results are relevent to the
+materials community and I recommend publication after a few comments are addressed.
 }
@@
 
-1) Concerning references: Page 8, line 7: "kinetic path via the gas phase (liquid -- > gas -- > crystal) has been completely overlooked so far" - does this statement refer to experiments?
+1) Concerning references: Page 8, line 7: "kinetic path via the gas phase (liquid $\rightarrow$ gas $\rightarrow$ crystal) has been completely overlooked so far" - does this statement refer to experiments?
 I believe that the literature has covered a wide range of theory on this topic, for example in articles by Evans et al or Renth et al with WCK Poon (2 articles in PRE, 2001).
@@
 also stated that spinodal decomposition takes place in the region G before crystallization takes place.
-Then they cited the simulation work by Soga et al.
+Then they cited the
+simulation work by Soga et al.
 (Ref.
-[XX]) and stated that spinodal-like instability leads to a texture of high and low density amorphous regions and then crystals subsequently formed in one of these regions.
+7) and stated that spinodal-like instability leads to a texture of high and low density amorphous regions and then crystals subsequently formed in one of these regions.
 This statement means that crystallites are formed in either the gas or the liquid phase.
 Thus, this is clearly different from what we found in our paper as a new pathway involving all the three phases in a non-trivial manner.
-Nonetheless, our statement "kinetic path via the gas phase (liquid -- > gas -- > crystal) has been completely overlooked so far" was obviously too strong.
-So We have replaced “completely” by “largely” and have cited these pioneering works, which considered the role of the complex free energy landscape for the first time.
+Nonetheless, our statement "kinetic path via the gas phase (liquid $\rightarrow$ gas $\rightarrow$ crystal) has been completely overlooked so far" was obviously too strong.
+We have replaced “completely” by “largely” and have cited these pioneering works, which considered the role of the complex free energy landscape for the first time.
 
@@
 We hope that this would make the situation much clearer.
-This is not enough according to Mathieu!!
+\textcolor{red}{Mathieu could you check this point?}
 
@@
 4) How is the local volume fraction defined? Is it the particle volume divided by the volume of the Voronoi cell?
+}
 
+\bigskip
+\doublespacing
+
+We apologize that these technical details were not sufficiently explained in the previous version of the manuscript.
+We have now added a section to the Supplementary Informations detailing the exact definition of crystalline particles
+(and the other states too). In brief, crystallinity is measured by means of bond orientational analysis, with order
+parameters designed to detect all relevant crystalline phases for colloidal/polymer mixtures (fcc, hcp and bcc).
+The amount of crystals always refers to number and not volume.
+
+The Referee is correct regarding the volume fraction. It is defined by calculating the Voronoi diagram of each configuration
+and computing the fraction of volume that each colloid occupies in its cell. We have also added this information to
+the manuscript.
+
+\vspace{1em}
+
+\singlespacing
+
+{\bf
 5) On pages 4 and 5, the authors refer to "random gelation universality class" and the associated exponent D=2.53.
@@
 
+By "random gelation universality class" we mean three-dimensional percolation universality class (which applies for example to randomly branched polymers).
 It is known that three-dimensional random percolation clusters and diffusion-limited aggregation clusters have both about the same fractal dimension of about 2.5, for which we have cited the review article by Herrmann.
-We agree that DLCA and RLCA provide smaller fractal dimensions; however, DLCA is observed in a much lower volume fraction with stronger bonding energy (after bonding there is little diffusion, which leads to the formation of really fractal objects).
-Our case is very different from such a limit.
-We hope that this clarifies the situation.
+We agree that DLCA and RLCA provide smaller fractal dimensions; however, DLCA is usually observed at much lower volume fraction and with stronger bonding energy (after bonding there is little diffusion, which leads to the formation of really fractal objects) than our case.
+For our starting volume fraction of about $\phi\approx 0.30$ and polymer concentrations, we have found that the fractal dimension
+to be more consistent with random percolation. We hope that this clarifies the situation.
 
@@
 It is as if the system itself changed its state point and shifted over a boundary that then allowed the crystallization to accelerate.
-(The increase in free particles occurs more at ~ 500 Brownian times occurs more rapidly than I think would be expected based on the kinetics of spinodal decomposition.) I think that additional observation of the phenomena of interest at different state points and/or evidence for replication at the indicated state points are needed to rule out idiosyncratic (time-dependent) physical chemical explanations for the abrupt increase in gas phase colloids.
+(The increase in free particles occurs more at ~ 500 Brownian times occurs more rapidly than I think would be expected based on the kinetics of spinodal decomposition.) I think that additional observation of the phenomena of interest at different state points and/or evidence for replication at the indicated
+state points are needed to rule out idiosyncratic (time-dependent) physical chemical explanations for the abrupt increase in gas phase colloids.
 }
@@
 
-We think that we are not clear enough about the explanations of our movies.
+We think that we have not been clear enough about the explanations of the Supplementary movies.
 Movie 1 and Movie 2 were made with different frame rates and there was an interval of 90 s between them.
-This is because the dynamic is rather fast in the initial stage driven by spinodal decomposition.
-But in the late stage, the dynamics slows down and thus we reduced the frame rate of 3D scan to cover the long-time period.
-We are afraid that the Reviewer’s impression on a sudden increase in free gas particles and a possible discontinuity between the two movies originate from this difference in the frame rate.
-Concerning the reproducibility of this kinetic pathway, we already made an independent experiment by using a sample with the similar composition which is inserted into an ordinary glass capillary tube and confirmed the reproducibility of this type of crystallization pathway.
-This we added in Supplementary Information.
+This is because the dynamics is rather fast in the initial stage driven by spinodal decomposition.
+But, in the late stage, the dynamics slows down and thus we reduced the frame rate of 3D scan to cover the long-time period.
+We are afraid that the Reviewer’s impression of a sudden increase in free gas particles and a possible discontinuity between the two movies originatse from this difference in the frame rate.
+
+To address the concern about the reproducibility of our kinetic pathway, we decided to make additional independent experiments
+by using a sample with similar composition, which is inserted into an ordinary glass capillary tube.
+We thus confirm the reproducibility of this type of crystallization pathway.
+We have added these additional experiments in the Supplementary Information.
+We would like to thank the Referee for stimulating
+this additional investigation.
 
@@
 
-The impression of the reviewer that the final structures are made of crystallites of finite size may come from that the percolated network structure cannot be seen by 2D sectioning.
+The impression of the reviewer that the final structures are made of crystallites of finite size may come from the fact that the percolated network structure cannot be seen by 2D sectioning.
+The Movies show only a thin slice of the system which corresponds to the particle that are in focus.
 To show the connectivity of the network clearly, we have also made Supplementary Movies of a thin slice with a finite thickness.
@@
 3) The data quality of Figure 1c, Figure 1d appear insufficient to classify the cp = 0.38 mg/mg state point as different from the others.
-The power law scaling in Figure 1c is generated over at most half a decade and the difference between 68% and 64% in Figure 1d at 12 nearest neighbors seems marginally resolvable given what the errors might be.
+The power law scaling in Figure 1c is generated over at most half a decade and the difference between 68\% and 64\% in Figure 1d at 12 nearest neighbors seems marginally resolvable given what the errors might be.
 The authors should add replications, errors, and significance testing if they wish to distinguish the dimensionality (Fig 1c) and maximum volume fraction (Fig 1d) of the 0.38 mg/ml state point from the others.
-
-4) In panel 4b, about ~ 25% of the particles are identified as in the surface state; however, transition probabilities are not recorded for this state in Figure 4c.
-From the SI movies, it seems that most of the kinetics are between surface and gas states at long times (as in Ostwald ripening).
-I did not find explanation in the paper about how these transition probabilities are defined and measured, and how the surface particles are excluded from the analysis.
-Also, do the kinetics measured satisfy detailed balance at long times?
-These data should include crystal-gas and crystal-liquid transition probabilities so that the author's mechanism of liquid to gas to crystal can be evaluated against the alternative that the sublimination/desublimination transitions are from crystal to gas to crystal.
-
-5) Looking at Figure 4b and 4c together: although the gas crystal transition probability is large, the abundance of gas particles is never very significant.
-In the key window between 2 and 6 thousand Brownian times, when the crystal forms, I do not see how Figure 4c discriminates between a liquid to gas to crystal scenario and a direct liquid to gas scenario.
-(I agree that at long times there are abundant gas to crystal and crystal to gas transitions.) But, at four thousand Brownian times, the low gas concentration relative to liquid concentration and the equivalence of the transition probabilities suggests to me that the direct liquid to crystal route is what actually builds this crystal.
-The authors describe this as an initial nucleation process; however, from these figures it appears that ~ 80% of the crystallization actually occurs in this window of initial nucleation, and this process is the typical one (reviewed for example in ref 20).
-
-The evidence for the formation of crystals in this paper is convincing (Figure 2d and Figure 4b).
-However, because of the weaknesses in the evidence presented in Figure 1c,d and Figure 4b,c, as well as from watching the supplemental movies, I do not think that the authors have adequately proven that they have observed a new crystallization pathway.
-First, at about ~ 4000 Brownian times, when the crystals form, it seems like a liquid to crystal transition pathway explains the data as well as the proposed liquid to gas to crystal pathway, and the bulk of the crystallization occurs at this point (rather than later through the proposed desublimation pathway).
-Second, at longer times, these solutions seem to be understandable as diffusing crystallites - the connection to gelation seems unnecessary to understand the physics.
-The late state dynamics instead seems like Ostwald ripening, in which smaller crystallites transition to larger one through sublimination; the liquid does not appear to play an intermediate role here and there is little evidence for the liquid to gas to crystal route that is the paper's primary message.
 }
@@
 
-We are afraid that the above interpretation basically originates from some misunderstandings, which is caused by our poor explanations on the details of Supplementary Movies and the lack of more 3D representations of our results.
+We can partly agree that the results of Fig. 1c/d would not be enough to differentiate the  cp = 0.38 mg/g state from the others
+state points, if considered alone. The spirit of these figures is to show all the information we can get before we utilize the bond orientational
+analysis to detect crystalline particles. Bond orientational analysis, as reported in Fig.2/3/4, confirms
+that indeed the state point cp = 0.38 mg/g is affected by crystallization. In particular, the dimensionality obtained
+with the structure factors, will be examined in more detail in Fig. 3b, where single-particle statistics allows us to
+get rid of all the noise.
+Regarding the volume fractions of Fig. 1d, we have performed the error analysis suggested by the Referee, by looking at the
+standard deviations of the volumes of particles with different neighbours. The results are reported in the updated Fig. 1d,
+where we indeed confirm that our results are statistically sound.
+
+\vspace{1em}
+
+\singlespacing
+
+{\bf
+4) In panel 4b, about ~ 25\% of the particles are identified as in the surface state; however, transition probabilities are not recorded for this state in Figure 4c.
+From the SI movies, it seems that most of the kinetics are between surface and gas states at long times (as in Ostwald ripening).
+I did not find explanation in the paper about how these transition probabilities are defined and measured, and how the surface particles are excluded from the analysis.
+Also, do the kinetics measured satisfy detailed balance at long times?
+These data should include crystal-gas and crystal-liquid transition probabilities so that the author's mechanism of liquid to gas to crystal can be evaluated against the alternative that the sublimination/desublimination transitions are from crystal to gas to crystal.
+}
+
+In the previous version of the manuscript, Fig. 4c included surface particles together with crystalline particles just to improve
+the readability of the graph, but we agree with the Referee that by doing so we didn't convincingly show that the Bergeron channel (liquid-to-gas)
+is more prominent than the Ostwald ripening channel (crystal-to-gas). 
+To include the full analysis, as requested by the Referee,
+we have added a section to Supplementary Information, with a new figure (Fig. S3) containing four panels that detail the transition
+probabilities between all states. In particular we note that, while being an active channel, the Ostwald ripening pathway is limited by
+the small conversion probability of surface particles into the gas phase.
+
+\vspace{1em}
+
+\singlespacing
+
+{\bf
+5) Looking at Figure 4b and 4c together: although the gas crystal transition probability is large, the abundance of gas particles is never very significant.
+In the key window between 2 and 6 thousand Brownian times, when the crystal forms, I do not see how Figure 4c discriminates between a liquid to gas to crystal scenario and a direct liquid to gas scenario.
+(I agree that at long times there are abundant gas to crystal and crystal to gas transitions.) But, at four thousand Brownian times, the low gas concentration relative to liquid concentration and the equivalence of the transition probabilities suggests to me that the direct liquid to crystal route is what actually builds this crystal.
+The authors describe this as an initial nucleation process; however, from these figures it appears that ~ 80\% of the crystallization actually occurs in this window of initial nucleation, and this process is the typical one (reviewed for example in ref 20).
+The evidence for the formation of crystals in this paper is convincing (Figure 2d and Figure 4b).
+However, because of the weaknesses in the evidence presented in Figure 1c,d and Figure 4b,c, as well as from watching the supplemental movies, I do not think that the authors have adequately proven that they have observed a new crystallization pathway.
+First, at about ~ 4000 Brownian times, when the crystals form, it seems like a liquid to crystal transition pathway explains the data as well as the proposed liquid to gas to crystal pathway, and the bulk of the crystallization occurs at this point (rather than later through the proposed desublimation pathway).
+Second, at longer times, these solutions seem to be understandable as diffusing crystallites - the connection to gelation seems unnecessary to understand the physics.
+The late state dynamics instead seems like Ostwald ripening, in which smaller crystallites transition to larger one through sublimination; the liquid does not appear to play an intermediate
+role here and there is little evidence for the liquid to gas to crystal route that is the paper's primary message.
+}
+
+\bigskip
+\doublespacing
+
+As mentioned before, we are afraid that the above interpretation basically originates from some misunderstandings, which is caused by our poor explanations on the different frame rates of Supplementary Movies and the lack of more 3D representations of our results.
 We hope that the above explanations and new Movies would convince the reviewer of the formation of a percolated network structure composed of crystals (not isolated crystals diffusing around) and the novel kinetic pathway leading to porous crystalline structures.
+Moreover, to address directly the Referee's concerns we have added a new analysis, whose results are reported in the new Fig. 4c.
+The analysis is based on a classification of the trajectory of
+each particle.
+By considering the different state transformation that a particle undergoes, we are able to classify its trajectory as belonging to one of three different crystallization channels: direct crystallization, Bergeron and Ostwald.
+Direct crystallization refers to trajectories in
+which gas or liquid particles transition to the crystal state without liquid evaporation or crystal de-sublimation in between the two states.
+The Bergeron process refers to trajectories in which liquid particles transition to the gas state before attaching to the nucleus. The Ostwald
+process refers to trajectories in which a crystal particle transitions to the gas state before crystallizing again.
+The results show that, over the whole process, 54\% of the trajectories crystallize directly, 34\% take the Bergeron route, and
+11\% instead can be attributed to Ostwald ripening. We hope that this analysis can convince the Referee that the Bergeron process is a
+fundamental new channel of crystallization, that plays an important role in the growth phase of crystallization.
 

### 2016-03-06 crystal-gel/nature/crystal_gel2.tex
Updated version.

@@
 \affiliation{Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}
+\affiliation{ {School of Mathematics, University of Bristol, Bristol BS8 1TW, United Kingdom} }
 \author{Mathieu Leocmach}
 \affiliation{Institut Lumière Matière, CNRS UMR 5306, Université Claude Bernard Lyon 1, Université de Lyon, Lyon, 69622 Villeurbanne Cedex, France}
-\author{Hajime Tanaka}
+\author{Hajime Tanaka  \footnote{e-mail: tanaka@iis.u-tokyo.ac.jp}}
 \affiliation{Institute of Industrial Science, University of Tokyo, 4-6-1 Komaba, Meguro-ku, Tokyo 153-8505, Japan}
@@
 {\bf
-Viscoelastic phase separation of colloidal suspensions can be interrupted either by the glass transition, with the formation of colloidal gels~\cite{poon2002,zaccarelli2007,piazza1994phase,verhaegh1997transient,tanaka1999colloid,lu2008gelation}, or by crystallization, which results in the formation of a crystalline network~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways,sabin2012}.
-But while the former process has been studied in detail, the latter has never been observed experimentally at a single-particle level and it is thus lacking a microscopic understanding.
+Viscoelastic phase separation of colloidal suspensions can be interrupted to form gels either by 
+%the glass transition, with the formation of colloidal gels~\cite{poon2002,zaccarelli2007,piazza1994phase,verhaegh1997transient,tanaka1999colloid,lu2008gelation}, or by crystallization, which results in the formation of a crystalline network~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways,sabin2012}.
+glass transition~\cite{poon2002,zaccarelli2007,piazza1994phase,verhaegh1997transient,tanaka1999colloid,lu2008gelation} or by crystallization~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways,sabin2012}.
+%But while the former process has been studied in detail, the latter has never been observed experimentally at a single-particle level and it is thus lacking a microscopic understanding. 
 With a new confocal microscopy protocol, we unveil a novel crystallization pathway to sponge-like porous structures in a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation.
@@
 desublimation, driving evaporation of nearby liquid.
-The latter processes via the gas phase has been overlooked so far.
-This process represents the colloidal analogue of the Bergeron process~\cite{glickman2000glossary}, which explains the formation of ice crystals in mixed phase clouds and is at the origin of rain.
+%The latter processes via the gas phase has been overlooked so far.
+%This process represents the colloidal analogue of the Bergeron process~\cite{glickman2000glossary}, which explains the formation of ice crystals in mixed phase clouds and is at the origin of rain.
 Our finding gives us full experimental access to the kinetic pathway of this poorly understood phenomenon important for climate science.
-The key is complex dynamic interplay involving the three phases upon simultaneous occurrence of two non-equilibrium phenomena, phase separation and crystallization.
+%The key is complex dynamic interplay involving the three phases upon simultaneous occurrence of two non-equilibrium phenomena, phase separation and crystallization. 
 We argue that similar crystal-gel states can be formed as the result of the gas-liquid phase separation of any crystallizable components, such as monoatomic and molecular systems, if dynamics of the liquid phase is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
-This novel mechanism not only realizes single-step spontaneous formation of nano-porous crystals of metals and semiconductors without dealloying \cite{erlebacher2001evolution}, which may be important for catalytic, optical, sensing, and filtration applications \cite{ding2004nanoporous,ding2009nanoporous, wittstock2010nanoporous}, but also sheds new light on crystal network formation in more complex systems such as magma \cite{philpotts1998role}, biominerals \cite{rousseau2005multiscale}, and foods \cite{deman1987fat}.}
+This mechanism will provide a novel single-step pathway to form nano-porous crystals of metals and semiconductors without dealloying \cite{erlebacher2001evolution}, which may be important for catalytic, optical, sensing, and filtration applications \cite{ding2004nanoporous,ding2009nanoporous, wittstock2010nanoporous}.
+%, but also sheds new light on crystal network formation in more complex systems such as magma \cite{philpotts1998role}, biominerals \cite{rousseau2005multiscale}, and foods \cite{deman1987fat}.}
+}
 \end{abstract}
@@
 The resulting transition from open to more compact networks occurs only through the breaking of the bonds that have accumulated more stress.
-The process was called stress-driven ageing~\cite{tanaka2007spontaneous}, which is characteristic of viscoelastic phase separation \cite{tanaka2000viscoelastic}.
-Below we provide the first experimental proof of this process.
+The process was called stress-driven ageing~\cite{
+%tanaka2007spontaneous}, which is characteristic of viscoelastic phase separation \cite{tanaka2000viscoelastic}. Below we provide the first experimental proof of this process.
 
@@
 The figure clearly shows that the network restructuring follows a peak in the number of bond-breaking events.
-We have thus found direct evidence of the stress-driven ageing process in colloidal gels that is a direct consequence of the formation of a thin network with the help of hydrodynamic interactions in the first stages of the aggregation process.
+We have thus found direct evidence of the stress-driven ageing process in colloidal gels that is a direct consequence of the formation of a thin network with the help of
+hydrodynamic interactions in the first stages of the aggregation process.
 
@@
 
-To gain additional insight, in Fig.~\ref{fig:phase_separation}d we plot the average local volume fraction, computed
-through Voronoi diagrams, as a function of the number of neighbours. For all state points, we can see that the average volume fraction of a colloidal
+To gain additional insight, in Fig.~\ref{fig:phase_separation}d we plot the average local volume fraction as a function of the number of neighbours.
+\textcolor{red}{The volume fraction is obtained by computing the Voronoi diagram of each configuration,
+which uniquely assigns a volume to each colloidal particle.}
+For all state points, we can see that the average volume fraction of a colloidal
 particle increases with its number of neighbours. Approaching close packing (12 neighbours), we clearly see two families of curves. The state point
@@
 
-Now we demonstrate that the volume growth is due to direct condensation of the gas phase on the crystalline seeds.
+\textcolor{red}{Next we analyse in detail the crystallization trajectory, distinguishing between the different process responsible for the
+crystallization of the phase separated network.
 Supplementary Movie 2 shows the confocal microscopy video for the crystallizing sample at $\phi\approx 0.30$ and $c_p=0.38$ mg/g,
@@
 From the trajectory, we extract slabs of 10 particle diameter thickness for configurations at different times, shown in Fig.~\ref{fig:transitions}a. 
-Particles are colored according to their phases (gas, liquid, crystal, and crystal surface). The figure shows the first nucleation 
-events inside the liquid network (left panel); then these crystalline nuclei grow directly from the gas phase (middle panel), and finally
-the different nuclei coalesce (right panel).
+Particles are coloured according to their phases (gas, liquid, and crystal). For the classification criteria see the \emph{Methods} section. The figure shows the first nucleation 
+events inside the liquid network (left panel).
+At the same time liquid regions that have not crystallized start evaporating, and the gas phase contributes to the growth of nuclei (middle panel).
+Finally the different nuclei coalesce (right panel).
+Note that the network pattern changes between $t=2055\,\tau_B$ (left panel) and $t=5577,\tau_B$ (middle panel): not all liquid
+branches of the network crystallize due to evaporation. The crystal-gel structure instead persists, as seen by comparing $t=5577,\tau_B$ (middle panel) and $t=12098,\tau_B$ (right panel).}
 
-%In the next section we show evidence for the fact that the volume growth is due to direct condensation of the gas phase on the crystalline seeds.
-%First we show in Fig.~\ref{fig:transitions}a slabs of 10 particle diameter thickness for configurations at $\phi\approx 0.30$ and $c_p=0.38$ at different times.
-%Particles are coloured according to their phase as follows. Green: hcp; red: fcc; blue: bcc; white: liquid or gas. The figure shows the first nucleation
-%events inside the liquid network (left panel); then these crystalline nuclei grow directly from the gas phase (middle panel), and finally
-%the different nuclei coalesce (right panel).
-
+\textcolor{red}{
 Next we show in Fig.~\ref{fig:transitions}b the fraction of particles in each different phase for the crystallization trajectory at
-$\phi\approx 0.30$ and $c_p=0.38$ mg/g, after the liquid-gas phase separation has occurred. The process of crystallization is characterized 
-by a steep decrease in liquid particles, as they transform in small crystalline nuclei inside the liquid domains. This decrease
+$\phi\approx 0.30$ and $c_p=0.38$ mg/g, after the liquid-gas phase separation has occurred.
+For the classification criteria see \emph{Methods}.
+The process of crystallization is characterized 
+by a steep decrease in liquid particles, as they transform into small crystalline nuclei inside the liquid domains. This decrease
 is then accompanied by an increase in the fraction of gas particles: as the first crystals start to reach the gas phase, liquid particles evaporate 
 to the gas phase due to the higher vapour pressure of the liquid phase compared to the crystalline phase. After the onset 
-of a steady-state gas population, the crystallization proceeds by adding surface particles from the gas phase to the small crystallites. To confirm this process we
-examine in Fig.~\ref{fig:transitions}c the transfer matrix probabilities $P_{\alpha\beta}$, where $\alpha$ and $\beta$ are each of the 
-phases, gas, liquid and crystal (which includes hcp, fcc and surface particles). Liquid particles undergo two processes: on one
-hand they transform to crystallites growing inside the liquid branches of the network, and also increasingly transform to gas particles.
-Gas particles first undergo a transition to liquid particles as the final stage of the liquid-gas phase separation, but once
-crystalline nuclei outgrow the liquid pockets in which they are formed, the probability for the transition from gas to liquid
-abruptly decreases. Instead, gas particles start depositing on the crystal, and at later stages the pace of the gas-to-crystal transformation (sublimation)
-is larger than the one of the liquid-to-crystal one.
+of a steady-state gas population, there are three growth channels for the crystal. Direct crystallization is the process by which 
+crystals grow by incorporating nearby liquid particles. In the Bergeron process, liquid droplets first evaporate and the resulting gas phase 
+contributes to the growth of crystalline regions. Ostwald ripening is instead the process by which small crystallites sublimate, and colloidal
+particles are transferred to larger nuclei.
+In order to assess the relative importance of these three growth channels, we analyse the trajectory, and, for every time frame, assign each particle a state between gas, liquid and crystal.
+In order to minimize short-term fluctuations, the state of each colloidal particle is time averaged for $50\,\tau_B$.
+We then measure the fraction of trajectories with different transition histories.
+Direct crystallization accounts for 54\% of particles trajectories in which
+gas or liquid particles transition to the crystal state without liquid evaporation or crystal de-sublimation. The Bergeron process accounts for 35\% of particle trajectories in which liquid particles
+transition to the gas state before crystallizing. Ostwald ripening accounts for 11\% of particle trajectories in which crystal particles transition to the gas state before returning
+to the crystal state. The different crystallization channels are depicted in the diagram of Fig.~\ref{fig:transitions}c.}
 
-We can thus conclude that the crystallization proceeds via two distinct mechanisms:
-(1) freezing of the fluid; (2) evaporation of the fluid and condensation of the vapour on the crystals, i.e., a coupled process of evaporation and sublimation.
-The first process dominates the initial stages of crystallization, while the second one becomes effective once
-crystalline nuclei have outgrown the liquid regions in which they form.
-Process (2), or the kinetic path via the gas phase (liquid$\rightarrow$gas$\rightarrow$crystal), has been completely overlooked so far, although it plays a crucial role
-not only in the kinetic pathway, but also in determining the morphology of the porous crystal, as we discussed above. We note that this is the crucial step  
-in the Bergeron process~\cite{glickman2000glossary,morrison2012resilience} (see below).
+\textcolor{red}{
+In Supplementary Information we confirm these results with a detailed analysis of transition probabilities, including a state for surface particles that are in contact.
+to crystalline particles.
+While the direct freezing of the fluid represents the major contribution to the nuclei growth, the kinetic path via the gas phase (gas$\rightarrow$crystal), also plays a crucial role,
+especially in determining the morphology of the porous crystal, as we discussed above. In this context the Bergeron process~\cite{glickman2000glossary,morrison2012resilience}
+plays a considerably more important role than Ostwald ripening.}
+
 % 
@@
  \centering
- \includegraphics[width=14cm]{figure2_n}
+ \includegraphics[width=12cm]{fig2}
  % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
-\caption{{\bf Percolated networks: {\bf a:} gel; {\bf b:} crystal-gel.} The gel state is obtained from the state
+\caption{\textcolor{red}{{\bf Percolated network structures.} {\bf a:} gel; {\bf b:} crystal-gel. The gel state is obtained from the state
 point $\phi\approx 0.30$ and $c_p=1.07$ mg/g, while the crystal-gel is obtained at the same $\phi$ but
-at lower polymer concentration, $c_p=0.38$ mg/g. The top row shows depicts the networks as a continuous field that has
+at lower polymer concentration, $c_p=0.38$ mg/g. On the left, the network is represented as a continuous field that has
 been obtained by applying a gaussian filter to the position of the particles (of width equal to the inter-particle distance) and
-plotting the surface which bounds the highest 20\% values of the field. We can clearly see not only the percolated nature of both network, 
+plotting the surface which bounds the highest 20 \% values of the field. We can clearly see not only the percolated nature of both networks, 
 but also the difference in the smoothness of the network surface between amorphous and crystal gels. 
-The lower row depicts slabs of 5 particle diameter thickness,
-with particles colored according to the following code: red (fcc), green (hcp), white (crystal surface particles), cyan (liquid particles), magenta (gas particles). 
-We note that very few liquid particles remain and the network is almost crystalline.} 
+On the right, slabs of 5 particle diameter thickness are plotted,
+with particles coloured with the same colour scheme as in Fig.~\ref{fig:transitions}: red (crystal), green (gas particles), white (liquid particles). 
+We note that very few liquid particles remain and the network is almost crystalline.}} 
 \label{fig:network}
@@
 
-\begin{figure*}
+\begin{figure*}[!t]
  \centering
- \includegraphics{crystal}
+ \includegraphics[width=14cm]{fig4}
  % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735
- \caption{
-{\bf Crystal gel formation.}  
+ \caption{\textcolor{red}{{\bf Crystal gel formation.}  
 {\bf a,} Reconstructions from confocal coordinates at $\phi\approx 0.30$ and $c_p=0.38$ mg/g.
 Depth of view is $5\sigma$.
-Particles are drawn to scale and colored according to their phase.
-Dark red: crystal; orange: surface of crystal; yellow: liquid; blue: gas.
+Particles are drawn to scale and coloured according to their phase.
+Red: crystal; white: liquid; green: gas.
 {\bf b,} Fraction of particles for each phase. Gray vertical lines indicate the times shown in panel a. 
-{\bf c,} Transition probabilities for the same process as in panel b. Here crystal includes surface particles.}
+{\bf c,} Each particle during its trajectory changes between three states: gas, liquid and crystal. Direct crystallization accounts for 54\% of particles trajectories in which
+gas or liquid particles transition to the crystal state without liquid evaporation or crystal de-sublimation. The Bergeron process accounts for 35\% of particle trajectories in which liquid particles
+transition to the gas state before crystallizing. Ostwald ripening accounts for 11\% of particle trajectories in which crystal particle transition to the gas state before returning
+to the crystal state.}}
 \label{fig:transitions}

### 2016-03-11 phasediag.py
generate fig:phase_separation from the new data

@@ -8,11 +8,11 @@ def xp2th(cp, phi, qR=0.1, cpov=1.):
     return y2piv(cp/cpov/alpha(vf2f(phi), qR2q(qR)), qR)
 
 #radius of gyration in nm
-R = 82.
+R = 87.#82.
 #Molecular weight (Dalton or g/mol)
 Mw = 3.6e6
 #colloid diameter in nm
-sigma = 2200.
+sigma = 2100.#2200.
 #density of the solvent in g/mL
 d = 1.24
 

### 2016-03-11 crystal-gel/nature/crystal_gel2.tex
Add binodal and liquid spinodal. Arrow to mark the end of the tie lines. Fluid and crystal regions as gray areas. Update figure caption.

@@
 {\bf a,} Phase diagram.
-Dashed lines are fluid-solid equilibrium coexistence boundaries.
-The continuous line is the gas side of the metastable gas-liquid spinodal region with the black diamond marking the critical point.
+Gray areas are equilibrium fluid and crystal regions.
+The dashed and continuous lines are the metastable gas-liquid binodal and spinodal respectively.
+The black diamond marks the critical point.
 Empty black triangles are experimental point showing a fluid behaviour.
 The other symbols are the experimental points showing gel behaviour, consistently used in all panels.
+The dotted lines are the gas-liquid tie lines determined experimentally, with arrow heads indicating measured gas and liquid compositions.
 {\bf b,} Radius of gyration ($R_g$) as a function of cluster size ($s$) for all state points at percolation. The

### 2016-03-11 crystal-gel/nature/Supplementary.tex
Equation of state

@@
 
-\vspace{5mm}
-\noindent
-\emph{\bf Supplementary Movie 2}.
+\section*{Supplementary Movie 2}
 This movie shows the late stage after Supplemental Movie 1 with scanning duration of 27000~s (12300~$\tau$). 
+Please note that there is a large difference in the frame rate between Movie 1 and 2.
+
+\section*{Definition of states}
+The configurations obtained by tracking the position of the colloids in the samples
+are analyzed in order to determine whether a colloidal particles is in the
+crystal, surface, liquid or gas state. In the following we describe the classification
+criteria
+
+\subsection*{Crystal}
+In order to detect crystalline particle we use bond orientational analysis,
+as described in Ref.~\cite{russo2013interplay}.
+A $(2l+1)$ dimensional complex vector ($\mathbf{q}_l$) is defined for each
+particle $i$ as $q_{lm}(i)=\frac{1}{N_b(i)}\sum_{j=1}^{N_b(i)}
+Y_{lm}(\mathbf{\hat{r}_{ij}})$, where we set $l=12$, and $m$ is an integer that
+runs from $m=-l$ to $m=l$. The functions $Y_{lm}$ are the spherical harmonics
+and $\mathbf{\hat{r}_{ij}}$ is the normalised vector from the oxygen of
+molecule $i$ to the one of molecule $j$.  The sum goes over the first
+$N_b(i)=12$ neighbours of molecule $i$. This choice accounts for
+the first coordination shell of close packed crystals (fcc or hcp).
+The scalar product between $q_{6,m}$ of two particles
+is defined as $\mathbf{q}_{6}(i)\cdot\mathbf{q}_{6}(j)=\sum_m q_{6,m}(i)q_{6,m}(j)$.
+For each pair $i$ and $j$ of neighbouring particles we define a \emph{connection}
+if the scalar product $(\mathbf{q}_{6}(i)/|\mathbf{q}_{6}(i)|)\cdot(\mathbf{q}_{6}(j)/|\mathbf{q}_{6}(j)|)>0.7$.
+A molecule is then identified as crystalline if it has at least $7$ connected neighbours.
+
+The criteria outlined above are commonly used to identify crystalline arrangements in hard sphere systems.
+
+\subsection*{Surface, gas and liquid particles}
+
+\begin{figure}[!t]
+ \centering
+ \includegraphics[width=10cm]{./fractions.pdf}
+ % fractions.pdf: 681x503 pixel, 72dpi, 24.02x17.74 cm, bb=0 0 681 503
+ \caption{{\bf Fraction of particles for each phase at $\phi\approx 0.30$ and $c_p=0.38$ mg/g}. The liquid
+ fraction is depicted as a dashed black line, the gas fraction as a dotted-dashed green line, the surface
+ fraction as a double-dotted-dashed blue line, while the crystal with a continuous red line.}
+ \label{fig:fractions}
+\end{figure}
+
+Surface, gas and liquid particles are identified from the subset of non-crystalline particles depending
+on their neighbouring particles. A particle with at least two neighbouring crystal particle is classified as
+\emph{surface}. A particle that is neither solid or surface, and has at least four neighbouring particles
+is classified as \emph{liquid}. A particle which is neither solid, surface or liquid is classified as \emph{gas}.
+Fig.~\ref{fig:fractions} plots the fraction of each state for the state point at $\phi\approx 0.30$ and $c_p=0.38$ mg/g
+
+The distinction between gas and liquid phases is consistent with the analysis of volume fraction and composition
+of the gas and liquid phases just after phase separation (and before crystallization). To analyse the fraction
+of liquid and gas after phase separation we convert the position of particles into a contiunous field smearing
+the position of each particle with a gaussian field of variance $5$ pixels, and then mapping the field
+to a cubic grid with $10$ pixels spacing. In this way liquid and gas regions can be identified by
+an appropriate threshold of the field.
+
+\begin{figure}[!t]
+ \centering
+ \includegraphics[width=10cm]{./field.pdf}
+ % field.pdf: 681x483 pixel, 72dpi, 24.02x17.04 cm, bb=0 0 681 483
+ \caption{{\bf Distribution function of the gaussian field at $t=490\,\tau_B$.} The gaussian field is applied to
+ a configuration just after phase separation and before crystallization. The vertical dashed line represents the
+ field threshold $f=0.06$ used to distinguish between gas and liquid particles.}
+ \label{fig:field}
+\end{figure}
+
+Fig.~\ref{fig:field} shows the probability distribution of the field, where two peaks clearly indicate the presence
+of gas/liquid coexistence. By choosing a threshold over the field of $f=0.06$ (dashed line) we compute the fraction
+of particles in the solid and gas phases at $t=490\,\tau_B$, just after phase separation. The results are consistent
+with Fig.~\ref{fig:fractions}, where the number fraction of particles in the liquid state accounts for more than
+$90\%$ of the particles after phase separation.
+
+\section*{Phase diagram}
+
+The phase diagram is obtained in the framework of the generalized free volume theory.
+The colloids are considered as hard spheres and the polymers as an ideal gas.
+The center of the polymer cannot penetrate within $R$ of the surface of a colloid.
+Therefore the ideal gas is constrained to a free volume smaller than the total volume and that depends on the configuration (volume fraction) of the colloids and the polymer/colloid size ratio.
+
+\subsection*{Equation of state}
+
+This theory relies on an equation of state (EoS) for hard spheres to deduce phase boundaries of the colloids+polymers system.
+Canonically, the Carnahan-Starling (CS) EoS is used for the fluid and the equation of state by Hall for the crystal.
+CS is accurate for a wide range of volume fraction, but lacks a divergence below 1 although a divergence of the pressure is expected at the jamming point (random close packing $\phi_J\approx 0.64$).
+However CS is accurate up to 0.56.
+Therefore CS (coupled will Hall EoS) always describes accurately fluid-crystal coexistence since the fluid volume fraction is at most 0.495.
+
+The gas spinodal line is given by a local maximum in free energy, a local property of the equation of state.
+Therefore, the locally correct CS EoS is able to describe accurately the gas side of the spinodal region up to 0.56.
+However if the critical point is situated at a larger volume fraction, CS fails to locate it properly.
+In the same way, the liquid side of the spinodal as described by CS can reach unphysical volume fractions (above close packing).
+Even more annoying, the binodal line, given by equating pressures and chemical potentials in the gas and the liquid, is not a local property.
+Therefore, the position of the gas side of the binodal depends on the behaviour of the equation of state on the liquid side.
+Failure of CS at high volume fraction implies inaccurate binodal lines both gas and liquid.
+
+Here, our small polymer/colloid size ratio pushes the gas-liquid critical point to high volume fractions, exemplifying the failure of CS EoS.
+Adding a divergence at Jamming into CS equation of state is not trivial.
+Attempts by Lefebre~\cite{LeFevre1972} and Kolafa~\cite{Kolafa2004} match CS at low density, have the right divergence close to jamming but fare less than CS at intermediate volume fractions (0.5-0.56), the exact range where our critical point lies.
+Here we decided to use an equation by Liu~\cite{Liu2006a} that is completely equivalent to CS up to 0.56 and then show good agreement with simulation data near jamming.
+
+\section*{Transition probabilities}
+\begin{figure}[!t]
+ \centering
+ \includegraphics[width=14cm]{./sfig1.pdf}
+ % sfig1.pdf: 1398x1026 pixel, 72dpi, 49.32x36.19 cm, bb=0 0 1398 1026
+ \caption{{\bf Transition probabilities during the formation of the crystal-gel.}
+ {\bf a,} Transition probabilities from the liquid state;
+ {\bf b,} Transition probabilities from the gas state;
+ {\bf c,} Transition probabilities from the surface state;
+ {\bf d,} Transition probabilities from the crystal state. In all panels the possible
+ transition states are surface (blue), liquid (black), gas (green) and crystal (red).
+ See the text for a definition of the states.
+ }
+ \label{fig:probabilities}
+\end{figure}
+
+In Fig.~\ref{fig:probabilities} we plot the transition probabilities between the gas, liquid, crystal
+and surface states, as defined in the previous Section. Each panel represents the transition probability
+of a different state: {\bf a} for the liquid state, {\bf b} for the gas state, {\bf c} for the surface state,
+and {\bf d} for the crystal state. The results indicate that the liquid phase preferentially transforms into
+the surface state (as a precursor to crystallization), which accounts for the \emph{direct crystallization channel}
+that was described in the main text. A high percentage of liquid transforms also in the gas phase. The gas phase itself
+transforms back into liquid or into surface particles, a process which
+is linked to the \emph{Bergeron process}. These results support the idea that both channels are active
+crystallization pathways. Direct transformation of either liquid or gas into crystalline states is almost absent,
+indicating that our surface state indeed captures the precursor particles that attach to nuclei and later crystallize.
+Surface particles in fact, first transform preferentially into the liquid state, but at later time instead transform
+more prominently into the crystal state. The transformation of crystals into surface state, represents the first
+step of the \emph{Ostwald process}, but this channel is limited by the small rate of conversion of surface particles
+into the gas state.
+
+\bibliographystyle{naturemag2}
+\bibliography{biblio}
 

### 2016-03-12 crystal-gel/nature/crystal_gel2.tex
Update sigma, R and volume fractions. More honest description of how we get them.

@@
 %In the following we discuss all the stages of the formation of the crystal gels: phase separation, stress-driven agein and crystallization.
-The experimental data are taken at two different volume fractions ($\phi\approx 0.13$ and $\phi\approx 0.30$) and for different values
-of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ mg/g for $\phi\approx 0.30$ and $c_p=0.82,1.36$ mg/g for $\phi\approx 0.13$.
-In Fig.~\ref{fig:phase_separation}a, we superimpose the state points with the phase diagram, which we calculated from generalised free volume theory~\cite{Fleer2008} using the experimental values for the size of the colloids, the radius of gyration of the polymer, and the polymer concentration (see Methods).
+The experimental data are taken at two different volume fractions ($\phi\approx 0.14$ and $\phi\approx 0.33$) and for different values
+of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ mg/g for $\phi\approx 0.33$ and $c_p=0.82,1.36$ mg/g for $\phi\approx 0.14$.
+In Fig.~\ref{fig:phase_separation}a, we superimpose the state points with a theoretical phase diagram, which we calculated from generalised free volume theory~\cite{Fleer2008} using the experimental values for the size of the colloids, the radius of gyration of the polymer, and the polymer concentration (see Supplementary Methods).
 For each state point, scans were made at early times (every $10$~s, i.e. $4\tau_B$) and at the late stage (every $30$~s) of the gelation process.
@@
 All samples share the same early stages of spinodal decomposition. 
-A typical phase ordering process observed at $\phi\approx 0.30$ and $c_p=0.38$ mg/g can be seen in Supplementary Movies 1 and 2. 
+A typical phase ordering process observed at $\phi\approx 0.33$ and $c_p=0.38$ mg/g can be seen in Supplementary Movies 1 and 2. 
 Due to strong dynamical asymmetry between colloids and the solvent \cite{tanaka1999colloid}, the colloidal particles start aggregating, 
@@
 The figure shows that at low wavenumber $q$, the structure factor displays fractal scaling compatible with Guinier law, $S(q)\sim q^{-D}$.
-But a difference in the fractal dimension $D$ between the $\phi\approx 0.30$ and $c_p=0.38$ mg/g state point and other state points starts to become visible. In fact, while states
+But a difference in the fractal dimension $D$ between the $\phi\approx 0.33$ and $c_p=0.38$ mg/g state point and other state points starts to become visible. In fact, while states
 with high polymer concentration retain the exponent $D=2.5(3)$, 
 which is the random gel universality class exponent, as we also observed in the early stages of the gelation process (see Fig.~\ref{fig:phase_separation}b),
-the state point with low polymer concentration ($\phi\approx 0.30$ and $c_p=0.38$ mg/g) displays the biggest deviation from that exponent. In the figure we also
+the state point with low polymer concentration ($\phi\approx 0.33$ and $c_p=0.38$ mg/g) displays the biggest deviation from that exponent. In the figure we also
 plot the exponent $D=3$, which corresponds to volume growth, which we will show in the following analysis to be the correct exponent for this state. 
 The difference is due to the onset of crystallization in the low-polymer concentration sample. This is already evident in the high $q$ behaviour of
-the structure factors (Fig.~\ref{fig:phase_separation}c), where the diffraction planes appear as sharper peaks for $\phi\approx 0.30$ and $c_p=0.38$ mg/g.
+the structure factors (Fig.~\ref{fig:phase_separation}c), where the diffraction planes appear as sharper peaks for $\phi\approx 0.33$ and $c_p=0.38$ mg/g.
 
@@
 particle increases with its number of neighbours. Approaching close packing (12 neighbours), we clearly see two families of curves. The state point
-at $\phi\approx 0.30$ and $c_p=0.38$ mg/g reaches close packing at a volume fraction of around $\sim 68\%$, which is in reasonable agreement with the
+at $\phi\approx 0.33$ and $c_p=0.38$ mg/g reaches close packing at a volume fraction of around $\sim 68\%$, which is in reasonable agreement with the
 volume fraction of the stable crystal phase~\cite{ilett1995phase}. All other state points, instead, reach close packing at a volume fraction $\sim 64\%$, which is
@@
 We should run the code of Mathieau with size detection}).
-This result clearly indicates that the state point at $\phi\approx 0.30$ and $c_p=0.38$ mg/g follows a different arrest mechanism, in which phase separation
+This result clearly indicates that the state point at $\phi\approx 0.33$ and $c_p=0.38$ mg/g follows a different arrest mechanism, in which phase separation
 is arrested by crystallization and not by glassiness. Although our estimation of the absolute volume fraction may involves systematic errors of a few \%, the trend should be robust~\cite{Poon2012}.
@@
 A direct analysis of the colloidal positions at long times confirms that indeed the new structure obtained at
-$\phi\approx 0.30$ and $c_p=0.38$ mg/g has different morphological properties than usual colloidal gels.
+$\phi\approx 0.33$ and $c_p=0.38$ mg/g has different morphological properties than usual colloidal gels.
 We directly compare the two structures in Fig.~\ref{fig:network}, where Gaussian filtering is used to depict the network structure as a continuous field.
@@
 To confirm the formation of crystalline seeds, we adopt bond orientational analysis to detect which particles are in crystalline environments~\cite{russo2013interplay}. 
-With this analysis we reveal that crystallization events indeed start occurring for the state point $\phi\approx 0.30$ and $c_p=0.38$ mg/g, which has the lowest
+With this analysis we reveal that crystallization events indeed start occurring for the state point $\phi\approx 0.33$ and $c_p=0.38$ mg/g, which has the lowest
 polymer concentration of all state points. This is shown in Fig.~\ref{fig:crystals}a, where the black line indicates the average size of the crystallites, 
@@
 %\end{itemize}
-All other state points show only negligible signs of crystallization. In particular, for state points with $\phi\approx 0.30$, increasing
+All other state points show only negligible signs of crystallization. In particular, for state points with $\phi\approx 0.33$, increasing
 polymer concentration drastically reduces the amount of crystals. This is in agreement with the
@@
 To confirm the presence of the two processes, we plot in Fig.~\ref{fig:crystals}b the gyration radius of individual crystalline nuclei for the state point 
-$\phi\approx 0.30$ and $c_p=0.38$ mg/g. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with 
+$\phi\approx 0.33$ and $c_p=0.38$ mg/g. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with 
 a fractal dimension close to random percolation ($D=2.53$), while at large sizes it scales as $D=3$, as in compact crystal growth. This demonstrates that, while 
@@
 crystallization of the phase separated network.
-Supplementary Movie 2 shows the confocal microscopy video for the crystallizing sample at $\phi\approx 0.30$ and $c_p=0.38$ mg/g,
+Supplementary Movie 2 shows the confocal microscopy video for the crystallizing sample at $\phi\approx 0.33$ and $c_p=0.38$ mg/g,
 where the process of crystal growth by direct attachment of gas particles can be observed directly. 
@@
 Next we show in Fig.~\ref{fig:transitions}b the fraction of particles in each different phase for the crystallization trajectory at
-$\phi\approx 0.30$ and $c_p=0.38$ mg/g, after the liquid-gas phase separation has occurred.
+$\phi\approx 0.33$ and $c_p=0.38$ mg/g, after the liquid-gas phase separation has occurred.
 For the classification criteria see \emph{Methods}.
@@
 Depending on the polymer concentration, there are two possible arrest mechanisms. 
-(a) Crystallization, observed at $\phi\approx 0.30$ and $c_p=0.38$ mg/g, small crystalline nuclei appear inside the liquid network, reach the surface 
+(a) Crystallization, observed at $\phi\approx 0.33$ and $c_p=0.38$ mg/g, small crystalline nuclei appear inside the liquid network, reach the surface 
  of the liquid branches, and then grow by addition of particles from the gas phase. The final structure is a network of crystal droplets,
@@
 We use \textsc{pmma} (poly(methyl methacrylate)) colloids sterically stabilized with methacryloxypropyl terminated \textsc{pdms}(poly(dimethyl siloxane)) and fluorescently labelled with rhodamine isothiocyanate chemically bonded to the \textsc{pmma}.
-The average colloid diameter is 2.2 $\mu$m, measured via the diffusion constant in dilute conditions without polymer. The Brownian time is $\tau_B \approx 2.3$ s. 
+The diffusion constant in dilute conditions without polymer allows us to estimate the colloid diameter to $2.3\pm 0.05\, \mu$m. The Brownian time is $\tau_B \approx 2.3\,$s. 
 We assess that the size distribution of our particle is Gaussian with a polydispersity below 5\% via direct confocal measurements~\cite{Leocmach2013}.
@@
 Experiments are conducted at 27 $^\circ$C, some 80 $^\circ$C above the theta temperature in this solvents mixture~\cite{Royall2007}.
-A Flory scaling of the measurements of~\cite{lu2008gelation} yields a radius of gyration $R_g=82$ nm, and thus the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.075$ and the overlap mass fraction of polymer $2.1$~mg/g.
+A Flory scaling of the measurements of~\cite{lu2008gelation} yields a radius of gyration $R_g=76\pm5$ nm.
+
+We obtain the best fit between theoretical phase diagram and experimental data for $\sigma=2.25\,\mu$m and $R_g=80$ nm (see Supplementary method).
+Therefore the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.071$ and the overlap mass fraction of polymer $2.25$~mg/g.
 
@@
 
-To account for imprecision in particle localisation close to contact, we consider two particles bonded if their center-to-center distance is within 2.3 $\mu$m $>\sigma+2R_g$.
+To account for imprecision in particle localisation close to contact, we consider two particles bonded if their center-to-center distance is within 2.3 $\mu$m $\approx\sigma+2R_g$.
 
@@
 \caption{\textcolor{red}{{\bf Percolated network structures.} {\bf a:} gel; {\bf b:} crystal-gel. The gel state is obtained from the state
-point $\phi\approx 0.30$ and $c_p=1.07$ mg/g, while the crystal-gel is obtained at the same $\phi$ but
+point $\phi\approx 0.33$ and $c_p=1.07$ mg/g, while the crystal-gel is obtained at the same $\phi$ but
 at lower polymer concentration, $c_p=0.38$ mg/g. On the left, the network is represented as a continuous field that has
@@
  % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
-\caption{{\bf Characterization of the crystal-gel.} {\bf a,} Time evolution of the average size of the crystals (blue line) and the number of crystals (red line) for the state point $\phi\approx 0.30$ and $c_p=0.38$ mg/g.
+\caption{{\bf Characterization of the crystal-gel.} {\bf a,} Time evolution of the average size of the crystals (blue line) and the number of crystals (red line) for the state point $\phi\approx 0.33$ and $c_p=0.38$ mg/g.
 {\bf b,} Radius of gyration of crystalline nuclei at the late stages of gelation for the same state point.}
@@
  \caption{\textcolor{red}{{\bf Crystal gel formation.}  
-{\bf a,} Reconstructions from confocal coordinates at $\phi\approx 0.30$ and $c_p=0.38$ mg/g.
+{\bf a,} Reconstructions from confocal coordinates at $\phi\approx 0.33$ and $c_p=0.38$ mg/g.
 Depth of view is $5\sigma$.

### 2016-03-12 crystal-gel/nature/crystal_gel2.tex
Remove marker for critical point. Swap solid and dashed lines between spinodal and binodal.

@@
 {\bf a,} Phase diagram.
-Gray areas are equilibrium fluid and crystal regions.
-The dashed and continuous lines are the metastable gas-liquid binodal and spinodal respectively.
-The black diamond marks the critical point.
 Empty black triangles are experimental point showing a fluid behaviour.
@@
 The dotted lines are the gas-liquid tie lines determined experimentally, with arrow heads indicating measured gas and liquid compositions.
+Gray areas are theoretical equilibrium fluid and crystal regions.
+The solid and dashed lines are the metastable gas-liquid binodal and spinodal respectively.
 {\bf b,} Radius of gyration ($R_g$) as a function of cluster size ($s$) for all state points at percolation. The

### 2016-03-13 nature/response.tex
Improved the response.

@@
 and for providing us with three Referee reports.
-We were happy to see that all Referees appreciated our work, and at least two of them suggested
-resubmitting it after some technical clarifications. Referee 1 describes our results as ``very nice and beautiful'' but does not recommend our work
+We were happy to see that all Referees appreciated our work, and at least two of them recommended it after some technical clarifications.
+For this new version of the manuscript we have conducted additional experiments and analysis,
+which we believe address all points raised by the Referee.
+
+Referee 1 describes our results as ``very nice and beautiful'' but does not recommend our work
 based on the claim that it is not novel. He refers in particular to a series of works by the Edinburgh group in which the kinetics of phase
@@
 
-To address the technical questions of Referee 2 and 3 we have conducted novel experiments, that clearly demonstrate the reproducibility
-of our work. In particular, we believe that the concerns of Referee 3 are simply due to the fact that the two movies in Supplementary Informations
-are shot with different framerates. This gives the impression of abundance of gas phase, that probably led the Referee into thinking that the Ostwald
-ripening process is dominating. We have now made this issue more clear, and are confident that it will answer the Referee's concern fully.
-We also broadened our analysis, to include a direct study of the trajectory history to determine the different crystallization
-channels. As a result, new sections have been added to the Supplementary Informations, providing for a in-depth demonstration of the generality of our results.
+Referee 2 recommended our work, and had only minor technical questions. The most important one was regarding the phase boundaries
+which were computed with free-volume theory. In the present manuscript we have strengthened our results with a Gaussian field analysis
+that shows full consistency between theory and experimental observations.
+
+Referee 3 main concern comes from her/his direct inspection of our supplemental movies, where she/he observed an anomalous abundance
+of gas particles, that led her/him to suggest that Ostwald ripening was the main crystallization channel. But this observation
+was misguided by the fact that the two supplemental movies are shot at very different framerates. This was explained in the movies captions,
+but there was no time stamp on the movies, and we believe that this was the source of confusion, for which we apologize. While the first movies covers
+the first minutes of the process, the second movie shows the evolution over more than 7 hours, giving the impression
+of an abrupt increase of gas particles between the two. Also the movies display only a thin slice of the sample, inside which
+the crystalline nuclei appear to be separate from each other, which led the Referee to suggest Ostwald ripening. But looking at the process in three dimensions,
+it is without any doubt that crystalline nuclei are connected in a porous network. We show this very clear three-dimensional representation in a new Supplementary Figure.
+Together with a novel analysis which determines the crystallization pathway directly by looking
+at the particle trajectories, we believe that the Referee's issue is now fully solved.
 
@@
 
-First We would like to thank the Referee for her/his careful reading of our manuscript, for her/his comments.
+First we would like to thank the Referee for her/his careful reading of our manuscript, for her/his comments.
 
@@
 
-First We would like to thank the Referee for her/his careful reading of our manuscript, for providing us with valuable comments. We are very glad to see such kind positive comments on our work.
+First we would like to thank the Referee for her/his careful reading of our manuscript, for providing us with valuable comments. We are very glad to see such kind positive comments on our work.
 
@@
 We are afraid that the Reviewer’s impression of a sudden increase in free gas particles and a possible discontinuity between the two movies originatse from this difference in the frame rate.
+To address this issue we have now included time stamps on both Supplemental Movies.
 
@@
 We thus confirm the reproducibility of this type of crystallization pathway.
-We have added these additional experiments in the Supplementary Information.
-We would like to thank the Referee for stimulating
+An example of crystal-gel porour network obtained through our new capillary experiments
+is found in Fig. SXXX. All details about the experiments are included the Supplementary Information. We would like to thank the Referee for stimulating
 this additional investigation.
@@
 The Movies show only a thin slice of the system which corresponds to the particle that are in focus.
-To show the connectivity of the network clearly, we have also made Supplementary Movies of a thin slice with a finite thickness.
+To show the connectivity of the network clearly, we have also made Supplementary Figure SXXX, highlighting the crystal network.
+We also include movies of a thin slice with a finite thickness. \textcolor{red}{Did we do this?} 
 We hope that this will clarify the situation and convince the reviewer of the fact that our sample indeed forms a porous network of crystals.
@@
 
+\bigskip
+\doublespacing
+
 In the previous version of the manuscript, Fig. 4c included surface particles together with crystalline particles just to improve
@@
 we have added a section to Supplementary Information, with a new figure (Fig. S3) containing four panels that detail the transition
-probabilities between all states. In particular we note that, while being an active channel, the Ostwald ripening pathway is limited by
+probabilities between all states. These transition probabilities are measures by reconstructing the particles' trajectories, and
+computing probability of transition between different states at each time.
+In particular we note that, while being an active channel, the Ostwald ripening pathway is limited by
 the small conversion probability of surface particles into the gas phase.

### 2016-03-13 crystal-gel/nature/crystal_gel2.tex
Small addition to text.

@@
 The other symbols are the experimental points showing gel behaviour, consistently used in all panels.
-The dotted lines are the gas-liquid tie lines determined experimentally, with arrow heads indicating measured gas and liquid compositions.
+The dotted lines are the gas-liquid tie lines determined experimentally, with arrow heads indicating measured gas and liquid compositions (see Supplementary Information on the procedure to extract the compositions from confocal images).
 Gray areas are theoretical equilibrium fluid and crystal regions.

### 2016-03-13 crystal-gel/nature/Supplementary.tex
Description of the iterative process

@@
 
+\subsection*{Adjusting theoretical and experimental phase diagrams}
+
+The volume fraction of a colloidal suspension is notoriously difficult to estimate experimentally~\cite{Poon2012}.
+It depends on the cube of the colloid diameter and is thus extremely sensitive to any error in size determination.
+Therefore here we determine the volume fraction directly from the phase behaviour and deduce a precise measure of the (effective) colloid diameter from it.
+
+Our two constraints is (i) to have the final local volume fraction of the 12-neighboured particles in the crystallising sample to be equal to the theoretical equilibrium crystal volume fraction (ii) to have the theoretical spinodal line to correspond to the gel region boundary.
+Starting from our initial estimates of $\sigma$ and $R$ we iteratively converge to $R=80$~nm, $\sigma = 2.25\,\mu$m and a crystal at $0.723$.
+
 \subsection*{Equation of state}
 
-This theory relies on an equation of state (EoS) for hard spheres to deduce phase boundaries of the colloids+polymers system.
+The free volume theory relies on an equation of state (EoS) for hard spheres to deduce phase boundaries of the colloids+polymers system.
 Canonically, the Carnahan-Starling (CS) EoS is used for the fluid and the equation of state by Hall for the crystal.
@@
 Therefore, the locally correct CS EoS is able to describe accurately the gas side of the spinodal region up to 0.56.
-However if the critical point is situated at a larger volume fraction, CS fails to locate it properly.
-In the same way, the liquid side of the spinodal as described by CS can reach unphysical volume fractions (above close packing).
-Even more annoying, the binodal line, given by equating pressures and chemical potentials in the gas and the liquid, is not a local property.
-Therefore, the position of the gas side of the binodal depends on the behaviour of the equation of state on the liquid side.
-Failure of CS at high volume fraction implies inaccurate binodal lines both gas and liquid.
+The iterative process described above can be done only using CS EoS.
 
-Here, our small polymer/colloid size ratio pushes the gas-liquid critical point to high volume fractions, exemplifying the failure of CS EoS.
+However, the liquid side of the binodal and spinodal lines obtained from CS EoS reach unphysically high volume fractions (above close packing), making comparisons with the measured liquid compositions impossible.
+
+Indeed our small polymer/colloid size ratio pushes the gas-liquid critical point to high volume fractions, exemplifying the failure of CS EoS.
 Adding a divergence at Jamming into CS equation of state is not trivial.
@@
 Here we decided to use an equation by Liu~\cite{Liu2006a} that is completely equivalent to CS up to 0.56 and then show good agreement with simulation data near jamming.
+
+The Liu EoS allows us to draw a phase diagram that is in qualitative agreement with the measured liquid compositions.
 

### 2016-03-13 nature/crystal_gel2.tex
Change Fig1d to time evolution of 12-neighboured particles local volume fraction. Use the asymptotic value for the crystallising sample as a definition for the crystal volume fraction.

@@
 The figure shows that at low wavenumber $q$, the structure factor displays fractal scaling compatible with Guinier law, $S(q)\sim q^{-D}$.
-But a difference in the fractal dimension $D$ between the $\phi\approx 0.30$ and $c_p=0.38$ mg/g state point and other state points starts to become visible. In fact, while states
+But a difference in the fractal dimension $D$ between the $\phi\approx 0.33$ and $c_p=0.38$ mg/g state point and other state points starts to become visible. In fact, while states
 with high polymer concentration retain the exponent $D=2.5(3)$, 
 which is the random gel universality class exponent, as we also observed in the early stages of the gelation process (see Fig.~\ref{fig:phase_separation}b),
-the state point with low polymer concentration ($\phi\approx 0.30$ and $c_p=0.38$ mg/g) displays the biggest deviation from that exponent. In the figure we also
+the state point with low polymer concentration ($\phi\approx 0.33$ and $c_p=0.38$ mg/g) displays the biggest deviation from that exponent. In the figure we also
 plot the exponent $D=3$, which corresponds to volume growth, which we will show in the following analysis to be the correct exponent for this state. 
 The difference is due to the onset of crystallization in the low-polymer concentration sample. This is already evident in the high $q$ behaviour of
-the structure factors (Fig.~\ref{fig:phase_separation}c), where the diffraction planes appear as sharper peaks for $\phi\approx 0.30$ and $c_p=0.38$ mg/g.
+the structure factors (Fig.~\ref{fig:phase_separation}c), where the diffraction planes appear as sharper peaks for $\phi\approx 0.33$ and $c_p=0.38$ mg/g.
 
-To gain additional insight, in Fig.~\ref{fig:phase_separation}d we plot the average local volume fraction, computed
-through Voronoi diagrams, as a function of the number of neighbours. For all state points, we can see that the average volume fraction of a colloidal
+To gain additional insight, in Fig.~\ref{fig:phase_separation}d we plot the average local volume fraction as a function of the number of neighbours.
+\textcolor{red}{The volume fraction is obtained by computing the Voronoi diagram of each configuration,
+which uniquely assigns a volume to each colloidal particle.}
+For all state points, we can see that the average volume fraction of a colloidal
 particle increases with its number of neighbours. Approaching close packing (12 neighbours), we clearly see two families of curves. The state point
-at $\phi\approx 0.30$ and $c_p=0.38$ mg/g reaches close packing at a volume fraction of around $\sim 68\%$, which is in reasonable agreement with the
+at $\phi\approx 0.33$ and $c_p=0.38$ mg/g reaches close packing at a volume fraction of around $\sim 68\%$, which is in reasonable agreement with the
 volume fraction of the stable crystal phase~\cite{ilett1995phase}. All other state points, instead, reach close packing at a volume fraction $\sim 64\%$, which is
@@
 Experiments are conducted at 27 $^\circ$C, some 80 $^\circ$C above the theta temperature in this solvents mixture~\cite{Royall2007}.
-A Flory scaling of the measurements of~\cite{lu2008gelation} yields a radius of gyration $R_g=82$ nm, and thus the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.075$ and the overlap mass fraction of polymer $2.1$~mg/g.
+A Flory scaling of the measurements of~\cite{lu2008gelation} yields a radius of gyration $R_g=76\pm5$ nm.
+
+We obtain the best fit between theoretical phase diagram and experimental data for $\sigma=2.24\,\mu$m and $R_g=80$ nm (see Supplementary method).
+Therefore the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.071$ and the overlap mass fraction of polymer $2.25$~mg/g.
 
@@
 {\bf a,} Phase diagram.
-Dashed lines are fluid-solid equilibrium coexistence boundaries.
-The continuous line is the gas side of the metastable gas-liquid spinodal region with the black diamond marking the critical point.
 Empty black triangles are experimental point showing a fluid behaviour.
 The other symbols are the experimental points showing gel behaviour, consistently used in all panels.
+The dotted lines are the gas-liquid tie lines determined experimentally, with arrow heads indicating measured gas and liquid compositions (see Supplementary Information on the procedure to extract the compositions from confocal images).
+Gray areas are theoretical equilibrium fluid and crystal regions.
+The solid and dashed lines are the metastable gas-liquid binodal and spinodal respectively.
 {\bf b,} Radius of gyration ($R_g$) as a function of cluster size ($s$) for all state points at percolation. The
@@
 {\bf c,} Structure factors for the late stages of the gelation process. The gray and black lines represents respectively random gelation and compact fractal dimensions.
-{\bf d,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle for the state points in the very late stage.
-Horizontal lines stress the value at 12 neighbours.
+{\bf d,} Late stage of the evolution of local volume fraction around colloidal particles having 12 nearest neighbours.
  }

### 2016-03-15 crystal-gel/nature/response.tex
Dear all,
Thank you for your efforts. I made corrections to the main text, the Supplementary Information, and the response letter. Please check them carefully. There are also some places we need to fix. Thanks!! Hajime

--- 2016-03-13 nature/response.tex
@@
-\documentclass[11pt]{article}
+\documentclass[11pt,a4paper]{article}
 \usepackage{graphicx}
@@
 \usepackage{setspace}
+
+\usepackage[top=35mm,bottom=35mm,left=30mm,right=30mm]{geometry}
 
@@
 \noindent
-Dear Editor,
+{\bf Dear Pep,}
 \vskip 0.3cm
-we would like to thank you for your comments to our manuscript ``A novel route to the spontaneous formation of porous crystals via viscoelastic phase separation'' (NM15113574),
+We would like to thank you for your comments to our manuscript ``A novel route to the spontaneous formation of porous crystals via viscoelastic phase separation'' (NM15113574),
 and for providing us with three Referee reports.
@@
 For this new version of the manuscript we have conducted additional experiments and analysis,
-which we believe address all points raised by the Referee.
+which we believe address all points raised by the Referees.
 
 Referee 1 describes our results as ``very nice and beautiful'' but does not recommend our work
-based on the claim that it is not novel. He refers in particular to a series of works by the Edinburgh group in which the kinetics of phase
+based on the claim that it is not novel. She/He refers in particular to a series of works by the Edinburgh group in which the kinetics of phase
 transitions is classified according to the free energy landscape. These work are by themselves very important and relevant, but contrary
-to the Referee claim, do not contain a description or results on the crystallization channel we describe in our work.
+to the Referee's claim, do not contain a description or results on the crystallization channel we describe in our work.
 Both Referees 2 and 3 have appreciated this fact.
@@
 
-Referee 3 main concern comes from her/his direct inspection of our supplemental movies, where she/he observed an anomalous abundance
+Referee 3's main concern comes from her/his direct inspection of our supplemental movies, where she/he observed an anomalous abundance
 of gas particles, that led her/him to suggest that Ostwald ripening was the main crystallization channel. But this observation
@@
 \noindent
-Yours Sincerely,
+{\bf Yours Sincerely,
 \vskip 0.3cm
@@
 \indent Mathieu Leocmach\\
-\indent Hajime Tanaka
+\indent Hajime Tanaka}
 
@@
 
-Below we will answer the comments one by one.
+We believe that many monoatomic and single-component molecular systems can satisfy all the conditions (i)-(iv) in a certain range of the temperature and pressure. However, the reviewer 
+is right in the sense that there may be a practical difficulty in realizing condition (iii), for example, for noble metals. 
+Condition (iii) is necessary to create a porous crystalline structure with large porosity. However, even if condition (iii) is not met, bicontinuous phase-separation structures 
+can be formed near a symmetry composition. Thus, we can still form a porous crystal structure even in such a case. We have added the following sentences to mention this. 
+We thank the reviewer for her/his thoughtful comment on this important point.
+
+``{\it For monoatomic systems such as noble metals, condition (iii) may not be satisfied easily. 
+To access glassy slow dynamics in a liquid phase, we may need a deep quench at a high pressure. However, we note that even without strong dynamic asymmetry 
+due to glassiness, bicontinuous phase separation can take place between 35-85 volume \% of the liquid phase in ordinary gas/liquid phase separation; 
+and, thus, porous crystalline structures can be formed, although a thin network structure may be difficult to be formed.}''
 
@@
 This statement means that crystallites are formed in either the gas or the liquid phase.
-Thus, this is clearly different from what we found in our paper as a new pathway involving all the three phases in a non-trivial manner.
-Nonetheless, our statement "kinetic path via the gas phase (liquid $\rightarrow$ gas $\rightarrow$ crystal) has been completely overlooked so far" was obviously too strong.
-We have replaced “completely” by “largely” and have cited these pioneering works, which considered the role of the complex free energy landscape for the first time.
+As kindly pointed out by the reviewer, thus, this is clearly different from what we found in our paper as a new pathway involving all the three phases in a non-trivial manner.
+Nonetheless, our statement ``kinetic path via the gas phase (liquid $\rightarrow$ gas $\rightarrow$ crystal) has been completely overlooked so far'' was obviously too strong.
+We have replaced ``completely'' by ``largely'' and have cited these pioneering works, which considered the role of the complex free energy landscape for the first time.
 
@@
 We hope that this would make the situation much clearer.
-\textcolor{red}{Mathieu could you check this point?}
+\textcolor{blue}{In relation to this, we have also re-estimated the volume fraction of each phase more accurately
+by a method described in Supplementary Information.
+We think that this increases the reliability of our estimate of the volume fraction.
+The end points of the above-mentioned tie lines are determined in this way.
+(Mathieu, John: please add any useful information.) }
 
@@
 
-By "random gelation universality class" we mean three-dimensional percolation universality class (which applies for example to randomly branched polymers).
+By ``random gelation universality class'' we mean three-dimensional percolation universality class (which applies for example to randomly branched polymers).
 It is known that three-dimensional random percolation clusters and diffusion-limited aggregation clusters have both about the same fractal dimension of about 2.5, for which we have cited the review article by Herrmann.
 We agree that DLCA and RLCA provide smaller fractal dimensions; however, DLCA is usually observed at much lower volume fraction and with stronger bonding energy (after bonding there is little diffusion, which leads to the formation of really fractal objects) than our case.
-For our starting volume fraction of about $\phi\approx 0.30$ and polymer concentrations, we have found that the fractal dimension
+For our starting volume fraction of about $\phi\approx 0.30$ and polymer concentrations, we have found that the fractal (or effective) dimension
 to be more consistent with random percolation. We hope that this clarifies the situation.
@@
 We thank the Reviewer for pointing out this mistake. We have corrected it as suggested.
+
+\vspace{1cm}
 
@@
 An example of crystal-gel porour network obtained through our new capillary experiments
-is found in Fig. SXXX. All details about the experiments are included the Supplementary Information. We would like to thank the Referee for stimulating
+is found in \textcolor{blue}{Fig. SXXX}. All details about the experiments are included the Supplementary Information. We would like to thank the Referee for stimulating
 this additional investigation.
@@
 The Movies show only a thin slice of the system which corresponds to the particle that are in focus.
+As can be seen in Fig. 2b (left panel), the crystal gel has a bicontinuous porous structure. 
 To show the connectivity of the network clearly, we have also made Supplementary Figure SXXX, highlighting the crystal network.
-We also include movies of a thin slice with a finite thickness. \textcolor{red}{Did we do this?} 
+We also include movies of a thin slice with a finite thickness. \textcolor{blue}{Please prepare movie(s).} 
 We hope that this will clarify the situation and convince the reviewer of the fact that our sample indeed forms a porous network of crystals.
@@
 In particular we note that, while being an active channel, the Ostwald ripening pathway is limited by
-the small conversion probability of surface particles into the gas phase.
+the small conversion probability of surface particles into the gas phase (please see also our reply to the next comment).
 
@@
 \doublespacing
-We have followed the kind suggestion of the reviewer.
+We have added a time stamp, although we did not merge the two movies due to the file size problem.  
+We have also made the difference in the frame rate more explicit in Supplementary Movie Captions.
 
@@
 
+\vspace{1cm}
+
 We hope that the Reviewer would find that the revised manuscript is now suitable for publication in Nature Materials.

### 2016-03-15 nature/Supplementary.tex
Dear all,
Thank you for your efforts. I made corrections to the main text, the Supplementary Information, and the response letter. Please check them carefully. There are also some places we need to fix. Thanks!! Hajime

@@
 
-\section*{Supplementary Movie 1}
-This movie shows the early stage of gelation process of $\phi\approx 0.30$ and $c_p=0.38$ mg/g, from the original confocal images at the same z-position as Fig.~3a.
-Total duration is 1160 s (530~$\tau$).
-We quickly started scanning the sample after exchanging the salt-reservoir with the solvent including salt.
-At the beginning, salt had not reached the scanning volume and colloids were dispersed by electrostatic repulsion.
-At 0:02 in the movie, salt diffused into the scanning volume and screened the surface charges of colloids.
-Then, colloids immediately initiated phase separation by short-ranged attractions.
-
-\section*{Supplementary Movie 2}
-This movie shows the late stage after Supplemental Movie 1 with scanning duration of 27000~s (12300~$\tau$). 
-Please note that there is a large difference in the frame rate between Movie 1 and 2.
-
 \section*{Definition of states}
 The configurations obtained by tracking the position of the colloids in the samples
-are analyzed in order to determine whether a colloidal particles is in the
+are analysed in order to determine whether a colloidal particles is in the
 crystal, surface, liquid or gas state. In the following we describe the classification
@@
 
-\begin{figure}[!t]
+\begin{figure}[!b]
  \centering
@@
 
+\section*{Phase diagram}
+
+The phase diagram is obtained in the framework of the generalized free volume theory.
+The colloids are considered as hard spheres and the polymers as an ideal gas.
+The center of the polymer cannot penetrate within $R$ of the surface of a colloid.
+Therefore the ideal gas is constrained to a free volume smaller than the total volume and that depends on the configuration (volume fraction) of the colloids and the polymer/colloid size ratio.
+
+\subsection*{Adjusting theoretical and experimental phase diagrams}
+
+The volume fraction of a colloidal suspension is notoriously difficult to estimate experimentally~\cite{Poon2012}.
+It depends on the cube of the colloid diameter and is thus extremely sensitive to any error in size determination.
+Therefore here we determine the volume fraction directly from the phase behaviour and deduce a precise measure of the (effective) colloid diameter from it.
+
+Our two constraints is (i) to have the final local volume fraction of the 12-neighboured particles in the crystallising sample to be equal to the theoretical equilibrium crystal volume fraction and (ii) to have the theoretical spinodal line to correspond to the gel region boundary.
+Starting from our initial estimates of $\sigma$ and $R$ we iteratively converge to $R=80$~nm, $\sigma = 2.24\,\mu$m and a crystal at $0.723$.
+
+\subsection*{Equation of state}
+
+The free volume theory relies on an equation of state (EoS) for hard spheres to deduce phase boundaries of the colloids+polymers system.
+Canonically, the Carnahan-Starling (CS) EoS is used for the fluid and the equation of state by Hall for the crystal.
+CS is accurate for a wide range of volume fraction, but lacks a divergence below 1 although a divergence of the pressure is expected at the jamming point (random close packing $\phi_J\approx 0.64$).
+However CS is accurate up to 0.56.
+Therefore CS (coupled will Hall EoS) always describes accurately fluid-crystal coexistence since the fluid volume fraction is at most 0.495.
+
+The gas spinodal line is given by a local maximum in free energy, a local property of the equation of state.
+Therefore, the locally correct CS EoS is able to describe accurately the gas side of the spinodal region up to 0.56.
+The iterative process described above can be done only using CS EoS.
+
+However, the liquid side of the binodal and spinodal lines obtained from CS EoS reach unphysically high volume fractions (above close packing), making comparisons with the measured liquid compositions impossible.
+
+Indeed our small polymer/colloid size ratio pushes the gas-liquid critical point to high volume fractions, exemplifying the failure of CS EoS.
+Adding a divergence at Jamming into CS equation of state is not trivial.
+Attempts by Lefebre~\cite{LeFevre1972} and Kolafa~\cite{Kolafa2004} match CS at low density, have the right divergence close to jamming but far less than CS at intermediate volume fractions (0.5-0.56), the exact range where our critical point lies.
+Here we decided to use an equation by Liu~\cite{Liu2006a} that is completely equivalent to CS up to 0.56 and then show good agreement with simulation data near jamming.
+
+The Liu EoS allows us to draw a phase diagram that is in qualitative agreement with the measured liquid compositions.
+
 \section*{Transition probabilities}
-\begin{figure}[!t]
+\begin{figure}[!b]
  \centering
@@
 Surface particles in fact, first transform preferentially into the liquid state, but at later time instead transform
-more prominently into the crystal state. The transformation of crystals into surface state, represents the first
+more prominently into the crystal state. The transformation of crystals into the surface state, represents the first
 step of the \emph{Ostwald process}, but this channel is limited by the small rate of conversion of surface particles
@@
 
+\clearpage
+\section*{Supplementary Movie 1}
+This movie shows the early stage of gelation process of $\phi\approx 0.30$ and $c_p=0.38$ mg/g, from the original confocal images at the same z-position as Fig.~3a.
+Total duration is 1160 s (530~$\tau$).
+We quickly started scanning the sample after exchanging the salt-reservoir with the solvent including salt.
+At the beginning, salt had not reached the scanning volume and colloids were dispersed by electrostatic repulsion.
+At 0:02 in the movie, salt diffused into the scanning volume and screened the surface charges of colloids.
+Then, colloids immediately initiated phase separation by short-ranged attractions.
+
+\section*{Supplementary Movie 2}
+This movie shows the late stage after Supplemental Movie 1 with scanning duration of 27000~s (12300~$\tau$). 
+Please note that there is a large difference in the frame rate between Movie 1 and 2.
+
 \end{document}

### 2016-03-15 nature/crystal_gel2.tex
Dear all,
Thank you for your efforts. I made corrections to the main text, the Supplementary Information, and the response letter. Please check them carefully. There are also some places we need to fix. Thanks!! Hajime

@@
 To gain additional insight, in Fig.~\ref{fig:phase_separation}d we plot the average local volume fraction as a function of the number of neighbours.
-\textcolor{red}{The volume fraction is obtained by computing the Voronoi diagram of each configuration,
+\textcolor{blue}{The volume fraction is obtained by computing the Voronoi diagram of each configuration,
 which uniquely assigns a volume to each colloidal particle.}
@@
 
-\textcolor{red}{Next we analyse in detail the crystallization trajectory, distinguishing between the different process responsible for the
+\textcolor{blue}{Next we analyse in detail the crystallization trajectory, distinguishing between the different processes responsible for the
 crystallization of the phase separated network.
@@
 From the trajectory, we extract slabs of 10 particle diameter thickness for configurations at different times, shown in Fig.~\ref{fig:transitions}a. 
-Particles are coloured according to their phases (gas, liquid, and crystal). For the classification criteria see the \emph{Methods} section. The figure shows the first nucleation 
+Particles are coloured according to their phases (gas, liquid, and crystal). For the classification criteria see Supplementary Information. The figure shows the first nucleation 
 events inside the liquid network (left panel).
@@
 Finally the different nuclei coalesce (right panel).
-Note that the network pattern changes between $t=2055\,\tau_B$ (left panel) and $t=5577,\tau_B$ (middle panel): not all liquid
-branches of the network crystallize due to evaporation. The crystal-gel structure instead persists, as seen by comparing $t=5577,\tau_B$ (middle panel) and $t=12098,\tau_B$ (right panel).}
+\textcolor{red}{Note that the network pattern changes its surface smoothness (??) between $t=2055\,\tau_B$ (left panel) and $t=5577,\tau_B$ (middle panel) (see, e.g., Fig2a):
+not all liquid branches of the network crystallize due to evaporation (?).
+The crystal-gel structure instead persists, as seen by comparing $t=5577,\tau_B$ (middle panel) and $t=12098,\tau_B$ (right panel).
+{\bf (Hajime: this part is not so clear for me.)}}}
 
-\textcolor{red}{
+\textcolor{blue}{
 Next we show in Fig.~\ref{fig:transitions}b the fraction of particles in each different phase for the crystallization trajectory at
 $\phi\approx 0.33$ and $c_p=0.38$ mg/g, after the liquid-gas phase separation has occurred.
-For the classification criteria see \emph{Methods}.
 The process of crystallization is characterized 
@@
 
-\textcolor{red}{
-In Supplementary Information we confirm these results with a detailed analysis of transition probabilities, including a state for surface particles that are in contact.
+\textcolor{blue}{
+In Supplementary Information we confirm these results with a detailed analysis of transition probabilities, including a state for surface particles that are in contact 
 to crystalline particles.
@@
 This can be seen, for example, by looking at the phase diagram of a Lennard-Jones liquid \cite{lodge1997brownian}, which represents many molecular systems 
-without specific directional interactions.
+without specific directional interactions. \textcolor{blue}{For monoatomic systems such as noble metals, condition (iii) may not be satisfied easily. 
+To access glassy slow dynamics in a liquid phase, we may need a deep quench at a high pressure. However, we note that even without strong dynamic asymmetry 
+due to glassiness, bicontinuous phase separation can take place between 35-85 volume \% of the liquid phase in ordinary gas/liquid phase separation; 
+and, thus, porous crystalline structures can be formed, although a thin network structure may be difficult to be formed.}
 
@@
 The other symbols are the experimental points showing gel behaviour, consistently used in all panels.
-The dotted lines are the gas-liquid tie lines determined experimentally, with arrow heads indicating measured gas and liquid compositions (see Supplementary Information on the procedure to extract the compositions from confocal images).
+\textcolor{blue}{The dotted lines are the gas-liquid tie lines determined experimentally, with arrow heads indicating measured gas and liquid compositions (see Supplementary Information on the procedure to extract the compositions from confocal images).
 Gray areas are theoretical equilibrium fluid and crystal regions.
-The solid and dashed lines are the metastable gas-liquid binodal and spinodal respectively.
+The solid and dashed lines are the metastable gas-liquid binodal and spinodal respectively.}
 {\bf b,} Radius of gyration ($R_g$) as a function of cluster size ($s$) for all state points at percolation. The
@@
 {\bf c,} Structure factors for the late stages of the gelation process. The gray and black lines represents respectively random gelation and compact fractal dimensions.
-{\bf d,} Late stage of the evolution of local volume fraction around colloidal particles having 12 nearest neighbours.
+\textcolor{blue}{{\bf d,} Late stage of the evolution of local volume fraction around colloidal particles having 12 nearest neighbours.}
  }
@@
  % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
-\caption{\textcolor{red}{{\bf Percolated network structures.} {\bf a:} gel; {\bf b:} crystal-gel. The gel state is obtained from the state
+\caption{\textcolor{blue}{{\bf Percolated network structures.} {\bf a:} gel; {\bf b:} crystal-gel. The gel state is obtained from the state
 point $\phi\approx 0.33$ and $c_p=1.07$ mg/g, while the crystal-gel is obtained at the same $\phi$ but
@@
 with particles coloured with the same colour scheme as in Fig.~\ref{fig:transitions}: red (crystal), green (gas particles), white (liquid particles). 
-We note that very few liquid particles remain and the network is almost crystalline.}} 
+We note that very few liquid particles remain and the network is almost crystalline. 
+\textcolor{red}{\bf Hajime: we need to put length and time information for each panel.} 
+}} 
 \label{fig:network}
@@
  % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735
- \caption{\textcolor{red}{{\bf Crystal gel formation.}  
+ \caption{\textcolor{blue}{{\bf Crystal gel formation.}  
 {\bf a,} Reconstructions from confocal coordinates at $\phi\approx 0.33$ and $c_p=0.38$ mg/g.
@@
 Red: crystal; white: liquid; green: gas.
+\textcolor{red}{\bf Hajime: we need to put length information for images.}
 {\bf b,} Fraction of particles for each phase. Gray vertical lines indicate the times shown in panel a. 

### 2016-03-16 nature/crystal_gel2.tex
Updated the text and added the capillary snapshots.

@@
 At the same time liquid regions that have not crystallized start evaporating, and the gas phase contributes to the growth of nuclei (middle panel).
-Finally the different nuclei coalesce (right panel).
-\textcolor{red}{Note that the network pattern changes its surface smoothness (??) between $t=2055\,\tau_B$ (left panel) and $t=5577,\tau_B$ (middle panel) (see, e.g., Fig2a):
-not all liquid branches of the network crystallize due to evaporation (?).
-The crystal-gel structure instead persists, as seen by comparing $t=5577,\tau_B$ (middle panel) and $t=12098,\tau_B$ (right panel).
-{\bf (Hajime: this part is not so clear for me.)}}}
+Finally the different nuclei coalesce (right panel).}
+\textcolor{red}{Note that the topology of the crystal network found at late times, i.e.
+$t=5577,\tau_B$ (middle panel) and $t=12098,\tau_B$ (right panel) in Fig.~\ref{fig:transitions}a, differs significantly from that of the liquid network at early stages, i.e.
+$t=2055\,\tau_B$ (left panel) in Fig.~\ref{fig:transitions}a.
+This suggests that the crystal network doesn't form only by direct freezing of the liquid, but also by other mechanisms which involve the evaporation of the liquid (Bergeron process) and the sublimation of small crystals (Ostwald ripening).
+In the following we will investigate these different crystallization mechanisms in depth.}
 
@@
 \caption{\textcolor{blue}{{\bf Percolated network structures.} {\bf a:} gel; {\bf b:} crystal-gel. The gel state is obtained from the state
-point $\phi\approx 0.33$ and $c_p=1.07$ mg/g, while the crystal-gel is obtained at the same $\phi$ but
-at lower polymer concentration, $c_p=0.38$ mg/g. On the left, the network is represented as a continuous field that has
+point $\phi\approx 0.33$ and $c_p=1.07$ mg/g at $t=400$ min after sample preparation, while the crystal-gel is obtained at the same $\phi$ but
+at lower polymer concentration, $c_p=0.38$ mg/g at $t=468$ min after sample preparation. On the left, the network is represented as a continuous field that has
 been obtained by applying a gaussian filter to the position of the particles (of width equal to the inter-particle distance) and
@@
 but also the difference in the smoothness of the network surface between amorphous and crystal gels. 
-On the right, slabs of 5 particle diameter thickness are plotted,
+On the right, slabs of 4.5 $\mu$m thickness and 41 $\mu$m width are plotted,
 with particles coloured with the same colour scheme as in Fig.~\ref{fig:transitions}: red (crystal), green (gas particles), white (liquid particles). 
 We note that very few liquid particles remain and the network is almost crystalline. 
-\textcolor{red}{\bf Hajime: we need to put length and time information for each panel.} 
 }} 
@@
 {\bf a,} Reconstructions from confocal coordinates at $\phi\approx 0.33$ and $c_p=0.38$ mg/g.
-Depth of view is $5\sigma$.
+Depth of view is 9.0 $\mu$m, while the later dimension is 41 $\mu$m.
 Particles are drawn to scale and coloured according to their phase.
 Red: crystal; white: liquid; green: gas.
-\textcolor{red}{\bf Hajime: we need to put length information for images.}
 {\bf b,} Fraction of particles for each phase. Gray vertical lines indicate the times shown in panel a. 

### 2016-03-18 crystal-gel/nature/crystal_gel2.tex
Hajime, Slight changes in the main text

--- 2016-03-12 crystal-gel/nature/crystal_gel2.tex
@@
 Finally the different nuclei coalesce (right panel).
-Note that the network pattern changes between $t=2055\,\tau_B$ (left panel) and $t=5577,\tau_B$ (middle panel): not all liquid
-branches of the network crystallize due to evaporation. The crystal-gel structure instead persists, as seen by comparing $t=5577,\tau_B$ (middle panel) and $t=12098,\tau_B$ (right panel).}
+Note that the topology of the crystal network found at late times, i.e.
+$t=5577\,\tau_B$ (middle panel) and $t=12098\,\tau_B$ (right panel) in Fig.~\ref{fig:transitions}a, differs significantly from that of the liquid network at early stages, i.e.
+$t=2055\,\tau_B$ (left panel) in Fig.~\ref{fig:transitions}a.
+This suggests that the crystal network does not form only by direct freezing of the liquid, but also by other mechanisms which involve the evaporation of the liquid (Bergeron process) and the sublimation of small crystals (Ostwald ripening).
+In the following we will investigate these different crystallization mechanisms in depth.}
 
-\textcolor{red}{
+\textcolor{blue}{
 Next we show in Fig.~\ref{fig:transitions}b the fraction of particles in each different phase for the crystallization trajectory at
 $\phi\approx 0.33$ and $c_p=0.38$ mg/g, after the liquid-gas phase separation has occurred.
-For the classification criteria see \emph{Methods}.
 The process of crystallization is characterized 
@@
 
-We obtain the best fit between theoretical phase diagram and experimental data for $\sigma=2.25\,\mu$m and $R_g=80$ nm (see Supplementary method).
+We obtain the best fit between theoretical phase diagram and experimental data for $\sigma=2.24\,\mu$m and $R_g=80$ nm (see Supplementary method).
 Therefore the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.071$ and the overlap mass fraction of polymer $2.25$~mg/g.
@@
 The other symbols are the experimental points showing gel behaviour, consistently used in all panels.
-The dotted lines are the gas-liquid tie lines determined experimentally, with arrow heads indicating measured gas and liquid compositions.
+\textcolor{blue}{The dotted lines are the gas-liquid tie lines determined experimentally, with arrow heads indicating measured gas and liquid compositions (see Supplementary Information on the procedure to extract the compositions from confocal images).
 Gray areas are theoretical equilibrium fluid and crystal regions.
-The solid and dashed lines are the metastable gas-liquid binodal and spinodal respectively.
+The solid and dashed lines are the metastable gas-liquid binodal and spinodal respectively.}
 {\bf b,} Radius of gyration ($R_g$) as a function of cluster size ($s$) for all state points at percolation. The
@@
 {\bf c,} Structure factors for the late stages of the gelation process. The gray and black lines represents respectively random gelation and compact fractal dimensions.
-{\bf d,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle for the state points in the very late stage.
-Horizontal lines stress the value at 12 neighbours.
+\textcolor{blue}{{\bf d,} Late stage of the evolution of local volume fraction around colloidal particles having 12 nearest neighbours.}
  }
@@
  % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
-\caption{\textcolor{red}{{\bf Percolated network structures.} {\bf a:} gel; {\bf b:} crystal-gel. The gel state is obtained from the state
-point $\phi\approx 0.33$ and $c_p=1.07$ mg/g, while the crystal-gel is obtained at the same $\phi$ but
-at lower polymer concentration, $c_p=0.38$ mg/g. On the left, the network is represented as a continuous field that has
+\caption{\textcolor{blue}{{\bf Percolated network structures.} {\bf a:} gel; {\bf b:} crystal-gel. The gel state is obtained from the state
+point $\phi\approx 0.33$ and $c_p=1.07$ mg/g at $t=400$ min after sample preparation, while the crystal-gel is obtained at the same $\phi$ but
+at lower polymer concentration, $c_p=0.38$ mg/g at $t=468$ min after sample preparation. On the left, the network is represented as a continuous field that has
 been obtained by applying a gaussian filter to the position of the particles (of width equal to the inter-particle distance) and
@@
 but also the difference in the smoothness of the network surface between amorphous and crystal gels. 
-On the right, slabs of 5 particle diameter thickness are plotted,
+On the right, slabs of 4.5 $\mu$m thickness and 41 $\mu$m width are plotted,
 with particles coloured with the same colour scheme as in Fig.~\ref{fig:transitions}: red (crystal), green (gas particles), white (liquid particles). 
-We note that very few liquid particles remain and the network is almost crystalline.}} 
+We note that very few liquid particles remain and the network is almost crystalline. 
+}} 
 \label{fig:network}
@@
  % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735
- \caption{\textcolor{red}{{\bf Crystal gel formation.}  
+ \caption{\textcolor{blue}{{\bf Crystal gel formation.}  
 {\bf a,} Reconstructions from confocal coordinates at $\phi\approx 0.33$ and $c_p=0.38$ mg/g.
-Depth of view is $5\sigma$.
+Depth of view is 9.0 $\mu$m, while the later dimension is 41 $\mu$m.
 Particles are drawn to scale and coloured according to their phase.

### 2016-03-18 crystal-gel/nature/response.tex
Matching theoretical and experimental phase boundaries. Time dependence of local volume fraction

@@
 \usepackage{epstopdf}
-\usepackage{color}
+\usepackage{xcolor}
 \usepackage{setspace}
@@
 
+\newenvironment{referee}%
+{\bigskip\singlespacing\bf}%
+{\par\bigskip}
+
 \begin{document}
 
-\doublespacing
+%\doublespacing
 
-\doublespacing
+%\doublespacing
 
@@
 Referee 2 recommended our work, and had only minor technical questions. The most important one was regarding the phase boundaries
-which were computed with free-volume theory. In the present manuscript we have strengthened our results with a Gaussian field analysis
-that shows full consistency between theory and experimental observations.
+which were computed with free-volume theory.
+In the present manuscript \textcolor{teal}{we took great care to match experimental and theoretical phase boundaries.
+In addition} we strengthened our results with a Gaussian field analysis
+that shows independently full consistency between theory and experimental observations.
 
@@
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 This manuscript describes experimental work on a colloid-polymer mixture that undergoes viscoelastic phase separation thereby forming a spinodal structure, which is interrupted by crystallization.
@@
 referee hesitate to recommend the manuscript for publication in Nature Materials due to a lack of novelty (in terms of material and fundamental insight) of the present work.
-}
-
-\bigskip
-\doublespacing
+\end{referee}
 
@@
 This is supported by the kind comments on this point made by Reviewer 2.
-The references predict the kinetic pathway from the free energy landscape, and the discussion
+The references predict the kinetic pathway from the free energy landscape \textcolor{teal}{in the case of triple coexistence}, and the discussion
 is based on macroscopic experiments under the influence of gravity. These experiments do not contain information on the kinetic pathway at a microscopic level.
-In the paper by Renth et al., the closest example to our crystalizzation mechanism is “Sample 5” belonging to the kinetic regime called “G”.
+In the paper by Renth et al., the closest example to our crystalizzation mechanism is ``Sample 5'' belonging to the kinetic regime called ``G''.
 In this case, the process is initiated by spinodal decomposition.
@@
 This statement means that crystallites are formed in either the gas or the liquid phase.
-Overall, the process is different from what we found in our paper, which is a pathway involving all the three phases in a non-trivial manner.
+Overall, the process is different from what we found in our paper, which is a pathway involving all the three phases in a non-trivial manner\textcolor{teal}{, with a liquid phase metastable with respect to crystallisation}.
 Unveiling the exact kinetic pathway of crystal-gel formation cannot
@@
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 The authors report on a very interesting and new result: systems with short-range attractions that undergo freezing from a gas phase can form a solid with short-rang crystalline order but long-range, gel-like amorphous (porous) structure.
@@
 materials community and I recommend publication after a few comments are addressed.
-}
-
-\bigskip
-\doublespacing
+\end{referee}
 
@@
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 The authors propose that this method might work in a variety of other materials and be useful as a generalizable, rational method to make materials that are crystalline yet porous.
@@
 Alternatively, if the authors have good arguments for why this could work for noble metals, then these could be a good addition to the article.
-}
-
-\bigskip
-\doublespacing
+\end{referee}
 
@@
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 Overall the methods are appropriate to the conclusions. The article is quite clear other than the points listed below.
-}
-
-\bigskip
-\doublespacing
+\end{referee}
 
@@
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 A few technical points:
@@
 One of those articles also reports experiments, which might be worth citing here (though I do not think those experiments are the same as these findings).
-}
+\end{referee}
 
-\bigskip
-\doublespacing
-We should have cited these pioneering works on the pathway whereby a system phase separate into macroscopically coexisting colloidal gas, liquid, and crystal regions and its link to the free energy landscape.
+We should have cited these pioneering works on the pathway whereby a system phase separate into macroscopically coexisting colloidal gas, liquid, and crystal regions \textcolor{teal}{(in our case the liquid is metastable to crystallisation)} and its link to the free energy landscape.
 We have carefully read these papers.
-In the paper by Renth et al., the closed example to ours is their “Sample 5” belonging to the kinetic regime called “G”.
+In the paper by Renth et al., the closed example to ours is their ``Sample 5'' belonging to the kinetic regime called ``G''.
 In this case, the process is initiated by spinodal decomposition.
@@
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 2) Page 5, second paragraph: the text states that the vol frac of 68\% corresponds to the vol frac of the stable crystal phase.
@@
 There may be good reasons for the difference between the measured 0.68 and the expected equilibrium value, but this should be discussed.
-}
+\end{referee}
 
-\bigskip
-\doublespacing
+\textcolor{teal}{The Referee here points a real inconsistency in our previous analysis.
+We tracked our mistake down to our estimate of the volume fraction stemming from the poor precision on the colloid diameter obtained either from diffusion constant measurement ($2.3\pm0.5\,\mu$m) or from the position of the first peak of $g(r)$ ($2.2\pm0.5\,\mu$m).
+In the present version, we instead matched the stable crystal phase volume fraction to the asymptotic value of the local volume fraction of 12-neighboured particles in the crystallising sample.
+This yields a much more precise value for the colloid diameter ($2.24\,\mu$m).
+Our method is detailed in Supplementary Information.}
 
-We were not clear enough on this point.
-In the experimental representation of the phase diagram (vertical axis is $c_p$), the tie lines are not horizontal.
-We have now drawn the slanted gas-crystal tie line in the figure as a guide.
-We hope that this would make the situation much clearer.
-\textcolor{blue}{In relation to this, we have also re-estimated the volume fraction of each phase more accurately
-by a method described in Supplementary Information.
-We think that this increases the reliability of our estimate of the volume fraction.
-The end points of the above-mentioned tie lines are determined in this way.
-(Mathieu, John: please add any useful information.) }
+\textcolor{teal}{To cross-check the validity of this method, we extended both our experimental measurements and theoretical predictions to be able to compare the phase boundaries in all samples.
+On the theory side, we computed the position of the liquid side of both binodal and spinodal lines.
+For this, we had use a hard sphere equation of state that describes well the approach to jamming.
+On the experimental side, we used a Gaussian approximation of the density field to measure the amount of gas and liquid and deduce the volume fraction in the liquid phase.
+The updated Fig 1a shows a good agreement between the liquid-gas tie lines measured experimentally and the predicted binodal line.}
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 3) How were crystalline particles defined?
@@
 4) How is the local volume fraction defined? Is it the particle volume divided by the volume of the Voronoi cell?
-}
-
-\bigskip
-\doublespacing
+\end{referee}
 
@@
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 5) On pages 4 and 5, the authors refer to "random gelation universality class" and the associated exponent D=2.53.
@@
 Please explain and justify this comparison.
-}
-
-\bigskip
-\doublespacing
+\end{referee}
 
@@
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 6) A small technical point: 'sublimation' is used to refer to the formation of a crystal from vapor at the bottom of page 7 and on page 8 (8 lines from the bottom).
 Sublimation is a transition from solid to vapor, not the other way.
-}
-
-\bigskip
-\doublespacing
+\end{referee}
 
@@
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 Tsurusawa et al.
@@
 I have the following specific comments:
-}
-
-\bigskip
-\doublespacing
+\end{referee}
 
@@
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 1) The evidence for this crystal gel mechanism appears to derive from measurements at a single state point.
@@
 state points are needed to rule out idiosyncratic (time-dependent) physical chemical explanations for the abrupt increase in gas phase colloids.
-}
-
-\bigskip
-\doublespacing
+\end{referee}
 
@@
 But, in the late stage, the dynamics slows down and thus we reduced the frame rate of 3D scan to cover the long-time period.
-We are afraid that the Reviewer’s impression of a sudden increase in free gas particles and a possible discontinuity between the two movies originatse from this difference in the frame rate.
+\textcolor{teal}{While the first movies covers
+the first minutes of the process, the second movie shows the evolution over more than 7 hours.} We are afraid that the Reviewer’s impression of a sudden increase in free gas particles and a possible discontinuity between the two movies originates from this difference in the frame rate.
 To address this issue we have now included time stamps on both Supplemental Movies.
@@
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 2) From the SI movie, I don't recognize the final reported structure to be a crystal gel.
@@
 The changes in crystal size at late times, when the liquid to gas to crystal mechanism is proposed, appear that they could be equally well described as a process in which larger crystallites grow at the expense of smaller ones, through the gas phase.
-
-}
-
-\bigskip
-\doublespacing
+\end{referee}
 
@@
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 3) The data quality of Figure 1c, Figure 1d appear insufficient to classify the cp = 0.38 mg/mg state point as different from the others.
@@
 The authors should add replications, errors, and significance testing if they wish to distinguish the dimensionality (Fig 1c) and maximum volume fraction (Fig 1d) of the 0.38 mg/ml state point from the others.
-}
+\end{referee}
 
-\bigskip
-\doublespacing
-
-We can partly agree that the results of Fig. 1c/d would not be enough to differentiate the  cp = 0.38 mg/g state from the others
+We can partly agree that the results of Fig. 1c and previous Fig. 1d would not be enough to differentiate the  cp = 0.38 mg/g state from the others
 state points, if considered alone. The spirit of these figures is to show all the information we can get before we utilize the bond orientational
@@
 that indeed the state point cp = 0.38 mg/g is affected by crystallization. In particular, the dimensionality obtained
-with the structure factors, will be examined in more detail in Fig. 3b, where single-particle statistics allows us to
-get rid of all the noise.
-Regarding the volume fractions of Fig. 1d, we have performed the error analysis suggested by the Referee, by looking at the
-standard deviations of the volumes of particles with different neighbours. The results are reported in the updated Fig. 1d,
-where we indeed confirm that our results are statistically sound.
+with the structure factors, will be examined in more detail in Fig. 3b, where single-particle statistics allows us to get rid of all the noise.
 
-\vspace{1em}
+\textcolor{teal}{That said, we went even further than the error analysis suggested by the Referee, by looking at the time dependence of the local volume fraction (only for particles having 12 neighbours).
+This is now shown as Fig.
+1d.
+Although the curves start after all samples underwent gas-liquid phase separation, only the the cp = 0.38 mg/g state shows an increase, up to an asymptotic value markedly higher than all other samples.
+Even without bond order analysis this behaviour points to a degree of compaction and phase ordering.
+We also included error bars as suggested by the Referee.}
 
-\singlespacing
-
-{\bf
+\begin{referee}
 4) In panel 4b, about ~ 25\% of the particles are identified as in the surface state; however, transition probabilities are not recorded for this state in Figure 4c.
@@
 These data should include crystal-gas and crystal-liquid transition probabilities so that the author's mechanism of liquid to gas to crystal can be evaluated against the alternative that the sublimination/desublimination transitions are from crystal to gas to crystal.
-}
-
-\bigskip
-\doublespacing
+\end{referee}
 
@@
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 5) Looking at Figure 4b and 4c together: although the gas crystal transition probability is large, the abundance of gas particles is never very significant.
@@
 role here and there is little evidence for the liquid to gas to crystal route that is the paper's primary message.
-}
-
-\bigskip
-\doublespacing
+\end{referee}
 
@@
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
-
+\begin{referee}
 Minor points:
@@
 A) I think the previous literature about Bernal spirals and other one-dimensional structures in gels (e.g. PRL 94 208301 2005) is relevant to the present study.
-}
-
-\bigskip
-\doublespacing
+\end{referee}
 
@@
 
-\vspace{1em}
+\begin{referee}
+B) The SI movies would be more effective if they were merged and a time stamp added; that would allow the reader to better compare them to Figure 4.
+\end{referee}
 
-\singlespacing
-
-{\bf
-B) The SI movies would be more effective if they were merged and a time stamp added; that would allow the reader to better compare them to Figure 4.
-}
-
-\bigskip
-\doublespacing
 We have added a time stamp, although we did not merge the two movies due to the file size problem.  
@@
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 C) What time point or condition is Figure 3b at? Please add this information to distinguish these data from Figure 1b.
-}
-
-\bigskip
-\doublespacing
+\end{referee}
 

### 2016-03-19 nature/response.tex
Following the update of the data for the 12 neighbours in time computed with the average specific volume instead of the average volume fraction, sigma is now 2.25um instead of 2.24um

@@
 \usepackage{epstopdf}
-\usepackage{color}
+\usepackage{xcolor}
 \usepackage{setspace}
+
+\usepackage[top=35mm,bottom=35mm,left=30mm,right=30mm]{geometry}
+
+\newenvironment{referee}%
+{\bigskip\singlespacing\bf}%
+{\par\bigskip}
 
@@
 
-\doublespacing
+%\doublespacing
 
-\doublespacing
+%\doublespacing
 
 \noindent
-Dear Editor,
+{\bf Dear Pep,}
 \vskip 0.3cm
-we would like to thank you for your comments to our manuscript ``A novel route to the spontaneous formation of porous crystals via viscoelastic phase separation'' (NM15113574),
+We would like to thank you for your comments to our manuscript ``A novel route to the spontaneous formation of porous crystals via viscoelastic phase separation'' (NM15113574),
 and for providing us with three Referee reports.
@@
 Referee 2 recommended our work, and had only minor technical questions. The most important one was regarding the phase boundaries
-which were computed with free-volume theory. In the present manuscript we have strengthened our results with a Gaussian field analysis
-that shows full consistency between theory and experimental observations.
+which were computed with free-volume theory.
+In the present manuscript \textcolor{teal}{we took great care to match experimental and theoretical phase boundaries.
+In addition} we strengthened our results with a Gaussian field analysis
+that shows independently full consistency between theory and experimental observations.
 
-Referee 3 main concern comes from her/his direct inspection of our supplemental movies, where she/he observed an anomalous abundance
+Referee 3's main concern comes from her/his direct inspection of our supplemental movies, where she/he observed an anomalous abundance
 of gas particles, that led her/him to suggest that Ostwald ripening was the main crystallization channel. But this observation
@@
 Alternatively, if the authors have good arguments for why this could work for noble metals, then these could be a good addition to the article.
-}
+\end{referee}
 
-\bigskip
-\doublespacing
+We believe that many monoatomic and single-component molecular systems can satisfy all the conditions (i)-(iv) in a certain range of the temperature and pressure. However, the reviewer 
+is right in the sense that there may be a practical difficulty in realizing condition (iii), for example, for noble metals. 
+Condition (iii) is necessary to create a porous crystalline structure with large porosity. However, even if condition (iii) is not met, bicontinuous phase-separation structures 
+can be formed near a symmetry composition. Thus, we can still form a porous crystal structure even in such a case. We have added the following sentences to mention this. 
+We thank the reviewer for her/his thoughtful comment on this important point.
+
+``{\it For monoatomic systems such as noble metals, condition (iii) may not be satisfied easily. 
+To access glassy slow dynamics in a liquid phase, we may need a deep quench at a high pressure. However, we note that even without strong dynamic asymmetry 
+due to glassiness, bicontinuous phase separation can take place between 35-85 volume \% of the liquid phase in ordinary gas/liquid phase separation; 
+and, thus, porous crystalline structures can be formed, although a thin network structure may be difficult to be formed.}''
+
+\begin{referee}
+Overall the methods are appropriate to the conclusions. The article is quite clear other than the points listed below.
+\end{referee}
 
@@
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
-Overall the methods are appropriate to the conclusions. The article is quite clear other than the points listed below.
-}
-
-\bigskip
-\doublespacing
-
-Below we will answer the comments one by one.
-
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 A few technical points:
@@
 This statement means that crystallites are formed in either the gas or the liquid phase.
-Thus, this is clearly different from what we found in our paper as a new pathway involving all the three phases in a non-trivial manner.
-Nonetheless, our statement "kinetic path via the gas phase (liquid $\rightarrow$ gas $\rightarrow$ crystal) has been completely overlooked so far" was obviously too strong.
-We have replaced “completely” by “largely” and have cited these pioneering works, which considered the role of the complex free energy landscape for the first time.
+As kindly pointed out by the reviewer, thus, this is clearly different from what we found in our paper as a new pathway involving all the three phases in a non-trivial manner.
+Nonetheless, our statement ``kinetic path via the gas phase (liquid $\rightarrow$ gas $\rightarrow$ crystal) has been completely overlooked so far'' was obviously too strong.
+We have replaced ``completely'' by ``largely'' and have cited these pioneering works, which considered the role of the complex free energy landscape for the first time.
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 2) Page 5, second paragraph: the text states that the vol frac of 68\% corresponds to the vol frac of the stable crystal phase.
@@
 There may be good reasons for the difference between the measured 0.68 and the expected equilibrium value, but this should be discussed.
-}
+\end{referee}
 
-\bigskip
-\doublespacing
+\textcolor{teal}{The Referee here points a real inconsistency in our previous analysis.
+We tracked our mistake down to our estimate of the volume fraction stemming from the poor precision on the colloid diameter obtained either from diffusion constant measurement ($2.3\pm0.5\,\mu$m) or from the position of the first peak of $g(r)$ ($2.2\pm0.5\,\mu$m).
+In the present version, we instead matched the stable crystal phase volume fraction to the asymptotic value of the local volume fraction of 12-neighboured particles in the crystallising sample.
+This yields a much more precise value for the colloid diameter ($2.25\,\mu$m).
+Our method is detailed in Supplementary Information.}
 
-We were not clear enough on this point.
-In the experimental representation of the phase diagram (vertical axis is $c_p$), the tie lines are not horizontal.
-We have now drawn the slanted gas-crystal tie line in the figure as a guide.
-We hope that this would make the situation much clearer.
-\textcolor{red}{Mathieu could you check this point?}
+\textcolor{teal}{To cross-check the validity of this method, we extended both our experimental measurements and theoretical predictions to be able to compare the phase boundaries in all samples.
+On the theory side, we computed the position of the liquid side of both binodal and spinodal lines.
+For this, we had use a hard sphere equation of state that describes well the approach to jamming.
+On the experimental side, we used a Gaussian approximation of the density field to measure the amount of gas and liquid and deduce the volume fraction in the liquid phase.
+The updated Fig 1a shows a good agreement between the liquid-gas tie lines measured experimentally and the predicted binodal line.}
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 3) How were crystalline particles defined?
@@
 Please explain and justify this comparison.
-}
+\end{referee}
 
-\bigskip
-\doublespacing
-
-By "random gelation universality class" we mean three-dimensional percolation universality class (which applies for example to randomly branched polymers).
+By ``random gelation universality class'' we mean three-dimensional percolation universality class (which applies for example to randomly branched polymers).
 It is known that three-dimensional random percolation clusters and diffusion-limited aggregation clusters have both about the same fractal dimension of about 2.5, for which we have cited the review article by Herrmann.
 We agree that DLCA and RLCA provide smaller fractal dimensions; however, DLCA is usually observed at much lower volume fraction and with stronger bonding energy (after bonding there is little diffusion, which leads to the formation of really fractal objects) than our case.
-For our starting volume fraction of about $\phi\approx 0.30$ and polymer concentrations, we have found that the fractal dimension
+For our starting volume fraction of about $\phi\approx 0.30$ and polymer concentrations, we have found that the fractal (or effective) dimension
 to be more consistent with random percolation. We hope that this clarifies the situation.
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 6) A small technical point: 'sublimation' is used to refer to the formation of a crystal from vapor at the bottom of page 7 and on page 8 (8 lines from the bottom).
 Sublimation is a transition from solid to vapor, not the other way.
-}
-
-\bigskip
-\doublespacing
+\end{referee}
 
 We thank the Reviewer for pointing out this mistake. We have corrected it as suggested.
+
+\vspace{1cm}
 
@@
 An example of crystal-gel porour network obtained through our new capillary experiments
-is found in Fig. SXXX. All details about the experiments are included the Supplementary Information. We would like to thank the Referee for stimulating
+is found in \textcolor{blue}{Fig. SXXX}. All details about the experiments are included the Supplementary Information. We would like to thank the Referee for stimulating
 this additional investigation.
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 2) From the SI movie, I don't recognize the final reported structure to be a crystal gel.
@@
 The Movies show only a thin slice of the system which corresponds to the particle that are in focus.
+As can be seen in Fig. 2b (left panel), the crystal gel has a bicontinuous porous structure. 
 To show the connectivity of the network clearly, we have also made Supplementary Figure SXXX, highlighting the crystal network.
-We also include movies of a thin slice with a finite thickness. \textcolor{red}{Did we do this?} 
+We also include movies of a thin slice with a finite thickness. \textcolor{blue}{Please prepare movie(s).} 
 We hope that this will clarify the situation and convince the reviewer of the fact that our sample indeed forms a porous network of crystals.
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 3) The data quality of Figure 1c, Figure 1d appear insufficient to classify the cp = 0.38 mg/mg state point as different from the others.
@@
 In particular we note that, while being an active channel, the Ostwald ripening pathway is limited by
-the small conversion probability of surface particles into the gas phase.
+the small conversion probability of surface particles into the gas phase (please see also our reply to the next comment).
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
+\begin{referee}
 5) Looking at Figure 4b and 4c together: although the gas crystal transition probability is large, the abundance of gas particles is never very significant.
@@
 
-\vspace{1em}
+\begin{referee}
+B) The SI movies would be more effective if they were merged and a time stamp added; that would allow the reader to better compare them to Figure 4.
+\end{referee}
 
-\singlespacing
+We have added a time stamp, although we did not merge the two movies due to the file size problem.  
+We have also made the difference in the frame rate more explicit in Supplementary Movie Captions.
 
-{\bf
-B) The SI movies would be more effective if they were merged and a time stamp added; that would allow the reader to better compare them to Figure 4.
-}
+\begin{referee}
+C) What time point or condition is Figure 3b at? Please add this information to distinguish these data from Figure 1b.
+\end{referee}
 
-\bigskip
-\doublespacing
 We have followed the kind suggestion of the reviewer.
 
-\vspace{1em}
-
-\singlespacing
-
-{\bf
-C) What time point or condition is Figure 3b at? Please add this information to distinguish these data from Figure 1b.
-}
-
-\bigskip
-\doublespacing
-
-We have followed the kind suggestion of the reviewer.
+\vspace{1cm}
 

### 2016-03-19 nature/crystal_gel2.tex
Following the update of the data for the 12 neighbours in time computed with the average specific volume instead of the average volume fraction, sigma is now 2.25um instead of 2.24um

@@
 At the same time liquid regions that have not crystallized start evaporating, and the gas phase contributes to the growth of nuclei (middle panel).
-Finally the different nuclei coalesce (right panel).}
-\textcolor{red}{Note that the topology of the crystal network found at late times, i.e.
-$t=5577,\tau_B$ (middle panel) and $t=12098,\tau_B$ (right panel) in Fig.~\ref{fig:transitions}a, differs significantly from that of the liquid network at early stages, i.e.
+Finally the different nuclei coalesce (right panel).
+Note that the topology of the crystal network found at late times, i.e.
+$t=5577\,\tau_B$ (middle panel) and $t=12098\,\tau_B$ (right panel) in Fig.~\ref{fig:transitions}a, differs significantly from that of the liquid network at early stages, i.e.
 $t=2055\,\tau_B$ (left panel) in Fig.~\ref{fig:transitions}a.
-This suggests that the crystal network doesn't form only by direct freezing of the liquid, but also by other mechanisms which involve the evaporation of the liquid (Bergeron process) and the sublimation of small crystals (Ostwald ripening).
+This suggests that the crystal network does not form only by direct freezing of the liquid, but also by other mechanisms which involve the evaporation of the liquid (Bergeron process) and the sublimation of small crystals (Ostwald ripening).
 In the following we will investigate these different crystallization mechanisms in depth.}
@@
 
-We obtain the best fit between theoretical phase diagram and experimental data for $\sigma=2.24\,\mu$m and $R_g=80$ nm (see Supplementary method).
+We obtain the best match between theoretical phase diagram and experimental data for $\sigma=2.25\,\mu$m and $R_g=80$ nm (see Supplementary method).
 Therefore the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.071$ and the overlap mass fraction of polymer $2.25$~mg/g.

### 2016-03-19 crystal-gel/nature/response.tex
Revert Fig1d to local volume fraction function of number of neighbours. Transfer the time evolution of the local volume fraction for 12-neighboured particles to supplementary.

@@
 In the present version, we instead matched the stable crystal phase volume fraction to the asymptotic value of the local volume fraction of 12-neighboured particles in the crystallising sample.
-This yields a much more precise value for the colloid diameter ($2.24\,\mu$m).
+This yields a much more precise value for the colloid diameter ($2.25\,\mu$m).
 Our method is detailed in Supplementary Information.}
@@
 
-\textcolor{teal}{That said, we went even further than the error analysis suggested by the Referee, by looking at the time dependence of the local volume fraction (only for particles having 12 neighbours).
-This is now shown as Fig.
-1d.
-Although the curves start after all samples underwent gas-liquid phase separation, only the the cp = 0.38 mg/g state shows an increase, up to an asymptotic value markedly higher than all other samples.
-Even without bond order analysis this behaviour points to a degree of compaction and phase ordering.
-We also included error bars as suggested by the Referee.}
+\textcolor{teal}{That said, we included error bars as suggested by the Referee.
+Although minute, the density difference between the cp = 0.38 mg/g state and the other samples is significant.
+Furthermore, we added a supplementary figure that shows the time dependence of the local volume fraction (only for particles having 12 neighbours).
+Although the curves start after all samples underwent gas-liquid phase separation, only the cp = 0.38 mg/g state shows an increase, up to an asymptotic value markedly higher than all other samples.
+Even without bond order analysis this behaviour points to a degree of compaction and phase ordering.}
 

### 2016-03-19 crystal-gel/nature/Supplementary.tex
Change Fig1d to time evolution of 12-neighboured particles local volume fraction. Use the asymptotic value for the crystallising sample as a definition for the crystal volume fraction.
Revert Fig1d to local volume fraction function of number of neighbours. Transfer the time evolution of the local volume fraction for 12-neighboured particles to supplementary.

@@
 
+\begin{figure}
+\includegraphics{local_phi12.pdf}
+\caption{Late stage of the evolution of local volume fraction around colloidal particles having 12 nearest neighbours.
+Samples are shown with the same color code as in Fig.1 of the main text.
+In particular red circles correspond to the crystalizing sample at $\phi=0.33$ and $cp=0.38\,$mg/g.}
+\label{fig:local_phi12}
+\end{figure}
+
 The volume fraction of a colloidal suspension is notoriously difficult to estimate experimentally~\cite{Poon2012}.
@@
 
-Our two constraints is (i) to have the final local volume fraction of the 12-neighboured particles in the crystallising sample to be equal to the theoretical equilibrium crystal volume fraction (ii) to have the theoretical spinodal line to correspond to the gel region boundary.
+Our two constraints is (i) to have the final local volume fraction of the 12-neighboured particles in the crystallising sample to be equal to the theoretical equilibrium crystal volume fraction and (ii) to have the theoretical spinodal line to correspond to the gel region boundary.
 Starting from our initial estimates of $\sigma$ and $R$ we iteratively converge to $R=80$~nm, $\sigma = 2.25\,\mu$m and a crystal at $0.723$.
@@
 Adding a divergence at Jamming into CS equation of state is not trivial.
-Attempts by Lefebre~\cite{LeFevre1972} and Kolafa~\cite{Kolafa2004} match CS at low density, have the right divergence close to jamming but fare less than CS at intermediate volume fractions (0.5-0.56), the exact range where our critical point lies.
+Attempts by Lefebre~\cite{LeFevre1972} and Kolafa~\cite{Kolafa2004} match CS at low density, have the right divergence close to jamming but far less than CS at intermediate volume fractions (0.5-0.56), the exact range where our critical point lies.
 Here we decided to use an equation by Liu~\cite{Liu2006a} that is completely equivalent to CS up to 0.56 and then show good agreement with simulation data near jamming.
@@
 \section*{Transition probabilities}
-\begin{figure}[!t]
+\begin{figure}[!b]
  \centering

### 2016-03-19 nature/crystal_gel2.tex
Revert Fig1d to local volume fraction function of number of neighbours. Transfer the time evolution of the local volume fraction for 12-neighboured particles to supplementary.

@@
 {\bf c,} Structure factors for the late stages of the gelation process. The gray and black lines represents respectively random gelation and compact fractal dimensions.
-\textcolor{blue}{{\bf d,} Late stage of the evolution of local volume fraction around colloidal particles having 12 nearest neighbours.}
+\textcolor{blue}{{\bf d,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle for the state points in the very late stage.}
  }

### 2016-03-21 crystal-gel/nature/crystal_gel2.tex
fig:capillary

@@
 To be able to follow individual trajectories, we perform a 3D scan every 10~s ($\approx 4\tau_B$) at early time and every 30 s later.
+
+The crystal-gel state was reproduced in a sealed cell with all the components of the suspension pre-mixed.
+Colloid volume fraction was virtually identical at $\approx 0.33$ and polymer concentration was $0.40\,$mg/g.
+\textcolor{blue}{See Supplementary Figure XXX.}
 

### 2016-03-21 nature/Supplementary.tex
Following the update of the data for the 12 neighbours in time computed with the average specific volume instead of the average volume fraction, sigma is now 2.25um instead of 2.24um
Insert S-fig:capillary Reference S-fig:local_phi12, S-fig:probabilities and S-fig:capillary in the text gaussian -> Gaussian Rewrite the description of S-fig:local_phi12 and fig:phase_separation

@@
 of liquid and gas after phase separation we convert the position of particles into a contiunous field smearing
-the position of each particle with a gaussian field of variance $5$ pixels, and then mapping the field
+the position of each particle with a Gaussian field of variance $5$ pixels, and then mapping the field
 to a cubic grid with $10$ pixels spacing. In this way liquid and gas regions can be identified by
@@
  % field.pdf: 681x483 pixel, 72dpi, 24.02x17.04 cm, bb=0 0 681 483
- \caption{{\bf Distribution function of the gaussian field at $t=490\,\tau_B$.} The gaussian field is applied to
+ \caption{{\bf Distribution function of the Gaussian field at $t=490\,\tau_B$.} The Gaussian field is applied to
  a configuration just after phase separation and before crystallization. The vertical dashed line represents the
@@
 
+\begin{figure}
+\includegraphics{local_phi12.pdf}%
+\caption{Late stage of the evolution of local volume fraction around colloidal particles having 12 nearest neighbours.
+Samples are shown with the same color code as in Fig.1 of the main text.
+In particular red circles correspond to the crystalizing sample at $\phi=0.33$ and $cp=0.38\,$mg/g.}%
+\label{fig:local_phi12}%
+\end{figure}
+
 The volume fraction of a colloidal suspension is notoriously difficult to estimate experimentally~\cite{Poon2012}.
@@
 Our two constraints is (i) to have the final local volume fraction of the 12-neighboured particles in the crystallising sample to be equal to the theoretical equilibrium crystal volume fraction and (ii) to have the theoretical spinodal line to correspond to the gel region boundary.
-Starting from our initial estimates of $\sigma$ and $R$ we iteratively converge to $R=80$~nm, $\sigma = 2.24\,\mu$m and a crystal at $0.723$.
+Starting from our initial estimates of $\sigma$ and $R$ we iteratively converge to $R=80$~nm, $\sigma = 2.25\,\mu$m and a crystal at $0.723$.
 
@@
 Here we decided to use an equation by Liu~\cite{Liu2006a} that is completely equivalent to CS up to 0.56 and then show good agreement with simulation data near jamming.
-
+%
 The Liu EoS allows us to draw a phase diagram that is in qualitative agreement with the measured liquid compositions.
@@
 \section*{Transition probabilities}
-\begin{figure}[!b]
+\begin{figure}
  \centering
@@
 
-\bibliographystyle{naturemag2}
-\bibliography{biblio}
+\begin{figure*}
+\includegraphics{capillary.pdf}
+\caption{\textbf{Capillary experiment reproducing the crystal-gel state} Colloid volume fraction is $\approx 0.33$ and polymer concentration is $0.40\,$mg/g.
+Top row: Confocal slices showing the initial network, crystallising network and final crystal-gel network.
+Bottom row: Reconstructions from confocal coordinates.
+Particles are drawn to scale and coloured according to their phase.
+Red: crystal; white: liquid; green: gas.
+To highlight the crystalline regions, liquid and gas particles are drawn half-sized.}
+\label{fig:capillary}
+\end{figure*}
 
@@
 
+\bibliographystyle{naturemag2}
+\bibliography{biblio}
+
 \end{document}

### 2016-03-21 nature/crystal_gel2.tex
Insert S-fig:capillary Reference S-fig:local_phi12, S-fig:probabilities and S-fig:capillary in the text gaussian -> Gaussian Rewrite the description of S-fig:local_phi12 and fig:phase_separation

@@
 \usepackage{color}
+\usepackage{xr}
+\externaldocument[S-]{Supplementary}
 \begin{document}
@@
 
-To gain additional insight, in Fig.~\ref{fig:phase_separation}d we plot the average local volume fraction as a function of the number of neighbours.
-\textcolor{blue}{The volume fraction is obtained by computing the Voronoi diagram of each configuration,
-which uniquely assigns a volume to each colloidal particle.}
-For all state points, we can see that the average volume fraction of a colloidal
-particle increases with its number of neighbours. Approaching close packing (12 neighbours), we clearly see two families of curves. The state point
-at $\phi\approx 0.33$ and $c_p=0.38$ mg/g reaches close packing at a volume fraction of around $\sim 68\%$, which is in reasonable agreement with the
-volume fraction of the stable crystal phase~\cite{ilett1995phase}. All other state points, instead, reach close packing at a volume fraction $\sim 64\%$, which is
-indeed close to the volume fraction of the attractive glass state~\cite{pham2002multiple}. 
+\textcolor{blue}{To gain additional insight, in Supplementary Figure~\ref{S-fig:local_phi12} we plot the late stage (after all samples already underwent gas-liquid phase separation) time evolution of the local volume fraction of closed packed particles (having 12 neighbours).
+The volume fraction is obtained by computing the Voronoi diagram of each configuration, which uniquely assigns a volume to each colloidal particle.
+Only the state point at $\phi\approx 0.33$ and $c_p = 0.38\,$mg/g state shows an increase, up to an asymptotic value that we identify with the composition of the stable crystal phase $\approx 0.72$.
+We plot the asymptotic average volume fraction as a function of the number of neighbours in Fig.~\ref{fig:phase_separation}d.
+Approaching close packing (12 neighbours), we clearly see two families of curves.
+The state point at $\phi\approx 0.33$ and $c_p=0.38\,$mg/g reaches close packing at the volume fraction $\approx 0.72$.
+By contrast all other state points, reach close packing at a markedly lower volume fraction, which is indeed close to the volume fraction of the attractive glass state~\cite{pham2002multiple}.
+This result clearly indicates that a single state point follows a different arrest mechanism, in which phase separation is arrested by crystallization and not by glassiness.}
+
+%To gain additional insight, in Fig.~\ref{fig:phase_separation}d we plot the average local volume fraction as a function of the number of neighbours.
+%\textcolor{blue}{The volume fraction is obtained by computing the Voronoi diagram of each configuration, which uniquely assigns a volume to each colloidal particle.}
+%For all state points, we can see that the average volume fraction of a colloidal
+%particle increases with its number of neighbours.
+Approaching close packing (12 neighbours), we clearly see two families of curves.
+\textcolor{blue}{The state point at $\phi\approx 0.33$ and $c_p=0.38\,$mg/g reaches close packing at the
+%volume fraction $\approx 0.72$ of the stable crystal phase.
+By contrast all other state points, reach close packing at a lower volume fraction, which is indeed close to the volume fraction of the attractive glass state~\cite{pham2002multiple}.
+%%~\cite{ilett1995phase}
+%% 
+%The time evolution  of the local volume fraction of particles having 12 neighbours is also informative, see Supplementary Figure~\ref{S-fig:local_phi12}.
+After all samples underwent gas-liquid phase separation, only the state point at $\phi\approx 0.33$ and $c_p = 0.38\,$mg/g state shows an increase, up to an asymptotic value markedly higher than all other samples.
 %(\textbf{the absolute values of these volume fraction are not totally convincing, probably because of a not optimal estimate of the size of the colloids.
 We should run the code of Mathieau with size detection}).
-This result clearly indicates that the state point at $\phi\approx 0.33$ and $c_p=0.38$ mg/g follows a different arrest mechanism, in which phase separation
-is arrested by crystallization and not by glassiness. Although our estimation of the absolute volume fraction may involves systematic errors of a few \%, the trend should be robust~\cite{Poon2012}.
+%This result clearly indicates that this state point} follows a different arrest mechanism, in which phase separation is arrested by crystallization and not by glassiness.
+%Although our estimation of the absolute volume fraction may involves systematic errors of a few \%, the trend should be robust~\cite{Poon2012}.
 
@@
 \textcolor{blue}{
-Next we show in Fig.~\ref{fig:transitions}b the fraction of particles in each different phase for the crystallization trajectory at
+In Fig.~\ref{fig:transitions}b, we show  the fraction of particles in each different phase for the crystallization trajectory at
 $\phi\approx 0.33$ and $c_p=0.38$ mg/g, after the liquid-gas phase separation has occurred.
@@
 In Supplementary Information we confirm these results with a detailed analysis of transition probabilities, including a state for surface particles that are in contact 
-to crystalline particles.
+to crystalline particles., see Supplementary Figure~\ref{S-fig:probabilities}.
 While the direct freezing of the fluid represents the major contribution to the nuclei growth, the kinetic path via the gas phase (gas$\rightarrow$crystal), also plays a crucial role,
@@
 
+Matching phase boundaries often leads to better determination of volume fractions than colloid diameter measurements~\cite{Poon2012}.
 We obtain the best match between theoretical phase diagram and experimental data for $\sigma=2.25\,\mu$m and $R_g=80$ nm (see Supplementary method).
@@
 To be able to follow individual trajectories, we perform a 3D scan every 10~s ($\approx 4\tau_B$) at early time and every 30 s later.
+
+The crystal-gel state is reproduced in a sealed cell with all the components of the suspension pre-mixed.
+Colloid volume fraction is virtually identical at $\approx 0.33$ and polymer concentration is $0.40\,$mg/g.
+See Supplementary Figure~\ref{S-fig:capillary}.
 
@@
 at lower polymer concentration, $c_p=0.38$ mg/g at $t=468$ min after sample preparation. On the left, the network is represented as a continuous field that has
-been obtained by applying a gaussian filter to the position of the particles (of width equal to the inter-particle distance) and
+been obtained by applying a Gaussian filter to the position of the particles (of width equal to the inter-particle distance) and
 plotting the surface which bounds the highest 20 \% values of the field. We can clearly see not only the percolated nature of both networks, 

### 2016-03-21 crystal-gel/nature/Supplementary.tex
Small additions.

@@
 \begin{figure}
-\includegraphics{local_phi12.pdf}
+\includegraphics{local_phi12.pdf}%
 \caption{Late stage of the evolution of local volume fraction around colloidal particles having 12 nearest neighbours.
 Samples are shown with the same color code as in Fig.1 of the main text.
-In particular red circles correspond to the crystalizing sample at $\phi=0.33$ and $cp=0.38\,$mg/g.}
-\label{fig:local_phi12}
+In particular red circles correspond to the crystalizing sample at $\phi=0.33$ and $cp=0.38\,$mg/g.}%
+\label{fig:local_phi12}%
 \end{figure}
@@
 
-Our two constraints is (i) to have the final local volume fraction of the 12-neighboured particles in the crystallising sample to be equal to the theoretical equilibrium crystal volume fraction and (ii) to have the theoretical spinodal line to correspond to the gel region boundary.
+Our two constraints are (i) to have the final local volume fraction of the 12-neighboured particles in the crystallising sample to be equal to the theoretical equilibrium crystal volume fraction and (ii) to have the theoretical spinodal line to correspond to the gel region boundary.
 Starting from our initial estimates of $\sigma$ and $R$ we iteratively converge to $R=80$~nm, $\sigma = 2.25\,\mu$m and a crystal at $0.723$.
@@
 
-\bibliographystyle{naturemag2}
-\bibliography{biblio}
+\section*{Capillary experiment}
 
-\clearpage
+\begin{figure*}[!t]
+\includegraphics{capillary.pdf}
+\caption{\textbf{Capillary experiment reproducing the crystal-gel state} Colloid volume fraction is $\approx 0.33$ and polymer concentration is $0.40\,$mg/g.
+Top row: Confocal slices showing the initial network, crystallising network and final crystal-gel network.
+Bottom row: Reconstructions from confocal coordinates.
+Particles are drawn to scale and coloured according to their phase.
+Red: crystal; white: liquid; green: gas.
+To highlight the crystalline regions, liquid and gas particles are drawn to a tenth of the crystalline particle size.}
+\label{fig:capillary}
+\end{figure*}
+
+The use of a reservoir of a salt solution through a semi-permeable membrane, allowed us to follow the entire microscopic dynamics in absence of spurious flows, from the first stages of phase separation to the formation of the crystal-gel state.
+In order to confirm the reproducibility of the crystal-gel state, we conducted experiments in a more traditional capillary setting, where all components of the suspension are pre-mixed in a sealed cell.
+A typical example is
+shown in Fig.~\ref{fig:capillary}, where a system with colloid volume fraction of $\approx 0.33$ and polymer concentration is $0.40\,$mg/g is considered.
+The figure shows the time evolution of the sample at three different times, just after phase separation ($t=340\,\tau_\beta$), and during
+crystal-gel formation ($t=2600\,\tau_\beta$ and $t=6650\,\tau_\beta$). The top-row shows images from quasi two-dimensional confocal slices, while the bottom-row shows
+the three-dimensional reconstruction after single-particle tracking, where crystalline particles are highlighted in both size and color (red).
+
 \section*{Supplementary Movie 1}

### 2016-03-21 crystal-gel/nature/crystal_gel2.tex
Small additions.

@@
 
-To gain additional insight, in Fig.~\ref{fig:phase_separation}d we plot the average local volume fraction as a function of the number of neighbours.
-\textcolor{blue}{The volume fraction is obtained by computing the Voronoi diagram of each configuration,
-which uniquely assigns a volume to each colloidal particle.}
-For all state points, we can see that the average volume fraction of a colloidal
-particle increases with its number of neighbours. Approaching close packing (12 neighbours), we clearly see two families of curves. The state point
-at $\phi\approx 0.33$ and $c_p=0.38$ mg/g reaches close packing at a volume fraction of around $\sim 68\%$, which is in reasonable agreement with the
-volume fraction of the stable crystal phase~\cite{ilett1995phase}. All other state points, instead, reach close packing at a volume fraction $\sim 64\%$, which is
-indeed close to the volume fraction of the attractive glass state~\cite{pham2002multiple}. 
+\textcolor{blue}{To gain additional insight, in Supplementary Figure~\ref{S-fig:local_phi12} we plot the late stage (after all samples already underwent gas-liquid phase separation) time evolution of the local volume fraction of closed packed particles (having 12 neighbours).
+The volume fraction is obtained by computing the Voronoi diagram of each configuration, which uniquely assigns a volume to each colloidal particle.
+Only the state point at $\phi\approx 0.33$ and $c_p = 0.38\,$mg/g state shows an increase, up to an asymptotic value that we identify with the composition of the stable crystal phase $\approx 0.72$.
+We plot the asymptotic average volume fraction as a function of the number of neighbours in Fig.~\ref{fig:phase_separation}d.
+Approaching close packing (12 neighbours), we clearly see two families of curves.
+The state point at $\phi\approx 0.33$ and $c_p=0.38\,$mg/g reaches close packing at the volume fraction $\approx 0.72$.
+By contrast all other state points, reach close packing at a markedly
+lower volume fraction, which is indeed close to the volume fraction of the attractive glass state~\cite{pham2002multiple}.
+This result clearly indicates that a single state point follows a different arrest mechanism, in which phase separation is arrested by crystallization and not by glassiness.}
+
+%To gain additional insight, in Fig.~\ref{fig:phase_separation}d we plot the average local volume fraction as a function of the number of neighbours.
+%\textcolor{blue}{The volume fraction is obtained by computing the Voronoi diagram of each configuration, which uniquely assigns a volume to each colloidal particle.}
+%For all state points, we can see that the average volume fraction of a colloidal
+%particle increases with its number of neighbours.
+Approaching close packing (12 neighbours), we clearly see two families of curves.
+\textcolor{blue}{The state point at $\phi\approx 0.33$ and $c_p=0.38\,$mg/g reaches close packing at the
+%volume fraction $\approx 0.72$ of the stable crystal phase.
+By contrast all other state points, reach close packing at a lower volume fraction, which is indeed close to the volume fraction of the attractive glass state~\cite{pham2002multiple}.
+%%~\cite{ilett1995phase}
+%% 
+%The time evolution  of the local volume fraction of particles having 12 neighbours is also informative, see Supplementary Figure~\ref{S-fig:local_phi12}.
+After all samples underwent gas-liquid phase separation, only the state point at $\phi\approx 0.33$ and $c_p = 0.38\,$mg/g state shows an increase, up to an asymptotic value markedly higher than all other samples.
 %(\textbf{the absolute values of these volume fraction are not totally convincing, probably because of a not optimal estimate of the size of the colloids.
 We should run the code of Mathieau with size detection}).
-This result clearly indicates that the state point at $\phi\approx 0.33$ and $c_p=0.38$ mg/g follows a different arrest mechanism, in which phase separation
-is arrested by crystallization and not by glassiness. Although our estimation of the absolute volume fraction may involves systematic errors of a few \%, the trend should be robust~\cite{Poon2012}.
+%This result clearly indicates that this state point} follows a different arrest mechanism, in which phase separation is arrested by crystallization and not by glassiness.
+%Although our estimation of the absolute volume fraction may involves systematic errors of a few \%, the trend should be robust~\cite{Poon2012}.
 
@@
 
-The crystal-gel state was reproduced in a sealed cell with all the components of the suspension pre-mixed.
-Colloid volume fraction was virtually identical at $\approx 0.33$ and polymer concentration was $0.40\,$mg/g.
-\textcolor{blue}{See Supplementary Figure XXX.}
+\textcolor{blue}{
+The crystal-gel state is reproduced in a sealed cell with all the components of the suspension pre-mixed.
+Colloid volume fraction is virtually identical at $\approx 0.33$ and polymer concentration is $0.40\,$mg/g.
+See Supplementary Figure~\ref{S-fig:capillary}.}
 

### 2016-03-21 nature/response.tex
Small additions.

@@
 Condition (iii) is necessary to create a porous crystalline structure with large porosity. However, even if condition (iii) is not met, bicontinuous phase-separation structures 
-can be formed near a symmetry composition. Thus, we can still form a porous crystal structure even in such a case. We have added the following sentences to mention this. 
+can be formed near a \textcolor{blue}{symmetric} composition.
+\textcolor{blue}{Thus, we believe a porous crystal structure could be formed even in such a case.} We have added the following sentences to mention this.
 We thank the reviewer for her/his thoughtful comment on this important point.
@@
 The subsequent theoretical paper by Evans et al.
-also stated that spinodal decomposition takes place in the region G before crystallization takes place.
-Then they cited the
+also stated that spinodal decomposition takes
+place in the region G before crystallization takes place. Then they cited the 
 simulation work by Soga et al.
@@
 An example of crystal-gel porour network obtained through our new capillary experiments
-is found in \textcolor{blue}{Fig. SXXX}. All details about the experiments are included the Supplementary Information. We would like to thank the Referee for stimulating
+is found in \textcolor{blue}{Fig. S5}. All details about the experiments are included the Supplementary Information. We would like to thank the Referee for stimulating
 this additional investigation.
@@
 As can be seen in Fig. 2b (left panel), the crystal gel has a bicontinuous porous structure. 
-To show the connectivity of the network clearly, we have also made Supplementary Figure SXXX, highlighting the crystal network.
+To show the connectivity of the network clearly, we have also made Supplementary Figure S5, highlighting the crystal network.
 We also include movies of a thin slice with a finite thickness. \textcolor{blue}{Please prepare movie(s).} 

### 2016-03-22 crystal-gel/nature/response.tex
Dear All,
Thank you again for your efforts. 
I went through all the files and made many small corrections. 
I did not put any color since you can see the differences from the previous files by using a function of SVN. I just put blue color in the response file (One is a point on which I would like to ask your opinion and the other is about a new movie). Actually, it is difficult to make the movie, it looks OK without it. Please check all modified parts in the three files. Thank you! Hajime

@@
 \vskip 0.3cm
-We would like to thank you for your comments to our manuscript ``A novel route to the spontaneous formation of porous crystals via viscoelastic phase separation'' (NM15113574),
+We would like to thank you for your encouraging suggestions to our manuscript 
+``A novel route to the spontaneous formation of porous crystals via viscoelastic phase separation'' (NM15113574),
 and for providing us with three Referee reports. 
@@
 which were computed with free-volume theory.
-In the present manuscript \textcolor{teal}{we took great care to match experimental and theoretical phase boundaries.
-In addition} we strengthened our results with a Gaussian field analysis
-that shows independently full consistency between theory and experimental observations.
+In the present manuscript we took great care to match experimental and theoretical phase boundaries.
+In addition, we strengthened our results with a Gaussian field analysis that shows independently full consistency between theory and experimental observations.
 
@@
 of gas particles, that led her/him to suggest that Ostwald ripening was the main crystallization channel. But this observation
-was misguided by the fact that the two supplemental movies are shot at very different framerates. This was explained in the movies captions,
-but there was no time stamp on the movies, and we believe that this was the source of confusion, for which we apologize. While the first movies covers
+was misguided by the fact that the two supplemental movies are shot at very different frame rates. This was explained in the movies captions,
+but there was no time stamp on the movies, and we believe that this was the source of confusion, for which we apologize. While the first movie covers
 the first minutes of the process, the second movie shows the evolution over more than 7 hours, giving the impression
@@
 the crystalline nuclei appear to be separate from each other, which led the Referee to suggest Ostwald ripening. But looking at the process in three dimensions,
-it is without any doubt that crystalline nuclei are connected in a porous network. We show this very clear three-dimensional representation in a new Supplementary Figure.
+it is without any doubt that crystalline nuclei are connected in a porous network. This is evident from a coarse-grained morphology shown in Fig. 2b. 
+We also provide a very clear three-dimensional particle-level representation in a new Supplementary Figure 5.  
 Together with a novel analysis which determines the crystallization pathway directly by looking 
@@
 This is supported by the kind comments on this point made by Reviewer 2.
-The references predict the kinetic pathway from the free energy landscape \textcolor{teal}{in the case of triple coexistence}, and the discussion
+The references predict the kinetic pathway from the free energy landscape in the case of triple coexistence, and the discussion
 is based on macroscopic experiments under the influence of gravity. These experiments do not contain information on the kinetic pathway at a microscopic level.
-In the paper by Renth et al., the closest example to our crystalizzation mechanism is ``Sample 5'' belonging to the kinetic regime called ``G''.
+In the paper by Renth et al., the closest example to our crystallization mechanism is ``Sample 5'' belonging to the kinetic regime called ``G''.
 In this case, the process is initiated by spinodal decomposition.
@@
 This statement means that crystallites are formed in either the gas or the liquid phase.
-Overall, the process is different from what we found in our paper, which is a pathway involving all the three phases in a non-trivial manner\textcolor{teal}{, with a liquid phase metastable with respect to crystallisation}.
+Overall, the process is different from what we found in our paper, which is a pathway involving all the three phases in a non-trivial manner, with a liquid phase metastable with respect to crystallisation.
 Unveiling the exact kinetic pathway of crystal-gel formation cannot
@@
 Therefore, I think the results are quite new.
-I think the results are relevent to the
+I think the results are relevant to the
 materials community and I recommend publication after a few comments are addressed.
@@
 For example, the authors propose that this could work for metals - this is mentioned in the abstract and in the conclusion, on page 9.
-Success in this area requires meeting the 4 conditions listed on pages 8-9 but this may not be possible with nobel metals (especially with regard to iii).
+Success in this area requires meeting the 4 conditions listed on pages 8-9 but this may not be possible with noble metals (especially with regard to iii).
 Choosing a different example might make this point more credible and the article more valuable.
@@
 
-We believe that many monoatomic and single-component molecular systems can satisfy all the conditions (i)-(iv) in a certain range of the temperature and pressure. However, the reviewer 
+We believe that many monoatomic and single-component molecular systems can in principle satisfy all the conditions (i)-(iv) in a certain range of the temperature and pressure. However, the reviewer 
 is right in the sense that there may be a practical difficulty in realizing condition (iii), for example, for noble metals. 
 Condition (iii) is necessary to create a porous crystalline structure with large porosity. However, even if condition (iii) is not met, bicontinuous phase-separation structures 
-can be formed near a symmetry composition. Thus, we can still form a porous crystal structure even in such a case. We have added the following sentences to mention this. 
+can be formed near a symmetric composition. Thus, we believe a porous crystal structure could be formed even in such a case. We have added the following sentences to mention this 
+with a new reference [42]. 
 We thank the reviewer for her/his thoughtful comment on this important point.
@@
 To access glassy slow dynamics in a liquid phase, we may need a deep quench at a high pressure. However, we note that even without strong dynamic asymmetry 
-due to glassiness, bicontinuous phase separation can take place between 35-85 volume \% of the liquid phase in ordinary gas/liquid phase separation; 
+due to glassiness, bicontinuous phase separation can take place between 35-85 volume \% of the liquid phase in ordinary gas/liquid phase separation [42]; 
 and, thus, porous crystalline structures can be formed, although a thin network structure may be difficult to be formed.}''
@@
 
-We should have cited these pioneering works on the pathway whereby a system phase separate into macroscopically coexisting colloidal gas, liquid, and crystal regions \textcolor{teal}{(in our case the liquid is metastable to crystallisation)} and its link to the free energy landscape.
+We should have cited these pioneering works on the pathway whereby a system phase separates into macroscopically coexisting colloidal gas, liquid, and crystal regions (in our case the liquid is metastable to crystallisation) and its link to the free energy landscape.
 We have carefully read these papers.
@@
 Nonetheless, our statement ``kinetic path via the gas phase (liquid $\rightarrow$ gas $\rightarrow$ crystal) has been completely overlooked so far'' was obviously too strong.
-We have replaced ``completely'' by ``largely'' and have cited these pioneering works, which considered the role of the complex free energy landscape for the first time.
+Thus, we have replaced ``completely'' by ``largely'' and have cited these pioneering works, which considered the role of the complex free energy landscape for the first time.
 
@@
 
-\textcolor{teal}{The Referee here points a real inconsistency in our previous analysis.
+The Referee here points a real inconsistency in our previous analysis.
 We tracked our mistake down to our estimate of the volume fraction stemming from the poor precision on the colloid diameter obtained either from diffusion constant measurement ($2.3\pm0.5\,\mu$m) or from the position of the first peak of $g(r)$ ($2.2\pm0.5\,\mu$m).
@@
 This yields a much more precise value for the colloid diameter ($2.25\,\mu$m).
-Our method is detailed in Supplementary Information.}
+Our method is detailed in Supplementary Information.
 
-\textcolor{teal}{To cross-check the validity of this method, we extended both our experimental measurements and theoretical predictions to be able to compare the phase boundaries in all samples.
+To cross-check the validity of this method, we extended both our experimental measurements and theoretical predictions to be able to compare the phase boundaries in all samples.
 On the theory side, we computed the position of the liquid side of both binodal and spinodal lines.
@@
 On the experimental side, we used a Gaussian approximation of the density field to measure the amount of gas and liquid and deduce the volume fraction in the liquid phase.
-The updated Fig 1a shows a good agreement between the liquid-gas tie lines measured experimentally and the predicted binodal line.}
+The updated Fig 1a shows a good agreement between the liquid-gas tie lines measured experimentally and the predicted binodal line.
+\textcolor{blue}{Here we note that the volume fraction of the dense phase is an average over a certain range of the number of neighbours
+shown in Fig. 1d.}
 
@@
 We agree that DLCA and RLCA provide smaller fractal dimensions; however, DLCA is usually observed at much lower volume fraction and with stronger bonding energy (after bonding there is little diffusion, which leads to the formation of really fractal objects) than our case.
-For our starting volume fraction of about $\phi\approx 0.30$ and polymer concentrations, we have found that the fractal (or effective) dimension
-to be more consistent with random percolation. We hope that this clarifies the situation.
+For our starting volume fraction of about $\phi\approx 0.30$ and polymer concentrations, we have found that the fractal (or effective) dimension to be more consistent with random percolation.
+We hope that this clarifies the situation.
 
@@
 
-We would like to thank the Referee for her/his careful reading of our
-manuscript, and for her/his comments.
+We would like to thank the Referee for her/his careful reading of our manuscript, and for her/his comments.
 
@@
 
-We think that we have not been clear enough about the explanations of the Supplementary movies.
+We think that we have not been clear enough about the explanations of the Supplementary Movies.
 Movie 1 and Movie 2 were made with different frame rates and there was an interval of 90 s between them.
@@
 But, in the late stage, the dynamics slows down and thus we reduced the frame rate of 3D scan to cover the long-time period.
-\textcolor{teal}{While the first movies covers
-the first minutes of the process, the second movie shows the evolution over more than 7 hours.} We are afraid that the Reviewer’s impression of a sudden increase in free gas particles and a possible discontinuity between the two movies originates from this difference in the frame rate.
-To address this issue we have now included time stamps on both Supplemental Movies.
+While the first movie covers the first minutes of the process, the second movie shows the evolution over more than 7 hours.
+We are afraid that the Reviewer’s impression of a sudden increase in free gas particles and a possible discontinuity between the two movies originates from this difference in the frame rate.
+To address this issue we have now included time stamps on both Supplemental Movies and explicitly stated the difference in the frame rate in Supplementary Movie 2 Caption.
 
 To address the concern about the reproducibility of our kinetic pathway, we decided to make additional independent experiments
-by using a sample with similar composition, which is inserted into an ordinary glass capillary tube.
+by using a sample with a similar composition, which is inserted into an ordinary glass capillary tube.
 We thus confirm the reproducibility of this type of crystallization pathway.
-An example of crystal-gel porour network obtained through our new capillary experiments
-is found in \textcolor{blue}{Fig. SXXX}. All details about the experiments are included the Supplementary Information. We would like to thank the Referee for stimulating
-this additional investigation.
+An example of crystal-gel porous  network obtained through our new capillary experiments is found in new Supplementary Fig.
+S5.
+All details about the experiments are also included the Supplementary Information.
+We would like to thank the Referee for stimulating this additional investigation.
 
@@
 As can be seen in Fig. 2b (left panel), the crystal gel has a bicontinuous porous structure. 
-To show the connectivity of the network clearly, we have also made Supplementary Figure SXXX, highlighting the crystal network.
-We also include movies of a thin slice with a finite thickness. \textcolor{blue}{Please prepare movie(s).} 
+To show the connectivity of the network clearly with a particle-level resolution, we have also made Supplementary Figure S5, highlighting the crystal network. 
+\textcolor{blue}{We also include a movie of a thin slice with a finite thickness. Please prepare a movie.} 
 We hope that this will clarify the situation and convince the reviewer of the fact that our sample indeed forms a porous network of crystals.
@@
 
-\textcolor{teal}{That said, we included error bars as suggested by the Referee.
+That said, we have included error bars as suggested by the Referee.
 Although minute, the density difference between the cp = 0.38 mg/g state and the other samples is significant.
-Furthermore, we added a supplementary figure that shows the time dependence of the local volume fraction (only for particles having 12 neighbours).
+Furthermore, we added a Supplementary Figure S3 that shows the time dependence of the local volume fraction (only for particles having 12 neighbours).
 Although the curves start after all samples underwent gas-liquid phase separation, only the cp = 0.38 mg/g state shows an increase, up to an asymptotic value markedly higher than all other samples.
-Even without bond order analysis this behaviour points to a degree of compaction and phase ordering.}
+Even without bond order analysis this behaviour points to a degree of compaction and phase ordering.
 
@@
 To include the full analysis, as requested by the Referee,
-we have added a section to Supplementary Information, with a new figure (Fig. S3) containing four panels that detail the transition
+we have added a section to Supplementary Information, with a new figure (Fig. S4) containing four panels that detail the transition 
 probabilities between all states. These transition probabilities are measures by reconstructing the particles' trajectories, and
@@
 As mentioned before, we are afraid that the above interpretation basically originates from some misunderstandings, which is caused by our poor explanations on the different frame rates of Supplementary Movies and the lack of more 3D representations of our results.
-We hope that the above explanations and new Movies would convince the reviewer of the formation of a percolated network structure composed of crystals (not isolated crystals diffusing around) and the novel kinetic pathway leading to porous crystalline structures.
+We hope that the above explanations and new Supplementary Fig.
+S5 and Movie would convince the reviewer of the formation of a percolated network structure composed of crystals (not isolated crystals diffusing around) and the novel kinetic pathway leading to porous crystalline structures (see also Fig.
+2b).
 Moreover, to address directly the Referee's concerns we have added a new analysis, whose results are reported in the new Fig. 4c.
@@
 We have added a time stamp, although we did not merge the two movies due to the file size problem.  
-We have also made the difference in the frame rate more explicit in Supplementary Movie Captions.
+We have also made the difference in the frame rate more explicit in Supplementary Movie 2 Caption.
 

### 2016-03-22 crystal-gel/nature/Supplementary.tex
Dear All,
Thank you again for your efforts. 
I went through all the files and made many small corrections. 
I did not put any color since you can see the differences from the previous files by using a function of SVN. I just put blue color in the response file (One is a point on which I would like to ask your opinion and the other is about a new movie). Actually, it is difficult to make the movie, it looks OK without it. Please check all modified parts in the three files. Thank you! Hajime

@@
 \section*{Definition of states}
-The configurations obtained by tracking the position of the colloids in the samples
-are analysed in order to determine whether a colloidal particles is in the
+The configurations obtained by tracking the position of colloids in the samples
+are analysed in order to determine whether a colloidal particle is in the
 crystal, surface, liquid or gas state. In the following we describe the classification
-criteria
+criteria.
 
@@
 if the scalar product $(\mathbf{q}_{6}(i)/|\mathbf{q}_{6}(i)|)\cdot(\mathbf{q}_{6}(j)/|\mathbf{q}_{6}(j)|)>0.7$.
-A molecule is then identified as crystalline if it has at least $7$ connected neighbours.
+A particle is then identified as crystalline if it has at least $7$ connected neighbours.
 
@@
 Surface, gas and liquid particles are identified from the subset of non-crystalline particles depending
-on their neighbouring particles. A particle with at least two neighbouring crystal particle is classified as
-\emph{surface}. A particle that is neither solid or surface, and has at least four neighbouring particles
-is classified as \emph{liquid}. A particle which is neither solid, surface or liquid is classified as \emph{gas}.
-Fig.~\ref{fig:fractions} plots the fraction of each state for the state point at $\phi\approx 0.30$ and $c_p=0.38$ mg/g
+on their neighbouring particles. A particle with at least two neighbouring crystal particles is classified as
+\emph{surface}. A particle that is neither solid nor surface, and has at least four neighbouring particles
+is classified as \emph{liquid}. A particle which is neither solid, surface nor liquid is classified as \emph{gas}.
+Figure~\ref{fig:fractions} plots the fraction of each state for the state point at $\phi\approx 0.30$ and $c_p=0.38$ mg/g
 
@@
 of the gas and liquid phases just after phase separation (and before crystallization). To analyse the fraction
-of liquid and gas after phase separation we convert the position of particles into a contiunous field smearing
+of liquid and gas after phase separation we convert the position of particles into a continuous field smearing
 the position of each particle with a Gaussian field of variance $5$ pixels, and then mapping the field
@@
  % field.pdf: 681x483 pixel, 72dpi, 24.02x17.04 cm, bb=0 0 681 483
- \caption{{\bf Distribution function of the Gaussian field at $t=490\,\tau_B$.} The Gaussian field is applied to
+ \caption{{\bf Distribution function of the Gaussian field at $t=490\,\tau_{\rm B}$.} The Gaussian field is applied to
  a configuration just after phase separation and before crystallization. The vertical dashed line represents the
@@
 
-Fig.~\ref{fig:field} shows the probability distribution of the field, where two peaks clearly indicate the presence
-of gas/liquid coexistence. By choosing a threshold over the field of $f=0.06$ (dashed line) we compute the fraction
-of particles in the solid and gas phases at $t=490\,\tau_B$, just after phase separation. The results are consistent
+Figure~\ref{fig:field} shows the probability distribution of the field, where two peaks clearly indicate the presence
+of gas/liquid coexistence. By choosing a threshold over the field of $f=0.06$ (vertical dashed line) we compute the fraction 
+of particles in the solid and gas phases at $t=490\,\tau_{\rm B}$, just after phase separation. The results are consistent
 with Fig.~\ref{fig:fractions}, where the number fraction of particles in the liquid state accounts for more than
-$90\%$ of the particles after phase separation.
+$90\%$ of the particles just after phase separation.
 
@@
 The colloids are considered as hard spheres and the polymers as an ideal gas.
-The center of the polymer cannot penetrate within $R$ of the surface of a colloid.
+The centre of mass of a polymer cannot penetrate within $R$ of the surface of a colloid.
 Therefore the ideal gas is constrained to a free volume smaller than the total volume and that depends on the configuration (volume fraction) of the colloids and the polymer/colloid size ratio.
@@
 \includegraphics{local_phi12.pdf}%
-\caption{Late stage of the evolution of local volume fraction around colloidal particles having 12 nearest neighbours.
-Samples are shown with the same color code as in Fig.1 of the main text.
-In particular red circles correspond to the crystalizing sample at $\phi=0.33$ and $cp=0.38\,$mg/g.}%
+\caption{{\bf Late stage of the evolution of the local volume fraction around colloidal particles having 12 nearest neighbours.} Samples are shown with the same colour code as in Fig.1 of the main text.
+In particular red circles correspond to the crystallizing sample at $\phi=0.33$ and $cp=0.38\,$mg/g.}%
 \label{fig:local_phi12}%
@@
 Our two constraints are (i) to have the final local volume fraction of the 12-neighboured particles in the crystallising sample to be equal to the theoretical equilibrium crystal volume fraction and (ii) to have the theoretical spinodal line to correspond to the gel region boundary.
-Starting from our initial estimates of $\sigma$ and $R$ we iteratively converge to $R=80$~nm, $\sigma = 2.25\,\mu$m and a crystal at $0.723$.
+Starting from our initial estimates of $\sigma$ and $R$ we iteratively converge to $R=80$~nm, $\sigma = 2.25\,\mu$m and a crystal (without defects) at $0.723$.
 
@@
  transition states are surface (blue), liquid (black), gas (green) and crystal (red).
- See the text for a definition of the states.
+ See the text for the definition of each state.
  }
@@
 \includegraphics{capillary.pdf}
-\caption{\textbf{Capillary experiment reproducing the crystal-gel state} Colloid volume fraction is $\approx 0.33$ and polymer concentration is $0.40\,$mg/g.
+\caption{\textbf{Capillary experiment reproducing the crystal-gel state.} The colloid volume fraction is $\approx 0.33$ and the polymer concentration is $0.40\,$mg/g.
 Top row: Confocal slices showing the initial network, crystallising network and final crystal-gel network.
@@
 A typical example is
-shown in Fig.~\ref{fig:capillary}, where a system with colloid volume fraction of $\approx 0.33$ and polymer concentration is $0.40\,$mg/g is considered.
-The figure shows the time evolution of the sample at three different times, just after phase separation ($t=340\,\tau_\beta$), and during
-crystal-gel formation ($t=2600\,\tau_\beta$ and $t=6650\,\tau_\beta$). The top-row shows images from quasi two-dimensional confocal slices, while the bottom-row shows
-the three-dimensional reconstruction after single-particle tracking, where crystalline particles are highlighted in both size and color (red).
+shown in Fig.~\ref{fig:capillary}, where a system with the colloid volume fraction of $\approx 0.33$ and the polymer concentration of $0.40\,$mg/g is considered.
+The figure shows the time evolution of the sample at three different times, just after phase separation ($t=340\,\tau_{\rm B}$), and during
+crystal-gel formation ($t=2600\,\tau_{\rm B}$ and $t=6650\,\tau_{\rm B}$). The top-row shows images from quasi two-dimensional confocal slices, while the bottom-row shows
+the three-dimensional reconstruction after single-particle tracking, where crystalline particles are highlighted in both size (large) and color (red).
+
+\bibliographystyle{naturemag2}
+\bibliography{biblio}
+
+\clearpage
 
@@
 
-\bibliographystyle{naturemag2}
-\bibliography{biblio}
-
 \end{document}

### 2016-03-22 crystal-gel/nature/crystal_gel2.tex
Dear All,
Thank you again for your efforts. 
I went through all the files and made many small corrections. 
I did not put any color since you can see the differences from the previous files by using a function of SVN. I just put blue color in the response file (One is a point on which I would like to ask your opinion and the other is about a new movie). Actually, it is difficult to make the movie, it looks OK without it. Please check all modified parts in the three files. Thank you! Hajime

@@
 
-\date{Received \today}
+\date{Received November 26, 2015}
 
@@
 Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface, i.e.
-desublimation, driving evaporation of nearby liquid.
+de-sublimation, driving evaporation of nearby liquid.
 %The latter processes via the gas phase has been overlooked so far.
@@
 Differently from the standard crystal-gel scenario, where crystals only nucleate and grow inside the dense branches of the gel, becoming dynamically arrested, we show that the kinetic
-pathways involving the gas phase (predominanlty the Bergeron process, but also Ostwald ripening) play a fundamental role. We also observe that increasing the strength of the interparticle attraction
+pathways involving the gas phase (predominantly the Bergeron process, but also Ostwald ripening) play a fundamental role. We also observe that increasing the strength of the interparticle attraction
 (by increasing the polymer concentration), instead of forming crystal gels, the system undergoes the classical gelation process, in which
@@
 on the details of our experiments).
-At $t=0$, the increase in the concentration of salt ions within a few Brownian times ($\tau_B$) screens the Coulombic repulsion between colloidal particles,
+At $t=0$, the increase in the concentration of salt ions within a few Brownian times ($\tau_{\rm B}$) screens the Coulomb repulsion between colloidal particles,
 which are then subject to attractive depletion forces, making the system thermodynamically unstable and leading to phase separation. 
@@
 In Fig.~\ref{fig:phase_separation}a, we superimpose the state points with a theoretical phase diagram, which we calculated from generalised free volume theory~\cite{Fleer2008} using the experimental values for the size of the colloids, the radius of gyration of the polymer, and the polymer concentration (see Supplementary Methods).
-For each state point, scans were made at early times (every $10$~s, i.e. $4\tau_B$) and at the late stage (every $30$~s) of the gelation process.
+For each state point, scans were made at early times (every $10$~s, i.e. $4\tau_{\rm B}$) and at the late stage (every $30$~s) of the gelation process.
 
@@
 In Fig.~\ref{fig:phase_separation}a, the state points which did not phase separate are indicated with open triangles, while
-the other symbols indicate samples undergoing phase separation. The (black) continuous line is the gas spinodal computed from free volume theory,
-and it is in good agreement with our experimental points (the small discrepancy at small colloidal volume fractions 
-is common in free volume theory)~\cite{Royall2007,lu2008gelation}. 
+the other symbols indicate samples undergoing phase separation. \textcolor{blue}{The black dashed line is the gas spinodal computed from free volume theory and the 
+black continuous line is the binodal line. They are in good agreement with our experimental points (the small discrepancy at small colloidal volume fractions 
+is common in free volume theory)~\cite{Royall2007,lu2008gelation}. }
 The topology of the network is similar to all our samples at the initial stages. 
@@
 as a function of the cluster size ($s$) at the percolation point. Two particles belong to the same cluster if their distance is
-within the first peak of the radial distribution function. All samples show the same scaling law, $R^2_{\rm g}\sim s^{2/D}$, which is compatible with
+within the first peak of the radial distribution function. All samples show the same scaling law, $R_{\rm g}\sim s^{1/D}$, which is compatible with
 the random gelation universality class exponent ($D=2.53$) as shown by the dashed line in Fig.~\ref{fig:phase_separation}b
 (we do not imply that the formation of the network follows the random gelation universality class, just that our
-fractal dimension is compatible with it).
+fractal (or effective) dimension is compatible with it).
 
@@
 The figure clearly shows that the network restructuring follows a peak in the number of bond-breaking events.
-We have thus found direct evidence of the stress-driven ageing process in colloidal gels that is a direct consequence of the formation of a thin network with the help of
-hydrodynamic interactions in the first stages of the aggregation process.
+We have thus found direct evidence of the stress-driven ageing process in colloidal gels that is a direct consequence of the formation of a thin network with the help of hydrodynamic interactions in the first stages of the aggregation process.
 
@@
 Calculation of the structure factor is done with the Hanning window function, to minimize boundary effects.
-The figure shows that at low wavenumber $q$, the structure factor displays fractal scaling compatible with Guinier law, $S(q)\sim q^{-D}$.
-But a difference in the fractal dimension $D$ between the $\phi\approx 0.33$ and $c_p=0.38$ mg/g state point and other state points starts to become visible. In fact, while states
+The figure shows that at low wavenumber $q$, the structure factor displays fractal scaling compatible with the Guinier law, $S(q)\sim q^{-D}$.
+But a difference in the fractal (or effective) dimension $D$ between the $\phi\approx 0.33$ and $c_p=0.38$ mg/g state point and other state points starts to become visible. In fact, while states
 with high polymer concentration retain the exponent $D=2.5(3)$, 
@@
 The volume fraction is obtained by computing the Voronoi diagram of each configuration, which uniquely assigns a volume to each colloidal particle.
-Only the state point at $\phi\approx 0.33$ and $c_p = 0.38\,$mg/g state shows an increase, up to an asymptotic value that we identify with the composition of the stable crystal phase $\approx 0.72$.
+Only the state point at $\phi\approx 0.33$ and $c_p = 0.38\,$mg/g shows an increase, up to an asymptotic value that we identify with the composition of the stable crystal phase $\approx 0.72$.
 We plot the asymptotic average volume fraction as a function of the number of neighbours in Fig.~\ref{fig:phase_separation}d.
@@
 The comparison shows that the new arrest mechanism produces a network structure with bigger pores, smoother surfaces, and thicker strands (panel b) compared
-to the colloidal gel sample (panel a). The bottom row shows the individual particle positions for slabs with a thickness of 5 particle diameters.
+to the colloidal gel sample (panel a). The right column shows the individual particle positions for slabs with a thickness of 5 particle diameters. 
 From the figure it is immediately clear that the network strands at low polymer concentration are crystalline. The new arrest mechanism thus
@@
 
-To confirm the formation of crystalline seeds, we adopt bond orientational analysis to detect which particles are in crystalline environments~\cite{russo2013interplay}. 
+To confirm the formation of crystalline seeds, we adopt bond orientational analysis to detect which particles are in crystalline environments~\cite{russo2013interplay} 
+\textcolor{blue}{(see Supplementary Information on the detail)}. 
 With this analysis we reveal that crystallization events indeed start occurring for the state point $\phi\approx 0.33$ and $c_p=0.38$ mg/g, which has the lowest
@@
 $\phi\approx 0.33$ and $c_p=0.38$ mg/g. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with 
-a fractal dimension close to random percolation ($D=2.53$), while at large sizes it scales as $D=3$, as in compact crystal growth. This demonstrates that, while 
+the fractal (or effective) dimension close to random percolation ($D=2.53$), while at large sizes it scales as $D=3$, as in compact crystal growth. This demonstrates that, while 
 small crystalline nuclei are nucleated inside the phase-separated liquid branches, once they reach the transverse size of the liquid branch, the growth 
@@
 From the trajectory, we extract slabs of 10 particle diameter thickness for configurations at different times, shown in Fig.~\ref{fig:transitions}a. 
-Particles are coloured according to their phases (gas, liquid, and crystal). For the classification criteria see Supplementary Information. The figure shows the first nucleation 
+Particles are coloured according to their phases (gas, liquid, and crystal). For the classification criteria, see Supplementary Information. The figure shows the first nucleation 
 events inside the liquid network (left panel).
@@
 Note that the topology of the crystal network found at late times, i.e.
-$t=5577\,\tau_B$ (middle panel) and $t=12098\,\tau_B$ (right panel) in Fig.~\ref{fig:transitions}a, differs significantly from that of the liquid network at early stages, i.e.
-$t=2055\,\tau_B$ (left panel) in Fig.~\ref{fig:transitions}a.
+$t=5577\,\tau_{\rm B}$ (middle panel) and $t=12098\,\tau_{\rm B}$ (right panel) in Fig.~\ref{fig:transitions}a, differs significantly from that of the liquid network at early stages, i.e.
+$t=2055\,\tau_{\rm B}$ (left panel) in Fig.~\ref{fig:transitions}a.
 This suggests that the crystal network does not form only by direct freezing of the liquid, but also by other mechanisms which involve the evaporation of the liquid (Bergeron process) and the sublimation of small crystals (Ostwald ripening).
@@
 In order to assess the relative importance of these three growth channels, we analyse the trajectory, and, for every time frame, assign each particle a state between gas, liquid and crystal.
-In order to minimize short-term fluctuations, the state of each colloidal particle is time averaged for $50\,\tau_B$.
+In order to minimize short-term fluctuations, the state of each colloidal particle is time averaged for $50\,\tau_{\rm B}$.
 We then measure the fraction of trajectories with different transition histories.
@@
 In Supplementary Information we confirm these results with a detailed analysis of transition probabilities, including a state for surface particles that are in contact 
-to crystalline particles., see Supplementary Figure~\ref{S-fig:probabilities}.
+to crystalline particles, see Supplementary Figure~\ref{S-fig:probabilities}.
 While the direct freezing of the fluid represents the major contribution to the nuclei growth, the kinetic path via the gas phase (gas$\rightarrow$crystal), also plays a crucial role,
@@
 In clouds there is a mixture of ice crystals and supercooled water. The vapour phase is in coexistence with the liquid phase, but is supersaturated
-with respect to the ice crystals. This causes the water droplets to evaporate and sublimate directly on the ice crystals. Our system can then
+with respect to the ice crystals. This causes the water droplets to evaporate and de-sublimate directly on the ice crystals. Our system can then
 be regarded as a colloidal analogue for this important process, which, for the first time, we can observe at the single-particle level.
@@
 To access glassy slow dynamics in a liquid phase, we may need a deep quench at a high pressure. However, we note that even without strong dynamic asymmetry 
-due to glassiness, bicontinuous phase separation can take place between 35-85 volume \% of the liquid phase in ordinary gas/liquid phase separation; 
+due to glassiness, bicontinuous phase separation can take place between 35-85 volume \% of the liquid phase in ordinary gas/liquid phase separation \cite{onuki2002}; 
 and, thus, porous crystalline structures can be formed, although a thin network structure may be difficult to be formed.}
@@
 We use \textsc{pmma} (poly(methyl methacrylate)) colloids sterically stabilized with methacryloxypropyl terminated \textsc{pdms}(poly(dimethyl siloxane)) and fluorescently labelled with rhodamine isothiocyanate chemically bonded to the \textsc{pmma}.
-The diffusion constant in dilute conditions without polymer allows us to estimate the colloid diameter to $2.3\pm 0.05\, \mu$m. The Brownian time is $\tau_B \approx 2.3\,$s. 
+The diffusion constant in dilute conditions without polymer allows us to estimate the colloid diameter to $2.3\pm 0.05\, \mu$m. The Brownian time is $\tau_{\rm B} \approx 2.3\,$s. 
 We assess that the size distribution of our particle is Gaussian with a polydispersity below 5\% via direct confocal measurements~\cite{Leocmach2013}.
@@
 
-Matching phase boundaries often leads to better determination of volume fractions than colloid diameter measurements~\cite{Poon2012}.
+\textcolor{blue}{Matching phase boundaries often leads to better determination of volume fractions than colloid diameter measurements~\cite{Poon2012}.
 We obtain the best match between theoretical phase diagram and experimental data for $\sigma=2.25\,\mu$m and $R_g=80$ nm (see Supplementary method).
-Therefore the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.071$ and the overlap mass fraction of polymer $2.25$~mg/g.
+Therefore the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.071$ and the overlap mass fraction of polymer $2.25$~mg/g.}
 
@@
 Screening by tetrabutylammonium bromide (Fluka) at saturated concentration brings down the Debye length to about several nanometers practically discarding the repulsion.
-Thus, salt injection can screen the Coulomb repulsion and make the polymer-induced depletion attraction effective, initiating phase separation.
+Thus, salt injection (see below) can screen the Coulomb repulsion and make the polymer-induced depletion attraction effective, initiating phase separation.
 
@@
 A mixture of colloids, polymer, and solvent without salt is  set in the first layer for microscopy observation.
-The second layer is a salt-reservoir with a half-opened structure, which allos us to exchange or insert a reservoir solution.
+The second layer is a salt-reservoir with a half-opened structure, which allows us to exchange or insert a reservoir solution.
 The volume ratio of the first and second compartments is approximately 1:100.
@@
 Salt diffused into the first layer within a few minutes, typically 2 minutes, screened the surface charges, and initiated colloidal aggregation.
+This allows us to observe the process without suffering from harmful turbulent flow.
 
@@
 Our spatial resolution is 192 nm / pixel.
-To be able to follow individual trajectories, we perform a 3D scan every 10~s ($\approx 4\tau_B$) at early time and every 30 s later.
+To be able to follow individual trajectories, we perform a 3D scan every 10~s ($\approx 4\tau_{\rm B}$) at early time and every 30 s later.
 
@@
 The crystal-gel state is reproduced in a sealed cell with all the components of the suspension pre-mixed.
-Colloid volume fraction is virtually identical at $\approx 0.33$ and polymer concentration is $0.40\,$mg/g.
+The colloid volume fraction is virtually identical at $\approx 0.33$ and polymer concentration is $0.40\,$mg/g.
 See Supplementary Figure~\ref{S-fig:capillary}.}
@@
 H.Ts. and J.R. contributed equally to this work. 
-H.Ta. conceived and supervised the project, H.Ts. performed experiments, J.R. analyzed the data, M.L. calculated the phase diagram, and all the authors discussed and wrote the manuscript.
+H.Ta. conceived and supervised the project, H.Ts. performed experiments, J.R. analysed the data, M.L. calculated the phase diagram, and all the authors discussed and wrote the manuscript.
 
@@
 {\bf c,} Structure factors for the late stages of the gelation process. The gray and black lines represents respectively random gelation and compact fractal dimensions.
-\textcolor{blue}{{\bf d,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle for the state points in the very late stage.}
+\textcolor{blue}{{\bf d,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle for the state points in the very late stage. 
+The two horizontal straight lines indicate the characteristic volume fractions in attractive systems, crystal ($\sim 0.72$) and glass ($\sim 0.64$). }
  }
@@
 \caption{{\bf Characterization of the crystal-gel.} {\bf a,} Time evolution of the average size of the crystals (blue line) and the number of crystals (red line) for the state point $\phi\approx 0.33$ and $c_p=0.38$ mg/g.
-{\bf b,} Radius of gyration of crystalline nuclei at the late stages of gelation for the same state point.}
+{\bf b,} Radius of gyration of crystalline nuclei at the late stages of gelation for the same state point.
+\textcolor{blue}{The gray and black straight lines have a slope of $1/2.53$ and $1/3$ respectively}} 
  \label{fig:crystals}
@@
 {\bf a,} Reconstructions from confocal coordinates at $\phi\approx 0.33$ and $c_p=0.38$ mg/g.
-Depth of view is 9.0 $\mu$m, while the later dimension is 41 $\mu$m.
+The depth of view is 9.0 $\mu$m, while the lateral dimension is 41 $\mu$m.
 Particles are drawn to scale and coloured according to their phase.

### 2016-03-22 nature/Supplementary.tex
spinodal = inflexion point of free energy not maximum

--- 2016-03-15 nature/Supplementary.tex
@@
 of the gas and liquid phases just after phase separation (and before crystallization). To analyse the fraction
-of liquid and gas after phase separation we convert the position of particles into a contiunous field smearing
-the position of each particle with a gaussian field of variance $5$ pixels, and then mapping the field
+of liquid and gas after phase separation we convert the position of particles into a continuous field smearing
+the position of each particle with a Gaussian field of variance $5$ pixels, and then mapping the field
 to a cubic grid with $10$ pixels spacing. In this way liquid and gas regions can be identified by
@@
  % field.pdf: 681x483 pixel, 72dpi, 24.02x17.04 cm, bb=0 0 681 483
- \caption{{\bf Distribution function of the gaussian field at $t=490\,\tau_B$.} The gaussian field is applied to
+ \caption{{\bf Distribution function of the Gaussian field at $t=490\,\tau_{\rm B}$.} The Gaussian field is applied to
  a configuration just after phase separation and before crystallization. The vertical dashed line represents the
@@
 \subsection*{Adjusting theoretical and experimental phase diagrams}
+
+\begin{figure}
+\includegraphics{local_phi12.pdf}%
+\caption{{\bf Late stage of the evolution of the local volume fraction around colloidal particles having 12 nearest neighbours.} Samples are shown with the same colour code as in Fig.1 of the main text.
+In particular red circles correspond to the crystallizing sample at $\phi=0.33$ and $cp=0.38\,$mg/g.}%
+\label{fig:local_phi12}%
+\end{figure}
 
@@
 
-Our two constraints is (i) to have the final local volume fraction of the 12-neighboured particles in the crystallising sample to be equal to the theoretical equilibrium crystal volume fraction and (ii) to have the theoretical spinodal line to correspond to the gel region boundary.
-Starting from our initial estimates of $\sigma$ and $R$ we iteratively converge to $R=80$~nm, $\sigma = 2.24\,\mu$m and a crystal at $0.723$.
+Our two constraints are (i) to have the final local volume fraction of the 12-neighboured particles in the crystallising sample to be equal to the theoretical equilibrium crystal volume fraction and (ii) to have the theoretical spinodal line to correspond to the gel region boundary.
+Starting from our initial estimates of $\sigma$ and $R$ we iteratively converge to $R=80$~nm, $\sigma = 2.25\,\mu$m and a crystal (without defects) at $0.723$.
 
@@
 
-The gas spinodal line is given by a local maximum in free energy, a local property of the equation of state.
+The gas spinodal line is given by an inflexion point of the free energy, a local property of the equation of state.
 Therefore, the locally correct CS EoS is able to describe accurately the gas side of the spinodal region up to 0.56.
@@
 
+\section*{Capillary experiment}
+
+\begin{figure*}[!t]
+\includegraphics{capillary.pdf}
+\caption{\textbf{Capillary experiment reproducing the crystal-gel state.} The colloid volume fraction is $\approx 0.33$ and the polymer concentration is $0.40\,$mg/g.
+Top row: Confocal slices showing the initial network, crystallising network and final crystal-gel network.
+Bottom row: Reconstructions from confocal coordinates.
+Particles are drawn to scale and coloured according to their phase.
+Red: crystal; white: liquid; green: gas.
+To highlight the crystalline regions, liquid and gas particles are drawn to a tenth of the crystalline particle size.}
+\label{fig:capillary}
+\end{figure*}
+
+The use of a reservoir of a salt solution through a semi-permeable membrane, allowed us to follow the entire microscopic dynamics in absence of spurious flows, from the first stages of phase separation to the formation of the crystal-gel state.
+In order to confirm the reproducibility of the crystal-gel state, we conducted experiments in a more traditional capillary setting, where all components of the suspension are pre-mixed in a sealed cell.
+A typical example is
+shown in Fig.~\ref{fig:capillary}, where a system with the colloid volume fraction of $\approx 0.33$ and the polymer concentration of $0.40\,$mg/g is considered.
+The figure shows the time evolution of the sample at three different times, just after phase separation ($t=340\,\tau_{\rm B}$), and during
+crystal-gel formation ($t=2600\,\tau_{\rm B}$ and $t=6650\,\tau_{\rm B}$). The top-row shows images from quasi two-dimensional confocal slices, while the bottom-row shows
+the three-dimensional reconstruction after single-particle tracking, where crystalline particles are highlighted in both size (large) and color (red).
+
 \bibliographystyle{naturemag2}
@@
 \clearpage
+
 \section*{Supplementary Movie 1}

### 2016-03-22 nature/response.tex
12 neighbours is a maximum in compaction and thus the dense phase can only have lower volume fraction

@@
 
-We believe that many monoatomic and single-component molecular systems can satisfy all the conditions (i)-(iv) in a certain range of the temperature and pressure. However, the reviewer 
+We believe that many monoatomic and single-component molecular systems can in principle satisfy all the conditions (i)-(iv) in a certain range of the temperature and pressure. However, the reviewer 
 is right in the sense that there may be a practical difficulty in realizing condition (iii), for example, for noble metals. 
 Condition (iii) is necessary to create a porous crystalline structure with large porosity. However, even if condition (iii) is not met, bicontinuous phase-separation structures 
-can be formed near a \textcolor{blue}{symmetric} composition.
-\textcolor{blue}{Thus, we believe a porous crystal structure could be formed even in such a case.} We have added the following sentences to mention this.
+can be formed near a symmetric composition. Thus, we believe a porous crystal structure could be formed even in such a case. We have added the following sentences to mention this 
+with a new reference [42]. 
 We thank the reviewer for her/his thoughtful comment on this important point.
@@
 On the experimental side, we used a Gaussian approximation of the density field to measure the amount of gas and liquid and deduce the volume fraction in the liquid phase.
-The updated Fig 1a shows a good agreement between the liquid-gas tie lines measured experimentally and the predicted binodal line.}
+The updated Fig 1a shows a good agreement between the liquid-gas tie lines measured experimentally and the predicted binodal line.
+\textcolor{blue}{Here we note that the volume fraction of the dense phase is an average over a certain range of the number of neighbours.
+As shown in Fig.
+1d.
+12 neighbours is a maximum in compaction and thus the dense phase can only have lower volume fraction.}
 
@@
 But, in the late stage, the dynamics slows down and thus we reduced the frame rate of 3D scan to cover the long-time period.
-\textcolor{teal}{While the first movies covers
-the first minutes of the process, the second movie shows the evolution over more than 7 hours.} We are afraid that the Reviewer’s impression of a sudden increase in free gas particles and a possible discontinuity between the two movies originates from this difference in the frame rate.
-To address this issue we have now included time stamps on both Supplemental Movies.
+While the first movie covers the first minutes of the process, the second movie shows the evolution over more than 7 hours.
+We are afraid that the Reviewer’s impression of a sudden increase in free gas particles and a possible discontinuity between the two movies originates from this difference in the frame rate.
+To address this issue we have now included time stamps on both Supplemental Movies and explicitly stated the difference in the frame rate in Supplementary Movie 2 Caption.
 
 To address the concern about the reproducibility of our kinetic pathway, we decided to make additional independent experiments
-by using a sample with similar composition, which is inserted into an ordinary glass capillary tube.
+by using a sample with a similar composition, which is inserted into an ordinary glass capillary tube.
 We thus confirm the reproducibility of this type of crystallization pathway.
-An example of crystal-gel porour network obtained through our new capillary experiments
-is found in \textcolor{blue}{Fig. S5}. All details about the experiments are included the Supplementary Information. We would like to thank the Referee for stimulating
-this additional investigation.
+An example of crystal-gel porous  network obtained through our new capillary experiments is found in new Supplementary Fig.
+S5.
+All details about the experiments are also included the Supplementary Information.
+We would like to thank the Referee for stimulating this additional investigation.
 
@@
 As can be seen in Fig. 2b (left panel), the crystal gel has a bicontinuous porous structure. 
-To show the connectivity of the network clearly, we have also made Supplementary Figure S5, highlighting the crystal network.
-We also include movies of a thin slice with a finite thickness. \textcolor{blue}{Please prepare movie(s).} 
+To show the connectivity of the network clearly with a particle-level resolution, we have also made Supplementary Figure S5, highlighting the crystal network. 
+\textcolor{blue}{We also include a movie of a thin slice with a finite thickness. Please prepare a movie.} 
 We hope that this will clarify the situation and convince the reviewer of the fact that our sample indeed forms a porous network of crystals.

### 2016-03-23 crystal-gel/nature/response.tex
New figures.

@@
 The updated Fig 1a shows a good agreement between the liquid-gas tie lines measured experimentally and the predicted binodal line.
-\textcolor{blue}{Here we note that the volume fraction of the dense phase is an average over a certain range of the number of neighbours
-shown in Fig. 1d.}
+\textcolor{blue}{Here we note that the volume fraction of the dense phase is an average over a certain range of the number of neighbours.
+As shown in Fig.
+1d.
+12 neighbours is a maximum in compaction and thus the dense phase can only have lower volume fraction.}
 
@@
 To show the connectivity of the network clearly with a particle-level resolution, we have also made Supplementary Figure S5, highlighting the crystal network. 
-\textcolor{blue}{We also include a movie of a thin slice with a finite thickness. Please prepare a movie.} 
+We also include a movie of a thin slice with a finite thickness (Supplementary Movie 3).
 We hope that this will clarify the situation and convince the reviewer of the fact that our sample indeed forms a porous network of crystals.

### 2016-03-23 nature/Supplementary.tex
New figures.

@@
 crystal-gel formation ($t=2600\,\tau_{\rm B}$ and $t=6650\,\tau_{\rm B}$). The top-row shows images from quasi two-dimensional confocal slices, while the bottom-row shows
-the three-dimensional reconstruction after single-particle tracking, where crystalline particles are highlighted in both size (large) and color (red).
+the three-dimensional reconstruction after single-particle tracking, where crystalline particles are highlighted in both size (large) and color (red), surface particles
+are depicted in blue, gas in green and liquid in white.
 
@@
 
+\section*{Supplementary Movie 3}
+Same as Supplementary Movie 2, but with computer rendered trajectories obtained from the tracking of the colloidal particles.
+Particles are colored according to the following code: crystal (red), surface (blue), gas (green) and liquid (white).
+For display purposes, crystalline particles and gas particles are respectively 400\% and 50\% bigger than other particles (liquid or surface).
+
 \end{document}

### 2016-03-23 nature/Supplementary.tex

@@
 Particles are drawn to scale and coloured according to their phase.
-Red: crystal; white: liquid; green: gas.
-To highlight the crystalline regions, liquid and gas particles are drawn to a tenth of the crystalline particle size.}
+Red: crystal; white: liquid; green: gas; surface: blue.
+To highlight the crystalline regions, liquid, gas, surface particles are drawn to a tenth of the crystalline particle size.}
 \label{fig:capillary}

### 2016-03-24 crystal-gel/nature/response.tex
Dear all,
I have made corrections associated with new Supplementary Movie 3. 
Please check them. 

If you agree with the inclusion of new Movie 3, I think that we are ready to resubmission!

How do you think?

Best wishes, 
Hajime

@@
 To show the connectivity of the network clearly with a particle-level resolution, we have also made Supplementary Figure S5, highlighting the crystal network. 
-We also include a movie of a thin slice with a finite thickness (Supplementary Movie 3).
+We also include a new movie of a thin slice with a finite thickness (Supplementary Movie 3).
 We hope that this will clarify the situation and convince the reviewer of the fact that our sample indeed forms a porous network of crystals.
@@
 We hope that the above explanations and new Supplementary Fig.
-S5 and Movie would convince the reviewer of the formation of a percolated network structure composed of crystals (not isolated crystals diffusing around) and the novel kinetic pathway leading to porous crystalline structures (see also Fig.
+S5 and Supplementary Movie 3 would convince the reviewer of the formation of a percolated network structure composed of crystals (not isolated crystals diffusing around) and the novel kinetic pathway leading to porous crystalline structures (see also Fig.
 2b).

### 2016-03-24 crystal-gel/nature/crystal_gel2.tex
Dear all,
I have made corrections associated with new Supplementary Movie 3. 
Please check them. 

If you agree with the inclusion of new Movie 3, I think that we are ready to resubmission!

How do you think?

Best wishes, 
Hajime

@@
 All samples share the same early stages of spinodal decomposition. 
-A typical phase ordering process observed at $\phi\approx 0.33$ and $c_p=0.38$ mg/g can be seen in Supplementary Movies 1 and 2. 
+A typical phase ordering process observed at $\phi\approx 0.33$ and $c_p=0.38$ mg/g can be seen in Supplementary Movies 1 and 2 (two-dimensional (2D)) 
+and 3 (three-dimensional (3D)). 
 Due to strong dynamical asymmetry between colloids and the solvent \cite{tanaka1999colloid}, the colloidal particles start aggregating, 
@@
 crystallization of the phase separated network.
-Supplementary Movie 2 shows the confocal microscopy video for the crystallizing sample at $\phi\approx 0.33$ and $c_p=0.38$ mg/g,
+Supplementary Movies 2 and 3D show the time-lapse videos of 2D confocal microscopy observation and 3D computer reconstruction, respectively, for the crystallizing sample at $\phi\approx 0.33$ and $c_p=0.38$ mg/g,
 where the process of crystal growth by direct attachment of gas particles can be observed directly. 

### 2016-03-24 nature/crystal_gel2.tex
Changed the position where we refer to the Movies onuki2002 -> Onuki2002

@@
 All samples share the same early stages of spinodal decomposition. 
-A typical phase ordering process observed at $\phi\approx 0.33$ and $c_p=0.38$ mg/g can be seen in Supplementary Movies 1 and 2. 
+A typical early stage phase ordering process observed at $\phi\approx 0.33$ and $c_p=0.38$ mg/g can be seen in Supplementary Movie 1.. 
 Due to strong dynamical asymmetry between colloids and the solvent \cite{tanaka1999colloid}, the colloidal particles start aggregating, 
@@
 The state point at $\phi\approx 0.33$ and $c_p=0.38\,$mg/g reaches close packing at the volume fraction $\approx 0.72$.
-By contrast all other state points, reach close packing at a markedly lower volume fraction, which is indeed close to the volume fraction of the attractive glass state~\cite{pham2002multiple}.
+By contrast all other state points, reach close packing at a markedly
+lower volume fraction, which is indeed close to the volume fraction of the attractive glass state~\cite{pham2002multiple}.
 This result clearly indicates that a single state point follows a different arrest mechanism, in which phase separation is arrested by crystallization and not by glassiness.}
@@
 
-To confirm the formation of crystalline seeds, we adopt bond orientational analysis to detect which particles are in crystalline environments~\cite{russo2013interplay}. 
+Supplementary Movie 2 shows the late stage (after the initial spinodal decomposition) evolution of a 2D confocal slice.
+Even on this raw data the growth of crystalline regions is obvious.
+More quantitatively, we adopt bond orientational analysis to detect which particles are in crystalline environments~\cite{russo2013interplay}
+\textcolor{blue}{(see Supplementary Information on the detail)}. 
 With this analysis we reveal that crystallization events indeed start occurring for the state point $\phi\approx 0.33$ and $c_p=0.38$ mg/g, which has the lowest
@@
 start occurring inside the liquid network, but eventually starts decreasing as the different crystallites grow and merge with each other.
+Supplementary Movie 3 shows a three-dimensional (3D) reconstruction  of the whole phase ordering sequence that emphasizes crystalline particles.
 %The average size of the crystals shows an interesting linear growth regime for $100<t<600$. This linear growth regime goes past the time
@@
 crystallization of the phase separated network.
-Supplementary Movie 2 shows the confocal microscopy video for the crystallizing sample at $\phi\approx 0.33$ and $c_p=0.38$ mg/g,
-where the process of crystal growth by direct attachment of gas particles can be observed directly. 
+%Supplementary Movies 2 and 3D show the time-lapse videos of 2D confocal microscopy observation and 3D computer reconstruction, respectively, for the crystallizing sample at $\phi\approx 0.33$ and $c_p=0.38$ mg/g,
+%where the process of crystal growth by direct attachment of gas particles can be observed directly. 
 From the trajectory, we extract slabs of 10 particle diameter thickness for configurations at different times, shown in Fig.~\ref{fig:transitions}a. 
-Particles are coloured according to their phases (gas, liquid, and crystal). For the classification criteria see Supplementary Information. The figure shows the first nucleation 
+Particles are coloured according to their phases (gas, liquid, and crystal). For the classification criteria, see Supplementary Information. The figure shows the first nucleation 
 events inside the liquid network (left panel).
@@
 To access glassy slow dynamics in a liquid phase, we may need a deep quench at a high pressure. However, we note that even without strong dynamic asymmetry 
-due to glassiness, bicontinuous phase separation can take place between 35-85 volume \% of the liquid phase in ordinary gas/liquid phase separation; 
+due to glassiness, bicontinuous phase separation can take place between 35-85 volume \% of the liquid phase in ordinary gas/liquid phase separation \cite{Onuki2002}; 
 and, thus, porous crystalline structures can be formed, although a thin network structure may be difficult to be formed.}
@@
 Our spatial resolution is 192 nm / pixel.
-To be able to follow individual trajectories, we perform a 3D scan every 10~s ($\approx 4\tau_B$) at early time and every 30 s later.
+To be able to follow individual trajectories, we perform a 3D scan every 10~s ($\approx 4\tau_{\rm B}$) at early time and every 30 s later.
 
+\textcolor{blue}{
 The crystal-gel state is reproduced in a sealed cell with all the components of the suspension pre-mixed.
-Colloid volume fraction is virtually identical at $\approx 0.33$ and polymer concentration is $0.40\,$mg/g.
-See Supplementary Figure~\ref{S-fig:capillary}.
+The colloid volume fraction is virtually identical at $\approx 0.33$ and polymer concentration is $0.40\,$mg/g.
+See Supplementary Figure~\ref{S-fig:capillary}.}
 

### 2016-03-24 nature/response.tex
Small addition.

@@
 it is without any doubt that crystalline nuclei are connected in a porous network. This is evident from a coarse-grained morphology shown in Fig. 2b. 
-We also provide a very clear three-dimensional particle-level representation in a new Supplementary Figure 5.  
+We also provide a very clear three-dimensional particle-level representation in a new Supplementary Figure 5, and a new 3D-rendered movie in Supplementary Movie 3.  
 Together with a novel analysis which determines the crystallization pathway directly by looking 
@@
 To show the connectivity of the network clearly with a particle-level resolution, we have also made Supplementary Figure S5, highlighting the crystal network. 
-\textcolor{blue}{We also include a movie of a thin slice with a finite thickness. Please prepare a movie.} 
+We also include a new movie of a thin slice with a finite thickness (Supplementary Movie 3).
 We hope that this will clarify the situation and convince the reviewer of the fact that our sample indeed forms a porous network of crystals.

### 2016-03-24 nature/response.tex
Took away blue parts in the response letter.

@@
 The updated Fig 1a shows a good agreement between the liquid-gas tie lines measured experimentally and the predicted binodal line.
-\textcolor{blue}{Here we note that the volume fraction of the dense phase is an average over a certain range of the number of neighbours.
+Here we note that the volume fraction of the dense phase is an average over a certain range of the number of neighbours.
 As shown in Fig.
 1d.
-12 neighbours is a maximum in compaction and thus the dense phase can only have lower volume fraction.}
+12 neighbours is a maximum in compaction and thus the dense phase can only have lower volume fraction.
 

### 2016-03-24 crystal-gel/nature/crystal_gel2.tex
New fig 4.

--- 2016-03-22 crystal-gel/nature/crystal_gel2.tex
@@
 
-To confirm the formation of crystalline seeds, we adopt bond orientational analysis to detect which particles are in crystalline environments~\cite{russo2013interplay} 
+Supplementary Movie 2 shows the late stage (after the initial spinodal decomposition) evolution of a 2D confocal slice.
+Even on this raw data the growth of crystalline regions is obvious.
+More quantitatively, we adopt bond orientational analysis to detect which particles are in crystalline environments~\cite{russo2013interplay}
 \textcolor{blue}{(see Supplementary Information on the detail)}. 
@@
 crystallization of the phase separated network.
-Supplementary Movie 2 shows the confocal microscopy video for the crystallizing sample at $\phi\approx 0.33$ and $c_p=0.38$ mg/g,
-where the process of crystal growth by direct attachment of gas particles can be observed directly. 
+%Supplementary Movies 2 and 3D show the time-lapse videos of 2D confocal microscopy observation and 3D computer reconstruction, respectively, for the crystallizing sample at $\phi\approx 0.33$ and $c_p=0.38$ mg/g,
+%where the process of crystal growth by direct attachment of gas particles can be observed directly. 
 From the trajectory, we extract slabs of 10 particle diameter thickness for configurations at different times, shown in Fig.~\ref{fig:transitions}a. 
@@
 To access glassy slow dynamics in a liquid phase, we may need a deep quench at a high pressure. However, we note that even without strong dynamic asymmetry 
-due to glassiness, bicontinuous phase separation can take place between 35-85 volume \% of the liquid phase in ordinary gas/liquid phase separation \cite{onuki2002}; 
+due to glassiness, bicontinuous phase separation can take place between 35-85 volume \% of the liquid phase in ordinary gas/liquid phase separation \cite{Onuki2002}; 
 and, thus, porous crystalline structures can be formed, although a thin network structure may be difficult to be formed.}
@@
 {\bf b,} Fraction of particles for each phase. Gray vertical lines indicate the times shown in panel a. 
-{\bf c,} Each particle during its trajectory changes between three states: gas, liquid and crystal. Direct crystallization accounts for 54\% of particles trajectories in which
-gas or liquid particles transition to the crystal state without liquid evaporation or crystal de-sublimation. The Bergeron process accounts for 35\% of particle trajectories in which liquid particles
-transition to the gas state before crystallizing. Ostwald ripening accounts for 11\% of particle trajectories in which crystal particle transition to the gas state before returning
+{\bf c,} Each particle during its trajectory changes between three states: gas, liquid and crystal. Direct crystallization (continuous arrow) accounts for 54\% of particles trajectories in which
+gas or liquid particles transition to the crystal state without liquid evaporation or crystal de-sublimation.
+The Bergeron process (dotted arrow) accounts for 35\% of particle trajectories in which liquid particles
+transition to the gas state before crystallizing. Ostwald ripening (dashed arrow) accounts for 11\% of particle trajectories in which crystal particle transition to the gas state before returning
 to the crystal state.}}

### 2016-03-25 crystal-gel/nature/crystal_gel2.tex
Just put blue color in the main text where we are mention Supplementary Movie 3.  Hajime

@@
 black continuous line is the binodal line. They are in good agreement with our experimental points (the small discrepancy at small colloidal volume fractions 
-is common in free volume theory)~\cite{Royall2007,lu2008gelation}. }
+is common in free volume theory)~\cite{Royall2007,lu2008gelation}.}
 The topology of the network is similar to all our samples at the initial stages. 
@@
 start occurring inside the liquid network, but eventually starts decreasing as the different crystallites grow and merge with each other.
-Supplementary Movie 3 shows a three-dimensional (3D) reconstruction  of the whole phase ordering sequence that emphasizes crystalline particles.
+\textcolor{blue}{Supplementary Movie 3 shows a three-dimensional (3D) reconstruction  of the whole phase ordering sequence that emphasizes crystalline particles.}
 %The average size of the crystals shows an interesting linear growth regime for $100<t<600$. This linear growth regime goes past the time

### 2016-06-08 presentation/talk_Roscoff.tex
Talk for Roscoff Bergeron process

\documentclass[xcolor=table]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[british]{babel}
\usepackage[super]{nth}
%\usetheme{Boadilla}
%\usecolortheme{rose}
%\usecolortheme{crane}
%\usefonttheme{structuresmallcapsserif}
%\setbeamertemplate{navigation symbols}{}

\definecolor{Main}{rgb}{0.74, 0.13, 0.19}
\definecolor{Accent1}{rgb}{0.76,0.36,0.13}
\definecolor{Accent2}{rgb}{0.54,0.1,0.4}

\mode<presentation>{\usetheme{ilm}}
%\usecolortheme{rose}
%\useinnertheme[shadow]{circles}
%\usecolortheme{whale}
%\useoutertheme{infolines}

%\usecolortheme[named=Accent1]{structure}

%\setbeamerfont{page number in head/foot}{size=\large}
%\setbeamercolor{page number in head/foot}{fg=Main}
%% page/total
%%\setbeamertemplate{footline}[frame number]
%% pas de total
%\setbeamertemplate{footline}{%
%    	\hfill%
%	\usebeamercolor[fg]{page number in head/foot}%
%	\usebeamerfont{page number in head/foot}%
%	\insertframenumber\kern1em\vskip2pt%
%}
%
%\setbeamersize{text margin left=1em}
%\setbeamersize{text margin right=1em}

\usepackage[overlay,absolute]{textpos}
\setlength{\TPHorizModule}{10mm}
\setlength{\TPVertModule}{\TPHorizModule}
\textblockorigin{10mm}{10mm} % start everything near the top-left corner
\setlength{\parindent}{0pt}

%font
\usepackage[T1]{fontenc}
\usepackage{times}
%\usepackage[oldstylenums]{kpfonts}

%\include{macros}

%proper math and math symbols
%\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{datenumber,fp}

\usepackage{siunitx}

\usepackage{tabu}
\usepackage{multirow}
\usepackage{booktabs}

% Allow the usage of graphics (.jpg, .png, etc.) in the document
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,backgrounds, calc, positioning, topaths,chains, intersections, decorations.markings, decorations.text, shapes.geometric, matrix,patterns,mindmap}
%\usetikzlibrary{positioning, patterns,topaths,chains,matrix}

\usepackage{pgfplots}
\usepackage{pgfplotstable}
\pgfplotsset{compat=1.9}
\usepgfplotslibrary{groupplots}
\usepgfplotslibrary{external}
\makeatletter
\newcommand*{\overlaynumber}{\number\beamer@slideinframe}
\tikzset{
  beamer externalizing/.style={%
    execute at end picture={%
      \tikzifexternalizing{%
        \ifbeamer@anotherslide
        \pgfexternalstorecommand{\string\global\string\beamer@anotherslidetrue}%
        \fi
      }{}%
    }%
  },
  external/optimize=false
}
\let\orig@tikzsetnextfilename=\tikzsetnextfilename
\renewcommand\tikzsetnextfilename[1]{\orig@tikzsetnextfilename{#1-\overlaynumber}}
\makeatother

\tikzset{every picture/.style={beamer externalizing}}
\tikzexternalize
\tikzsetexternalprefix{fig_presentation/}
%\tikzset{external/optimize=false}
%\tikzset{external/force remake}

%link or play movies
\usepackage{multimedia}

%beamer related package

\usepackage{todonotes}
\presetkeys{todonotes}{inline}{}

%bibliography
\usepackage[style=authoryear-comp, language=british,eprint=false, url=false, doi=false, sortcites=true, sorting=none, isbn=false, firstinits=true,maxcitenames=6]{biblatex}
%minimal citations
\AtEveryCitekey{%
	\clearfield{title}
	\clearfield{pages}
	\clearfield{volume}
	\clearfield{number}
	\clearfield{month}}
\newcommand{\myfullcite}[1]{{\scriptsize\fullcite{#1}}}
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{%
  \printtext{\bibstring{in}\intitlepunct}}}
%\bibliography{biblio}

\newcolumntype{P}[1]{>{\raggedright}p{#1}}

\institute[iLM]{iLM, Univ. Claude Bernard Lyon 1 and CNRS, France}
\title[crystal gel]{A novel route to the spontaneous formation of porous crystals via viscoelastic phase separation}
\author[M. Leocmach]{Mathieu Leocmach}
\date{30 June 2016}
\titlegraphic{
	\begin{tabu}{X[c]X[c]X[c]}
		\includegraphics[height=3\baselineskip]{Tsurusawa}&
		\includegraphics[height=3\baselineskip]{John}&
		\includegraphics[height=3\baselineskip]{Tanaka}\\
		Hideyo Tsurusawa & John Russo & Hajime Tanaka\\
		U. Tokyo & U. Bristol & U. Tokyo\\
	\end{tabu}

	%\vfill
	%\includegraphics[height=2\baselineskip]{logo_ens-lyon}\quad
	%\includegraphics[height=2\baselineskip]{logo_ums_grand}\quad
	%\includegraphics[height=2\baselineskip]{../../Yaourt/CNRSfilaire-Q}\quad
	%\includegraphics[height=2\baselineskip]{CRPP}\quad
	%\includegraphics[height=2\baselineskip,clip=true, trim=6mm 14mm 6mm 0]{NEW-Logo-ERC-OUTLINE}
	}

\newlength{\mylength}

%\includeonly{creep_beamer}

\begin{document}
\tikzset{every mark/.append style={scale=0.8}}
\pgfplotsset{every axis/.append style={footnotesize}}

\pgfplotscreateplotcyclelist{earthy}{%
{red!40!black,mark=o},
{red!60!black,mark=triangle, every mark/.append style={rotate=180}},
{red!80!black,mark=square},
{red,mark=triangle},
{red!80!yellow, mark=diamond},
red!60!yellow,
red!40!yellow,
}

\AtBeginSection[]{
	\addtocounter{framenumber}{-1}
	\begin{frame}[plain]
		\tableofcontents[currentsection, hideothersubsections]
	\end{frame}
}

\begin{frame}{\pgfuseimage{cnrs-logo}\hspace*{0.3cm}\pgfuseimage{ucbl-logo}}%[plain]
	\titlepage
\end{frame}

\setcounter{framenumber}{0}

\begin{frame}{Let it snow!}
\hfill Wegener-\alert{Bergeron}-Findeisen \alert{process}
\begin{columns}[b]
\column{0.4\textwidth}
\includegraphics[width=\textwidth,clip=true,trim=0 0 0 1cm]{Pitter1.jpg}

\vspace{1em}\includegraphics[width=\textwidth]{bergeron.jpg}
\column{0.40\textwidth}
\includegraphics[width=\textwidth,clip=true,trim=0 0 0.8mm 0]{queries-figure-1.jpg}

\vspace{1em}\includegraphics[width=\textwidth]{OKC-Fallstreak.jpg}
\end{columns}
\end{frame}

\begin{frame}{Wavelength}
	\tikzsetnextfilename{lambda}
	%\input{fig_soundbite_palavas/lambda.pgf}
	\textcolor{ilmorange}{$\blacktriangleright$} $\lambda\propto e$ compatible with nucleation rate

\end{frame}

\appendix
\newcounter{finalframe}
\setcounter{finalframe}{\value{framenumber}}

\begin{frame}[plain]
\end{frame}

\setcounter{framenumber}{\value{finalframe}}
\end{document}

### 2016-06-09 crystal-gel/presentation/talk_Roscoff.tex
why study crystal nucleation
Colloid + Polymer

--- 2016-06-08 presentation/talk_Roscoff.tex
@@
 
+\begin{frame}{Crystallisation}
+	\tikzsetnextfilename{whyx}%
+	\input{fig_presentation/whyx.pgf}
+
+	\begin{block}{Nucleation}
+	\begin{itemize}
+		\item 10-1000 molecules $\rightarrow$ difficult to observe
+		\item classical nucleation theory (CNT): fluid $\xrightarrow{\text{1 step}}$ crystal
+	\end{itemize}
+	\end{block}
+\end{frame}
+
 \begin{frame}{Let it snow!}
@@
 
-\begin{frame}{Wavelength}
-	\tikzsetnextfilename{lambda}
-	%\input{fig_soundbite_palavas/lambda.pgf}
-	\textcolor{ilmorange}{$\blacktriangleright$} $\lambda\propto e$ compatible with nucleation rate
+\begin{frame}{Colloid + Polymer}
 
+\begin{columns}
+	\column{0.45\textwidth}
+	\begin{tikzpicture}
+		\let\mrad\relax
+		\newlength\mrad
+		\setlength{\mrad}{1em}
+
+		%schemas interactions
+
+		%repulsion
+		\begin{scope}[xshift=2\mrad]
+		\node[circle, inner sep=0, minimum width=4\mrad, inner color=Main, outer color=white]  (puri) {}; 
+		\foreach \x in {0,30,...,330}{
+			\draw[decoration={random steps,segment length=0.1\mrad, amplitude=0.08\mrad},decorate] (0,0) -- (\x:1\mrad);
+		}
+		\fill[gray, radius=0.8\mrad] circle;
+		\foreach \x in {15,75,...,315}{
+			\node[Accent2, font=\tiny] at (\x:0.6\mrad) {+};
+		}
+		\end{scope}
+
+		%spheres collantes
+		\begin{scope}[xshift=\textwidth-2.5\mrad]
+		\node[draw,circle, inner sep=0, minimum width=2.5\mrad, green!50!black, dashed]  (colli) {}; 
+		\draw[green!50!black, radius=1.25\mrad, dashed] (1.6\mrad,0) circle;
+		\fill[gray, radius=0.8\mrad] circle;
+		\fill[gray, radius=0.8\mrad] (1.6\mrad,0) circle;
+		\end{scope}
+
+		\path (puri) -- (colli) coordinate[midway] (m);
+
+		%spheres dures
+		\begin{scope}[shift=(m)]
+		\node[circle, inner sep=0, minimum width=2.2\mrad, inner color=Main, outer color=white]  (duri) {}; 
+		\foreach \x in {0,30,...,330}{
+			\draw[decoration={random steps,segment length=0.1\mrad, amplitude=0.08\mrad},decorate] (0,0) -- (\x:1\mrad);
+		}
+		\fill[gray, radius=0.8\mrad] circle;
+		\foreach \x in {15,75,...,315}{
+			\node[Accent2, font=\tiny] at (\x:0.6\mrad) {+};
+		}
+		\end{scope}
+
+		%fleches
+		\draw[->] (puri) -- (duri) node[midway,above,font=\footnotesize]{ions};
+		\draw[->] (duri) -- (colli)node[midway, above,font=\footnotesize]{PS};
+		% \draw[->] (puri) -- (ps) -- (depli);
+
+		%titres
+		\node[text width=6em,anchor=base west, inner xsep=0] (pur) at (puri.west|-puri.north) {repulsion};
+
+		\node[text width=6em,anchor=base, align=center] (dur) at (duri|-pur.base) {hard};
+
+		\node[text width=6em, anchor=base east, inner xsep=0,xshift=1.5\mrad, align=right] (coll) at (colli.east|-pur.base) {depletion};
+	\end{tikzpicture}
+
+	\bigskip
+	\includegraphics[width=\textwidth]{/home/mathieu/Documents/Recherches/tex/CNRS-2015/Renth_sedim.jpg}\\
+	\textit{\footnotesize Renth 1999}
+	\begin{itemize}
+	\item triple coexistence
+	\item various kinetic paths
+	\end{itemize}
+
+	\column{0.45\textwidth}
+	\includegraphics[width=\textwidth,clip=true,trim=0 0 0 4.5cm]{Renth_diag.pdf}
+	\end{columns}
 \end{frame}

### 2016-06-13 crystal-gel/presentation/talk_Roscoff.tex
biblio crystal-gel
experimental slide

@@
 	\column{0.45\textwidth}
+	\tikzsetnextfilename{interactions}%
 	\begin{tikzpicture}
@@
 		\node[text width=6em, anchor=base east, inner xsep=0,xshift=1.5\mrad, align=right] (coll) at (colli.east|-pur.base) {depletion};
-	\end{tikzpicture}
+	\end{tikzpicture}\\
 
 	\bigskip
-	\includegraphics[width=\textwidth]{/home/mathieu/Documents/Recherches/tex/CNRS-2015/Renth_sedim.jpg}\\
-	\textit{\footnotesize Renth 1999}
+	\includegraphics[width=\textwidth]{../../CNRS-2015/Renth_sedim.jpg}\\
+	\textit{\footnotesize Poon PRL 1999, Renth PRE 2001}
 	\begin{itemize}
@@
 	\includegraphics[width=\textwidth,clip=true,trim=0 0 0 4.5cm]{Renth_diag.pdf}
+	\end{columns}
+\end{frame}
+
+\begin{frame}{Another class of gel? Crystal-gel?}
+	\begin{itemize}
+	\item Phase separation arrested by crystallisation
+	\begin{description}
+		\item[polymer blend] Tanaka \& Nishi \textit{PRL} 1985
+		\item[simulations] Soga \textit{J.Chem. Phys.} 1999; Fortini \textit{PRE} 2008 
+	\end{description}
+	\item Spinodal then crystallisation
+	\begin{description}
+		\item[simulations] Pérez \textit{Langmuir} 2011
+		\item[microgravity] (triple coexistence) Sabin \textit{PRL} 2012
+	\end{description}
+	\end{itemize}
+	\includegraphics[width=\textwidth]{micrograv_Sabin2012}
+\end{frame}
+
+\begin{frame}{Experimental}
+How to do that in 3D at particle level \alert{without a space station}?
+\begin{itemize}
+\item index matched + \alert{\SI{2.3}{\micro\metre} colloids} + confocal
+\item density match + $T$-control
+\item preparation protocol without flow
+\end{itemize}
+\bigskip
+\begin{columns}
+	\column{0.265\textwidth}
+		Colloids+polymers\\
+		{\footnotesize no screening
+		\begin{itemize}
+			\item Dispersed
+			\item Homogeneous
+		\end{itemize}
+		\bigskip
+		\begin{beamercolorbox}[rounded=true]{block body}PMMA + PS\\
+		in decalin + CHB $\epsilon_r\approx5$\\
+		$\kappa^{-1}>\SI{10}{\micro\metre}$
+		\end{beamercolorbox}}
+	\column{0.4\textwidth}
+		\tikzsetnextfilename{chemin_gelification}%
+		\input{chemin_gelification.pgf}
+	\column{0.335\textwidth}
+		\tikzsetnextfilename{cell}%
+		\input{cell.pgf}
 \end{columns}

### 2016-06-13 presentation/talk_Roscoff.tex
Exploring phase diagram Outlier at later times Real space: crystal gel

@@
 
-\begin{frame}{Wavelength}
-	\tikzsetnextfilename{lambda}
-	%\input{fig_soundbite_palavas/lambda.pgf}
-	\textcolor{ilmorange}{$\blacktriangleright$} $\lambda\propto e$ compatible with nucleation rate
+\begin{frame}{Colloid + Polymer}
 
+\begin{columns}
+	\column{0.45\textwidth}
+	\tikzsetnextfilename{interactions}%
+	\begin{tikzpicture}
+		\let\mrad\relax
+		\newlength\mrad
+		\setlength{\mrad}{1em}
+
+		%schemas interactions
+
+		%repulsion
+		\begin{scope}[xshift=2\mrad]
+		\node[circle, inner sep=0, minimum width=4\mrad, inner color=Main, outer color=white]  (puri) {}; 
+		\foreach \x in {0,30,...,330}{
+			\draw[decoration={random steps,segment length=0.1\mrad, amplitude=0.08\mrad},decorate] (0,0) -- (\x:1\mrad);
+		}
+		\fill[gray, radius=0.8\mrad] circle;
+		\foreach \x in {15,75,...,315}{
+			\node[Accent2, font=\tiny] at (\x:0.6\mrad) {+};
+		}
+		\end{scope}
+
+		%spheres collantes
+		\begin{scope}[xshift=\textwidth-2.5\mrad]
+		\node[draw,circle, inner sep=0, minimum width=2.5\mrad, green!50!black, dashed]  (colli) {}; 
+		\draw[green!50!black, radius=1.25\mrad, dashed] (1.6\mrad,0) circle;
+		\fill[gray, radius=0.8\mrad] circle;
+		\fill[gray, radius=0.8\mrad] (1.6\mrad,0) circle;
+		\end{scope}
+
+		\path (puri) -- (colli) coordinate[midway] (m);
+
+		%spheres dures
+		\begin{scope}[shift=(m)]
+		\node[circle, inner sep=0, minimum width=2.2\mrad, inner color=Main, outer color=white]  (duri) {}; 
+		\foreach \x in {0,30,...,330}{
+			\draw[decoration={random steps,segment length=0.1\mrad, amplitude=0.08\mrad},decorate] (0,0) -- (\x:1\mrad);
+		}
+		\fill[gray, radius=0.8\mrad] circle;
+		\foreach \x in {15,75,...,315}{
+			\node[Accent2, font=\tiny] at (\x:0.6\mrad) {+};
+		}
+		\end{scope}
+
+		%fleches
+		\draw[->] (puri) -- (duri) node[midway,above,font=\footnotesize]{ions};
+		\draw[->] (duri) -- (colli)node[midway, above,font=\footnotesize]{PS};
+		% \draw[->] (puri) -- (ps) -- (depli);
+
+		%titres
+		\node[text width=6em,anchor=base west, inner xsep=0] (pur) at (puri.west|-puri.north) {repulsion};
+
+		\node[text width=6em,anchor=base, align=center] (dur) at (duri|-pur.base) {hard};
+
+		\node[text width=6em, anchor=base east, inner xsep=0,xshift=1.5\mrad, align=right] (coll) at (colli.east|-pur.base) {depletion};
+	\end{tikzpicture}\\
+
+	\bigskip
+	\includegraphics[width=\textwidth]{../../CNRS-2015/Renth_sedim.jpg}\\
+	\textit{\footnotesize Poon PRL 1999, Renth PRE 2001}
+	\begin{itemize}
+	\item triple coexistence
+	\item various kinetic paths
+	\end{itemize}
+
+	\column{0.45\textwidth}
+	\includegraphics[width=\textwidth,clip=true,trim=0 0 0 4.5cm]{Renth_diag.pdf}
+	\end{columns}
+\end{frame}
+
+\begin{frame}{Another class of gel? Crystal-gel?}
+	\begin{itemize}
+	\item Phase separation arrested by crystallisation
+	\begin{description}
+		\item[polymer blend] Tanaka \& Nishi \textit{PRL} 1985
+		\item[simulations] Soga \textit{J.Chem. Phys.} 1999; Fortini \textit{PRE} 2008 
+	\end{description}
+	\item Spinodal then crystallisation
+	\begin{description}
+		\item[simulations] Pérez \textit{Langmuir} 2011
+		\item[microgravity] (triple coexistence) Sabin \textit{PRL} 2012
+	\end{description}
+	\end{itemize}
+	\includegraphics[width=\textwidth]{micrograv_Sabin2012}
+\end{frame}
+
+\begin{frame}{Experimental}
+How to do that in 3D at particle level \alert{without a space station}?
+\begin{itemize}
+\item index matched + \alert{\SI{2.3}{\micro\metre} colloids} + confocal
+\item density match + $T$-control
+\item preparation protocol without flow
+\end{itemize}
+\bigskip
+\begin{columns}
+	\column{0.265\textwidth}
+		Colloids+polymers\\
+		{\footnotesize no screening
+		\begin{itemize}
+			\item Dispersed
+			\item Homogeneous
+		\end{itemize}
+		\bigskip
+		\begin{beamercolorbox}[rounded=true]{block body}PMMA + PS\\
+		in decalin + CHB $\epsilon_r\approx5$\\
+		$\kappa^{-1}>\SI{10}{\micro\metre}$
+		\end{beamercolorbox}}
+	\column{0.4\textwidth}
+		\tikzsetnextfilename{chemin_gelification}%
+		\input{chemin_gelification.pgf}
+	\column{0.335\textwidth}
+		\tikzsetnextfilename{cell}%
+		\input{cell.pgf}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Exploring phase diagram}
+	\vspace{\baselineskip}
+	\tikzsetnextfilename{phdiag_cluster}%
+	\begin{tikzpicture}[
+		every axis/.style={
+		xlabel absolute, every axis x label/.append style={anchor=base, yshift=-1em},
+		%ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
+		}]
+	\pgfplotscreateplotcyclelist{samples}{%
+	blue,mark=square*,\\%
+	cyan,mark=square\\%
+	red, thick, mark=*\\%
+	red!65!yellow, mark=otimes\\%
+	red!30!yellow, mark=oplus\\%
+	red!65!black, mark=o\\%
+	}%
+	\begin{groupplot}[
+		group style={
+				group name=g, group size=2 by 2,
+				horizontal sep=4em,
+				vertical sep=3em,
+				%x descriptions at=edge bottom,
+			},
+		cycle list name=samples,
+		width=0.51\textwidth,
+		height=0.5\linewidth,
+		axis on top,
+	]
+	\nextgroupplot[
+		title={phase diagram},
+		xlabel={$\phi$}, ylabel={$c_p$ (\si{\milli\gram/\gram})},
+		xmin=0,xmax=0.74,
+		ymin=0, ymax=1.9, ytick={0,0.5,...,2},
+		mark options={solid}
+		]
+		\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
+		\addplot file {/data/Documents/tex/svn/crystal-gel/sample0.phd};
+		\addplot file {/data/Documents/tex/svn/crystal-gel/sample1.phd};
+		\addplot file {/data/Documents/tex/svn/crystal-gel/sample2.phd};
+		\addplot file {/data/Documents/tex/svn/crystal-gel/sample3.phd};
+		\addplot file {/data/Documents/tex/svn/crystal-gel/sample4.phd};
+		\addplot file {/data/Documents/tex/svn/crystal-gel/sample5.phd};
+		\end{scope}
+		\addplot[black, dashed,no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
+		\addplot[black, dashed,no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_sl.phd};
+		\addplot[black, no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_bl.phd};
+		\addplot[black, no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_bg.phd};
+		\addplot[fill=lightgray!50, no marks, draw=none] file {/data/Documents/tex/svn/crystal-gel/fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
+		\addplot[fill=lightgray!50, no marks, draw=none] file {/data/Documents/tex/svn/crystal-gel/fluidcrystal_x.phd} node[above left] (x) at (rel axis cs:1,0.5) {Crystal}  -| (rel axis cs:1,0) \closedcycle;
+		%\draw[gray] (axis cs:0.371290311343, 0.12299370925) -- (axis cs:0.680497991395,0.0372636873433);
+		%\draw[gray, dotted] (axis cs:3.40152108798e-05,0.624739653238) -- (axis cs:0.722831684246, 0.0726528862772);
+		\draw[->] (x.base) --(rel axis cs:0.98,0.01);
+		\addplot[only marks, mark=triangle] file{/data/Documents/tex/svn/crystal-gel/fluid_samples.phd};
+
+	\nextgroupplot[
+		title={cluster size distribution},
+		xlabel={$s$}, ylabel={$R_g/\sigma$},
+		xmin=1, xmax=4e3,
+		ymin=0.5, ymax=30,
+		xmode=log, ymode=log, 
+		clip mode=individual,
+		]
+		\addplot file{/data/Documents/tex/svn/crystal-gel/data/fig1b_sqrt_1.dat};
+		\addplot file{/data/Documents/tex/svn/crystal-gel/data/fig1b_sqrt_2.dat};
+		\addplot file{/data/Documents/tex/svn/crystal-gel/data/fig1b_sqrt_3.dat};
+		\addplot file{/data/Documents/tex/svn/crystal-gel/data/fig1b_sqrt_4.dat};
+		\addplot file{/data/Documents/tex/svn/crystal-gel/data/fig1b_sqrt_5.dat};
+		\addplot[gray, ultra thick, domain=1:4e3] {0.629*x^(1/2.53)} node [left] at (rel axis cs:1,0.9){$D=2.53$};
+
+	\end{groupplot}
+	%\draw(g c1r1.outer south west) -- +(\textwidth,0);
+	\end{tikzpicture}
+	\begin{itemize}
+	\item Good agreement with theoretical phase diagram
+	\item At percolation time, compatible with random gelation
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Outlier at later times}
+	\vspace{\baselineskip}
+	\tikzsetnextfilename{Sq_localphi}%
+	\begin{tikzpicture}[
+		every axis/.style={
+		xlabel absolute, every axis x label/.append style={anchor=base, yshift=-1em},
+		%ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
+		}]
+	\pgfplotscreateplotcyclelist{samples}{%
+	blue,mark=square*,\\%
+	cyan,mark=square\\%
+	red, thick, mark=*\\%
+	red!65!yellow, mark=otimes\\%
+	red!30!yellow, mark=oplus\\%
+	red!65!black, mark=o\\%
+	}%
+	\begin{groupplot}[
+		group style={
+				group name=g, group size=2 by 2,
+				horizontal sep=4em,
+				vertical sep=3em,
+				%x descriptions at=edge bottom,
+			},
+		cycle list name=samples,
+		width=0.51\textwidth,
+		height=0.5\linewidth,
+		axis on top,
+	]
+	\nextgroupplot[
+		title={structure factor},
+		xlabel={$q\sigma$}, ylabel={$S(q)$},
+		xmode=log, ymode=log, xmin=0.2, xmax=20, ymin=0.1,
+		restrict x to domain=-1.8:3, unbounded coords=discard, 
+		%clip mode=individual
+		]
+		\addplot+[clip mode=individual] file{/data/Documents/tex/svn/crystal-gel/data/fig3c_1.dat};
+		\addplot+[clip mode=individual] file{/data/Documents/tex/svn/crystal-gel/data/fig3c_2.dat};
+		\addplot file{/data/Documents/tex/svn/crystal-gel/data/fig3c_3.dat};
+		\addplot+[clip mode=individual] file{/data/Documents/tex/svn/crystal-gel/data/fig3c_4.dat};
+		\addplot+[clip mode=individual] file{/data/Documents/tex/svn/crystal-gel/data/fig3c_5.dat};
+		\addplot[black, ultra thick, domain=0.1:3] {2.14605*x^(-3)} node[pos=0.92,left=0.5em] {$D=3$};
+		\addplot[gray, ultra thick, domain=0.1:3] {3.98797*x^(-2.53)} node [pos=0.65, above right] {$D=2.53$};
+
+	\nextgroupplot[
+		title={local volume fraction},
+		xlabel={\# neighbours}, ylabel={$\phi$ [\%]},
+		xmin=0, ymin=0, ytick={0,10,30,50},
+		extra y ticks={64, 72}, extra y tick style={grid=major}, extra y tick label={},
+		legend pos=north west, legend style={font=\footnotesize},
+		no markers,
+		/pgfplots/error bars/y dir=both, /pgfplots/error bars/y explicit,
+		]
+		\addplot table[y error=err]{/data/Documents/tex/svn/crystal-gel/data/vol_err_phi_0.10_cp_0.82.scaled};
+		\addplot table[y error=err]{/data/Documents/tex/svn/crystal-gel/data/vol_err_phi_0.10_cp_1.36.scaled};
+		\addplot table[y error=err]{/data/Documents/tex/svn/crystal-gel/data/vol_err_phi_0.25_cp_0.38.scaled};
+		\addplot table[y error=err]{/data/Documents/tex/svn/crystal-gel/data/vol_err_phi_0.25_cp_0.48.scaled};
+		\addplot table[y error=err]{/data/Documents/tex/svn/crystal-gel/data/vol_err_phi_0.25_cp_0.57.scaled};
+
+	\end{groupplot}
+	%\draw(g c1r1.outer south west) -- +(\textwidth,0);
+	\end{tikzpicture}
+	\begin{itemize}
+	\item Final stucture compatible with random gelation
+	\item Glassy local $\phi$ for 12-neighboured particles
+	\end{itemize}
+	\alert{Except} \textcolor{red}{$\bullet$}: compact structure and higher local $\phi$. %\textcolor{ilmorange}{$\Rightarrow$} Crystal?
+\end{frame}
+
+\begin{frame}{Real space: crystal gel!}
+	\tikzsetnextfilename{phdiag_arrows}%
+	\begin{tikzpicture}[
+			every axis/.style={
+			xlabel absolute, every axis x label/.append style={anchor=base, yshift=-1em},
+			%ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
+			}]
+		\pgfplotscreateplotcyclelist{samples}{%
+		blue,mark=square*,\\%
+		cyan,mark=square\\%
+		red, thick, mark=*\\%
+		red!65!yellow, mark=otimes\\%
+		red!30!yellow, mark=oplus\\%
+		red!65!black, mark=o\\%
+		}%
+		\begin{axis}[
+			cycle list name=samples,
+			width=0.51\textwidth,
+			height=0.5\linewidth,
+			axis on top,
+			xlabel={$\phi$}, ylabel={$c_p$ (\si{\milli\gram/\gram})},
+			xmin=0,xmax=0.74,
+			ymin=0, ymax=1.9, ytick={0,0.5,...,2},
+			mark options={solid}
+			]
+			\begin{scope}[/pgfplots/skip coords between index={0}{1}, /pgfplots/skip coords between index={2}{3}]%[draw=none,mark repeat=2,mark phase=2]
+			\addplot table {/data/Documents/tex/svn/crystal-gel/sample0.phd};
+			\addplot table {/data/Documents/tex/svn/crystal-gel/sample1.phd};
+			\addplot table {/data/Documents/tex/svn/crystal-gel/sample2.phd}[->] -- (rel axis cs:0.95,0.2);
+			\addplot table {/data/Documents/tex/svn/crystal-gel/sample3.phd};
+			\addplot table {/data/Documents/tex/svn/crystal-gel/sample4.phd};
+			\addplot table {/data/Documents/tex/svn/crystal-gel/sample5.phd}[->] -- (rel axis cs:0.95,0.7);
+			\end{scope}
+			\addplot[black, dashed,no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
+			\addplot[black, dashed,no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_sl.phd};
+			\addplot[black, no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_bl.phd};
+			\addplot[black, no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_bg.phd};
+			\addplot[fill=lightgray!50, no marks, draw=none] file {/data/Documents/tex/svn/crystal-gel/fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {/data/Documents/tex/svn/crystal-gel/fluidcrystal_x.phd} %node[above left] (x) at (rel axis cs:1,0.5) {Crystal}  
+			-| (rel axis cs:1,0) \closedcycle;
+			%\draw[->] (x.base) --(rel axis cs:0.98,0.01);
+			\addplot[only marks, mark=triangle] file{/data/Documents/tex/svn/crystal-gel/fluid_samples.phd};
+	\end{axis}
+	\end{tikzpicture}
+	\includegraphics[height=11\baselineskip,clip=true,trim=0.5cm 0 0 0]{/data/Documents/tex/svn/crystal-gel/nature/fig2.pdf}
+
+	\bigskip
+	Compact structure and higher local $\phi$ \alert{because crystallized}
 \end{frame}

### 2016-06-13 presentation/talk_Roscoff.tex
Crystals outgrow liquids Dynamics: Bergeron > Ostwald

@@
 
+\begin{frame}{Crystals outgrow liquid}
+\vspace{\baselineskip}
+\tikzsetnextfilename{crystal_fates}%
+\begin{tikzpicture}
+\begin{groupplot}[%
+			group style={
+				group name=g, group size=2 by 1,
+				horizontal sep=4em,
+				},
+		width=0.5\linewidth,
+		%height=0.5\linewidth,
+	]
+	\nextgroupplot[
+		title={crystals},
+		xlabel={$t/10^3\tau_B$}, ylabel=\#,
+		xmin=0, xmax=13.0, ymin=0,
+		no marks,
+		legend style={cells={anchor=base west}},
+		]
+	\addplot table[x expr={\thisrowno{0}/2.3e3}]{/data/Documents/tex/svn/crystal-gel/data/fig3a_1.dat} node[left] at (rel axis cs:1,0.2) {size};
+	\addplot table[x expr={\thisrowno{0}/2.3e3}]{/data/Documents/tex/svn/crystal-gel/data/fig3a_2.dat} node[left] at (rel axis cs:1,0.6) {number};
+	%\legend{average size, average number};
+
+	\nextgroupplot[
+		title={cluster size distribution},
+		xlabel={$s$}, ylabel=$R_g/\sigma$,
+		xmode=log, ymode=log,
+		ytick={0.5, 1, 2, 4}, yticklabels={$\frac{1}{2}$, 1, 2, 4},
+		domain=1:5e2,
+		]
+	\addplot[only marks, red, thick, mark=*] file{/data/Documents/tex/svn/crystal-gel/data/fig3d_sqrt_1.dat};
+	\addplot[black, ultra thick] {0.522*x^(1/3)} node[below left] at (axis cs:9e2,2.5) {$D=3$};
+	\addplot[gray, ultra thick] {0.426*x^(1/2.53)} node[pos=0.08,below right] {$D=2.53$};
+\end{groupplot}
+\end{tikzpicture}
+\begin{center}
+\includegraphics[height=6.5\baselineskip]{pathsX.pdf}
+\end{center}
+\end{frame}
+
+\begin{frame}{Dynamics: Bergeron > Ostwald}
+\vspace{\baselineskip}
+\includegraphics[width=\textwidth]{/data/Documents/tex/svn/crystal-gel/nature/fig4.pdf}
+\end{frame}
+
 \appendix

### 2016-06-14 crystal-gel/presentation/talk_Roscoff.tex
Conclusion

@@
 
+\begin{frame}{Exploring phase diagram}
+	\vspace{\baselineskip}
+	\tikzsetnextfilename{phdiag_cluster}%
+	\begin{tikzpicture}[
+		every axis/.style={
+		xlabel absolute, every axis x label/.append style={anchor=base, yshift=-1em},
+		%ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
+		}]
+	\pgfplotscreateplotcyclelist{samples}{%
+	blue,mark=square*,\\%
+	cyan,mark=square\\%
+	red, thick, mark=*\\%
+	red!65!yellow, mark=otimes\\%
+	red!30!yellow, mark=oplus\\%
+	red!65!black, mark=o\\%
+	}%
+	\begin{groupplot}[
+		group style={
+				group name=g, group size=2 by 2,
+				horizontal sep=4em,
+				vertical sep=3em,
+				%x descriptions at=edge bottom,
+			},
+		cycle list name=samples,
+		width=0.51\textwidth,
+		height=0.5\linewidth,
+		axis on top,
+	]
+	\nextgroupplot[
+		title={phase diagram},
+		xlabel={$\phi$}, ylabel={$c_p$ (\si{\milli\gram/\gram})},
+		xmin=0,xmax=0.74,
+		ymin=0, ymax=1.9, ytick={0,0.5,...,2},
+		mark options={solid}
+		]
+		\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
+		\addplot file {/data/Documents/tex/svn/crystal-gel/sample0.phd};
+		\addplot file {/data/Documents/tex/svn/crystal-gel/sample1.phd};
+		\addplot file {/data/Documents/tex/svn/crystal-gel/sample2.phd};
+		\addplot file {/data/Documents/tex/svn/crystal-gel/sample3.phd};
+		\addplot file {/data/Documents/tex/svn/crystal-gel/sample4.phd};
+		\addplot file {/data/Documents/tex/svn/crystal-gel/sample5.phd};
+		\end{scope}
+		\addplot[black, dashed,no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
+		\addplot[black, dashed,no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_sl.phd};
+		\addplot[black, no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_bl.phd};
+		\addplot[black, no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_bg.phd};
+		\addplot[fill=lightgray!50, no marks, draw=none] file {/data/Documents/tex/svn/crystal-gel/fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
+		\addplot[fill=lightgray!50, no marks, draw=none] file {/data/Documents/tex/svn/crystal-gel/fluidcrystal_x.phd} node[above left] (x) at (rel axis cs:1,0.5) {Crystal}  -| (rel axis cs:1,0) \closedcycle;
+		%\draw[gray] (axis cs:0.371290311343, 0.12299370925) -- (axis cs:0.680497991395,0.0372636873433);
+		%\draw[gray, dotted] (axis cs:3.40152108798e-05,0.624739653238) -- (axis cs:0.722831684246, 0.0726528862772);
+		\draw[->] (x.base) --(rel axis cs:0.98,0.01);
+		\addplot[only marks, mark=triangle] file{/data/Documents/tex/svn/crystal-gel/fluid_samples.phd};
+
+	\nextgroupplot[
+		title={cluster size distribution},
+		xlabel={$s$}, ylabel={$R_g/\sigma$},
+		xmin=1, xmax=4e3,
+		ymin=0.5, ymax=30,
+		xmode=log, ymode=log, 
+		clip mode=individual,
+		]
+		\addplot file{/data/Documents/tex/svn/crystal-gel/data/fig1b_sqrt_1.dat};
+		\addplot file{/data/Documents/tex/svn/crystal-gel/data/fig1b_sqrt_2.dat};
+		\addplot file{/data/Documents/tex/svn/crystal-gel/data/fig1b_sqrt_3.dat};
+		\addplot file{/data/Documents/tex/svn/crystal-gel/data/fig1b_sqrt_4.dat};
+		\addplot file{/data/Documents/tex/svn/crystal-gel/data/fig1b_sqrt_5.dat};
+		\addplot[gray, ultra thick, domain=1:4e3] {0.629*x^(1/2.53)} node [left] at (rel axis cs:1,0.9){$D=2.53$};
+
+	\end{groupplot}
+	%\draw(g c1r1.outer south west) -- +(\textwidth,0);
+	\end{tikzpicture}
+	\begin{itemize}
+	\item Good agreement with theoretical phase diagram
+	\item At percolation time, compatible with random gelation
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Outlier at later times}
+	\vspace{\baselineskip}
+	\tikzsetnextfilename{Sq_localphi}%
+	\begin{tikzpicture}[
+		every axis/.style={
+		xlabel absolute, every axis x label/.append style={anchor=base, yshift=-1em},
+		%ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
+		}]
+	\pgfplotscreateplotcyclelist{samples}{%
+	blue,mark=square*,\\%
+	cyan,mark=square\\%
+	red, thick, mark=*\\%
+	red!65!yellow, mark=otimes\\%
+	red!30!yellow, mark=oplus\\%
+	red!65!black, mark=o\\%
+	}%
+	\begin{groupplot}[
+		group style={
+				group name=g, group size=2 by 2,
+				horizontal sep=4em,
+				vertical sep=3em,
+				%x descriptions at=edge bottom,
+			},
+		cycle list name=samples,
+		width=0.51\textwidth,
+		height=0.5\linewidth,
+		axis on top,
+	]
+	\nextgroupplot[
+		title={structure factor},
+		xlabel={$q\sigma$}, ylabel={$S(q)$},
+		xmode=log, ymode=log, xmin=0.2, xmax=20, ymin=0.1,
+		restrict x to domain=-1.8:3, unbounded coords=discard, 
+		%clip mode=individual
+		]
+		\addplot+[clip mode=individual] file{/data/Documents/tex/svn/crystal-gel/data/fig3c_1.dat};
+		\addplot+[clip mode=individual] file{/data/Documents/tex/svn/crystal-gel/data/fig3c_2.dat};
+		\addplot file{/data/Documents/tex/svn/crystal-gel/data/fig3c_3.dat};
+		\addplot+[clip mode=individual] file{/data/Documents/tex/svn/crystal-gel/data/fig3c_4.dat};
+		\addplot+[clip mode=individual] file{/data/Documents/tex/svn/crystal-gel/data/fig3c_5.dat};
+		\addplot[black, ultra thick, domain=0.1:3] {2.14605*x^(-3)} node[pos=0.92,left=0.5em] {$D=3$};
+		\addplot[gray, ultra thick, domain=0.1:3] {3.98797*x^(-2.53)} node [pos=0.65, above right] {$D=2.53$};
+
+	\nextgroupplot[
+		title={local volume fraction},
+		xlabel={\# neighbours}, ylabel={$\phi$ [\%]},
+		xmin=0, ymin=0, ytick={0,10,30,50},
+		extra y ticks={64, 72}, extra y tick style={grid=major}, extra y tick label={},
+		legend pos=north west, legend style={font=\footnotesize},
+		no markers,
+		/pgfplots/error bars/y dir=both, /pgfplots/error bars/y explicit,
+		]
+		\addplot table[y error=err]{/data/Documents/tex/svn/crystal-gel/data/vol_err_phi_0.10_cp_0.82.scaled};
+		\addplot table[y error=err]{/data/Documents/tex/svn/crystal-gel/data/vol_err_phi_0.10_cp_1.36.scaled};
+		\addplot table[y error=err]{/data/Documents/tex/svn/crystal-gel/data/vol_err_phi_0.25_cp_0.38.scaled};
+		\addplot table[y error=err]{/data/Documents/tex/svn/crystal-gel/data/vol_err_phi_0.25_cp_0.48.scaled};
+		\addplot table[y error=err]{/data/Documents/tex/svn/crystal-gel/data/vol_err_phi_0.25_cp_0.57.scaled};
+
+	\end{groupplot}
+	%\draw(g c1r1.outer south west) -- +(\textwidth,0);
+	\end{tikzpicture}
+	\begin{itemize}
+	\item Final stucture compatible with random gelation
+	\item Glassy local $\phi$ for 12-neighboured particles
+	\end{itemize}
+	\alert{Except} \textcolor{red}{$\bullet$}: compact structure and higher local $\phi$. %\textcolor{ilmorange}{$\Rightarrow$} Crystal?
+\end{frame}
+
+\begin{frame}{Real space: crystal gel!}
+	\tikzsetnextfilename{phdiag_arrows}%
+	\begin{tikzpicture}[
+			every axis/.style={
+			xlabel absolute, every axis x label/.append style={anchor=base, yshift=-1em},
+			%ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
+			}]
+		\pgfplotscreateplotcyclelist{samples}{%
+		blue,mark=square*,\\%
+		cyan,mark=square\\%
+		red, thick, mark=*\\%
+		red!65!yellow, mark=otimes\\%
+		red!30!yellow, mark=oplus\\%
+		red!65!black, mark=o\\%
+		}%
+		\begin{axis}[
+			cycle list name=samples,
+			width=0.51\textwidth,
+			height=0.5\linewidth,
+			axis on top,
+			xlabel={$\phi$}, ylabel={$c_p$ (\si{\milli\gram/\gram})},
+			xmin=0,xmax=0.74,
+			ymin=0, ymax=1.9, ytick={0,0.5,...,2},
+			mark options={solid}
+			]
+			\begin{scope}[/pgfplots/skip coords between index={0}{1}, /pgfplots/skip coords between index={2}{3}]%[draw=none,mark repeat=2,mark phase=2]
+			\addplot table {/data/Documents/tex/svn/crystal-gel/sample0.phd};
+			\addplot table {/data/Documents/tex/svn/crystal-gel/sample1.phd};
+			\addplot table {/data/Documents/tex/svn/crystal-gel/sample2.phd}[->] -- (rel axis cs:0.95,0.2);
+			\addplot table {/data/Documents/tex/svn/crystal-gel/sample3.phd};
+			\addplot table {/data/Documents/tex/svn/crystal-gel/sample4.phd};
+			\addplot table {/data/Documents/tex/svn/crystal-gel/sample5.phd}[->] -- (rel axis cs:0.95,0.7);
+			\end{scope}
+			\addplot[black, dashed,no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
+			\addplot[black, dashed,no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_sl.phd};
+			\addplot[black, no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_bl.phd};
+			\addplot[black, no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_bg.phd};
+			\addplot[fill=lightgray!50, no marks, draw=none] file {/data/Documents/tex/svn/crystal-gel/fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {/data/Documents/tex/svn/crystal-gel/fluidcrystal_x.phd} %node[above left] (x) at (rel axis cs:1,0.5) {Crystal}  
+			-| (rel axis cs:1,0) \closedcycle;
+			%\draw[->] (x.base) --(rel axis cs:0.98,0.01);
+			\addplot[only marks, mark=triangle] file{/data/Documents/tex/svn/crystal-gel/fluid_samples.phd};
+	\end{axis}
+	\end{tikzpicture}
+	\includegraphics[height=11\baselineskip,clip=true,trim=0.5cm 0 0 0]{/data/Documents/tex/svn/crystal-gel/nature/fig2.pdf}
+
+	\bigskip
+	Compact structure and higher local $\phi$ \alert{because crystallized}
+\end{frame}
+
+\begin{frame}{Crystals outgrow liquid}
+\vspace{\baselineskip}
+\tikzsetnextfilename{crystal_fates}%
+\begin{tikzpicture}
+\begin{groupplot}[%
+			group style={
+				group name=g, group size=2 by 1,
+				horizontal sep=4em,
+				},
+		width=0.5\linewidth,
+		%height=0.5\linewidth,
+	]
+	\nextgroupplot[
+		title={nuclei},
+		xlabel={$t/10^3\tau_B$}, ylabel=\#,
+		xmin=0, xmax=13.0, ymin=0,
+		no marks,
+		legend style={cells={anchor=base west}},
+		]
+	\addplot table[x expr={\thisrowno{0}/2.3e3}]{/data/Documents/tex/svn/crystal-gel/data/fig3a_1.dat} node[left] at (rel axis cs:1,0.2) {avg size};
+	\addplot table[x expr={\thisrowno{0}/2.3e3}]{/data/Documents/tex/svn/crystal-gel/data/fig3a_2.dat} node[left] at (rel axis cs:1,0.6) {number};
+	%\legend{average size, average number};
+
+	\nextgroupplot[
+		title={cluster size distribution},
+		xlabel={$s$}, ylabel=$R_g/\sigma$,
+		xmode=log, ymode=log,
+		ytick={0.5, 1, 2, 4}, yticklabels={$\frac{1}{2}$, 1, 2, 4},
+		domain=1:5e2,
+		]
+	\addplot[only marks, red, thick, mark=*] file{/data/Documents/tex/svn/crystal-gel/data/fig3d_sqrt_1.dat};
+	\addplot[black, ultra thick] {0.522*x^(1/3)} node[below left] at (axis cs:9e2,2.5) {$D=3$};
+	\addplot[gray, ultra thick] {0.426*x^(1/2.53)} node[pos=0.08,below right] {$D=2.53$};
+\end{groupplot}
+\end{tikzpicture}
+\begin{center}
+\tikzsetnextfilename{pathsX}\vspace{-0.5em}\input{pathsX.pgf}
+\end{center}
+\end{frame}
+
+\begin{frame}{Dynamics: Bergeron > Ostwald}
+\vspace{\baselineskip}
+\includegraphics[width=\textwidth]{/data/Documents/tex/svn/crystal-gel/nature/fig4.pdf}
+\end{frame}
+
+\begin{frame}{Conclusion}
+	\begin{itemize}
+		\item A particle level colloidal analogue to the Bergeron process
+		\item Generic "crystal-gel" process
+		\begin{enumerate}
+			\item gas-liquid phase separation below melting point
+		of a crystal
+			\item weak or little frustration against crystallization
+			\item dynamical slowing down in a supercooled liquid state
+			\begin{itemize}
+				\item viscoelastic phase separation
+				\item network of the minority liquid phase
+			\end{itemize}
+			\item low supercooling: crystallisation > vitrification
+		\end{enumerate}
+		\item Gels of poor glass formers (monoatomic systems)
+		\item Porous material with a large area of crystalline planes
+		\begin{itemize}
+			\item catalysis, sensing, filtration, super-capacitor\ldots
+			\item single step
+		\end{itemize}
+	\end{itemize}
+\end{frame}
+
 \appendix

### 2016-06-14 crystal-gel/talk_Roscoff.tex
Change tree structure

--- 2016-06-14 crystal-gel/presentation/talk_Roscoff.tex
@@
 	\begin{tabu}{X[c]X[c]X[c]}
-		\includegraphics[height=3\baselineskip]{Tsurusawa}&
-		\includegraphics[height=3\baselineskip]{John}&
-		\includegraphics[height=3\baselineskip]{Tanaka}\\
+		\includegraphics[height=3\baselineskip]{presentation/Tsurusawa}&
+		\includegraphics[height=3\baselineskip]{presentation/John}&
+		\includegraphics[height=3\baselineskip]{presentation/Tanaka}\\
 		Hideyo Tsurusawa & John Russo & Hajime Tanaka\\
@@
 	\tikzsetnextfilename{whyx}%
-	\input{fig_presentation/whyx.pgf}
+	\input{presentation/whyx.pgf}
 
@@
 	\column{0.4\textwidth}
-	\includegraphics[width=\textwidth,clip=true,trim=0 0 0 1cm]{Pitter1.jpg}
+	\includegraphics[width=\textwidth,clip=true,trim=0 0 0 1cm]{presentation/Pitter1.jpg}
 
-	\vspace{1em}\includegraphics[width=\textwidth]{bergeron.jpg}
+	\vspace{1em}\includegraphics[width=\textwidth]{presentation/bergeron.jpg}
 	\column{0.40\textwidth}
-	\includegraphics[width=\textwidth,clip=true,trim=0 0 0.8mm 0]{queries-figure-1.jpg}
+	\includegraphics[width=\textwidth,clip=true,trim=0 0 0.8mm 0]{presentation/queries-figure-1.jpg}
 
-	\vspace{1em}\includegraphics[width=\textwidth]{OKC-Fallstreak.jpg}
+	\vspace{1em}\includegraphics[width=\textwidth]{presentation/OKC-Fallstreak.jpg}
 	\end{columns}
@@
 	\bigskip
-	\includegraphics[width=\textwidth]{../../CNRS-2015/Renth_sedim.jpg}\\
+	\includegraphics[width=\textwidth]{../CNRS-2015/Renth_sedim.jpg}\\
 	\textit{\footnotesize Poon PRL 1999, Renth PRE 2001}
@@
 	\column{0.45\textwidth}
-	\includegraphics[width=\textwidth,clip=true,trim=0 0 0 4.5cm]{Renth_diag.pdf}
+	\includegraphics[width=\textwidth,clip=true,trim=0 0 0 4.5cm]{presentation/Renth_diag.pdf}
 	\end{columns}
@@
 	\end{itemize}
-	\includegraphics[width=\textwidth]{micrograv_Sabin2012}
+	\includegraphics[width=\textwidth]{presentation/micrograv_Sabin2012}
 \end{frame}
@@
 		\tikzsetnextfilename{chemin_gelification}%
-		\input{chemin_gelification.pgf}
+		\input{presentation/chemin_gelification.pgf}
 	\column{0.335\textwidth}
 		\tikzsetnextfilename{cell}%
-		\input{cell.pgf}
+		\input{presentation/cell.pgf}
 \end{columns}
@@
 		\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
-		\addplot file {/data/Documents/tex/svn/crystal-gel/sample0.phd};
-		\addplot file {/data/Documents/tex/svn/crystal-gel/sample1.phd};
-		\addplot file {/data/Documents/tex/svn/crystal-gel/sample2.phd};
-		\addplot file {/data/Documents/tex/svn/crystal-gel/sample3.phd};
-		\addplot file {/data/Documents/tex/svn/crystal-gel/sample4.phd};
-		\addplot file {/data/Documents/tex/svn/crystal-gel/sample5.phd};
+		\addplot file {sample0.phd};
+		\addplot file {sample1.phd};
+		\addplot file {sample2.phd};
+		\addplot file {sample3.phd};
+		\addplot file {sample4.phd};
+		\addplot file {sample5.phd};
 		\end{scope}
-		\addplot[black, dashed,no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
-		\addplot[black, dashed,no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_sl.phd};
-		\addplot[black, no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_bl.phd};
-		\addplot[black, no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_bg.phd};
-		\addplot[fill=lightgray!50, no marks, draw=none] file {/data/Documents/tex/svn/crystal-gel/fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
-		\addplot[fill=lightgray!50, no marks, draw=none] file {/data/Documents/tex/svn/crystal-gel/fluidcrystal_x.phd} node[above left] (x) at (rel axis cs:1,0.5) {Crystal}  -| (rel axis cs:1,0) \closedcycle;
+		\addplot[black, dashed,no marks] file {gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
+		\addplot[black, dashed,no marks] file {gasliquid_sl.phd};
+		\addplot[black, no marks] file {gasliquid_bl.phd};
+		\addplot[black, no marks] file {gasliquid_bg.phd};
+		\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
+		\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd} node[above left] (x) at (rel axis cs:1,0.5) {Crystal}  -| (rel axis cs:1,0) \closedcycle;
 		%\draw[gray] (axis cs:0.371290311343, 0.12299370925) -- (axis cs:0.680497991395,0.0372636873433);
@@
 		\draw[->] (x.base) --(rel axis cs:0.98,0.01);
-		\addplot[only marks, mark=triangle] file{/data/Documents/tex/svn/crystal-gel/fluid_samples.phd};
+		\addplot[only marks, mark=triangle] file{fluid_samples.phd};
 
@@
 		]
-		\addplot file{/data/Documents/tex/svn/crystal-gel/data/fig1b_sqrt_1.dat};
-		\addplot file{/data/Documents/tex/svn/crystal-gel/data/fig1b_sqrt_2.dat};
-		\addplot file{/data/Documents/tex/svn/crystal-gel/data/fig1b_sqrt_3.dat};
-		\addplot file{/data/Documents/tex/svn/crystal-gel/data/fig1b_sqrt_4.dat};
-		\addplot file{/data/Documents/tex/svn/crystal-gel/data/fig1b_sqrt_5.dat};
+		\addplot file{data/fig1b_sqrt_1.dat};
+		\addplot file{data/fig1b_sqrt_2.dat};
+		\addplot file{data/fig1b_sqrt_3.dat};
+		\addplot file{data/fig1b_sqrt_4.dat};
+		\addplot file{data/fig1b_sqrt_5.dat};
 		\addplot[gray, ultra thick, domain=1:4e3] {0.629*x^(1/2.53)} node [left] at (rel axis cs:1,0.9){$D=2.53$};
@@
 		]
-		\addplot+[clip mode=individual] file{/data/Documents/tex/svn/crystal-gel/data/fig3c_1.dat};
-		\addplot+[clip mode=individual] file{/data/Documents/tex/svn/crystal-gel/data/fig3c_2.dat};
-		\addplot file{/data/Documents/tex/svn/crystal-gel/data/fig3c_3.dat};
-		\addplot+[clip mode=individual] file{/data/Documents/tex/svn/crystal-gel/data/fig3c_4.dat};
-		\addplot+[clip mode=individual] file{/data/Documents/tex/svn/crystal-gel/data/fig3c_5.dat};
+		\addplot+[clip mode=individual] file{data/fig3c_1.dat};
+		\addplot+[clip mode=individual] file{data/fig3c_2.dat};
+		\addplot file{data/fig3c_3.dat};
+		\addplot+[clip mode=individual] file{data/fig3c_4.dat};
+		\addplot+[clip mode=individual] file{data/fig3c_5.dat};
 		\addplot[black, ultra thick, domain=0.1:3] {2.14605*x^(-3)} node[pos=0.92,left=0.5em] {$D=3$};
@@
 		]
-		\addplot table[y error=err]{/data/Documents/tex/svn/crystal-gel/data/vol_err_phi_0.10_cp_0.82.scaled};
-		\addplot table[y error=err]{/data/Documents/tex/svn/crystal-gel/data/vol_err_phi_0.10_cp_1.36.scaled};
-		\addplot table[y error=err]{/data/Documents/tex/svn/crystal-gel/data/vol_err_phi_0.25_cp_0.38.scaled};
-		\addplot table[y error=err]{/data/Documents/tex/svn/crystal-gel/data/vol_err_phi_0.25_cp_0.48.scaled};
-		\addplot table[y error=err]{/data/Documents/tex/svn/crystal-gel/data/vol_err_phi_0.25_cp_0.57.scaled};
+		\addplot table[y error=err]{data/vol_err_phi_0.10_cp_0.82.scaled};
+		\addplot table[y error=err]{data/vol_err_phi_0.10_cp_1.36.scaled};
+		\addplot table[y error=err]{data/vol_err_phi_0.25_cp_0.38.scaled};
+		\addplot table[y error=err]{data/vol_err_phi_0.25_cp_0.48.scaled};
+		\addplot table[y error=err]{data/vol_err_phi_0.25_cp_0.57.scaled};
 
@@
 			\begin{scope}[/pgfplots/skip coords between index={0}{1}, /pgfplots/skip coords between index={2}{3}]%[draw=none,mark repeat=2,mark phase=2]
-			\addplot table {/data/Documents/tex/svn/crystal-gel/sample0.phd};
-			\addplot table {/data/Documents/tex/svn/crystal-gel/sample1.phd};
-			\addplot table {/data/Documents/tex/svn/crystal-gel/sample2.phd}[->] -- (rel axis cs:0.95,0.2);
-			\addplot table {/data/Documents/tex/svn/crystal-gel/sample3.phd};
-			\addplot table {/data/Documents/tex/svn/crystal-gel/sample4.phd};
-			\addplot table {/data/Documents/tex/svn/crystal-gel/sample5.phd}[->] -- (rel axis cs:0.95,0.7);
+			\addplot table {sample0.phd};
+			\addplot table {sample1.phd};
+			\addplot table {sample2.phd}[->] -- (rel axis cs:0.95,0.2);
+			\addplot table {sample3.phd};
+			\addplot table {sample4.phd};
+			\addplot table {sample5.phd}[->] -- (rel axis cs:0.95,0.7);
 			\end{scope}
-			\addplot[black, dashed,no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
-			\addplot[black, dashed,no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_sl.phd};
-			\addplot[black, no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_bl.phd};
-			\addplot[black, no marks] file {/data/Documents/tex/svn/crystal-gel/gasliquid_bg.phd};
-			\addplot[fill=lightgray!50, no marks, draw=none] file {/data/Documents/tex/svn/crystal-gel/fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
-			\addplot[fill=lightgray!50, no marks, draw=none] file {/data/Documents/tex/svn/crystal-gel/fluidcrystal_x.phd} %node[above left] (x) at (rel axis cs:1,0.5) {Crystal}  
+			\addplot[black, dashed,no marks] file {gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
+			\addplot[black, dashed,no marks] file {gasliquid_sl.phd};
+			\addplot[black, no marks] file {gasliquid_bl.phd};
+			\addplot[black, no marks] file {gasliquid_bg.phd};
+			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd} %node[above left] (x) at (rel axis cs:1,0.5) {Crystal}  
 			-| (rel axis cs:1,0) \closedcycle;
 			%\draw[->] (x.base) --(rel axis cs:0.98,0.01);
-			\addplot[only marks, mark=triangle] file{/data/Documents/tex/svn/crystal-gel/fluid_samples.phd};
+			\addplot[only marks, mark=triangle] file{fluid_samples.phd};
 	\end{axis}
 	\end{tikzpicture}
-	\includegraphics[height=11\baselineskip,clip=true,trim=0.5cm 0 0 0]{/data/Documents/tex/svn/crystal-gel/nature/fig2.pdf}
+	\includegraphics[height=11\baselineskip,clip=true,trim=0.5cm 0 0 0]{nature/fig2.pdf}
 
@@
 		]
-	\addplot table[x expr={\thisrowno{0}/2.3e3}]{/data/Documents/tex/svn/crystal-gel/data/fig3a_1.dat} node[left] at (rel axis cs:1,0.2) {avg size};
-	\addplot table[x expr={\thisrowno{0}/2.3e3}]{/data/Documents/tex/svn/crystal-gel/data/fig3a_2.dat} node[left] at (rel axis cs:1,0.6) {number};
+	\addplot table[x expr={\thisrowno{0}/2.3e3}]{data/fig3a_1.dat} node[left] at (rel axis cs:1,0.2) {avg size};
+	\addplot table[x expr={\thisrowno{0}/2.3e3}]{data/fig3a_2.dat} node[left] at (rel axis cs:1,0.6) {number};
 	%\legend{average size, average number};
@@
 		]
-	\addplot[only marks, red, thick, mark=*] file{/data/Documents/tex/svn/crystal-gel/data/fig3d_sqrt_1.dat};
+	\addplot[only marks, red, thick, mark=*] file{data/fig3d_sqrt_1.dat};
 	\addplot[black, ultra thick] {0.522*x^(1/3)} node[below left] at (axis cs:9e2,2.5) {$D=3$};
@@
 \begin{center}
-\tikzsetnextfilename{pathsX}\vspace{-0.5em}\input{pathsX.pgf}
+\tikzsetnextfilename{pathsX}\vspace{-0.5em}\input{presentation/pathsX.pgf}
 \end{center}
@@
 \vspace{\baselineskip}
-\includegraphics[width=\textwidth]{/data/Documents/tex/svn/crystal-gel/nature/fig4.pdf}
+\includegraphics[width=\textwidth]{nature/fig4.pdf}
 \end{frame}

### 2016-06-17 talk_Roscoff.tex
starts to explain VPS

--- 2016-06-14 crystal-gel/talk_Roscoff.tex
@@
 	\includegraphics[width=\textwidth,clip=true,trim=0 0 0 4.5cm]{presentation/Renth_diag.pdf}
+	\end{columns}
+\end{frame}
+
+\begin{frame}{Gel formation}
+	\begin{columns}
+		\column{0.5\textwidth}
+		\begin{tikzpicture}[
+			every axis/.style={
+			xlabel absolute, every axis x label/.append style={anchor=base, yshift=-0.5em},
+			ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
+			}]
+		\pgfplotscreateplotcyclelist{samples}{%
+		blue,mark=square*,\\%
+		cyan,mark=square\\%
+		red, thick, mark=*\\%
+		red!65!yellow, mark=otimes\\%
+		red!30!yellow, mark=oplus\\%
+		red!65!black, mark=o\\%
+		}%
+		\begin{groupplot}[
+			group style={
+					group name=g, group size=1 by 2,
+					horizontal sep=4em,
+					vertical sep=0.5em,
+					x descriptions at=edge bottom,
+				},
+			cycle list name=samples,
+			width=\textwidth,
+			%height=0.5\linewidth,
+			axis on top,
+		]
+		\nextgroupplot[
+			ylabel={$\Pi v$},
+			xmin=0,xmax=0.74,
+			ymin=0, ymax=2400, %ytick={0,0.5,...,2},
+			mark options={solid}
+			]
+			\addplot[black, dashed,no marks] file {res_gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
+			\addplot[black, dashed,no marks] file {res_gasliquid_sl.phd};
+			\addplot[black, no marks] file {res_gasliquid_bl.phd};
+			\addplot[black, no marks] file {res_gasliquid_bg.phd};
+			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_x.phd} node[above left, inner ysep=0, inner xsep=1pt] (x) at (rel axis cs:1,0) {Crystal}  -| (rel axis cs:1,0) \closedcycle;
+			\node at (axis cs:0.4,1500) {Gel};
+
+		\nextgroupplot[
+			xlabel={$\phi$}, ylabel={$c_p$ (\si{\milli\gram/\gram})},
+			xmin=0,xmax=0.74,
+			ymin=0, ymax=1.9, ytick={0,0.5,...,2},
+			mark options={solid}
+			]
+			\addplot[black, dashed,no marks] file {gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
+			\addplot[black, dashed,no marks] file {gasliquid_sl.phd};
+			\addplot[black, no marks] file {gasliquid_bl.phd};
+			\addplot[black, no marks] file {gasliquid_bg.phd};
+			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd} node[above left] (x) at (rel axis cs:1,0.5) {Crystal}  -| (rel axis cs:1,0) \closedcycle;
+			\draw[->] (x.base) --(rel axis cs:0.98,0.01);
+		\end{groupplot}
+		%\draw(g c1r1.outer south west) -- +(\textwidth,0);
+		\end{tikzpicture}
+		\column{0.5\textwidth}
+		Small polymer/colloid size ratio
+		\begin{itemize}
+		\item metastable gas-liquid
+		\item spinodal + glass $\Rightarrow$ gel\\ \hfill\textit{\footnotesize Lu et al. Nature~2008}
+		\item continuous minority phase
+		\item viscoelastic phase sep.\\ \hfill\textit{\footnotesize Tanaka J.Phys.Cond.Mat.~2000}
+		\end{itemize}
+		\begin{tikzpicture}
+	\draw[->] (0,0) -- +(\textwidth,0) node[pos=0.1, label=-90:\textcolor{ilmorange}{$\tau_1$}](t1) {|} node[pos=0.5, label=-90:\textcolor{ilmcolor}{$\tau_2$}] (t2){|};
+	\fill[ilmcolor] ($(t1)!0.1!45:(t2)$) coordinate(cse) rectangle ($(t1)!0.9!45:(t2)$) coordinate(cnw);
+	\clip (cse) rectangle (cnw);
+	\fill[ilmorange] ++($(cse)!0.5!(cnw)$) 
+		+(60:0.06\textwidth) circle (0.045\textwidth) 
+		+(180:0.06\textwidth) circle (0.045\textwidth) 
+		+(-60:0.06\textwidth) circle (0.045\textwidth) 
+		+(0:0.12\textwidth) circle (0.045\textwidth)
+		+(120:0.12\textwidth) circle (0.045\textwidth)
+		+(-120:0.12\textwidth) circle (0.045\textwidth)
+		+(45:0.16\textwidth) circle (0.045\textwidth)
+		+(-45:0.16\textwidth) circle (0.045\textwidth);
+	\end{tikzpicture}
 	\end{columns}

### 2016-06-17 crystal-gel/talk_Roscoff.tex
finish VPS description

@@
 	\includegraphics[width=\textwidth,clip=true,trim=0 0 0 4.5cm]{presentation/Renth_diag.pdf}
+	\end{columns}
+\end{frame}
+
+\begin{frame}{Gel formation}
+	\begin{columns}
+		\column{0.5\textwidth}
+		\vspace{\baselineskip}
+		\tikzsetnextfilename{phdiagth}%
+		\begin{tikzpicture}[
+			every axis/.style={
+			xlabel absolute, every axis x label/.append style={anchor=base, yshift=-0.5em},
+			ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
+			}]
+		\pgfplotscreateplotcyclelist{samples}{%
+		blue,mark=square*,\\%
+		cyan,mark=square\\%
+		red, thick, mark=*\\%
+		red!65!yellow, mark=otimes\\%
+		red!30!yellow, mark=oplus\\%
+		red!65!black, mark=o\\%
+		}%
+		\begin{groupplot}[
+			group style={
+					group name=g, group size=1 by 2,
+					horizontal sep=4em,
+					vertical sep=0.5em,
+					x descriptions at=edge bottom,
+				},
+			cycle list name=samples,
+			width=\textwidth,
+			%height=0.5\linewidth,
+			axis on top,
+		]
+		\nextgroupplot[
+			ylabel={$\Pi v$},
+			xmin=0,xmax=0.74,
+			ymin=0, ymax=2400, %ytick={0,0.5,...,2},
+			mark options={solid}
+			]
+			\addplot[black, dashed,no marks] file {res_gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
+			\addplot[black, dashed,no marks] file {res_gasliquid_sl.phd};
+			\addplot[black, no marks] file {res_gasliquid_bl.phd};
+			\addplot[black, no marks] file {res_gasliquid_bg.phd};
+			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_x.phd} node[above left, inner ysep=0, inner xsep=1pt] (x) at (rel axis cs:1,0) {Crystal}  -| (rel axis cs:1,0) \closedcycle;
+			\node at (axis cs:0.4,1500) {Gel};
+
+		\nextgroupplot[
+			xlabel={$\phi$}, ylabel={$c_p$ (\si{\milli\gram/\gram})},
+			xmin=0,xmax=0.74,
+			ymin=0, ymax=1.9, ytick={0,0.5,...,2},
+			mark options={solid}
+			]
+			\addplot[black, dashed,no marks] file {gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
+			\addplot[black, dashed,no marks] file {gasliquid_sl.phd};
+			\addplot[black, no marks] file {gasliquid_bl.phd};
+			\addplot[black, no marks] file {gasliquid_bg.phd};
+			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd} node[above left] (x) at (rel axis cs:1,0.5) {Crystal}  -| (rel axis cs:1,0) \closedcycle;
+			\draw[->] (x.base) --(rel axis cs:0.98,0.01);
+		\end{groupplot}
+		%\draw(g c1r1.outer south west) -- +(\textwidth,0);
+		\end{tikzpicture}
+		\column{0.5\textwidth}
+		Small polymer/colloid size ratio
+		\begin{itemize}
+		\item metastable gas-liquid
+		\item spinodal + glass $\Rightarrow$ gel\\ \hfill\textit{\footnotesize Lu et al. Nature~2008}
+		\item continuous minority phase
+		\item viscoelastic phase sep.\\ \hfill\textit{\footnotesize Tanaka J.Phys.Cond.Mat.~2000}
+		\item due to dynamic assymetry
+		\end{itemize}
+		\tikzsetnextfilename{vps}%
+		\begin{tikzpicture}
+	\draw[->] (0,0) -- +(\textwidth,0) node[pos=0.05, label=-90:\textcolor{ilmorange}{$\tau_1$}](t1) {|} node[pos=0.54, label=-90:\textcolor{ilmcolor}{$\tau_2$}] (t2){|} coordinate[pos=1] (tf) node[anchor=north east,font=\footnotesize]{time};
+	\let\side\relax
+	\newlength\side
+	\setlength{\side}{0.275\textwidth}
+	\fill[ilmcolor!80!ilmorange] ($(t1)!0.1!45:(t2)$) coordinate(cse) rectangle +(\side,\side) coordinate(cnw);
+	\begin{scope}
+	\clip (cse) rectangle (cnw);
+	\fill[ilmorange!80!ilmcolor] ++($(cse)!0.5!(cnw)$) 
+		+(60:0.3\side) circle (0.22\side) 
+		+(180:0.3\side) circle (0.22\side) 
+		+(-60:0.3\side) circle (0.22\side) 
+		+(0:0.6\side) circle (0.22\side)
+		+(120:0.6\side) circle (0.22\side)
+		+(-120:0.6\side) circle (0.22\side)
+		+(45:0.8\side) circle (0.22\side)
+		+(-45:0.8\side) circle (0.22\side);
+	\end{scope}
+	\fill[ilmcolor] (cse-|t2) ++(-0.5\side,0) coordinate(cse) rectangle +(\side,\side) coordinate(cnw);
+	\begin{scope}
+	\clip (cse) rectangle (cnw);
+	\fill[ilmorange] ++($(cse)!0.5!(cnw)$) 
+		+(60:0.3\side) circle (0.26\side) 
+		+(180:0.3\side) circle (0.26\side) 
+		+(-60:0.3\side) circle (0.25\side) 
+		+(0:0.6\side) circle (0.25\side)
+		+(120:0.6\side) circle (0.25\side)
+		+(-120:0.6\side) circle (0.25\side)
+		+(45:0.8\side) circle (0.25\side)
+		+(-45:0.8\side) circle (0.25\side)
+		%enhance broken bond
+		+(120:0.15\side) circle (0.07\side);
+	\end{scope}
+	\fill[ilmorange] (cse-|tf) ++(-\side,0) coordinate(cse) rectangle +(\side,\side) coordinate(cnw);
+	\begin{scope}
+	\clip (cse) rectangle (cnw);
+	\fill[ilmcolor] ++($(cse)!0.5!(cnw)$) circle (0.07\side)
+		+(0:0.3\side) circle (0.07\side) 
+		+(120:0.3\side) circle (0.07\side) 
+		+(-120:0.3\side) circle (0.07\side) 
+		+(-60:0.6\side) circle (0.07\side)
+		+(60:0.6\side) circle (0.07\side)
+		++(0:0.3\side)
+		+(-60:0.3\side) circle (0.07\side)
+		+(60:0.3\side) circle (0.07\side)
+		+(120:0.6\side) circle (0.07\side)
+		+(-120:0.6\side) circle (0.07\side)
+		++(180:0.9\side)
+		+(-60:0.3\side) circle (0.07\side)
+		+(60:0.3\side) circle (0.07\side);
+	\end{scope}
+	\end{tikzpicture}
+
+	\hfill$\textcolor{ilmcolor}{\tau_2} \rightarrow +\infty$ {\usebeamercolor[fg]{itemize item}\usebeamertemplate{itemize item}} arrest 
 	\end{columns}

### 2016-06-20 crystal-gel/talk_Roscoff.tex
backslides

@@
 
-\usepackage{siunitx}
+\usepackage[separate-uncertainty = true]{siunitx}
 
@@
 
+\begin{frame}{Reproducibility in capillary}
+\hfill $\phi\approx 0.33$, $c_P=\SI{0.40}{\milli\gram/\gram}$
+
+\includegraphics[width=\textwidth]{nature/capillary}
+\end{frame}
+
+\begin{frame}{Transition probabilities}
+\vspace{\baselineskip}
+\includegraphics[width=\textwidth]{nature/sfig1.pdf}
+\end{frame}
+
+\begin{frame}{Evolution local volume fraction}
+\vspace{\baselineskip}
+12-neighboured particles
+\begin{center}
+\includegraphics{nature/local_phi12.pdf}
+\end{center}
+\end{frame}
+
+\begin{frame}{phase definition}
+\begin{description}
+\item[crystal]
+	\begin{itemize} 
+	\item particle $i$ has $N_b(i)=12$ neighbours
+	\item $\displaystyle q_{6 m}(i)=\frac{1}{N_b(i)}\sum_{j=1}^{N_b(i)}
+Y_{6 m}(\mathbf{r_{ij}}),\quad -6 <m<6$
+	\item crystalline bond $(i,j)$: \[(\mathbf{q}_{6}(i)/|\mathbf{q}_{6}(i)|)\cdot(\mathbf{q}_{6}(j)/|\mathbf{q}_{6}(j)|)>0.7\]
+	\item crystalline particle: $7$ crystalline bonds
+	\end{itemize}
+\item[surface] two neighbouring crystal particles
+\item[liquid] at least four neighbouring particles
+\item[gas]
+\end{description}
+\end{frame}
+
+\begin{frame}{Colloid and polymer sizes}
+Direct measurement \alert{too imprecise}
+\begin{itemize}
+\item diffusion constant $\Rightarrow R= \SI{2.3 +- 0.05}{\micro\metre}$
+\item Flory scaling litterature measurements $\Rightarrow R_g=\SI{76. +- 50.}{\nano\metre}$
+\end{itemize}
+
+\begin{block}{Adjusting theoretical and experimental phase diagrams}
+\begin{enumerate}
+\item to have the final local volume fraction of the 12-neighboured particles in the crystallising sample to be equal to the theoretical equilibrium crystal volume fraction 
+\item to have the theoretical spinodal line to correspond to the gel region boundary.
+\end{enumerate}
+\end{block}
+Starting from our initial estimates of $\sigma$ and $R$ we iteratively converge to $R=\SI{80}{\nano\metre}$, $\sigma = \SI{2.25}{\micro\metre}$ and a crystal (without defects) at $0.723$.
+\end{frame}
+
 \setcounter{framenumber}{\value{finalframe}}

### 2016-06-20 talk_Roscoff.tex
backslides

@@
 		\column{0.5\textwidth}
+		\vspace{\baselineskip}
+		\tikzsetnextfilename{phdiagth}%
 		\begin{tikzpicture}[
@@
 		\item viscoelastic phase sep.\\ \hfill\textit{\footnotesize Tanaka J.Phys.Cond.Mat.~2000}
+		\item due to dynamic assymetry
 		\end{itemize}
+		\tikzsetnextfilename{vps}%
 		\begin{tikzpicture}
-	\draw[->] (0,0) -- +(\textwidth,0) node[pos=0.1, label=-90:\textcolor{ilmorange}{$\tau_1$}](t1) {|} node[pos=0.5, label=-90:\textcolor{ilmcolor}{$\tau_2$}] (t2){|};
-	\fill[ilmcolor] ($(t1)!0.1!45:(t2)$) coordinate(cse) rectangle ($(t1)!0.9!45:(t2)$) coordinate(cnw);
+	\draw[->] (0,0) -- +(\textwidth,0) node[pos=0.05, label=-90:\textcolor{ilmorange}{$\tau_1$}](t1) {|} node[pos=0.54, label=-90:\textcolor{ilmcolor}{$\tau_2$}] (t2){|} coordinate[pos=1] (tf) node[anchor=north east,font=\footnotesize]{time};
+	\let\side\relax
+	\newlength\side
+	\setlength{\side}{0.275\textwidth}
+	\fill[ilmcolor!80!ilmorange] ($(t1)!0.1!45:(t2)$) coordinate(cse) rectangle +(\side,\side) coordinate(cnw);
+	\begin{scope}
+	\clip (cse) rectangle (cnw);
+	\fill[ilmorange!80!ilmcolor] ++($(cse)!0.5!(cnw)$) 
+		+(60:0.3\side) circle (0.22\side) 
+		+(180:0.3\side) circle (0.22\side) 
+		+(-60:0.3\side) circle (0.22\side) 
+		+(0:0.6\side) circle (0.22\side)
+		+(120:0.6\side) circle (0.22\side)
+		+(-120:0.6\side) circle (0.22\side)
+		+(45:0.8\side) circle (0.22\side)
+		+(-45:0.8\side) circle (0.22\side);
+	\end{scope}
+	\fill[ilmcolor] (cse-|t2) ++(-0.5\side,0) coordinate(cse) rectangle +(\side,\side) coordinate(cnw);
+	\begin{scope}
 	\clip (cse) rectangle (cnw);
 	\fill[ilmorange] ++($(cse)!0.5!(cnw)$) 
-		+(60:0.06\textwidth) circle (0.045\textwidth) 
-		+(180:0.06\textwidth) circle (0.045\textwidth) 
-		+(-60:0.06\textwidth) circle (0.045\textwidth) 
-		+(0:0.12\textwidth) circle (0.045\textwidth)
-		+(120:0.12\textwidth) circle (0.045\textwidth)
-		+(-120:0.12\textwidth) circle (0.045\textwidth)
-		+(45:0.16\textwidth) circle (0.045\textwidth)
-		+(-45:0.16\textwidth) circle (0.045\textwidth);
+		+(60:0.3\side) circle (0.26\side) 
+		+(180:0.3\side) circle (0.26\side) 
+		+(-60:0.3\side) circle (0.25\side) 
+		+(0:0.6\side) circle (0.25\side)
+		+(120:0.6\side) circle (0.25\side)
+		+(-120:0.6\side) circle (0.25\side)
+		+(45:0.8\side) circle (0.25\side)
+		+(-45:0.8\side) circle (0.25\side)
+		%enhance broken bond
+		+(120:0.15\side) circle (0.07\side);
+	\end{scope}
+	\fill[ilmorange] (cse-|tf) ++(-\side,0) coordinate(cse) rectangle +(\side,\side) coordinate(cnw);
+	\begin{scope}
+	\clip (cse) rectangle (cnw);
+	\fill[ilmcolor] ++($(cse)!0.5!(cnw)$) circle (0.07\side)
+		+(0:0.3\side) circle (0.07\side) 
+		+(120:0.3\side) circle (0.07\side) 
+		+(-120:0.3\side) circle (0.07\side) 
+		+(-60:0.6\side) circle (0.07\side)
+		+(60:0.6\side) circle (0.07\side)
+		++(0:0.3\side)
+		+(-60:0.3\side) circle (0.07\side)
+		+(60:0.3\side) circle (0.07\side)
+		+(120:0.6\side) circle (0.07\side)
+		+(-120:0.6\side) circle (0.07\side)
+		++(180:0.9\side)
+		+(-60:0.3\side) circle (0.07\side)
+		+(60:0.3\side) circle (0.07\side);
+	\end{scope}
 	\end{tikzpicture}
+
+	\hfill$\textcolor{ilmcolor}{\tau_2} \rightarrow +\infty$ {\usebeamercolor[fg]{itemize item}\usebeamertemplate{itemize item}} arrest 
 	\end{columns}
@@
 
+\begin{frame}{Reproducibility in capillary}
+	\hfill $\phi\approx 0.33$, $c_P=\SI{0.40}{\milli\gram/\gram}$
+
+	\includegraphics[width=\textwidth]{nature/capillary}
+\end{frame}
+
+\begin{frame}{Transition probabilities}
+	\vspace{\baselineskip}
+	\includegraphics[width=\textwidth]{nature/sfig1.pdf}
+\end{frame}
+
+\begin{frame}{Evolution local volume fraction}
+	\vspace{\baselineskip}
+	12-neighboured particles
+	\begin{center}
+	\includegraphics{nature/local_phi12.pdf}
+	\end{center}
+\end{frame}
+
+\begin{frame}{phase definition}
+	\begin{description}
+	\item[crystal]
+		\begin{itemize} 
+		\item particle $i$ has $N_b(i)=12$ neighbours
+		\item $\displaystyle q_{6 m}(i)=\frac{1}{N_b(i)}\sum_{j=1}^{N_b(i)}
+	Y_{6 m}(\mathbf{r_{ij}}),\quad -6 <m<6$
+		\item crystalline bond $(i,j)$: \[(\mathbf{q}_{6}(i)/|\mathbf{q}_{6}(i)|)\cdot(\mathbf{q}_{6}(j)/|\mathbf{q}_{6}(j)|)>0.7\]
+		\item crystalline particle: $7$ crystalline bonds
+		\end{itemize}
+	\item[surface] two neighbouring crystal particles
+	\item[liquid] at least four neighbouring particles
+	\item[gas]
+	\end{description}
+\end{frame}
+
+\begin{frame}{Colloid and polymer sizes}
+	Direct measurement \alert{too imprecise}
+	\begin{itemize}
+	\item diffusion constant $\Rightarrow R= \SI{2.3 +- 0.05}{\micro\metre}$
+	\item Flory scaling litterature measurements $\Rightarrow R_g=\SI{76. +- 50.}{\nano\metre}$
+	\end{itemize}
+
+	\begin{block}{Adjusting theoretical and experimental phase diagrams}
+	\begin{enumerate}
+	\item to have the final local volume fraction of the 12-neighboured particles in the crystallising sample to be equal to the theoretical equilibrium crystal volume fraction 
+	\item to have the theoretical spinodal line to correspond to the gel region boundary.
+	\end{enumerate}
+	\end{block}
+	Starting from our initial estimates of $\sigma$ and $R$ we iteratively converge to $R=\SI{80}{\nano\metre}$, $\sigma = \SI{2.25}{\micro\metre}$ and a crystal (without defects) at $0.723$.
+\end{frame}
+
+\begin{frame}{Carnahan-Starling is not enough}
+	\tikzsetnextfilename{CSphasediag}%
+	\begin{tikzpicture}[
+		every axis/.style={
+			xlabel absolute, every axis x label/.append style={anchor=base, yshift=-0.5em},
+			ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
+			}]
+		\pgfplotscreateplotcyclelist{samples}{%
+		blue,mark=square*,\\%
+		cyan,mark=square\\%
+		red, thick, mark=*\\%
+		red!65!yellow, mark=otimes\\%
+		red!30!yellow, mark=oplus\\%
+		red!65!black, mark=o\\%
+		}%
+		\begin{groupplot}[
+			group style={
+					group name=g, group size=2 by 2,
+					horizontal sep=0.5em,
+					vertical sep=0.5em,
+					x descriptions at=edge bottom,
+					y descriptions at=edge left,
+				},
+			cycle list name=samples,
+			width=0.5\textwidth,
+			%height=0.5\linewidth,
+			axis on top,
+			xlabel={$\phi$},
+		]
+		\nextgroupplot[
+			ylabel={attraction},
+			xmin=0,xmax=0.74,
+			ymin=0, ymax=2400, %ytick={0,0.5,...,2},
+			mark options={solid}
+			]
+			\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
+		\addplot file {res_sample0.phd};
+		\addplot file {res_sample1.phd};
+		\addplot file {res_sample2.phd};
+		\addplot file {res_sample3.phd};
+		\addplot file {res_sample4.phd};
+		\addplot file {res_sample5.phd};
+		\end{scope}
+			\addplot[black, dashed,no marks] file {res_gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
+			\addplot[black, dashed,no marks] file {res_gasliquid_sl.phd};
+			\addplot[black, no marks] file {res_gasliquid_bl.phd};
+			\addplot[black, no marks] file {res_gasliquid_bg.phd};
+			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_x.phd}  -| (rel axis cs:1,0) \closedcycle;
+
+	\nextgroupplot[
+			xmin=0,xmax=0.74,
+			ymin=0, ymax=2400, %ytick={0,0.5,...,2},
+			mark options={solid}
+			]
+			\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
+		\addplot file {res_sample0.phd};
+		\addplot file {res_sample1.phd};
+		\addplot file {res_sample2.phd};
+		\addplot file {res_sample3.phd};
+		\addplot file {res_sample4.phd};
+		\addplot file {res_sample5.phd};
+		\end{scope}
+			\addplot[black, dashed,no marks] file {res_gasliquidCS_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
+			\addplot[black, dashed,no marks] file {res_gasliquidCS_sl.phd};
+			\addplot[black, no marks,skip coords between index={10}{100}] file {res_gasliquidCS_bl.phd};
+			\addplot[black, no marks,skip coords between index={10}{100}] file {res_gasliquidCS_bg.phd};
+			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_x.phd}  -| (rel axis cs:1,0) \closedcycle;
+
+		\nextgroupplot[
+			ylabel={P concentr. (\si{\milli\gram/\gram})},
+			xmin=0,xmax=0.74,
+			ymin=0, ymax=1.9, ytick={0,0.5,...,2},
+			mark options={solid}
+			]
+			\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
+			\addplot file {sample0.phd};
+			\addplot file {sample1.phd};
+			\addplot file {sample2.phd};
+			\addplot file {sample3.phd};
+			\addplot file {sample4.phd};
+			\addplot file {sample5.phd};
+			\end{scope}
+			\addplot[black, dashed,no marks] file {gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
+			\addplot[black, dashed,no marks] file {gasliquid_sl.phd};
+			\addplot[black, no marks] file {gasliquid_bl.phd};
+			\addplot[black, no marks] file {gasliquid_bg.phd};
+			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd}  -| (rel axis cs:1,0) \closedcycle;
+			\addplot[only marks, mark=triangle] file{fluid_samples.phd};
+
+		\nextgroupplot[
+			xmin=0,xmax=0.74,
+			ymin=0, ymax=1.9, ytick={0,0.5,...,2},
+			mark options={solid}
+			]
+			\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
+			\addplot file {sample0.phd};
+			\addplot file {sample1.phd};
+			\addplot file {sample2.phd};
+			\addplot file {sample3.phd};
+			\addplot file {sample4.phd};
+			\addplot file {sample5.phd};
+			\end{scope}
+			\addplot[black, dashed,no marks] file {gasliquidCS_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
+			\addplot[black, dashed,no marks] file {gasliquidCS_sl.phd};
+			\addplot[black, no marks,skip coords between index={10}{100}] file {gasliquidCS_bl.phd};
+			\addplot[black, no marks,skip coords between index={10}{100}] file {gasliquidCS_bg.phd};
+			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd}  -| (rel axis cs:1,0) \closedcycle;
+			\addplot[only marks, mark=triangle] file{fluid_samples.phd};
+	\end{groupplot}
+	\end{tikzpicture}
+\end{frame}
+
 \setcounter{framenumber}{\value{finalframe}}

### 2016-06-20 talk_Roscoff.tex
Attraction instead of Piv, etc.

@@
 
-\begin{frame}{Colloid + Polymer}
+\begin{frame}{Big atoms: Colloid + Polymer}
 
@@
 		\nextgroupplot[
-			ylabel={$\Pi v$},
+			ylabel={attraction},
 			xmin=0,xmax=0.74,
@@
 		\nextgroupplot[
-			xlabel={$\phi$}, ylabel={$c_p$ (\si{\milli\gram/\gram})},
+			xlabel={colloid volume fraction}, ylabel={P concentration  (\si{\milli\gram/\gram})},
 			xmin=0,xmax=0.74,

### 2016-06-20 crystal-gel/talk_Roscoff.tex
remove capitals in phase names. Note explicitely gas and liquid

@@
 			]
+			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {fluid} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_x.phd} node[above left, inner ysep=0, inner xsep=1pt] (x) at (rel axis cs:1,0) {crystal}  -| (rel axis cs:1,0) \closedcycle;
 			\addplot[black, dashed,no marks] file {res_gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
 			\addplot[black, dashed,no marks] file {res_gasliquid_sl.phd};
-			\addplot[black, no marks] file {res_gasliquid_bl.phd};
-			\addplot[black, no marks] file {res_gasliquid_bg.phd};
-			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
-			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_x.phd} node[above left, inner ysep=0, inner xsep=1pt] (x) at (rel axis cs:1,0) {Crystal}  -| (rel axis cs:1,0) \closedcycle;
-			\node at (axis cs:0.4,1500) {Gel};
+			\addplot[black, no marks] file {res_gasliquid_bl.phd} node[pos=0.5,sloped,below] {liquid};
+			\addplot[black, no marks] file {res_gasliquid_bg.phd} node[pos=0.1,sloped,below] {gas};
+
+			\node at (axis cs:0.35,2000) {g-l spinodal};
 
 		\nextgroupplot[
-			xlabel={$\phi$}, ylabel={$c_p$ (\si{\milli\gram/\gram})},
+			xlabel={colloid volume fraction}, ylabel={P concent.  (\si{\milli\gram/\gram})},
 			xmin=0,xmax=0.74,
@@
 			]
+			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {fluid} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd} %node[above left, inner ysep=0, inner xsep=1pt] (x) at (rel axis cs:1,0) {Crystal}  
+-| (rel axis cs:1,0) \closedcycle;
 			\addplot[black, dashed,no marks] file {gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
 			\addplot[black, dashed,no marks] file {gasliquid_sl.phd};
-			\addplot[black, no marks] file {gasliquid_bl.phd};
-			\addplot[black, no marks] file {gasliquid_bg.phd};
-			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
-			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd} node[above left] (x) at (rel axis cs:1,0.5) {Crystal}  -| (rel axis cs:1,0) \closedcycle;
-			\draw[->] (x.base) --(rel axis cs:0.98,0.01);
+			\addplot[black, no marks] file {gasliquid_bl.phd} node[pos=1,sloped,below] {liquid};
+			\addplot[black, no marks] file {gasliquid_bg.phd} node[pos=0.3,sloped,below] {gas};
+
+			%\draw[->] (x.base) --(rel axis cs:0.98,0.01);
 		\end{groupplot}
+		\node[anchor=north west, rotate=90, inner xsep=0,lightgray] (x) at (g c1r2.south east) {crystal};
 		%\draw(g c1r1.outer south west) -- +(\textwidth,0);
@@
 		]
-		\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
+		\begin{scope}[dotted, thick, mark repeat=2,mark phase=2,<->]
 		\addplot file {sample0.phd};
@@
 		\end{scope}
+		\addplot[only marks, mark=triangle] file{fluid_samples.phd};
+		\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {fluid} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd} %node[above left, inner ysep=0, inner xsep=1pt] (x) at (rel axis cs:1,0) {Crystal}  
+-| (rel axis cs:1,0) \closedcycle;
 			\addplot[black, dashed,no marks] file {gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
 			\addplot[black, dashed,no marks] file {gasliquid_sl.phd};
-		\addplot[black, no marks] file {gasliquid_bl.phd};
-		\addplot[black, no marks] file {gasliquid_bg.phd};
-		\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
-		\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd} node[above left] (x) at (rel axis cs:1,0.5) {Crystal}  -| (rel axis cs:1,0) \closedcycle;
-		%\draw[gray] (axis cs:0.371290311343, 0.12299370925) -- (axis cs:0.680497991395,0.0372636873433);
-		%\draw[gray, dotted] (axis cs:3.40152108798e-05,0.624739653238) -- (axis cs:0.722831684246, 0.0726528862772);
-		\draw[->] (x.base) --(rel axis cs:0.98,0.01);
-		\addplot[only marks, mark=triangle] file{fluid_samples.phd};
+			\addplot[black, no marks] file {gasliquid_bl.phd} node[pos=1,sloped,below] {liquid};
+			\addplot[black, no marks] file {gasliquid_bg.phd} node[pos=0.3,sloped,below] {gas};
 
@@
 	\end{groupplot}
+	\node[anchor=north west, rotate=90, inner xsep=0,lightgray] (x) at (g c1r1.south east) {crystal};
 	%\draw(g c1r1.outer south west) -- +(\textwidth,0);
@@
 			\addplot[black, no marks] file {gasliquid_bg.phd};
-			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {fluid} -- (rel axis cs:0,0) \closedcycle;
 			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd} %node[above left] (x) at (rel axis cs:1,0.5) {Crystal}  
@@
 
+\begin{frame}{Carnahan-Starling is not enough}
+	\tikzsetnextfilename{CSphasediag}%
+	\begin{tikzpicture}[
+		every axis/.style={
+			xlabel absolute, every axis x label/.append style={anchor=base, yshift=-0.5em},
+			ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
+			}]
+		\pgfplotscreateplotcyclelist{samples}{%
+		blue,mark=square*,\\%
+		cyan,mark=square\\%
+		red, thick, mark=*\\%
+		red!65!yellow, mark=otimes\\%
+		red!30!yellow, mark=oplus\\%
+		red!65!black, mark=o\\%
+		}%
+		\begin{groupplot}[
+			group style={
+					group name=g, group size=2 by 2,
+					horizontal sep=0.5em,
+					vertical sep=0.5em,
+					x descriptions at=edge bottom,
+					y descriptions at=edge left,
+				},
+			cycle list name=samples,
+			width=0.5\textwidth,
+			%height=0.5\linewidth,
+			axis on top,
+			xlabel={$\phi$},
+		]
+		\nextgroupplot[
+			ylabel={attraction},
+			xmin=0,xmax=0.74,
+			ymin=0, ymax=2400, %ytick={0,0.5,...,2},
+			mark options={solid}
+			]
+			\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
+		\addplot file {res_sample0.phd};
+		\addplot file {res_sample1.phd};
+		\addplot file {res_sample2.phd};
+		\addplot file {res_sample3.phd};
+		\addplot file {res_sample4.phd};
+		\addplot file {res_sample5.phd};
+		\end{scope}
+			\addplot[black, dashed,no marks] file {res_gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
+			\addplot[black, dashed,no marks] file {res_gasliquid_sl.phd};
+			\addplot[black, no marks] file {res_gasliquid_bl.phd};
+			\addplot[black, no marks] file {res_gasliquid_bg.phd};
+			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_x.phd}  -| (rel axis cs:1,0) \closedcycle;
+
+	\nextgroupplot[
+			xmin=0,xmax=0.74,
+			ymin=0, ymax=2400, %ytick={0,0.5,...,2},
+			mark options={solid}
+			]
+			\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
+		\addplot file {res_sample0.phd};
+		\addplot file {res_sample1.phd};
+		\addplot file {res_sample2.phd};
+		\addplot file {res_sample3.phd};
+		\addplot file {res_sample4.phd};
+		\addplot file {res_sample5.phd};
+		\end{scope}
+			\addplot[black, dashed,no marks] file {res_gasliquidCS_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
+			\addplot[black, dashed,no marks] file {res_gasliquidCS_sl.phd};
+			\addplot[black, no marks,skip coords between index={10}{100}] file {res_gasliquidCS_bl.phd};
+			\addplot[black, no marks,skip coords between index={10}{100}] file {res_gasliquidCS_bg.phd};
+			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_x.phd}  -| (rel axis cs:1,0) \closedcycle;
+
+		\nextgroupplot[
+			ylabel={P concentr. (\si{\milli\gram/\gram})},
+			xmin=0,xmax=0.74,
+			ymin=0, ymax=1.9, ytick={0,0.5,...,2},
+			mark options={solid}
+			]
+			\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
+			\addplot file {sample0.phd};
+			\addplot file {sample1.phd};
+			\addplot file {sample2.phd};
+			\addplot file {sample3.phd};
+			\addplot file {sample4.phd};
+			\addplot file {sample5.phd};
+			\end{scope}
+			\addplot[black, dashed,no marks] file {gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
+			\addplot[black, dashed,no marks] file {gasliquid_sl.phd};
+			\addplot[black, no marks] file {gasliquid_bl.phd};
+			\addplot[black, no marks] file {gasliquid_bg.phd};
+			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd}  -| (rel axis cs:1,0) \closedcycle;
+			\addplot[only marks, mark=triangle] file{fluid_samples.phd};
+
+		\nextgroupplot[
+			xmin=0,xmax=0.74,
+			ymin=0, ymax=1.9, ytick={0,0.5,...,2},
+			mark options={solid}
+			]
+			\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
+			\addplot file {sample0.phd};
+			\addplot file {sample1.phd};
+			\addplot file {sample2.phd};
+			\addplot file {sample3.phd};
+			\addplot file {sample4.phd};
+			\addplot file {sample5.phd};
+			\end{scope}
+			\addplot[black, dashed,no marks] file {gasliquidCS_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
+			\addplot[black, dashed,no marks] file {gasliquidCS_sl.phd};
+			\addplot[black, no marks,skip coords between index={10}{100}] file {gasliquidCS_bl.phd};
+			\addplot[black, no marks,skip coords between index={10}{100}] file {gasliquidCS_bg.phd};
+			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd}  -| (rel axis cs:1,0) \closedcycle;
+			\addplot[only marks, mark=triangle] file{fluid_samples.phd};
+	\end{groupplot}
+	\end{tikzpicture}
+\end{frame}
+
 \setcounter{framenumber}{\value{finalframe}}

### 2016-06-20 crystal-gel/talk_Roscoff.tex
Single side with overlays to show ~fig1, with phase diagram always visible.

@@
 \presetkeys{todonotes}{inline}{}
+
+\tikzset{onslide/.code args={<#1>#2}{%
+  \only<#1>{\pgfkeysalso{#2}} %
+}}%
 
@@
 
-\begin{frame}{Exploring phase diagram}
+\begin{frame}{Exploring the phase diagram}
 \vspace{\baselineskip}
-	\tikzsetnextfilename{phdiag_cluster}%
+	\tikzsetnextfilename{phdiag_panels}%
 	\begin{tikzpicture}[
@@
 		xlabel absolute, every axis x label/.append style={anchor=base, yshift=-1em},
-		%ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
+		ylabel absolute, every axis y label/.append style={anchor=base, yshift=-1em}
 		}]
@@
 		]
-		\begin{scope}[dotted, thick, mark repeat=2,mark phase=2,<->]
+		\begin{scope}[
+			onslide={<1,3->{/pgfplots/skip coords between index={0}{1}, /pgfplots/skip coords between index={2}{3}}},
+			onslide={<2>{dotted, thick, mark repeat=2,mark phase=2,<->}}]
 		\addplot file {sample0.phd};
@@
 
+	\only<3>{
 	\nextgroupplot[
@@
 		\addplot[gray, ultra thick, domain=1:4e3] {0.629*x^(1/2.53)} node [left] at (rel axis cs:1,0.9){$D=2.53$};
-
-	\end{groupplot}
-	\node[anchor=north west, rotate=90, inner xsep=0,lightgray] (x) at (g c1r1.south east) {crystal};
-	%\draw(g c1r1.outer south west) -- +(\textwidth,0);
-	\end{tikzpicture}
-	\begin{itemize}
-	\item Good agreement with theoretical phase diagram
-	\item At percolation time, compatible with random gelation
-	\end{itemize}
-\end{frame}
-
-\begin{frame}{Outlier at later times}
-	\vspace{\baselineskip}
-	\tikzsetnextfilename{Sq_localphi}%
-	\begin{tikzpicture}[
-		every axis/.style={
-		xlabel absolute, every axis x label/.append style={anchor=base, yshift=-1em},
-		%ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
-		}]
-	\pgfplotscreateplotcyclelist{samples}{%
-	blue,mark=square*,\\%
-	cyan,mark=square\\%
-	red, thick, mark=*\\%
-	red!65!yellow, mark=otimes\\%
-	red!30!yellow, mark=oplus\\%
-	red!65!black, mark=o\\%
-	}%
-	\begin{groupplot}[
-		group style={
-				group name=g, group size=2 by 2,
-				horizontal sep=4em,
-				vertical sep=3em,
-				%x descriptions at=edge bottom,
-			},
-		cycle list name=samples,
-		width=0.51\textwidth,
-		height=0.5\linewidth,
-		axis on top,
-	]
+	}
+	\only<4>{
 	\nextgroupplot[
@@
 		\addplot[gray, ultra thick, domain=0.1:3] {3.98797*x^(-2.53)} node [pos=0.65, above right] {$D=2.53$};
-
+	}
+	\only<5>{
 	\nextgroupplot[
@@
 		\addplot table[y error=err]{data/vol_err_phi_0.25_cp_0.57.scaled};
+	}
 
 	\end{groupplot}
+	\node[anchor=north west, rotate=90, inner xsep=0,lightgray] (x) at (g c1r1.south east) {crystal};
 	%\draw(g c1r1.outer south west) -- +(\textwidth,0);
 	\end{tikzpicture}
+	\vspace{-0.5em}
 	\begin{itemize}
-	\item Final stucture compatible with random gelation
-	\item Glassy local $\phi$ for 12-neighboured particles
+	\item<2-> Good agreement with theoretical phase diagram
+	\item<3-> At percolation time, compatible with random gelation
+	\item<4-> Final stucture compatible with random gelation \alert{Except} \textcolor{red}{$\bullet$}
+	\item<5-> Glassy local $\phi$ for 12-neighboured particles \alert{Except} \textcolor{red}{$\bullet$}
 	\end{itemize}
-	\alert{Except} \textcolor{red}{$\bullet$}: compact structure and higher local $\phi$. %\textcolor{ilmorange}{$\Rightarrow$} Crystal?
 \end{frame}

### 2016-06-20 talk_Roscoff.tex
evolution of local phi for 12-neighboured particles as yet another overlay rather than a backslide

@@
 			]
+			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {fluid} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_x.phd} node[above left, inner ysep=0, inner xsep=1pt] (x) at (rel axis cs:1,0) {crystal}  -| (rel axis cs:1,0) \closedcycle;
 			\addplot[black, dashed,no marks] file {res_gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
 			\addplot[black, dashed,no marks] file {res_gasliquid_sl.phd};
-			\addplot[black, no marks] file {res_gasliquid_bl.phd};
-			\addplot[black, no marks] file {res_gasliquid_bg.phd};
-			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
-			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_x.phd} node[above left, inner ysep=0, inner xsep=1pt] (x) at (rel axis cs:1,0) {Crystal}  -| (rel axis cs:1,0) \closedcycle;
-			\node at (axis cs:0.4,1500) {Gel};
+			\addplot[black, no marks] file {res_gasliquid_bl.phd} node[pos=0.5,sloped,below] {liquid};
+			\addplot[black, no marks] file {res_gasliquid_bg.phd} node[pos=0.1,sloped,below] {gas};
+
+			\node at (axis cs:0.35,2000) {g-l spinodal};
 
 		\nextgroupplot[
-			xlabel={colloid volume fraction}, ylabel={P concentration  (\si{\milli\gram/\gram})},
+			xlabel={colloid volume fraction}, ylabel={P concent.  (\si{\milli\gram/\gram})},
 			xmin=0,xmax=0.74,
@@
 		]
-		\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
+		\begin{scope}[
+			onslide={<1,3->{/pgfplots/skip coords between index={0}{1}, /pgfplots/skip coords between index={2}{3}}},
+			onslide={<2>{dotted, thick, mark repeat=2,mark phase=2,<->}}]
 		\addplot file {sample0.phd};
@@
 		\end{scope}
+		\addplot[only marks, mark=triangle] file{fluid_samples.phd};
+		\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {fluid} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd} %node[above left, inner ysep=0, inner xsep=1pt] (x) at (rel axis cs:1,0) {Crystal}  
+-| (rel axis cs:1,0) \closedcycle;
 			\addplot[black, dashed,no marks] file {gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
 			\addplot[black, dashed,no marks] file {gasliquid_sl.phd};
-		\addplot[black, no marks] file {gasliquid_bl.phd};
-		\addplot[black, no marks] file {gasliquid_bg.phd};
-		\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
-		\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd} node[above left] (x) at (rel axis cs:1,0.5) {Crystal}  -| (rel axis cs:1,0) \closedcycle;
-		%\draw[gray] (axis cs:0.371290311343, 0.12299370925) -- (axis cs:0.680497991395,0.0372636873433);
-		%\draw[gray, dotted] (axis cs:3.40152108798e-05,0.624739653238) -- (axis cs:0.722831684246, 0.0726528862772);
-		\draw[->] (x.base) --(rel axis cs:0.98,0.01);
-		\addplot[only marks, mark=triangle] file{fluid_samples.phd};
+			\addplot[black, no marks] file {gasliquid_bl.phd} node[pos=1,sloped,below] {liquid};
+			\addplot[black, no marks] file {gasliquid_bg.phd} node[pos=0.3,sloped,below] {gas};
 
+	\only<3>{
 	\nextgroupplot[
@@
 		\addplot[gray, ultra thick, domain=1:4e3] {0.629*x^(1/2.53)} node [left] at (rel axis cs:1,0.9){$D=2.53$};
-
-	\end{groupplot}
-	%\draw(g c1r1.outer south west) -- +(\textwidth,0);
-	\end{tikzpicture}
-	\begin{itemize}
-	\item Good agreement with theoretical phase diagram
-	\item At percolation time, compatible with random gelation
-	\end{itemize}
-\end{frame}
-
-\begin{frame}{Outlier at later times}
-	\vspace{\baselineskip}
-	\tikzsetnextfilename{Sq_localphi}%
-	\begin{tikzpicture}[
-		every axis/.style={
-		xlabel absolute, every axis x label/.append style={anchor=base, yshift=-1em},
-		%ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
-		}]
-	\pgfplotscreateplotcyclelist{samples}{%
-	blue,mark=square*,\\%
-	cyan,mark=square\\%
-	red, thick, mark=*\\%
-	red!65!yellow, mark=otimes\\%
-	red!30!yellow, mark=oplus\\%
-	red!65!black, mark=o\\%
-	}%
-	\begin{groupplot}[
-		group style={
-				group name=g, group size=2 by 2,
-				horizontal sep=4em,
-				vertical sep=3em,
-				%x descriptions at=edge bottom,
-			},
-		cycle list name=samples,
-		width=0.51\textwidth,
-		height=0.5\linewidth,
-		axis on top,
-	]
+	}
+	\only<4>{
 	\nextgroupplot[
@@
 		\addplot table[y error=err]{data/vol_err_phi_0.25_cp_0.57.scaled};
-
+	}
+	\only<6>{
+	\nextgroupplot[
+		title={local $\phi$ 12-ngb particles},
+		xlabel=$t/10^3\tau_B$, ylabel={$\phi$ [\%]},
+		xmin=0,
+		/pgfplots/error bars/y dir=both, /pgfplots/error bars/y explicit,
+		]
+		\addplot table[y error=err]{data/mediaphi_phi0.10_cp0.82.scaled};
+		\addplot table[y error=err]{data/mediaphi_phi0.10_cp1.36.scaled};
+		\addplot table[y error=err]{data/mediaphi_phi0.25_cp0.38.scaled};
+		\addplot table[y error=err]{data/mediaphi_phi0.25_cp0.48.scaled};
+		\addplot table[y error=err]{data/mediaphi_phi0.25_cp0.57.scaled};
+	}
 	\end{groupplot}
+	\node[anchor=north west, rotate=90, inner xsep=0,lightgray] (x) at (g c1r1.south east) {crystal};
 	%\draw(g c1r1.outer south west) -- +(\textwidth,0);
 	\end{tikzpicture}
+	\vspace{-0.5em}
 	\begin{itemize}
-	\item Final stucture compatible with random gelation
-	\item Glassy local $\phi$ for 12-neighboured particles
+	\item<2-> Good agreement with theoretical phase diagram
+	\item<3-> At percolation time, compatible with random gelation
+	\item<4-> Final stucture compatible with random gelation \alert{Except} \textcolor{red}{$\bullet$}
+	\item<5-> Glassy local $\phi$ for 12-neighboured particles \alert{Except} \textcolor{red}{$\bullet$}
 	\end{itemize}
-	\alert{Except} \textcolor{red}{$\bullet$}: compact structure and higher local $\phi$. %\textcolor{ilmorange}{$\Rightarrow$} Crystal?
 \end{frame}
@@
 	\includegraphics[width=\textwidth]{nature/sfig1.pdf}
-\end{frame}
-
-\begin{frame}{Evolution local volume fraction}
-	\vspace{\baselineskip}
-	12-neighboured particles
-	\begin{center}
-	\includegraphics{nature/local_phi12.pdf}
-	\end{center}
 \end{frame}

### 2016-06-20 crystal-gel/talk_Roscoff.tex
Move "Another class of gel? Crystal-gel?" from intro to after crystal gel detected "Crystals outgrow metastable liquid" with graphs + experimental slices "Dynamics: Bergeron > Ostwald" with pathX scheme and triangle (redone in tikz)

@@
 
-\begin{frame}{Another class of gel? Crystal-gel?}
-	\begin{itemize}
-	\item Phase separation arrested by crystallisation
-	\begin{description}
-		\item[polymer blend] Tanaka \& Nishi \textit{PRL} 1985
-		\item[simulations] Soga \textit{J.Chem. Phys.} 1999; Fortini \textit{PRE} 2008 
-	\end{description}
-	\item Spinodal then crystallisation
-	\begin{description}
-		\item[simulations] Pérez \textit{Langmuir} 2011
-		\item[microgravity] (triple coexistence) Sabin \textit{PRL} 2012
-	\end{description}
-	\end{itemize}
-	\includegraphics[width=\textwidth]{presentation/micrograv_Sabin2012}
-\end{frame}
-
 \begin{frame}{Experimental}
@@
 		extra y ticks={64, 72}, extra y tick style={grid=major}, extra y tick label={},
-		legend pos=north west, legend style={font=\footnotesize},
 		no markers,
@@
 	}
-
+	\only<6>{
+	\nextgroupplot[
+		title={local $\phi$ 12-ngb particles},
+		xlabel=$t/10^3\tau_B$, ylabel={$\phi$ [\%]},
+		xmin=0,
+		/pgfplots/error bars/y dir=both, /pgfplots/error bars/y explicit,
+		]
+		\addplot table[y error=err]{data/mediaphi_phi0.10_cp0.82.scaled};
+		\addplot table[y error=err]{data/mediaphi_phi0.10_cp1.36.scaled};
+		\addplot table[y error=err]{data/mediaphi_phi0.25_cp0.38.scaled};
+		\addplot table[y error=err]{data/mediaphi_phi0.25_cp0.48.scaled};
+		\addplot table[y error=err]{data/mediaphi_phi0.25_cp0.57.scaled};
+	}
 	\end{groupplot}
 	\node[anchor=north west, rotate=90, inner xsep=0,lightgray] (x) at (g c1r1.south east) {crystal};
-	%\draw(g c1r1.outer south west) -- +(\textwidth,0);
 	\end{tikzpicture}
@@
 
-\begin{frame}{Crystals outgrow liquid}
+\begin{frame}{Another class of gel? Crystal-gel?}
+	\begin{itemize}
+	\item Phase separation arrested by crystallisation
+	\begin{description}
+		\item[polymer blend] Tanaka \& Nishi \textit{PRL} 1985
+		\item[simulations] Soga \textit{J.Chem. Phys.} 1999; Fortini \textit{PRE} 2008 
+	\end{description}
+	\item Spinodal then crystallisation
+	\begin{description}
+		\item[simulations] Pérez \textit{Langmuir} 2011
+		\item[microgravity] (triple coexistence) Sabin \textit{PRL} 2012
+	\end{description}
+	\end{itemize}
+	\includegraphics[width=\textwidth]{presentation/micrograv_Sabin2012}
+\end{frame}
+
+\begin{frame}{Crystals outgrow metastable liquid}
 	\vspace{\baselineskip}
@@
 			%height=0.5\linewidth,
+			xlabel absolute, every axis x label/.append style={anchor=base, font=\footnotesize, yshift=-0.5em},
 		]
@@
 	\end{tikzpicture}
-\begin{center}
-\tikzsetnextfilename{pathsX}\vspace{-0.5em}\input{presentation/pathsX.pgf}
-\end{center}
+
+	\hfill\includegraphics[width=0.95\textwidth, clip=true, trim=0 20cm 0 0]{nature/fig4.pdf}\hfill
 \end{frame}
@@
 	\vspace{\baselineskip}
-\includegraphics[width=\textwidth]{nature/fig4.pdf}
+	\begin{center}
+	\tikzsetnextfilename{pathsX}\vspace{-0.5em}\input{presentation/pathsX.pgf}
+	\end{center}
+	\begin{columns}
+	\column{0.5\textwidth}
+	\begin{itemize}
+	\item Individual trajectories
+	\item Probability of path to X
+	\end{itemize}
+	\column{0.5\textwidth}
+	\tikzsetnextfilename{pathsX2}%
+	\begin{tikzpicture}
+	\node[regular polygon, regular polygon sides=3, minimum width=5em] (tri){};
+	\node[circle, minimum width=3.5em, fill=green!40, left=1em of tri.corner 2] (gas) {gas};
+	\node[circle, minimum width=3.5em, draw, fill=white, right =1em of tri.corner 3] (liq) {liquid};
+	\node[regular polygon, regular polygon sides=6, fill=red,minimum width=3.5em, inner sep=0, above=0 of tri.corner 1] (X) {crystal};
+	\draw[->, very thick] ($(tri.corner 2)+(0,0.25em)$) -- ($(liq.west)+(0,0.25em)$) node[midway, above]{54\%}-- (X);
+	\draw[->, very thick, dotted] ($(tri.corner 3)+(0,-0.25em)$) -- ($(gas.east)+(0,-0.25em)$) node[midway, below]{35\%}-- (X);
+	\draw[->, very thick, dashed] (X) edge [in=190,out=-130,loop] node[left]{11\%} ();
+	\end{tikzpicture}
+	\end{columns}
 \end{frame}

### 2016-06-21 talk_Roscoff.tex
More compréhensive slide about colloids as big atoms Gel formation: show phase sep on the phase diagram, liquid and gas in slow and fast colors, introduce glass transition as an overlay

@@
 \begin{frame}{Big atoms: Colloid + Polymer}
-
 	\begin{columns}
-	\column{0.45\textwidth}
+	\column{0.5\textwidth}
+	\begin{itemize}
+		\item visible, trackable
+		\item tunable interactions
+	\end{itemize}
+	\medskip
 	\tikzsetnextfilename{interactions}%
-	\begin{tikzpicture}
+	\begin{tikzpicture}[every node/.style={font=\footnotesize}]
 		\let\mrad\relax
@@
 		\node[text width=6em, anchor=base east, inner xsep=0,xshift=1.5\mrad, align=right] (coll) at (colli.east|-pur.base) {depletion};
-	\end{tikzpicture}\\
-
-	\bigskip
-	\includegraphics[width=\textwidth]{../CNRS-2015/Renth_sedim.jpg}\\
-	\textit{\footnotesize Poon PRL 1999, Renth PRE 2001}
+	\end{tikzpicture}
+	\begin{block}{Various phases}
 		\begin{itemize}
-	\item triple coexistence
-	\item various kinetic paths
+			\item triple coexistence\\
+			\textit{\footnotesize Poon PRL 1999}
+			\item gas-liquid metastable\\\hfill to gas-crystal
 		\end{itemize}
-
-	\column{0.45\textwidth}
-	\includegraphics[width=\textwidth,clip=true,trim=0 0 0 4.5cm]{presentation/Renth_diag.pdf}
+	\end{block}
+	\column{0.5\textwidth}
+	\vspace{0.5\baselineskip}
+	\hfill\tikzsetnextfilename{sedim_glx}%
+	\begin{tikzpicture}
+		\node[inner sep=0] (sedim){\includegraphics[height=17\baselineskip, clip=true, trim=0 250 0 0]{presentation/sedim_gas_liq_crystal_p.png}};
+		\path[every node/.style={left}] (sedim.north west) -- (sedim.south west) node[pos=0.1] {gas} node[pos=0.45] {liquid} node[pos=0.8] {crystal} node[pos=0.15, ilmorange, font=\itshape] {low viscosity} node[pos=0.5, ilmcolor, font=\itshape] {high viscosity};
+		\node[anchor=north east, font=\scriptsize, inner sep=1pt] at (sedim.south east) {Paddy Royall};
+	\end{tikzpicture}
 	\end{columns}
@@
 			}]
-		\pgfplotscreateplotcyclelist{samples}{%
-		blue,mark=square*,\\%
-		cyan,mark=square\\%
-		red, thick, mark=*\\%
-		red!65!yellow, mark=otimes\\%
-		red!30!yellow, mark=oplus\\%
-		red!65!black, mark=o\\%
-		}%
+% 		\pgfplotscreateplotcyclelist{samples}{%
+% 		blue,mark=square*,\\%
+% 		cyan,mark=square\\%
+% 		red, thick, mark=*\\%
+% 		red!65!yellow, mark=otimes\\%
+% 		red!30!yellow, mark=oplus\\%
+% 		red!65!black, mark=o\\%
+% 		}%
 		\begin{groupplot}[
@@
 				},
-			cycle list name=samples,
+			%cycle list name=samples,
 			width=\textwidth,
@@
 			]
-			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {fluid} -- (rel axis cs:0,0) \closedcycle;
+			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {fluid} -- (rel axis cs:0,0) coordinate (O) \closedcycle;
 			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_x.phd} node[above left, inner ysep=0, inner xsep=1pt] (x) at (rel axis cs:1,0) {crystal}  -| (rel axis cs:1,0) \closedcycle;
@@
 			\addplot[black, dashed,no marks] file {res_gasliquid_sl.phd};
-			\addplot[black, no marks] file {res_gasliquid_bl.phd} node[pos=0.5,sloped,below] {liquid};
-			\addplot[black, no marks] file {res_gasliquid_bg.phd} node[pos=0.1,sloped,below] {gas};
+			\addplot[black, no marks] file {res_gasliquid_bl.phd} node[pos=0.5,sloped,below,ilmcolor] (l) {liquid};
+			\addplot[black, no marks] file {res_gasliquid_bg.phd} node[pos=0.1,sloped,below,ilmorange] {gas};
 
 			\node at (axis cs:0.35,2000) {g-l spinodal};
+			\draw[<->, thick, dotted] (l) -- (l -| O) node[pos=0.5,fill] {};
+			%glass
+			\only<2>{\addplot[pattern=crosshatch, pattern color=ilmcolor,no marks, draw=none] coordinates {(0.56,0) (0.6,500) (0.56,2400)}  -- (rel axis cs:1,1)  \closedcycle;}
 
@@
 			\addplot[black, dashed,no marks] file {gasliquid_sl.phd};
-			\addplot[black, no marks] file {gasliquid_bl.phd} node[pos=1,sloped,below] {liquid};
-			\addplot[black, no marks] file {gasliquid_bg.phd} node[pos=0.3,sloped,below] {gas};
+			\addplot[black, no marks] file {gasliquid_bl.phd} node[pos=1,sloped,below,ilmcolor] (l) {liquid};
+			\addplot[black, no marks] file {gasliquid_bg.phd} node[pos=0.3,sloped,below,ilmorange] {gas};
+			\draw[<->, thick, dotted] (l.north) -- (rel axis cs:0,1) node[pos=0.5,fill] {};
+			%glass
+			\only<2>{\addplot[pattern=crosshatch, pattern color=ilmcolor,no marks, draw=none] coordinates {(0.56,0) (0.6,0.125) (0.56,2)}  -- (rel axis cs:1,1)  \closedcycle;}
 
@@
 		\column{0.5\textwidth}
-		Small polymer/colloid size ratio
-		\begin{itemize}
-		\item metastable gas-liquid
-		\item spinodal + glass $\Rightarrow$ gel\\ \hfill\textit{\footnotesize Lu et al. Nature~2008}
-		\item continuous minority phase
-		\item viscoelastic phase sep.\\ \hfill\textit{\footnotesize Tanaka J.Phys.Cond.Mat.~2000}
-		\item due to dynamic assymetry
-		\end{itemize}
+		%Small polymer/colloid size ratio
+		\structure{Dynamic assymetry}
 		\tikzsetnextfilename{vps}%
@@
 	\draw[->] (0,0) -- +(\textwidth,0) node[pos=0.05, label=-90:\textcolor{ilmorange}{$\tau_1$}](t1) {|} node[pos=0.54, label=-90:\textcolor{ilmcolor}{$\tau_2$}] (t2){|} coordinate[pos=1] (tf) node[anchor=north east,font=\footnotesize]{time};
+	\node[below, font=\scriptsize, text width=0.4\textwidth, align=center] at ($(t1)!0.5!(t2)$) {continuous \textcolor{ilmcolor}{minority} phase};
 	\let\side\relax
@@
 	\end{tikzpicture}
-
-	\hfill$\textcolor{ilmcolor}{\tau_2} \rightarrow +\infty$ {\usebeamercolor[fg]{itemize item}\usebeamertemplate{itemize item}} arrest 
+	\begin{itemize}
+		\item viscoelastic phase sep.\\ \hfill\textit{\footnotesize Tanaka J.Phys.Cond.Mat.~2000}
+		\pause
+		\item glass transition $\textcolor{ilmcolor}{\tau_2} \rightarrow +\infty$
+		\item arrested phase separation
+		\item spinodal + glass $\Rightarrow$ gel\\ \hfill\textit{\footnotesize Lu et al. Nature~2008}
+	\end{itemize}	
 	\end{columns}
-\end{frame}
-
-\begin{frame}{Another class of gel? Crystal-gel?}
-	\begin{itemize}
-	\item Phase separation arrested by crystallisation
-	\begin{description}
-		\item[polymer blend] Tanaka \& Nishi \textit{PRL} 1985
-		\item[simulations] Soga \textit{J.Chem. Phys.} 1999; Fortini \textit{PRE} 2008 
-	\end{description}
-	\item Spinodal then crystallisation
-	\begin{description}
-		\item[simulations] Pérez \textit{Langmuir} 2011
-		\item[microgravity] (triple coexistence) Sabin \textit{PRL} 2012
-	\end{description}
-	\end{itemize}
-	\includegraphics[width=\textwidth]{presentation/micrograv_Sabin2012}
 \end{frame}
@@
 	\node[anchor=north west, rotate=90, inner xsep=0,lightgray] (x) at (g c1r1.south east) {crystal};
-	%\draw(g c1r1.outer south west) -- +(\textwidth,0);
 	\end{tikzpicture}

### 2016-06-21 talk_Roscoff.tex
Experimental: whole process, 10^4 particles, from t=0, to t>10^4 tau_B
fix nuclei size distribution

@@
 \begin{frame}{Experimental}
-How to do that in 3D at particle level \alert{without a space station}?
+	\vspace{\baselineskip}
+	How to observe the whole process in 3D at particle level?
 	\begin{itemize}
@@
 			\input{presentation/cell.pgf}
+			\vspace{-\baselineskip}
+			\begin{itemize}
+			\item $\sim 10^4$ particles
+			\item from $t=0$
+			\item to $t>10^4\tau_B$
+			\end{itemize}
 	\end{columns}
@@
 		\nextgroupplot[
-			title={cluster size distribution},
+			title={nuclei size distribution},
 			xlabel={$s$}, ylabel=$R_g/\sigma$,
-			xmode=log, ymode=log,
+			xmode=log, ymode=log, ymin=0.3,
 			ytick={0.5, 1, 2, 4}, yticklabels={$\frac{1}{2}$, 1, 2, 4},
-			domain=1:5e2,
+			domain=1.1:5e2,
 			]
+		\draw[dashed] (rel axis cs:0.5,0) -- (rel axis cs:0.5,1);
 		\addplot[only marks, red, thick, mark=*] file{data/fig3d_sqrt_1.dat};
-		\addplot[black, ultra thick] {0.522*x^(1/3)} node[below left] at (axis cs:9e2,2.5) {$D=3$};
-		\addplot[gray, ultra thick] {0.426*x^(1/2.53)} node[pos=0.08,below right] {$D=2.53$};
+		\addplot[black, ultra thick] {0.522*x^(1/3)} node[below left] at (axis cs:9e2,2.5) {$D=3$} node[anchor=north east, font=\footnotesize, text width=0.1\textwidth] at (rel axis cs:1,0.5) {outgrow\linebreak network};
+		\addplot[gray, ultra thick] {0.426*x^(1/2.53)} node[anchor=south west, inner sep=1pt] at (rel axis cs:0,0){$D=2.53$} node[anchor=north east, font=\footnotesize, text width=0.2\textwidth, align=right] at (rel axis cs:0.45,1) {within\linebreak network};
+
 	\end{groupplot}

### 2016-06-21 crystal-gel/talk_Roscoff.tex
phase inversion

@@
 	\draw[->] (0,0) -- +(\textwidth,0) node[pos=0.05, label=-90:\textcolor{ilmorange}{$\tau_1$}](t1) {|} node[pos=0.54, label=-90:\textcolor{ilmcolor}{$\tau_2$}] (t2){|} coordinate[pos=1] (tf) node[anchor=north east,font=\footnotesize]{time};
+	\node[below, font=\scriptsize, text width=0.4\textwidth, align=center] at ($(t1)!0.5!(t2)$) {continuous \textcolor{ilmcolor}{minority} phase};
+	\node[below, font=\scriptsize, text width=0.3\textwidth, align=center] at ($(t2)!0.4!(tf)$) {phase inversion};
 	\let\side\relax
@@
 	\end{tikzpicture}
-
-	\hfill$\textcolor{ilmcolor}{\tau_2} \rightarrow +\infty$ {\usebeamercolor[fg]{itemize item}\usebeamertemplate{itemize item}} arrest 
+	\begin{itemize}
+		\item viscoelastic phase sep.\\ \hfill\textit{\footnotesize Tanaka J.Phys.Cond.Mat.~2000}
+		\pause
+		\item glass transition $\textcolor{ilmcolor}{\tau_2} \rightarrow +\infty$
+		\item arrested phase separation
+		\item spinodal + glass $\Rightarrow$ gel\\ \hfill\textit{\footnotesize Lu et al. Nature~2008}
+	\end{itemize}	
 	\end{columns}

### 2016-06-22 talk_Roscoff.tex
liquid in lightgray use relative positionning in pathsX

@@
 	\node[below, font=\scriptsize, text width=0.4\textwidth, align=center] at ($(t1)!0.5!(t2)$) {continuous \textcolor{ilmcolor}{minority} phase};
+	\node[below, font=\scriptsize, text width=0.3\textwidth, align=center] at ($(t2)!0.4!(tf)$) {phase inversion};
 	\let\side\relax
@@
 	\begin{tikzpicture}
+	\colorlet{gas}{green!40}%
+	\colorlet{crystal}{red}%
+	\colorlet{liquid}{lightgray}%
 	\node[regular polygon, regular polygon sides=3, minimum width=5em] (tri){};
-	\node[circle, minimum width=3.5em, fill=green!40, left=1em of tri.corner 2] (gas) {gas};
-	\node[circle, minimum width=3.5em, draw, fill=white, right =1em of tri.corner 3] (liq) {liquid};
-	\node[regular polygon, regular polygon sides=6, fill=red,minimum width=3.5em, inner sep=0, above=0 of tri.corner 1] (X) {crystal};
+	\node[circle, minimum width=3.5em, fill=gas, left=1em of tri.corner 2] (gas) {gas};
+	\node[circle, minimum width=3.5em, fill=liquid, right =1em of tri.corner 3] (liq) {liquid};
+	\node[regular polygon, regular polygon sides=6, fill=crystal,minimum width=3.5em, inner sep=0, above=0 of tri.corner 1] (X) {crystal};
 	\draw[->, very thick] ($(tri.corner 2)+(0,0.25em)$) -- ($(liq.west)+(0,0.25em)$) node[midway, above]{54\%}-- (X);

### 2016-06-22 crystal-gel/talk_Roscoff.tex
use the color scheme of iLM to differentiate the phases in pathsX. More friendly to color blind.

@@
 	\begin{tikzpicture}
+		\colorlet{an}{black}
+		\colorlet{gas}{ilmorange!50!-an}%
+		\colorlet{crystal}{ilmcolor!50!-an}%
+		\colorlet{liquid}{ilmcolor!90!-an}%
 		\node[regular polygon, regular polygon sides=3, minimum width=5em] (tri){};
-	\node[circle, minimum width=3.5em, fill=green!40, left=1em of tri.corner 2] (gas) {gas};
-	\node[circle, minimum width=3.5em, draw, fill=white, right =1em of tri.corner 3] (liq) {liquid};
-	\node[regular polygon, regular polygon sides=6, fill=red,minimum width=3.5em, inner sep=0, above=0 of tri.corner 1] (X) {crystal};
+		\node[circle, minimum width=3.5em, fill=gas, left=1em of tri.corner 2] (gas) {gas};
+		\node[circle, minimum width=3.5em, fill=liquid, right =1em of tri.corner 3] (liq) {liquid};
+		\node[regular polygon, regular polygon sides=6, fill=crystal,minimum width=3.5em, inner sep=0, above=0 of tri.corner 1] (X) {crystal};
 		\draw[->, very thick] ($(tri.corner 2)+(0,0.25em)$) -- ($(liq.west)+(0,0.25em)$) node[midway, above]{54\%}-- (X);

### 2016-06-23 data/slicesILM.py
lighter color for the liquid than for the crystal generate povray images with coherent color scheme that replace the call to nature/fig4.pdf

@@ -1,0 +1,30 @@
+from colloids.povray import *
+import subprocess
+
+#4 -> LIQUID
+#5 -> GAS
+#3 -> SURFACE
+#0 -> BCC
+#1 -> HCP
+#2 -> FCC
+
+for t in [120, 390, 890]:
+    pos = np.loadtxt('pos_%03d'%t, skiprows=1)
+    sl = (pos[:,-1]>100) & (pos[:,-1]<150)
+    pos = pos[sl]
+    phase = np.loadtxt('studyphase_%03d'%t, dtype=int, usecols=[1])[sl]
+    
+    f = File("sliceILM_t%03d.pov"%t,"colors.inc", "header.inc")
+    crystals = Union(*[
+        Sphere( (x,y,z), 5)
+        for x,y,z in pos[phase<3]
+        ]+[Texture(Pigment(color=(0.694,  0.145,  0.623)))])
+    f.write(crystals)
+    for i, c in zip([3,4,5], [(0.847,0.57,0.81), (0.847,0.57,0.81), (1.,  0.455,  0.156)]):
+        f.write(Union(*[
+            Sphere( (x,y,z), 5)
+            for x,y,z in pos[phase==i]
+            ]+[Texture(Pigment(color=c))
+            ]))
+    f.file.flush()
+    subprocess.call(['povray', '-O../presentation/sliceILM_t%03d.png'%t, '+W256', '+H256' , 'sliceILM_t%03d.pov'%t])

### 2016-06-24 crystal-gel/data/slicesILM.py
thinner slices (100<z<125) data from highcp sample to generate the slice through povray

@@ -8,23 +8,24 @@ import subprocess
 #1 -> HCP
 #2 -> FCC
 
-for t in [120, 390, 890]:
-    pos = np.loadtxt('pos_%03d'%t, skiprows=1)
-    sl = (pos[:,-1]>100) & (pos[:,-1]<150)
-    pos = pos[sl]
-    phase = np.loadtxt('studyphase_%03d'%t, dtype=int, usecols=[1])[sl]
-    
-    f = File("sliceILM_t%03d.pov"%t,"colors.inc", "header.inc")
-    crystals = Union(*[
-        Sphere( (x,y,z), 5)
-        for x,y,z in pos[phase<3]
-        ]+[Texture(Pigment(color=(0.694,  0.145,  0.623)))])
-    f.write(crystals)
-    for i, c in zip([3,4,5], [(0.847,0.57,0.81), (0.847,0.57,0.81), (1.,  0.455,  0.156)]):
-        f.write(Union(*[
+for sample, times in zip(['', '_highcp'], [[120, 390, 890], [800]]):
+    for t in times:
+        pos = np.loadtxt('pos%s_%03d'%(sample, t), skiprows=1)
+        sl = (pos[:,-1]>100) & (pos[:,-1]<125)
+        pos = pos[sl]
+        phase = np.loadtxt('studyphase%s_%03d'%(sample, t), dtype=int, usecols=[1])[sl]
+        
+        f = File("sliceILM%s_t%03d.pov"%(sample, t),"colors.inc", "header.inc")
+        crystals = Union(*[
             Sphere( (x,y,z), 5)
-            for x,y,z in pos[phase==i]
-            ]+[Texture(Pigment(color=c))
-            ]))
-    f.file.flush()
-    subprocess.call(['povray', '-O../presentation/sliceILM_t%03d.png'%t, '+W256', '+H256' , 'sliceILM_t%03d.pov'%t])
+            for x,y,z in pos[phase<3]
+            ]+[Texture(Pigment(color=(0.694,  0.145,  0.623)))])
+        f.write(crystals)
+        for i, c in zip([3,4,5], [(0.847,0.57,0.81), (0.847,0.57,0.81), (1.,  0.455,  0.156)]):
+            f.write(Union(*[
+                Sphere( (x,y,z), 5)
+                for x,y,z in pos[phase==i]
+                ]+[Texture(Pigment(color=c))
+                ]))
+        f.file.flush()
[…]

### 2016-06-24 talk_Roscoff.tex
lighter color for the liquid than for the crystal generate povray images with coherent color scheme that replace the call to nature/fig4.pdf
thinner slices (100<z<125) data from highcp sample to generate the slice through povray

@@
 		every axis/.style={
-		xlabel absolute, every axis x label/.append style={anchor=base, yshift=-1em},
-		ylabel absolute, every axis y label/.append style={anchor=base, yshift=-1em}
+		xlabel absolute, every axis x label/.append style={anchor=base, yshift=-1em, inner sep=0},
+		ylabel absolute, every axis y label/.append style={anchor=base, yshift=-1em, inner sep=0}
 		}]
@@
 	\end{tikzpicture}
-	\vspace{-0.5em}
+	%\vspace{-0.5em}
 	\begin{itemize}
@@
 			every axis/.style={
-			xlabel absolute, every axis x label/.append style={anchor=base, yshift=-1em},
-			%ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
+			xlabel absolute, every axis x label/.append style={anchor=base, yshift=-1em, inner sep=0},
+			ylabel absolute, every axis y label/.append style={anchor=base, yshift=-1em, inner sep=0}
 			}]
@@
 		\begin{axis}[
+			name=ax,
 			cycle list name=samples,
@@
 	\end{axis}
+	\let\ml\relax%
+	\newlength\ml%
+	\setlength{\ml}{0.25\textwidth}%
+	\matrix[matrix of nodes, inner sep=0, above left= 0 and -\textwidth of ax.outer south west, row sep=0.5em, column sep=0.3em,ampersand replacement=\&] (m){
+		\includegraphics[width=\ml]{presentation/highcpfinal_gaussian.jpg}\&
+		\includegraphics[width=\ml]{presentation/sliceILM_highcp_t800.png}\\
+		\includegraphics[width=\ml]{presentation/lowcpfinal_gaussian.jpg}\&
+		\includegraphics[width=\ml]{presentation/sliceILM_t890.png}\\
+	};
+	\matrix[
+	matrix of nodes, inner sep=0, column sep=0.2em, below left= 0.3em and 0 of m.south east, font=\scriptsize, ampersand replacement=\&,
+	every odd column/.style={circle, inner sep=0, minimum width= 0.6em, anchor=south west}
+	] {
+	|[ball color=ilmcolor]|\&crystal\&
+	|[ball color=ilmcolor!50!white]|\&liquid\&
+	|[ball color=ilmorange]|\&gas
+	\\};
 	\end{tikzpicture}
-	\includegraphics[height=11\baselineskip,clip=true,trim=0.5cm 0 0 0]{nature/fig2.pdf}
 
@@
 			xmin=0, xmax=13.0, ymin=0,
+			extra x ticks={2.11, 5.63, 12.13}, extra x tick style={grid=major}, extra x tick labels={},
 			no marks,
@@
 
-	\hfill\includegraphics[width=0.95\textwidth, clip=true, trim=0 20cm 0 0]{nature/fig4.pdf}\hfill
+	\let\imw\relax
+	\newlength\imw
+	\setlength{\imw}{0.29\linewidth}
+	\begin{scriptsize}
+	\begin{tabular}{ccc}
+	\includegraphics[width=\imw]{presentation/sliceILM_t120.png}&
+	\includegraphics[width=\imw]{presentation/sliceILM_t390.png}&
+	\includegraphics[width=\imw]{presentation/sliceILM_t890.png}\\
+	$t=2110\tau_B$ & $t=5630\tau_B$ & $t=12130\tau_B$\\
+	\end{tabular}
+	\end{scriptsize}
 \end{frame}
@@
 	\begin{tikzpicture}
-	\colorlet{gas}{green!40}%
-	\colorlet{crystal}{red}%
-	\colorlet{liquid}{lightgray}%
+		\colorlet{an}{black}
+		\colorlet{gas}{ilmorange!50!-an}%
+		\colorlet{crystal}{ilmcolor!90!-an}%
+		\colorlet{liquid}{ilmcolor!50!-an}%
 		\node[regular polygon, regular polygon sides=3, minimum width=5em] (tri){};
@@
 		\item A particle level colloidal analogue to the Bergeron process
-		\item Generic "crystal-gel" process
+		\item Generic ``crystal-gel'' process
 		\begin{enumerate}

### 2016-07-09 crystal-gel/talk_statphys.tex
12 min version for statphys. More stremlined intro, no interruption for citations of other instances of crystallised networks

--- 2016-06-24 talk_Roscoff.tex
@@
 
-\institute[iLM]{iLM, Univ. Claude Bernard Lyon 1 and CNRS, France}
+\institute[iLM]{Univ Lyon, Université Claude Bernard Lyon 1, CNRS, Institut Lumière Matière}
 \title[crystal gel]{A novel route to the spontaneous formation of porous crystals via viscoelastic phase separation}
-\author[M. Leocmach]{Mathieu Leocmach}
-\date{30 June 2016}
+\author[M.
+Leocmach]{Mathieu Leocmach \hfill{\usebeamerfont{normal text}\texttt{\usebeamercolor[fg]{normal text}\footnotesize @LamSonLeoc\includegraphics[height=1em]{Twitter_Bird}}}\vspace{-\baselineskip}}
+\date{18 July 2016}
 \titlegraphic{
@@
 
-\begin{frame}{Crystallisation}
-	\tikzsetnextfilename{whyx}%
-	\input{presentation/whyx.pgf}
+\begin{frame}<1>[label=intro]{How does it crystallise?}
+	\begin{description}[Understand]
+		\item[Understand] natural phenomena
+		\item[Engineer] technological applications
+	\end{description}
 
-	\begin{block}{Nucleation}
+	\begin{block}{What determines the ensemble properties of crystallites?}
+	\vspace{-0.5\baselineskip}\begin{columns}
+	\column{0.45\textwidth}%
 		\begin{itemize}
-		\item 10-1000 molecules $\rightarrow$ difficult to observe
-		\item classical nucleation theory (CNT): fluid $\xrightarrow{\text{1 step}}$ crystal
+			\item polymorph type
+			\item shape
 		\end{itemize}
+	\column{0.45\textwidth}
+	\begin{itemize}
+			\item spatial arrangement
+			\item size distribution
+		\end{itemize}
+	\end{columns}
 	\end{block}
+
+	\begin{itemize}
+		\item Not always simple nucleation and growth scenario
+		\begin{itemize}
+			\item<2> Example: Bergeron process
+		\end{itemize}
+		\item<2> \structure{More detailed view of the processes needed}
+	\end{itemize}
 \end{frame}
@@
 \end{frame}
+
+\againframe<2>{intro}
 
@@
 
-\begin{frame}{Another class of gel? Crystal-gel?}
-	\begin{itemize}
-	\item Phase separation arrested by crystallisation
-	\begin{description}
-		\item[polymer blend] Tanaka \& Nishi \textit{PRL} 1985
-		\item[simulations] Soga \textit{J.Chem. Phys.} 1999; Fortini \textit{PRE} 2008 
-	\end{description}
-	\item Spinodal then crystallisation
-	\begin{description}
-		\item[simulations] Pérez \textit{Langmuir} 2011
-		\item[microgravity] (triple coexistence) Sabin \textit{PRL} 2012
-	\end{description}
-	\end{itemize}
-	\includegraphics[width=\textwidth]{presentation/micrograv_Sabin2012}
-\end{frame}
-
 \begin{frame}{Crystals outgrow metastable liquid}
@@
 \begin{frame}[plain]
+\end{frame}
+
+\begin{frame}{Another class of gel? Crystal-gel?}
+	\begin{itemize}
+	\item Phase separation arrested by crystallisation
+	\begin{description}
+		\item[polymer blend] Tanaka \& Nishi \textit{PRL} 1985
+		\item[simulations] Soga \textit{J.Chem. Phys.} 1999; Fortini \textit{PRE} 2008 
+	\end{description}
+	\item Spinodal then crystallisation
+	\begin{description}
+		\item[simulations] Pérez \textit{Langmuir} 2011
+		\item[microgravity] (triple coexistence) Sabin \textit{PRL} 2012
+	\end{description}
+	\end{itemize}
+	\includegraphics[width=\textwidth]{presentation/micrograv_Sabin2012}
 \end{frame}

### 2016-07-19 crystal-gel/nature/Supplementary.tex

@@
 \centerline{\bf \large Supplementary Information for}
-\centerline{\bf \large``A novel route to the spontaneous formation of porous crystals via}
-\centerline{\bf \large viscoelastic phase separation''}
+\centerline{\bf \large``A novel physical mechanism for the spontaneous formation }
+\centerline{\bf \large of porous crystals via viscoelastic phase separation''}
 \vspace{0.5cm}
@@
 \section*{Supplementary Movie 2}
-This movie shows the late stage after Supplemental Movie 1 with scanning duration of 27000~s (12300~$\tau$). 
-Please note that there is a large difference in the frame rate between Movie 1 and 2.
+This movie shows the late stage after Supplemental Movie 1 with scanning duration of 27000~s (12300~$\tau$). Please note that there is a large difference in the frame rate between Movie 1 and 2.
+
+\section*{Supplementary Movie 3}
+Same as Supplementary Movie 2, but with computer rendered trajectories obtained from the tracking of the colloidal particles.
+Particles are coloured according to the following code: crystal (red), surface (blue), gas (green) and liquid (white).
+For display purposes, crystalline particles and gas particles are respectively 400\% and 50\% bigger than other particles (liquid or surface).
 

### 2016-07-19 crystal-gel/nature/crystal_gel3.tex

--- 2016-03-24 crystal-gel/nature/crystal_gel2.tex
@@
 
-\title{A novel route to the spontaneous formation of porous crystals via viscoelastic phase separation}
+\title{A novel physical mechanism for the spontaneous formation of porous crystals via viscoelastic phase separation}
 
@@
 
-\date{Received November 26, 2015}
+\date{Received July 20, 2016}
 
@@
 %the glass transition, with the formation of colloidal gels~\cite{poon2002,zaccarelli2007,piazza1994phase,verhaegh1997transient,tanaka1999colloid,lu2008gelation}, or by crystallization, which results in the formation of a crystalline network~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways,sabin2012}.
-glass transition~\cite{poon2002,zaccarelli2007,piazza1994phase,verhaegh1997transient,tanaka1999colloid,lu2008gelation} or by crystallization~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways,sabin2012}.
+glass transition~\cite{poon2002,zaccarelli2007,piazza1994phase,verhaegh1997transient,tanaka1999colloid,lu2008gelation} or by crystallization~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways,sabin2012,zhang2012non}.
 %But while the former process has been studied in detail, the latter has never been observed experimentally at a single-particle level and it is thus lacking a microscopic understanding. 
-With a new confocal microscopy protocol, we unveil a novel crystallization pathway to sponge-like porous structures in a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation.
-We show that crystallization leads to a network structure different from the original phase separation pattern.
+With a new confocal microscopy protocol, we unveil a novel crystallization pathway to sponge-like porous crystal structures in a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation.
+We show that this crystallization pathway leads to a network structure different from the original phase separation pattern.
 Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface, i.e.
@@
 Indeed the possibility of a different class of gels, which are stabilized by crystallization,
-was suggested by numerical simulations~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways} and observed in microgravity experiments \cite{sabin2012}. 
+was suggested by numerical simulations~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways} and observed in experiments \cite{sabin2012,zhang2012non}. 
 For colloidal systems, we can in principle access structural evolution in real time at a single-particle level; however, there has so far been no confocal microscopy  
-study on the dynamical process of crystal-gel formation. 
+studies on the dynamical process of crystal-gel formation on a microscopic level, which makes the microscopic mechanism of crystal-gel formation elusive. 
 For confocal microscopy experiments on colloid gelation, colloid-polymer mixtures have so far been used as a model system. 
@@
 Differently from the standard crystal-gel scenario, where crystals only nucleate and grow inside the dense branches of the gel, becoming dynamically arrested, we show that the kinetic
-pathways involving the gas phase (predominantly the Bergeron process, but also Ostwald ripening) play a fundamental role. We also observe that increasing the strength of the interparticle attraction
+pathways involving the gas phase (predominantly the Bergeron process, but also Ostwald ripening) play a fundamental role. 
+We reveal that the final crystal-gel network structure has smoother interface than the ordinary gels formed by vitrification because of this novel crystal growth mechanism.  
+We also observe that increasing the strength of the interparticle attraction
 (by increasing the polymer concentration), instead of forming crystal gels, the system undergoes the classical gelation process, in which
@@
 All samples share the same early stages of spinodal decomposition. 
-A typical early stage phase ordering process observed at $\phi\approx 0.33$ and $c_p=0.38$ mg/g can be seen in Supplementary Movie 1.. 
+A typical early stage phase ordering process observed at $\phi\approx 0.33$ and $c_p=0.38$ mg/g can be seen in Supplementary Movie 1.  
 Due to strong dynamical asymmetry between colloids and the solvent \cite{tanaka1999colloid}, the colloidal particles start aggregating, 
@@
 In Fig.~\ref{fig:phase_separation}a, the state points which did not phase separate are indicated with open triangles, while
-the other symbols indicate samples undergoing phase separation. \textcolor{blue}{The black dashed line is the gas spinodal computed from free volume theory and the 
+the other symbols indicate samples undergoing phase separation. The black dashed line is the gas spinodal computed from free volume theory and the 
 black continuous line is the binodal line. They are in good agreement with our experimental points (the small discrepancy at small colloidal volume fractions 
-is common in free volume theory)~\cite{Royall2007,lu2008gelation}. }
+is common in free volume theory)~\cite{Royall2007,lu2008gelation}.
 The topology of the network is similar to all our samples at the initial stages. 
@@
 
-\textcolor{blue}{To gain additional insight, in Supplementary Figure~\ref{S-fig:local_phi12} we plot the late stage (after all samples already underwent gas-liquid phase separation) time evolution of the local volume fraction of closed packed particles (having 12 neighbours).
+To gain additional insight, in Supplementary Figure~\ref{S-fig:local_phi12} we plot the late stage (after all samples already underwent gas-liquid phase separation) time evolution of the local volume fraction of closed packed particles (having 12 neighbours).
 The volume fraction is obtained by computing the Voronoi diagram of each configuration, which uniquely assigns a volume to each colloidal particle.
@@
 lower volume fraction, which is indeed close to the volume fraction of the attractive glass state~\cite{pham2002multiple}.
-This result clearly indicates that a single state point follows a different arrest mechanism, in which phase separation is arrested by crystallization and not by glassiness.}
+This result clearly indicates that a single state point follows a different arrest mechanism, in which phase separation is arrested by crystallization and not by glassiness.
 
@@
 More quantitatively, we adopt bond orientational analysis to detect which particles are in crystalline environments~\cite{russo2013interplay}
-\textcolor{blue}{(see Supplementary Information on the detail)}. 
+(see Supplementary Information on the detail). 
 With this analysis we reveal that crystallization events indeed start occurring for the state point $\phi\approx 0.33$ and $c_p=0.38$ mg/g, which has the lowest
@@
 Supplementary Movie 3 shows a three-dimensional (3D) reconstruction  of the whole phase ordering sequence that emphasizes crystalline particles.
+Although crystal clusters are isolated in this movie, they are bridged by the dense colloidal liquid phase and involved in the percolated gel network, as can be seen in Fig.
+\ref{fig:network}b.
+We can see that crystal clusters grow by condensation from the colloidal gas phase. 
 %The average size of the crystals shows an interesting linear growth regime for $100<t<600$. This linear growth regime goes past the time
@@
 
-\textcolor{blue}{Next we analyse in detail the crystallization trajectory, distinguishing between the different processes responsible for the
+Next we analyse in detail the crystallization trajectory, distinguishing between the different processes responsible for the
 crystallization of the phase separated network.
@@
 This suggests that the crystal network does not form only by direct freezing of the liquid, but also by other mechanisms which involve the evaporation of the liquid (Bergeron process) and the sublimation of small crystals (Ostwald ripening).
-In the following we will investigate these different crystallization mechanisms in depth.}
+In the following we will investigate these different crystallization mechanisms in depth.
 
-\textcolor{blue}{
 In Fig.~\ref{fig:transitions}b, we show  the fraction of particles in each different phase for the crystallization trajectory at
@@
 transition to the gas state before crystallizing. Ostwald ripening accounts for 11\% of particle trajectories in which crystal particles transition to the gas state before returning
-to the crystal state. The different crystallization channels are depicted in the diagram of Fig.~\ref{fig:transitions}c.}
+to the crystal state. The different crystallization channels are depicted in the diagram of Fig.~\ref{fig:transitions}c.
 
-\textcolor{blue}{
 In Supplementary Information we confirm these results with a detailed analysis of transition probabilities, including a state for surface particles that are in contact 
@@
 especially in determining the morphology of the porous crystal, as we discussed above. In this context the Bergeron process~\cite{glickman2000glossary,morrison2012resilience}
-plays a considerably more important role than Ostwald ripening.}
+plays a considerably more important role than Ostwald ripening.
 
@@
 This can be seen, for example, by looking at the phase diagram of a Lennard-Jones liquid \cite{lodge1997brownian}, which represents many molecular systems 
-without specific directional interactions. \textcolor{blue}{For monoatomic systems such as noble metals, condition (iii) may not be satisfied easily. 
+without specific directional interactions. For monoatomic systems such as noble metals, condition (iii) may not be satisfied easily. 
 To access glassy slow dynamics in a liquid phase, we may need a deep quench at a high pressure. However, we note that even without strong dynamic asymmetry 
 due to glassiness, bicontinuous phase separation can take place between 35-85 volume \% of the liquid phase in ordinary gas/liquid phase separation \cite{Onuki2002}; 
-and, thus, porous crystalline structures can be formed, although a thin network structure may be difficult to be formed.}
+and, thus, porous crystalline structures can be formed, although a thin network structure may be difficult to be formed.
 
@@
 
-\textcolor{blue}{Matching phase boundaries often leads to better determination of volume fractions than colloid diameter measurements~\cite{Poon2012}.
+Matching phase boundaries often leads to better determination of volume fractions than colloid diameter measurements~\cite{Poon2012}.
 We obtain the best match between theoretical phase diagram and experimental data for $\sigma=2.25\,\mu$m and $R_g=80$ nm (see Supplementary method).
-Therefore the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.071$ and the overlap mass fraction of polymer $2.25$~mg/g.}
+Therefore the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.071$ and the overlap mass fraction of polymer $2.25$~mg/g.
 
@@
 
-\textcolor{blue}{
 The crystal-gel state is reproduced in a sealed cell with all the components of the suspension pre-mixed.
 The colloid volume fraction is virtually identical at $\approx 0.33$ and polymer concentration is $0.40\,$mg/g.
-See Supplementary Figure~\ref{S-fig:capillary}.}
+See Supplementary Figure~\ref{S-fig:capillary}.
 
@@
 The other symbols are the experimental points showing gel behaviour, consistently used in all panels.
-\textcolor{blue}{The dotted lines are the gas-liquid tie lines determined experimentally, with arrow heads indicating measured gas and liquid compositions (see Supplementary Information on the procedure to extract the compositions from confocal images).
+The dotted lines are the gas-liquid tie lines determined experimentally, with arrow heads indicating measured gas and liquid compositions (see Supplementary Information on the procedure to extract the compositions from confocal images).
 Gray areas are theoretical equilibrium fluid and crystal regions.
-The solid and dashed lines are the metastable gas-liquid binodal and spinodal respectively.}
+The solid and dashed lines are the metastable gas-liquid binodal and spinodal respectively.
 {\bf b,} Radius of gyration ($R_g$) as a function of cluster size ($s$) for all state points at percolation. The
@@
 {\bf c,} Structure factors for the late stages of the gelation process. The gray and black lines represents respectively random gelation and compact fractal dimensions.
-\textcolor{blue}{{\bf d,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle for the state points in the very late stage. 
-The two horizontal straight lines indicate the characteristic volume fractions in attractive systems, crystal ($\sim 0.72$) and glass ($\sim 0.64$). }
+{\bf d,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle for the state points in the very late stage. 
+The two horizontal straight lines indicate the characteristic volume fractions in attractive systems, crystal ($\sim 0.72$) and glass ($\sim 0.64$). 
  }
@@
  % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
-\caption{\textcolor{blue}{{\bf Percolated network structures.} {\bf a:} gel; {\bf b:} crystal-gel. The gel state is obtained from the state
+\caption{{\bf Percolated network structures.} {\bf a:} gel; {\bf b:} crystal-gel. The gel state is obtained from the state
 point $\phi\approx 0.33$ and $c_p=1.07$ mg/g at $t=400$ min after sample preparation, while the crystal-gel is obtained at the same $\phi$ but
@@
 We note that very few liquid particles remain and the network is almost crystalline. 
-}} 
+} 
 \label{fig:network}
@@
 {\bf b,} Radius of gyration of crystalline nuclei at the late stages of gelation for the same state point.
-\textcolor{blue}{The gray and black straight lines have a slope of $1/2.53$ and $1/3$ respectively}} 
+The gray and black straight lines have a slope of $1/2.53$ and $1/3$ respectively} 
  \label{fig:crystals}
@@
  % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735
- \caption{\textcolor{blue}{{\bf Crystal gel formation.}  
+ \caption{{\bf Crystal gel formation.}  
 {\bf a,} Reconstructions from confocal coordinates at $\phi\approx 0.33$ and $c_p=0.38$ mg/g.
@@
 transition to the gas state before crystallizing. Ostwald ripening (dashed arrow) accounts for 11\% of particle trajectories in which crystal particle transition to the gas state before returning
-to the crystal state.}}
+to the crystal state.}
 \label{fig:transitions}

### 2016-12-12 crystal-gel/nature/crystal_gel3.tex
Merge John's edits: - added temporarily some sections, so it's easier for us to navigate and understand the flow of the manuscript (we can remove the section names before submitting). - added fig:bond_breaking - insist on the role of initial rearrangements of the network to make nucleation possible

@@
 %But while the former process has been studied in detail, the latter has never been observed experimentally at a single-particle level and it is thus lacking a microscopic understanding. 
-With a new confocal microscopy protocol, we unveil a novel crystallization pathway to sponge-like porous crystal structures in a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation.
-We show that this crystallization pathway leads to a network structure different from the original phase separation pattern.
+With a new confocal microscopy protocol, we unveil a novel crystallization pathway to sponge-like porous structures in a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation.
+\textcolor{blue}{
+We follow the entire kinetics of phase separation, from the homogeneous phase, percolation, stress driven aging, and until the formation of porous crystals.}
+We show that crystallization leads to a network structure different from the original phase separation pattern.
 Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface, i.e.
@@
 \maketitle
+
+\subsection{Introduction}
 
@@
 
+\textcolor{blue}{
 In this Letter, we show that, for low polymer concentration, the system can undergo a novel crystallization pathway that leads to the formation
 of crystalline droplets.
-These droplets originate from crystal seeds that form inside the liquid branches of the spinodal aggregate of network morphology, and then grow by direct condensation from the gas phase.
-The process is analogous to the Bergeron process of ice formation in mixed phase clouds, where ice droplets grow at the
-expense of the supercooled liquid droplets due to their lower saturated vapour pressure. 
-This is speculated to be the primary mechanism for the formation of raindrops. 
-Differently from the standard crystal-gel scenario, where crystals only nucleate and grow inside the dense branches of the gel, becoming dynamically arrested, we show that the kinetic
-pathways involving the gas phase (predominantly the Bergeron process, but also Ostwald ripening) play a fundamental role. 
+These droplets originate from crystal seeds that form inside the liquid branches of the spinodal aggregate of network morphology, and then grow with three different growth channels.
+(i) growth within the dense branches of the gel ;
+(ii) Bergeron process, analogous to ice formation in mixed phase clouds, where ice droplets grow at the
+expense of the supercooled liquid droplets due to their lower saturated vapour pressure ;
+(iii) Ostwald ripening.
+Differently from the standard crystal-gel scenario, where crystals only nucleate and grow inside the dense branches of the gel, becoming dynamically arrested, we show that the kinetic pathways involving the gas phase (predominantly the Bergeron process, but also Ostwald ripening) play an important role.
+We show that a necessary condition for the formation of crystal gels is the stress-driven aging of the network, that is possible only at low polymer concentration, and which releases the stresses
+that build-up during the network formation due to hydrodynamic interactions, allowing for an increase in the number of average neighbours.
+If the average concentration of particles is high enough, this relaxation triggers the nucleation process.
 We reveal that the final crystal-gel network structure has smoother interface than the ordinary gels formed by vitrification because of this novel crystal growth mechanism.  
-We also observe that increasing the strength of the interparticle attraction
+On the other side, we also observe that increasing the strength of the interparticle attraction
 (by increasing the polymer concentration), instead of forming crystal gels, the system undergoes the classical gelation process, in which
 phase separation is interrupted by vitrification~\cite{verhaegh1997transient,tanaka1999colloid,Royall2007,lu2008gelation}. 
+}
+
+\subsection{Samples}
 
@@
 
+\subsection{Early times: percolation}
+
 As described above and also in Methods, salt injection initiates liquid-gas phase separation of the colloidal suspension. 
@@
 
-%%%%
-%Previous simulation results~\cite{tanaka2000,tanaka2007spontaneous,furukawa2010key} have suggested that hydrodynamic interactions play a key role in the process of colloidal gelation.
-Without hydrodynamic interactions, particles have the tendency to aggregate in compact structures and subsequently form thick network structures.
+\subsection{Intermediate times: stress-driven aging}
+
+\textcolor{blue}
+{
+At intermediate times, after percolation has started, hydrodynamic interactions play a fundamental role~\cite{tanaka2000,tanaka2007spontaneous,furukawa2010key}.
+Without hydrodynamic interactions, as often assumed in simulations, particles have the tendency to aggregate in compact structures and subsequently form thick network structures.
 This also limits the possibility to form arrested gel states only at relatively high volume fractions.
@@
 Thus, hydrodynamic interactions lead to the formation of gels that are very far from equilibrium and under a strong thermodynamic driving force towards more stable compact structures.
-The resulting transition from open to more compact networks occurs only through the breaking of the bonds that have accumulated more stress.
-The process was called stress-driven ageing~\cite{
-%tanaka2007spontaneous}, which is characteristic of viscoelastic phase separation \cite{tanaka2000viscoelastic}. Below we provide the first experimental proof of this process.
+The resulting transition from open to more compact networks occurs through the breaking of the bonds that have accumulated more stress.
+The process was called stress-driven ageing~\cite{tanaka2007spontaneous}, and is characteristic of viscoelastic phase separation \cite{tanaka2000viscoelastic}.
+Whether the network undergoes such restructuring or not depends on the strength of the bonds, i.e.
+the polymer concentration.
+In Fig.~\ref{fig:bond_breaking} we plot the bond breaking probability, $p_\text{bond}$, and the average coordination, $\bar{n}$, as a function of the time elapsed since salt injection ($t=0$).
+At high polymer concentrations (panels \textbf{a}, \textbf{b}), $p_\text{break}$ decreases monotonically, until a stationaty state is reached where $\bar{n}<5$.
+At low polymer concentrations (panels \textbf{c}, \textbf{d}), the initial decay of $p_\text{break}$ to a network with $\bar{n}<5$ is followed by a secondary decay to a more compact network with $\bar{n}>6$.
+The build up of mechanical stress in the first transient network can relax to a more compact network only when when bonds are weak enough, i.e.
+at low polymer concentrations.
+}
 
+\textcolor{blue}{
+It is worth noting that the two-step beheviour of $p_\text{break}$ and $\bar{n}$ is not due to crystallization, as this only starts after the network reorganization.
+For example, the fraction of crystalline particles for all frames in Fig.~\ref{fig:bond_breaking}d is always below $0.3\%$.
+The increase in number of bonds after network reorganization is a necessary condition for crystallization, as we will show in the following, while in its absence the network forms low-density arrested states (gels).
+ }
 %In Fig.~\ref{fig:stress}a, we plot the fraction of liquid and gas particles as a function of time during phase separation at the state points of $\phi\approx 0.13$.
@@
 %%%
+\subsection{Late times: dynamic-arrest vs crystallization}
 
-While sharing similar structural properties at early times, the samples show significantly different behaviours in the later stages of the phase separation process.
+The samples show significantly different behaviours in the later stages of the phase separation process.
 To examine this in more detail, we plot in Fig.~\ref{fig:phase_separation}c the structure factor for the final stages of the gelation process for all state points.
@@
 
-While nucleation always occurs inside the liquid branches of the phase-separating network, we suggest here the presence of a new growth channel
-which was neither considered nor observed before.
+\subsection{Crystal-growth channels}
+
+While nucleation always occurs inside the liquid branches of the phase-separating network, crystal-growth can proceed through different growth channels.
 In the following we will show that the growth mechanism of the crystal is not only due to filling of the spinodal liquid network, as was observed in Refs.~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways} for Brownian Dynamics simulations, but also involves
-evaporation of the liquid phase and subsequent direct condensation of the gas phase on the crystalline seeds. 
-To confirm the presence of the two processes, we plot in Fig.~\ref{fig:crystals}b the gyration radius of individual crystalline nuclei for the state point 
+The Bergeron process and Ostwald ripening. 
+In Fig.~\ref{fig:crystals}b the gyration radius of individual crystalline nuclei for the state point 
 $\phi\approx 0.33$ and $c_p=0.38$ mg/g. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with 
@@
 (iii) dynamical slowing down in a supercooled liquid state, which is necessary to induce viscoelastic phase separation leading to the formation 
-of a network structure of the minority liquid phase, and (iv) the degree of supercooling is low enough to avoid a high rate of densification of the liquid phase 
-leading to vitrification.  
+of a network structure of the minority liquid phase, and (iv) the degree of supercooling is low enough
+\textcolor{blue}
+{
+to allow bond-breaking events that avoid the vitrification of the liquid phase.  
+}
 Many monoatomic and single-component molecular systems can satisfy all these conditions in a certain range of the temperature and pressure. 
@@
 
+\begin{figure}
+ \centering
+ \includegraphics[width=16cm]{./bond_breaking_figure.pdf}
+ % bond_breaking_figure.pdf: 1553x1066 pixel, 72dpi, 54.79x37.61 cm, bb=0 0 1553 1066
+ \caption{{\bf Stress driven aging.} Time evolution of the average coordination of colloidal particles $\bar{n}$ (black continuous line) and the bond-breaking probability $p_\text{break}$ (red dashed line) for state points $(\phi;c_p)$: {\bf a} $(0.14;1.36)$, {\bf b} $(0.33;1.07)$, {\bf c} $(0,14;0.82)$, {\bf d} $(0.33;0.38)$.
+$p_\text{break}$ is measured as the probability of bonds breaking in a time interval of $\Delta t=10\,s$.
+For high polymer concentration, panels {\bf a} and {\bf b}, $p_\text{break}$ decreases monotonically until it reaches a stationary state at long times, with $\bar{n}$ saturating at an average of less than $5$ neighbours.
+For low polymer concentration, panels {\bf c} and {\bf d}, $p_\text{break}$ follows a similar decay at intermediate times, with $\bar{n}$ less than $5$ neighbours, but which is then followed by a second decay to new configuration, where $\bar{n}$ becomes greater than $5$ neighbours.}
+ \label{fig:bond_breaking}
+\end{figure}
+
+\clearpage
+
 \begin{figure*}[!t]

### 2016-12-12 crystal-gel/nature/crystal_gel3.tex
State more clearly what we show in the paper, in the order of gelation and crystallisation process

@@
 \usepackage{graphicx,amsmath}
-\usepackage{color}
+\usepackage{xcolor}
 \usepackage{xr}
@@
 
-Crystallization plays a fundamental role in many processes occurring in nature, such as cloud composition, and in technological applications
+Crystallization plays a fundamental role in many processes occurring in nature, such as \textcolor{blue}{ice formation in atmospheric clouds}, and in technological applications
 that are at the core of the chemical, pharmaceutical, and food industries.
@@
 
-Under usual conditions, the nucleation stage involves a very small number of molecules, of the order of 10-1000 molecules.
+\textcolor{brown}{Mathieu: When I presented this story, I was often told that our initial focus on nucleation was misleading since our main result is about growth.
+What about trimming down this part?
+Or focussing more on growth rather than nucleation?
+}Under usual conditions, the nucleation stage involves a very small number of molecules, of the order of 10-1000 molecules.
 This has severely limited the possibility to observe the nucleation process directly, and has also posed limits on
@@
 
-Unlike the above standard scenario of colloidal gelation, it is known that phase separation can also be arrested by crystallization, if the process takes place below the melting point of a component of a mixture~\cite{tanaka1985new}.
+Unlike the above standard scenario of colloidal gelation, it is known that phase separation can also be arrested by crystallization, if the process takes place below the melting point of \textcolor{blue}{one of the two separating phases}~\cite{tanaka1985new}.
 Indeed the possibility of a different class of gels, which are stabilized by crystallization,
@@
 For confocal microscopy experiments on colloid gelation, colloid-polymer mixtures have so far been used as a model system. 
-The main difficulty consists in the preparation protocol of such a colloid-polymer mixture, where before observation of its dynamics shaking or applying shear is necessary to prepare an initial homogeneous state.
+The main difficulty consists in the preparation protocol of such a colloid-polymer mixture where, before observation of its dynamics, shaking or applying shear is necessary to prepare an initial homogeneous state.
 This introduces turbulent flows at the beginning of the process, and also does not allow the
-observation of the initial stages of crystallization. This initial perturbation may even alter the selection of the final non-equilibrium arrested state.  
+observation of the initial stages of crystallization \textcolor{brown}{Is it crystallization of gelation here in this context?}.
+This initial perturbation may even alter the selection of the final non-equilibrium arrested state.
 In our experiments we instead succeed in building a novel protocol that allows the time evolution of the system to be observed directly
@@
 \textcolor{blue}{
-In this Letter, we show that, for low polymer concentration, the system can undergo a novel crystallization pathway that leads to the formation
-of crystalline droplets.
-These droplets originate from crystal seeds that form inside the liquid branches of the spinodal aggregate of network morphology, and then grow with three different growth channels.
+In this Letter, we show that the system can undergo a novel crystallization pathway that leads to the formation of interconnected crystalline droplets. 
+These droplets originate from crystal seeds that form inside the liquid branches of the spinodal aggregate of network morphology.
+We show that a necessary condition for the nucleation of crystal seeds is the stress-driven aging of the network which releases the stresses that build-up during the network formation due to hydrodynamic interactions, allowing for an increase in the number of average neighbours.
+If the average concentration of particles is high enough, this relaxation triggers the nucleation process. 
+Stress-driven aging is possible only at low polymer concentration. 
+Indeed, at higher polymer concentration (stronger interparticle attraction) we observe that, instead of forming crystal gels, the system undergoes the classical gelation process, in which
+phase separation is interrupted by vitrification~\cite{verhaegh1997transient,tanaka1999colloid,Royall2007,lu2008gelation}.
+We then show that the growth of the crystal seeds occurs by three different routes:
 (i) growth within the dense branches of the gel ;
@@
 Differently from the standard crystal-gel scenario, where crystals only nucleate and grow inside the dense branches of the gel, becoming dynamically arrested, we show that the kinetic pathways involving the gas phase (predominantly the Bergeron process, but also Ostwald ripening) play an important role.
-We show that a necessary condition for the formation of crystal gels is the stress-driven aging of the network, that is possible only at low polymer concentration, and which releases the stresses
-that build-up during the network formation due to hydrodynamic interactions, allowing for an increase in the number of average neighbours.
-If the average concentration of particles is high enough, this relaxation triggers the nucleation process.
 We reveal that the final crystal-gel network structure has smoother interface than the ordinary gels formed by vitrification because of this novel crystal growth mechanism.
-On the other side, we also observe that increasing the strength of the interparticle attraction
-(by increasing the polymer concentration), instead of forming crystal gels, the system undergoes the classical gelation process, in which
-phase separation is interrupted by vitrification~\cite{verhaegh1997transient,tanaka1999colloid,Royall2007,lu2008gelation}. 
 }
@@
 Without hydrodynamic interactions, as often assumed in simulations, particles have the tendency to aggregate in compact structures and subsequently form thick network structures.
-This also limits the possibility to form arrested gel states only at relatively high volume fractions.
-With hydrodynamic interactions, particles first form a transient gel even at very low volume fractions, and only later they increase the number of nearest neighbours to minimize the energy of the structure.
+This also limits to relatively high volume fractions the possibility to form arrested gel states.
+With hydrodynamic interactions, particles first form a transient gel even at very low volume fractions, and the number of nearest neighbours increases only later to minimize the energy of the structure.
 Thus, hydrodynamic interactions lead to the formation of gels that are very far from equilibrium and under a strong thermodynamic driving force towards more stable compact structures.
 The resulting transition from open to more compact networks occurs through the breaking of the bonds that have accumulated more stress.
-The process was called stress-driven ageing~\cite{tanaka2007spontaneous}, and is characteristic of viscoelastic phase separation \cite{tanaka2000viscoelastic}.
+This process has been called stress-driven ageing~\cite{tanaka2007spontaneous}, and is characteristic of viscoelastic phase separation \cite{tanaka2000viscoelastic}.
 Whether the network undergoes such restructuring or not depends on the strength of the bonds, i.e.
@@
 In Fig.~\ref{fig:bond_breaking} we plot the bond breaking probability, $p_\text{bond}$, and the average coordination, $\bar{n}$, as a function of the time elapsed since salt injection ($t=0$).
-At high polymer concentrations (panels \textbf{a}, \textbf{b}), $p_\text{break}$ decreases monotonically, until a stationaty state is reached where $\bar{n}<5$.
+At high polymer concentrations (panels \textbf{a}, \textbf{b}), $p_\text{break}$ decreases monotonically, until a stationary state is reached where $\bar{n}<5$.
 At low polymer concentrations (panels \textbf{c}, \textbf{d}), the initial decay of $p_\text{break}$ to a network with $\bar{n}<5$ is followed by a secondary decay to a more compact network with $\bar{n}>6$.
-The build up of mechanical stress in the first transient network can relax to a more compact network only when when bonds are weak enough, i.e.
+This means that mechanical stress built up in the first transient network can relax to a more compact network only when bonds are weak enough, i.e.
 at low polymer concentrations.
@@
  \centering
- \includegraphics[width=16cm]{./bond_breaking_figure.pdf}
+ %\includegraphics[width=16cm]{./bond_breaking_figure.pdf}
  % bond_breaking_figure.pdf: 1553x1066 pixel, 72dpi, 54.79x37.61 cm, bb=0 0 1553 1066

### 2016-12-12 journee_nucleation.tex
présentation pour la journée nucléation

--- 2016-07-09 crystal-gel/talk_statphys.tex
@@
+\pdfminorversion=4
 \documentclass[xcolor=table]{beamer}
@@
 
-\institute[iLM]{Univ Lyon, Université Claude Bernard Lyon 1, CNRS, Institut Lumière Matière}
+\institute[iLM]{Equipe Liquides et Interfaces}%{Univ Lyon, Université Claude Bernard Lyon 1, CNRS, Institut Lumière Matière}
 \title[crystal gel]{A novel route to the spontaneous formation of porous crystals via viscoelastic phase separation}
 \author[M.
-Leocmach]{Mathieu Leocmach \hfill{\usebeamerfont{normal text}\texttt{\usebeamercolor[fg]{normal text}\footnotesize @LamSonLeoc\includegraphics[height=1em]{Twitter_Bird}}}\vspace{-\baselineskip}}
-\date{18 July 2016}
+Leocmach]{Mathieu Leocmach \hfill{\usebeamerfont{normal text}\texttt{\usebeamercolor[fg]{normal text}\footnotesize @LamSonLeoc\includegraphics[height=1em]{Twitter_Bird}}}\vspace{-0.5\baselineskip}}
+\date{16 Nov 2016}
 \titlegraphic{
@@
 		\includegraphics[width=\ml]{presentation/highcpfinal_gaussian.jpg}\&
-		\includegraphics[width=\ml]{presentation/sliceILM_highcp_t800.png}\\
+		\includegraphics[width=\ml]{presentation/sliceILM_highcp_t800.jpg}\\
 		\includegraphics[width=\ml]{presentation/lowcpfinal_gaussian.jpg}\&
-		\includegraphics[width=\ml]{presentation/sliceILM_t890.png}\\
+		\includegraphics[width=\ml]{presentation/sliceILM_t890.jpg}\\
 	};
@@
 	\begin{tabular}{ccc}
-	\includegraphics[width=\imw]{presentation/sliceILM_t120.png}&
-	\includegraphics[width=\imw]{presentation/sliceILM_t390.png}&
-	\includegraphics[width=\imw]{presentation/sliceILM_t890.png}\\
+	\includegraphics[width=\imw]{presentation/sliceILM_t120.jpg}&
+	\includegraphics[width=\imw]{presentation/sliceILM_t390.jpg}&
+	\includegraphics[width=\imw]{presentation/sliceILM_t890.jpg}\\
 	$t=2110\tau_B$ & $t=5630\tau_B$ & $t=12130\tau_B$\\

### 2016-12-12 crystal-gel/nature/Supplementary.tex
Update colloid size from knowledge of actual pixel size. Impacts the main text, supplementary, phase diagram and local volume fraction

@@
 Our two constraints are (i) to have the final local volume fraction of the 12-neighboured particles in the crystallising sample to be equal to the theoretical equilibrium crystal volume fraction and (ii) to have the theoretical spinodal line to correspond to the gel region boundary.
-Starting from our initial estimates of $\sigma$ and $R$ we iteratively converge to $R=80$~nm, $\sigma = 2.25\,\mu$m and a crystal (without defects) at $0.723$.
+Starting from our initial estimates of $\sigma$ and $R$ we iteratively converge to $R=80$~nm, $\sigma = 2.21\,\mu$m and a crystal (without defects) at $0.723$.
 

### 2016-12-12 nature/crystal_gel3.tex
Update colloid size from knowledge of actual pixel size. Impacts the main text, supplementary, phase diagram and local volume fraction

--- 2016-12-12 crystal-gel/nature/crystal_gel3.tex
@@
 To examine this in more detail, we plot in Fig.~\ref{fig:phase_separation}c the structure factor for the final stages of the gelation process for all state points.
-Calculation of the structure factor is done with the Hanning window function, to minimize boundary effects.
+%Calculation of the structure factor is done with the Hanning window function, to minimize boundary effects.
 The figure shows that at low wavenumber $q$, the structure factor displays fractal scaling compatible with the Guinier law, $S(q)\sim q^{-D}$.
@@
 Matching phase boundaries often leads to better determination of volume fractions than colloid diameter measurements~\cite{Poon2012}.
-We obtain the best match between theoretical phase diagram and experimental data for $\sigma=2.25\,\mu$m and $R_g=80$ nm (see Supplementary method).
-Therefore the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.071$ and the overlap mass fraction of polymer $2.25$~mg/g.
+We obtain the best match between theoretical phase diagram and experimental data for $\sigma=2.21\,\mu$m and $R_g=80$ nm (see Supplementary method).
+Therefore the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.072$ and the overlap mass fraction of polymer $2.25$~mg/g.
 
@@
 
-To account for imprecision in particle localisation close to contact, we consider two particles bonded if their center-to-center distance is within 2.3 $\mu$m $\approx\sigma+2R_g$.
+\textcolor{blue}{Calculation of the structure factor is done with the Hanning window function, to minimize boundary effects.
+To account for imprecision in particle localisation close to contact, we consider two particles bonded if their center-to-center distance is within 2.4~$\mu$m, which is the range of the potential ($\sigma+2R_g$) plus our worse estimate of localisation inaccuracy (0.15~pixels)~\cite{Leocmach2013}.}
 

### 2016-12-12 crystal-gel/nature/crystal_gel3.tex
Correct Methods section: 0.1µm is permeable to polymers "Since we have access to individual particle trajectories, we can directly assess the relative importance of these three growth channels"

@@
 To examine this in more detail, we plot in Fig.~\ref{fig:phase_separation}c the structure factor for the final stages of the gelation process for all state points.
-Calculation of the structure factor is done with the Hanning window function, to minimize boundary effects.
-The figure shows that at low wavenumber $q$, the structure factor displays fractal scaling compatible with the Guinier law, $S(q)\sim q^{-D}$.
+%Calculation of the structure factor is done with the Hanning window function, to minimize boundary effects.
+At low wavenumber $q$, the structure factor displays fractal scaling compatible with the Guinier law, $S(q)\sim q^{-D}$.
 But a difference in the fractal (or effective) dimension $D$ between the $\phi\approx 0.33$ and $c_p=0.38$ mg/g state point and other state points starts to become visible. In fact, while states
@@
 
-A direct analysis of the colloidal positions at long times confirms that indeed the new structure obtained at
+A direct analysis of the colloidal positions at late times confirms that indeed the new structure obtained at
 $\phi\approx 0.33$ and $c_p=0.38$ mg/g has different morphological properties than usual colloidal gels.
@@
 
-\subsection{Crystal-growth channels}
+\subsection{Crystal-growth routes}
 
-While nucleation always occurs inside the liquid branches of the phase-separating network, crystal-growth can proceed through different growth channels.
-In the following we will show that the growth mechanism of the crystal is not only due to filling of the spinodal liquid network, as was observed in Refs.~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways} for Brownian Dynamics simulations, but also involves
-The Bergeron process and Ostwald ripening. 
-In Fig.~\ref{fig:crystals}b the gyration radius of individual crystalline nuclei for the state point 
+While nucleation always occurs inside the liquid branches of the phase-separating network, crystal-growth can proceed through different growth \textcolor{blue}{routes}.
+In the following we will show that the growth mechanism of the crystal is not only due to filling of the spinodal liquid network, as was observed in Refs.~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways} for Brownian Dynamics simulations, but also involves the Bergeron process and Ostwald ripening.
+In Fig.~\ref{fig:crystals}b we show the gyration radius of individual crystalline nuclei for the state point 
 $\phi\approx 0.33$ and $c_p=0.38$ mg/g. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with 
@@
 is then accompanied by an increase in the fraction of gas particles: as the first crystals start to reach the gas phase, liquid particles evaporate 
-to the gas phase due to the higher vapour pressure of the liquid phase compared to the crystalline phase. After the onset 
-of a steady-state gas population, there are three growth channels for the crystal. Direct crystallization is the process by which 
-crystals grow by incorporating nearby liquid particles. In the Bergeron process, liquid droplets first evaporate and the resulting gas phase 
-contributes to the growth of crystalline regions. Ostwald ripening is instead the process by which small crystallites sublimate, and colloidal
-particles are transferred to larger nuclei.
-In order to assess the relative importance of these three growth channels, we analyse the trajectory, and, for every time frame, assign each particle a state between gas, liquid and crystal.
+to the gas phase due to the higher vapour pressure of the liquid phase compared to the crystalline phase. 
+After the onset of a steady-state gas population, there are three growth routes for the crystal. 
+Direct crystallization is the process by which crystals grow by incorporating nearby liquid particles. 
+In the Bergeron process, liquid droplets first evaporate and the resulting gas phase contributes to the growth of crystalline regions.
+Ostwald ripening is instead the process by which small crystallites sublimate, and colloidal particles are transferred to larger nuclei.
+\textcolor{blue}{Since we have access to individual particle trajectories, we can directly assess the relative importance of these three growth channels.}
+For every time frame, we assign each particle a state between gas, liquid and crystal.
 In order to minimize short-term fluctuations, the state of each colloidal particle is time averaged for $50\,\tau_{\rm B}$.
 We then measure the fraction of trajectories with different transition histories.
-Direct crystallization accounts for 54\% of particles trajectories in which
-gas or liquid particles transition to the crystal state without liquid evaporation or crystal de-sublimation. The Bergeron process accounts for 35\% of particle trajectories in which liquid particles
-transition to the gas state before crystallizing. Ostwald ripening accounts for 11\% of particle trajectories in which crystal particles transition to the gas state before returning
-to the crystal state. The different crystallization channels are depicted in the diagram of Fig.~\ref{fig:transitions}c.
+The different crystallization routes are depicted in the diagram of Fig.~\ref{fig:transitions}c.
+Direct crystallization accounts for 54\% of particles trajectories in which gas or liquid particles transition to the crystal state without liquid evaporation or crystal de-sublimation. 
+The Bergeron process accounts for 35\% of particle trajectories in which liquid particles transition to the gas state before crystallizing. 
+Ostwald ripening accounts for 11\% of particle trajectories in which crystal particles transition to the gas state before returning to the crystal state.
 
@@
 
-%\subsection*{Discussion and conclusions}
+\subsection*{Discussion and conclusions}
 The first stages of gelation always involve spinodal decomposition with the formation
@@
 of a network structure of the minority liquid phase, and (iv) the degree of supercooling is low enough
-\textcolor{blue}
-{
-to allow bond-breaking events that avoid the vitrification of the liquid phase.  
-}
+\textcolor{blue}{to allow bond-breaking events that avoid the vitrification of the liquid phase.}
 Many monoatomic and single-component molecular systems can satisfy all these conditions in a certain range of the temperature and pressure. 
@@
 
-To induce short-ranged depletion attraction, we use polystyrene (TOSOH) of molecular weight 3.8 Mdalton as non-adsorbing polymer.
+To induce short-ranged depletion attraction, we use polystyrene (TOSOH) as non-adsorbing polymer of molecular weight 3.8 MDa.
 Experiments are conducted at 27 $^\circ$C, some 80 $^\circ$C above the theta temperature in this solvents mixture~\cite{Royall2007}.
@@
 Matching phase boundaries often leads to better determination of volume fractions than colloid diameter measurements~\cite{Poon2012}.
-We obtain the best match between theoretical phase diagram and experimental data for $\sigma=2.25\,\mu$m and $R_g=80$ nm (see Supplementary method).
-Therefore the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.071$ and the overlap mass fraction of polymer $2.25$~mg/g.
+We obtain the best match between theoretical phase diagram and experimental data for $\sigma=2.21\,\mu$m and $R_g=80$ nm (see Supplementary method).
+Therefore the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.072$ and the overlap mass fraction of polymer $2.25$~mg/g.
 
-In the absence of salt, the Debye length is expected to reach a few $\mu$m and the (weakly) charged colloids experience a long range electrostatic repulsion.
+In the absence of salt, the Debye length is expected to reach a few micrometers and the (weakly) charged colloids experience a long range electrostatic repulsion.
 We confirm that colloids never come close enough to feel the short-ranged attraction.
-Screening by tetrabutylammonium bromide (Fluka) at saturated concentration brings down the Debye length to about several nanometers practically discarding the repulsion.
-Thus, salt injection (see below) can screen the Coulomb repulsion and make the polymer-induced depletion attraction effective, initiating phase separation.
+Screening by tetrabutylammonium bromide (Fluka) at saturated concentration brings down the Debye length to about 100~nm practically discarding the repulsion.
+Thus, salt introduction can screen the Coulomb repulsion and make the polymer-induced depletion attraction effective, initiating phase separation.
 
-Our sample cell has two layer compartments separated by membrane filters.
-The first layer is a sample container with thickness of 200 $\mu$m.
-A mixture of colloids, polymer, and solvent without salt is  set in the first layer for microscopy observation.
-The second layer is a salt-reservoir with a half-opened structure, which allows us to exchange or insert a reservoir solution.
-The volume ratio of the first and second compartments is approximately 1:100.
-These two compartments are separated by a membrane filter with its pore size of 0.1 $\mu$m, which passes only salt ions.
+\textcolor{blue}{To avoid solvent flow upon salt addition, our sample cell has two layer compartments separated by a membrane filter of pore size of 0.1 $\mu$m, permeable only to polymers and the salt ions.
+The top layer is 200 $\mu$m thick and contains the sample: a mixture of colloids, polymer, and solvent without salt.
+The bottom layer is a salt-reservoir with a half-opened structure, which allows us to exchange or insert a reservoir solution.
+The volume ratio of the first and second compartments is approximately 1:100.}
 
-In our experiments, the reservoir solution was initially a polymer solution without salt and electrostatic repulsion by the unscreened surface charges inhibited colloidal aggregation.
-Under microscopy observation, we quickly exchanged the reservoir solution into a polymer solution with salt at saturated concentration (4 mM) and sealed the half-opened reservoir with cover glass to avoid evaporation of solvent.
-Salt diffused into the first layer within a few minutes, typically 2 minutes, screened the surface charges, and initiated colloidal aggregation.
+In our experiments, the reservoir solution is initially a polymer solution without salt and electrostatic repulsion by the unscreened surface charges inhibited colloidal aggregation.
+Under microscopy observation, we quickly exchange the reservoir solution to a polymer solution with salt at saturated concentration (4 mM) and seal the half-opened reservoir with cover glass to avoid evaporation of solvent.
+Salt diffuse into the first layer within a few minutes, typically 2 minutes, screen the surface charges, and initiat colloidal aggregation.
 This allows us to observe the process without suffering from harmful turbulent flow.
 
-The data are collected on a Leica SP5 confocal microscope, using 532 nm laser excitation.
+The data are collected on an upright Leica SP5 confocal microscope, using 532 nm laser excitation.
 The scanning volume is 98 $\times$ 98 $\times$ 53 $\mu$m$^3$, which contains $\sim 10^4$ colloid particles.

### 2016-12-12 nature/crystal_gel3.tex
Remake fig:bond_breaking from John's data

@@
 %Calculation of the structure factor is done with the Hanning window function, to minimize boundary effects.
-The figure shows that at low wavenumber $q$, the structure factor displays fractal scaling compatible with the Guinier law, $S(q)\sim q^{-D}$.
+At low wavenumber $q$, the structure factor displays fractal scaling compatible with the Guinier law, $S(q)\sim q^{-D}$.
 But a difference in the fractal (or effective) dimension $D$ between the $\phi\approx 0.33$ and $c_p=0.38$ mg/g state point and other state points starts to become visible. In fact, while states
@@
 
-In the absence of salt, the Debye length is expected to reach a few $\mu$m and the (weakly) charged colloids experience a long range electrostatic repulsion.
+In the absence of salt, the Debye length is expected to reach a few micrometers and the (weakly) charged colloids experience a long range electrostatic repulsion.
 We confirm that colloids never come close enough to feel the short-ranged attraction.
-Screening by tetrabutylammonium bromide (Fluka) at saturated concentration brings down the Debye length to about several nanometers practically discarding the repulsion.
-Thus, salt injection (see below) can screen the Coulomb repulsion and make the polymer-induced depletion attraction effective, initiating phase separation.
+Screening by tetrabutylammonium bromide (Fluka) at saturated concentration brings down the Debye length to about 100~nm practically discarding the repulsion.
+Thus, salt introduction can screen the Coulomb repulsion and make the polymer-induced depletion attraction effective, initiating phase separation.
 
-Our sample cell has two layer compartments separated by membrane filters.
-The first layer is a sample container with thickness of 200 $\mu$m.
-A mixture of colloids, polymer, and solvent without salt is  set in the first layer for microscopy observation.
-The second layer is a salt-reservoir with a half-opened structure, which allows us to exchange or insert a reservoir solution.
-The volume ratio of the first and second compartments is approximately 1:100.
-These two compartments are separated by a membrane filter with its pore size of 0.1 $\mu$m, which passes only salt ions.
+\textcolor{blue}{To avoid solvent flow upon salt addition, our sample cell has two layer compartments separated by a membrane filter of pore size of 0.1 $\mu$m, permeable only to polymers and the salt ions.
+The top layer is 200 $\mu$m thick and contains the sample: a mixture of colloids, polymer, and solvent without salt.
+The bottom layer is a salt-reservoir with a half-opened structure, which allows us to exchange or insert a reservoir solution.
+The volume ratio of the first and second compartments is approximately 1:100.}
 
-In our experiments, the reservoir solution was initially a polymer solution without salt and electrostatic repulsion by the unscreened surface charges inhibited colloidal aggregation.
-Under microscopy observation, we quickly exchanged the reservoir solution into a polymer solution with salt at saturated concentration (4 mM) and sealed the half-opened reservoir with cover glass to avoid evaporation of solvent.
-Salt diffused into the first layer within a few minutes, typically 2 minutes, screened the surface charges, and initiated colloidal aggregation.
+In our experiments, the reservoir solution is initially a polymer solution without salt and electrostatic repulsion by the unscreened surface charges inhibited colloidal aggregation.
+Under microscopy observation, we quickly exchange the reservoir solution to a polymer solution with salt at saturated concentration (4 mM) and seal the half-opened reservoir with cover glass to avoid evaporation of solvent.
+Salt diffuse into the first layer within a few minutes, typically 2 minutes, screen the surface charges, and initiat colloidal aggregation.
 This allows us to observe the process without suffering from harmful turbulent flow.
 
-The data are collected on a Leica SP5 confocal microscope, using 532 nm laser excitation.
+The data are collected on an upright Leica SP5 confocal microscope, using 532 nm laser excitation.
 The scanning volume is 98 $\times$ 98 $\times$ 53 $\mu$m$^3$, which contains $\sim 10^4$ colloid particles.
@@
  \centering
- %\includegraphics[width=16cm]{./bond_breaking_figure.pdf}
- % bond_breaking_figure.pdf: 1553x1066 pixel, 72dpi, 54.79x37.61 cm, bb=0 0 1553 1066
- \caption{{\bf Stress driven aging.} Time evolution of the average coordination of colloidal particles $\bar{n}$ (black continuous line) and the bond-breaking probability $p_\text{break}$ (red dashed line) for state points $(\phi;c_p)$: {\bf a} $(0.14;1.36)$, {\bf b} $(0.33;1.07)$, {\bf c} $(0,14;0.82)$, {\bf d} $(0.33;0.38)$.
+ \includegraphics[width=16cm]{bond_breaking.pdf}
+ \caption{{\bf Stress driven aging.} Time evolution of the average coordination of colloidal particles $\bar{n}$ (black continuous line) and the bond-breaking probability $p_\text{break}$ (red dashed line) for four gel samples.
+Symbols relates each sample to its position on the phase diagram Fig~\ref{fig:phase_separation}a.
 $p_\text{break}$ is measured as the probability of bonds breaking in a time interval of $\Delta t=10\,s$.

### 2016-12-13 crystal-gel/nature/crystal_gel3.tex
twocolumn layout fig:network_formation as the two first panels of fig:phase_separation

@@
-\documentclass[preprint,amsmath,amssymb,superscriptaddress]{revtex4-1}
+\documentclass[twocolumn,amsmath,amssymb,superscriptaddress]{revtex4-1}
 \usepackage{graphicx,amsmath}
@@
 \begin{document}
-\draft
+%\draft
 
@@
 
-\textcolor{brown}{Mathieu: When I presented this story, I was often told that our initial focus on nucleation was misleading since our main result is about growth.
+%\textcolor{brown}{Mathieu: When I presented this story, I was often told that our initial focus on nucleation was misleading since our main result is about growth.
 What about trimming down this part?
-Or focussing more on growth rather than nucleation?
-}Under usual conditions, the nucleation stage involves a very small number of molecules, of the order of 10-1000 molecules.
+Or focussing more on growth rather than nucleation?}
+Under usual conditions, the nucleation stage involves a very small number of molecules, of the order of 10-1000 molecules.
 This has severely limited the possibility to observe the nucleation process directly, and has also posed limits on
@@
 \subsection{Samples}
+\begin{figure}
+ \includegraphics{network_formation.pdf}
+ \caption{{\bf Network formation.} 
+Phase diagram.
+Empty black triangles are experimental point showing a fluid behaviour.
+The other symbols are the experimental points showing gel behaviour, consistently used in all figures.
+The dotted lines are the gas-liquid tie lines determined experimentally, with arrow heads indicating measured gas and liquid compositions (see Supplementary Information on the procedure to extract the compositions from confocal images).
+Gray areas are theoretical equilibrium fluid and crystal regions.
+The solid and dashed lines are the metastable gas-liquid binodal and spinodal respectively.
+\textbf{(Inset)} Radius of gyration ($R_g$) as a function of cluster size ($s$) for all state points at percolation time.
+The gray line represents the fractal dimension of the random gelation universality class.
+ }
+ \label{fig:network_formation}
+\end{figure}
 
@@
 of the polymer concentration, $c_p=0.38,0.48,0.57,1.07$ mg/g for $\phi\approx 0.33$ and $c_p=0.82,1.36$ mg/g for $\phi\approx 0.14$.
-In Fig.~\ref{fig:phase_separation}a, we superimpose the state points with a theoretical phase diagram, which we calculated from generalised free volume theory~\cite{Fleer2008} using the experimental values for the size of the colloids, the radius of gyration of the polymer, and the polymer concentration (see Supplementary Methods).
+In Fig.~\ref{fig:network_formation}, we superimpose the state points with a theoretical phase diagram, which we calculated from generalised free volume theory~\cite{Fleer2008} using the experimental values for the size of the colloids, the radius of gyration of the polymer, and the polymer concentration (see Supplementary Methods).
 For each state point, scans were made at early times (every $10$~s, i.e. $4\tau_{\rm B}$) and at the late stage (every $30$~s) of the gelation process.
@@
 eventually forming a percolating network. Thus, a dense network (liquid) coexists with freely diffusing monomers (gas).
-In Fig.~\ref{fig:phase_separation}a, the state points which did not phase separate are indicated with open triangles, while
+In Fig.~\ref{fig:network_formation}, the state points which did not phase separate are indicated with open triangles, while
 the other symbols indicate samples undergoing phase separation. The black dashed line is the gas spinodal computed from free volume theory and the 
@@
 The topology of the network is similar to all our samples at the initial stages. 
-This is shown for example in Fig.~\ref{fig:phase_separation}b, where we plot the radius of gyration of colloidal clusters ($R_g$)
+This is shown for example in the inset of Fig.~\ref{fig:network_formation}, where we plot the radius of gyration of colloidal clusters ($R_g$)
 as a function of the cluster size ($s$) at the percolation point. Two particles belong to the same cluster if their distance is
 within the first peak of the radial distribution function. All samples show the same scaling law, $R_{\rm g}\sim s^{1/D}$, which is compatible with
-the random gelation universality class exponent ($D=2.53$) as shown by the dashed line in Fig.~\ref{fig:phase_separation}b
+the random gelation universality class exponent ($D=2.53$) as shown by the dashed line in the inset of Fig.~\ref{fig:network_formation}
 (we do not imply that the formation of the network follows the random gelation universality class, just that our
@@
 with high polymer concentration retain the exponent $D=2.5(3)$, 
-which is the random gel universality class exponent, as we also observed in the early stages of the gelation process (see Fig.~\ref{fig:phase_separation}b),
+which is the random gel universality class exponent, as we also observed in the early stages of the gelation process (see inset of Fig.~\ref{fig:network_formation}),
 the state point with low polymer concentration ($\phi\approx 0.33$ and $c_p=0.38$ mg/g) displays the biggest deviation from that exponent. In the figure we also
@@
  \centering
- %\includegraphics[width=16cm]{./bond_breaking_figure.pdf}
- % bond_breaking_figure.pdf: 1553x1066 pixel, 72dpi, 54.79x37.61 cm, bb=0 0 1553 1066
- \caption{{\bf Stress driven aging.} Time evolution of the average coordination of colloidal particles $\bar{n}$ (black continuous line) and the bond-breaking probability $p_\text{break}$ (red dashed line) for state points $(\phi;c_p)$: {\bf a} $(0.14;1.36)$, {\bf b} $(0.33;1.07)$, {\bf c} $(0,14;0.82)$, {\bf d} $(0.33;0.38)$.
+ \includegraphics[width=16cm]{bond_breaking.pdf}
+ \caption{{\bf Stress driven aging.} Time evolution of the average coordination of colloidal particles $\bar{n}$ (black continuous line) and the bond-breaking probability $p_\text{break}$ (red dashed line) for four gel samples.
+Symbols relates each sample to its position on the phase diagram Fig~\ref{fig:network_formation}.
 $p_\text{break}$ is measured as the probability of bonds breaking in a time interval of $\Delta t=10\,s$.

### 2016-12-13 crystal-gel/nature/crystal_gel3.tex
reduce fig:bond_breaking to a single column in a twocolumn layout
fig:late_structure including the two last panels of fig:phase_separation and fig:local_phi12 name phi_loc and phi_loc(12)

@@
 
+\begin{figure}
+ \includegraphics{bond_breaking.pdf}
+ \caption{{\bf Stress driven aging.} Time evolution of the average coordination of colloidal particles $\bar{n}$ (black continuous line) and the bond-breaking probability $p_\text{break}$ (red dashed line) for four gel samples.
+Symbols relates each sample to its position on the phase diagram Fig~\ref{fig:network_formation}.
+$p_\text{break}$ is measured as the probability of bonds breaking in a time interval of $\Delta t=10\,s$.
+For high polymer concentration, panels {\bf a} and {\bf b}, $p_\text{break}$ decreases monotonically until it reaches a stationary state at long times, with $\bar{n}$ saturating at an average of less than $5$ neighbours.
+For low polymer concentration, panels {\bf c} and {\bf d}, $p_\text{break}$ follows a similar decay at intermediate times, with $\bar{n}$ less than $5$ neighbours, but which is then followed by a second decay to new configuration, where $\bar{n}$ becomes greater than $5$ neighbours.}
+ \label{fig:bond_breaking}
+\end{figure}
+
 \textcolor{blue}
@@
 
+\begin{figure*}
+ \includegraphics{late_structure.pdf}
+ \caption{{\bf Structure at late times.} 
+{\bf a,} Structure factors for the late stages of the gelation process. The gray and black lines represents respectively random gelation and compact fractal dimensions.
+{\bf b,} Late stage of the evolution of the local volume fraction around colloidal particles having 12 nearest neighbours.
+{\bf c,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle for the state points in the very late stage. 
+The two horizontal straight lines indicate the characteristic volume fractions in attractive systems, crystal ($\sim 0.72$) and glass ($\sim 0.64$).
+Samples are shown with the same colour code as in Fig.~\ref{fig:network_formation}. In particular red circles correspond to the crystallizing sample at $\phi=0.33$ and $cp=0.38\,$mg/g.
+ }
+ \label{fig:late_structure}
+\end{figure*}
+
 The samples show significantly different behaviours in the later stages of the phase separation process.
-To examine this in more detail, we plot in Fig.~\ref{fig:phase_separation}c the structure factor for the final stages of the gelation process for all state points.
+To examine this in more detail, we plot in Fig.~\ref{fig:late_structure}a the structure factor for the final stages of the gelation process for all state points.
 %Calculation of the structure factor is done with the Hanning window function, to minimize boundary effects.
@@
 The difference is due to the onset of crystallization in the low-polymer concentration sample. This is already evident in the high $q$ behaviour of
-the structure factors (Fig.~\ref{fig:phase_separation}c), where the diffraction planes appear as sharper peaks for $\phi\approx 0.33$ and $c_p=0.38$ mg/g.
+the structure factors (Fig.~\ref{fig:late_structure}a), where the diffraction planes appear as sharper peaks for $\phi\approx 0.33$ and $c_p=0.38$ mg/g.
 
-To gain additional insight, in Supplementary Figure~\ref{S-fig:local_phi12} we plot the late stage (after all samples already underwent gas-liquid phase separation) time evolution of the local volume fraction of closed packed particles (having 12 neighbours).
+To gain additional insight, in Fig.~\ref{fig:late_structure}b we plot the late stage (after all samples already underwent gas-liquid phase separation) time evolution of the local volume fraction $\phi_\mathrm{loc}$ of closed packed particles (having 12 neighbours).
 The volume fraction is obtained by computing the Voronoi diagram of each configuration, which uniquely assigns a volume to each colloidal particle.
 Only the state point at $\phi\approx 0.33$ and $c_p = 0.38\,$mg/g shows an increase, up to an asymptotic value that we identify with the composition of the stable crystal phase $\approx 0.72$.
-We plot the asymptotic average volume fraction as a function of the number of neighbours in Fig.~\ref{fig:phase_separation}d.
+We plot the asymptotic average volume fraction as a function of the number of neighbours in Fig.~\ref{fig:late_structure}c.
 Approaching close packing (12 neighbours), we clearly see two families of curves.
@@
 
-%To gain additional insight, in Fig.~\ref{fig:phase_separation}d we plot the average local volume fraction as a function of the number of neighbours.
+%To gain additional insight, in Fig.~\ref{fig:late_structure}c we plot the average local volume fraction as a function of the number of neighbours.
 %\textcolor{blue}{The volume fraction is obtained by computing the Voronoi diagram of each configuration, which uniquely assigns a volume to each colloidal particle.}
@@
 follows a volume growth with the formation of isotropic crystal droplets. The same scaling law was suggested in the analysis of the
-structure factors, Fig.~\ref{fig:phase_separation}c, but here it is shown directly by the analysis of the gyration radius.
+structure factors, Fig.~\ref{fig:late_structure}a, but here it is shown directly by the analysis of the gyration radius.
 
@@
  \centering
- \includegraphics{phase_separation.pdf}
- \caption{{\bf Different routes of viscoelastic phase separation.} 
-{\bf a,} Phase diagram.
-Empty black triangles are experimental point showing a fluid behaviour.
-The other symbols are the experimental points showing gel behaviour, consistently used in all panels.
-The dotted lines are the gas-liquid tie lines determined experimentally, with arrow heads indicating measured gas and liquid compositions (see Supplementary Information on the procedure to extract the compositions from confocal images).
-Gray areas are theoretical equilibrium fluid and crystal regions.
-The solid and dashed lines are the metastable gas-liquid binodal and spinodal respectively.
-{\bf b,} Radius of gyration ($R_g$) as a function of cluster size ($s$) for all state points at percolation. The
- gray line represents the fractal dimension of the random gelation universality class.
-{\bf c,} Structure factors for the late stages of the gelation process. The gray and black lines represents respectively random gelation and compact fractal dimensions.
-{\bf d,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle for the state points in the very late stage. 
-The two horizontal straight lines indicate the characteristic volume fractions in attractive systems, crystal ($\sim 0.72$) and glass ($\sim 0.64$).
- }
- \label{fig:phase_separation}
-\end{figure*}
-
-\clearpage
-
-\begin{figure}
- \centering
- \includegraphics[width=16cm]{bond_breaking.pdf}
- \caption{{\bf Stress driven aging.} Time evolution of the average coordination of colloidal particles $\bar{n}$ (black continuous line) and the bond-breaking probability $p_\text{break}$ (red dashed line) for four gel samples.
-Symbols relates each sample to its position on the phase diagram Fig~\ref{fig:network_formation}.
-$p_\text{break}$ is measured as the probability of bonds breaking in a time interval of $\Delta t=10\,s$.
-For high polymer concentration, panels {\bf a} and {\bf b}, $p_\text{break}$ decreases monotonically until it reaches a stationary state at long times, with $\bar{n}$ saturating at an average of less than $5$ neighbours.
-For low polymer concentration, panels {\bf c} and {\bf d}, $p_\text{break}$ follows a similar decay at intermediate times, with $\bar{n}$ less than $5$ neighbours, but which is then followed by a second decay to new configuration, where $\bar{n}$ becomes greater than $5$ neighbours.}
- \label{fig:bond_breaking}
-\end{figure}
-
-\clearpage
-
-\begin{figure*}[!t]
- \centering
  \includegraphics[width=12cm]{fig2}

### 2016-12-13 crystal-gel/nature/crystal_gel3.tex
move fig:network in the text and scale it to one column. Move the method of coarse-grained representation to Methods

@@
 
-% 
-% 
-% \begin{figure}
-%  \centering
-%  \framebox{Phase diagram\hfill}
-%  \includegraphics[width=8cm]{oldfig/fig3b}
-%  \includegraphics[width=8cm]{oldfig/fig3c}
-%  % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
-%  \caption{{\bf Two types of arrest mechanism leading to gel formation.} \textbf{a,} Phase diagram \textbf{b,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle.
-%  Different lines represent the different state points. {\bf c,} Structure factors for the late stages of the gelation process.} 
-%  \label{fig:statepoints}
-% \end{figure}
+\begin{figure}
+ \includegraphics[width=\linewidth]{fig2}
+ % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
+\caption{{\bf Percolated network structures.} {\bf a:} gel; {\bf b:} crystal-gel. The gel state is obtained from the state
+point $\phi\approx 0.33$ and $c_p=1.07$ mg/g at $t=400$ min after sample preparation, while the crystal-gel is obtained at the same $\phi$ but
+at lower polymer concentration, $c_p=0.38$ mg/g at $t=468$ min after sample preparation. On the left, the network is represented in a coarse-grained fashion (see Methods). 
+We can clearly see not only the percolated nature of both networks, but also the difference in the smoothness of the network surface between amorphous and crystal gels. 
+On the right, slabs of 4.5 $\mu$m thickness and 41 $\mu$m width are plotted, with particles coloured with the same colour scheme as in Fig.~\ref{fig:transitions} red (crystal), green (gas particles), white (liquid particles).
+We note that very few liquid particles remain in \textbf{b} and the network is almost crystalline. 
+} 
+\label{fig:network}
+\end{figure}
 
@@
 $\phi\approx 0.33$ and $c_p=0.38$ mg/g has different morphological properties than usual colloidal gels.
-We directly compare the two structures in Fig.~\ref{fig:network}, where Gaussian filtering is used to depict the network structure as a continuous field.
+We directly compare the two structures in Fig.~\ref{fig:network}, where Gaussian filtering is used to depict the network structure as a continuous field (see Methods).
 Panel a presents the results for high
@@
 
+\textcolor{blue}{Coarse-grained representation of the network is obtained by applying a Gaussian filter to the position of the particles (of width equal to the inter-particle distance) and plotting the surface which bounds the highest 20 \% values of the field.}
+
 \bibliographystyle{naturemag2}
@@
 The authors declare no competing financial interests
-
-\clearpage
-
-\begin{figure*}[!t]
- \centering
- \includegraphics[width=12cm]{fig2}
- % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
-\caption{{\bf Percolated network structures.} {\bf a:} gel; {\bf b:} crystal-gel. The gel state is obtained from the state
-point $\phi\approx 0.33$ and $c_p=1.07$ mg/g at $t=400$ min after sample preparation, while the crystal-gel is obtained at the same $\phi$ but
-at lower polymer concentration, $c_p=0.38$ mg/g at $t=468$ min after sample preparation. On the left, the network is represented as a continuous field that has
-been obtained by applying a Gaussian filter to the position of the particles (of width equal to the inter-particle distance) and
-plotting the surface which bounds the highest 20 \% values of the field. We can clearly see not only the percolated nature of both networks, 
-but also the difference in the smoothness of the network surface between amorphous and crystal gels. 
-On the right, slabs of 4.5 $\mu$m thickness and 41 $\mu$m width are plotted,
-with particles coloured with the same colour scheme as in Fig.~\ref{fig:transitions}: red (crystal), green (gas particles), white (liquid particles). 
-We note that very few liquid particles remain and the network is almost crystalline. 
-} 
-\label{fig:network}
-\end{figure*}
 

### 2016-12-13 nature/crystal_gel3.tex
Fit fig:crystals on one column by putting the time evolution in inset. Decimate time evolution and rescale time to 10^3 tau_B in the data (faster plot)

@@
 \subsection{Intermediate times: stress-driven aging}
+
+\begin{figure}
+ \includegraphics{bond_breaking.pdf}
+ \caption{{\bf Stress driven aging.} Time evolution of the average coordination of colloidal particles $\bar{n}$ (black continuous line) and the bond-breaking probability $p_\text{break}$ (red dashed line) for four gel samples.
+Symbols relates each sample to its position on the phase diagram Fig~\ref{fig:network_formation}.
+$p_\text{break}$ is measured as the probability of bonds breaking in a time interval of $\Delta t=10\,s$.
+For high polymer concentration, panels {\bf a} and {\bf b}, $p_\text{break}$ decreases monotonically until it reaches a stationary state at long times, with $\bar{n}$ saturating at an average of less than $5$ neighbours.
+For low polymer concentration, panels {\bf c} and {\bf d}, $p_\text{break}$ follows a similar decay at intermediate times, with $\bar{n}$ less than $5$ neighbours, but which is then followed by a second decay to new configuration, where $\bar{n}$ becomes greater than $5$ neighbours.}
+ \label{fig:bond_breaking}
+\end{figure}
 
@@
 
+\begin{figure}
+ \includegraphics{characterisation}
+ % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
+\caption{{\bf Characterization of the crystal-gel.} 
+Radius of gyration of crystalline nuclei at the late stages of gelation for the state point $\phi\approx 0.33$ and $c_p=0.38$ mg/g.
+The gray and black straight lines have a slope of $1/2.53$ and $1/3$ respectively
+\textbf{(Inset)} Time evolution of the average size of the crystals (black line) and the number of crystals (red dashed line) for the same state point.
+} 
+ \label{fig:crystals}
+\end{figure}
+
 Supplementary Movie 2 shows the late stage (after the initial spinodal decomposition) evolution of a 2D confocal slice.
@@
 With this analysis we reveal that crystallization events indeed start occurring for the state point $\phi\approx 0.33$ and $c_p=0.38$ mg/g, which has the lowest
-polymer concentration of all state points. This is shown in Fig.~\ref{fig:crystals}a, where the black line indicates the average size of the crystallites, 
+polymer concentration of all state points. This is shown in the inset of Fig.~\ref{fig:crystals}, where the black line indicates the average size of the crystallites, 
 while the red line indicates the average number of crystallites as a function of time. The number of crystallites first rapidly increases as nucleation events
@@
 In the following we will show that the growth mechanism of the crystal is not only due to filling of the spinodal liquid network, as was observed in Refs.~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways} for Brownian Dynamics simulations, but also involves the Bergeron process and Ostwald ripening.
-In Fig.~\ref{fig:crystals}b we show the gyration radius of individual crystalline nuclei for the state point 
+In Fig.~\ref{fig:crystals} we show the gyration radius of individual crystalline nuclei for the state point 
 $\phi\approx 0.33$ and $c_p=0.38$ mg/g. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with 
@@
  \centering
- \includegraphics{phase_separation.pdf}
- \caption{{\bf Different routes of viscoelastic phase separation.} 
-{\bf a,} Phase diagram.
-Empty black triangles are experimental point showing a fluid behaviour.
-The other symbols are the experimental points showing gel behaviour, consistently used in all panels.
-The dotted lines are the gas-liquid tie lines determined experimentally, with arrow heads indicating measured gas and liquid compositions (see Supplementary Information on the procedure to extract the compositions from confocal images).
-Gray areas are theoretical equilibrium fluid and crystal regions.
-The solid and dashed lines are the metastable gas-liquid binodal and spinodal respectively.
-{\bf b,} Radius of gyration ($R_g$) as a function of cluster size ($s$) for all state points at percolation. The
- gray line represents the fractal dimension of the random gelation universality class.
-{\bf c,} Structure factors for the late stages of the gelation process. The gray and black lines represents respectively random gelation and compact fractal dimensions.
-{\bf d,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle for the state points in the very late stage. 
-The two horizontal straight lines indicate the characteristic volume fractions in attractive systems, crystal ($\sim 0.72$) and glass ($\sim 0.64$).
- }
- \label{fig:phase_separation}
-\end{figure*}
-
-\clearpage
-
-\begin{figure}
- \centering
- \includegraphics[width=16cm]{bond_breaking.pdf}
- \caption{{\bf Stress driven aging.} Time evolution of the average coordination of colloidal particles $\bar{n}$ (black continuous line) and the bond-breaking probability $p_\text{break}$ (red dashed line) for four gel samples.
-Symbols relates each sample to its position on the phase diagram Fig~\ref{fig:phase_separation}a.
-$p_\text{break}$ is measured as the probability of bonds breaking in a time interval of $\Delta t=10\,s$.
-For high polymer concentration, panels {\bf a} and {\bf b}, $p_\text{break}$ decreases monotonically until it reaches a stationary state at long times, with $\bar{n}$ saturating at an average of less than $5$ neighbours.
-For low polymer concentration, panels {\bf c} and {\bf d}, $p_\text{break}$ follows a similar decay at intermediate times, with $\bar{n}$ less than $5$ neighbours, but which is then followed by a second decay to new configuration, where $\bar{n}$ becomes greater than $5$ neighbours.}
- \label{fig:bond_breaking}
-\end{figure}
-
-\clearpage
-
-\begin{figure*}[!t]
- \centering
- \includegraphics[width=12cm]{fig2}
- % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
-\caption{{\bf Percolated network structures.} {\bf a:} gel; {\bf b:} crystal-gel. The gel state is obtained from the state
-point $\phi\approx 0.33$ and $c_p=1.07$ mg/g at $t=400$ min after sample preparation, while the crystal-gel is obtained at the same $\phi$ but
-at lower polymer concentration, $c_p=0.38$ mg/g at $t=468$ min after sample preparation. On the left, the network is represented as a continuous field that has
-been obtained by applying a Gaussian filter to the position of the particles (of width equal to the inter-particle distance) and
-plotting the surface which bounds the highest 20 \% values of the field. We can clearly see not only the percolated nature of both networks, 
-but also the difference in the smoothness of the network surface between amorphous and crystal gels. 
-On the right, slabs of 4.5 $\mu$m thickness and 41 $\mu$m width are plotted,
-with particles coloured with the same colour scheme as in Fig.~\ref{fig:transitions}: red (crystal), green (gas particles), white (liquid particles). 
-We note that very few liquid particles remain and the network is almost crystalline. 
-} 
-\label{fig:network}
-\end{figure*}
-
-\clearpage
-
-\begin{figure}[!t]
- \centering
- \includegraphics{characterisation}
- % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
-\caption{{\bf Characterization of the crystal-gel.} {\bf a,} Time evolution of the average size of the crystals (blue line) and the number of crystals (red line) for the state point $\phi\approx 0.33$ and $c_p=0.38$ mg/g.
-{\bf b,} Radius of gyration of crystalline nuclei at the late stages of gelation for the same state point.
-The gray and black straight lines have a slope of $1/2.53$ and $1/3$ respectively} 
- \label{fig:crystals}
-\end{figure}
-
-\clearpage
-
-\begin{figure*}[!t]
- \centering
  \includegraphics[width=14cm]{fig4}

### 2016-12-13 crystal-gel/nature/crystal_gel3.tex
fig:transitions in the text

@@
 
+\begin{figure*}
+ \centering
+ \includegraphics[width=14cm]{fig4}
+ % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735
+ \caption{{\bf Crystal gel formation.}  
+{\bf a,} Reconstructions from confocal coordinates at $\phi\approx 0.33$ and $c_p=0.38$ mg/g.
+The depth of view is 9.0 $\mu$m, while the lateral dimension is 41 $\mu$m.
+Particles are drawn to scale and coloured according to their phase.
+Red: crystal; white: liquid; green: gas.
+{\bf b,} Fraction of particles for each phase. Gray vertical lines indicate the times shown in panel a. 
+{\bf c,} Each particle during its trajectory changes between three states: gas, liquid and crystal. Direct crystallization (continuous arrow) accounts for 54\% of particles trajectories in which
+gas or liquid particles transition to the crystal state without liquid evaporation or crystal de-sublimation.
+The Bergeron process (dotted arrow) accounts for 35\% of particle trajectories in which liquid particles
+transition to the gas state before crystallizing. Ostwald ripening (dashed arrow) accounts for 11\% of particle trajectories in which crystal particle transition to the gas state before returning
+to the crystal state.}
+\label{fig:transitions}
+\end{figure*}
+
 While nucleation always occurs inside the liquid branches of the phase-separating network, crystal-growth can proceed through different growth \textcolor{blue}{routes}.
 In the following we will show that the growth mechanism of the crystal is not only due to filling of the spinodal liquid network, as was observed in Refs.~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways} for Brownian Dynamics simulations, but also involves the Bergeron process and Ostwald ripening.
-In Fig.~\ref{fig:crystals}b we show the gyration radius of individual crystalline nuclei for the state point 
+In Fig.~\ref{fig:crystals} we show the gyration radius of individual crystalline nuclei for the state point 
 $\phi\approx 0.33$ and $c_p=0.38$ mg/g. The results show that the crystal growth follows two different scaling laws: at small crystalline sizes it scales with 
@@
 
-\subsection*{Discussion and conclusions}
+\subsection{Discussion and conclusions}
 The first stages of gelation always involve spinodal decomposition with the formation
@@
 
-\bibliographystyle{naturemag2}
-\bibliography{biblio}
-
-\vspace{1cm}
+\bigskip
 \noindent
@@
 
-\vspace{0.3cm}
+\medskip
 \noindent
@@
 
-\clearpage
-
-\begin{figure}[!t]
- \centering
- \includegraphics{characterisation}
- % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
-\caption{{\bf Characterization of the crystal-gel.} {\bf a,} Time evolution of the average size of the crystals (blue line) and the number of crystals (red line) for the state point $\phi\approx 0.33$ and $c_p=0.38$ mg/g.
-{\bf b,} Radius of gyration of crystalline nuclei at the late stages of gelation for the same state point.
-The gray and black straight lines have a slope of $1/2.53$ and $1/3$ respectively} 
- \label{fig:crystals}
-\end{figure}
-
-\clearpage
-
-\begin{figure*}[!t]
- \centering
- \includegraphics[width=14cm]{fig4}
- % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735
- \caption{{\bf Crystal gel formation.}  
-{\bf a,} Reconstructions from confocal coordinates at $\phi\approx 0.33$ and $c_p=0.38$ mg/g.
-The depth of view is 9.0 $\mu$m, while the lateral dimension is 41 $\mu$m.
-Particles are drawn to scale and coloured according to their phase.
-Red: crystal; white: liquid; green: gas.
-{\bf b,} Fraction of particles for each phase. Gray vertical lines indicate the times shown in panel a. 
-{\bf c,} Each particle during its trajectory changes between three states: gas, liquid and crystal. Direct crystallization (continuous arrow) accounts for 54\% of particles trajectories in which
-gas or liquid particles transition to the crystal state without liquid evaporation or crystal de-sublimation.
-The Bergeron process (dotted arrow) accounts for 35\% of particle trajectories in which liquid particles
-transition to the gas state before crystallizing. Ostwald ripening (dashed arrow) accounts for 11\% of particle trajectories in which crystal particle transition to the gas state before returning
-to the crystal state.}
-\label{fig:transitions}
-\end{figure*}
+\bibliographystyle{naturemag2}
+\bibliography{biblio}
 

### 2016-12-13 crystal-gel/nature/crystal_gel3.tex
shift local phi of 12 neighboured particles to correspond to agregation starting time

@@
 This is shown for example in the inset of Fig.~\ref{fig:network_formation}, where we plot the radius of gyration of colloidal clusters ($R_g$)
-as a function of the cluster size ($s$) at the percolation point. Two particles belong to the same cluster if their distance is
-within the first peak of the radial distribution function. All samples show the same scaling law, $R_{\rm g}\sim s^{1/D}$, which is compatible with
-the random gelation universality class exponent ($D=2.53$) as shown by the dashed line in the inset of Fig.~\ref{fig:network_formation}
-(we do not imply that the formation of the network follows the random gelation universality class, just that our
-fractal (or effective) dimension is compatible with it).
+as a function of the cluster size ($s$) at the percolation time, see Methods. All samples show the same scaling law, $R_{\rm g}\sim s^{1/D}$, which is compatible with
+the random gelation universality class exponent ($D=2.53$) as shown by the gray line in the inset of Fig.~\ref{fig:network_formation}.
+Here we do not imply that the formation of the network follows the random gelation universality class, just that our
+fractal (or effective) dimension is compatible with it.
 
@@
 
-\vspace{1cm}
-\noindent
-{\bf METHODS}
+\section*{METHODS}
 
-\noindent
-{\bf Experiments}
+\subsection*{Experiments}
 
@@
 
-\noindent
-{\bf Structural analysis}
+\subsection*{Structural analysis}
 
@@
 
-To detect percolation, two different methods have been employed. In the first one we detect percolation
-by looking at which frame the cluster size distribution has a more extended power-law decay. In the second technique,
-we simply measure the spatial extent of the largest cluster and consider it percolating when it is comparable to the
-size of the field of view of the microscope. Both methods lead to essentially identical percolation time for each state point.
+To detect percolation, two different methods have been employed. In the first one we detect percolation by looking at which frame the cluster size distribution has a more extended power-law decay. 
+In the second technique, we simply measure the spatial extent of the largest cluster and consider it percolating when it is comparable to the size of the field of view of the microscope. 
+Both methods lead to essentially identical percolation time for each state point.
 
@@
 H.Ts. and J.R. contributed equally to this work. 
-H.Ta. conceived and supervised the project, H.Ts. performed experiments, J.R. analysed the data, M.L. calculated the phase diagram, and all the authors discussed and wrote the manuscript.
+H.Ta. conceived and supervised the project, H.Ts. performed experiments, J.R. analysed the data, M.L. linked experiments and analysis, and all the authors discussed and wrote the manuscript.
 

### 2016-12-15 crystal-gel/nature/crystal_gel_r.tex
Redeclare Sensei's changes from a copy of the previous file

--- 2016-12-13 crystal-gel/nature/crystal_gel3.tex
@@
-\documentclass[twocolumn,amsmath,amssymb,superscriptaddress]{revtex4-1}
+\documentclass[preprint,amsmath,amssymb,superscriptaddress]{revtex4-1}
+%\documentclass[twocolumn,amsmath,amssymb,superscriptaddress]{revtex4-1}
 \usepackage{graphicx,amsmath}
@@
 %the glass transition, with the formation of colloidal gels~\cite{poon2002,zaccarelli2007,piazza1994phase,verhaegh1997transient,tanaka1999colloid,lu2008gelation}, or by crystallization, which results in the formation of a crystalline network~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways,sabin2012}.
-glass transition~\cite{poon2002,zaccarelli2007,piazza1994phase,verhaegh1997transient,tanaka1999colloid,lu2008gelation} or by crystallization~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways,sabin2012,zhang2012non}.
+glass transition
+%~\cite{poon2002,zaccarelli2007,piazza1994phase,verhaegh1997transient,tanaka1999colloid,lu2008gelation} 
+or by crystallization. 
+%~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways,sabin2012,zhang2012non}. 
 %But while the former process has been studied in detail, the latter has never been observed experimentally at a single-particle level and it is thus lacking a microscopic understanding. 
-With a new confocal microscopy protocol, we unveil a novel crystallization pathway to sponge-like porous structures in a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation.
-\textcolor{blue}{
-We follow the entire kinetics of phase separation, from the homogeneous phase, percolation, stress driven aging, and until the formation of porous crystals.}
+With a new confocal microscopy protocol, \textcolor{blue}{we successfully follow the entire kinetics of phase separation at a single-particle level, from the homogeneous phase, percolation, stress driven ageing, and until the formation of the final gel structure.
+This allows us to unveil a novel crystallization pathway to sponge-like porous crystal structures in a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation and to reveal its physical mechanism.}
+%\textcolor{blue}{We follow the entire kinetics of phase separation, from the homogeneous phase, percolation, stress driven aging, and until the formation of porous crystals.}
 We show that crystallization leads to a network structure different from the original phase separation pattern. 
-Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface, i.e.
-de-sublimation, driving evaporation of nearby liquid.
+Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface, i.e. de-sublimation, driving evaporation of nearby liquid. 
 %The latter processes via the gas phase has been overlooked so far.
@@
 We argue that similar crystal-gel states can be formed as the result of the gas-liquid phase separation of any crystallizable components, such as monoatomic and molecular systems, if dynamics of the liquid phase is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
-This mechanism will provide a novel single-step pathway to form nano-porous crystals of metals and semiconductors without dealloying \cite{erlebacher2001evolution}, which may be important for catalytic, optical, sensing, and filtration applications \cite{ding2004nanoporous,ding2009nanoporous, wittstock2010nanoporous}.
+This mechanism will provide a novel single-step pathway to form nano-porous crystals of metals and semiconductors without dealloying, %\cite{erlebacher2001evolution},
+which may be important for catalytic, optical, sensing, and filtration applications. 
+%\cite{ding2004nanoporous,ding2009nanoporous, wittstock2010nanoporous}.
 %, but also sheds new light on crystal network formation in more complex systems such as magma \cite{philpotts1998role}, biominerals \cite{rousseau2005multiscale}, and foods \cite{deman1987fat}.}
@@
 
-\subsection{Introduction}
+%\subsection{Introduction}
 
-Crystallization plays a fundamental role in many processes occurring in nature, such as \textcolor{blue}{ice formation in atmospheric clouds}, and in technological applications
+Crystallization plays a fundamental role in many processes occurring in nature, such as \textcolor{blue}{ice formation in atmospheric clouds \cite{glickman2000glossary,morrison2012resilience}}, and in technological applications
 that are at the core of the chemical, pharmaceutical, and food industries.
@@
 Under usual conditions, the nucleation stage involves a very small number of molecules, of the order of 10-1000 molecules.
-This has severely limited the possibility to observe the nucleation process directly, and has also posed limits on
-models based on macroscopic thermodynamic properties. 
-The inability to directly observe the nucleation process is also a limit to our understanding of its kinetic pathway. 
+This has severely limited the possibility to \textcolor{blue}{experimentally} observe the  \textcolor{blue}{crystallization} process directly, and has also posed limits on models based on macroscopic thermodynamic properties.
+The inability to directly observe the  \textcolor{blue}{crystallization} process is also a limit to our understanding of its kinetic pathway. 
 Classical Nucleation Theory assumes the crystallization processes to occur in one step,
@@
 resolution, and at the same time scales over which nucleation takes place. In colloids with short-range attractions the liquid-gas transition 
-becomes metastable with respect to crystallization~\cite{anderson2002insights,lekkerkerker2011colloids}, and can form gels
-by viscoelastic phase separation \cite{tanaka1999colloid,tanaka2000viscoelastic}, with the coexistence of a dense liquid phase with slow dynamics and a dilute gas phase with fast dynamics. 
+becomes metastable with respect to crystallization~\cite{anderson2002insights,lekkerkerker2011colloids}, and can form gels.  
+~\cite{poon2002,zaccarelli2007,piazza1994phase,verhaegh1997transient,lu2008gelation}. 
+\textcolor{blue}{This gel formation process can be regarded as viscoelastic phase separation  \cite{tanaka1999colloid,tanaka2000viscoelastic} into a dense liquid phase with slow dynamics and a dilute gas phase with fast dynamics.}
 This difference in the viscoelastic properties between the two phases allows the system to form a space-spanning 
@@
 of the dense liquid phase~\cite{pusey1993dynamics,piazza1994phase,ilett1995phase,verhaegh1997transient,tanaka1999colloid,foffi2002,buzzaccaro2007sticky,zaccarelli2007,lu2008gelation,zaccarelli2008gelation,testard2011}.
-More precisely, the dynamical stabilization is due to percolation of locally favoured structures, which are locally stable non-crystalline structures \cite{royall2008g}.
+\textcolor{blue}{For thin colloidal gels}, the dynamical stabilization is due to percolation of locally favoured structures, which are locally stable non-crystalline structures \cite{royall2008g}.
 
@@
 was suggested by numerical simulations~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways} and observed in experiments \cite{sabin2012,zhang2012non}.
+\textcolor{blue}{Such crystallization upon gel formation was also observed
+for a system of oppositely charged colloids \cite{sanz2008gel,sanz2008out}.}  
 For colloidal systems, we can in principle access structural evolution in real time at a single-particle level; however, there has so far been no confocal microscopy  
@@
 This introduces turbulent flows at the beginning of the process, and also does not allow the
-observation of the initial stages of crystallization \textcolor{brown}{Is it crystallization of gelation here in this context?}.
-This initial perturbation may even alter the selection of the final non-equilibrium arrested state.
+observation of the initial stages of  \textcolor{blue}{gelation}. This initial perturbation may even alter the selection of the final non-equilibrium arrested state.  
 In our experiments we instead succeed in building a novel protocol that allows the time evolution of the system to be observed directly
@@
 We show that a necessary condition for the nucleation of crystal seeds is the stress-driven aging of the network which releases the stresses that build-up during the network formation due to hydrodynamic interactions, allowing for an increase in the number of average neighbours.
-If the average concentration of particles is high enough, this relaxation triggers the nucleation process. 
-Stress-driven aging is possible only at low polymer concentration. 
-Indeed, at higher polymer concentration (stronger interparticle attraction) we observe that, instead of forming crystal gels, the system undergoes the classical gelation process, in which
+If the average concentration of particles is high enough, this stress relaxation triggers the nucleation process of crystals. 
+Stress-driven ageing is possible only at low polymer concentration. 
+Indeed, at higher polymer concentration (i.e., under stronger interparticle attraction) we observe that, instead of forming crystal gels, the system undergoes the classical gelation process, in which
 phase separation is interrupted by vitrification~\cite{verhaegh1997transient,tanaka1999colloid,Royall2007,lu2008gelation}.
 We then show that the growth of the crystal seeds occurs by three different routes:
-(i) growth within the dense branches of the gel ;
-(ii) Bergeron process, analogous to ice formation in mixed phase clouds, where ice droplets grow at the
-expense of the supercooled liquid droplets due to their lower saturated vapour pressure ;
+(i) growth within the dense branches of the gel;
+(ii) Bergeron process, analogous to ice formation in mixed phase clouds \cite{glickman2000glossary,morrison2012resilience}, where ice droplets grow at the
+expense of the supercooled liquid droplets due to their lower saturated vapour pressure;
 (iii) Ostwald ripening.
@@
 
-\subsection{Samples}
-\begin{figure}
- \includegraphics{network_formation.pdf}
- \caption{{\bf Network formation.} 
-Phase diagram.
-Empty black triangles are experimental point showing a fluid behaviour.
+%\subsection{Samples}
+\begin{figure}[h]
+ \includegraphics[width=8cm]{network_formation.pdf}
+ \caption{\textcolor{blue}{{\bf Phase behaviour.}} 
+Empty black triangles are experimental points showing fluid behaviour.
 The other symbols are the experimental points showing gel behaviour, consistently used in all figures.
@@
 \textbf{(Inset)} Radius of gyration ($R_g$) as a function of cluster size ($s$) for all state points at percolation time.
-The gray line represents the fractal dimension of the random gelation universality class.
+The gray line represents the fractal dimension of the random gelation universality class \textcolor{blue}{($D=2.53$)}.
  }
@@
 
-\subsection{Early times: percolation}
+%\subsection{Early times: percolation}
 
@@
 the random gelation universality class exponent ($D=2.53$) as shown by the gray line in the inset of Fig.~\ref{fig:network_formation}.
-Here we do not imply that the formation of the network follows the random gelation universality class, just that our
+Here we do not imply that the formation of the network follows the random gelation universality class, \textcolor{blue}{but} just that our
 fractal (or effective) dimension is compatible with it.
 
-\subsection{Intermediate times: stress-driven aging}
+%\subsection{Intermediate times: stress-driven aging}
 
 \begin{figure}
- \includegraphics{bond_breaking.pdf}
+ \includegraphics[width=10cm]{bond_breaking.pdf}
  \caption{{\bf Stress driven aging.} Time evolution of the average coordination of colloidal particles $\bar{n}$ (black continuous line) and the bond-breaking probability $p_\text{break}$ (red dashed line) for four gel samples.
-Symbols relates each sample to its position on the phase diagram Fig~\ref{fig:network_formation}.
-$p_\text{break}$ is measured as the probability of bonds breaking in a time interval of $\Delta t=10\,s$.
+Symbols relate each sample to its position on the phase diagram \textcolor{blue}{in} Fig.~\ref{fig:network_formation}.
+$p_\text{break}$ is measured as the probability of bond breaking in a time interval of $\Delta t=10$ s.
 For high polymer concentration, panels {\bf a} and {\bf b}, $p_\text{break}$ decreases monotonically until it reaches a stationary state at long times, with $\bar{n}$ saturating at an average of less than $5$ neighbours.
@@
 {
-At intermediate times, after percolation has started, hydrodynamic interactions play a fundamental role~\cite{tanaka2000,tanaka2007spontaneous,furukawa2010key}.
+\textcolor{red}{From initial to intermediate times, before} percolation has started, hydrodynamic interactions play a fundamental role~\cite{tanaka2000,tanaka2007spontaneous,furukawa2010key}.
 Without hydrodynamic interactions, as often assumed in simulations, particles have the tendency to aggregate in compact structures and subsequently form thick network structures.
@@
 %%%
-\subsection{Late times: dynamic-arrest vs crystallization}
+%\subsection{Late times: dynamic-arrest vs crystallization}
 
 \begin{figure*}
- \includegraphics{late_structure.pdf}
- \caption{{\bf Structure at late times.} 
-{\bf a,} Structure factors for the late stages of the gelation process. The gray and black lines represents respectively random gelation and compact fractal dimensions.
-{\bf b,} Late stage of the evolution of the local volume fraction around colloidal particles having 12 nearest neighbours.
-{\bf c,} Local volume fraction around a colloidal particle, as a function of the number of nearest neighbours of the particle for the state points in the very late stage. 
+ \includegraphics[width=16cm]{late_structure.pdf}
+ \caption{\textcolor{blue}{{\bf Structural evolution at late times.}} 
+{\bf a,} Structure factors for the late stages of the gelation process. The gray and black lines represent respectively random gelation and compact fractal dimensions.
+{\bf b,} Late stage of the evolution of the local volume fraction \textcolor{blue}{$\phi_{\rm loc}$(12)} around colloidal particles having 12 nearest neighbours.
+{\bf c,} Local volume fraction \textcolor{blue}{$\phi_{\rm loc}$} around a colloidal particle, as a function of the number of nearest neighbours of the particle for the state points in the very late stage.
 The two horizontal straight lines indicate the characteristic volume fractions in attractive systems, crystal ($\sim 0.72$) and glass ($\sim 0.64$).
-Samples are shown with the same colour code as in Fig.~\ref{fig:network_formation}. In particular red circles correspond to the crystallizing sample at $\phi=0.33$ and $cp=0.38\,$mg/g.
+Samples are shown with the same colour code as in Fig.~\ref{fig:network_formation}. In particular red circles correspond to the crystallizing sample at $\phi=0.33$ and $c_p=0.38\,$mg/g.
  }
@@
 
-The samples show significantly different behaviours in the later stages of the phase separation process.
+\textcolor{blue}{The samples share the same early stages of the phase separation process, but show significantly different behaviours in the later stages.}
 To examine this in more detail, we plot in Fig.~\ref{fig:late_structure}a the structure factor for the final stages of the gelation process for all state points.
@@
 The state point at $\phi\approx 0.33$ and $c_p=0.38\,$mg/g reaches close packing at the volume fraction $\approx 0.72$.
-By contrast all other state points, reach close packing at a markedly
+By contrast, all other state points reach close packing at a markedly
 lower volume fraction, which is indeed close to the volume fraction of the attractive glass state~\cite{pham2002multiple}.
@@
 \begin{figure}
- \includegraphics[width=\linewidth]{fig2}
+ \includegraphics[width=10cm]{fig2}
  % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
@@
 We can clearly see not only the percolated nature of both networks, but also the difference in the smoothness of the network surface between amorphous and crystal gels. 
-On the right, slabs of 4.5 $\mu$m thickness and 41 $\mu$m width are plotted, with particles coloured with the same colour scheme as in Fig.~\ref{fig:transitions} red (crystal), green (gas particles), white (liquid particles).
+On the right, slabs of 4.5 $\mu$m thickness and 41 $\mu$m width are plotted, with particles coloured with the same colour scheme as in Fig.~\ref{fig:transitions}: red (crystal), green (gas particles), white (liquid particles).
 We note that very few liquid particles remain in \textbf{b} and the network is almost crystalline. 
@@
 polymer concentration gel, $c_p=1.07$ mg/g; panel b presents the results for the low polymer concentration sample, $c_p=0.38$ mg/g.
-The comparison shows that the new arrest mechanism produces a network structure with bigger pores, smoother surfaces, and thicker strands (panel b) compared
+The comparison shows that the new arrest mechanism produces a \textcolor{blue}{beaded} network structure with bigger pores, smoother surfaces \textcolor{blue}{(consistently with $D=3$)}, and thicker strands (panel b) compared
 to the colloidal gel sample (panel a). The right column shows the individual particle positions for slabs with a thickness of 5 particle diameters. 
 From the figure it is immediately clear that the network strands at low polymer concentration are crystalline. The new arrest mechanism thus
-involve the formation of a crystal-gel network. In order to explain the morphological differences between colloidal gels and crystal-gels,
+involve the formation of a crystal-gel network.
+In order to explain the morphological differences between colloidal gels \textcolor{blue}{(networks made of thin strands)} and crystal-gels \textcolor{blue}{(networks made of thick crystal beads)},
 we now consider the kinetics of formation in our samples.
@@
 \begin{figure}
- \includegraphics{characterisation}
+ \includegraphics[width=8cm]{characterisation}
  % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
@@
 Radius of gyration of crystalline nuclei at the late stages of gelation for the state point $\phi\approx 0.33$ and $c_p=0.38$ mg/g.
-The gray and black straight lines have a slope of $1/2.53$ and $1/3$ respectively
-\textbf{(Inset)} Time evolution of the average size of the crystals (black line) and the number of crystals (red dashed line) for the same state point.
+The gray and black straight lines have a slope of $1/2.53$ and $1/3$ respectively. 
+\textbf{(Inset)} Time evolution of the average size of the crystals \textcolor{blue}{(red dashed line)} and the number of crystals 
+\textcolor{blue}{(black line)} for the same state point.
 } 
@@
 With this analysis we reveal that crystallization events indeed start occurring for the state point $\phi\approx 0.33$ and $c_p=0.38$ mg/g, which has the lowest
-polymer concentration of all state points. This is shown in the inset of Fig.~\ref{fig:crystals}, where the black line indicates the average size of the crystallites, 
-while the red line indicates the average number of crystallites as a function of time. The number of crystallites first rapidly increases as nucleation events
+polymer concentration of all state points. This is shown in the inset of Fig.~\ref{fig:crystals}, where the \textcolor{blue}{red dashed} line indicates the average size of the crystallites, 
+while the \textcolor{blue}{black} line indicates the average number of crystallites as a function of time. The number of crystallites first rapidly increases as nucleation events
 start occurring inside the liquid network, but eventually starts decreasing as the different crystallites grow and merge with each other.
@@
 
-\subsection{Crystal-growth routes}
+%\subsection{Crystal-growth routes}
 
@@
  \centering
- \includegraphics[width=14cm]{fig4}
+ \includegraphics[width=12cm]{fig4}
  % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735
@@
 
-While nucleation always occurs inside the liquid branches of the phase-separating network, crystal-growth can proceed through different growth \textcolor{blue}{routes}.
+While nucleation always occurs inside the liquid branches of the phase-separating network, crystal growth can proceed through different growth \textcolor{blue}{routes}.
 In the following we will show that the growth mechanism of the crystal is not only due to filling of the spinodal liquid network, as was observed in Refs.~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways} for Brownian Dynamics simulations, but also involves the Bergeron process and Ostwald ripening.
@@
 events inside the liquid network (left panel).
-At the same time liquid regions that have not crystallized start evaporating, and the gas phase contributes to the growth of nuclei (middle panel).
+At the same time liquid regions that have not crystallized start evaporating, and the gas phase contributes to the growth of crystal nuclei (middle panel).
 Finally the different nuclei coalesce (right panel).
@@
 especially in determining the morphology of the porous crystal, as we discussed above. In this context the Bergeron process~\cite{glickman2000glossary,morrison2012resilience}
-plays a considerably more important role than Ostwald ripening.
+plays a considerably more important role than Ostwald ripening \textcolor{blue}{and is responsible for the beaded network morphology}.
 
@@
 
-\subsection{Discussion and conclusions}
+%\subsection{Discussion and conclusions}
 The first stages of gelation always involve spinodal decomposition with the formation
@@
 %
-%We will consider two particles bonded if their center-to-center distance is within $12$ pixels. The radius of a colloidal
+%We will consider two particles bonded if their centre-to-centre distance is within $12$ pixels. The radius of a colloidal
 %particle is $5.365$ pixels.
@@
 \textcolor{blue}{Calculation of the structure factor is done with the Hanning window function, to minimize boundary effects.
-To account for imprecision in particle localisation close to contact, we consider two particles bonded if their center-to-center distance is within 2.4~$\mu$m, which is the range of the potential ($\sigma+2R_g$) plus our worse estimate of localisation inaccuracy (0.15~pixels)~\cite{Leocmach2013}.}
+To account for imprecision in particle localisation close to contact, we consider two particles bonded if their centre-to-centre distance is within 2.4~$\mu$m, which is the range of the potential ($\sigma+2R_g$) plus our worse estimate of localisation inaccuracy (0.15~pixels)~\cite{Leocmach2013}.}
 

### 2016-12-15 nature/crystal_gel_r.tex
Update fig:network and fig:transitions with a better color scheme: purple for crystal, gray for liquid and orange for gas. Switch columns and rows in fig:network

--- 2016-12-15 crystal-gel/nature/crystal_gel_r.tex
@@
 \begin{figure}
- \includegraphics[width=10cm]{fig2}
+ \includegraphics[width=10cm]{network}
  % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
-\caption{{\bf Percolated network structures.} {\bf a:} gel; {\bf b:} crystal-gel. The gel state is obtained from the state
-point $\phi\approx 0.33$ and $c_p=1.07$ mg/g at $t=400$ min after sample preparation, while the crystal-gel is obtained at the same $\phi$ but
-at lower polymer concentration, $c_p=0.38$ mg/g at $t=468$ min after sample preparation. On the left, the network is represented in a coarse-grained fashion (see Methods). 
-We can clearly see not only the percolated nature of both networks, but also the difference in the smoothness of the network surface between amorphous and crystal gels. 
-On the right, slabs of 4.5 $\mu$m thickness and 41 $\mu$m width are plotted, with particles coloured with the same colour scheme as in Fig.~\ref{fig:transitions}: red (crystal), green (gas particles), white (liquid particles).
-We note that very few liquid particles remain in \textbf{b} and the network is almost crystalline. 
+\caption{{\bf Percolated network structures.} 
+{\bf a, b:} the network is represented in a coarse-grained fashion (see Methods).
+{\bf c, d:} slabs of 4.5 $\mu$m thickness and 41 $\mu$m width. Particles are drawn to scale and coloured according to their phase. Purple: crystal; gray: liquid; orange: gas. 
+{\bf a, c:} a gel obtained from the state point $\phi\approx 0.33$ and $c_p=1.07$ mg/g at $t=400$ min after sample preparation. 
+{\bf b, d:} a crystal-gel at the same $\phi$ but at lower polymer concentration, $c_p=0.38$ mg/g at $t=468$ min after sample preparation. 
+On the top row we can clearly see not only the percolated nature of both networks, but also the difference in the smoothness of the network surface between amorphous and crystal gels. 
+We note that very few liquid particles remain in \textbf{d} and the network is almost crystalline. 
 } 
@@
 $\phi\approx 0.33$ and $c_p=0.38$ mg/g has different morphological properties than usual colloidal gels.
-We directly compare the two structures in Fig.~\ref{fig:network}, where Gaussian filtering is used to depict the network structure as a continuous field (see Methods).
-Panel a presents the results for high
-polymer concentration gel, $c_p=1.07$ mg/g; panel b presents the results for the low polymer concentration sample, $c_p=0.38$ mg/g.
+We directly compare the two structures in Fig.~\ref{fig:network}a and b, where Gaussian filtering is used to depict the network structure as a continuous field (see Methods).
+Panel~a presents the results for high polymer concentration gel, $c_p=1.07$ mg/g; panel~b presents the results for the low polymer concentration sample, $c_p=0.38$ mg/g.
 The comparison shows that the new arrest mechanism produces a \textcolor{blue}{beaded} network structure with bigger pores, smoother surfaces \textcolor{blue}{(consistently with $D=3$)}, and thicker strands (panel b) compared
-to the colloidal gel sample (panel a). The right column shows the individual particle positions for slabs with a thickness of 5 particle diameters. 
+to the colloidal gel sample (panel a). The bottom row shows the individual particle positions for slabs with a thickness of 5 particle diameters. 
 From the figure it is immediately clear that the network strands at low polymer concentration are crystalline. The new arrest mechanism thus
@@
 Although crystal clusters are isolated in this movie, they are bridged by the dense colloidal liquid phase and involved in the percolated gel network, as can be seen in Fig.
-\ref{fig:network}b.
+\ref{fig:network}b and d.
 We can see that crystal clusters grow by condensation from the colloidal gas phase. 
@@
  \centering
- \includegraphics[width=12cm]{fig4}
+ \includegraphics[width=16cm]{transitions}
  % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735
@@
 Particles are drawn to scale and coloured according to their phase.
-Red: crystal; white: liquid; green: gas.
+Purple: crystal; gray: liquid; orange: gas.
 {\bf b,} Fraction of particles for each phase. Gray vertical lines indicate the times shown in panel a. 

### 2016-12-15 crystal-gel/data/slices_article.py
Update fig:network and fig:transitions with a better color scheme: purple for crystal, gray for liquid and orange for gas. Switch columns and rows in fig:network

@@ -8,7 +8,7 @@ from matplotlib.pyplot import imread, imsave
 #0 -> BCC
 #1 -> HCP
 #2 -> FCC
-label = 'ILM'
+label = '_article'
 
 for sample, times in zip(['', '_highcp'], [[120, 390, 890], [800]]):
     for t in times:
@@ -23,7 +23,7 @@ for sample, times in zip(['', '_highcp'], [[120, 390, 890], [800]]):
             for x,y,z in pos[phase<3]
             ]+[Texture(Pigment(color=(0.694,  0.145,  0.623)))])
         f.write(crystals)
-        for i, c in zip([3,4,5], [(0.847,0.57,0.81), (0.847,0.57,0.81), (1.,  0.455,  0.156)]):
+        for i, c in zip([3,4,5], [(0.5,0.5,0.5), (0.5,0.5,0.5), (1.,  0.455,  0.156)]):
             f.write(Union(*[
                 Sphere( (x,y,z), 5)
                 for x,y,z in pos[phase==i]

### 2016-12-15 nature/crystal_gel_r.tex
recall that high Cp do no undergo stress-driven ageing

@@
 
-A direct analysis of the colloidal positions at late times confirms that indeed the new structure obtained at
-$\phi\approx 0.33$ and $c_p=0.38$ mg/g has different morphological properties than usual colloidal gels.
+A direct analysis of the colloidal positions at late times confirms that indeed the new structure obtained at $\phi\approx 0.33$ and $c_p=0.38$ mg/g has different morphological properties than usual colloidal gels.
 We directly compare the two structures in Fig.~\ref{fig:network}a and b, where Gaussian filtering is used to depict the network structure as a continuous field (see Methods).
-Panel~a presents the results for high polymer concentration gel, $c_p=1.07$ mg/g; panel~b presents the results for the low polymer concentration sample, $c_p=0.38$ mg/g.
+Panel~a presents the results for high polymer concentration gel, $c_p=1.07$ mg/g \textcolor{blue}{that never undergoes stress-driven ageing}; panel~b presents the results for the low polymer concentration sample, $c_p=0.38$ mg/g.
 The comparison shows that the new arrest mechanism produces a \textcolor{blue}{beaded} network structure with bigger pores, smoother surfaces \textcolor{blue}{(consistently with $D=3$)}, and thicker strands (panel b) compared

### 2016-12-16 crystal-gel/nature/crystal_gel3.tex
John edits.

@@
 
-\title{A novel physical mechanism for the spontaneous formation of porous crystals via viscoelastic phase separation}
+%\title{A novel physical mechanism for the spontaneous formation of porous crystals via viscoelastic phase separation}
+\title{\textcolor{red}{Formation of porous crystals via viscoelastic phase separation}}
 
@@
 With a new confocal microscopy protocol, we unveil a novel crystallization pathway to sponge-like porous structures in a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation.
-\textcolor{blue}{
-We follow the entire kinetics of phase separation, from the homogeneous phase, percolation, stress driven aging, and until the formation of porous crystals.}
+\textcolor{red}{
+We follow the entire kinetics of phase separation from the homogeneous phase to one of its arrested states: at high-polymer concentration this is the gel-phase, but at low-polymer concentration ($c_p\sim 0.38$ mg/g), we find that the liquid phase undergoes a structural rearrangement to a state with higher connectivity, which at high volume fractions ($\phi\sim 0.33$) crystallizes into a porous structure.}
 We show that crystallization leads to a network structure different from the original phase separation pattern.
@@
 lower volume fraction, which is indeed close to the volume fraction of the attractive glass state~\cite{pham2002multiple}.
-This result clearly indicates that a single state point follows a different arrest mechanism, in which phase separation is arrested by crystallization and not by glassiness.
+\textcolor{red}{These structural analysis provide some evidence that the state point $\phi\approx 0.33$ and $c_p=0.38\,$mg/g could be following a different arrest mechanism, in which phase separation is arrested by crystallization and not by glassiness.
+In the reminder of this manuscript we investigate this state more closely, to confirm these early results.}
 

### 2016-12-16 nature/response3.tex
John edits.

\documentclass[11pt,a4paper]{article}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage{xcolor}
\usepackage{setspace}

\usepackage[top=35mm,bottom=35mm,left=30mm,right=30mm]{geometry}

\newenvironment{referee}%
{\bigskip\singlespacing\bf}%
{\par\bigskip}

\begin{document}

%\doublespacing

%\doublespacing

\noindent
{\bf Dear Pep,}
\vskip 0.3cm
We would like to thank you for your providing us with two Referee reports for our manuscript (NM15113574). We understand your decision to reject the manuscript, but, after careful consideration of the Referee reports, we couldn't avoid thinking that they are mostly supportive of our work, and the only Referee (Referee \#1) that is somewhat against publication is basing his argument on a probable misunderstanding of the true novelty of the first part of our work.

We decided to write to you again to clarify that the Referee might have missed some very important points. In general, both Referees seem to have appreciated the improvements
we made after the first round of reviews: Reviewer \#1 calls the work \emph{very beautiful and interesting}, while Referee \#3 goes into the details of explaining how we actually satisfactorily clarified his previous doubts, with just a couple of minor points left. Reviewer \#2 was very positive during the first round, but unfortunately didn't submit a second report.
Overall we were able to find only two criticism in the reports.
\begin{itemize}
 \item Referee \#1 laments that the first part of the paper describes spinodal decomposition without showing anything new from previous studies (but he admits the novelty of the study of the different crystallisation channels). But this is not the case, as the crystallisation of the liquid network requires an additional step, called stress-driven aging, that was never observed experimentally before.
In our new version of the manuscript we spent considerable time developing a new analysis, that shows how spinodal decomposition proceeds differently depending on the polymer concentration, and that the aging of the network due to mechanical stress is a necessary condition for the subsequent crystallization process. We think that this completely invalidates the point of the Referee, because in addition to different crystallization channels, our study unveils a new link between the aging of the liquid network and the subsequent crystallization process.
 \item Referee \#3 suggests that we better acknowledge that the crystal-gel is found only in a quite specific region of the phase diagram, and this also became the main motivation for the rejection of our manuscript. We also fundamentally disagree with this idea. It's true that crystal-gel states don't appear everywhere in the phase diagram, but this doesn't mean that they are hard to obtain. Rather the opposite, there is a reproducible region around the critical point were these states can be consistently obtained, and we were able to reproduce these findings both in our new cell design, and in ordinary capillaries, showing that the region of crystal-gel formation is well within the experimental range. Moreover the extent of this region can be widened by changing the location of the critical point, which in our system is controlled by the length of the non-adsorbing polymer: moving the critical point to lower polymer concentrations opens the window where bonds can rearrange before the intervening glass transition. So we argue that the region of formation of crystal-gel porous structures is not only easily accessible, but also controllable.
\end{itemize}

We hope that you'll see that both criticisms are not indicative of any weaknesses in our work. We also took advantage of the new suggestions to improve the manuscript even further. In particular, concerning the point raised by Referee \#1 we have introduced a new type of analysis, which looks at the bond-breaking probability as a function of time, during the whole process of spinodal decomposition. Our new results show that, at low polymer-concentrations, the network undergoes a structural reorganisation, which results in more compact structures, that are a necessary condition for the appearance of the first crystalline nuclei. To our knowledge this is the first experimental demonstration of this process: it is not captured by computer simulations, as they generally neglect hydrodynamic interactions which play a fundamental role in this kinetic process, and no previous experiment has had the capability of tracking particles during the early stages of spinodal decomposition.

We are rather positive that the new version of the manuscript would put the Referees in a position to better understand the full extent of our results, and
so we hope that you might give it another chance. If that is is not the case, we are still grateful for the opportunity that the comments gave us for improving the manuscript in a meaningful way.

\vskip 0.8cm

\noindent
{\bf Yours Sincerely,
\vskip 0.3cm
\indent Hideyo Tsurusawa\\
\indent John Russo\\
\indent Mathieu Leocmach\\
\indent Hajime Tanaka}

\clearpage

\noindent
\begin{Large}
{\bf Response to Referee \#1}
\end{Large}

\vskip 0.5cm

We would like to thank the Referee for her/his comments. We address these points in the following response.

\begin{referee}
The authors have improved their manuscript according to the comments of all referees, which has increased the readability of the paper. The novelty of the paper is also much better explained, which is in fact described not earlier than on page 7 of the manuscript.
The first 7 pages of the paper describes one of the arrest mechanisms of spinodal decomposition due to crystallization, which was already observed in experiments and simulations.
\end{referee}

We thank the Referee for her/his appreciation of the improvements made during our last revision of the manuscript. The Referee comments on the first 7 pages being not original, and simply describing the process of spinodal decomposition as previously observed in experiments and simulations. But we believe that our study offers novel insights also in the description of the early stages of the phase separation process. This is a very important point, as there is a clear link between the early-stage kinetics of the liquid network, and the late-stage crystallization process. In order to highlight this connection, in this new version of the manuscript we introduce a new type of analysis. In summary, we show that our work represents the first experimental demonstration of a process called \emph{stress-driven aging}. Most past simulation works have neglected, or not fully taken into account, the role of hydrodynamic interactions (HI) in the phase separation process. As was shows for example in Ref.~\cite{furukawa2010key}, the phase separation process without HI produces rather compact structures, which would then easily crystallize to form compact nuclei. This is not what happens if one take HI into consideration. Due to the incompressibility condition, the liquid network has a lower-connectivity, and the viscoelastic phase separation results in the accumulation of mechanical stress on these liquid bonds. To things can happen depending on the polymer concentration. At high-polymer concentration, the bonds are not-easily broken, and the network which is first formed cannot undergo structural reorganisation. This inevitably leads to gelation. On the other side, if the polymer concentration is low enough, the liquid network can reorganise with a rather sharp transition to a more compact network. This condition is necessary for the appearance of the first crystalline nuclei. In our new manuscript, we are able to follow the kinetics of these processes by looking at the time evolution of the bond-breaking probability. The different network states are in fact characterised by different bond-breaking probabilities and by different connectives. In Fig.~2a,b we show these quantities for high-polymer concentration samples, where spinodal decomposition is the only process that occurs. At low-polymer concentrations (Fig.~2c,d) instead, the network undergoes a sharp transition to a state with lower bond-breaking probability and higher connectivity. This transition is the stress-driven aging process, and increases the connectivity of the network. Despite being theoretically predicted, Ref.~\cite{tanaka2000,tanaka2007spontaneous}, this transition was not directly observed before.
We agree that the previous version of the manuscript was lacking on this part, so the Referee comments inspired us to develop the new analysis presented in Fig.2.

\begin{referee}
What is new is that crystallization not only takes place in the spinodal liquid branches, but can also occur via the Bergeron process in which the liquid evaporates, and the gas phase crystallizes again or via Ostwald ripening, i.e., melting of crystallites into the gas phase and recrystallization. Both processes make sense and can be explained that the crystallites should be surrounded by the gas phase due to the non-existence of a liquid-crystal coexistence as already speculated by Renth et al. It would be worthwhile to mention this more explicitly in the paper. In addition, the authors mention on page 3, that for confocal microscopy studies on colloidal gelation, the colloid-polymer mixture have so far been used. There are also confocal microscopy experiments performed on oppositely charged colloids,
which is worthwhile mentioning as well, Sanz et al., J. Phys. condens. Matt. 20, 494247 (2008), Sanz et al., J. Phys. Chem. B 112, 10861 (2008), where the effect of range and attraction strength have been studied on the different arrest mechanisms of the gels, including the crystallization of spinodal structures. Although the work is very beautiful and interesting, the referee has still strong doubts if the work provides sufficient novel insights to warrant publication in Nature Materials. First of all, the title suggests that a completely new route is found to form porous crystals. Crystallisation of spinodal networks has already been found in previous works and so the present work doesn't provide a new route, only another crystallization mechanism when the crystals get larger than the thickness of the branches. Secondly, the abstract suggests a new confocal microscopy protocol to study phase separation, which is in fact based on a salt injection that changes the interactions in
situ. Finally, the last sentence of the abstract is also misleading as this new route is not a single-step pathway to form nanoporous crystals, it involves spinodal decomposition, crystallization of the branches, and subsequently evaporation of the liquid and crystallization or melting of crystalline parts and recrystallization.
\end{referee}

In the new version of the manuscript we have tried to incorporate all the Referee's previous suggestions. We have added the recommended references to the new version of the manuscript~\textcolor{red}{I remember the sensei saying that he had included some of the References; do we need any else?}. We simplified the title~\textcolor{red}{This is not necessary, of course, but as of lately I'm leaning towards simple titles, so I let you decide}. Regarding the use of a novel salt--injection we believe this to be the first time this cell design has been used to study the formation of porous crystals in gels. \textcolor{red}{Could you please confirm and comment this?}.

\clearpage

\noindent
\begin{Large}
{\bf Response to Referee \#3}
\end{Large}

\vskip 0.5cm

We would like to thank the Referee for her/his comments. In the following we respond to the new points raised.

\begin{referee}
I was aware of the different frame rates on the two movies but I agree that the time stamps are helpful. Figure S5 addresses my concern about reproducibility at the single state point, and the different geometry itself is helpful too. I think the authors should be up front in the manuscript that the mechanism discussed is found in a very specific window (from the figures and SI, I think it is $\phi\sim 0.33$ and $c_p~0.39$ (plus minus) 0.01 mg/g) by moving this point into the abstract.
\end{referee}

In this new version of the manuscript we follow the Referee's recommendation and write explicitly in the abstract the condition for the formation of crystal-gels. We stress that the conditions for crystal-gel formation are not stringent and well within the experimental accessible states, as we previously demonstrated in our reproducibility tests inside a capillary.

\begin{referee}
The addition of the SI movie 3 and the discussion of Figure 2b addresses this concern for me. It's the percolation that is critical to the authors claim of a crystal gel - the percolated structure needs to persist over long times, otherwise, for example, the gel wouldn't have a finite low frequency elastic modulus. I think SI move 3 does show this.

3) The data quality of Figure 1c, Figure 1d appear insufficient to classify the cp = 0.38 mg/mg state point as different from the others. The authors should add replications, errors, and significance testing if they wish to distinguish the dimensionality (Fig 1c) and maximum volume fraction (Fig 1d) of the 0.38 mg/ml state point from the others.

I continue to think that the evidence for different behavior in the cp = 0.38 mg/g state point from Figure 1 c,d is weak. Supplemental Figure 3 is more convincing that the final state points are different. 
\end{referee}

We fully agree that the evidence from Fig.~1 c,d was not conclusive, but the goal of those figures was not to provide conclusive evidence, but to provide a comprehensive study of the early-stages of spinodal decomposition. In this context the Fig.~1 c,d only provide hints that the cp = 0.38 mg/g state point behaves differently from other state points. The rest of the manuscript and the Supplementary materials is devoted to providing the conclusive evidence that that is the case. We have made this more clear in the new manuscript.

\textcolor{red}{The Referee really likes Supplementary Fig3, so at this point if we want to answer him fully, we should probably replace Fig. 3c with Supplementary Figure 3. What do you think?}

\begin{referee}
(Yet, in the new section, the authors write: "This result clearly indicates that a single state point follows a different arrest mechanism, in which phase separation is arrested by crystallization and not by glassiness." Is that really true? It seems overstated. Don't Figure 1 c,d and SF3 just establish that the end point is different? They don't say anything about the mechanism, which would require kinetic analysis, as comes later. And it doesn't say anything about arrest, which would require dynamics.)
\end{referee}

In the new version of the manuscript we have removed that sentence, and replaced it with:

\emph{
These structural analysis provide some evidence that the state point $\phi\approx 0.33$ and $c_p=0.38\,$mg/g could be following a different arrest mechanism, in which phase separation is arrested by crystallization and not by glassiness. In the reminder of this manuscript we investigate this state more closely, to confirm these early results.}

We also agree that our previous manuscript didn't say much about the actual kinetic mechanism by which crystal-gels form. This point was also raised by Referee \#1 in noting that we didn't provide any additional insight in the phase separation process. In order to fully address this issue, we have implemented a kinetic analysis of the trajectories at all state points, computing the probability of bond-breaking between consecutive frames. This new kinetic analysis reveals that at low-polymer concentrations a new mechanism (stress-driven aging) is taking place, which has the effect of increasing the average coordination of the colloidal particles in the liquid phase.
We believe this to be a novel observation of this phenomena. Most past simulation works have neglected, or not fully taken into account, the role of hydrodynamic interactions (HI) in the phase separation process. As was shows for example in Ref.~\cite{furukawa2010key}, the phase separation process without HI produces rather compact structures, which would then easily crystallize to form compact nuclei. This is not what happens if one take HI into consideration. Due to the incompressibility condition, the liquid network has a lower-connectivity, and the viscoelastic phase separation results in the accumulation of mechanical stress on these liquid bonds. To things can happen depending on the polymer concentration. At high-polymer concentration, the bonds are not-easily broken, and the network which is first formed cannot undergo structural reorganisation. This inevitably leads to gelation. On the other side, if the polymer concentration is low enough, the liquid network can reorganise with a rather sharp transition to a more compact network. This condition is necessary for the appearance of the first crystalline nuclei. In our new manuscript, we are able to follow the kinetics of these processes by looking at the time evolution of the bond-breaking probability. The different network states are in fact characterised by different bond-breaking probabilities and by different connectives. In Fig.~2a,b we show these quantities for high-polymer concentration samples, where spinodal decomposition is the only process that occurs. At low-polymer concentrations (Fig.~2c,d) instead, the network undergoes a sharp transition to a state with lower bond-breaking probability and higher connectivity. This transition is the stress-driven aging process, and increases the connectivity of the network. Despite being theoretically predicted, Ref.~\cite{tanaka2000,tanaka2007spontaneous}, this transition was not directly observed before.

\begin{referee}
 The revised Figure 4, along with Figure S4, are very helpful to resolving the relative contributions of the three crystalline pathways. In Figure 4b,c, I still think that surface particles should somehow be distinguished - they represent a distinct state and a kind of consistency check, because the different crystallization pathways should all be mediated by this state. Without including the surface states in the primary figure, I don't see how the reader can inspect Figure S4, and get to the 54/34/11 abundances for direct, Bergeron and Ostwald processes. These abundances are the crucial claim of the paper.

In summary the confirmatory experiments (Fig S5 and Movie S3), the additional analysis of the crystalline state (Fig 3S) and the full transition analysis (rev Fig 4 and F S4) offer significant new support of the authors' claims; the authors have something new relative to prior literature. At this point, I think that the authors should: (i) better acknowledge that what is observed here is found in a quite specific region of the state diagram ($phi\sim 0.33$ and $c_p\sim 0.39$ {plus minus} 0.01 mg/g); (ii) better connect Figure 4 and Figure S4 by accounting for surface states in the former figure.

\end{referee}

\textcolor{red}{The Referee insists on having surface particles represented in Fig.6b. Even if I think that the figure becomes more complicated this way, maybe we should just update the figure. I'm sending Mathieu the dataset with the fraction of surface particles and liquid particles in case you agree on this change. If you do, I'll also update the description in the manuscript.}

In the new version of the manuscript we follow the Referee's recommendation and include in Fig. 6b the fraction of surface particles, and update the description accordingly.
% We remind that the 54/34/11 abundances was obtained by looking at the phase history of each colloidal particle. Fig. 6b and Supplementary Fig. 4 don't contain information about the phase history of individual particles, just about the total number of particles in each phase, and the time-dependence of the transition probabilities. Of course these quantities are closely connected, and we hope to have provided a detailed description of the process.

\clearpage

\bibliographystyle{naturemag2}
\bibliography{biblio}

\end{document}

### 2016-12-17 crystal-gel/nature/crystal_gel_r.tex

@@
 
-\title{A novel physical mechanism for the spontaneous formation of porous crystals via viscoelastic phase separation}
+\title{Formation of porous crystals via viscoelastic phase separation}
 
@@
 %But while the former process has been studied in detail, the latter has never been observed experimentally at a single-particle level and it is thus lacking a microscopic understanding. 
-With a new confocal microscopy protocol, \textcolor{blue}{we successfully follow the entire kinetics of phase separation at a single-particle level, from the homogeneous phase, percolation, stress driven ageing, and until the formation of the final gel structure.
-This allows us to unveil a novel crystallization pathway to sponge-like porous crystal structures in a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation and to reveal its physical mechanism.}
+With a new confocal microscopy protocol, \textcolor{blue}{we follow the entire kinetics of phase separation, from the homogeneous phase, percolation, stress driven ageing, and until the formation of the final structure.
+This allows us to unveil a novel crystallization pathway to sponge-like porous crystal structures in a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation and to reveal its physical mechanism.
 %\textcolor{blue}{We follow the entire kinetics of phase separation, from the homogeneous phase, percolation, stress driven aging, and until the formation of porous crystals.}
-We show that crystallization leads to a network structure different from the original phase separation pattern. 
+We clarify the physical requirements necessary for this to happen and show that crystallization leads to a network structure different from the original phase separation pattern.} 
 Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface, i.e. de-sublimation, driving evaporation of nearby liquid. 
@@
 We argue that similar crystal-gel states can be formed as the result of the gas-liquid phase separation of any crystallizable components, such as monoatomic and molecular systems, if dynamics of the liquid phase is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
-This mechanism will provide a novel single-step pathway to form nano-porous crystals of metals and semiconductors without dealloying, %\cite{erlebacher2001evolution},
+This mechanism will provide a novel pathway to form nano-porous crystals of metals and semiconductors without dealloying, %\cite{erlebacher2001evolution},
 which may be important for catalytic, optical, sensing, and filtration applications. 
@@
 Under usual conditions, the nucleation stage involves a very small number of molecules, of the order of 10-1000 molecules.
-This has severely limited the possibility to \textcolor{blue}{experimentally} observe the  \textcolor{blue}{crystallization} process directly, and has also posed limits on models based on macroscopic thermodynamic properties.
+This has severely limited the possibility to observe the  \textcolor{blue}{crystallization} process directly, and has also posed limits on models based on macroscopic thermodynamic properties. 
 The inability to directly observe the  \textcolor{blue}{crystallization} process is also a limit to our understanding of its kinetic pathway. 
@@
 of the dense liquid phase~\cite{pusey1993dynamics,piazza1994phase,ilett1995phase,verhaegh1997transient,tanaka1999colloid,foffi2002,buzzaccaro2007sticky,zaccarelli2007,lu2008gelation,zaccarelli2008gelation,testard2011}.
-\textcolor{blue}{For thin colloidal gels}, the dynamical stabilization is due to percolation of locally favoured structures, which are locally stable non-crystalline structures \cite{royall2008g}.
+More precisely, the dynamical stabilization is due to percolation of locally favoured structures, which are locally stable non-crystalline structures \cite{royall2008g}.
 
@@
 
-Unlike the above standard scenario of colloidal gelation, it is known that phase separation can also be arrested by crystallization, if the process takes place below the melting point of \textcolor{blue}{one of the two separating phases}~\cite{tanaka1985new}.
+\textcolor{blue}{Unlike the above standard scenario of colloidal gelation, it is known that phase separation can also be accomapnied by crystallization \cite{poon1999cpm,renth2001phase}, if the process takes place below the melting point of one of the two separating phases}~\cite{tanaka1985new}.
 Indeed the possibility of a different class of gels, which are stabilized by crystallization,
@@
  \caption{{\bf Stress driven aging.} Time evolution of the average coordination of colloidal particles $\bar{n}$ (black continuous line) and the bond-breaking probability $p_\text{break}$ (red dashed line) for four gel samples.
-Symbols relate each sample to its position on the phase diagram \textcolor{blue}{in} Fig.~\ref{fig:network_formation}.
-$p_\text{break}$ is measured as the probability of bond breaking in a time interval of $\Delta t=10$ s.
+Symbols relates each sample to its position on the phase diagram \textcolor{blue}{in} Fig.~\ref{fig:network_formation}.
+$p_\text{break}$ is measured as the probability of bonds breaking in a time interval of $\Delta t=10$ s.
 For high polymer concentration, panels {\bf a} and {\bf b}, $p_\text{break}$ decreases monotonically until it reaches a stationary state at long times, with $\bar{n}$ saturating at an average of less than $5$ neighbours.
@@
 {
-\textcolor{red}{From initial to intermediate times, before} percolation has started, hydrodynamic interactions play a fundamental role~\cite{tanaka2000,tanaka2007spontaneous,furukawa2010key}.
+At intermediate times, after percolation has started, hydrodynamic interactions play a fundamental role~\cite{tanaka2000,tanaka2007spontaneous,furukawa2010key}.
 Without hydrodynamic interactions, as often assumed in simulations, particles have the tendency to aggregate in compact structures and subsequently form thick network structures.
@@
 
-\textcolor{blue}{The samples share the same early stages of the phase separation process, but show significantly different behaviours in the later stages.}
+\textcolor{blue}{The samples shares the same early stages of of the phase separation process, but show significantly different behaviours in the later stages.}
 To examine this in more detail, we plot in Fig.~\ref{fig:late_structure}a the structure factor for the final stages of the gelation process for all state points.
@@
 lower volume fraction, which is indeed close to the volume fraction of the attractive glass state~\cite{pham2002multiple}.
-This result clearly indicates that a single state point follows a different arrest mechanism, in which phase separation is arrested by crystallization and not by glassiness.
+\textcolor{blue}{These structural analysis provide some evidence that the state point $\phi\approx 0.33$ and $c_p=0.38\,$mg/g could be following a different arrest mechanism, in which phase separation is arrested by crystallization and not by glassiness.
+In the reminder of this manuscript we investigate this state more closely, to confirm these early results.}
 
@@
 \begin{figure}
- \includegraphics[width=10cm]{fig2}
+ \includegraphics[width=10cm]{network}
  % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
@@
 polymer concentration gel, $c_p=1.07$ mg/g; panel b presents the results for the low polymer concentration sample, $c_p=0.38$ mg/g.
-The comparison shows that the new arrest mechanism produces a \textcolor{blue}{beaded} network structure with bigger pores, smoother surfaces \textcolor{blue}{(consistently with $D=3$)}, and thicker strands (panel b) compared
+The comparison shows that the new arrest mechanism produces a network structure with bigger pores, smoother surfaces \textcolor{blue}{(consistently with $D=3$)}, and thicker strands (panel b) compared
 to the colloidal gel sample (panel a). The right column shows the individual particle positions for slabs with a thickness of 5 particle diameters. 
 From the figure it is immediately clear that the network strands at low polymer concentration are crystalline. The new arrest mechanism thus
-involve the formation of a crystal-gel network.
-In order to explain the morphological differences between colloidal gels \textcolor{blue}{(networks made of thin strands)} and crystal-gels \textcolor{blue}{(networks made of thick crystal beads)},
+involve the formation of a crystal-gel network. In order to explain the morphological differences between colloidal gels and crystal-gels,
 we now consider the kinetics of formation in our samples.
@@
  \centering
- \includegraphics[width=12cm]{fig4}
+ \includegraphics[width=12cm]{transitions}
  % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735
@@
 Ostwald ripening accounts for 11\% of particle trajectories in which crystal particles transition to the gas state before returning to the crystal state.
+\textcolor{blue}{Here we note that the Bergeron and Ostwald ripening processes both take place only when the crystallites are
+surrounded by the gas phase due to the non-existence of a liquid-crystal coexistence.
+Such a gas-crystal coexistence was speculated  by Poon and his coworkers on the basis of free-energy argument \cite{poon1999cpm,renth2001phase}.}
 
@@
 especially in determining the morphology of the porous crystal, as we discussed above. In this context the Bergeron process~\cite{glickman2000glossary,morrison2012resilience}
-plays a considerably more important role than Ostwald ripening \textcolor{blue}{and is responsible for the beaded network morphology}.
+plays a considerably more important role than Ostwald ripening.
 
@@
 Depending on the polymer concentration, there are two possible arrest mechanisms. 
-(a) Crystallization, observed at $\phi\approx 0.33$ and $c_p=0.38$ mg/g, small crystalline nuclei appear inside the liquid network, reach the surface 
+\textcolor{blue}{(a) Crystallization: small crystalline nuclei appear inside the liquid network, reach the surface 
  of the liquid branches, and then grow by addition of particles from the gas phase. The final structure is a network of crystal droplets,
  as confirmed by the fractal dimension of the branches, the volume fraction of the particles within the branches, and bond orientational analysis. 
-(b) Dynamic arrest, particle arrest when the dynamics inside the liquid branch becomes slow, which should happen at the intersection of the
- glass line with the liquid side of the coexistence curve.
+(b) Dynamic arrest: particle arrest when the dynamics inside the liquid branch becomes slow, which should happen at the intersection of the  glass line with the liquid side of the coexistence curve.}
+
+\textcolor{blue}{In our system, mechanism (a) is operative around $\phi\approx 0.33$ and $c_p=0.38$ mg/g. 
+The physical conditions required for the crystal-gel formation revealed in our study indicates that the extent of the $\phi$-$c_p$ region where mechanism (a) is operative can be widened by changing the location of the critical point, which in our system is controlled by the size of the non-adsorbing polymer: moving the critical point to lower polymer concentrations opens the window where bonds can rearrange before the intervening glass transition.
+So we argue that the region of formation of crystal-gel porous structures is not only easily accessible, but also controllable.}
 

### 2016-12-17 crystal-gel/nature/response3.tex

--- 2016-12-16 nature/response3.tex
@@
 \noindent
-{\bf Dear Pep,}
-\vskip 0.3cm
+{\bf Dear Pep and/or a newly assigned Editor in Nature Materials,}
+\vskip 0.7cm
 We would like to thank you for your providing us with two Referee reports for our manuscript (NM15113574).
-We understand your decision to reject the manuscript, but, after careful consideration of the Referee reports, we couldn't avoid thinking that they are mostly supportive of our work, and the only Referee (Referee \#1) that is somewhat against publication is basing his argument on a probable misunderstanding of the true novelty of the first part of our work.
+We understand your decision to reject the manuscript, but, after careful consideration of the Referee reports, we couldn't avoid thinking that they are mostly supportive of our work, and the only Referee (Referee \#1) that is somewhat against publication is basing his/her argument on a probable misunderstanding of the true novelty of the first part of our work.
 
@@
 \begin{itemize}
- \item Referee \#1 laments that the first part of the paper describes spinodal decomposition without showing anything new from previous studies (but he admits the novelty of the study of the different crystallisation channels).
-But this is not the case, as the crystallisation of the liquid network requires an additional step, called stress-driven aging, that was never observed experimentally before.
-In our new version of the manuscript we spent considerable time developing a new analysis, that shows how spinodal decomposition proceeds differently depending on the polymer concentration, and that the aging of the network due to mechanical stress is a necessary condition for the subsequent crystallization process.
-We think that this completely invalidates the point of the Referee, because in addition to different crystallization channels, our study unveils a new link between the aging of the liquid network and the subsequent crystallization process.
+ \item Referee \#1 laments that the first part of the paper describes spinodal decomposition without showing anything new from previous studies (but he/she admits the novelty of the study of the different crystallisation channels).
+But this is not the case, as the crystallisation of the liquid network requires an additional step, called stress-driven ageing, that was never observed experimentally before.
+In our new version of the manuscript we spent considerable time developing a new analysis, that shows how spinodal decomposition proceeds differently depending on the polymer concentration, and that the ageing of the network due to mechanical stress is a necessary condition for the subsequent crystallization process.
+We think that this completely invalidates the point of the Referee, because in addition to different crystallization channels, our study unveils a new link between the ageing of the liquid network and the subsequent crystallization process.
  \item Referee \#3 suggests that we better acknowledge that the crystal-gel is found only in a quite specific region of the phase diagram, and this also became the main motivation for the rejection of our manuscript.
@@
 
+\vskip 0.3cm
+
+Since the manuscript now includes the new analysis mentioned above and is expanded with 6 figures, we would like to resubmit it as an Article rather than a Letter. 
+\textit{
+\vskip 0.3cm}
+
 We are rather positive that the new version of the manuscript would put the Referees in a position to better understand the full extent of our results, and
 so we hope that you might give it another chance. If that is is not the case, we are still grateful for the opportunity that the comments gave us for improving the manuscript in a meaningful way.
+
+\vskip 0.5cm
+
+Warm thanks to you for your kind consideration and cooperation.
 
@@
 
-We would like to thank the Referee for her/his comments. We address these points in the following response.
+We would like to thank the Referee for her/his comments. 
+We are glad to see that he/she called the work \emph{very beautiful and interesting}. 
+We address the concerns of the Referee in the following response.
 
@@
 Most past simulation works have neglected, or not fully taken into account, the role of hydrodynamic interactions (HI) in the phase separation process.
-As was shows for example in Ref.~\cite{furukawa2010key}, the phase separation process without HI produces rather compact structures, which would then easily crystallize to form compact nuclei.
+As was shown for example in Ref.~\cite{furukawa2010key}, the phase separation process without HI produces rather compact structures, which would then easily crystallize to form compact nuclei.
 This is not what happens if one take HI into consideration.
 Due to the incompressibility condition, the liquid network has a lower-connectivity, and the viscoelastic phase separation results in the accumulation of mechanical stress on these liquid bonds.
-To things can happen depending on the polymer concentration.
+Two things can happen depending on the polymer concentration.
 At high-polymer concentration, the bonds are not-easily broken, and the network which is first formed cannot undergo structural reorganisation.
@@
 Despite being theoretically predicted, Ref.~\cite{tanaka2000,tanaka2007spontaneous}, this transition was not directly observed before.
-We agree that the previous version of the manuscript was lacking on this part, so the Referee comments inspired us to develop the new analysis presented in Fig.2.
+We agree that the previous version of the manuscript was lacking on this part, so the Referee comments inspired us to develop the new analysis presented in new Fig. 2.
 
@@
 In the new version of the manuscript we have tried to incorporate all the Referee's previous suggestions. 
-We have added the recommended references to the new version of the manuscript~\textcolor{red}{I remember the sensei saying that he had included some of the References; do we need any else?}.
-We simplified the title~\textcolor{red}{This is not necessary, of course, but as of lately I'm leaning towards simple titles, so I let you decide}.
-Regarding the use of a novel salt--injection we believe this to be the first time this cell design has been used to study the formation of porous crystals in gels.
-\textcolor{red}{Could you please confirm and comment this?}.
+Following the suggestion, we have explicitly mentioned the work of Poon and coworkers in connection to the Bergeron and Ostwald ripening processes.  
+We have also added the two recommended references by Sanz et al. to the new version of the manuscript. 
+On the doubt of the referee concerning the novelty of our work, we hope that our new analysis explained above provides a 
+new physical insight on the physical condition required for the formation of porous crystal gels besides its new kinetic route. 
+We also simplified the title to avoid misunderstanding of our message.
+Regarding the use of a novel salt--injection we believe this to be the first time this cell design has been applied for confocal microscopy observation at a single-particle level and
+used to study the formation of porous crystals in gels, 
+which is the key to access the early stage of viscoelastic phase separation at a single particle resolution with full particle tracking. 
+For example, the new analysis shown in new Fig. 2 is only possible with this new experimental protocol and full particle tracking.  
+Finally, the last sentence has been modified, following the kind suggestion of the Referee.
+
+\vskip 1cm
+
+We hope that the Referee will find that the new analysis have revealed a new mechanism behind 
+crystal-gel formation and increased the novelty of our manuscript considerably and 
+the revised manuscript is now suitable for publication in Nature Materials.
 
@@
 
-We would like to thank the Referee for her/his comments. In the following we respond to the new points raised.
+We would like to thank the Referee for his/her comments. In the following we respond to the new points raised.
 
@@
 
-In this new version of the manuscript we follow the Referee's recommendation and write explicitly in the abstract the condition for the formation of crystal-gels.
+Following the Referee's recommendation, we have put the following sentence in the abstract: We clarify the physical requirements necessary for this to happen.
+Although this does not explicitly specify the points we observed crystal gels, our new analysis (see below)
+has revealed the physical conditions required for observing crystal-gel formation.
+We have also discussed this issue in our revised manuscript with the new analysis shown in a new Fig.
+2.
 We stress that the conditions for crystal-gel formation are not stringent and well within the experimental accessible states, as we previously demonstrated in our reproducibility tests inside a capillary.
+Our new results show that, at low polymer-concentrations, the network undergoes a structural reorganisation, which results in more compact structures, that are a necessary condition for the appearance of the first crystalline nuclei.
+To our knowledge, this is the first experimental demonstration of this process: it is not captured by previous computer simulations, as they generally neglect hydrodynamic interactions which play a fundamental role in this kinetic process, and no previous experiment has had the capability of tracking particles during the early stages of spinodal decomposition.
 
@@
 In this context the Fig.~1 c,d only provide hints that the cp = 0.38 mg/g state point behaves differently from other state points.
-The rest of the manuscript and the Supplementary materials is devoted to providing the conclusive evidence that that is the case.
-We have made this more clear in the new manuscript.
-
-\textcolor{red}{The Referee really likes Supplementary Fig3, so at this point if we want to answer him fully, we should probably replace Fig. 3c with Supplementary Figure 3. What do you think?}
+Following the Referee's kind suggestion, we have now moved previous Fig. S3 to a new Fig. 3b. 
+We hope that this will make the difference of the cp = 0.38 mg/mg state point from the others clearer.
 
@@
 This new kinetic analysis reveals that at low-polymer concentrations a new mechanism (stress-driven aging) is taking place, which has the effect of increasing the average coordination of the colloidal particles in the liquid phase.
-We believe this to be a novel observation of this phenomena.
+We believe this to be a novel observation of this phenomenon.
 Most past simulation works have neglected, or not fully taken into account, the role of hydrodynamic interactions (HI) in the phase separation process.
@@
 This transition is the stress-driven aging process, and increases the connectivity of the network.
-Despite being theoretically predicted, Ref.~\cite{tanaka2000,tanaka2007spontaneous}, this transition was not directly observed before.
+Despite being theoretically predicted, Ref.~\cite{tanaka2000,tanaka2007spontaneous}, this transition was not directly observed before in experiments.
+We stress that for the experimental observation we need to follow particles at a single-particle level from the beginning of phase separation.
+This was made possible by a new experimental protocol of salt injection combined with confocal microscopy observation.
 
@@
 
+We are glad to hear that the Referee thought that the previous revisions offered significant new support of the authors' claims. 
+On the comment (i), as we explained above, we have clarified the physical conditions required for crystal-gel formation to take place.
+We have also discussed how we can widen the region where this is to be observed (see page 14).
+(ii) Following the suggestion, in the new version of the manuscript we follow the Referee's recommendation and include in Fig.
+6b the fraction of surface particles, and update the description accordingly.
+
 \textcolor{red}{The Referee insists on having surface particles represented in Fig.6b.
@@
 
-In the new version of the manuscript we follow the Referee's recommendation and include in Fig. 6b the fraction of surface particles, and update the description accordingly.
-% We remind that the 54/34/11 abundances was obtained by looking at the phase history of each colloidal particle.
+\textcolor{red}{We remind that the 54/34/11 abundances was obtained by looking at the phase history of each colloidal particle.
 Fig.
@@
 4 don't contain information about the phase history of individual particles, just about the total number of particles in each phase, and the time-dependence of the transition probabilities.
-Of course these quantities are closely connected, and we hope to have provided a detailed description of the process.
+Of course these quantities are closely connected, and we hope to have provided a detailed description of the process (??).}
+
+\vskip 5mm
+
+We hope that the Referee will find that the new analysis have revealed a new mechanism behind 
+crystal-gel formation and increased the novelty of our manuscript considerably and 
+the revised manuscript is now suitable for publication in Nature Materials.
 

### 2016-12-17 nature/crystal_gel_r.tex

--- 2016-12-15 crystal-gel/nature/crystal_gel_r.tex
@@
 %But while the former process has been studied in detail, the latter has never been observed experimentally at a single-particle level and it is thus lacking a microscopic understanding. 
-With a new confocal microscopy protocol, \textcolor{blue}{we successfully follow the entire kinetics of phase separation at a single-particle level, from the homogeneous phase, percolation, stress driven ageing, and until the formation of the final gel structure.
-This allows us to unveil a novel crystallization pathway to sponge-like porous crystal structures in a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation and to reveal its physical mechanism.}
-%\textcolor{blue}{We follow the entire kinetics of phase separation, from the homogeneous phase, percolation, stress driven aging, and until the formation of porous crystals.}
-We show that crystallization leads to a network structure different from the original phase separation pattern. 
+With a new confocal microscopy protocol, \textcolor{blue}{we follow the entire kinetics of phase separation, from the homogeneous phase, percolation, stress driven ageing, and until the formation of the final structure.
+This allows us to unveil a novel crystallization pathway to sponge-like porous crystal structures in a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation and to reveal its physical mechanism.
+We clarify the physical requirements necessary for this to happen and show that crystallization leads to a network structure different from the original phase separation pattern.} 
 Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface, i.e. de-sublimation, driving evaporation of nearby liquid. 
@@
 Matching phase boundaries often leads to better determination of volume fractions than colloid diameter measurements~\cite{Poon2012}.
-We obtain the best match between theoretical phase diagram and experimental data for $\sigma=2.21\,\mu$m and $R_g=80$ nm (see Supplementary method).
+We obtain the best match between theoretical phase diagram and experimental data for $\sigma=2.21\,\mu$m and $R_g=80$ nm (see Supplementary Method).
 Therefore the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.072$ and the overlap mass fraction of polymer $2.25$~mg/g.
@@
 
+\bibliographystyle{naturemag2}
+\bibliography{biblio}
+
 \bigskip

### 2016-12-17 crystal-gel/nature/Supplementary.tex

@@
 \centerline{\bf \large Supplementary Information for}
-\centerline{\bf \large``A novel physical mechanism for the spontaneous formation }
-\centerline{\bf \large of porous crystals via viscoelastic phase separation''}
+\centerline{\bf \large``Formation of porous crystals via viscoelastic phase separation'' }
 \vspace{0.5cm}
@@
 
-\begin{figure}
-\includegraphics{local_phi12.pdf}%
-\caption{{\bf Late stage of the evolution of the local volume fraction around colloidal particles having 12 nearest neighbours.} Samples are shown with the same colour code as in Fig.1 of the main text.
-In particular red circles correspond to the crystallizing sample at $\phi=0.33$ and $cp=0.38\,$mg/g.}%
-\label{fig:local_phi12}%
-\end{figure}
+%\begin{figure}
+%\includegraphics{local_phi12.pdf}%
+%\caption{{\bf Late stage of the evolution of the local volume fraction around colloidal particles having 12 nearest neighbours.} Samples are shown with the same colour code as in Fig.1 of the main text.
+In particular red circles correspond to the crystallizing sample at $\phi=0.33$ and $c_p=0.38\,$mg/g.}%
+%\label{fig:local_phi12}%
+%\end{figure}
 
@@
 \includegraphics{capillary.pdf}
-\caption{\textbf{Capillary experiment reproducing the crystal-gel state.} The colloid volume fraction is $\approx 0.33$ and the polymer concentration is $0.40\,$mg/g.
+\caption{\textbf{Capillary experiment reproducing the crystal-gel state.} The colloid volume fraction is $\approx 0.33$ and the polymer concentration $c_p$ is $0.40\,$mg/g.
 Top row: Confocal slices showing the initial network, crystallising network and final crystal-gel network.
@@
 
+\clearpage
 \bibliographystyle{naturemag2}
@@
 \section*{Supplementary Movie 1}
-This movie shows the early stage of gelation process of $\phi\approx 0.30$ and $c_p=0.38$ mg/g, from the original confocal images at the same z-position as Fig.~3a.
+This movie shows the early stage of gelation process of $\phi\approx 0.30$ and $c_p=0.38$ mg/g, from the original confocal images at the same z-position as Fig.~4d.
 Total duration is 1160 s (530~$\tau$).

### 2016-12-17 crystal-gel/nature/response3.tex

@@
 We would like to thank you for your providing us with two Referee reports for our manuscript (NM15113574). 
-We understand your decision to reject the manuscript, but, after careful consideration of the Referee reports, we couldn't avoid thinking that they are mostly supportive of our work, and the only Referee (Referee \#1) that is somewhat against publication is basing his/her argument on a probable misunderstanding of the true novelty of the first part of our work.
+We are glad to see that both Referees 1 and 3 seem to have appreciated the improvements 
+we made after the first round of reviews: Reviewer \#1 calls the work \emph{very beautiful and interesting}, while Referee \#3 goes into the details of explaining how we actually satisfactorily clarified his/her previous doubts, with just a couple of minor points left.
+Reviewer \#2 was very positive during the first round.
 
-We decided to write to you again to clarify that the Referee might have missed some very important points. In general, both Referees seem to have appreciated the improvements 
-we made after the first round of reviews: Reviewer \#1 calls the work \emph{very beautiful and interesting}, while Referee \#3 goes into the details of explaining how we actually satisfactorily clarified his previous doubts, with just a couple of minor points left.
-Reviewer \#2 was very positive during the first round, but unfortunately didn't submit a second report.
-Overall we were able to find only two criticism in the reports.
+After receiving the comments of the Referees' reports, we have taken them seriously and made an effort to address them properly. 
+We summarize how we have addressed the critical comments of the Referees.
 \begin{itemize}
- \item Referee \#1 laments that the first part of the paper describes spinodal decomposition without showing anything new from previous studies (but he/she admits the novelty of the study of the different crystallisation channels).
-But this is not the case, as the crystallisation of the liquid network requires an additional step, called stress-driven ageing, that was never observed experimentally before.
-In our new version of the manuscript we spent considerable time developing a new analysis, that shows how spinodal decomposition proceeds differently depending on the polymer concentration, and that the ageing of the network due to mechanical stress is a necessary condition for the subsequent crystallization process.
-We think that this completely invalidates the point of the Referee, because in addition to different crystallization channels, our study unveils a new link between the ageing of the liquid network and the subsequent crystallization process.
- \item Referee \#3 suggests that we better acknowledge that the crystal-gel is found only in a quite specific region of the phase diagram, and this also became the main motivation for the rejection of our manuscript.
-We also fundamentally disagree with this idea.
-It's true that crystal-gel states don't appear everywhere in the phase diagram, but this doesn't mean that they are hard to obtain.
-Rather the opposite, there is a reproducible region around the critical point were these states can be consistently obtained, and we were able to reproduce these findings both in our new cell design, and in ordinary capillaries, showing that the region of crystal-gel formation is well within the experimental range.
-Moreover the extent of this region can be widened by changing the location of the critical point, which in our system is controlled by the length of the non-adsorbing polymer: moving the critical point to lower polymer concentrations opens the window where bonds can rearrange before the intervening glass transition.
-So we argue that the region of formation of crystal-gel porous structures is not only easily accessible, but also controllable.
+ \item Referee \#1 seemed to think that the first part of the paper describes spinodal decomposition without showing anything new from previous studies (but he/she admits the novelty of the study of the different crystallisation channels).
+We are afraid that there may be some misunderstanding here. 
+This is because the crystallisation of the liquid network requires an additional step, called stress-driven ageing, 
+that was never observed experimentally before. However, this criticism makes us realize that this point was not so clear 
+in the previous manuscript. 
+To improve this, we spent considerable time developing a new analysis, which shows how spinodal decomposition proceeds differently depending on the polymer concentration, and that the ageing of the network due to mechanical stress is a necessary condition for the subsequent crystallization process.
+We think that this has considerably increased the impact of our work, because in addition to different crystallization channels, this additional work unveils a new link between the ageing of the liquid network and the subsequent crystallization process.
+\item Referee \#3 suggests that we better acknowledge that the crystal-gel is found only in a quite specific region of the phase diagram. 
+It's certainly true that crystal-gel states don't appear everywhere in the phase diagram, but this doesn't mean that they are hard to obtain. 
+The Referee's comment led us to consider this problem in more depth. 
+First we already showed that there is a reproducible region around the critical point were these states can be consistently obtained, and we were able to reproduce these findings both in our new cell design, and in ordinary capillaries, showing that the region of crystal-gel formation is well within the experimental range.
+Moreover, thanks to the above-mentioned new analysis, we have successfully unveil the physical requirements for crystal-gel observation.
+This mechanism suggests that the extent of this region can be widened by changing the location of the critical point, which in our system is controlled by the size of the non-adsorbing polymer: moving the critical point to lower polymer concentrations opens the window where bonds can rearrange before the intervening glass transition.
+So we may say that the region of formation of crystal-gel porous structures is not only easily accessible, but also controllable.
 \end{itemize}
 
-We hope that you'll see that both criticisms are not indicative of any weaknesses in our work.
+We hope that the above explanation have addressed the comments of the Referees satisfactorily.
 We also took advantage of the new suggestions to improve the manuscript even further.
-In particular, concerning the point raised by Referee \#1 we have introduced a new type of analysis, which looks at the bond-breaking probability as a function of time, during the whole process of spinodal decomposition.
+In particular, as mentioned above, we have introduced a new type of analysis, which looks at the bond-breaking probability as a function of time, during the whole process of spinodal decomposition.
 Our new results show that, at low polymer-concentrations, the network undergoes a structural reorganisation, which results in more compact structures, that are a necessary condition for the appearance of the first crystalline nuclei.
@@
 
-We are rather positive that the new version of the manuscript would put the Referees in a position to better understand the full extent of our results, and
-so we hope that you might give it another chance. If that is is not the case, we are still grateful for the opportunity that the comments gave us for improving the manuscript in a meaningful way.
+We hope that you and your reviewers will find that the revised manuscript is now suitable for publication in Nature Materials.
 

### 2016-12-17 crystal-gel/nature/response3.tex
John's modifications.

@@
 We are glad to see that both Referees 1 and 3 seem to have appreciated the improvements 
-we made after the first round of reviews: Reviewer \#1 calls the work \emph{very beautiful and interesting}, while Referee \#3 goes into the details of explaining how we actually satisfactorily clarified his/her previous doubts, with just a couple of minor points left.
+we made after the first round of reviews: Reviewer \#1 calls the work \emph{very beautiful and interesting}, while Referee \#3 goes into the details of explaining how we actually satisfactorily clarified her/his previous doubts, with just a couple of minor points left.
 Reviewer \#2 was very positive during the first round.
@@
 \begin{itemize}
- \item Referee \#1 seemed to think that the first part of the paper describes spinodal decomposition without showing anything new from previous studies (but he/she admits the novelty of the study of the different crystallisation channels).
+ \item Referee \#1 seemed to think that the first part of the paper describes spinodal decomposition without showing anything new from previous studies (but she/he admits the novelty of the study of the different crystallzation channels).
 We are afraid that there may be some misunderstanding here. 
@@
 First we already showed that there is a reproducible region around the critical point were these states can be consistently obtained, and we were able to reproduce these findings both in our new cell design, and in ordinary capillaries, showing that the region of crystal-gel formation is well within the experimental range.
-Moreover, thanks to the above-mentioned new analysis, we have successfully unveil the physical requirements for crystal-gel observation.
+Moreover, thanks to the above-mentioned new analysis, we have successfully unveiled the physical requirements for crystal-gel observation.
 This mechanism suggests that the extent of this region can be widened by changing the location of the critical point, which in our system is controlled by the size of the non-adsorbing polymer: moving the critical point to lower polymer concentrations opens the window where bonds can rearrange before the intervening glass transition.
@@
 Most past simulation works have neglected, or not fully taken into account, the role of hydrodynamic interactions (HI) in the phase separation process.
-As was shown for example in Ref.~\cite{furukawa2010key}, the phase separation process without HI produces rather compact structures, which would then easily crystallize to form compact nuclei.
+As was shown for example in Ref.~\cite{
+furukawa2010key}, the phase separation process without HI produces rather compact structures, which would then easily crystallize to form compact nuclei.
 This is not what happens if one take HI into consideration.
@@
 This condition is necessary for the appearance of the first crystalline nuclei.
-In our new manuscript, we are able to follow the kinetics of these processes by looking at the time evolution of the bond-breaking probability.
+In our new manuscript, we are able to follow the kinetics of these processes by looking at the time evolution of the bond-breaking
+probability.
 The different network states are in fact characterised by different bond-breaking probabilities and by different connectives.
@@
 In this context the Fig.~1 c,d only provide hints that the cp = 0.38 mg/g state point behaves differently from other state points.
+Direct evidence comes from the use of bond-orientational order parameters, presented in Fig.
+5 and 6.
 Following the Referee's kind suggestion, we have now moved previous Fig. S3 to a new Fig. 3b. 
@@
 Due to the incompressibility condition, the liquid network has a lower-connectivity, and the viscoelastic phase separation results in the accumulation of mechanical stress on these liquid bonds.
-To things can happen depending on the polymer concentration.
+Two things can happen depending on the polymer concentration.
 At high-polymer concentration, the bonds are not-easily broken, and the network which is first formed cannot undergo structural reorganisation.
 This inevitably leads to gelation.
-On the other side, if the polymer concentration is low enough, the liquid network can reorganise with a rather sharp transition to a more compact network.
+On the other side, if the polymer concentration is low enough, the liquid network can reorganise with a rather sharp
+transition to a more compact network.
 This condition is necessary for the appearance of the first crystalline nuclei.

### 2016-12-17 nature/crystal_gel_r.tex
John's modifications.

@@
 %But while the former process has been studied in detail, the latter has never been observed experimentally at a single-particle level and it is thus lacking a microscopic understanding. 
-With a new confocal microscopy protocol, \textcolor{blue}{we follow the entire kinetics of phase separation, from the homogeneous phase, percolation, stress driven ageing, and until the formation of the final structure.
-This allows us to unveil a novel crystallization pathway to sponge-like porous crystal structures in a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation and to reveal its physical mechanism.
-We clarify the physical requirements necessary for this to happen and show that crystallization leads to a network structure different from the original phase separation pattern.} 
-Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface, i.e. de-sublimation, driving evaporation of nearby liquid. 
+With a new confocal microscopy protocol, \textcolor{blue}{we follow the entire kinetics of phase separation, from the homogeneous phase to the formation of different arrested states.
+This allows us to unveil a novel crystallization pathway to sponge-like porous crystal structures in a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation.
+We show that crystallization is preceeded by a structural reorganization of the liquid phase, called stress-driven aging.
+Nucleation starts inside this liquid network, but crystals grow past it by direct condensation of the gas phase on their surface, i.e.
+de-sublimation, driving evaporation of nearby liquid.
+This produces a network structure different from the original phase separation pattern.}
 %The latter processes via the gas phase has been overlooked so far.

### 2016-12-19 crystal-gel/nature/crystal_gel_r.tex
Reintroduce some changes I made before r2660 and that were delete by non updating users. In particular the change in colour scheme for crystal, liquid and gas.

@@
 %But while the former process has been studied in detail, the latter has never been observed experimentally at a single-particle level and it is thus lacking a microscopic understanding. 
-With a new confocal microscopy protocol, \textcolor{blue}{we follow the entire kinetics of phase separation, from the homogeneous phase, percolation, stress driven ageing, and until the formation of the final structure.
-This allows us to unveil a novel crystallization pathway to sponge-like porous crystal structures in a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation and to reveal its physical mechanism.
-%\textcolor{blue}{We follow the entire kinetics of phase separation, from the homogeneous phase, percolation, stress driven aging, and until the formation of porous crystals.}
-We clarify the physical requirements necessary for this to happen and show that crystallization leads to a network structure different from the original phase separation pattern.} 
-Crystals nucleate inside the liquid network, but grow past it by direct condensation of the gas phase on their surface, i.e. de-sublimation, driving evaporation of nearby liquid. 
+With a new confocal microscopy protocol, \textcolor{blue}{we follow the entire kinetics of phase separation, from the homogeneous phase to the formation of different arrested states.
+This allows us to unveil a novel crystallization pathway to sponge-like porous crystal structures in a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation.
+We show that crystallization is preceeded by a structural reorganization of the liquid phase, called stress-driven aging.
+Nucleation starts inside this liquid network, but crystals grow past it by direct condensation of the gas phase on their surface, i.e.
+de-sublimation, driving evaporation of nearby liquid.
+This produces a network structure different from the original phase separation pattern.}
 %The latter processes via the gas phase has been overlooked so far.
@@
  \caption{{\bf Stress driven aging.} Time evolution of the average coordination of colloidal particles $\bar{n}$ (black continuous line) and the bond-breaking probability $p_\text{break}$ (red dashed line) for four gel samples.
-Symbols relates each sample to its position on the phase diagram \textcolor{blue}{in} Fig.~\ref{fig:network_formation}.
-$p_\text{break}$ is measured as the probability of bonds breaking in a time interval of $\Delta t=10$ s.
+Symbols relate each sample to its position on the phase diagram \textcolor{blue}{in} Fig.~\ref{fig:network_formation}.
+$p_\text{break}$ is measured as the probability of bond breaking in a time interval of $\Delta t=10$ s.
 For high polymer concentration, panels {\bf a} and {\bf b}, $p_\text{break}$ decreases monotonically until it reaches a stationary state at long times, with $\bar{n}$ saturating at an average of less than $5$ neighbours.
@@
 {
-At intermediate times, after percolation has started, hydrodynamic interactions play a fundamental role~\cite{tanaka2000,tanaka2007spontaneous,furukawa2010key}.
+\textcolor{red}{From initial to intermediate times, before percolation is complete}, hydrodynamic interactions play a fundamental role~\cite{tanaka2000,tanaka2007spontaneous,furukawa2010key}.
 Without hydrodynamic interactions, as often assumed in simulations, particles have the tendency to aggregate in compact structures and subsequently form thick network structures.
@@
 
-\textcolor{blue}{The samples shares the same early stages of of the phase separation process, but show significantly different behaviours in the later stages.}
+\textcolor{blue}{The samples share the same early stages of the phase separation process, but show significantly different behaviours in the later stages.}
 To examine this in more detail, we plot in Fig.~\ref{fig:late_structure}a the structure factor for the final stages of the gelation process for all state points.
@@
  % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
-\caption{{\bf Percolated network structures.} {\bf a:} gel; {\bf b:} crystal-gel. The gel state is obtained from the state
-point $\phi\approx 0.33$ and $c_p=1.07$ mg/g at $t=400$ min after sample preparation, while the crystal-gel is obtained at the same $\phi$ but
-at lower polymer concentration, $c_p=0.38$ mg/g at $t=468$ min after sample preparation. On the left, the network is represented in a coarse-grained fashion (see Methods). 
-We can clearly see not only the percolated nature of both networks, but also the difference in the smoothness of the network surface between amorphous and crystal gels. 
-On the right, slabs of 4.5 $\mu$m thickness and 41 $\mu$m width are plotted, with particles coloured with the same colour scheme as in Fig.~\ref{fig:transitions}: red (crystal), green (gas particles), white (liquid particles).
-We note that very few liquid particles remain in \textbf{b} and the network is almost crystalline. 
+\caption{{\bf Percolated network structures.} 
+{\bf a, b:} the network is represented in a coarse-grained fashion (see Methods).
+{\bf c, d:} slabs of 4.5 $\mu$m thickness and 41 $\mu$m width. Particles are drawn to scale and coloured according to their phase. Purple: crystal; gray: liquid; orange: gas. 
+{\bf a, c:} a gel obtained from the state point $\phi\approx 0.33$ and $c_p=1.07$ mg/g at $t=400$ min after sample preparation. 
+{\bf b, d:} a crystal-gel at the same $\phi$ but at lower polymer concentration, $c_p=0.38$ mg/g at $t=468$ min after sample preparation. 
+On the top row we can clearly see not only the percolated nature of both networks, but also the difference in the smoothness of the network surface between amorphous and crystal gels. 
+We note that very few liquid particles remain in \textbf{d} and the network is almost crystalline. 
 } 
@@
 
-A direct analysis of the colloidal positions at late times confirms that indeed the new structure obtained at
-$\phi\approx 0.33$ and $c_p=0.38$ mg/g has different morphological properties than usual colloidal gels.
-We directly compare the two structures in Fig.~\ref{fig:network}, where Gaussian filtering is used to depict the network structure as a continuous field (see Methods).
-Panel a presents the results for high
-polymer concentration gel, $c_p=1.07$ mg/g; panel b presents the results for the low polymer concentration sample, $c_p=0.38$ mg/g.
-The comparison shows that the new arrest mechanism produces a network structure with bigger pores, smoother surfaces \textcolor{blue}{(consistently with $D=3$)}, and thicker strands (panel b) compared
-to the colloidal gel sample (panel a). The right column shows the individual particle positions for slabs with a thickness of 5 particle diameters. 
+A direct analysis of the colloidal positions at late times confirms that indeed the new structure obtained at $\phi\approx 0.33$ and $c_p=0.38$ mg/g has different morphological properties than usual colloidal gels.
+We directly compare the two structures in Fig.~\ref{fig:network}a and b, where Gaussian filtering is used to depict the network structure as a continuous field (see Methods).
+Panel~a presents the results for high polymer concentration gel, $c_p=1.07$ mg/g \textcolor{blue}{that never undergoes stress-driven ageing}; panel~b presents the results for the low polymer concentration sample, $c_p=0.38$ mg/g.
+The comparison shows that the new arrest mechanism produces a \textcolor{blue}{beaded} network structure with bigger pores, smoother surfaces \textcolor{blue}{(consistently with $D=3$)}, and thicker strands (panel b) compared
+to the colloidal gel sample (panel a). The bottom row shows the individual particle positions for slabs with a thickness of 5 particle diameters. 
 From the figure it is immediately clear that the network strands at low polymer concentration are crystalline. The new arrest mechanism thus
-involve the formation of a crystal-gel network. In order to explain the morphological differences between colloidal gels and crystal-gels,
+involve the formation of a crystal-gel network.
+In order to explain the morphological differences between colloidal gels \textcolor{blue}{(networks made of thin strands)} and crystal-gels \textcolor{blue}{(networks made of thick crystal beads)},
 we now consider the kinetics of formation in our samples.
@@
  \centering
- \includegraphics[width=12cm]{transitions}
+ \includegraphics[width=16cm]{transitions}
  % snapshots.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=0 -1 2256 735

### 2016-12-19 nature/crystal_gel_r.tex
Add surface particles: - draw them in dark purple - differentiate from liquid in evolution of fraction of particles - sketch a surface layer around the hexagon sketching the crystal

@@
  % crystal_size.eps: 0x0 pixel, 300dpi, 0.00x0.00 cm, bb=(atend)
-\caption{{\bf Percolated network structures.} {\bf a:} gel; {\bf b:} crystal-gel. The gel state is obtained from the state
-point $\phi\approx 0.33$ and $c_p=1.07$ mg/g at $t=400$ min after sample preparation, while the crystal-gel is obtained at the same $\phi$ but
-at lower polymer concentration, $c_p=0.38$ mg/g at $t=468$ min after sample preparation. On the left, the network is represented in a coarse-grained fashion (see Methods). 
-We can clearly see not only the percolated nature of both networks, but also the difference in the smoothness of the network surface between amorphous and crystal gels. 
-On the right, slabs of 4.5 $\mu$m thickness and 41 $\mu$m width are plotted, with particles coloured with the same colour scheme as in Fig.~\ref{fig:transitions}: red (crystal), green (gas particles), white (liquid particles).
-We note that very few liquid particles remain in \textbf{b} and the network is almost crystalline. 
+\caption{{\bf Percolated network structures.} 
+{\bf a, b:} the network is represented in a coarse-grained fashion (see Methods).
+{\bf c, d:} slabs of 4.5 $\mu$m thickness and 41 $\mu$m width.
+Particles are drawn to scale and coloured according to their phase.
+Purple: crystal; dark purple: crystal surface; gray: liquid; orange: gas.
+{\bf a, c:} a gel obtained from the state point $\phi\approx 0.33$ and $c_p=1.07$ mg/g at $t=400$ min after sample preparation. 
+{\bf b, d:} a crystal-gel at the same $\phi$ but at lower polymer concentration, $c_p=0.38$ mg/g at $t=468$ min after sample preparation. 
+On the top row we can clearly see not only the percolated nature of both networks, but also the difference in the smoothness of the network surface between amorphous and crystal gels. 
+We note that very few liquid particles remain in \textbf{d} and the network is almost crystalline. 
 } 
@@
 Particles are drawn to scale and coloured according to their phase.
-Red: crystal; white: liquid; green: gas.
+Purple: crystal; dark purple: crystal surface; gray: liquid; orange: gas.
 {\bf b,} Fraction of particles for each phase. Gray vertical lines indicate the times shown in panel a. 
-{\bf c,} Each particle during its trajectory changes between three states: gas, liquid and crystal. Direct crystallization (continuous arrow) accounts for 54\% of particles trajectories in which
-gas or liquid particles transition to the crystal state without liquid evaporation or crystal de-sublimation.
-The Bergeron process (dotted arrow) accounts for 35\% of particle trajectories in which liquid particles
-transition to the gas state before crystallizing. Ostwald ripening (dashed arrow) accounts for 11\% of particle trajectories in which crystal particle transition to the gas state before returning
-to the crystal state.}
+{\bf c,} Sketch of the three routes by which a particle can end up in a crystal state. 
+Probabilities are computed from the history of each single trajectory. 
+A trajectory without liquid evaporation or crystal sublimation is counted as direct crystallization (continuous arrow, 54\%). 
+If a liquid particle evaporates and then de-sublimates it counts in the Bergeron process (dotted arrow, 35\%). 
+If sublimation is followed by de-sublimation it is Ostwald ripening (dashed arrow, 11\%). 
+Most trajectories proceed via the surface state.}
 \label{fig:transitions}

### 2016-12-19 nature/response3.tex
Add surface particles: - draw them in dark purple - differentiate from liquid in evolution of fraction of particles - sketch a surface layer around the hexagon sketching the crystal

@@
 \noindent
-{\bf Dear Pep,}
-\vskip 0.3cm
+{\bf Dear Pep and/or a newly assigned Editor in Nature Materials,}
+\vskip 0.7cm
 We would like to thank you for your providing us with two Referee reports for our manuscript (NM15113574). 
-We understand your decision to reject the manuscript, but, after careful consideration of the Referee reports, we couldn't avoid thinking that they are mostly supportive of our work, and the only Referee (Referee \#1) that is somewhat against publication is basing his argument on a probable misunderstanding of the true novelty of the first part of our work.
+We are glad to see that both Referees 1 and 3 seem to have appreciated the improvements 
+we made after the first round of reviews: Reviewer \#1 calls the work \emph{very beautiful and interesting}, while Referee \#3 goes into the details of explaining how we actually satisfactorily clarified her/his previous doubts, with just a couple of minor points left.
+Reviewer \#2 was very positive during the first round.
 
-We decided to write to you again to clarify that the Referee might have missed some very important points. In general, both Referees seem to have appreciated the improvements
-we made after the first round of reviews: Reviewer \#1 calls the work \emph{very beautiful and interesting}, while Referee \#3 goes into the details of explaining how we actually satisfactorily clarified his previous doubts, with just a couple of minor points left.
-Reviewer \#2 was very positive during the first round, but unfortunately didn't submit a second report.
-Overall we were able to find only two criticism in the reports.
+After receiving the comments of the Referees' reports, we have taken them seriously and made an effort to address them properly. 
+We summarize how we have addressed the critical comments of the Referees.
 \begin{itemize}
- \item Referee \#1 laments that the first part of the paper describes spinodal decomposition without showing anything new from previous studies (but he admits the novelty of the study of the different crystallisation channels).
-But this is not the case, as the crystallisation of the liquid network requires an additional step, called stress-driven aging, that was never observed experimentally before.
-In our new version of the manuscript we spent considerable time developing a new analysis, that shows how spinodal decomposition proceeds differently depending on the polymer concentration, and that the aging of the network due to mechanical stress is a necessary condition for the subsequent crystallization process.
-We think that this completely invalidates the point of the Referee, because in addition to different crystallization channels, our study unveils a new link between the aging of the liquid network and the subsequent crystallization process.
- \item Referee \#3 suggests that we better acknowledge that the crystal-gel is found only in a quite specific region of the phase diagram, and this also became the main motivation for the rejection of our manuscript.
-We also fundamentally disagree with this idea.
-It's true that crystal-gel states don't appear everywhere in the phase diagram, but this doesn't mean that they are hard to obtain.
-Rather the opposite, there is a reproducible region around the critical point were these states can be consistently obtained, and we were able to reproduce these findings both in our new cell design, and in ordinary capillaries, showing that the region of crystal-gel formation is well within the experimental range.
-Moreover the extent of this region can be widened by changing the location of the critical point, which in our system is controlled by the length of the non-adsorbing polymer: moving the critical point to lower polymer concentrations opens the window where bonds can rearrange before the intervening glass transition.
-So we argue that the region of formation of crystal-gel porous structures is not only easily accessible, but also controllable.
+ \item Referee \#1 seemed to think that the first part of the paper describes spinodal decomposition without showing anything new from previous studies (but she/he admits the novelty of the study of the different crystallzation channels).
+We are afraid that there may be some misunderstanding here. 
+This is because the crystallisation of the liquid network requires an additional step, called stress-driven ageing, 
+that was never observed experimentally before. However, this criticism makes us realize that this point was not so clear 
+in the previous manuscript. 
+To improve this, we spent considerable time developing a new analysis, which shows how spinodal decomposition proceeds differently depending on the polymer concentration, and that the ageing of the network due to mechanical stress is a necessary condition for the subsequent crystallization process.
+We think that this has considerably increased the impact of our work, because in addition to different crystallization channels, this additional work unveils a new link between the ageing of the liquid network and the subsequent crystallization process.
+\item Referee \#3 suggests that we better acknowledge that the crystal-gel is found only in a quite specific region of the phase diagram. 
+It's certainly true that crystal-gel states don't appear everywhere in the phase diagram, but this doesn't mean that they are hard to obtain. 
+The Referee's comment led us to consider this problem in more depth. 
+First we already showed that there is a reproducible region around the critical point were these states can be consistently obtained, and we were able to reproduce these findings both in our new cell design, and in ordinary capillaries, showing that the region of crystal-gel formation is well within the experimental range.
+Moreover, thanks to the above-mentioned new analysis, we have successfully unveiled the physical requirements for crystal-gel observation.
+This mechanism suggests that the extent of this region can be widened by changing the location of the critical point, which in our system is controlled by the size of the non-adsorbing polymer: moving the critical point to lower polymer concentrations opens the window where bonds can rearrange before the intervening glass transition.
+So we may say that the region of formation of crystal-gel porous structures is not only easily accessible, but also controllable.
 \end{itemize}
 
-We hope that you'll see that both criticisms are not indicative of any weaknesses in our work.
+We hope that the above explanation have addressed the comments of the Referees satisfactorily.
 We also took advantage of the new suggestions to improve the manuscript even further.
-In particular, concerning the point raised by Referee \#1 we have introduced a new type of analysis, which looks at the bond-breaking probability as a function of time, during the whole process of spinodal decomposition.
+In particular, as mentioned above, we have introduced a new type of analysis, which looks at the bond-breaking probability as a function of time, during the whole process of spinodal decomposition.
 Our new results show that, at low polymer-concentrations, the network undergoes a structural reorganisation, which results in more compact structures, that are a necessary condition for the appearance of the first crystalline nuclei.
@@
 
-We are rather positive that the new version of the manuscript would put the Referees in a position to better understand the full extent of our results, and
-so we hope that you might give it another chance. If that is is not the case, we are still grateful for the opportunity that the comments gave us for improving the manuscript in a meaningful way.
+\vskip 0.3cm
+
+Since the manuscript now includes the new analysis mentioned above and is expanded with 6 figures, we would like to resubmit it as an Article rather than a Letter. 
+\textit{
+\vskip 0.3cm}
+
+We hope that you and your reviewers will find that the revised manuscript is now suitable for publication in Nature Materials.
+
+\vskip 0.5cm
+
+Warm thanks to you for your kind consideration and cooperation.
 
@@
 Most past simulation works have neglected, or not fully taken into account, the role of hydrodynamic interactions (HI) in the phase separation process.
-As was shows for example in Ref.~\cite{furukawa2010key}, the phase separation process without HI produces rather compact structures, which would then easily crystallize to form compact nuclei.
-This is not what happens if one take HI into consideration.
+As was shown for example in Ref.~\cite{
+furukawa2010key}, the phase separation process without HI produces rather compact structures, which would then easily crystallize to form compact nuclei.
+This is not what happens if one takes HI into consideration.
 Due to the incompressibility condition, the liquid network has a lower-connectivity, and the viscoelastic phase separation results in the accumulation of mechanical stress on these liquid bonds.
-To things can happen depending on the polymer concentration.
+Two things can happen depending on the polymer concentration.
 At high-polymer concentration, the bonds are not-easily broken, and the network which is first formed cannot undergo structural reorganisation.
@@
 This condition is necessary for the appearance of the first crystalline nuclei.
-In our new manuscript, we are able to follow the kinetics of these processes by looking at the time evolution of the bond-breaking probability.
-The different network states are in fact characterised by different bond-breaking probabilities and by different connectives.
+In our new manuscript, we are able to follow the kinetics of these processes by looking at the time evolution of the bond-breaking
+probability.
+The different network states are in fact characterised by different bond-breaking probabilities and by different coordination numbers.
 In Fig.~2a,b we show these quantities for high-polymer concentration samples, where spinodal decomposition is the only process that occurs.
@@
 In this context the Fig.~1 c,d only provide hints that the cp = 0.38 mg/g state point behaves differently from other state points.
-The rest of the manuscript and the Supplementary materials is devoted to providing the conclusive evidence that that is the case.
-We have made this more clear in the new manuscript.
-
-\textcolor{red}{The Referee really likes Supplementary Fig3, so at this point if we want to answer him fully, we should probably replace Fig. 3c with Supplementary Figure 3. What do you think?}
+Direct evidence comes from the use of bond-orientational order parameters, presented in Fig.
+5 and 6.
+Following the Referee's kind suggestion, we have now moved previous Fig. S3 to a new Fig. 3b. 
+We hope that this will make the difference of the cp = 0.38 mg/mg state point from the others clearer.
 
@@
 Due to the incompressibility condition, the liquid network has a lower-connectivity, and the viscoelastic phase separation results in the accumulation of mechanical stress on these liquid bonds.
-To things can happen depending on the polymer concentration.
+Two things can happen depending on the polymer concentration.
 At high-polymer concentration, the bonds are not-easily broken, and the network which is first formed cannot undergo structural reorganisation.
 This inevitably leads to gelation.
-On the other side, if the polymer concentration is low enough, the liquid network can reorganise with a rather sharp transition to a more compact network.
+On the other side, if the polymer concentration is low enough, the liquid network can reorganise with a rather sharp
+transition to a more compact network.
 This condition is necessary for the appearance of the first crystalline nuclei.
 In our new manuscript, we are able to follow the kinetics of these processes by looking at the time evolution of the bond-breaking probability.
-The different network states are in fact characterised by different bond-breaking probabilities and by different connectives.
+The different network states are in fact characterised by different bond-breaking probabilities and by different coordination numbers.
 In Fig.~2a,b we show these quantities for high-polymer concentration samples, where spinodal decomposition is the only process that occurs.

### 2016-12-19 crystal-gel/data/slices_article.py
Add surface particles: - draw them in dark purple - differentiate from liquid in evolution of fraction of particles - sketch a surface layer around the hexagon sketching the crystal

@@ -23,7 +23,7 @@ for sample, times in zip(['', '_highcp'], [[120, 390, 890], [800]]):
             for x,y,z in pos[phase<3]
             ]+[Texture(Pigment(color=(0.694,  0.145,  0.623)))])
         f.write(crystals)
-        for i, c in zip([3,4,5], [(0.5,0.5,0.5), (0.5,0.5,0.5), (1.,  0.455,  0.156)]):
+        for i, c in zip([3,4,5], [(0.347,0.07,0.311), (0.5,0.5,0.5), (1.,  0.455,  0.156)]):
             f.write(Union(*[
                 Sphere( (x,y,z), 5)
                 for x,y,z in pos[phase==i]

### 2016-12-19 crystal-gel/nature/crystal_gel_r.tex
John: added small description of new figure

@@
 {\bf a, b:} the network is represented in a coarse-grained fashion (see Methods).
-{\bf c, d:} slabs of 4.5 $\mu$m thickness and 41 $\mu$m width. Particles are drawn to scale and coloured according to their phase. Purple: crystal; gray: liquid; orange: gas. 
+{\bf c, d:} slabs of 4.5 $\mu$m thickness and 41 $\mu$m width.
+Particles are drawn to scale and coloured according to their phase.
+Purple: crystal; dark purple: crystal surface; gray: liquid; orange: gas.
 {\bf a, c:} a gel obtained from the state point $\phi\approx 0.33$ and $c_p=1.07$ mg/g at $t=400$ min after sample preparation. 
@@
 Particles are drawn to scale and coloured according to their phase.
-Purple: crystal; gray: liquid; orange: gas.
+Purple: crystal; dark purple: crystal surface; gray: liquid; orange: gas.
 {\bf b,} Fraction of particles for each phase. Gray vertical lines indicate the times shown in panel a. 
-{\bf c,} Each particle during its trajectory changes between three states: gas, liquid and crystal. Direct crystallization (continuous arrow) accounts for 54\% of particles trajectories in which
-gas or liquid particles transition to the crystal state without liquid evaporation or crystal de-sublimation.
-The Bergeron process (dotted arrow) accounts for 35\% of particle trajectories in which liquid particles
-transition to the gas state before crystallizing. Ostwald ripening (dashed arrow) accounts for 11\% of particle trajectories in which crystal particle transition to the gas state before returning
-to the crystal state.}
+{\bf c,} Sketch of the three routes by which a particle can end up in a crystal state. 
+Probabilities are computed from the history of each single trajectory. 
+A trajectory without liquid evaporation or crystal sublimation is counted as direct crystallization (continuous arrow, 54\%). 
+If a liquid particle evaporates and then de-sublimates it counts in the Bergeron process (dotted arrow, 35\%). 
+If sublimation is followed by de-sublimation it is Ostwald ripening (dashed arrow, 11\%). 
+Most trajectories proceed via the surface state.}
 \label{fig:transitions}
@@
 to the gas phase due to the higher vapour pressure of the liquid phase compared to the crystalline phase.
+\textcolor{blue}{The crystalline nuclei are initially composed mostly of surface particles, but as the growth of nuclei becomes more compact the number of surface particles slowly decreases, while bulk crystals keep increasing}.
 After the onset of a steady-state gas population, there are three growth routes for the crystal. 

### 2016-12-19 crystal-gel/nature/crystal_gel_r.tex
John

@@
 $\phi\approx 0.33$ and $c_p=0.38$ mg/g, after the liquid-gas phase separation has occurred.
+\textcolor{blue}{Surface, gas and liquid particles are identified from the subset of non-crystalline particles depending
+on their neighbouring particles. A particle with at least two neighbouring crystal particles is classified as
+\emph{surface}. A particle that is neither solid nor surface, and has at least four neighbouring particles
+is classified as \emph{liquid}. A particle which is neither solid, surface nor liquid is classified as \emph{gas}.}
 The process of crystallization is characterized 

### 2016-12-20 crystal-gel/nature/Supplementary.tex

@@
 
-\begin{figure}[!b]
- \centering
- \includegraphics[width=10cm]{./fractions.pdf}
- % fractions.pdf: 681x503 pixel, 72dpi, 24.02x17.74 cm, bb=0 0 681 503
- \caption{{\bf Fraction of particles for each phase at $\phi\approx 0.30$ and $c_p=0.38$ mg/g}. The liquid
- fraction is depicted as a dashed black line, the gas fraction as a dotted-dashed green line, the surface
- fraction as a double-dotted-dashed blue line, while the crystal with a continuous red line.}
- \label{fig:fractions}
-\end{figure}
+%\begin{figure}[!b]
+% \centering
+% \includegraphics[width=10cm]{./fractions.pdf}
+% \caption{{\bf Fraction of particles for each phase at $\phi\approx 0.30$ and $c_p=0.38$ mg/g}.
+The liquid  fraction is depicted as a dashed black line, the gas fraction as a dotted-dashed green line, the surface  fraction as a double-dotted-dashed blue line, while the crystal with a continuous red line.}
+% \label{fig:fractions}
+%\end{figure}
 
@@
 is classified as \emph{liquid}. A particle which is neither solid, surface nor liquid is classified as \emph{gas}. 
-Figure~\ref{fig:fractions} plots the fraction of each state for the state point at $\phi\approx 0.30$ and $c_p=0.38$ mg/g
+In Fig. 6b, we plot the fraction of each state for the state point at $\phi\approx 0.30$ and $c_p=0.38$ mg/g.
 
@@
 
-The phase diagram is obtained in the framework of the generalized free volume theory.
+The phase diagram shown in Fig.
+1 is obtained in the framework of the generalized free volume theory.
 The colloids are considered as hard spheres and the polymers as an ideal gas.
@@
 
-\clearpage
+%\clearpage
 \bibliographystyle{naturemag2}

### 2016-12-20 crystal-gel/nature/cover_letter_NM.tex

--- 2016-12-17 crystal-gel/nature/response3.tex
@@
 
-We decided to write to you again to clarify that the Referee might have missed some very important points. In general, both Referees seem to have appreciated the improvements 
-we made after the first round of reviews: Reviewer \#1 calls the work \emph{very beautiful and interesting}, while Referee \#3 goes into the details of explaining how we actually satisfactorily clarified his previous doubts, with just a couple of minor points left.
+We decided to write to you again to clarify that the Referees might have missed some very important points. In general, both Referees seem to have appreciated the improvements 
+we made after the first round of reviews: Reviewer \#1 calls the work \emph{very beautiful and interesting}, while Referee \#3 goes into the details of explaining how we actually satisfactorily clarified his/her previous doubts, with just a couple of minor points left.
 Reviewer \#2 was very positive during the first round, but unfortunately didn't submit a second report.
-Overall we were able to find only two criticism in the reports.
+Overall we were able to find only two criticisms in the reports.
 \begin{itemize}
@@
 But this is not the case, as the crystallisation of the liquid network requires an additional step, called stress-driven ageing, that was never observed experimentally before.
-In our new version of the manuscript we spent considerable time developing a new analysis, that shows how spinodal decomposition proceeds differently depending on the polymer concentration, and that the ageing of the network due to mechanical stress is a necessary condition for the subsequent crystallization process.
+To show this clearly, we spent considerable time developing a new analysis, which demonstrates how spinodal decomposition proceeds differently depending on the polymer concentration, and that the ageing of the network due to mechanical stress is a necessary condition for the subsequent crystallization process.
 We think that this completely invalidates the point of the Referee, because in addition to different crystallization channels, our study unveils a new link between the ageing of the liquid network and the subsequent crystallization process.
@@
 Rather the opposite, there is a reproducible region around the critical point were these states can be consistently obtained, and we were able to reproduce these findings both in our new cell design, and in ordinary capillaries, showing that the region of crystal-gel formation is well within the experimental range.
-Moreover the extent of this region can be widened by changing the location of the critical point, which in our system is controlled by the length of the non-adsorbing polymer: moving the critical point to lower polymer concentrations opens the window where bonds can rearrange before the intervening glass transition.
+Moreover the extent of this region can be widened by changing the location of the critical point, which in our system is controlled by the size of the non-adsorbing polymer: moving the critical point to lower polymer concentrations opens the window where bonds can rearrange before the intervening glass transition.
 So we argue that the region of formation of crystal-gel porous structures is not only easily accessible, but also controllable.
@@
 To our knowledge this is the first experimental demonstration of this process: it is not captured by computer simulations, as they generally neglect hydrodynamic interactions which play a fundamental role in this kinetic process, and no previous experiment has had the capability of tracking particles during the early stages of spinodal decomposition.
+
+\vskip 0.3cm
+
+Could you see our Rebuttal Letter for the detailed responses to the Referees' comments and criticisms?
 
@@
 \indent Hajime Tanaka}
+
+\end{document}
 

### 2016-12-20 crystal-gel/nature/crystal_gel_r.tex

@@
 
-Crystallization plays a fundamental role in many processes occurring in nature, such as \textcolor{blue}{ice formation in atmospheric clouds \cite{glickman2000glossary,morrison2012resilience}}, and in technological applications
+Crystallization plays a fundamental role in many processes occurring in nature, such as ice formation in atmospheric clouds \cite{glickman2000glossary,morrison2012resilience}, and in technological applications
 that are at the core of the chemical, pharmaceutical, and food industries.
@@
 Under usual conditions, the nucleation stage involves a very small number of molecules, of the order of 10-1000 molecules.
-This has severely limited the possibility to observe the  \textcolor{blue}{crystallization} process directly, and has also posed limits on models based on macroscopic thermodynamic properties. 
-The inability to directly observe the  \textcolor{blue}{crystallization} process is also a limit to our understanding of its kinetic pathway. 
+This has severely limited the possibility to observe the crystallization process directly, and has also posed limits on models based on macroscopic thermodynamic properties. 
+The inability to directly observe the crystallization process is also a limit to our understanding of its kinetic pathway. 
 Classical Nucleation Theory assumes the crystallization processes to occur in one step,
@@
 ~\cite{poon2002,zaccarelli2007,piazza1994phase,verhaegh1997transient,lu2008gelation}. 
-\textcolor{blue}{This gel formation process can be regarded as viscoelastic phase separation  \cite{tanaka1999colloid,tanaka2000viscoelastic} into a dense liquid phase with slow dynamics and a dilute gas phase with fast dynamics.}
+This gel formation process can be regarded as viscoelastic phase separation  \cite{tanaka1999colloid,tanaka2000viscoelastic} into a dense liquid phase with slow dynamics and a dilute gas phase with fast dynamics.
 This difference in the viscoelastic properties between the two phases allows the system to form a space-spanning 
@@
 This introduces turbulent flows at the beginning of the process, and also does not allow the
-observation of the initial stages of  \textcolor{blue}{gelation}. This initial perturbation may even alter the selection of the final non-equilibrium arrested state.  
+observation of the initial stages of gelation. This initial perturbation may even alter the selection of the final non-equilibrium arrested state.  
 In our experiments we instead succeed in building a novel protocol that allows the time evolution of the system to be observed directly
@@
  \includegraphics[width=8cm]{network_formation.pdf}
- \caption{\textcolor{blue}{{\bf Phase behaviour.}} 
+ \caption{{\bf Phase behaviour.} 
 Empty black triangles are experimental points showing fluid behaviour.
@@
 the random gelation universality class exponent ($D=2.53$) as shown by the gray line in the inset of Fig.~\ref{fig:network_formation}.
-Here we do not imply that the formation of the network follows the random gelation universality class, \textcolor{blue}{but} just that our
+Here we do not imply that the formation of the network follows the random gelation universality class, but just that our
 fractal (or effective) dimension is compatible with it.
@@
  \caption{{\bf Stress driven aging.} Time evolution of the average coordination of colloidal particles $\bar{n}$ (black continuous line) and the bond-breaking probability $p_\text{break}$ (red dashed line) for four gel samples.
-Symbols relate each sample to its position on the phase diagram \textcolor{blue}{in} Fig.~\ref{fig:network_formation}.
+Symbols relate each sample to its position on the phase diagram in Fig.~\ref{fig:network_formation}.
 $p_\text{break}$ is measured as the probability of bond breaking in a time interval of $\Delta t=10$ s.
@@
  \includegraphics[width=16cm]{late_structure.pdf}
- \caption{\textcolor{blue}{{\bf Structural evolution at late times.}} 
+ \caption{{\bf Structural evolution at late times.} 
 {\bf a,} Structure factors for the late stages of the gelation process. The gray and black lines represent respectively random gelation and compact fractal dimensions.
-{\bf b,} Late stage of the evolution of the local volume fraction \textcolor{blue}{$\phi_{\rm loc}$(12)} around colloidal particles having 12 nearest neighbours.
-{\bf c,} Local volume fraction \textcolor{blue}{$\phi_{\rm loc}$} around a colloidal particle, as a function of the number of nearest neighbours of the particle for the state points in the very late stage.
+{\bf b,} Late stage of the evolution of the local volume fraction $\phi_{\rm loc}$(12) around colloidal particles having 12 nearest neighbours.
+{\bf c,} Local volume fraction $\phi_{\rm loc}$ around a colloidal particle, as a function of the number of nearest neighbours of the particle for the state points in the very late stage. 
 The two horizontal straight lines indicate the characteristic volume fractions in attractive systems, crystal ($\sim 0.72$) and glass ($\sim 0.64$).
@@
 The gray and black straight lines have a slope of $1/2.53$ and $1/3$ respectively. 
-\textbf{(Inset)} Time evolution of the average size of the crystals \textcolor{blue}{(red dashed line)} and the number of crystals 
-\textcolor{blue}{(black line)} for the same state point.
+\textbf{(Inset)} Time evolution of the average size of the crystals (red dashed line) and the number of crystals 
+(black line) for the same state point.
 } 
@@
 With this analysis we reveal that crystallization events indeed start occurring for the state point $\phi\approx 0.33$ and $c_p=0.38$ mg/g, which has the lowest
-polymer concentration of all state points. This is shown in the inset of Fig.~\ref{fig:crystals}, where the \textcolor{blue}{red dashed} line indicates the average size of the crystallites, 
-while the \textcolor{blue}{black} line indicates the average number of crystallites as a function of time. The number of crystallites first rapidly increases as nucleation events
+polymer concentration of all state points. This is shown in the inset of Fig.~\ref{fig:crystals}, where the red dashed line indicates the average size of the crystallites, 
+while the black line indicates the average number of crystallites as a function of time. The number of crystallites first rapidly increases as nucleation events
 start occurring inside the liquid network, but eventually starts decreasing as the different crystallites grow and merge with each other.
@@
 
-While nucleation always occurs inside the liquid branches of the phase-separating network, crystal growth can proceed through different growth \textcolor{blue}{routes}.
+While nucleation always occurs inside the liquid branches of the phase-separating network, crystal growth can proceed through different growth routes.
 In the following we will show that the growth mechanism of the crystal is not only due to filling of the spinodal liquid network, as was observed in Refs.~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways} for Brownian Dynamics simulations, but also involves the Bergeron process and Ostwald ripening.
@@
 
+%\vspace{5mm}
+%\noindent
+%{\bf Data Availability.}
+
+%The data that support the plots within this paper and other findings of this study are available upon reasonable request.
+
 \section*{METHODS}

### 2016-12-20 crystal-gel/nature/RebuttalLetter.tex

\documentclass[11pt,a4paper]{article}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage{xcolor}
\usepackage{setspace}

\usepackage[top=35mm,bottom=35mm,left=30mm,right=30mm]{geometry}

\newenvironment{referee}%
{\bigskip\singlespacing\bf}%
{\par\bigskip}

\begin{document}

%\doublespacing

%\doublespacing

\noindent
{\bf Dear Editor,}
\vskip 0.7cm
We would like to thank you for your providing us with two Referee reports for our manuscript (NM15113574). 
We are glad to see that both Referees 1 and 3 seem to have appreciated the improvements 
we made after the first round of reviews: Reviewer \#1 calls the work \emph{very beautiful and interesting}, while Referee \#3 goes into the details of explaining how we actually satisfactorily clarified her/his previous doubts, with just a couple of minor points left. Reviewer \#2 was very positive during the first round.

After receiving the comments of the Referees' reports, we have taken them seriously and made an effort to address them properly. 
We summarize how we have addressed the critical comments of the Referees.
\begin{itemize}
 \item Referee \#1 seemed to think that the first part of the paper describes spinodal decomposition without showing anything new from previous studies (but she/he admits the novelty of the study of the different crystallzation channels). 
We are afraid that there may be some misunderstanding here. 
This is because the crystallisation of the liquid network requires an additional step, called stress-driven ageing, 
that was never observed experimentally before. However, this criticism makes us realize that this point was not so clear 
in the previous manuscript. 
To improve this, we spent considerable time developing a new analysis, which shows how spinodal decomposition proceeds differently depending on the polymer concentration, and that the ageing of the network due to mechanical stress is a necessary condition for the subsequent crystallization process. We think that this has considerably increased the impact of our work, because in addition to different crystallization channels, this additional work unveils a new link between the ageing of the liquid network and the subsequent crystallization process.
\item Referee \#3 suggests that we better acknowledge that the crystal-gel is found only in a quite specific region of the phase diagram. 
It's certainly true that crystal-gel states don't appear everywhere in the phase diagram, but this doesn't mean that they are hard to obtain. 
The Referee's comment led us to consider this problem in more depth. 
First we already showed that there is a reproducible region around the critical point were these states can be consistently obtained, and we were able to reproduce these findings both in our new cell design, and in ordinary capillaries, showing that the region of crystal-gel formation is well within the experimental range. Moreover, thanks to the above-mentioned new analysis, we have successfully unveiled the physical requirements for crystal-gel observation. This mechanism suggests that the extent of this region can be widened by changing the location of the critical point, which in our system is controlled by the size of the non-adsorbing polymer: moving the critical point to lower polymer concentrations opens the window where bonds can rearrange before the intervening glass transition. So we may say that the region of formation of crystal-gel porous structures is not only easily accessible, but also controllable.
\end{itemize}

We hope that the above explanation have addressed the comments of the Referees satisfactorily. We also took advantage of the new suggestions to improve the manuscript even further. In particular, as mentioned above, we have introduced a new type of analysis, which looks at the bond-breaking probability as a function of time, during the whole process of spinodal decomposition. Our new results show that, at low polymer-concentrations, the network undergoes a structural reorganisation, which results in more compact structures, that are a necessary condition for the appearance of the first crystalline nuclei. To our knowledge this is the first experimental demonstration of this process: it is not captured by computer simulations, as they generally neglect hydrodynamic interactions which play a fundamental role in this kinetic process, and no previous experiment has had the capability of tracking particles during the early stages of spinodal decomposition.

\vskip 0.3cm

Since the manuscript now includes the new analysis mentioned above and is expanded with 6 figures, we would like to resubmit it as an Article rather than a Letter. 
\textit{
\vskip 0.3cm}

We hope that you and your reviewers will find that the revised manuscript is now suitable for publication in Nature Materials.

\vskip 0.5cm

Warm thanks to you for your kind consideration and cooperation.

\vskip 0.8cm

\noindent
{\bf Yours Sincerely,
\vskip 0.3cm
\indent Hideyo Tsurusawa\\
\indent John Russo\\
\indent Mathieu Leocmach\\
\indent Hajime Tanaka}

\clearpage

\noindent
\begin{Large}
{\bf Response to Referee \#1}
\end{Large}

\vskip 0.5cm

We would like to thank the Referee for her/his comments. 
We are glad to see that he/she called the work \emph{very beautiful and interesting}. 
We address the concerns of the Referee in the following response.

\begin{referee}
The authors have improved their manuscript according to the comments of all referees, which has increased the readability of the paper. The novelty of the paper is also much better explained, which is in fact described not earlier than on page 7 of the manuscript.
The first 7 pages of the paper describes one of the arrest mechanisms of spinodal decomposition due to crystallization, which was already observed in experiments and simulations.
\end{referee}

We thank the Referee for her/his appreciation of the improvements made during our last revision of the manuscript. The Referee comments on the first 7 pages being not original, and simply describing the process of spinodal decomposition as previously observed in experiments and simulations. But we believe that our study offers novel insights also in the description of the early stages of the phase separation process. This is a very important point, as there is a clear link between the early-stage kinetics of the liquid network, and the late-stage crystallization process. In order to highlight this connection, in this new version of the manuscript we introduce a new type of analysis. In summary, we show that our work represents the first experimental demonstration of a process called \emph{stress-driven aging}. Most past simulation works have neglected, or not fully taken into account, the role of hydrodynamic interactions (HI) in the phase separation process. As was shown for example in Ref.~\cite{
furukawa2010key}, the phase separation process without HI produces rather compact structures, which would then easily crystallize to form compact nuclei. This is not what happens if one takes HI into consideration. Due to the incompressibility condition, the liquid network has a lower-connectivity, and the viscoelastic phase separation results in the accumulation of mechanical stress on these liquid bonds. Two things can happen depending on the polymer concentration. At high-polymer concentration, the bonds are not-easily broken, and the network which is first formed cannot undergo structural reorganisation. This inevitably leads to gelation. On the other side, if the polymer concentration is low enough, the liquid network can reorganise with a rather sharp transition to a more compact network. This condition is necessary for the appearance of the first crystalline nuclei. In our new manuscript, we are able to follow the kinetics of these processes by looking at the time evolution of the bond-breaking 
probability. The different network states are in fact characterised by different bond-breaking probabilities and by different coordination numbers. In Fig.~2a,b we show these quantities for high-polymer concentration samples, where spinodal decomposition is the only process that occurs. At low-polymer concentrations (Fig.~2c,d) instead, the network undergoes a sharp transition to a state with lower bond-breaking probability and higher connectivity. This transition is the stress-driven aging process, and increases the connectivity of the network. Despite being theoretically predicted, Ref.~\cite{tanaka2000,tanaka2007spontaneous}, this transition was not directly observed before.
We agree that the previous version of the manuscript was lacking on this part, so the Referee comments inspired us to develop the new analysis presented in new Fig. 2.

\begin{referee}
What is new is that crystallization not only takes place in the spinodal liquid branches, but can also occur via the Bergeron process in which the liquid evaporates, and the gas phase crystallizes again or via Ostwald ripening, i.e., melting of crystallites into the gas phase and recrystallization. Both processes make sense and can be explained that the crystallites should be surrounded by the gas phase due to the non-existence of a liquid-crystal coexistence as already speculated by Renth et al. It would be worthwhile to mention this more explicitly in the paper. In addition, the authors mention on page 3, that for confocal microscopy studies on colloidal gelation, the colloid-polymer mixture have so far been used. There are also confocal microscopy experiments performed on oppositely charged colloids,
which is worthwhile mentioning as well, Sanz et al., J. Phys. condens. Matt. 20, 494247 (2008), Sanz et al., J. Phys. Chem. B 112, 10861 (2008), where the effect of range and attraction strength have been studied on the different arrest mechanisms of the gels, including the crystallization of spinodal structures. Although the work is very beautiful and interesting, the referee has still strong doubts if the work provides sufficient novel insights to warrant publication in Nature Materials. First of all, the title suggests that a completely new route is found to form porous crystals. Crystallisation of spinodal networks has already been found in previous works and so the present work doesn't provide a new route, only another crystallization mechanism when the crystals get larger than the thickness of the branches. Secondly, the abstract suggests a new confocal microscopy protocol to study phase separation, which is in fact based on a salt injection that changes the interactions in
situ. Finally, the last sentence of the abstract is also misleading as this new route is not a single-step pathway to form nanoporous crystals, it involves spinodal decomposition, crystallization of the branches, and subsequently evaporation of the liquid and crystallization or melting of crystalline parts and recrystallization.
\end{referee}

In the new version of the manuscript we have tried to incorporate all the Referee's previous suggestions. 
Following the suggestion, we have explicitly mentioned the work of Poon and coworkers in connection to the Bergeron and Ostwald ripening processes.  
We have also added the two recommended references by Sanz et al. to the new version of the manuscript. 
On the doubt of the referee concerning the novelty of our work, we hope that our new analysis explained above provides a 
new physical insight on the physical condition required for the formation of porous crystal gels besides its new kinetic route. 
We also simplified the title to avoid misunderstanding of our message. Regarding the use of a novel salt--injection we believe this to be the first time this cell design has been applied for confocal microscopy observation at a single-particle level and 
used to study the formation of porous crystals in gels, 
which is the key to access the early stage of viscoelastic phase separation at a single particle resolution with full particle tracking. 
For example, the new analysis shown in new Fig. 2 is only possible with this new experimental protocol and full particle tracking.  
Finally, the last sentence has been modified, following the kind suggestion of the Referee.

\vskip 0.5cm

We hope that the Referee will find that the new analysis have revealed a new mechanism behind 
crystal-gel formation and increased the novelty of our manuscript considerably and 
the revised manuscript is now suitable for publication in Nature Materials.

\clearpage

\noindent
\begin{Large}
{\bf Response to Referee \#3}
\end{Large}

\vskip 0.5cm

We would like to thank the Referee for his/her comments. In the following we respond to the new points raised.

\begin{referee}
I was aware of the different frame rates on the two movies but I agree that the time stamps are helpful. Figure S5 addresses my concern about reproducibility at the single state point, and the different geometry itself is helpful too. I think the authors should be up front in the manuscript that the mechanism discussed is found in a very specific window (from the figures and SI, I think it is $\phi\sim 0.33$ and $c_p~0.39$ (plus minus) 0.01 mg/g) by moving this point into the abstract.
\end{referee}

Following the Referee's recommendation, we have put the following sentence in the abstract: We clarify the physical requirements necessary for this to happen. Although this does not explicitly specify the points we observed crystal gels, our new analysis (see below) 
has revealed the physical conditions required for observing crystal-gel formation. We have also discussed this issue in our revised manuscript with the new analysis shown in a new Fig. 2. We stress that the conditions for crystal-gel formation are not stringent and well within the experimental accessible states, as we previously demonstrated in our reproducibility tests inside a capillary. Our new results show that, at low polymer-concentrations, the network undergoes a structural reorganisation, which results in more compact structures, that are a necessary condition for the appearance of the first crystalline nuclei. To our knowledge, this is the first experimental demonstration of this process: it is not captured by previous computer simulations, as they generally neglect hydrodynamic interactions which play a fundamental role in this kinetic process, and no previous experiment has had the capability of tracking particles during the early stages of spinodal decomposition.

\begin{referee}
The addition of the SI movie 3 and the discussion of Figure 2b addresses this concern for me. It's the percolation that is critical to the authors claim of a crystal gel - the percolated structure needs to persist over long times, otherwise, for example, the gel wouldn't have a finite low frequency elastic modulus. I think SI move 3 does show this.

3) The data quality of Figure 1c, Figure 1d appear insufficient to classify the cp = 0.38 mg/mg state point as different from the others. The authors should add replications, errors, and significance testing if they wish to distinguish the dimensionality (Fig 1c) and maximum volume fraction (Fig 1d) of the 0.38 mg/ml state point from the others.

I continue to think that the evidence for different behavior in the cp = 0.38 mg/g state point from Figure 1 c,d is weak. Supplemental Figure 3 is more convincing that the final state points are different. 
\end{referee}

We fully agree that the evidence from Fig.~1 c,d was not conclusive, but the goal of those figures was not to provide conclusive evidence, but to provide a comprehensive study of the early-stages of spinodal decomposition. In this context the Fig.~1 c,d only provide hints that the cp = 0.38 mg/g state point behaves differently from other state points. Direct evidence comes from the use of bond-orientational order parameters, presented in Fig. 5 and 6. 
Following the Referee's kind suggestion, we have now moved previous Fig. S3 to a new Fig. 3b. 
We hope that this will make the difference of the cp = 0.38 mg/mg state point from the others clearer.

\begin{referee}
(Yet, in the new section, the authors write: "This result clearly indicates that a single state point follows a different arrest mechanism, in which phase separation is arrested by crystallization and not by glassiness." Is that really true? It seems overstated. Don't Figure 1 c,d and SF3 just establish that the end point is different? They don't say anything about the mechanism, which would require kinetic analysis, as comes later. And it doesn't say anything about arrest, which would require dynamics.)
\end{referee}

In the new version of the manuscript we have removed that sentence, and replaced it with:

\emph{
These structural analysis provide some evidence that the state point $\phi\approx 0.33$ and $c_p=0.38\,$mg/g could be following a different arrest mechanism, in which phase separation is arrested by crystallization and not by glassiness. In the reminder of this manuscript we investigate this state more closely, to confirm these early results.}

We also agree that our previous manuscript didn't say much about the actual kinetic mechanism by which crystal-gels form. This point was also raised by Referee \#1 in noting that we didn't provide any additional insight in the phase separation process. In order to fully address this issue, we have implemented a kinetic analysis of the trajectories at all state points, computing the probability of bond-breaking between consecutive frames. This new kinetic analysis reveals that at low-polymer concentrations a new mechanism (stress-driven aging) is taking place, which has the effect of increasing the average coordination of the colloidal particles in the liquid phase.
We believe this to be a novel observation of this phenomenon. Most past simulation works have neglected, or not fully taken into account, the role of hydrodynamic interactions (HI) in the phase separation process. As was shows for example in Ref.~\cite{furukawa2010key}, the phase separation process without HI produces rather compact structures, which would then easily crystallize to form compact nuclei. This is not what happens if one take HI into consideration. Due to the incompressibility condition, the liquid network has a lower-connectivity, and the viscoelastic phase separation results in the accumulation of mechanical stress on these liquid bonds. Two things can happen depending on the polymer concentration. At high-polymer concentration, the bonds are not-easily broken, and the network which is first formed cannot undergo structural reorganisation. This inevitably leads to gelation. On the other side, if the polymer concentration is low enough, the liquid network can reorganise with a rather sharp 
transition to a more compact network. This condition is necessary for the appearance of the first crystalline nuclei. In our new manuscript, we are able to follow the kinetics of these processes by looking at the time evolution of the bond-breaking probability. The different network states are in fact characterised by different bond-breaking probabilities and by different coordination numbers. In Fig.~2a,b we show these quantities for high-polymer concentration samples, where spinodal decomposition is the only process that occurs. At low-polymer concentrations (Fig.~2c,d) instead, the network undergoes a sharp transition to a state with lower bond-breaking probability and higher connectivity. This transition is the stress-driven aging process, and increases the connectivity of the network. Despite being theoretically predicted, Ref.~\cite{tanaka2000,tanaka2007spontaneous}, this transition was not directly observed before in experiments. 
We stress that for the experimental observation we need to follow particles at a single-particle level from the beginning of phase separation. This was made possible by a new experimental protocol of salt injection combined with confocal microscopy observation.

\begin{referee}
 The revised Figure 4, along with Figure S4, are very helpful to resolving the relative contributions of the three crystalline pathways. In Figure 4b,c, I still think that surface particles should somehow be distinguished - they represent a distinct state and a kind of consistency check, because the different crystallization pathways should all be mediated by this state. Without including the surface states in the primary figure, I don't see how the reader can inspect Figure S4, and get to the 54/34/11 abundances for direct, Bergeron and Ostwald processes. These abundances are the crucial claim of the paper.

In summary the confirmatory experiments (Fig S5 and Movie S3), the additional analysis of the crystalline state (Fig 3S) and the full transition analysis (rev Fig 4 and F S4) offer significant new support of the authors' claims; the authors have something new relative to prior literature. At this point, I think that the authors should: (i) better acknowledge that what is observed here is found in a quite specific region of the state diagram ($phi\sim 0.33$ and $c_p\sim 0.39$ {plus minus} 0.01 mg/g); (ii) better connect Figure 4 and Figure S4 by accounting for surface states in the former figure.

\end{referee}

We are glad to hear that the Referee thought that the previous revisions offered significant new support of the authors' claims. 
On the comment (i), as we explained above, we have clarified the physical conditions required for crystal-gel formation to take place. We have also discussed how we can widen the region where this is to be observed (see page 14). 
(ii) Following the suggestion, in the new version of the manuscript we follow the Referee's recommendation and include in Fig. 6b the fraction of surface particles, and update the description accordingly.

%\textcolor{red}{The Referee insists on having surface particles represented in Fig.6b. Even if I think that the figure becomes more complicated this way, maybe we should just update the figure. I'm sending Mathieu the dataset with the fraction of surface particles and liquid particles in case you agree on this change. If you do, I'll also update the description in the manuscript.}

%\textcolor{red}{We remind that the 54/34/11 abundances was obtained by looking at the phase history of each colloidal particle. Fig. 6b and Supplementary Fig. 4 don't contain information about the phase history of individual particles, just about the total number of particles in each phase, and the time-dependence of the transition probabilities. Of course these quantities are closely connected, and we hope to have provided a detailed description of the process (??).}

\vskip 5mm

We hope that the Referee will find that the new analysis have revealed a new mechanism behind 
crystal-gel formation and increased the novelty of our manuscript considerably and 
the revised manuscript is now suitable for publication in Nature Materials.

\clearpage

\bibliographystyle{naturemag2}
\bibliography{biblio}

\end{document}

### 2016-12-20 crystal-gel/nature/RebuttalLetter.tex

@@
 Our new results show that, at low polymer-concentrations, the network undergoes a structural reorganisation, which results in more compact structures, that are a necessary condition for the appearance of the first crystalline nuclei.
-To our knowledge this is the first experimental demonstration of this process: it is not captured by computer simulations, as they generally neglect hydrodynamic interactions which play a fundamental role in this kinetic process, and no previous experiment has had the capability of tracking particles during the early stages of spinodal decomposition.
+To our knowledge this is the first experimental demonstration of this process: it is also not easy to be captured by computer simulations, as they often neglect hydrodynamic interactions which play a fundamental role in this kinetic process, and no previous experiment has had the capability of tracking particles during the early stages of spinodal decomposition.
 
@@
 
-We would like to thank the Referee for her/his comments. 
+We would like to thank the Referee for his/her comments. 
 We are glad to see that he/she called the work \emph{very beautiful and interesting}. 
@@
 
-We thank the Referee for her/his appreciation of the improvements made during our last revision of the manuscript.
+We thank the Referee for his/her appreciation of the improvements made during our last revision of the manuscript.
 The Referee comments on the first 7 pages being not original, and simply describing the process of spinodal decomposition as previously observed in experiments and simulations.
@@
 For example, the new analysis shown in new Fig. 2 is only possible with this new experimental protocol and full particle tracking.  
-Finally, the last sentence has been modified, following the kind suggestion of the Referee.
+Finally, the last sentence of the abstract has been modified, following the kind suggestion of the Referee.
 
@@
 Our new results show that, at low polymer-concentrations, the network undergoes a structural reorganisation, which results in more compact structures, that are a necessary condition for the appearance of the first crystalline nuclei.
-To our knowledge, this is the first experimental demonstration of this process: it is not captured by previous computer simulations, as they generally neglect hydrodynamic interactions which play a fundamental role in this kinetic process, and no previous experiment has had the capability of tracking particles during the early stages of spinodal decomposition.
+To our knowledge, this is the first experimental demonstration of this process: it is also not easy to be captured by computer simulations, as they often neglect hydrodynamic interactions which play a fundamental role in this kinetic process, and no previous experiment has had the capability of tracking particles during the early stages of spinodal decomposition.
 
@@
 
-3) The data quality of Figure 1c, Figure 1d appear insufficient to classify the cp = 0.38 mg/mg state point as different from the others.
+3) The data quality of Figure 1c, Figure 1d appear insufficient to classify the $c_p$ = 0.38 mg/mg state point as different from the others.
 The authors should add replications, errors, and significance testing if they wish to distinguish the dimensionality (Fig 1c) and maximum volume fraction (Fig 1d) of the 0.38 mg/ml state point from the others.
 
-I continue to think that the evidence for different behavior in the cp = 0.38 mg/g state point from Figure 1 c,d is weak.
+I continue to think that the evidence for different behavior in the $c_p$ = 0.38 mg/g state point from Figure 1 c,d is weak.
 Supplemental Figure 3 is more convincing that the final state points are different.
@@
 We fully agree that the evidence from Fig.~1 c,d was not conclusive, but the goal of those figures was not to provide conclusive evidence, but to provide a comprehensive study of the early-stages of spinodal decomposition.
-In this context the Fig.~1 c,d only provide hints that the cp = 0.38 mg/g state point behaves differently from other state points.
+In this context the Fig.~1 c,d only provide hints that the $c_p$ = 0.38 mg/g state point behaves differently from other state points.
 Direct evidence comes from the use of bond-orientational order parameters, presented in Fig.
@@
 Following the Referee's kind suggestion, we have now moved previous Fig. S3 to a new Fig. 3b. 
-We hope that this will make the difference of the cp = 0.38 mg/mg state point from the others clearer.
+We hope that this will make the difference of the $c_p$ = 0.38 mg/mg state point from the others clearer.
 

### 2016-12-20 crystal-gel/nature/cover_letter_NM.tex

@@
 \noindent
-{\bf Dear Pep and/or a newly assigned Editor in Nature Materials,}
+{\bf Dear Iskra,}
 \vskip 0.7cm
-We would like to thank you for your providing us with two Referee reports for our manuscript (NM15113574).
-We understand your decision to reject the manuscript, but, after careful consideration of the Referee reports, we couldn't avoid thinking that they are mostly supportive of our work, and the only Referee (Referee \#1) that is somewhat against publication is basing his/her argument on a probable misunderstanding of the true novelty of the first part of our work.
+
+We would like to resubmit our manuscript (NM15113574), which was considered for publication in Nature Materials but rejected on the basis of reviewers' reports. 
+This manuscript was taken care by Pep and the final decision was made by him just after he moved to Nature Biomedical Engineering as a chief editor. 
+Since we learned in the Nature Materials webpage that you handle manuscripts on soft matter and your scientific background is closely related to the subject of our manuscript, we thought that it may be appropriate to contact you.
+
+First of all, we would like to thank Pep for providing us with two Referee reports for our manuscript. 
+We understand his decision to reject the manuscript, but, after careful consideration of the Referee reports, we couldn't avoid thinking that they are mostly supportive of our work, and the only Referee (Referee \#1) that is somewhat against publication is basing his/her argument on a probable misunderstanding of the true novelty of the first part of our work.
 

### 2016-12-20 crystal-gel/nature/crystal_gel_r.tex

@@
 resolution, and at the same time scales over which nucleation takes place. In colloids with short-range attractions the liquid-gas transition 
-becomes metastable with respect to crystallization~\cite{anderson2002insights,lekkerkerker2011colloids}, and can form gels.  
+becomes metastable with respect to crystallization~\cite{anderson2002insights,lekkerkerker2011colloids}, and can form gels   
 ~\cite{poon2002,zaccarelli2007,piazza1994phase,verhaegh1997transient,lu2008gelation}. 
@@
 \textcolor{blue}{
-In this Letter, we show that the system can undergo a novel crystallization pathway that leads to the formation of interconnected crystalline droplets. 
+In this Article, we show that the system can undergo a novel crystallization pathway that leads to the formation of interconnected crystalline droplets. 
 These droplets originate from crystal seeds that form inside the liquid branches of the spinodal aggregate of network morphology.
@@
 We then show that the growth of the crystal seeds occurs by three different routes:
-(i) growth within the dense branches of the gel;
+(i) Growth within the dense branches of the gel;
 (ii) Bergeron process, analogous to ice formation in mixed phase clouds \cite{glickman2000glossary,morrison2012resilience}, where ice droplets grow at the
@@
 the random gelation universality class exponent ($D=2.53$) as shown by the gray line in the inset of Fig.~\ref{fig:network_formation}.
-Here we do not imply that the formation of the network follows the random gelation universality class, but just that our
-fractal (or effective) dimension is compatible with it.
+\textcolor{blue}{Here we do not imply that the formation of the network follows the random gelation universality class, but just that our
+fractal (or effective) dimension is compatible with it.}
 
@@
 {
-\textcolor{red}{From initial to intermediate times, before percolation is complete}, hydrodynamic interactions play a fundamental role~\cite{tanaka2000,tanaka2007spontaneous,furukawa2010key}.
+From initial to intermediate times, before percolation is complete, hydrodynamic interactions play a fundamental role~\cite{tanaka2000,tanaka2007spontaneous,furukawa2010key}.
 Without hydrodynamic interactions, as often assumed in simulations, particles have the tendency to aggregate in compact structures and subsequently form thick network structures.
@@
 with high polymer concentration retain the exponent $D=2.5(3)$, 
-which is the random gel universality class exponent, as we also observed in the early stages of the gelation process (see inset of Fig.~\ref{fig:network_formation}),
+which is the random gel universality class exponent, as we also observed in the early stages of the gelation process (see the inset of Fig.~\ref{fig:network_formation}),
 the state point with low polymer concentration ($\phi\approx 0.33$ and $c_p=0.38$ mg/g) displays the biggest deviation from that exponent. In the figure we also
@@
 
-To gain additional insight, in Fig.~\ref{fig:late_structure}b we plot the late stage (after all samples already underwent gas-liquid phase separation) time evolution of the local volume fraction $\phi_\mathrm{loc}$ of closed packed particles (having 12 neighbours).
+To gain additional insight, in Fig.~\ref{fig:late_structure}b we plot the late stage (after all samples already underwent gas-liquid phase separation) time evolution of the local volume fraction $\phi_\mathrm{loc}(12)$ of closed packed particles (having 12 neighbours).
 The volume fraction is obtained by computing the Voronoi diagram of each configuration, which uniquely assigns a volume to each colloidal particle.
@@
 Usually such nano-porous materials are formed via at least two steps: for example, phase separation and dealloying of a mixture
-\cite{erlebacher2001evolution}. Unlike such a method, our novel mechanism allows us to spontaneously form sponge-like nano-porous crystals \emph{in a single step}, 
+\cite{erlebacher2001evolution}. Unlike such a method, our novel mechanism allows us to spontaneously form sponge-like nano-porous crystals \textcolor{blue}{\emph{in a continuous process}}, 
 which may have an impact on many applications. 

### 2016-12-20 nature/RebuttalLetter.tex
Explain more the novelty of the reservoir cell

--- 2016-12-20 crystal-gel/nature/RebuttalLetter.tex
@@
 We also simplified the title to avoid misunderstanding of our message.
-Regarding the use of a novel salt--injection we believe this to be the first time this cell design has been applied for confocal microscopy observation at a single-particle level and
-used to study the formation of porous crystals in gels, 
-which is the key to access the early stage of viscoelastic phase separation at a single particle resolution with full particle tracking. 
-For example, the new analysis shown in new Fig. 2 is only possible with this new experimental protocol and full particle tracking.  
+
+Regarding our experimental protocol, it is indeed reminiscent of salt injection, but differs in two important ways.
+(i) Rather than an injection salts are introduced through a membrane.
+This dialysis protocol allows salt introduction without external flow.
+(ii) We are in a regime where, upon salt introduction, charges are screened throughout the sample almost instantaneously, that is to say in a few Brownian times.
+This is due to the short length the ions have to diffuse from the reservoir to the top of the sample and to the large diffusion constant ratio between the ions and our large colloidal particles.
+We believe this to be the first time this cell design has been applied for confocal microscopy observation at a single-particle level and used to study the formation of porous crystals in gels, which is the key to access the early stage of viscoelastic phase separation at a single particle resolution with full particle tracking.
+For example, the new analysis shown in new Fig.~2 is only possible with this new experimental protocol and full particle tracking.  
 Finally, the last sentence of the abstract has been modified, following the kind suggestion of the Referee.
@@
 
-We hope that the Referee will find that the new analysis have revealed a new mechanism behind 
+We hope that the Referee will find that the new analysis has revealed a new mechanism behind 
 crystal-gel formation and increased the novelty of our manuscript considerably and 

### 2016-12-20 crystal-gel/nature/crystal_gel_r.tex
rephrase crystal growth implying less surface

@@
 to the gas phase due to the higher vapour pressure of the liquid phase compared to the crystalline phase.
-\textcolor{blue}{The crystalline nuclei are initially composed mostly of surface particles, but as the growth of nuclei becomes more compact the number of surface particles slowly decreases, while bulk crystals keep increasing}.
+\textcolor{blue}{The crystalline nuclei are initially composed mostly of surface particles, but as the nuclei grow they becomes more compact and merge together such that the number of surface particles slowly decreases, while bulk crystals keep increasing}.
 After the onset of a steady-state gas population, there are three growth routes for the crystal. 

### 2016-12-20 nature/Supplementary.tex
Dear all,
I have made corrections associated with new Supplementary Movie 3. 
Please check them. 

If you agree with the inclusion of new Movie 3, I think that we are ready to resubmission!

How do you think?

Best wishes, 
Hajime
Merge movies 1 and 2 Rename 3 as 2

@@
 \centerline{\bf \large Supplementary Information for}
-\centerline{\bf \large``A novel route to the spontaneous formation of porous crystals via}
-\centerline{\bf \large viscoelastic phase separation''}
+\centerline{\bf \large``Formation of porous crystals via viscoelastic phase separation'' }
 \vspace{0.5cm}
@@
 
+%\clearpage
 \bibliographystyle{naturemag2}
@@
 \section*{Supplementary Movie 1}
-This movie shows the early stage of gelation process of $\phi\approx 0.30$ and $c_p=0.38$ mg/g, from the original confocal images at the same z-position as Fig.~3a.
-Total duration is 1160 s (530~$\tau$).
+This movie shows the gelation process of $\phi\approx 0.30$ and $c_p=0.38$ mg/g, from the original confocal images at the same z-position as Fig.~4d.
+As stressed by the time stamps, we use two very different frame rates for the first 12~s of the movie (1160~s in real time or 530~$\tau_B$) and the remaining 15~s (27000~s in real time or 12300~$\tau_B$).
 We quickly started scanning the sample after exchanging the salt-reservoir with the solvent including salt.
 At the beginning, salt had not reached the scanning volume and colloids were dispersed by electrostatic repulsion.
-At 0:02 in the movie, salt diffused into the scanning volume and screened the surface charges of colloids.
+At time stamp $t=0$, salt diffused into the scanning volume and screened the surface charges of colloids.
 Then, colloids immediately initiated phase separation by short-ranged attractions.
+At later times crystalline order is visible.
 
 \section*{Supplementary Movie 2}
-This movie shows the late stage after Supplemental Movie 1 with scanning duration of 27000~s (12300~$\tau$). 
-Please note that there is a large difference in the frame rate between Movie 1 and 2.
-
-\section*{Supplementary Movie 3}
-Same as Supplementary Movie 2, but with computer rendered trajectories obtained from the tracking of the colloidal particles.
-Particles are colored according to the following code: crystal (red), surface (blue), gas (green) and liquid (white).
+Same as the end of Supplementary Movie 1, but with computer rendered trajectories obtained from the tracking of the colloidal particles.
+Particles are coloured according to the following code: crystal (red), surface (blue), gas (green) and liquid (white).
 For display purposes, crystalline particles and gas particles are respectively 400\% and 50\% bigger than other particles (liquid or surface).

### 2016-12-20 nature/crystal_gel_r.tex
Merge movies 1 and 2 Rename 3 as 2

@@
 resolution, and at the same time scales over which nucleation takes place. In colloids with short-range attractions the liquid-gas transition 
-becomes metastable with respect to crystallization~\cite{anderson2002insights,lekkerkerker2011colloids}, and can form gels.  
+becomes metastable with respect to crystallization~\cite{anderson2002insights,lekkerkerker2011colloids}, and can form gels   
 ~\cite{poon2002,zaccarelli2007,piazza1994phase,verhaegh1997transient,lu2008gelation}. 
-\textcolor{blue}{This gel formation process can be regarded as viscoelastic phase separation  \cite{tanaka1999colloid,tanaka2000viscoelastic} into a dense liquid phase with slow dynamics and a dilute gas phase with fast dynamics.}
+This gel formation process can be regarded as viscoelastic phase separation  \cite{tanaka1999colloid,tanaka2000viscoelastic} into a dense liquid phase with slow dynamics and a dilute gas phase with fast dynamics.
 This difference in the viscoelastic properties between the two phases allows the system to form a space-spanning 
@@
 All samples share the same early stages of spinodal decomposition. 
-A typical early stage phase ordering process observed at $\phi\approx 0.33$ and $c_p=0.38$ mg/g can be seen in Supplementary Movie 1.
+A typical early stage phase ordering process observed at $\phi\approx 0.33$ and $c_p=0.38$ mg/g can be seen in at the beginning of Supplementary Movie~1.
 Due to strong dynamical asymmetry between colloids and the solvent \cite{tanaka1999colloid}, the colloidal particles start aggregating, 
@@
 the random gelation universality class exponent ($D=2.53$) as shown by the gray line in the inset of Fig.~\ref{fig:network_formation}.
-Here we do not imply that the formation of the network follows the random gelation universality class, \textcolor{blue}{but} just that our
-fractal (or effective) dimension is compatible with it.
+\textcolor{blue}{Here we do not imply that the formation of the network follows the random gelation universality class, but just that our
+fractal (or effective) dimension is compatible with it.}
 
@@
 
-Supplementary Movie 2 shows the late stage (after the initial spinodal decomposition) evolution of a 2D confocal slice.
+The end of Supplementary Movie~1 shows the late stage (after the initial spinodal decomposition) evolution of a 2D confocal slice.
 Even on this raw data the growth of crystalline regions is obvious.
@@
 With this analysis we reveal that crystallization events indeed start occurring for the state point $\phi\approx 0.33$ and $c_p=0.38$ mg/g, which has the lowest
-polymer concentration of all state points. This is shown in the inset of Fig.~\ref{fig:crystals}, where the \textcolor{blue}{red dashed} line indicates the average size of the crystallites, 
-while the \textcolor{blue}{black} line indicates the average number of crystallites as a function of time. The number of crystallites first rapidly increases as nucleation events
+polymer concentration of all state points. This is shown in the inset of Fig.~\ref{fig:crystals}, where the red dashed line indicates the average size of the crystallites, 
+while the black line indicates the average number of crystallites as a function of time. The number of crystallites first rapidly increases as nucleation events
 start occurring inside the liquid network, but eventually starts decreasing as the different crystallites grow and merge with each other.
-Supplementary Movie 3 shows a three-dimensional (3D) reconstruction  of the whole phase ordering sequence that emphasizes crystalline particles.
+Supplementary Movie~2 shows a three-dimensional (3D) reconstruction  of the whole phase ordering sequence that emphasizes crystalline particles.
 Although crystal clusters are isolated in this movie, they are bridged by the dense colloidal liquid phase and involved in the percolated gel network, as can be seen in Fig.
@@
 to the gas phase due to the higher vapour pressure of the liquid phase compared to the crystalline phase.
+\textcolor{blue}{The crystalline nuclei are initially composed mostly of surface particles, but as the nuclei grow they becomes more compact and merge together such that the number of surface particles slowly decreases, while bulk crystals keep increasing}.
 After the onset of a steady-state gas population, there are three growth routes for the crystal. 
@@
 which includes magma \cite{philpotts1998role}, biominerals \cite{rousseau2005multiscale}, and foods \cite{deman1987fat}.
+
+%\vspace{5mm}
+%\noindent
+%{\bf Data Availability.}
+
+%The data that support the plots within this paper and other findings of this study are available upon reasonable request.
 

### 2016-12-20 nature/Supplementary.tex

@@
 \section*{Supplementary Movie 2}
-Same as the end of Supplementary Movie 1, but with computer rendered trajectories obtained from the tracking of the colloidal particles.
+This movie shows the 3D structural evolution in the gelation process of $\phi\approx 0.30$ and $c_p=0.40$ mg/g in a capillary cell. 
+We show computer rendered trajectories obtained from the tracking of the colloidal particles.
 Particles are coloured according to the following code: crystal (red), surface (blue), gas (green) and liquid (white).

### 2016-12-21 crystal-gel/nature/RebuttalLetter.tex

@@
 we made after the first round of reviews: Reviewer \#1 calls the work \emph{very beautiful and interesting}, while Referee \#3 goes into the details of explaining how we actually satisfactorily clarified her/his previous doubts, with just a couple of minor points left.
-Reviewer \#2 was very positive during the first round.
+Reviewer \#2 supported publication after some minor revisions in the first round of review.
 
@@
 To our knowledge this is the first experimental demonstration of this process: it is also not easy to be captured by computer simulations, as they often neglect hydrodynamic interactions which play a fundamental role in this kinetic process, and no previous experiment has had the capability of tracking particles during the early stages of spinodal decomposition.
+The absence of any harmful flow upon the initiation of phase separation and any drift can be clearly seen in Supplementary Movie 1.
 
@@
 We also simplified the title to avoid misunderstanding of our message.
-Regarding the use of a novel salt--injection we believe this to be the first time this cell design has been applied for confocal microscopy observation at a single-particle level and
-used to study the formation of porous crystals in gels, 
-which is the key to access the early stage of viscoelastic phase separation at a single particle resolution with full particle tracking. 
-For example, the new analysis shown in new Fig. 2 is only possible with this new experimental protocol and full particle tracking.  
+
+Regarding our experimental protocol, it is indeed reminiscent of salt injection, but differs in two important ways.
+(i) Rather than an injection salts are introduced through a membrane.
+This dialysis protocol allows salt introduction without external flow.
+(ii) We are in a regime where, upon salt introduction, charges are screened throughout the sample almost instantaneously, that is to say in a few Brownian times.
+This is due to the short length the ions have to diffuse from the reservoir to the top of the sample and to the large diffusion constant ratio between the ions and our large colloidal particles.
+We believe this to be the first time this cell design has been applied for confocal microscopy observation at a single-particle level and used to study the formation of porous crystals in gels, which is the key to access the early stage of viscoelastic phase separation at a single particle resolution with full particle tracking.
+For example, the new analysis shown in new Fig.~2 is only possible with this new experimental protocol and full particle tracking.  
+The absence of any harmful flow upon the initiation of phase separation and any drift can be clearly seen in Supplementary Movie 1.  
 Finally, the last sentence of the abstract has been modified, following the kind suggestion of the Referee.
@@
 To our knowledge, this is the first experimental demonstration of this process: it is also not easy to be captured by computer simulations, as they often neglect hydrodynamic interactions which play a fundamental role in this kinetic process, and no previous experiment has had the capability of tracking particles during the early stages of spinodal decomposition.
+The absence of any harmful flow upon the initiation of phase separation and any drift can be clearly seen in Supplementary Movie 1.
 
@@
 In order to fully address this issue, we have implemented a kinetic analysis of the trajectories at all state points, computing the probability of bond-breaking between consecutive frames.
-This new kinetic analysis reveals that at low-polymer concentrations a new mechanism (stress-driven aging) is taking place, which has the effect of increasing the average coordination of the colloidal particles in the liquid phase.
+This new kinetic analysis reveals that at low-polymer concentrations a new mechanism (stress-driven ageing) is taking place, which has the effect of increasing the average coordination of the colloidal particles in the liquid phase.
 We believe this to be a novel observation of this phenomenon.

### 2016-12-21 nature/crystal_gel_r.tex

@@
 {\bf
-Viscoelastic phase separation of colloidal suspensions can be interrupted to form gels either by 
-%the glass transition, with the formation of colloidal gels~\cite{poon2002,zaccarelli2007,piazza1994phase,verhaegh1997transient,tanaka1999colloid,lu2008gelation}, or by crystallization, which results in the formation of a crystalline network~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways,sabin2012}.
-glass transition
-%~\cite{poon2002,zaccarelli2007,piazza1994phase,verhaegh1997transient,tanaka1999colloid,lu2008gelation} 
-or by crystallization. 
-%~\cite{soga1999metastable,fortini2008crystallization,perez2011pathways,sabin2012,zhang2012non}. 
-%But while the former process has been studied in detail, the latter has never been observed experimentally at a single-particle level and it is thus lacking a microscopic understanding. 
+Viscoelastic phase separation of colloidal suspensions can be interrupted to form gels either by glass transition or by crystallization.
 With a new confocal microscopy protocol, \textcolor{blue}{we follow the entire kinetics of phase separation, from the homogeneous phase to the formation of different arrested states.
 This allows us to unveil a novel crystallization pathway to sponge-like porous crystal structures in a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation.
-We show that crystallization is preceeded by a structural reorganization of the liquid phase, called stress-driven aging.
+We show that crystallization is preceded by a structural reorganization of the liquid phase, called stress-driven ageing.
 Nucleation starts inside this liquid network, but crystals grow past it by direct condensation of the gas phase on their surface, i.e.
 de-sublimation, driving evaporation of nearby liquid.
-This produces a network structure different from the original phase separation pattern.}
-%The latter processes via the gas phase has been overlooked so far.
-%This process represents the colloidal analogue of the Bergeron process~\cite{glickman2000glossary}, which explains the formation of ice crystals in mixed phase clouds and is at the origin of rain.
-Our finding gives us full experimental access to the kinetic pathway of this poorly understood phenomenon important for climate science.
-%The key is complex dynamic interplay involving the three phases upon simultaneous occurrence of two non-equilibrium phenomena, phase separation and crystallization. 
-We argue that similar crystal-gel states can be formed as the result of the gas-liquid phase separation of any crystallizable components, such as monoatomic and molecular systems, if dynamics of the liquid phase is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
-This mechanism will provide a novel pathway to form nano-porous crystals of metals and semiconductors without dealloying, %\cite{erlebacher2001evolution},
-which may be important for catalytic, optical, sensing, and filtration applications. 
-%\cite{ding2004nanoporous,ding2009nanoporous, wittstock2010nanoporous}.
-%, but also sheds new light on crystal network formation in more complex systems such as magma \cite{philpotts1998role}, biominerals \cite{rousseau2005multiscale}, and foods \cite{deman1987fat}.}
+This produces a network structure different from the original phase separation pattern.} We argue that similar crystal-gel states can be formed as the result of the gas-liquid phase separation of any crystallizable components, such as monoatomic and molecular systems, if dynamics of the liquid phase is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
+This mechanism will provide a novel pathway to form nano-porous crystals of metals and semiconductors without dealloying, which may be important for catalytic, optical, sensing, and filtration applications.
 }
 \end{abstract}
+
+%Viscoelastic phase separation of colloidal suspensions can be interrupted to form gels either by glass transition or by crystallization.
+With a new confocal microscopy protocol, we follow the entire kinetics of phase separation, from the homogeneous phase to the formation of different arrested states.
+This allows us to unveil a novel crystallization pathway to sponge-like porous crystal structures in a colloid-polymer mixture undergoing viscoelastic gas-liquid phase separation.
+We show that crystallization is preceded by a structural reorganization of the liquid phase, called stress-driven ageing.
+Nucleation starts inside this liquid network, but crystals grow past it by direct condensation of the gas phase on their surface, i.e.
+de-sublimation, driving evaporation of nearby liquid.
+This produces a network structure different from the original phase separation pattern.
+We argue that similar crystal-gel states can be formed as the result of the gas-liquid phase separation of any crystallizable components, such as monoatomic and molecular systems, if dynamics of the liquid phase is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
+This mechanism will provide a novel pathway to form nano-porous crystals of metals and semiconductors without dealloying, which may be important for catalytic, optical, sensing, and filtration applications.
 
@@
 These droplets originate from crystal seeds that form inside the liquid branches of the spinodal aggregate of network morphology.
-We show that a necessary condition for the nucleation of crystal seeds is the stress-driven aging of the network which releases the stresses that build-up during the network formation due to hydrodynamic interactions, allowing for an increase in the number of average neighbours.
+We show that a necessary condition for the nucleation of crystal seeds is the stress-driven ageing of the network which releases the stresses that build-up during the network formation due to hydrodynamic interactions, allowing for an increase in the number of average neighbours.
 If the average concentration of particles is high enough, this stress relaxation triggers the nucleation process of crystals. 
@@
 As described above and also in Methods, salt injection initiates liquid-gas phase separation of the colloidal suspension. 
+\textcolor{blue}{We can confirm the absence of any harmful flow upon the initiation of phase separation and any drift in 
+Supplementary Movie~1.}  
 All samples share the same early stages of spinodal decomposition. 
@@
 \textcolor{blue}{
-It is worth noting that the two-step beheviour of $p_\text{break}$ and $\bar{n}$ is not due to crystallization, as this only starts after the network reorganization.
+It is worth noting that the two-step behaviour of $p_\text{break}$ and $\bar{n}$ is not due to crystallization, as this only starts after the network reorganization.
 For example, the fraction of crystalline particles for all frames in Fig.~\ref{fig:bond_breaking}d is always below $0.3\%$.

### 2017-01-26 nature/RebuttalLetter.tex

@@
 For example, the new analysis shown in new Fig.~2 is only possible with this new experimental protocol and full particle tracking.  
+The absence of any harmful flow upon the initiation of phase separation and any drift can be clearly seen in Supplementary Movie 1.  
 Finally, the last sentence of the abstract has been modified, following the kind suggestion of the Referee.
@@
 \begin{referee}
-I was aware of the different frame rates on the two movies but I agree that the time stamps are helpful.
+\noindent
+1) I was aware of the different frame rates on the two movies but I agree that the time stamps are helpful.
 Figure S5 addresses my concern about reproducibility at the single state point, and the different geometry itself is helpful too.
@@
 
-Following the Referee's recommendation, we have put the following sentence in the abstract: We clarify the physical requirements necessary for this to happen.
-Although this does not explicitly specify the points we observed crystal gels, our new analysis (see below)
+\textcolor{blue}{We are glad to hear that the reviewer found that the time stamps are helpful and also Fig. S5 addresses his/her concern.}     
+Following the Referee's recommendation, we have put the following sentence in the abstract: ``We clarify the physical requirements necessary for this to happen.'' Although this does not explicitly specify the points we observed crystal gels, our new analysis (see below)
 has revealed the physical conditions required for observing crystal-gel formation.
@@
 To our knowledge, this is the first experimental demonstration of this process: it is also not easy to be captured by computer simulations, as they often neglect hydrodynamic interactions which play a fundamental role in this kinetic process, and no previous experiment has had the capability of tracking particles during the early stages of spinodal decomposition.
+The absence of any harmful flow upon the initiation of phase separation and any drift can be clearly seen in Supplementary Movie 1.
 
 \begin{referee}
-The addition of the SI movie 3 and the discussion of Figure 2b addresses this concern for me.
+2) The addition of the SI movie 3 and the discussion of Figure 2b addresses this concern for me.
 It's the percolation that is critical to the authors claim of a crystal gel - the percolated structure needs to persist over long times, otherwise, for example, the gel wouldn't have a finite low frequency elastic modulus.
 I think SI move 3 does show this.
+\end{referee}
 
-3) The data quality of Figure 1c, Figure 1d appear insufficient to classify the $c_p$ = 0.38 mg/mg state point as different from the others.
+\textcolor{blue}{We are glad to hear that the reviewer thought that SI movie 3 and the discussion of Fig. 2b have addressed his/her concern and supported our claim of a crystal gel.}
+
+\begin{referee}
+3-1) The data quality of Figure 1c, Figure 1d appear insufficient to classify the $c_p$ = 0.38 mg/mg state point as different from the others.
 The authors should add replications, errors, and significance testing if they wish to distinguish the dimensionality (Fig 1c) and maximum volume fraction (Fig 1d) of the 0.38 mg/ml state point from the others.
@@
 \begin{referee}
-(Yet, in the new section, the authors write: "This result clearly indicates that a single state point follows a different arrest mechanism, in which phase separation is arrested by crystallization and not by glassiness." Is that really true?
+3-2) (Yet, in the new section, the authors write: "This result clearly indicates that a single state point follows a different arrest mechanism, in which phase separation is arrested by crystallization and not by glassiness." Is that really true?
 It seems overstated.
@@
 \begin{referee}
- The revised Figure 4, along with Figure S4, are very helpful to resolving the relative contributions of the three crystalline pathways.
+4) Figure S4 is critical to the interpretation, including information about all the transition pathways.
+\end{referee}
+
+\textcolor{blue}{We are glad to know that the reviewer thought that Fig. S4 is useful for supporting the interpretation.}
+
+\begin{referee}
+ 5) The revised Figure 4, along with Figure S4, are very helpful to resolving the relative contributions of the three crystalline pathways.
 In Figure 4b,c, I still think that surface particles should somehow be distinguished - they represent a distinct state and a kind of consistency check, because the different crystallization pathways should all be mediated by this state.
@@
 These abundances are the crucial claim of the paper.
+ \end{referee}
 
+ \textcolor{blue}{We are glad to hear that the revised Fig.
+4, along with Fig.
+S4, are very useful to resolving the relative contributions of the three crystalline pathways.
+Following the kind suggestion of the reviewer, we have now included the fraction of surface particles in Fig.
+6b and added the related discussion.}
+
+\begin{referee}
 In summary the confirmatory experiments (Fig S5 and Movie S3), the additional analysis of the crystalline state (Fig 3S) and the full transition analysis (rev Fig 4 and F S4) offer significant new support of the authors' claims; the authors have something new relative to prior literature.
@@
 We have also discussed how we can widen the region where this is to be observed (see page 14).
-(ii) Following the suggestion, in the new version of the manuscript we follow the Referee's recommendation and include in Fig.
+(ii)  \textcolor{blue}{As mentioned above,} following the suggestion, in the new version of the manuscript we follow the Referee's recommendation and include in Fig.
 6b the fraction of surface particles, and update the description accordingly.

### 2017-01-26 nature/crystal_gel_r.tex

@@
 de-sublimation, driving evaporation of nearby liquid.
-This produces a network structure different from the original phase separation pattern.} We argue that similar crystal-gel states can be formed as the result of the gas-liquid phase separation of any crystallizable components, such as monoatomic and molecular systems, if dynamics of the liquid phase is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
+This produces a network structure different from the original phase separation pattern.
+We clarify the physical requirements necessary for this to happen.} We argue that similar crystal-gel states can be formed as the result of the gas-liquid phase separation of any crystallizable components, such as monoatomic and molecular systems, if dynamics of the liquid phase is slow enough to induce viscoelastic phase separation, but fast enough to prevent immediate vitrification.
 This mechanism will provide a novel pathway to form nano-porous crystals of metals and semiconductors without dealloying, which may be important for catalytic, optical, sensing, and filtration applications.

### 2017-01-26 nature/RebuttalLetter.tex

@@
 Figure S5 addresses my concern about reproducibility at the single state point, and the different geometry itself is helpful too.
-I think the authors should be up front in the manuscript that the mechanism discussed is found in a very specific window (from the figures and SI, I think it is $\phi\sim 0.33$ and $c_p~0.39$ (plus minus) 0.01 mg/g) by moving this point into the abstract.
+I think the authors should be up front in the manuscript that the mechanism discussed is found in a very specific window (from the figures and SI, I think it is $\phi\sim 0.33$ and $c_p \sim0.39$ (plus minus) 0.01 mg/g) by moving this point into the abstract.
 \end{referee}
 
-\textcolor{blue}{We are glad to hear that the reviewer found that the time stamps are helpful and also Fig. S5 addresses his/her concern.}     
+We are glad to hear that the reviewer found that the time stamps are helpful and also Fig. S5 addresses his/her concern.      
 Following the Referee's recommendation, we have put the following sentence in the abstract: ``We clarify the physical requirements necessary for this to happen.'' Although this does not explicitly specify the points we observed crystal gels, our new analysis (see below)
@@
 
-\textcolor{blue}{We are glad to hear that the reviewer thought that SI movie 3 and the discussion of Fig. 2b have addressed his/her concern and supported our claim of a crystal gel.}
+We are glad to hear that the reviewer thought that SI movie 3 and the discussion of Fig. 2b have addressed his/her concern and supported our claim of a crystal gel.
 
@@
 
-\textcolor{blue}{We are glad to know that the reviewer thought that Fig. S4 is useful for supporting the interpretation.}
+We are glad to know that the reviewer thought that Fig. S4 is useful for supporting the interpretation.
 
@@
 
- \textcolor{blue}{We are glad to hear that the revised Fig.
+We are glad to hear that the revised Fig.
 4, along with Fig.
@@
 Following the kind suggestion of the reviewer, we have now included the fraction of surface particles in Fig.
-6b and added the related discussion.}
+6b and added the related discussion.
 
@@
 We have also discussed how we can widen the region where this is to be observed (see page 14).
-(ii)  \textcolor{blue}{As mentioned above,} following the suggestion, in the new version of the manuscript we follow the Referee's recommendation and include in Fig.
+(ii)  As mentioned above, following the suggestion, in the new version of the manuscript we follow the Referee's recommendation and include in Fig.
 6b the fraction of surface particles, and update the description accordingly.
